### 2013-10-09 sigplanconf-template.tex
adding paper

%-----------------------------------------------------------------------------
%
%			Edwin's and Franck Paper !
%  uses the template for sigplanconf LaTeX Class
%
%-----------------------------------------------------------------------------

\documentclass{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\newcommand{\cL}{{\cal L}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{PLPV ’14, Co-located with POPL 2014}{January 21, 2014, San Diego, CA, USA.}
\copyrightyear{2013} 
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
\preprintfooter{My paper for PLPV’14}
\doi{nnnnnnn.nnnnnnn}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Using dependent types for writting correct by construction tactics}
\subtitle{An exemple for a solver over algebraics structures}

\authorinfo{Edwin Brady\and Franck Slama}
           {University of St Andrews, United-Kingdom}
           {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}

\maketitle

\begin{abstract}
In the quest of certified software, at least two main approachs are possible. The usual way is generally to write a program, and then to prove its correctness afterwards. An alternative is to ensure the correction by construction, i.e., by having such precise types that all the properties we want are already ensured. Both approach have their benefits. The usual way is naturally good for dealing with a lot of properties to ensure, one proved after the other, and all are proved after the definition of the function. The correct by construction approach is not so good for dealing with lot of properties to ensure, but one of its merits is that the writing of the program and the proof is simultaneously -forcing the programmer to think about correctness while programming-, and every bit of code has its "proof part", making therefore a strong link between the function and its proof of correctness. If the code of the function changes, that will be much easier to change the correspong bit in the proof. The usual approach usually fails for this purpose, since a script of proof done afterwards in a proof mode is not easy to maintain, mainly because these kind of scripts are not easy to read.
In this paper, we propose to describe as an exemple a correct by construction implementation of a solver of equalities over some algebraic structures (magma, semigroup, monoid, group, ring and commutative ring). This kind of automatic solvers are useful for proof assistants because these proofs of equalities usually consist of a long and not really interesting serie of rewriting. Thanksfully, for each of the algebraics structures we mentionned, there exists a complete and decidable algorithm which determines if a given equality is true, and which constructs a proof of this fact if appropriate. Our approach is to work by reflection -enabling to write the tactic in the language itself-, and to ensure its correness by construction, thanks to our extensively use of dependent types. Also, our approach is fairly  generic and avoid a lot of duplication for the solvers over the different structures.

%Previous abstract, too much focus on the solver, and not enought on the correct by construction part, wich is our contribution

%In the quest of certified sofware, one problem still remains: how can we make the process of proving easier? Some auxiliary proofs can be long and are not part of the interesting bit. For these cases, we would like to be able to let the machine find the proof for us. In this paper  we focus on a specific kind of  proofs : proofs of equalities over groups and rings. For this kind of structures, there exists a complete and decidable algorithm which determines if a given equality is true, and constructs a proof of this fact if so. However, since we want the tactic to be safe, we need to produce a proof of correction of this implementation. In our approach, this proof is obtained for free, because the tactic is built correct by construction, thanks to our extensive use of dependent types. Our approach is to work by reflection, and to normalize both side of the equality with a serie of rewritings that uses the axioms of the algebraic structures. At every step of the rewriting, a proof of equality between the new and the preivous terms is generated. Our approach is also generic, since it works for groups or ring, whether or not commutative. This example shows that the use of dependent types can help building software correct by construction.
\end{abstract}

\category{F.4.1}{Mathematical logic and formal languages}
			{Mathematical logic}
			[Proof theory]
\category{F.3.1}{Logics and meanings of programs}
			{Specifying and Verifying and Reasoning about Programs}
			[Mechanical verification]
\category{D.3.3}{Programming Languages}
			{Language Constructs and Features}
			[Abstract data types]

% general terms are not compulsory anymore, 
% you may leave them out
\terms
Languages, Verification

\keywords
dependent types, proof automation, reflection

\section{Introduction}

As soon as we start to work with numbers on a proof assistant, we are facing the need to prove that a certain term is equal to another. Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties. These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for exemple), or by axioms if we are working with an abstract structure.
These proofs of equality are usually done by hand, and are made with a potentially long serie of rewriting, using the available properties that we have. For exemple, if x and y are relatives numbers, and if we want to show that :
(x + y) * z = (z * x) = (y * z), we can first use the fact the symmetry of *, saying that
for all a b, a * b = b * c
With the use of this property, we can rewrite the left part into the term z * (x + y).
If now we use the distributivity of * on +, which says that :
forall a b c, a * (b+c) = a*b + a*c,
we now obtain the term x*z + y*z for the left side.
If we use again the symmetry of * on the subterm (x*z), the left side of the equality becomes : z*x + y*z, which is what we want.

This kind of proofs therefore consists of a potentially long serie of rewriting, every rewriting step using one property of the theory. Without some specific automation, this serie of rewriting is done by the user of the proof assistant. This is time consuming, and a little change in the left or the right hand side of the equality can invalidate completely the proof. The reusability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term. In this paper, we describe a certified implementation of a solver for algebraics structures, for the Idris language. Idris is a relatively new purely functional programming language with full dependent types. Idris also supports tactic based proofs.

For our goal, we focus on some specific theories, which are the folowing algebraics structures : magmas, semi-group, monoid, group, abelian group, ring and commutative ring.
For these structures, this is effectively possible to decide if a given equality is true, and to produce the proof of the equality. This result comes directly from the fact that there exist a normal form for every element of these structures. The general idea is therefore to normalize both side of the equality, and then to compare them using the usual syntactical Leibniz's equality.
This approach was followed by [paper Coq ring] for implementing a ring solver for the Coq proof assistant.
Our contribution mainly consits of two parts :
	- We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procudure, and proving afterwards that this function effectively computes a normal form. For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters. 
	- We try to be as general generic as possible, and re-use as much as possible the code of normalization from the strucures below in the upper structures. For exemple developping an expression by using the axiom of distribituvty works exactly the same in a commutative ring and in a ring.

\section{A brief overview of Idris}
Idris is a purely functional

Idris is a purely funcitonal programming language fairly recent. Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not desegned as a proof asssiant, but has a real programming language. One of its more interesting features -and the one that enables to effectively manipulate logical statments and proofs in the language- is its full and effective support for dependent types. Especially, Idris has a particularly good dependent pattern maching, which is something mandatory for writting code correct by construction.

\section{Normalization and polynomials in normal form}
% Ici on donne l'idée générale. Nom de la section peut etre à revoir
Lots of text.

More text.

\section{Datatypes}

Lots of text.

\subsection{Hierarchy of algebraic theorys}

Lots of text.

The exemple with Nat is a monoid

More text.

\subsection{Reflection in an abstract syntax}

Lots of text.

More text.

\section{Normalization procedures}

\subsection{Normalization for terms in a magma}

\subsection{Normalization for terms in a semi-group}

\subsection{Normalization for terms in a monoid}

\subsection{Normalization for terms in a group}

\subsection{Normalization for terms in a ring}

\subsection{Normalization for terms in a commutative ring}

\section{Coating}
%enrobage

\section{Metaification}

\section{Conclusion}

\section{References}

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

More text.

Lots of text.

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

\acks

Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}

\end{document}

%                       Revision History
%                       -------- -------
%  Date         Person  Ver.    Change
%  ----         ------  ----    ------

%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

### 2013-10-09 paper/paper.tex
re-added paper

--- 2013-10-09 sigplanconf-template.tex
@@
 \usepackage{graphicx}
+\usepackage{listings}
 \newcommand{\cL}{{\cal L}}
@@
            {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}
+\bibliographystyle{abbrvnat}
 
@@
 
-\begin{abstract}
-In the quest of certified software, at least two main approachs are possible.
-The usual way is generally to write a program, and then to prove its correctness afterwards.
-An alternative is to ensure the correction by construction, i.e., by having such precise types that all the properties we want are already ensured.
-Both approach have their benefits.
-The usual way is naturally good for dealing with a lot of properties to ensure, one proved after the other, and all are proved after the definition of the function.
-The correct by construction approach is not so good for dealing with lot of properties to ensure, but one of its merits is that the writing of the program and the proof is simultaneously -forcing the programmer to think about correctness while programming-, and every bit of code has its "proof part", making therefore a strong link between the function and its proof of correctness.
-If the code of the function changes, that will be much easier to change the correspong bit in the proof.
-The usual approach usually fails for this purpose, since a script of proof done afterwards in a proof mode is not easy to maintain, mainly because these kind of scripts are not easy to read.
-In this paper, we propose to describe as an exemple a correct by construction implementation of a solver of equalities over some algebraic structures (magma, semigroup, monoid, group, ring and commutative ring).
-This kind of automatic solvers are useful for proof assistants because these proofs of equalities usually consist of a long and not really interesting serie of rewriting.
-Thanksfully, for each of the algebraics structures we mentionned, there exists a complete and decidable algorithm which determines if a given equality is true, and which constructs a proof of this fact if appropriate.
-Our approach is to work by reflection -enabling to write the tactic in the language itself-, and to ensure its correness by construction, thanks to our extensively use of dependent types.
-Also, our approach is fairly  generic and avoid a lot of duplication for the solvers over the different structures.
-
-%Previous abstract, too much focus on the solver, and not enought on the correct by construction part, wich is our contribution
-
-%In the quest of certified sofware, one problem still remains: how can we make the process of proving easier?
-Some auxiliary proofs can be long and are not part of the interesting bit.
-For these cases, we would like to be able to let the machine find the proof for us.
-In this paper  we focus on a specific kind of  proofs : proofs of equalities over groups and rings.
-For this kind of structures, there exists a complete and decidable algorithm which determines if a given equality is true, and constructs a proof of this fact if so.
-However, since we want the tactic to be safe, we need to produce a proof of correction of this implementation.
-In our approach, this proof is obtained for free, because the tactic is built correct by construction, thanks to our extensive use of dependent types.
-Our approach is to work by reflection, and to normalize both side of the equality with a serie of rewritings that uses the axioms of the algebraic structures.
-At every step of the rewriting, a proof of equality between the new and the preivous terms is generated.
-Our approach is also generic, since it works for groups or ring, whether or not commutative.
-This example shows that the use of dependent types can help building software correct by construction.
-\end{abstract}
-
-\category{F.4.1}{Mathematical logic and formal languages}
-			{Mathematical logic}
-			[Proof theory]
-\category{F.3.1}{Logics and meanings of programs}
-			{Specifying and Verifying and Reasoning about Programs}
-			[Mechanical verification]
-\category{D.3.3}{Programming Languages}
-			{Language Constructs and Features}
-			[Abstract data types]
-
-% general terms are not compulsory anymore, 
-% you may leave them out
-\terms
-Languages, Verification
-
-\keywords
-dependent types, proof automation, reflection
+\input{./content/abstract.tex}
 
 \section{Introduction}
+\nocite{*}
 
@@
 These proofs of equality are usually done by hand, and are made with a potentially long serie of rewriting, using the available properties that we have.
-For exemple, if x and y are relatives numbers, and if we want to show that :
-(x + y) * z = (z * x) = (y * z), we can first use the fact the symmetry of *, saying that
+For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
+$(x + y) * z = (z * x) = (y * z)$, we can first use the fact the symmetry of $*$, saying that
 for all a b, a * b = b * c
@@
 If now we use the distributivity of * on +, which says that :
-forall a b c, a * (b+c) = a*b + a*c,
+$\forall a b c, a * (b+c) = a*b + a*c$,
 we now obtain the term x*z + y*z for the left side.
@@
 \section{A brief overview of Idris}
-Idris is a purely functional
 
-Idris is a purely funcitonal programming language fairly recent.
-Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not desegned as a proof asssiant, but has a real programming language.
+Idris is a purely functional programming language fairly recent.
+Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not designed to be a proof asssiant, but as a real programming language easily usable by non-logicians.
 One of its more interesting features -and the one that enables to effectively manipulate logical statments and proofs in the language- is its full and effective support for dependent types.
-Especially, Idris has a particularly good dependent pattern maching, which is something mandatory for writting code correct by construction.
+Especially, Idris has a particularly efficient dependent pattern maching, which is something mandatory for writting code correct by construction.
+
+\subsection{Defining inductive types and functions in Idris}
+
+Idris is all influenced by haskell in its syntax.
+The definition of an inductive type is done using the keyword 'data', followed by the name of the datatype, its type, and the list of constructors.
+For exemple, one can define the natural numbers like that in Idris :
+
+\begin{lstlisting}[caption=inductive datatype, captionpos=b, label=lst1:haskell2]
+{-
+
+member :: Eq a => a -> [a] -> Bool
+
+-}
+
+data MyNat : Type where
+	O : MyNat
+	S : MyNat -> MyNat
+\end{lstlisting}
+
+The definition of a function has exactly the same syntax as Haskell, except the use of ":" for specifying the type of the function, instead of "::", reserved for list concatenation.
+The first reason is that the symbol ":" has been used in type theory since a lot longer than "::" in Haskell.
+The second is that everything has a type, but everything is not a list, so it makes more sence to use the shorter for something that appears regularly.
+Here is for exemple the definition of a plus function :
+
+\begin{lstlisting}[caption=recursive function, captionpos=b, label=lst2:haskell2]
+myplus : MyNat -> MyNat -> MyNat
+myplus O y = y
+myplus (S px) y = S (myplus px y) 
+\end{lstlisting}
+
+\subsection{dependent types}
+Idris supports dependent types. This mainly means two things :
+- inductive datatypes can be indexed over values
+- the type of a function can be such has the type of its codomain dependend on the value of one of its argument.
+
+We will illustrate both with the usual exemple on vector. vector are a list with embedded size : the size is now a part of the datatype.
+
+Because the size is now a part of the datatype, it is now possible to
 
@@
 
-\section{References}
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-More text.
-
-Lots of text.
-
-\appendix
-\section{Appendix Title}
-
-This is the text of the appendix, if you need one.
-
 \acks
@@
 
-\bibliographystyle{abbrvnat}
+\bibliography{biblio}
 
-% The bibliography should be embedded for final submission.
+\appendix
+\section{Appendix Title}
 
-\begin{thebibliography}{}
-\softraggedright
-
-\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
-P. Q. Smith, and X. Y. Jones. ...reference text...
-
-\end{thebibliography}
+This is the text of the appendix, if you need one.
 

### 2013-10-10 paper/content/introduction.tex
added intro and idris overview

\section{Introduction}

A very common situation when using a proof assistant is to need to prove that one term is equal to another within some theories. Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties. These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for example), or by axioms if we are working with an abstract structure.
These proofs of equality are usually done by hand, and are made with a potentially long sequence of rewriting, using the available properties that we have. For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
$(x + y) * z = (z * x) + (y * z)$, we can first use the symmetry of $*$, saying that
$\forall a b, a * b = b * a$.
With the use of this property, we can rewrite the left part into the term $z * (x + y)$.
If now we use the distributivity of $*$ over $+$, which says that :
$\forall a b c, a * (b+c) = a*b + a*c$,
we now obtain the term $x*z + y*z$ for the left side.
If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes : $z*x + y*z$, which is what we want.

Thus, this kind of proofs consists of a potentially long sequence of rewriting, every rewriting step using one property of the theory. Without some specific automation, this sequence of rewriting is done by the user of the proof assistant. This is time consuming, and a little change in the left or the right hand side of the equality can invalidate completely the proof. The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term. In this paper, we describe a certified implementation of a solver for algebraic structures, for the Idris language. Idris is a relatively new purely functional and dependently typed programming language with full dependent types. Idris also supports tactic based proofs.

For our goal, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group, Abelian group, ring and commutative ring.
For these structures, this is effectively possible to decide if a given equality is true, and to produce the proof of the equality is appropriate. This result comes directly from the fact that there exist a normal form for every element of these structures. The general idea is therefore to normalize both side of the equality, and then to compare the resulting terms using the usual syntactical Leibniz's equality.
This approach was followed by [paper Coq ring] for implementing a ring solver for the Coq proof assistant.

Our contribution mainly consists of two parts :
	- We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form. For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters for assuring the correctness of the method. 
	- We try to be as general generic as possible, and re-use as much as possible the code of normalization from the strucures below in the upper structures. For example developing an expression by using the axiom of distributivity works exactly the same in a commutative ring and in a ring, and the code should be reusable.

### 2013-10-10 paper/paper.tex
added intro and idris overview

@@
 %
-%			Edwin's and Franck Paper !
+%			Edwin's and Franck's Paper !
 %  uses the template for sigplanconf LaTeX Class
@@
 \copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
-\preprintfooter{My paper for PLPV’14}
+\preprintfooter{Our paper for PLPV’14}
 \doi{nnnnnnn.nnnnnnn}
@@
 
-\title{Using dependent types for writting correct by construction tactics}
-\subtitle{An exemple for a solver over algebraics structures}
+\title{Proof by reflection for equalities in algebraic structures}
+%\subtitle{An example for a collection of tactics solving equalities in algebraic structures}
 
@@
 
+\nocite{*}
+
 \input{./content/abstract.tex}
 
-\section{Introduction}
-\nocite{*}
+\input{./content/introduction.tex}
 
-As soon as we start to work with numbers on a proof assistant, we are facing the need to prove that a certain term is equal to another.
-Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties.
-These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for exemple), or by axioms if we are working with an abstract structure.
-These proofs of equality are usually done by hand, and are made with a potentially long serie of rewriting, using the available properties that we have.
-For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
-$(x + y) * z = (z * x) = (y * z)$, we can first use the fact the symmetry of $*$, saying that
-for all a b, a * b = b * c
-With the use of this property, we can rewrite the left part into the term z * (x + y).
-If now we use the distributivity of * on +, which says that :
-$\forall a b c, a * (b+c) = a*b + a*c$,
-we now obtain the term x*z + y*z for the left side.
-If we use again the symmetry of * on the subterm (x*z), the left side of the equality becomes : z*x + y*z, which is what we want.
-
-This kind of proofs therefore consists of a potentially long serie of rewriting, every rewriting step using one property of the theory.
-Without some specific automation, this serie of rewriting is done by the user of the proof assistant.
-This is time consuming, and a little change in the left or the right hand side of the equality can invalidate completely the proof.
-The reusability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
-In this paper, we describe a certified implementation of a solver for algebraics structures, for the Idris language.
-Idris is a relatively new purely functional programming language with full dependent types.
-Idris also supports tactic based proofs.
-
-For our goal, we focus on some specific theories, which are the folowing algebraics structures : magmas, semi-group, monoid, group, abelian group, ring and commutative ring.
-For these structures, this is effectively possible to decide if a given equality is true, and to produce the proof of the equality.
-This result comes directly from the fact that there exist a normal form for every element of these structures.
-The general idea is therefore to normalize both side of the equality, and then to compare them using the usual syntactical Leibniz's equality.
-This approach was followed by [paper Coq ring] for implementing a ring solver for the Coq proof assistant.
-Our contribution mainly consits of two parts :
-	- We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procudure, and proving afterwards that this function effectively computes a normal form.
-For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters.
-	- We try to be as general generic as possible, and re-use as much as possible the code of normalization from the strucures below in the upper structures.
-For exemple developping an expression by using the axiom of distribituvty works exactly the same in a commutative ring and in a ring.
-
-\section{A brief overview of Idris}
-
-Idris is a purely functional programming language fairly recent.
-Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not designed to be a proof asssiant, but as a real programming language easily usable by non-logicians.
-One of its more interesting features -and the one that enables to effectively manipulate logical statments and proofs in the language- is its full and effective support for dependent types.
-Especially, Idris has a particularly efficient dependent pattern maching, which is something mandatory for writting code correct by construction.
-
-\subsection{Defining inductive types and functions in Idris}
-
-Idris is all influenced by haskell in its syntax.
-The definition of an inductive type is done using the keyword 'data', followed by the name of the datatype, its type, and the list of constructors.
-For exemple, one can define the natural numbers like that in Idris :
-
-\begin{lstlisting}[caption=inductive datatype, captionpos=b, label=lst1:haskell2]
-{-
-
-member :: Eq a => a -> [a] -> Bool
-
--}
-
-data MyNat : Type where
-	O : MyNat
-	S : MyNat -> MyNat
-\end{lstlisting}
-
-The definition of a function has exactly the same syntax as Haskell, except the use of ":" for specifying the type of the function, instead of "::", reserved for list concatenation.
-The first reason is that the symbol ":" has been used in type theory since a lot longer than "::" in Haskell.
-The second is that everything has a type, but everything is not a list, so it makes more sence to use the shorter for something that appears regularly.
-Here is for exemple the definition of a plus function :
-
-\begin{lstlisting}[caption=recursive function, captionpos=b, label=lst2:haskell2]
-myplus : MyNat -> MyNat -> MyNat
-myplus O y = y
-myplus (S px) y = S (myplus px y) 
-\end{lstlisting}
-
-\subsection{dependent types}
-Idris supports dependent types. This mainly means two things :
-- inductive datatypes can be indexed over values
-- the type of a function can be such has the type of its codomain dependend on the value of one of its argument.
-
-We will illustrate both with the usual exemple on vector. vector are a list with embedded size : the size is now a part of the datatype.
-
-Because the size is now a part of the datatype, it is now possible to
+\input{./content/idrisOverview}
 

### 2013-10-10 paper/content/introduction.tex
added section overview of method

@@
 
-A very common situation when using a proof assistant is to need to prove that one term is equal to another within some theories.
+A very common situation in formal certification is to need to prove that one term is equal to another within some theories.
 Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties.
@@
 The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
-In this paper, we describe a certified implementation of a solver for algebraic structures, for the Idris language.
+In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
 Idris is a relatively new purely functional and dependently typed programming language with full dependent types.
@@
 For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters for assuring the correctness of the method.
-	- We try to be as general generic as possible, and re-use as much as possible the code of normalization from the strucures below in the upper structures.
-For example developing an expression by using the axiom of distributivity works exactly the same in a commutative ring and in a ring, and the code should be reusable.
+	- We try to be as generic as possible, and re-use as much as possible the code of normalization from the structures below in the upper structures.
+For example developing an expression by using the axiom of distributivity works exactly the same in a commutative ring and in a ring, and this part of the reduction should be factorisable.

### 2014-06-11 newPaper/paper.tex
added one file

--- 2013-10-10 paper/paper.tex
@@
 
-\conferenceinfo{PLPV ’14, Co-located with POPL 2014}{January 21, 2014, San Diego, CA, USA.}
-\copyrightyear{2013} 
+\conferenceinfo{POPL ’15}{January 15-17, 2015, Mumbai, India.}
+\copyrightyear{2014} 
 \copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
-\preprintfooter{Our paper for PLPV’14}
+\preprintfooter{Our paper for POPL '15}
 \doi{nnnnnnn.nnnnnnn}
@@
 
-\authorinfo{Edwin Brady\and Franck Slama}
-           {University of St Andrews, United-Kingdom}
-           {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}
+%\authorinfo{Edwin Brady\and Franck Slama}
+%           {University of St Andrews, United-Kingdom}
+%           {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}
+\authorinfo{Author's name omitted for submission}
+           {University of ..., Country}
+           {email address omitted for submission}
+
 \bibliographystyle{abbrvnat}
@@
 
+Plan : \\
+- Intro (1 page): problem + contribution : NOT SPECIFIC ENOUGH YET. TOO GENERAL, TOO BORING. HOW TO CAPTURE THE INTEREST ? \\
+- The problem on a simplified idea (2 pages) : Assoc : that will present the reflection and correct by construction sides on a little example in order to build the intuition.  \\
+- The idea (2 pages)  : the different typeclass AND reflected data types AND functions decideEq \\
+- The details (4/5 pages): Normalizations functions in Magma, Semi-Group, monoid, Group \\ 
+- (1/2 pages) : Coating and automatic metaification \\
+- Related work (1 page): We may be slower (I'll have to check), but this is NICER.
+The proof of correctness of the  implementation of the normalization is trivial to obtain (and this is what we want to get !) \\
+- Conclusion (1 page)
+
 \input{./content/introduction.tex}
-
-\input{./content/idrisOverview}
-
-\section{Normalization and polynomials in normal form}
-% Ici on donne l'idée générale. Nom de la section peut etre à revoir
-Lots of text.
-
-More text.
-
-\section{Datatypes}
-
-Lots of text.
-
-\subsection{Hierarchy of algebraic theorys}
-
-Lots of text.
-
-The exemple with Nat is a monoid
-
-More text.
-
-\subsection{Reflection in an abstract syntax}
-
-Lots of text.
-
-More text.
 

### 2014-06-11 newPaper/content/idrisOverview.tex
new paper just started

\section{A brief overview of Idris}

Idris is a purely functional programming language fairly recent. Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not intended to be a proof assistant, but as a real programming language usable by non-logicians. One of its more interesting features -and the one that enables to effectively manipulate logical statements and proofs in the language- is its full and effective support for dependent types. Especially, Idris has a particularly efficient dependent pattern matching, which is something mandatory for writing code correct by construction.

\subsection{defining inductive types and functions in Idris}

Idris has been a lot influenced by haskell in its syntax.
The definition of an inductive type is done using the keyword 'data', followed by the name of the data type, its type, and the list of constructors.
For example, one can define the natural numbers like that in Idris :

\begin{lstlisting}[caption=inductive data type, captionpos=b, label=lst1:haskell2]
data Nat : Type where
	O : Nat
	S : Nat -> Nat
\end{lstlisting}

The definition of a function has exactly the same syntax as Haskell, except the use of ":" for specifying the type of the function, instead of "::", used for list concatenation, since ":" is the standard notation for that in type theory.
Here is for example the definition of a plus function :

\begin{lstlisting}[caption=recursive function, captionpos=b, label=lst2:haskell2]
plus : Nat -> Nat -> Nat
plus O y = y
plus (S px) y = S (plus px y) 
\end{lstlisting}

\subsection{dependent types}
Idris supports dependent types. This mainly means two things :
\begin{itemize}
\item inductive data types can be indexed over values
\item the type of a function can be such as the type of its co domain depends on the value of one of its argument.
\end{itemize}

We will illustrate both features with the usual example on vector. Vector have the same structure as a list, but with embedded size, which is now a part of the data type.
\begin{lstlisting}[caption=a family of types indexed by a natural number, captionpos=b, label=lst3:haskell2]
data Vect : Nat -> Type -> Type where
	Nil : Vect O a
	(::) : a -> Vect n a -> Vect (S n) a
\end{lstlisting}

Because the size is now a part of the data type, it becomes now possible to write easily a function that operates on vectors and that requires some property about the size of the vectors.
\begin{lstlisting}[caption=a function having a dependent type, captionpos=b, label=lst4:haskell2]
(++) : Vect m a -> Vect n a -> Vect (m + n) a
(++) Nil ys = ys
(++) (x::xs) ys = x :: xs ++ ys
\end{lstlisting}

When we write a function in this way, we say that the type of the function is a precise specification, and this is clearly one of the benefits of dependent types. If one tries to define the function and by mistake write for the second clause :
\begin{lstlisting}[captionpos=b, label=lst5:haskell2]
(++) (x::xs) ys = xs ++ ys
\end{lstlisting}
The compiler will complain with the following error message : "Can't unify (plus n n) with (S (plus n n))". Idris has realized that the returned expression for this clause doesn't have the expected size. Working with dependent types is therefore a good way for ensure properties about functions.

\subsection{type classes}
We often want to define functions which work for several different data types. For example, we would like arithmetic operators to work on natural numbers, relatives number and floating number. We would also like equality to work on the majority of data types. 
Type class are a construct that supports ad hoc polymorphism. This is achieved by adding constraints to type variables in parametrically polymorphic types.
For example, we can define the type class Show, and every structure that we can display on the screen will have to implement this type class :
\begin{lstlisting}[caption=a simple type class, captionpos=b, label=lst6:haskell2]
class Show a where {
	show : a -> String;
}
\end{lstlisting}
This generates a function 
\begin{lstlisting}[captionpos=b, label=lst7:haskell2]
show : Show a => a -> String
\end{lstlisting}
i.e. a function which will print any element of type $a$ under the constraint that $a$ is an instance of Show.

Type classes can be extended. Here is an example with a first type class Eq for equality and a second one called Ord that any sortable structure will have to implement.

\begin{lstlisting}[caption=extension of typeclass, captionpos=b, label=lst8:haskell2]
class Eq a where
	(==) : a -> a -> Bool

data Ordering = LT | EQ | GT

class Eq a => Ord a where
	compare : a -> a -> Ordering
	(<) : a -> a -> Bool
	(>) : a -> a -> Bool
	(<=) : a -> a -> Bool
	(>=) : a -> a -> Bool
	max : a -> a -> a
	min : a -> a -> a
\end{lstlisting}

### 2014-06-11 newPaper/content/introduction.tex
new paper just started

\section{Introduction}

A very common situation in formal certification is to need to prove that one term is equal to another within some theories. Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties. These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for example), or by axioms if we are working with an abstract axiomatised structure. 
These proofs are usually omitted "on the paper" for the ordinary every day maths, but are required by the proof assistant,  no matter how obvious they are. So they are done by hand, and are made with a potentially long sequence of rewriting, using the available properties that we have. For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
$(x + y) * z = (z * x) + (y * z)$, we can first use the symmetry of $*$, saying that
$\forall a b, a * b = b * a$.
With the use of this property, we can rewrite the left part into the term $z * (x + y)$.
If now we use the distributivity of $*$ over $+$, which says that :
$\forall a b c, a * (b+c) = a*b + a*c$,
we now obtain the term $x*z + y*z$ as the left side.
If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes : $z*x + y*z$, which is what we want, ie, the same thing as the right hand side.

Thus, this kind of proofs consists of a potentially long sequence of rewriting, every rewriting step using one property of the theory. Without some specific automation, this sequence of rewriting is done by the user of the proof assistant. This is time consuming, and a little change in the left or the right hand side of the equality can invalidate completely the proof. The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term. In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language. Idris is a relatively new purely functional and dependently typed programming language with full dependent types, which also supports tactic based proofs.

For our goal, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group and Abelian group.
The user will just have to prove that his structure satisfies the properties of an algebraic structure, and he will be allowed to invoke the corresponding tactic.

We will work by reflection for the implementation of each tactic, and the reason for this will be discussed in section [Blabla]. Working by reflection for implementing tactics has been done several times, as in [paper Coq ring] for the implementation of a ring solver for the Coq proof assistant. We will compare with other implementation of this category of tactics in section [Blabla].
% NOT THE RIGHT PLACE FOR THAT
%For these structures, this is effectively possible to decide if a given equality is true, and to produce the proof of the %equality is %appropriate. This result comes directly from the fact that there exist a normal form for every element of %these structures. The general %idea is therefore to normalize both side of the equality, and then to compare the %resulting terms using the usual syntactical Leibniz's %equality.
%Our contribution mainly consists of three points :
\begin{itemize}
	\item 1) We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form. For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters for assuring the correctness of the method. 
	\item 2) This collection of tactics can be written in a "correct by construction" style because we use an original representation of reflected terms which encodes the denoted -concrete- value;
	\item 3) We develop a hierarchy of tactics which enables to be increase a lot the reusability of each solver : we reuse each solver for the solvers of the structures above. For example simplifiying neutral elements is only implemented at the Monoid level, and each level above reuse it. As will discussed in section [blabla], reusability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive. For example, reusing the monoid solver for the group solver is not trivial, since we lose the possibility to express the negations $-x$ and minus $x-y$ operations.
\end{itemize}

### 2014-06-13 newPaper/paper.tex
plan for the paper written

@@
 
-Plan : \\
-- Intro (1 page): problem + contribution : NOT SPECIFIC ENOUGH YET. TOO GENERAL, TOO BORING. HOW TO CAPTURE THE INTEREST ? \\
-- The problem on a simplified idea (2 pages) : Assoc : that will present the reflection and correct by construction sides on a little example in order to build the intuition.  \\
-- The idea (2 pages)  : the different typeclass AND reflected data types AND functions decideEq \\
-- The details (4/5 pages): Normalizations functions in Magma, Semi-Group, monoid, Group \\ 
-- (1/2 pages) : Coating and automatic metaification \\
-- Related work (1 page): We may be slower (I'll have to check), but this is NICER.
-The proof of correctness of the  implementation of the normalization is trivial to obtain (and this is what we want to get !) \\
-- Conclusion (1 page)
-
+Remark : To do : making the introduction more SPECIFIC, with CRUNCHY things : what we would like to do, no general assertions like "we often have to prove equalities...". (cf Simon P J talk)
 \input{./content/introduction.tex}
 
-\section{Normalization procedures}
+\section {The problem explained on a smaller example (1/2 pages)}
 
-\subsection{Normalization for terms in a magma}
+TO DO : With the example of assoc for list, we develop the notion of normalization of terms, before using syntactical equality, in order to decide equality betweens lists (with associative concatenation).
 
-\subsection{Normalization for terms in a semi-group}
+\section {Back to the general problem : A hierachy of tactics (2/3 pages)}
 
-\subsection{Normalization for terms in a monoid}
+	\subsection {Hierarchy of typeclass}
 
-\subsection{Normalization for terms in a group}
+	\subsection {Reflected terms}
 
-\subsection{Normalization for terms in a ring}
+	\subsection {Deciding equality}
 
-\subsection{Normalization for terms in a commutative ring}
+\section {The details : normalization functions (4/5 pages)}
 
-\section{Coating}
-%enrobage
+	\subsection {Normalization of terms in a Magma}
 
-\section{Metaification}
+	\subsection {Normalization of terms in a Semi-Group}
 
-\section{Conclusion}
+	\subsection {Normalization of terms in a Monoid}
+
+	\subsection {Normalization of terms in a Group}
+
+\section {Auxiliary problems : Automatic metaification and coating (1/2 pages)}
+
+	\subsection {Automatic metaification}
+
+	\subsection {Coating : transforming this stuff into real tactics}
+
+\section {Related work (1 page)}
+
+\section {Conlusion (0,5/1 page)}
 
@@
 
-Acknowledgments, if needed.
+Acknowledgments : Chris, Mattus...
 

### 2014-06-16 newPaper/paper.tex
file Autoassoc with change for constant list

@@
 
+We will start with the study of a smaller problem in which we aim to deal with lists, and associativity of list concatenation.
+For example, we would like to be able to generate automatically proofs of facts like $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$. \\
+For this smaller problem, we have decided to only work with the associativity of concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$.
+So, in definitive, in this section, we want to write a decision procedure, able to say if to two expressions composed of lists and concatenation of lists are equal, when "equal" has the meaning "syntactically equal or equal thanks to associativity".
+
+The general idea -which will also apply for the real problem- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
+Of course, the normalization will have to fulfil the following property :
+$\forall l, norm\ l = l$ [1], which means that after normalization we obtain a list which is provably equal to the input's list. \\
+After normalization, we will compare the two resulting lists, and if they are equal, we will have obtained the corresponding proof.  \\
+Indeed :
+$norm\ l1 = norm\ l2$ will imply $l1=l2$ by [1].
+
+In fact, this is not really how it will be, because as for the real problem, we will work by reflection.
+The datatype presented in the Listing 1 is a first approximation what we will use to reflect out terms composed of lists and concatenation of lists.
+
+\begin{lstlisting}[caption=Reflected lists, captionpos=b, label=lst1:haskell2]
+data Expr : Type where
+	App  : Expr -> Expr -> Expr 
+	Var  : (i : Fin n) -> Expr
+    ENil : Expr
+\end{lstlisting}
+
 \section {Back to the general problem : A hierachy of tactics (2/3 pages)}

### 2014-06-18 newPaper/paper.tex
arrrgh

--- 2014-06-13 newPaper/paper.tex
@@
 
-\section {The problem explained on a smaller example (1/2 pages)}
+\section {A simpler problem : a little tactic correct by construction for lists (1/2 pages)}
 
-TO DO : With the example of assoc for list, we develop the notion of normalization of terms, before using syntactical equality, in order to decide equality betweens lists (with associative concatenation).
+We will start with the study of a smaller problem in which we aim to deal with variable lists, and associativity of list concatenation.
+For example, we would like to be able to automatically generate proofs of facts like $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$. \\
+For this smaller problem, we have decided to only work with the property of associativity of the concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$. \\
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of variable lists and concatenation of variable lists are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity".
+
+The general idea -which will also apply for the real bigger problem explained in the next sections- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
+Of course, the normalization will have to fulfil the following property :
+$\forall l, norm\ l = l$ [1], which means that after normalization we obtain a list which is provably equal to the input list (potentially using multiple time the property of associativity!). \\
+After normalization, we will compare the two resulting lists, and if they are equal, we will have obtained the desired proof.  \\
+Indeed :
+$norm\ l1 = norm\ l2$ will imply $l1=l2$ by [1].
+
+In fact, this is not exactly how it will work, because we will work by reflection, as we will do for the real problem later.
+The datatype presented in the Listing 1 is a first approximation of our reflected lists composed of variable lists, concatenation of lists and empty list.
+
+\begin{lstlisting}[caption=Reflected lists, captionpos=b, label=lst1:haskell2]
+data Expr : (Vect n (List a)) -> Type where
+	App  : {n:Nat} -> {G : Vect n (List a)} -> Expr G -> Expr G -> Expr G 
+	Var  : {n:Nat} -> (G : Vect n (List a)) -> (i : Fin n) -> Expr G
+   ENil : {n:Nat} -> (G : Vect n (List a)) -> Expr G 
+\end{lstlisting}
+
+For this smaller problem, we want to prove equalities in which lists are always universally quantified, so we only need to represent variable (a universally quantified list), the constant list Nil, and the concatenation between lists.
+Variables are represent using a "De Brujin" index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
+
+The type Expr is indexed over a vector of lists (we will usually call it G), that represents our context.
+For example, if we want to prove $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, then we will have to encode $(l1 ++ l2) ++ (l3 ++ l4)$ and $(l1 ++ (l2 ++ l3)) + l4$ in a context of four elements.
+The first element of this context denotes the variable $l1$, the second denotes $l2$, and so on.
+Thus, the left hand side will be encoded by :
+\begin{lstlisting}[caption=Reflected LHS, captionpos=b, label=lst1:haskell2]
+App (App (Var G fZ) 
+			(Var G (fS fZ))) 
+    (App (Var G (fS (fS fZ))) 
+    	   (Var G (fS (fS (fS fZ))))).
+\end{lstlisting}
+
+We need now to define what will be the normalization procedure.
+If it is a function which takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards :
+$\forall\ le1,\ interpretation\ (reduce le1)\ =\ interpretation\ le1$
+where $interpretation$ is a function computing the interpretation of an Expr, that is to say, the list this Expr is encoding.
+
+This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way we compute the interpretation.
+For this little example, the reduction will not be too heavy, but in the next sections, when we will have more properties than only associativity to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+\\
+To avoid these two sources of complexity, we will :
+- Use dependent types in order to directly capture the concrete list that an Expr is encoding
+- Write a "correct by construction" reduction procedure.
+The proof of preservation of the interpretation will be computed bit by bit at the same time as the function will produce the normalized expression.
+
+\begin{lstlisting}[caption=New reflected lists, captionpos=b, label=lst1:haskell2]
+using (x : List a, y : List a, G : Vect n (List a))
+  data Expr : (G : Vect n (List a)) -> List a -> Type where
+       App  : Expr G x -> Expr G y -> Expr G (x ++ y)
+       Var  : (i : Fin n) -> Expr G (index i G)
+       ENil : Expr G []
+\end{lstlisting}
+
+The reduction procedure is now supposed to be written on a "correct by construction" way, which mean that no additional proof should required after the definition of the function.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indewed over the real - concrete list : a term of type $Expr G l$ is the encoding of the list $l$.
+Thus, we can write the type of $reduce$ like this : \\
+$reduce\ :\ Expr\ G\ x\ ->\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
+The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair composed by an $Expr G x'$ which is the new encoded term indexed iver the new concrete list, and a proof that old and new lists are equal.
 

### 2014-06-19 newPaper/paper.tex
some progress on the smaller problem

--- 2014-06-13 newPaper/paper.tex
@@
 
-\section {The problem explained on a smaller example (1/2 pages)}
+\section {A simpler problem : a small correct by construction tactic for lists (1/2 pages)}
 
-TO DO : With the example of assoc for list, we develop the notion of normalization of terms, before using syntactical equality, in order to decide equality betweens lists (with associative concatenation).
+In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universaly quantified lists, and associativity of list concatenation.
+For example, we would like to be able to automatically generate proofs of facts like $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$ [Example 1]. \\
+For this smaller problem, we have decided to only work with the property of associativity of the concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$. \\
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified lists and concatenation of these lists are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity".
+\\
+
+When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
+For the example 1, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $l1=l2$.
+The general idea -which will also apply for the real bigger problem explained in the next sections- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
+The goal of the normalization is of course to compute a canonical representation for a list $l$, such that any other list provably equal to $l$ will have the same canonical representation.
+That's why it will be possible to decide the equality by simply comparing the two normal forms with a week syntactical equality.
+\\
+\\
+Of course, the normalization will have to fulfil the following correctness property :
+$\forall l, norm\ l = l$ [1], which means that after normalization we obtain a list which is provably equal to the input list (potentially using the property of associativity multiple times). \\
+And after normalization, we will compare the two resulting lists with a standard syntactical equality, and if they are equal, we will have obtained the desired proof.  \\
+Indeed :
+$norm\ l1 = norm\ l2$ will imply $l1=l2$ by using the correctness lemma [1] of $norm$ two times. \\
+\\
+In fact, this is not exactly how it will work.
+This way would not be easily feasible, because in the LHS and RHS of $l1=l2$, we have potentially variables which have been universally quantified.
+And the normalization function would need to do different treatments for a "variable list" (a list which has been universally quantified) and for a constant, empty list.
+This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality.
+For this reason, we will work by reflection, as we will do for for the real problem later.
+That means that we will define a datatype which will be used as an encoding of lists, or more precisely, an encoding of lists composed by "variable lists", Nil, and concatenation of the previous.
+This datatype will allow us to know the internal structure of the list, ie, where the variables and constants are.
+Indeed, this datatype will allow pattern matching.
+Previously, we were only able to pattern match a list against the constructor $Nil$ and $Cons$, which wasn't interesting for our goal.
+With the first approximation of the datatype $Expr$ presentend in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and Enil, which give us the informations we want.
+
+\begin{lstlisting}[caption=First version of reflected lists, captionpos=b, label=lst1:haskell2]
+data Expr : (Vect n (List a)) -> Type where
+  App  : {n:Nat} -> {G : Vect n (List a)} -> 
+         Expr G -> Expr G -> Expr G 
+  Var  : {n:Nat} -> (G : Vect n (List a)) -> 
+         (i : Fin n) -> Expr G
+  ENil : {n:Nat} -> (G : Vect n (List a)) -> 
+         Expr G 
+\end{lstlisting}
+
+For this smaller problem, we want to prove equalities in which lists are always universally quantified, so we only need to represent variable (a universally quantified list), the constant list Nil, and the concatenation between lists.
+Variables are represent using a De Brujin-like index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
+
+The type Expr is indexed over a vector of lists (we will usually name it G), that represents our context.
+For example, if in the example 1, we will have to encode $(l1 ++ l2) ++ (l3 ++ l4)$ and $(l1 ++ (l2 ++ l3)) + l4$ in a context where four elements are declared.
+The first element of this context denotes the variable $l1$, the second denotes $l2$, and so on.
+Thus, the left hand side will be encoded by :
+\begin{lstlisting}[caption=Reflected LHS of example 1, captionpos=b, label=lst1:haskell2]
+App (App (Var G fZ) 
+         (Var G (fS fZ))) 
+    (App (Var G (fS (fS fZ))) 
+         (Var G (fS (fS (fS fZ))))).
+\end{lstlisting}
+
+We need now to define what will be the normalization procedure.
+If it is, as probably expected, a function which takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards : \\
+$\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
+where $interpretation$ is a function computing the interpretation of an Expr, that is to say, the list this Expr is encoding.
+This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way we compute the interpretation.
+For this little example, the reduction will not be too heavy, but in the next sections, when we will have more properties than only associativity to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+\\
+\\
+To avoid these two sources of complexity, we will : \\
+- Use dependent types in order to directly capture and embed the concrete list that an Expr is encoding \\
+- Write a "correct by construction" reduction procedure.
+The proof of preservation of the interpretation will be computed "bit by bit" at the same time as the function will produce the normalized expression (by rewriting the term thanks to the property of associativity).
+
+\begin{lstlisting}[caption=New reflected lists, captionpos=b, label=lst1:haskell2]
+using (x : List a, y : List a, 
+       G : Vect n (List a))
+  data Expr : (G : Vect n (List a)) -> 
+              List a -> Type where
+       App  : Expr G x -> Expr G y -> 
+              Expr G (x ++ y)
+       Var  : (i : Fin n) -> 
+              Expr G (index i G)
+       ENil : Expr G []
+\end{lstlisting}
+For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the list $x$".
+When an expression is a variable, the denoted list is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
+Also, the $ENil$ expression denotes the $Nil$ list.
+Finally, if $e1$ is an expression encoding the list $x$, and $e2$ is an expression encoding the list $y$, then the expression $App\ e1\ e2$ denotes the list $(x ++ y)$.
+
+The reduction procedure is now supposed to be written on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real (concrete) list : a term of type $Expr\ G\ l$ is the encoding of the list $l$.
+Thus, we can write the type of $reduce$ like this : \\
+$reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
+The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete list, and a proof that old and new lists are equal.
+Note that this function can't simply produce an $Expr\ G\ x$, because the list on which the resulting expression will be indexed is not necessary syntactically equal to the original list. 
+In fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
+The reason is simple : when we try to automatically prove $x=y$, we will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$.
+We will normalize $e1$, and this will give a new list $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
+We will do the same with $e2$, and we will get a new list $y$, an expression $e2':Expr\ G\ y'$, and a proof of $y=y'$.
+\\
+Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality : \\
+$eqExpr\ :\ (e\ :\ Expr\ G\ x)\ \rightarrow\ (e'\ :\ Expr\ G\ y)\ \rightarrow\ Maybe\ (e= e')$ \\
+And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{we are working with the hetherogenous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
+By rewriting the two previously obtained equality $x=x'$ and $y=y'$ in the new equality $x'=y'$, we can get a proof of $x=y$.
 

### 2014-06-19 newPaper/paper.tex
progress on the paper

@@
 This way would not be easily feasible, because in the LHS and RHS of $l1=l2$, we have potentially variables which have been universally quantified.
-And the normalization function would need to do different treatments for a "variable list" (a list which has been universally quantified) and for a constant, empty list.
-This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality.
+And the normalization function would need to do different treatments for a "variable list" (a list which has been universally quantified) and for a constant empty list.
+This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality : this is a meta information of the AST that we can't access.
 For this reason, we will work by reflection, as we will do for for the real problem later.
@@
 Previously, we were only able to pattern match a list against the constructor $Nil$ and $Cons$, which wasn't interesting for our goal.
-With the first approximation of the datatype $Expr$ presentend in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and Enil, which give us the informations we want.
+With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and Enil, which give us the informations we want.
 
@@
 We need now to define what will be the normalization procedure.
-If it is, as probably expected, a function which takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards : \\
+If, as it is probably expected, this function simply takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards, corresponding to the correcness property mentioned before : \\
 $\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
@@
 To avoid these two sources of complexity, we will : \\
-- Use dependent types in order to directly capture and embed the concrete list that an Expr is encoding \\
+- Use dependent types in order to directly capture and embed the concrete list that an Expr is encoding. Thus, it won't be necessary to define the $interpretation$ function. \\
 - Write a "correct by construction" reduction procedure.
@@
 $reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
-The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete list, and a proof that old and new lists are equal.
+The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete list, and a proof that old and new -real- lists are equal.
 Note that this function can't simply produce an $Expr\ G\ x$, because the list on which the resulting expression will be indexed is not necessary syntactically equal to the original list. 
 In fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
-The reason is simple : when we try to automatically prove $x=y$, we will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$.
+The reason is simple : when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
+\\
+We will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
 We will normalize $e1$, and this will give a new list $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
@@
 \\
-Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality : \\
-$eqExpr\ :\ (e\ :\ Expr\ G\ x)\ \rightarrow\ (e'\ :\ Expr\ G\ y)\ \rightarrow\ Maybe\ (e= e')$ \\
-And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{we are working with the hetherogenous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
+Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality :
+\begin{lstlisting}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
+  eqExpr : (e : Expr G x) -> (e' : Expr G y) 
+  	       -> Maybe (e = e')
+  eqExpr (App x y) (App x' y') 
+  		with (eqExpr x x', eqExpr y y')
+    eqExpr (App x y) (App x y)   
+    	| (Just refl, Just refl) = Just refl
+    eqExpr (App x y) (App x' y') 
+    	| _ = Nothing
+  eqExpr (Var i) (Var j) with (decEq i j)
+    eqExpr (Var i) (Var i) 
+    	| (Yes refl) = Just refl
+    eqExpr (Var i) (Var j) 
+    	| _ = Nothing
+  eqExpr ENil ENil = Just refl
+  eqExpr _ _ = Nothing
+\end{lstlisting}
+
+And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{we are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
 By rewriting the two previously obtained equality $x=x'$ and $y=y'$ in the new equality $x'=y'$, we can get a proof of $x=y$.
+
+\begin{lstlisting}[caption=Build proof, captionpos=b, label=lst1:haskell2]
+  buildProof : {x : List a} -> {y : List a} 
+               -> Expr G ln -> Expr G rn ->
+               (x = ln) -> (y = rn) ->
+               Maybe (x = y) 
+  buildProof e e' lp rp with (eqExpr e e')
+    buildProof e e lp rp  | Just refl = ?bp1
+    buildProof e e' lp rp | Nothing = Nothing
+\end{lstlisting}
+As we said, the proof for the metavariable is just the rewriting of the two equalities :
+\begin{lstlisting}[caption=Build proof metavirable, captionpos=b, label=lst1:haskell2]
+  bp1 = proof {
+  intros;
+  refine Just;
+  rewrite sym lp;
+  rewrite sym rp;
+  trivial;
+}  
+\end{lstlisting}
+Finally, the main function which tries to prove the equality $x=y$ is simply made of the reduction of the two metaified terms reflecting the left and the right hand side, and of the composition of the two obtained proofs thanks to $buildProof$:
+\begin{lstlisting}[caption=testEq, captionpos=b, label=lst1:haskell2]
+  testEq : Expr G x -> Expr G y 
+           -> Maybe (x = y)
+  testEq l r = let (ln ** (l', lPrf)) = reduce l in 
+               let (rn ** (r', rPrf)) = reduce r in
+                   buildProof l' r' lPrf rPrf
+\end{lstlisting}
+Now, we need to define the function reduce.
+To do that, we have to fix a canonical representation of associative lists.
+We decide that the right associative form will be the canonical representation.
+Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform it in the form $l1 ++ (l2 ++ (l3 ++(...
+ln)))$.
+To do so, we can define a new datatype which captures this property.
+Thus it will be easier to get sure that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of depend types when, like here, it doesn't introduce many complications.
+\begin{lstlisting}[caption=New reflected lists, captionpos=b, label=lst1:haskell2]
+  data RExpr : (G : Vect n (List a)) -> 
+                List a -> Type where
+       RApp : RExpr G x -> (i : Fin n) -> 
+              RExpr G (x ++ index i G)
+       RNil : RExpr G []
+\end{lstlisting}
+This datatype has only two constructors. In fact, it combines the previous $Var$ and $App$ constructors so that it becomes impossible to write an expression which isn't right associative.
 

### 2014-06-20 newPaper/content/introduction.tex
paper continued

@@
 we now obtain the term $x*z + y*z$ as the left side.
-If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes : $z*x + y*z$, which is what we want, ie, the same thing as the right hand side.
+If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes : $z*x + y*z$, and that finishes the proof because it is the same thing as the right hand side.
 
@@
 Without some specific automation, this sequence of rewriting is done by the user of the proof assistant.
-This is time consuming, and a little change in the left or the right hand side of the equality can invalidate completely the proof.
+This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
 The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
+
+Instead, what we would like to have :
+INSERT CRUCHY EXAMPLE HERE : 
+a = b by cool\_tactic
+
 In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
 Idris is a relatively new purely functional and dependently typed programming language with full dependent types, which also supports tactic based proofs.
+Idris will be briefly defined the the [Section X] (at the end of the paper)
 
@@
 We will compare with other implementation of this category of tactics in section [Blabla].
+\\
 % NOT THE RIGHT PLACE FOR THAT
@@
 The general %idea is therefore to normalize both side of the equality, and then to compare the %resulting terms using the usual syntactical Leibniz's %equality.
-%Our contribution mainly consists of three points :
+\\
+Our contribution mainly consists of three points :
 \begin{itemize}

### 2014-06-20 newPaper/paper.tex
paper continued

@@
 
-\section {A simpler problem : a small correct by construction tactic for lists (1/2 pages)}
+\section {A smaller problem : a correct by construction tactic for the associativity of lists concatenation}
 
@@
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-For the example 1, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $l1=l2$.
-The general idea -which will also apply for the real bigger problem explained in the next sections- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
-The goal of the normalization is of course to compute a canonical representation for a list $l$, such that any other list provably equal to $l$ will have the same canonical representation.
+For the example 1, after abstraction of the variables, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $l1=l2$.
+The general idea -which will also apply for the real bigger problem detailed in the next sections- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
+The goal of the normalization is of course to compute a canonical representation for a list $l$, such that any other list provably equal to $l$ (by using the associativity) will have the same canonical representation.
 That's why it will be possible to decide the equality by simply comparing the two normal forms with a week syntactical equality.
@@
 In fact, this is not exactly how it will work.
-This way would not be easily feasible, because in the LHS and RHS of $l1=l2$, we have potentially variables which have been universally quantified.
+This way would not be easily feasible, because in the LHS and RHS of $l1=l2$, we potentially have variables which have been universally quantified.
 And the normalization function would need to do different treatments for a "variable list" (a list which has been universally quantified) and for a constant empty list.
@@
 Previously, we were only able to pattern match a list against the constructor $Nil$ and $Cons$, which wasn't interesting for our goal.
-With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and Enil, which give us the informations we want.
+With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and ENil, which gives us the informations we want.
 
@@
 
-\begin{lstlisting}[caption=New reflected lists, captionpos=b, label=lst1:haskell2]
+\begin{lstlisting}[caption=Second version of reflected lists with embedded denotation, captionpos=b, label=lst1:haskell2]
 using (x : List a, y : List a, 
@@
 The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete list, and a proof that old and new -real- lists are equal.
-Note that this function can't simply produce an $Expr\ G\ x$, because the list on which the resulting expression will be indexed is not necessary syntactically equal to the original list. 
+Note that this function can't simply produce an $Expr\ G\ x$, because the list on which the resulting expression will be indexed is not necessary syntactically equal to the original list since it can use the property of associativity.
 In fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
@@
 \\
-Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality :
+Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality because they are in their normal form :
 \begin{lstlisting}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
@@
 
-And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{we are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
+And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
 By rewriting the two previously obtained equality $x=x'$ and $y=y'$ in the new equality $x'=y'$, we can get a proof of $x=y$.
 
-\begin{lstlisting}[caption=Build proof, captionpos=b, label=lst1:haskell2]
+\begin{lstlisting}[caption=Building the proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
   buildProof : {x : List a} -> {y : List a} 
@@
            -> Maybe (x = y)
-  testEq l r = let (ln ** (l', lPrf)) = reduce l in 
+  testEq l r = 
+     let (ln ** (l', lPrf)) = reduce l in 
      let (rn ** (r', rPrf)) = reduce r in
@@
 To do that, we have to fix a canonical representation of associative lists.
-We decide that the right associative form will be the canonical representation.
-Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform it in the form $l1 ++ (l2 ++ (l3 ++(...
-ln)))$.
+We decide that the left associative form will be the canonical representation.
+Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform the underlying list in the form $(...((l1 ++ l2) ++ l3) ...
+++ ln)$.
 To do so, we can define a new datatype which captures this property.
-Thus it will be easier to get sure that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of depend types when, like here, it doesn't introduce many complications.
-\begin{lstlisting}[caption=New reflected lists, captionpos=b, label=lst1:haskell2]
-  data RExpr : (G : Vect n (List a)) -> 
+Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.
+\begin{lstlisting}[caption=New reflected lists : left associative, captionpos=b, label=lst1:haskell2]
+  data LExpr : (G : Vect n (List a)) -> 
                List a -> Type where
-       RApp : RExpr G x -> (i : Fin n) -> 
-              RExpr G (x ++ index i G)
-       RNil : RExpr G []
+       LApp : LExpr G x -> (i : Fin n) -> 
+              LExpr G (x ++ index i G)
+       LNil : LExpr G []
 \end{lstlisting}
-This datatype has only two constructors. In fact, it combines the previous $Var$ and $App$ constructors so that it becomes impossible to write an expression which isn't right associative.
+This datatype has only two constructors. In fact, it combines the previous $Var$ and $App$ constructors so that it becomes impossible to write an expression which isn't left associative.
+
+As part of the normalization, we will write a function $expr\_l$ converting an $Expr\ G\ x$ to a $LExpr\ G\ x'$ and producing a proof that $x=x'$.
+This function will therefore use the property of associativity multiple times, in order to obtain the expected fully left associative form.
+Of course, because by using this new datatype $LExpr$ we've changed the representation of our encoded lists, we will need to convert back an $LExpr\ G\ x$ to an $Expr\ G\ x$.
+The function $l\_expr$ will do this easy task.
+
+\begin{lstlisting}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
+  expr_l : Expr G x -> 
+               (x' ** (LExpr G x', x = x'))
+  expr_l ENil = (_ ** (LNil, refl))
+  expr_l (Var i) = (_ ** (LApp RNil i, refl))
+  expr_l (App ex ey) = 
+  	let (xl ** (xr, xprf)) = expr_l ex in
+  	let (yl ** (yr, yprf)) = expr_l ey in
+   	   appLExpr _ _ xr yr xprf yprf
+    where 
+       appLExpr : (x', y' : List a) ->
+             LExpr G x -> LExpr G y -> 
+             (x' = x) -> (y' = y) ->
+             (w' ** (LExpr G w', x' ++ y' = w'))
+       appLExpr x' y' rxs (LApp e i) xprf yprf =
+          let (xs ** (rec, prf)) = 
+             appLExpr _ _ rxs e refl refl in
+             (_ ** (LApp rec i, ?appLExpr1))
+       appLExpr x' y' rxs LNil xprf yprf = 
+          (_ ** (rxs, ?appLExpr2))
+
+  l_expr : LExpr G x -> Expr G x
+  l_expr LNil = ENil
+  l_expr (LApp xs i) = App (r_expr xs) (Var i)
+\end{lstlisting}
+
+We can now define the reduction, which is just the composition of the two previous functions :
+
+\begin{lstlisting}[caption=Reduction function, captionpos=b, label=lst1:haskell2]
+  reduce : Expr G x -> (x' ** (Expr G x', x = x'))
+  reduce e = let (x' ** (e', prf)) = expr_l e in
+                 (x' ** (l_expr e', prf))
+\end{lstlisting}
+
+ADDING EXAMPLE OF HOW TO USE THIS LITTLE TACTIC HERE.
 

### 2014-06-20 newPaper/paper.tex
paper continued... again...

@@
 \usepackage{listings}
+\usepackage[literate]{idrislang} 
 \newcommand{\cL}{{\cal L}}
+
+%\lstset {captionpos=b}
 
@@
 For example, we would like to be able to automatically generate proofs of facts like $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$ [Example 1]. \\
-For this smaller problem, we have decided to only work with the property of associativity of the concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$. \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified lists and concatenation of these lists are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity".
+For this smaller problem, we have decided to only work with the associativity of the concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$ and with the fact that Nil is a neutral element for the concatenation : $\forall l, l ++ Nil = l$.
+Of course, this work could be extended by dealing with more axioms, and this is exactly what we will do in the next sections when we will generalize this problem.
+\\
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified lists and concatenation of these lists are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the "neutrality" of Nil".
 \\
@@
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-For the example 1, after abstraction of the variables, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $l1=l2$.
-The general idea -which will also apply for the real bigger problem detailed in the next sections- will be to normalize both sides of the "potential equality" $l1=l2$, and afterwards to compare them using Leibniz syntactical equality.
+For the example 1, after abstraction of the variables, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $x=y$.
+The general idea -which will also apply for the real bigger problem detailed in the next sections- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
 The goal of the normalization is of course to compute a canonical representation for a list $l$, such that any other list provably equal to $l$ (by using the associativity) will have the same canonical representation.
@@
 Indeed :
-$norm\ l1 = norm\ l2$ will imply $l1=l2$ by using the correctness lemma [1] of $norm$ two times. \\
+$norm\ x = norm\ y$ will imply $x=y$ by using the correctness lemma [1] of $norm$ two times. \\
 \\
@@
 For this reason, we will work by reflection, as we will do for for the real problem later.
-That means that we will define a datatype which will be used as an encoding of lists, or more precisely, an encoding of lists composed by "variable lists", Nil, and concatenation of the previous.
+That means that we will define a datatype which will be used as an encoding of lists, or more precisely, an encoding of lists composed of "variable lists", Nil, and concatenation of the previous.
 This datatype will allow us to know the internal structure of the list, ie, where the variables and constants are.
@@
 
-\begin{lstlisting}[caption=First version of reflected lists, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=First version of reflected lists, captionpos=b, label=lst1:haskell2]
 data Expr : (Vect n (List a)) -> Type where
@@
          Expr G 
-\end{lstlisting}
+\end{code}
 
@@
 Thus, the left hand side will be encoded by :
-\begin{lstlisting}[caption=Reflected LHS of example 1, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=Reflected LHS of example 1, captionpos=b, label=lst1:haskell2]
 App (App (Var G fZ) 
@@
     (App (Var G (fS (fS fZ))) 
-         (Var G (fS (fS (fS fZ))))).
-\end{lstlisting}
+         (Var G (fS (fS (fS fZ)))))
+\end{code}
 
@@
 
-\begin{lstlisting}[caption=Second version of reflected lists with embedded denotation, captionpos=b, label=lst1:haskell2]
+The type for reflected lists therefore becomes :
+\begin{code}[caption=Second version of reflected lists with embedded denotation, captionpos=b, label=lst1:haskell2]
 using (x : List a, y : List a, 
@@
        ENil : Expr G []
-\end{lstlisting}
+\end{code}
 For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the list $x$".
@@
 Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality because they are in their normal form :
-\begin{lstlisting}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
+
+\begin{code}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
   eqExpr : (e : Expr G x) -> (e' : Expr G y) 
@@
   eqExpr _ _ = Nothing
-\end{lstlisting}
+\end{code}
 
-And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is if A=B}, and therefore $x'=y'$
+And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$
 By rewriting the two previously obtained equality $x=x'$ and $y=y'$ in the new equality $x'=y'$, we can get a proof of $x=y$.
 
-\begin{lstlisting}[caption=Building the proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=Building the desired proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
   buildProof : {x : List a} -> {y : List a} 
@@
     buildProof e e' lp rp | Nothing = Nothing
-\end{lstlisting}
+\end{code}
+
 As we said, the proof for the metavariable is just the rewriting of the two equalities :
-\begin{lstlisting}[caption=Build proof metavirable, captionpos=b, label=lst1:haskell2]
+
+\begin{code}[caption=buildProof metavariable, captionpos=b, label=lst1:haskell2]
   bp1 = proof {
@@
   rewrite sym rp;
-  trivial;
+  exact refl;
 }  
-\end{lstlisting}
-Finally, the main function which tries to prove the equality $x=y$ is simply made of the reduction of the two metaified terms reflecting the left and the right hand side, and of the composition of the two obtained proofs thanks to $buildProof$:
-\begin{lstlisting}[caption=testEq, captionpos=b, label=lst1:haskell2]
+\end{code}
+Finally, the main function which tries to prove the equality $x=y$ is simply made of the reduction of the two metaified terms reflecting the left and the right hand side, and of the composition of the two obtained proofs thanks to the function $buildProof$:
+\begin{code}[caption=testEq, captionpos=b, label=lst1:haskell2]
   testEq : Expr G x -> Expr G y 
@@
         buildProof l' r' lPrf rPrf
-\end{lstlisting}
+\end{code}
 Now, we need to define the function reduce.
@@
 ++ ln)$.
-To do so, we can define a new datatype which captures this property.
+To do so, one possibility is to define a new datatype which captures this property, and to write a function going from $Expr$ to this new type.
 Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.
-\begin{lstlisting}[caption=New reflected lists : left associative, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=Reflected left associative lists, captionpos=b, label=lst1:haskell2]
   data LExpr : (G : Vect n (List a)) -> 
@@
        LNil : LExpr G []
-\end{lstlisting}
+\end{code}
 This datatype has only two constructors. In fact, it combines the previous $Var$ and $App$ constructors so that it becomes impossible to write an expression which isn't left associative.
@@
 
-\begin{lstlisting}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
   expr_l : Expr G x -> 
@@
   expr_l ENil = (_ ** (LNil, refl))
-  expr_l (Var i) = (_ ** (LApp RNil i, refl))
+  expr_l (Var i) = (_ ** 
+                   (LApp RNil i, refl))
   expr_l (App ex ey) = 
@@
   l_expr (LApp xs i) = App (r_expr xs) (Var i)
-\end{lstlisting}
+\end{code}
 
-We can now define the reduction, which is just the composition of the two previous functions :
+We notice that for transforming the list into its left associative equivalent representation, we've effectively needed to know where the variables and the Nil constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
+\\
+\\
+We've got two metavariables to prove.
+The metavariable $appRExpr1$ requires us to prove the goal : $x' ++ y' = xs ++ index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x2\ ++\ index\ i\ G)$ and $(prf\ :\ x\ ++\ x2\ =\ xs)$.
+Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
 
-\begin{lstlisting}[caption=Reduction function, captionpos=b, label=lst1:haskell2]
-  reduce : Expr G x -> (x' ** (Expr G x', x = x'))
-  reduce e = let (x' ** (e', prf)) = expr_l e in
+\begin{code}[caption=Proof of the metavariable appRExpr1, captionpos=b, label=lst1:haskell2]
+AutoAssoc.appRExpr1 = proof {
+  intros;
+  rewrite sym xprf;
+  rewrite sym yprf;
+  rewrite prf;
+  rewrite sym 
+       (appendAssociative x x2 (index i G));
+  exact refl;
+}
+\end{code}
+
+And the metavariable appRExpr2 uses the fact that Nil is a neutral element for the concatenation.
+
+\begin{code}[caption=Proof of the metavariable appRExpr2, captionpos=b, label=lst1:haskell2]
+appRExpr2 = proof {
+  intros;
+  rewrite xprf;
+  rewrite sym yprf;
+  rewrite appendNilRightNeutral x';
+  exact refl;
+}
+\end{code}
+
+We can now define the reduction, which is just the composition of the two previous functions $expr\_l$ and $l\_expr$:
+
+\begin{code}[caption=Reduction function, captionpos=b, label=lst1:haskell2]
+  reduce : Expr G x -> 
+           (x' ** (Expr G x', x = x'))
+  reduce e = 
+     let (x' ** (e', prf)) = expr_l e in
         (x' ** (l_expr e', prf))
-\end{lstlisting}
+\end{code}
 
-ADDING EXAMPLE OF HOW TO USE THIS LITTLE TACTIC HERE.
+At the moment, what we've got is not exactly a real tactic, in the sense that we only have a function which produces a value of type $Maybe (x = y)$.
+A real tactic would be a coating of this function, which could properly fail when the two terms are not equal.
+However, here, when $x\ne y$, the function $testEq$ will simply produce the value $Nothing$.
+\\
+It's now time to see how to use this minimalist "tactic".
+Let's define two expressions $e1$ and $e2$, respectively representing the lists $((xs ++ ys) ++ (xs ++ zs))$ and $(xs ++ ((ys ++ xs) ++ zs))$ in the context $[xs, ys, zs]$ of three abstracted variables.
+
+\begin{code}[caption=Two test expressions, captionpos=b, label=lst1:haskell2]
+  e1 : (xs, ys, zs : List a) -> 
+           Expr [xs, ys, zs] 
+           ((xs ++ ys) ++ (xs ++ zs))
+  e1 xs ys zs = 
+     App (App (Var fZ) (Var (fS fZ))) 
+         (App (Var fZ) (Var (fS (fS fZ))))
+
+  e2 : (xs, ys, zs : List a) -> 
+           Expr [xs, ys, zs] 
+           (xs ++ ((ys ++ xs) ++ zs))
+  e2 xs ys zs = 
+     App (Var fZ) 
+         (App (App (Var (fS fZ)) (Var fZ)) 
+              (Var (fS (fS fZ))))
+\end{code}
+
+The lists denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this by using $testEq$.
+\begin{code}[caption=Test of equality betwen e1 and e2, captionpos=b, label=lst1:haskell2]
+  e1_e2_testEq : (xs, ys, zs : List a) ->
+          Maybe (((xs ++ ys) ++ (xs ++ zs)) = 
+                 (xs ++ ((ys ++ xs) ++ zs)))
+  e1_e2_testEq xs ys zs = 
+     testEq (e1 xs ys zs) (e2 xs ys zs) 
+\end{code}
+
+And if we ask for the evaluation of this term, we should obtain $Just$ and a proof of equality between the two underlying lists.
+\begin{code}[caption=Evaluation of the result, captionpos=b, label=lst1:haskell2]
+\xs => \ys => \zs => e1_e2_testEq {a=Int} xs ys zs
+
+\xs => \ys => \zs => Just (replace (sym (replace 
+(sym (replace (replace (appendNilRightNeutral xs)
+ refl (replace (sym (appendAssociative xs [] ys)) 
+ refl))) (replace (sym (replace (replace 
+ (appendNilRightNeutral [...] refl)))) refl)) 
+: (xs : List Int) -> (ys : List Int) -> 
+  (zs : List Int) -> 
+  Maybe ((xs ++ ys) ++ xs ++ zs 
+         = xs ++ (ys ++ xs) + zs)
+\end{code}
+
+And we effectively get what we wanted : a proof of equality between the two lists.
+As expected, this proof uses the properties of associativity ($appendAssociative$) and the property of neutrality of $Nil$ for $append$ ($appendNilRightAssociative$).
 
@@
 
-Acknowledgments : Chris, Mattus...
+Acknowledgments : Chris, Mattus for discussions. Jan for help in latex.
 

### 2014-06-24 newPaper/paper.tex
paper continued

@@
 
-\section {Back to the general problem : A hierachy of tactics (2/3 pages)}
+\section {Back to the general problem : A hierarchy of tactics (2/3 pages)}
+
+Now that we have introduced our ideas on this little example, it is time to apply them for our main goal : the implementation of a hierarchy of tactics proving equalities in algebraic structures.
+Very often, the properties available on a given type are the one of a well known structure : magma, semi-group, monoid, group, commutative-group...
+We will no longer only work with the properties of associativity and neutral elements for Lists as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.
+In a Magma, we only have a $Plus$ operation, and no axioms with it.
+With a Semi-Group, we will have one axiom about this Plus : the property of associativity.
+In a Monoid, we will have one more property : the fact that we have a neutral element.
+In fact, this property will be expressed as 3 facts : there is a value $Zero$, and we've got $\forall x, Zero + x = x$ and $\forall x, x + Zero = x$.
+\\
+
+We will define a hierarchy of typeclasses.
+All our tactics (the group solver, the monoid solver...) will be able to work on any type, being given that an instance of the corresponding typeclass is provided.
 
 \subsection {Hierarchy of typeclass}
+
+All our tactics will require to have a way of testing the equality between elements of the set, that is to say, a way to test equality between constants.
+For this reason, we define a notion of "set", which only requires this equality.
+All the algebraic structures above will extend this typeclass :
+
+\begin{code}[caption=Set, captionpos=b, label=lst1:haskell2]
+class Set c where
+    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
+\end{code}
+
+Note that this equality is a "week" equality : it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different.
+That's quite natural, since we want to generate proof of equality, and not to generate counter examples for dis-equalities, which is another problem.
+
+Of course, there will be no tactic associated to $Set$, since we have no operations and not properties associated to this structure.
+Therefore, the equality in a $Set$ is just the syntactical equality, and can be proven with $refl$.
+
+The first structure, almost trivial, is the magma.
+A magma has just a $Plus$ operation which let us computing constants, and no axioms about this operation.
+Thus, as we will see later, the normalization will be simply a computation between constants inside parenthesis.
+
+\begin{code}[caption=Magma, captionpos=b, label=lst1:haskell2]   
+class Set c => Magma c where
+    Plus : c -> c -> c
+\end{code}
+
+bbb
+
+\begin{code}[caption=Semi-Group, captionpos=b, label=lst1:haskell2]   
+class Magma c => SemiGroup c where
+    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) -> (Plus (Plus c1 c2) c3 = Plus c1 (Plus c2 c3))
+\end{code}
+
+ccc
+
+\begin{code}[caption=Monoid, captionpos=b, label=lst1:haskell2]   
+class SemiGroup c => Monoid c where
+	Zero : c    
+    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 = c1)    
+    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero = c1)
+\end{code}
+
+ddd
+
+\begin{code}[caption=Group, captionpos=b, label=lst1:haskell2]   
+hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) -> c -> c -> Type
+hasSymmetric c p a b = (Plus a b = Zero, Plus b a = the c Zero)
+
+class dataTypes.Monoid c => dataTypes.Group c where
+	Minus : c -> c ->
+	Neg : c -> c
+    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) --Minus should not be primitive and should be simplifiable
+    Plus_inverse : (c1:c) -> hasSymmetric c (%instance) c1 (Neg c1) -- Every element 'x' has a symmetric which is (Neg x)
+\end{code}
 

### 2014-06-25 newPaper/paper.tex
changing name VariableA to Variable

@@
 
-bbb
+A bit more interesting is the Semi-Group typeclass. A semi-Group is a Magma (ie, it still has a $plus$ operation), but it also has the property of associativity for this operation.
 
@@
 
-ccc
+The Monoid, for its part, is a Semi-Group with the property of neutral element.
 
@@
 
-ddd
+A Group is a Monoid, with two new operations.
+The binary $Minus$, and the unary $Neg$.
+One axiom says that the $Minus$ can always be simplified with the $Neg$ and the $Plus$.
+That means that $Minus$ is not a "primitive" operation of a Group.
+The second axiom, which is the most important one for a Group, is the fact that any $c1$ admits $Neg c1$ as an inverse, where being an inverse if define just under.
 
 \begin{code}[caption=Group, captionpos=b, label=lst1:haskell2]   
-hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) -> c -> c -> Type
-hasSymmetric c p a b = (Plus a b = Zero, Plus b a = the c Zero)
+-- This is just a conjunctive predicate
+hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) 
+                -> c -> c -> Type
+hasSymmetric c p a b = 
+         (Plus a b = Zero, Plus b a = Zero)
 
 class dataTypes.Monoid c => dataTypes.Group c where
-	Minus : c -> c ->
+	Minus : c -> c -> c
 	Neg : c -> c
-    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) --Minus should not be primitive and should be simplifiable
-    Plus_inverse : (c1:c) -> hasSymmetric c (%instance) c1 (Neg c1) -- Every element 'x' has a symmetric which is (Neg x)
+	Minus_simpl : (c1:c) -> (c2:c) -> 
+	             Minus c1 c2 = Plus c1 (Neg c2) 
+	Plus_inverse : (c1:c) -> 
+	             hasSymmetric c (instance) c1 (Neg c1)
 \end{code}
 
+This hierarchy can be extended without difficulty with Abelian groups (with the axiom of commutativity), Ring, Commutative Rings...
+
+The important thing to remember is that these typeclass will somehow be used as predicates on types.
+In order to call a solver on a type $T$, the user of the system will have to satisfy the corresponding typeclass by providing an instance of it for his type $T$, ie, he will have to prove that the corresponding properties effectively hold for $T$.
+
+Note that the properties required by these typeclass for a specific type $T$ will either be proven if the operations on $T$ are real computable definitions, or axiomatised if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$) are just axioms.
+
+As discussed in the section above, the algorithm of normalization will not directly use the concrete value of type $T$, but a reflected term, indexed over the concrete value.
+
 	\subsection {Reflected terms}
+
+We will have one datatype for reflected terms in each algebraic structure.
+
+\begin{code}[caption=Reflected terms in a Magma, captionpos=b, label=lst1:haskell2]  
+data ExprMa : Magma c -> (Vect n c) -> c -> Type where
+    ConstMa : (p : Magma c) -> (g:Vect n c) -> (c1:c)  -> ExprMa p g c1 
+    PlusMa : {p : Magma c} -> {g:Vect n c}  -> {c1:c} -> {c2:c} -> ExprMa p g c1 -> ExprMa p g c2 -> ExprMa p g (Plus c1 c2) 
+    VarMa : (p:Magma c) -> {g:Vect n c} -> {c1:c} -> VariableA (magma_eq_as_elem_of_set p) g c1 -> ExprMa p g c1
+\end{code}
 

### 2014-06-25 newPaper/paper.tex
paper continued... I AM LATE ARRRGH

@@
 
-We will have one datatype for reflected terms in each algebraic structure.
+We will have one datatype for reflected terms in each algebraic structure. \\
+Blablabla
 
 \begin{code}[caption=Reflected terms in a Magma, captionpos=b, label=lst1:haskell2]  
-data ExprMa : Magma c -> (Vect n c) -> c -> Type where
-    ConstMa : (p : Magma c) -> (g:Vect n c) -> (c1:c)  -> ExprMa p g c1 
-    PlusMa : {p : Magma c} -> {g:Vect n c}  -> {c1:c} -> {c2:c} -> ExprMa p g c1 -> ExprMa p g c2 -> ExprMa p g (Plus c1 c2) 
-    VarMa : (p:Magma c) -> {g:Vect n c} -> {c1:c} -> VariableA (magma_eq_as_elem_of_set p) g c1 -> ExprMa p g c1
+data ExprMa : Magma c -> (Vect n c) -> c -> 
+              Type where
+    ConstMa : (p : Magma c) -> (g:Vect n c) 
+               -> (c1:c)  -> ExprMa p g c1 
+    PlusMa : {p : Magma c} -> {g:Vect n c}
+               -> {c1:c} -> {c2:c} 
+               -> ExprMa p g c1 -> ExprMa p g c2 
+               -> ExprMa p g (Plus c1 c2) 
+    VarMa : (p:Magma c) -> {g:Vect n c}
+               -> {c1:c} 
+               -> VariableA (get_eq_magma p) g c1
+               -> ExprMa p g c1
 \end{code}
+
+For the moment, the definition of $Variable$ is just the following:
+
+\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
+data Variable : {c:Type} -> {n:Nat} -> 
+         (c_equal : (c1:c) -> (c2:c) -> Maybe(c1=c2)) 
+      -> (Vect n c) -> c -> Type where
+    RealVariable :
+             (c_equal:(c1:c)->(c2:c)->Maybe(c1=c2))
+             -> (g:Vect n c) -> (i:Fin n) 
+             -> Variable c_equal g (index_reverse i g) 
+\end{code}
+
+And $get\_eq\_magma$ is just a function which returns the equality test of the $Set$ level.
 

### 2014-06-26 newPaper/paper.tex
paper

@@
 
-We will define a hierarchy of typeclasses.
+We will construct a hierarchy of typeclasses.
 All our tactics (the group solver, the monoid solver...) will be able to work on any type, being given that an instance of the corresponding typeclass is provided.
@@
 
-All our tactics will require to have a way of testing the equality between elements of the set, that is to say, a way to test equality between constants.
+All our tactics will require to have a way of testing the equality between elements of the underlying set, that is to say, a way to test equality between constants.
 For this reason, we define a notion of "set", which only requires this equality.
@@
 Of course, there will be no tactic associated to $Set$, since we have no operations and not properties associated to this structure.
-Therefore, the equality in a $Set$ is just the syntactical equality, and can be proven with $refl$.
+Therefore, the equality in a $Set$ is just the syntactical equality, and can simply be proven with $refl$.
 
-The first structure, almost trivial, is the magma.
-A magma has just a $Plus$ operation which let us computing constants, and no axioms about this operation.
+The first real structure, almost trivial, is the magma.
+A magma has just a $Plus$ operation which let us computing constants, and has no axioms about this operation.
 Thus, as we will see later, the normalization will be simply a computation between constants inside parenthesis.
@@
 
-A bit more interesting is the Semi-Group typeclass. A semi-Group is a Magma (ie, it still has a $plus$ operation), but it also has the property of associativity for this operation.
+A bit more interesting is the Semi-Group typeclass. A semi-Group is a Magma (ie, it still has a $plus$ operation), but moreover it has the property of associativity for this operation.
 
@@
 class Magma c => SemiGroup c where
-    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) -> (Plus (Plus c1 c2) c3 = Plus c1 (Plus c2 c3))
+    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
+           -> (Plus (Plus c1 c2) c3 
+                = Plus c1 (Plus c2 c3))
 \end{code}
@@
     Zero : c    
-    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 = c1)    
-    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero = c1)
+    Plus_neutral_1 : (c1:c) 
+            -> (Plus Zero c1 = c1)    
+    Plus_neutral_2 : (c1:c) 
+             -> (Plus c1 Zero = c1)
 \end{code}
@@
 A Group is a Monoid, with two new operations.
-The binary $Minus$, and the unary $Neg$.
+The binary operation $Minus$, and the unary operation $Neg$.
 One axiom says that the $Minus$ can always be simplified with the $Neg$ and the $Plus$.
-That means that $Minus$ is not a "primitive" operation of a Group.
-The second axiom, which is the most important one for a Group, is the fact that any $c1$ admits $Neg c1$ as an inverse, where being an inverse if define just under.
+That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$.
+The second axiom, which is the most important one for a Group, is the fact that any $c1$ admits $Neg\ c1$ as an inverse, where being an inverse if the following property :
 
@@
 
-Note that the properties required by these typeclass for a specific type $T$ will either be proven if the operations on $T$ are real computable definitions, or axiomatised if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$) are just axioms.
+Note that the properties required by these typeclass for a specific type $T$ will either be proven if the operations on $T$ are real computable definitions, or axiomatised if the user is working on an axiomatised theory where the operations ($Plus$, $Neg$) are just axioms.
 
@@
 We will have one datatype for reflected terms in each algebraic structure. \\
-Blablabla
+Each of these datatype is parametrised on a type $c$ : the real type on which we want to prove equalities.
+It is also indexed over an instance of the corresponding typeclass for $c$ (we usally call it $p$, because it looks like a proof saying that the structure $c$ has the desired properties), and indexed over a context (a vector $g$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the encoded value.
+A magma is only equiped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constantes, variables, and addition.
 
@@
                -> {c1:c} -> {c2:c} 
-               -> ExprMa p g c1 -> ExprMa p g c2 
+               -> ExprMa p g c1 
+               -> ExprMa p g c2 
                -> ExprMa p g (Plus c1 c2) 
@@
 
+When we encode a constant $c1$ in a context $\Gamma$, we use the constructor $ConstMa$ to produce a term of type $ExprMa\ p\ \Gamma\ c1$ : the index representing the concrete value is simply the constant $c1$.
+If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the courant instance $p$.
+
+We've used $get\_eq\_magma$ in the constructor $VarMa$, and this is just a function which returns the equality test of the $Set$ level.
 For the moment, the definition of $Variable$ is just the following:
@@
 
-And $get\_eq\_magma$ is just a function which returns the equality test of the $Set$ level.
+This type will be extended with some other constructors later on.
+Because there is no more operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
+However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :
+
+\begin{code}[caption=Reflected terms in a Group, captionpos=b, label=lst1:haskell2]
+data ExprG :  Group c -> (Vect n c) -> c -> 
+              Type where
+    ConstG : (p : dataTypes.Group c) 
+       -> (g:Vect n c)
+       -> (c1:c) -> ExprG p g c1
+    PlusG : {p : Group c} -> {g:Vect n c} 
+       -> {c1:c} -> {c2:c} 
+       -> ExprG p g c1 
+       -> ExprG p g c2 
+       -> ExprG p g (Plus c1 c2)
+    MinusG : {p : Group c} -> {g:Vect n c} 
+       -> {c1:c} -> {c2:c} 
+       -> ExprG p g c1 
+       -> ExprG p g c2 
+       -> ExprG p g (Minus c1 c2)
+    NegG : {p : Group c} -> {g:Vect n c} 
+       -> {c1:c} -> ExprG p g c1 
+       -> ExprG p g (Neg c1)
+    VarG : (p : Group c) -> {g:Vect n c} 
+       -> {c1:c}
+       -> Variable (get_eq_group p) Neg g c1 
+       -> ExprG p g c1
+\end{code}
 

### 2014-06-27 newPaper/paper.tex
paper continued. I have to be consistent in my notations, and afterwards to adapt the code

@@
   buildProof : {x : List a} -> {y : List a} 
-               -> Expr G ln -> Expr G rn ->
-               (x = ln) -> (y = rn) ->
-               Maybe (x = y) 
-  buildProof e e' lp rp with (eqExpr e e')
-    buildProof e e lp rp  | Just refl = ?bp1
-    buildProof e e' lp rp | Nothing = Nothing
+               -> {x' : List a} -> {y' : List a} 
+               -> Expr G x' -> Expr G y'
+               -> (x = x') -> (y = y')
+               -> Maybe (x = y) 
+  buildProof e1 e2 p1 p2 with (eqExpr e1 e2)
+    buildProof e1 e1 p1 p2  | Just refl =?bp1
+    buildProof e1 e2 p1 p2 | Nothing = Nothing
 \end{code}
@@
   refine Just;
-  rewrite sym lp;
-  rewrite sym rp;
+  rewrite sym p1;
+  rewrite sym p2;
   exact refl;
@@
   testEq l r = 
-     let (ln ** (l', lPrf)) = reduce l in 
-     let (rn ** (r', rPrf)) = reduce r in
-        buildProof l' r' lPrf rPrf
+     let (x' ** (l', p1)) = reduce l in 
+     let (y' ** (r', p2)) = reduce r in
+        buildProof l' r' p1 p2
 \end{code}
@@
 \begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
-data Variable : {c:Type} -> {n:Nat} -> 
-         (c_equal : (c1:c) -> (c2:c) -> Maybe(c1=c2)) 
+data Variable : {c:Type} -> {n:Nat}
+  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
   -> (Vect n c) -> c -> Type where
-    RealVariable :
-             (c_equal:(c1:c)->(c2:c)->Maybe(c1=c2))
+    RealVariable : (c_equal:(c1:c)->(c2:c)
+                    ->Maybe(c1=c2))
           -> (g:Vect n c) -> (i:Fin n) 
-             -> Variable c_equal g (index_reverse i g) 
+          -> Variable c_equal g (index i g) 
 \end{code}
 
-This type will be extended with some other constructors later on.
+This type will be extended with some other constructors later on. For the moment, we can only express a "real variable". \\
+\\
 Because there is no more operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
@@
 
+Of course, the index of type $c$ (the real value represented by the expression) is always expressed with the oeprations available in the typeclass $p$, and here for a Group they are$Plus$, $Minus$ and $Neg$.
+
+	\subsection { A bit of notation}
+The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (making $c$ a Semi-Group, a Monoid, a Group...).
+The fact that $c$ fulfil the specification of an algebraic structure is expressed as an instance of the corresponding typeclass, and this instance will be denoted as $p$.
+The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ g\ x$ : $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$.
+We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
+Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ g\ x'$ and a proof of $p1:x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ g\ y'$ and a proof of $p2:y=y'$.
+
 	\subsection {Deciding equality}
+
+Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
+This is fact a nice property that all these algebraic structures admit a canonical represent for any element.
+Without this property, it would become a lot more complicated to decide the equality without bruteforcing the
+These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
+
+The profile of these function will always be the same :
+\begin{code}[caption=Type of the reduction function for terms reflecting elements in a Group, captionpos=b, label=lst1:haskell2]
+	groupReduce : {c:Type} -> {n:Nat} 
+	  -> (p:Group c) -> {g:Vect n c} -> {x:c}
+	  -> (ExprG p g x) 
+	  -> (x' ** (ExprG p g x', x=x'))
+\end{code}
+
+This function will have more job to do in structures with multiple axioms (like Group), than it has in the easier structure underneath.
+The detail of these functions will be given in the next section.
+For the moment, we just assume that we've got such a function for each structure, and we want to decide equality between elements.
+We continue to give the details for the Group, but the principle is exactly the same with the other structures.
+
+We want to write the following function :
+\begin{code}[caption=Type of the function for deciding equality between elements in a Group, captionpos=b, label=lst1:haskell2]
+	groupDecideEq : (p:Group c) -> {g:Vect n c} 
+	   -> {x : c} -> {y : c} 
+	   -> (ExprG p g x) -> (ExprG p g y) 
+	   -> Maybe (x = y)
+\end{code}
+
+The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
+Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by "composing" the two proofs $p1$ and $p2$.
+
+\begin{code}[caption=Decides if two elements of a group are equal, captionpos=b, label=lst1:haskell2]
+groupDecideEq p e1 e2 =
+  let (x' ** (e1', p1)) = 
+    groupReduce p e1 in
+  let (y' ** (e2', p2)) = 
+    groupReduce p e2 in
+	     buildProofGroup p e1' e2' p1 p2
+\end{code}
+
+The syntactical test of equality between $e1'$ and $e2'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$.
+
+\begin{code}[caption=Composes the two proofs if the NF are the same, captionpos=b, label=lst1:haskell2]
+buildProofGroup : (p:Group c) -> {g:Vect n c} 
+  -> {x : c} -> {y : c} 
+  -> {x':c} -> {y':c} 
+  -> (ExprG p g x') 
+  -> (ExprG p g y') 
+  -> (x = x') -> (y = y') 
+  -> (Maybe (x = y))
+buildProofGroup p e1 e2 p1 p2 
+       with (exprG_eq p _ e1 e2)
+	buildProofGroup p e1 e1 p1 p2 | Just refl = 
+	         ?MbuildProofGroup
+	buildProofGroup p e1 e2 p1 p2 | Nothing =
+	          Nothing
+\end{code}
+
+The proof for the metavariable MbuildProofGroup is exactly the same as the corresponding one (called $bp$) in the smallest example showed in section 2.
 

### 2014-07-02 newPaper/content/introduction.tex
continued..;

@@
 
-A very common situation in formal certification is to need to prove that one term is equal to another within some theories.
-Of course, they are usually not syntactically equal -this case would be trivial-, but they are equal according to a set of properties.
+Proving that one term is equal to another within some theories is something quite common in formal certification.
+Of course, the two things are usually not syntactically equal --this case would be trivial--, but they are equal according to a set of properties.
 These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for example), or by axioms if we are working with an abstract axiomatised structure.
-These proofs are usually omitted "on the paper" for the ordinary every day maths, but are required by the proof assistant,  no matter how obvious they are.
-So they are done by hand, and are made with a potentially long sequence of rewriting, using the available properties that we have.
+These proofs are usually omitted "on the paper" for the ordinary every day maths, but are required by proof assistants, no matter how obvious they are.
+Currently, in the Idris programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, using the available properties that we have.
 For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
@@
 With the use of this property, we can rewrite the left part into the term $z * (x + y)$.
-If now we use the distributivity of $*$ over $+$, which says that :
+If now we use the distributivity of $*$ over $+$, which says that
 $\forall a b c, a * (b+c) = a*b + a*c$,
 we now obtain the term $x*z + y*z$ as the left side.
-If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes : $z*x + y*z$, and that finishes the proof because it is the same thing as the right hand side.
+If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes $z*x + y*z$, and that finishes the proof because it is the same thing as the right hand side.
 
@@
 
-Instead, what we would like to have :
-INSERT CRUCHY EXAMPLE HERE : 
-a = b by cool\_tactic
+Algebraic structures like monoid, group, commutative group and ring have a very nice property : every element in them has a canonical represent.
+Thus these kind of proofs can be automatically produced by a tactic by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality).
 
+If we've proven that relative numbers follow the structure of some algebraic structure, this is the code that we would like to be able to write :
+
+\begin{code}[caption=Code we'd like to write, captionpos=b, label=lst1:haskell2]
+lemma : (x:ZZ) -> (y:ZZ) -> (z:ZZ) 
+        -> (x+y)*z = (z*x)+(y*z)
+lemma x y z = algebraicSolver ZZ
+\end{code}
+
+Algebraic solvers, like ring solvers, are already implemented for various proof assistants, including Coq and Agda.
 In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
-Idris is a relatively new purely functional and dependently typed programming language with full dependent types, which also supports tactic based proofs.
-Idris will be briefly defined the the [Section X] (at the end of the paper)
+Idris is a relatively new purely functional and dependently typed programming language with full dependent types, which also supports tactic based proofs  (Idris will be briefly defined the the [Section X] (at the end of the paper)).
+The interesting thing that we're going to present in this paper lies in the way we construct it, in a "correct by construction" style.
 
@@
 
-We will work by reflection for the implementation of each tactic, and the reason for this will be discussed in section [Blabla].
+Each tactic will be implemented by reflection, and the reason for this will be discussed in section [Blabla].
 Working by reflection for implementing tactics has been done several times, as in [paper Coq ring] for the implementation of a ring solver for the Coq proof assistant.
-We will compare with other implementation of this category of tactics in section [Blabla].
+We will compare our approach with other implementation of this kind of tactics in section [Blabla].
 \\
-% NOT THE RIGHT PLACE FOR THAT
-%For these structures, this is effectively possible to decide if a given equality is true, and to produce the proof of the %equality is %appropriate.
-This result comes directly from the fact that there exist a normal form for every element of %these structures.
-The general %idea is therefore to normalize both side of the equality, and then to compare the %resulting terms using the usual syntactical Leibniz's %equality.
 \\
@@
 	\item 1) We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
-For achieving this goal, we are extensively using dependent types in order to capture the interesting properties that matters for assuring the correctness of the method.
-	\item 2) This collection of tactics can be written in a "correct by construction" style because we use an original representation of reflected terms which encodes the denoted -concrete- value;
-	\item 3) We develop a hierarchy of tactics which enables to be increase a lot the reusability of each solver : we reuse each solver for the solvers of the structures above.
-For example simplifiying neutral elements is only implemented at the Monoid level, and each level above reuse it.
-As will discussed in section [blabla], reusability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.
+The approach will be presented in section 2 on a smaller problem, and on section 3 for the complete hierarchy of algebraic structures.
+Details of the normalization procedures are given in section 4.
+	\item 2) For achieving the first goal, we are extensively using dependent types in order to capture the interesting properties that matter for assuring the correctness of the method.
+We use an original representation of reflected terms which embeds the denoted --concrete-- value represented.
+The reflected terms become indexed over the concrete ones.
+These representations are presented in section 2 for the toy example, and in section 3 for the real problem.
+	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.
+For example simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuse it.
+As it will be discussed in section 4, reusability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.
 For example, reusing the monoid solver for the group solver is not trivial, since we lose the possibility to express the negations $-x$ and minus $x-y$ operations.

### 2014-07-02 newPaper/paper.tex
continued..;

@@
 
-\title{Proof by reflection for equalities in algebraic structures}
+\title{A correct by construction hierarchy of reflective tactics for equalities in algebraic structures}
 %\subtitle{An example for a collection of tactics solving equalities in algebraic structures}
@@
 
-Remark : To do : making the introduction more SPECIFIC, with CRUNCHY things : what we would like to do, no general assertions like "we often have to prove equalities...". (cf Simon P J talk)
 \input{./content/introduction.tex}

### 2014-07-03 newPaper/content/introduction.tex
introduction changed to a nicer example

@@
 These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for example), or by axioms if we are working with an abstract axiomatised structure.
-These proofs are usually omitted "on the paper" for the ordinary every day maths, but are required by proof assistants, no matter how obvious they are.
-Currently, in the Idris programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, using the available properties that we have.
-For example, if $x$ and $y$ are relatives numbers, and if we want to show that :
-$(x + y) * z = (z * x) + (y * z)$, we can first use the symmetry of $*$, saying that
-$\forall a b, a * b = b * a$.
-With the use of this property, we can rewrite the left part into the term $z * (x + y)$.
-If now we use the distributivity of $*$ over $+$, which says that
-$\forall a b c, a * (b+c) = a*b + a*c$,
-we now obtain the term $x*z + y*z$ as the left side.
-If we use again the symmetry of $*$ on the subterm $(x*z)$, the left side of the equality becomes $z*x + y*z$, and that finishes the proof because it is the same thing as the right hand side.
+These proofs are usually omitted "on the paper" for the ordinary every day maths, but they are required by proof assistants, no matter how obvious they are.
+Currently, in the Idris\footnote{Idris is a relatively new, purely functional, and dependently typed programming language with full dependent types, which also supports tactic based proofs  (Idris will be briefly defined the the [Section X] (at the end of the paper))} programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
+These proofs are the every day routine we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
+The indices can be for example natural numbers, and the lemmas we often have to prove about these indices are proven by some simple --but potentially long-- rewritings which use the properties of natural number and the usual operations : symmetry of +, commutativity of +, etc.
+
+\subsection{A motivating little example}
+Let's define the types of bit and binary number, both indexed over the value they represent, expressed in base ten as a natural number.
+\begin{code}[caption=Bit and binary number, captionpos=b, label=lst1:haskell2]
+data Bit : Nat -> Type where
+     b0 : Bit Z
+     b1 : Bit (S Z)
+
+data Binary : (width : Nat) -> (value : Nat) 
+              -> Type where
+     zero : Binary Z Z
+     (#) : Binary w v -> Bit bit 
+           -> Binary (S w) (bit + 2 * v)
+\end{code}
+
+This type of binary numbers\footnote{This representation allows to represent a binary number of width zero with the constructor $zero$, but instead we could have decided to replace this constructor with a constructor taking one bit as parameter and producing a binary number of width one.
+The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value represented by $bin$ plus the value represented by the added digit.
+
+It will be useful to be able to pad a binary number expressed with $w$ bits to a binary number with $S\ w$ digits : 
+\begin{code}[caption=Padding, captionpos=b, label=lst1:haskell2]
+pad : Binary w n -> Binary (S w) n
+pad zero = zero # b0
+pad (num # x) = pad num # x
+\end{code}
+Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition.
+
+Now, we're tempted to write the addition between two binary numbers.
+We start with an auxiliary function which adds three bits (the third one will play the role of the carry), and produces the two bits of the result.
+
+\begin{code}[caption=Addition of three bits, captionpos=b, label=lst1:haskell2]
+addBit : Bit x -> Bit y -> Bit c ->
+          (bX ** (bY ** 
+              (Bit bX, Bit bY, 
+                  c + x + y = bY + 2 * bX)))
+addBit b0 b0 b0 = (_ ** (_ ** (b0, b0, _)))
+addBit b0 b0 b1 = (_ ** (_ ** (b0, b1, _)))
+addBit b0 b1 b0 = (_ ** (_ ** (b0, b1, _)))
+addBit b0 b1 b1 = (_ ** (_ ** (b1, b0, _)))
+addBit b1 b0 b0 = (_ ** (_ ** (b0, b1, _)))
+addBit b1 b0 b1 = (_ ** (_ ** (b1, b0, _)))
+addBit b1 b1 b0 = (_ ** (_ ** (b1, b0, _)))
+addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, _)))
+\end{code}
+
+In fact, this function produces a dependent pair.
+The first component is the value $bX$ represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
+This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing the two bits produced and the proof that these two bits are effectively encoding the result of the addition we had to perform.
+For example, with the line corresponding to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
+
+Again, we note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : that two bits produced effectively represent the addition of the three bits given in input.
+
+Now, we would like to define a function adding two binary numbers and a bit (used as a carry).
+We demand the two binary number to be expressed with the same width $w$, and we produce a result of the same width, which of course should represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
+Thus, we're tempted to write :
+\begin{code}[caption=Addition of two binary numbers, captionpos=b, label=lst1:haskell2]
+adc : Binary w x -> Binary w y -> Bit c 
+     -> Binary (S w) (c + x + y)
+adc zero zero carry = zero # carry
+adc (numx # bX) (numy # bY) carry
+   = let (v1 ** (v2 ** (carry0, lsb, _))) = 
+      addBit bX bY carry in
+          adc numx numy carry0 # lsb
+\end{code}
+Unfortunately, the types are mismatching for both of the two lines.
+The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type $Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)$ and we're providing a term of type $Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))$ [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
+The problem is even bigger for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
+For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
+Using a provisional definitions here will make the definition being accepted, but we will have to prove that the term provided can be converted into a term which has the right type.
+However, for both lines, it is possible to prove the equality between the expected and the provided types, so it is possible to convert the type of the provided term into the expected term.
+These proofs will be made of a series of rewriting, and each rewriting will use a property about natural numbers and the usual operations.
+FOr example, the proof for the second line look like this :
+\begin{code}[caption=Proof by hand of correspondence between expected and provided types for the second line of the definition of adc, captionpos=b, label=lst1:haskell2]
+Main.adc_lemma_2 = proof {
+    intros;
+    rewrite sym (plusZeroRightNeutral x);
+    [...]
+    rewrite sym (plusAssociative (plus c (plus bit0 (plus v v))) bit1 (plus v1 v1));
+    rewrite (plusAssociative c (plus bit0 (plus v v)) bit1);
+    rewrite plusCommutative bit1 (plus v v);
+    [...]
+    rewrite (plusAssociative (plus (plus x v) v1) (plus x v) v1);
+    trivial;
+}
+\end{code}
 
@@
 This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
+This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will requires us to replace this proof by another one doing the same job : assuring the correspondance betwen the expected index and the given index.
 The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
@@
 
-If we've proven that relative numbers follow the structure of some algebraic structure, this is the code that we would like to be able to write :
+If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), and the associativity and commutativity of $+$.
+Thus, we're rewriting a term by using the properties of a commutative group.
+If we've already proven that natural numbers follow the structure of commutative group, this is the code that we would like to be able to write :
 
-\begin{code}[caption=Code we'd like to write, captionpos=b, label=lst1:haskell2]
-lemma : (x:ZZ) -> (y:ZZ) -> (z:ZZ) 
-        -> (x+y)*z = (z*x)+(y*z)
-lemma x y z = algebraicSolver ZZ
+\begin{code}[caption=Code we'd like to write for the addition of two binary numbers, captionpos=b, label=lst1:haskell2]
+adc : Binary w x -> Binary w y -> Bit c 
+     -> Binary (S w) (c + x + y)
+adc zero zero carry = zero # carry 
+        by CommutativeGroupSolver
+adc (numx # bX) (numy # bY) carry
+   = let (v1 ** (v2 ** (carry0, lsb, _))) = 
+      addBit bX bY carry in
+          adc numx numy carry0 # lsb 
+        by CommutativeGroupSolver
 \end{code}
+We would like to let the solver for commutative groups to find the proof on its own, and thereby, there will be no extra lemmas to prove afterwards.
+
+\subsection{Our contribution}
 
@@
 In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
-Idris is a relatively new purely functional and dependently typed programming language with full dependent types, which also supports tactic based proofs  (Idris will be briefly defined the the [Section X] (at the end of the paper)).
 The interesting thing that we're going to present in this paper lies in the way we construct it, in a "correct by construction" style.

### 2014-07-03 newPaper/content/introduction.tex
paper...

@@
 Of course, the two things are usually not syntactically equal --this case would be trivial--, but they are equal according to a set of properties.
-These properties are either obtained by implementation (the implementation of + for relatives numbers is symmetric for example), or by axioms if we are working with an abstract axiomatised structure.
+These properties are either obtained by implementation (the implementation of + for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.
 These proofs are usually omitted "on the paper" for the ordinary every day maths, but they are required by proof assistants, no matter how obvious they are.
 Currently, in the Idris\footnote{Idris is a relatively new, purely functional, and dependently typed programming language with full dependent types, which also supports tactic based proofs  (Idris will be briefly defined the the [Section X] (at the end of the paper))} programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
-These proofs are the every day routine we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
-The indices can be for example natural numbers, and the lemmas we often have to prove about these indices are proven by some simple --but potentially long-- rewritings which use the properties of natural number and the usual operations : symmetry of +, commutativity of +, etc.
+These proofs are the every day routine when we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
+The indices can be for example natural numbers, and in this case the lemmas that we have to prove about these indices are arithmetic statements which use the properties of natural numbers and their usual operations : symmetry and commutativity for +, existence of neutral element for +, etc.
 
 \subsection{A motivating little example}
-Let's define the types of bit and binary number, both indexed over the value they represent, expressed in base ten as a natural number.
+We introduce a little example which shows how these proofs appear when we index a type by values like natural numbers.
+Let's define the types of $Bit$ and $Binary$ number, both indexed over the value they represent, expressed in base ten as a natural number.
 \begin{code}[caption=Bit and binary number, captionpos=b, label=lst1:haskell2]
@@
 \end{code}
-Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition.
+Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition (the $value$ is still $n$).
 
 Now, we're tempted to write the addition between two binary numbers.
-We start with an auxiliary function which adds three bits (the third one will play the role of the carry), and produces the two bits of the result.
+We start with an auxiliary function which adds three bits (the third one will latter play the role of the carry), and produces the two bits of the result.
 
@@
 
-Again, we note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : that two bits produced effectively represent the addition of the three bits given in input.
+Again, we note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : that the two bits produced effectively represent the addition of the three bits given in input.
 
-Now, we would like to define a function adding two binary numbers and a bit (used as a carry).
-We demand the two binary number to be expressed with the same width $w$, and we produce a result of the same width, which of course should represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
+Now, we want to define a function adding two binary numbers and a bit (used as a carry).
+We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit, and this result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
 Thus, we're tempted to write :
@@
 \end{code}
-Unfortunately, the types are mismatching for both of the two lines.
-The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type $Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)$ and we're providing a term of type $Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))$ [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
-The problem is even bigger for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
-For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
-Using a provisional definitions here will make the definition being accepted, but we will have to prove that the term provided can be converted into a term which has the right type.
-However, for both lines, it is possible to prove the equality between the expected and the provided types, so it is possible to convert the type of the provided term into the expected term.
+Unfortunately, the types are mismatching for the two cases expressed by this pattern matching.
+The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] and we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\] [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
+The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
+For this kind of situation, Idris provides the possibility to write a provisional definition --provisional definitions are described with the brief overview of the language in page XX-- with the syntax "?=".
+Using a provisional definitions here will make the definition being accepted, but we will have to prove afterwards that the term provided can be converted into a term which has the right type.
+However, for both of these two patterns, it is possible to prove the equality between the expected and the provided types, and this is even stronger than we need.
+Using this equality of types it will be trivial to produce a term of the expected type by using the provided term and by simply rewriting its type with this proof.
 These proofs will be made of a series of rewriting, and each rewriting will use a property about natural numbers and the usual operations.
-FOr example, the proof for the second line look like this :
-\begin{code}[caption=Proof by hand of correspondence between expected and provided types for the second line of the definition of adc, captionpos=b, label=lst1:haskell2]
-Main.adc_lemma_2 = proof {
+For example, the proof for the second pattern look like this :
+\begin{code}[caption=Proof by hand of the correspondence between expected and provided types for the second line of the definition of adc, captionpos=b, label=lst1:haskell2]
+adc_lemma_2 = proof {
     intros;
@@
 
-Algebraic structures like monoid, group, commutative group and ring have a very nice property : every element in them has a canonical represent.
-Thus these kind of proofs can be automatically produced by a tactic by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality).
-
 If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), and the associativity and commutativity of $+$.
@@
 \end{code}
-We would like to let the solver for commutative groups to find the proof on its own, and thereby, there will be no extra lemmas to prove afterwards.
+We would like to let a solver for commutative groups to find the proof on its own, and thereby, there will be no extra lemmas to prove afterwards.
+This is a something doable because algebraic structures like monoid, group, commutative group and ring have all a very nice property : every element in them has a canonical represent.
+Thus these kind of proofs can be automatically produced by a tactic by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality).
 
-\subsection{Our contribution}
+\subsection{Our contributions}
 
@@
 
-For our goal, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group and Abelian group.
+In this paper, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group and Abelian group.
 The user will just have to prove that his structure satisfies the properties of an algebraic structure, and he will be allowed to invoke the corresponding tactic.

### 2014-07-05 newPaper/paper.tex
paper...
changed simplified problem with natural numbers, as in the introduction

@@
 
-\title{A correct by construction hierarchy of reflective tactics for equalities in algebraic structures}
+\title{A correct by construction hierarchy of reflective tactics for proving equalities in algebraic structures}
 %\subtitle{An example for a collection of tactics solving equalities in algebraic structures}
@@
 
-In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universaly quantified lists, and associativity of list concatenation.
-For example, we would like to be able to automatically generate proofs of facts like $\forall l1\ l2\ l3 : List,\ (l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$ [Example 1]. \\
-For this smaller problem, we have decided to only work with the associativity of the concatenation : $\forall l1\ l2\ l3,\ (l1 ++ l2) ++ l3 = l1 ++ (l2 ++ l3)$ and with the fact that Nil is a neutral element for the concatenation : $\forall l, l ++ Nil = l$.
-Of course, this work could be extended by dealing with more axioms, and this is exactly what we will do in the next sections when we will generalize this problem.
+In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universally quantified natural numbers, the the properties of associativity and neutral element for the operation plus.
+For example, we would like to be able to automatically generate proofs of facts like $\forall x1\ x2\ x3 : List,\ (x1 + l2) ++ (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
+For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) ++ x3 = x1 ++ (x2 + x3)$ and with the fact that Z is a neutral element on the right side of the plus : $\forall x, x ++ Z = x$.
+Of course, this work could be extended by dealing with more axioms, and this is exactly what we will do in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified lists and concatenation of these lists are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the "neutrality" of Nil".
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the "neutrality" of Z".
 \\
@@
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-For the example 1, after abstraction of the variables, the goal becomes simply $(l1 ++ l2) ++ (l3 ++ l4) = (l1 ++ (l2 ++ l3)) + l4$, which is something of the general form $x=y$.
+For the example 1, after abstraction of the variables, the goal becomes simply $(x1 + x2) ++ (x3 + l4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
 The general idea -which will also apply for the real bigger problem detailed in the next sections- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
-The goal of the normalization is of course to compute a canonical representation for a list $l$, such that any other list provably equal to $l$ (by using the associativity) will have the same canonical representation.
+The goal of the normalization is of course to compute a canonical representation for a list $x$, such that any other list provably equal to $x$ (by using the two available properties) will have the same canonical representation.
 That's why it will be possible to decide the equality by simply comparing the two normal forms with a week syntactical equality.
@@
 Of course, the normalization will have to fulfil the following correctness property :
-$\forall l, norm\ l = l$ [1], which means that after normalization we obtain a list which is provably equal to the input list (potentially using the property of associativity multiple times). \\
-And after normalization, we will compare the two resulting lists with a standard syntactical equality, and if they are equal, we will have obtained the desired proof.  \\
+$\forall x, norm\ x = x$ [1], which means that after normalization we obtain a list which is provably equal to the input list (potentially using our properties multiple times). \\
+And after normalization, we will compare the two resulting terms with a standard syntactical equality, and if they are equal, we will have obtained the desired proof.  \\
 Indeed :
@@
 In fact, this is not exactly how it will work.
-This way would not be easily feasible, because in the LHS and RHS of $l1=l2$, we potentially have variables which have been universally quantified.
-And the normalization function would need to do different treatments for a "variable list" (a list which has been universally quantified) and for a constant empty list.
+This way would not be easily feasible, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
+And the normalization function would need to do different treatments for a "variable natural number" (a number which has been universally quantified) and for a constant $Z$.
 This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality : this is a meta information of the AST that we can't access.
 For this reason, we will work by reflection, as we will do for for the real problem later.
-That means that we will define a datatype which will be used as an encoding of lists, or more precisely, an encoding of lists composed of "variable lists", Nil, and concatenation of the previous.
-This datatype will allow us to know the internal structure of the list, ie, where the variables and constants are.
+That means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
+This datatype will allow us to know the internal structure of the number, ie, where the variables and constants are.
 Indeed, this datatype will allow pattern matching.
-Previously, we were only able to pattern match a list against the constructor $Nil$ and $Cons$, which wasn't interesting for our goal.
-With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of list against the constructor App, Var and ENil, which gives us the informations we want.
+Previously, we were only able to pattern match a list against the constructor $Z$ and $S$, which wasn't interesting for our goal.
+With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of number against the constructor Plus, Var and Zero, which gives us the informations we want.
 
-\begin{code}[caption=First version of reflected lists, captionpos=b, label=lst1:haskell2]
-data Expr : (Vect n (List a)) -> Type where
-  App  : {n:Nat} -> {G : Vect n (List a)} -> 
+\begin{code}[caption=First version of reflected numbers, captionpos=b, label=lst1:haskell2]
+data Expr : (G : Vect n Nat) -> Type where
+     Plus : {n:Nat} -> {G:Vect n Nat} -> 
             Expr G -> Expr G -> Expr G
-  Var  : {n:Nat} -> (G : Vect n (List a)) -> 
+     Var  : {n:Nat} -> {G:Vect n Nat} -> 
             (i : Fin n) -> Expr G
-  ENil : {n:Nat} -> (G : Vect n (List a)) -> 
-         Expr G 
+     Zero : {n:Nat} -> (G:Vect n Nat) -> 
+            Expr G Z
 \end{code}
 
-For this smaller problem, we want to prove equalities in which lists are always universally quantified, so we only need to represent variable (a universally quantified list), the constant list Nil, and the concatenation between lists.
+For this smaller problem, we want to prove equalities in which numbers are always universally quantified, so we only need to represent variable (a universally quantified natural), the constant Z, and the addition.
 Variables are represent using a De Brujin-like index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
 
-The type Expr is indexed over a vector of lists (we will usually name it G), that represents our context.
-For example, if in the example 1, we will have to encode $(l1 ++ l2) ++ (l3 ++ l4)$ and $(l1 ++ (l2 ++ l3)) + l4$ in a context where four elements are declared.
-The first element of this context denotes the variable $l1$, the second denotes $l2$, and so on.
+The type Expr is indexed over a vector of numbers (we will usually name it $G$), that represents our context.
+For example, if in the example 1, we will have to encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are declared.
+The first element of this context denotes the variable $x1$, the second denotes $x2$, and so on.
 Thus, the left hand side will be encoded by :
@@
 
-We need now to define what will be the normalization procedure.
-If, as it is probably expected, this function simply takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards, corresponding to the correcness property mentioned before : \\
+We now need to define what will be the normalization procedure.
+If, as it is probably expected, this function simply takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards, corresponding to the correctness property mentioned before : \\
 $\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
@@
 This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way we compute the interpretation.
-For this little example, the reduction will not be too heavy, but in the next sections, when we will have more properties than only associativity to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+For this little example, the reduction will not be too heavy, but in the next sections, when we will have more properties than only associativity and neutral element to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
 \\
@@
 To avoid these two sources of complexity, we will : \\
-- Use dependent types in order to directly capture and embed the concrete list that an Expr is encoding. Thus, it won't be necessary to define the $interpretation$ function. \\
+- Use dependent types in order to directly capture and embed the concrete number that an Expr is encoding. Thus, it won't be necessary to define the $interpretation$ function. \\
 - Write a "correct by construction" reduction procedure.
-The proof of preservation of the interpretation will be computed "bit by bit" at the same time as the function will produce the normalized expression (by rewriting the term thanks to the property of associativity).
+The proof of preservation of the interpretation will be computed "bit by bit" at the same time as the function will produce the normalized expression (by rewriting the term thanks to the properties of associativity and neutral element).
 
@@
 \begin{code}[caption=Second version of reflected lists with embedded denotation, captionpos=b, label=lst1:haskell2]
-using (x : List a, y : List a, 
-       G : Vect n (List a))
-  data Expr : (G : Vect n (List a)) -> 
-              List a -> Type where
-       App  : Expr G x -> Expr G y -> 
-              Expr G (x ++ y)
-       Var  : (i : Fin n) -> 
-              Expr G (index i G)
-       ENil : Expr G []
+using (x : Nat, y : Nat, G : Vect n Nat)
+  data Expr : (G : Vect n Nat) -> Nat -> Type where
+       Plus : Expr G x -> Expr G y -> Expr G (x + y)
+       Var  : (i : Fin n) -> Expr G (index i G)
+       Zero : Expr G Z
 \end{code}
-For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the list $x$".
-When an expression is a variable, the denoted list is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
-Also, the $ENil$ expression denotes the $Nil$ list.
-Finally, if $e1$ is an expression encoding the list $x$, and $e2$ is an expression encoding the list $y$, then the expression $App\ e1\ e2$ denotes the list $(x ++ y)$.
+For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the number $x$".
+When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
+Also, the $Zero$ expression denotes the natural number $Z$.
+Finally, if $e1$ is an expression encoding the number $x$, and $e2$ is an expression encoding the number $y$, then the expression $Plus\ e1\ e2$ denotes the number $(x + y)$.
 
 The reduction procedure is now supposed to be written on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
-Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real (concrete) list : a term of type $Expr\ G\ l$ is the encoding of the list $l$.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- list : a term of type $Expr\ G\ x$ is the encoding of the number $x$.
 Thus, we can write the type of $reduce$ like this : \\
 $reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
-The function $reduce$ produces a dependent pair : the new concrete list $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete list, and a proof that old and new -real- lists are equal.
-Note that this function can't simply produce an $Expr\ G\ x$, because the list on which the resulting expression will be indexed is not necessary syntactically equal to the original list since it can use the property of associativity.
-In fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
+The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
+Note that this function can't simply produce an $Expr\ G\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since it can use the two available properties.
+And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
 The reason is simple : when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
@@
 We will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
-We will normalize $e1$, and this will give a new list $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
-We will do the same with $e2$, and we will get a new list $y$, an expression $e2':Expr\ G\ y'$, and a proof of $y=y'$.
+We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
+We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ G\ y'$, and a proof of $y=y'$.
 \\
@@
 \begin{code}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
-  eqExpr : (e : Expr G x) -> (e' : Expr G y) 
-  	       -> Maybe (e = e')
-  eqExpr (App x y) (App x' y') 
-          with (eqExpr x x', eqExpr y y')
-    eqExpr (App x y) (App x y)   
-    	| (Just refl, Just refl) = Just refl
-    eqExpr (App x y) (App x' y') 
-    	| _ = Nothing
-  eqExpr (Var i) (Var j) with (decEq i j)
-    eqExpr (Var i) (Var i) 
-    	| (Yes refl) = Just refl
-    eqExpr (Var i) (Var j) 
-    	| _ = Nothing
-  eqExpr ENil ENil = Just refl
+  eqExpr : (e : Expr G xs) -> (e' : Expr G ys) ->
+           Maybe (e = e')
+  eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
+    eqExpr (Plus x y) (Plus x y) | (Just refl, Just refl) = Just refl
+    eqExpr (Plus x y) (Plus x' y') | _ = Nothing
+  eqExpr (Var i) (Var j) with (prim__syntactic_eq _ _ i j)
+    eqExpr (Var i) (Var i) | (Just refl) = Just refl
+    eqExpr (Var i) (Var j) | _ = Nothing
+  eqExpr Zero Zero = Just refl
   eqExpr _ _ = Nothing
@@
 \begin{code}[caption=Building the desired proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
-  buildProof : {x : List a} -> {y : List a} 
-               -> {x' : List a} -> {y' : List a} 
-               -> Expr G x' -> Expr G y'
-               -> (x = x') -> (y = y')
-               -> Maybe (x = y) 
-  buildProof e1 e2 p1 p2 with (eqExpr e1 e2)
-    buildProof e1 e1 p1 p2  | Just refl =?bp1
-    buildProof e1 e2 p1 p2 | Nothing = Nothing
+  buildProof : {xs : Nat} -> {ys : Nat} ->
+               Expr G ln -> Expr G rn ->
+               (xs = ln) -> (ys = rn) -> Maybe (xs = ys)
+  buildProof e e' lp rp with (eqExpr e e')
+    buildProof e e lp rp | Just refl = ?bp1
+    buildProof e e' lp rp | Nothing = Nothing
 \end{code}
@@
 Now, we need to define the function reduce.
-To do that, we have to fix a canonical representation of associative lists.
+To do that, we have to fix a canonical representation of associative natural numbers.
 We decide that the left associative form will be the canonical representation.
-Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform the underlying list in the form $(...((l1 ++ l2) ++ l3) ...
-++ ln)$.
+Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
++ xn)$.
 To do so, one possibility is to define a new datatype which captures this property, and to write a function going from $Expr$ to this new type.
 Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.
-\begin{code}[caption=Reflected left associative lists, captionpos=b, label=lst1:haskell2]
-  data LExpr : (G : Vect n (List a)) -> 
-               List a -> Type where
-       LApp : LExpr G x -> (i : Fin n) -> 
-              LExpr G (x ++ index i G)
-       LNil : LExpr G []
+\begin{code}[caption=Reflected left associative numbers, captionpos=b, label=lst1:haskell2]
+  data LExpr : (G : Vect n Nat) -> Nat -> Type where
+       LPlus : LExpr G x -> (i : Fin n) -> LExpr G (x + index i G)
+       LZero : LExpr G Z
 \end{code}
-This datatype has only two constructors. In fact, it combines the previous $Var$ and $App$ constructors so that it becomes impossible to write an expression which isn't left associative.
+This datatype has only two constructors. In fact, it combines the previous $Var$ and $Plus$ constructors so that it becomes impossible to write an expression which isn't left associative.
 
@@
 \begin{code}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
-  expr_l : Expr G x -> 
-               (x' ** (LExpr G x', x = x'))
-  expr_l ENil = (_ ** (LNil, refl))
-  expr_l (Var i) = (_ ** 
-                   (LApp RNil i, refl))
-  expr_l (App ex ey) = 
-  	let (xl ** (xr, xprf)) = expr_l ex in
+  expr_l : Expr G xs -
+        > (xs' ** (LExpr G xs', xs = xs'))
+  expr_l Zero = (_ ** (LZero, refl))
+  expr_l (Var i) = (_ ** (LPlus LZero i, refl))
+  expr_l (Plus ex ey) = let (xl ** (xr, xprf)) = expr_l ex in
                        let (yl ** (yr, yprf)) = expr_l ey in
-   	   appLExpr _ _ xr yr xprf yprf
+                               PlusLExpr _ _ xr yr xprf yprf
     where
-       appLExpr : (x', y' : List a) ->
-         LExpr G x -> LExpr G y -> 
-         (x' = x) -> (y' = y) ->
-         (w' ** (LExpr G w', x' ++ y' = w'))
-       appLExpr x' y' rxs (LApp e i) xprf yprf =
-          let (xs ** (rec, prf)) = 
-             appLExpr _ _ rxs e refl refl in
-             (_ ** (LApp rec i, ?appLExpr1))
-       appLExpr x' y' rxs LNil xprf yprf = 
-          (_ ** (rxs, ?appLExpr2))
+      PlusLExpr : (x', y' : Nat) ->
+                 {G : Vect n Nat} -> {x, y : Nat} ->
+                 LExpr G x -> LExpr G y -> (x' = x) -> (y' = y) ->
+                 (w' ** (LExpr G w', x' + y' = w'))
+      PlusLExpr x' y' rx (LPlus e i) xprf yprf
+         = let (x ** (rec, prf)) = PlusLExpr _ _ rx e refl refl in
+               (_ ** (LPlus rec i, ?PlusLExpr1))
+      PlusLExpr x' y' rx LZero xprf yprf = (_ ** (rx, ?PlusLExpr2))
 
-  l_expr : LExpr G x -> Expr G x
-  l_expr LNil = ENil
-  l_expr (LApp xs i) = App (r_expr xs) (Var i)
+  l_expr : LExpr G xs -> Expr G xs
+  l_expr LZero = Zero
+  l_expr (LPlus x i) = Plus (l_expr x) (Var i)
+
 \end{code}
 
-We notice that for transforming the list into its left associative equivalent representation, we've effectively needed to know where the variables and the Nil constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
+We notice that for transforming the number into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
 \\
@@
 We've got two metavariables to prove.
-The metavariable $appRExpr1$ requires us to prove the goal : $x' ++ y' = xs ++ index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x2\ ++\ index\ i\ G)$ and $(prf\ :\ x\ ++\ x2\ =\ xs)$.
+The metavariable $appLExpr1$ requires us to prove the goal : $x' + y' = x + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x2\ ++\ index\ i\ G)$ and $(prf\ :\ x\ ++\ x2\ =\ x)$.
 Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
 
-\begin{code}[caption=Proof of the metavariable appRExpr1, captionpos=b, label=lst1:haskell2]
-AutoAssoc.appRExpr1 = proof {
+\begin{code}[caption=Proof of the metavariable appLExpr1, captionpos=b, label=lst1:haskell2]
+PlusRExpr1 = proof {
   intros;
@@
   rewrite sym 
-       (appendAssociative x x2 (index i G));
+       (plusAssociative x x2 (index i G));
   exact refl;
@@
 
-And the metavariable appRExpr2 uses the fact that Nil is a neutral element for the concatenation.
+And the metavariable appRExpr2 uses the fact that $Z$ is a neutral element for the concatenation.
 
-\begin{code}[caption=Proof of the metavariable appRExpr2, captionpos=b, label=lst1:haskell2]
-appRExpr2 = proof {
-  intros;
-  rewrite xprf;
-  rewrite sym yprf;
-  rewrite appendNilRightNeutral x';
-  exact refl;
-}
+\begin{code}[caption=Proof of the metavariable appLExpr2, captionpos=b, label=lst1:haskell2]
+PlusLExpr2 = proof
+  intros
+  rewrite xprf 
+  rewrite (sym yprf)
+  mrefine a_plus_zero
 \end{code}
@@
 It's now time to see how to use this minimalist "tactic".
-Let's define two expressions $e1$ and $e2$, respectively representing the lists $((xs ++ ys) ++ (xs ++ zs))$ and $(xs ++ ((ys ++ xs) ++ zs))$ in the context $[xs, ys, zs]$ of three abstracted variables.
+Let's define two expressions $e1$ and $e2$, respectively representing the numbers $((x + ys) + (x + z))$ and $(x + ((y + x) + z))$ in the context $[x, y, z]$ of three abstracted variables.
 
 \begin{code}[caption=Two test expressions, captionpos=b, label=lst1:haskell2]
-  e1 : (xs, ys, zs : List a) -> 
-           Expr [xs, ys, zs] 
-           ((xs ++ ys) ++ (xs ++ zs))
-  e1 xs ys zs = 
-     App (App (Var fZ) (Var (fS fZ))) 
-         (App (Var fZ) (Var (fS (fS fZ))))
+  e1 : (xs, ys, zs : Nat) -> 
+           Expr [xs, ys, zs] ((xs + ys) + (xs + zs))
+  e1 xs ys zs = Plus (Plus (Var fZ) (Var (fS fZ))) 
+                        (Plus (Var fZ) (Var (fS (fS fZ))))
 
-  e2 : (xs, ys, zs : List a) -> 
-           Expr [xs, ys, zs] 
-           (xs ++ ((ys ++ xs) ++ zs))
-  e2 xs ys zs = 
-     App (Var fZ) 
-         (App (App (Var (fS fZ)) (Var fZ)) 
-              (Var (fS (fS fZ))))
+  e2 : (xs, ys, zs : Nat) -> 
+           Expr [xs, ys, zs] (xs + ((ys + xs) + zs))
+  e2 xs ys zs = Plus (Var fZ) 
+         (Plus (Plus (Var (fS fZ)) (Var fZ)) (Var (fS (fS fZ))))
 \end{code}
 
-The lists denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this by using $testEq$.
+The numbers denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this by using $testEq$.
 \begin{code}[caption=Test of equality betwen e1 and e2, captionpos=b, label=lst1:haskell2]
-  e1_e2_testEq : (xs, ys, zs : List a) ->
-          Maybe (((xs ++ ys) ++ (xs ++ zs)) = 
-                 (xs ++ ((ys ++ xs) ++ zs)))
-  e1_e2_testEq xs ys zs = 
-     testEq (e1 xs ys zs) (e2 xs ys zs) 
+  e1_e2_testEq : (xs, ys, zs : Nat) ->
+          Maybe (((xs + ys) + (xs + zs)) = (xs + ((ys + xs) + zs)))
+  e1_e2_testEq xs ys zs = testEq (e1 xs ys zs) (e2 xs ys zs)
 \end{code}
 
-And if we ask for the evaluation of this term, we should obtain $Just$ and a proof of equality between the two underlying lists.
+And if we ask for the evaluation of this term, we should obtain $Just$ and a proof of equality between the two underlying numbers.
 \begin{code}[caption=Evaluation of the result, captionpos=b, label=lst1:haskell2]
-\xs => \ys => \zs => e1_e2_testEq {a=Int} xs ys zs
+\xs => \ys => \zs => e1_e2_testEq xs ys zs
 
 \xs => \ys => \zs => Just (replace (sym (replace 
-(sym (replace (replace (appendNilRightNeutral xs)
- refl (replace (sym (appendAssociative xs [] ys)) 
+(sym (replace (replace (plusZeroRightNeutral xs)
+ refl (replace (sym (plusAssociative xs [] ys)) 
  refl))) (replace (sym (replace (replace 
- (appendNilRightNeutral [...] refl)))) refl)) 
+ (plusZeroRightNeutral [...] refl)))) refl)) 
 : (xs : List Int) -> (ys : List Int) -> 
@@
 
-And we effectively get what we wanted : a proof of equality between the two lists.
-As expected, this proof uses the properties of associativity ($appendAssociative$) and the property of neutrality of $Nil$ for $append$ ($appendNilRightAssociative$).
+And we effectively get the proof of equality we wanted.
+As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightAssociative$).
 

### 2014-07-06 newPaper/paper.tex
paper continued...

@@
 In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universally quantified natural numbers, the the properties of associativity and neutral element for the operation plus.
-For example, we would like to be able to automatically generate proofs of facts like $\forall x1\ x2\ x3 : List,\ (x1 + l2) ++ (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
-For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) ++ x3 = x1 ++ (x2 + x3)$ and with the fact that Z is a neutral element on the right side of the plus : $\forall x, x ++ Z = x$.
-Of course, this work could be extended by dealing with more axioms, and this is exactly what we will do in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
+For example, we would like to be able to automatically generate proofs of facts like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
+For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a neutral element on the right side of the plus : $\forall x, x + Z = x$.
+Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the "neutrality" of Z".
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the neutrality of Z".
 \\
 
+\subsection{Working by reflexion}
+
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-For the example 1, after abstraction of the variables, the goal becomes simply $(x1 + x2) ++ (x3 + l4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
-The general idea -which will also apply for the real bigger problem detailed in the next sections- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
-The goal of the normalization is of course to compute a canonical representation for a list $x$, such that any other list provably equal to $x$ (by using the two available properties) will have the same canonical representation.
+On the example 1, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
+The general idea --which will also apply for the more general and bigger problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
+The goal of the normalization is of course to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
 That's why it will be possible to decide the equality by simply comparing the two normal forms with a week syntactical equality.
@@
 Of course, the normalization will have to fulfil the following correctness property :
-$\forall x, norm\ x = x$ [1], which means that after normalization we obtain a list which is provably equal to the input list (potentially using our properties multiple times). \\
-And after normalization, we will compare the two resulting terms with a standard syntactical equality, and if they are equal, we will have obtained the desired proof.  \\
+$\forall x, norm\ x = x$ [1], which means that after normalization we obtain a number which is provably equal to the input number (potentially using our properties multiple times). \\
+And after normalization, we will compare the two resulting terms with a standard syntactical equality, and if they are equal, we will get the desired proof.  \\
 Indeed :
@@
 This way would not be easily feasible, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
-And the normalization function would need to do different treatments for a "variable natural number" (a number which has been universally quantified) and for a constant $Z$.
-This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality : this is a meta information of the AST that we can't access.
-For this reason, we will work by reflection, as we will do for for the real problem later.
-That means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
-This datatype will allow us to know the internal structure of the number, ie, where the variables and constants are.
+And the normalization function needs to do different treatments for a "variable natural number" (a number which has been universally quantified) and for the constant $Z$.
+This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality : this is a meta information of the ASTs representing the two terms that we can't access.
+For this reason, we will work by reflection, as we will do for the hierarchy of tactics in the next section.
+In this smaller example, it means that means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
+This datatype will allow us to know the internal structure of a number, ie, where the variables and constants are.
 Indeed, this datatype will allow pattern matching.
-Previously, we were only able to pattern match a list against the constructor $Z$ and $S$, which wasn't interesting for our goal.
-With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of number against the constructor Plus, Var and Zero, which gives us the informations we want.
+Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't interesting for our goal.
+With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
 
@@
 
-For this smaller problem, we want to prove equalities in which numbers are always universally quantified, so we only need to represent variable (a universally quantified natural), the constant Z, and the addition.
+For this smaller problem, we want to prove equalities in which numbers are always universally quantified, so we only need to represent variables (universally quantified natural numbers), the constant Z, and additions.
 Variables are represent using a De Brujin-like index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
 
-The type Expr is indexed over a vector of numbers (we will usually name it $G$), that represents our context.
-For example, if in the example 1, we will have to encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are declared.
+The type Expr is indexed over a vector of numbers (we will usually name it $G$), which represents our context.
+For example, in the example 1, we will have to encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are declared.
 The first element of this context denotes the variable $x1$, the second denotes $x2$, and so on.
@@
 $\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
-where $interpretation$ is a function computing the interpretation of an Expr, that is to say, the list this Expr is encoding.
-This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way we compute the interpretation.
-For this little example, the reduction will not be too heavy, but in the next sections, when we will have more properties than only associativity and neutral element to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+where $interpretation$ is a function computing the interpretation of an Expr in a context $G$, that is to say, the number that this Expr is encoding.
+This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
+For this little example, the reduction will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
 \\
@@
 - Write a "correct by construction" reduction procedure.
-The proof of preservation of the interpretation will be computed "bit by bit" at the same time as the function will produce the normalized expression (by rewriting the term thanks to the properties of associativity and neutral element).
+The proof of preservation of the interpretation will be computed bit by bit at the same time as the function will produce the normalized expression (by rewriting the term by using the two properties).
 
-The type for reflected lists therefore becomes :
-\begin{code}[caption=Second version of reflected lists with embedded denotation, captionpos=b, label=lst1:haskell2]
+Thus, the type for reflected numbers becomes :
+\begin{code}[caption=Second version of reflected number with embedded denotation, captionpos=b, label=lst1:haskell2]
 using (x : Nat, y : Nat, G : Vect n Nat)
-  data Expr : (G : Vect n Nat) -> Nat -> Type where
-       Plus : Expr G x -> Expr G y -> Expr G (x + y)
-       Var  : (i : Fin n) -> Expr G (index i G)
+  data Expr : (Vect n Nat) -> Nat 
+               -> Type where
+       Plus : Expr G x -> Expr G y -> 
+              Expr G (x + y)
+       Var  : (i : Fin n) -> 
+              Expr G (index i G)
        Zero : Expr G Z
 \end{code}
-For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the number $x$".
+For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $G$".
 When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
@@
 
+\subsection{A correct by construction approach}
+
 The reduction procedure is now supposed to be written on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
-Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- list : a term of type $Expr\ G\ x$ is the encoding of the number $x$.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ G\ x$ is the encoding of the number $x$.
 Thus, we can write the type of $reduce$ like this : \\
 $reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
-The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed by the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
-Note that this function can't simply produce an $Expr\ G\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since it can use the two available properties.
+The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
+Note that this function can't simply produce an $Expr\ G\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
 And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
@@
 \\
-Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality because they are in their normal form :
+Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality because these two expressions are in normal form :
 
-\begin{code}[caption=Syntactical equality, captionpos=b, label=lst1:haskell2]
-  eqExpr : (e : Expr G xs) -> (e' : Expr G ys) ->
-           Maybe (e = e')
-  eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
-    eqExpr (Plus x y) (Plus x y) | (Just refl, Just refl) = Just refl
-    eqExpr (Plus x y) (Plus x' y') | _ = Nothing
-  eqExpr (Var i) (Var j) with (prim__syntactic_eq _ _ i j)
-    eqExpr (Var i) (Var i) | (Just refl) = Just refl
-    eqExpr (Var i) (Var j) | _ = Nothing
+\begin{code}[caption=Syntactical equality between reflected terms, captionpos=b, label=lst1:haskell2]
+eqExpr : (e : Expr G xs) -> (e' : Expr G ys) 
+         -> Maybe (e = e')
+eqExpr (Plus x y) (Plus x' y') 
+        with (eqExpr x x', eqExpr y y')
+  eqExpr (Plus x y) (Plus x y) 
+        | (Just refl, Just refl) = Just refl
+  eqExpr (Plus x y) (Plus x' y') 
+        | _ = Nothing
+eqExpr (Var i) (Var j) with (decEq i j)
+  eqExpr (Var i) (Var i) 
+        | (Yes refl) = Just refl
+  eqExpr (Var i) (Var j) 
+        | _ = Nothing
 eqExpr Zero Zero = Just refl
@@
 
-And of course, if $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$
-By rewriting the two previously obtained equality $x=x'$ and $y=y'$ in the new equality $x'=y'$, we can get a proof of $x=y$.
+Now, if the two normalised expressions $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
+By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$.
 
 \begin{code}[caption=Building the desired proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
-  buildProof : {xs : Nat} -> {ys : Nat} ->
-               Expr G ln -> Expr G rn ->
-               (xs = ln) -> (ys = rn) -> Maybe (xs = ys)
+buildProof : {x : Nat} -> {y : Nat} 
+       -> Expr G x' -> Expr G y' 
+       -> (x = x') -> (y = y') 
+       -> Maybe (x = y)
 buildProof e e' lp rp with (eqExpr e e')
@@
 
-As we said, the proof for the metavariable is just the rewriting of the two equalities :
+As we mentioned, the proof for the metavariable is just a rewriting of the two equalities :
 
@@
 \end{code}
-Finally, the main function which tries to prove the equality $x=y$ is simply made of the reduction of the two metaified terms reflecting the left and the right hand side, and of the composition of the two obtained proofs thanks to the function $buildProof$:
+Finally, the main function which tries to prove the equality $x=y$ simply has to reduce the two metaified terms reflecting the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs we just obtained :
 \begin{code}[caption=testEq, captionpos=b, label=lst1:haskell2]
@@
 \begin{code}[caption=Reflected left associative numbers, captionpos=b, label=lst1:haskell2]
-  data LExpr : (G : Vect n Nat) -> Nat -> Type where
-       LPlus : LExpr G x -> (i : Fin n) -> LExpr G (x + index i G)
+data LExpr : (G : Vect n Nat) -> Nat 
+             -> Type where
+     LPlus : LExpr G x -> (i : Fin n) 
+             -> LExpr G (x + index i G)
      LZero : LExpr G Z
@@
 
-As part of the normalization, we will write a function $expr\_l$ converting an $Expr\ G\ x$ to a $LExpr\ G\ x'$ and producing a proof that $x=x'$.
-This function will therefore use the property of associativity multiple times, in order to obtain the expected fully left associative form.
+As part of the normalization, we will write a function $expr\_l$ which converts an $Expr\ G\ x$ to a $LExpr\ G\ x'$ and which produces a proof of $x=x'$.
+This function will therefore use the two available properties multiple times --and especially the property of associativity--, in order to obtain the expected fully left associative form.
 Of course, because by using this new datatype $LExpr$ we've changed the representation of our encoded lists, we will need to convert back an $LExpr\ G\ x$ to an $Expr\ G\ x$.
@@
 \begin{code}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
-  expr_l : Expr G xs -
-        > (xs' ** (LExpr G xs', xs = xs'))
+expr_l : Expr G x 
+         -> (x' ** (LExpr G x', x = x'))
 expr_l Zero = (_ ** (LZero, refl))
 expr_l (Var i) = (_ ** (LPlus LZero i, refl))
-  expr_l (Plus ex ey) = let (xl ** (xr, xprf)) = expr_l ex in
+expr_l (Plus ex ey) = 
+  let (xl ** (xr, xprf)) = expr_l ex in
   let (yl ** (yr, yprf)) = expr_l ey in
-                               PlusLExpr _ _ xr yr xprf yprf
+    plusLExpr _ _ xr yr xprf yprf
       where 
-      PlusLExpr : (x', y' : Nat) ->
-                 {G : Vect n Nat} -> {x, y : Nat} ->
-                 LExpr G x -> LExpr G y -> (x' = x) -> (y' = y) ->
-                 (w' ** (LExpr G w', x' + y' = w'))
-      PlusLExpr x' y' rx (LPlus e i) xprf yprf
-         = let (x ** (rec, prf)) = PlusLExpr _ _ rx e refl refl in
-               (_ ** (LPlus rec i, ?PlusLExpr1))
-      PlusLExpr x' y' rx LZero xprf yprf = (_ ** (rx, ?PlusLExpr2))
+      plusLExpr : (x', y' : Nat)
+            -> {G : Vect n Nat} -> {x, y : Nat} 
+            -> LExpr G x -> LExpr G y 
+            -> (x' = x) -> (y' = y) 
+            -> (w' ** (LExpr G w', x'+y'=w'))
+      plusLExpr x' y' rx (LPlus e i) xprf yprf =
+        let (x ** (rec, prf)) = 
+          plusLExpr _ _ rx e refl refl in
+          (_ ** (LPlus rec i, ?plusLExpr1))
+      plusLExpr x' y' rx LZero xprf yprf =
+        (_ ** (rx, ?plusLExpr2))
 
-  l_expr : LExpr G xs -> Expr G xs
+l_expr : LExpr G x -> Expr G x
 l_expr LZero = Zero
 l_expr (LPlus x i) = Plus (l_expr x) (Var i)
-
 \end{code}
 
-We notice that for transforming the number into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
+We notice that for transforming the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
 \\
@@
 We've got two metavariables to prove.
-The metavariable $appLExpr1$ requires us to prove the goal : $x' + y' = x + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x2\ ++\ index\ i\ G)$ and $(prf\ :\ x\ ++\ x2\ =\ x)$.
+The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = x2 + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ G)$ and $(prf\ :\ x\ +\ x1\ =\ x2)$.
 Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
 
-\begin{code}[caption=Proof of the metavariable appLExpr1, captionpos=b, label=lst1:haskell2]
-PlusRExpr1 = proof {
+\begin{code}[caption=Proof of the metavariable plusLExpr1, captionpos=b, label=lst1:haskell2]
+plusLExpr1 = proof {
   intros;
@@
   rewrite prf;
-  rewrite sym 
-       (plusAssociative x x2 (index i G));
-  exact refl;
+  mrefine plusAssociative;
 }
@@
 
-And the metavariable appRExpr2 uses the fact that $Z$ is a neutral element for the concatenation.
+And the metavariable plusLExpr2 uses the fact that $Z$ is a neutral element for the addition.
 
-\begin{code}[caption=Proof of the metavariable appLExpr2, captionpos=b, label=lst1:haskell2]
-PlusLExpr2 = proof
+\begin{code}[caption=Proof of the metavariable plusLExpr2, captionpos=b, label=lst1:haskell2]
+plusLExpr2 = proof
   intros
@@
 \\
+
+\subsection{Usage of the "tactic"}
+
 It's now time to see how to use this minimalist "tactic".
-Let's define two expressions $e1$ and $e2$, respectively representing the numbers $((x + ys) + (x + z))$ and $(x + ((y + x) + z))$ in the context $[x, y, z]$ of three abstracted variables.
+Let's define two expressions $e1$ and $e2$, respectively representing the numbers $((x + y) + (x + z))$ and $(x + ((y + x) + z))$ in the context $[x, y, z]$ of three abstracted variables.
 
 \begin{code}[caption=Two test expressions, captionpos=b, label=lst1:haskell2]
-  e1 : (xs, ys, zs : Nat) -> 
-           Expr [xs, ys, zs] ((xs + ys) + (xs + zs))
-  e1 xs ys zs = Plus (Plus (Var fZ) (Var (fS fZ))) 
-                        (Plus (Var fZ) (Var (fS (fS fZ))))
+e1 : (x, y, z : Nat) 
+    -> Expr [x, y, z] ((x+y) + (x+z))
+e1 x y z = Plus (Plus (Var fZ) 
+                      (Var (fS fZ))) 
+                (Plus (Var fZ) 
+                      (Var (fS (fS fZ))))
 
-  e2 : (xs, ys, zs : Nat) -> 
-           Expr [xs, ys, zs] (xs + ((ys + xs) + zs))
-  e2 xs ys zs = Plus (Var fZ) 
-         (Plus (Plus (Var (fS fZ)) (Var fZ)) (Var (fS (fS fZ))))
+e2 : (x, y, z : Nat) 
+     -> Expr [x, y, z] (x + ((y + x) + z))
+e2 x y z = Plus (Var fZ) 
+                (Plus (Plus (Var (fS fZ)) 
+                            (Var fZ)) 
+                      (Var (fS (fS fZ))))
 \end{code}
@@
 \begin{code}[caption=Test of equality betwen e1 and e2, captionpos=b, label=lst1:haskell2]
-  e1_e2_testEq : (xs, ys, zs : Nat) ->
-          Maybe (((xs + ys) + (xs + zs)) = (xs + ((ys + xs) + zs)))
-  e1_e2_testEq xs ys zs = testEq (e1 xs ys zs) (e2 xs ys zs)
+e1_e2_testEq : (x, y, z : Nat) 
+       -> Maybe (((x + y) + (x + z)) 
+                  = (x + ((y + x) + z)))
+e1_e2_testEq x y z = testEq (e1 x y z)
+                            (e2 x y z)
 \end{code}
@@
 \begin{code}[caption=Evaluation of the result, captionpos=b, label=lst1:haskell2]
-\xs => \ys => \zs => e1_e2_testEq xs ys zs
+\x => \y => \z => e1_e2_testEq x y z
 
-\xs => \ys => \zs => Just (replace (sym (replace 
-(sym (replace (replace (plusZeroRightNeutral xs)
- refl (replace (sym (plusAssociative xs [] ys)) 
+\x => \y => \z => Just (replace (sym (replace 
+(sym (replace (replace (plusZeroRightNeutral x)
+ refl (replace (sym (plusAssociative x [] y)) 
  refl))) (replace (sym (replace (replace 
  (plusZeroRightNeutral [...] refl)))) refl)) 
-: (xs : List Int) -> (ys : List Int) -> 
-  (zs : List Int) -> 
-  Maybe ((xs ++ ys) ++ xs ++ zs 
-         = xs ++ (ys ++ xs) + zs)
+: (x : List Int) -> (y : List Int) -> 
+  (z : List Int) -> 
+  Maybe ((x + y) + x + z 
+         = x+ (y + x) + z)
 \end{code}

### 2014-07-07 newPaper/paper.tex
paper continued

@@
       plusLExpr x' y' rx (LPlus e i) xprf yprf =
-        let (x ** (rec, prf)) = 
+        let (xrec ** (rec, prf)) = 
           plusLExpr _ _ rx e refl refl in
@@
 We've got two metavariables to prove.
-The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = x2 + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ G)$ and $(prf\ :\ x\ +\ x1\ =\ x2)$.
+The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ G)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
 Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
@@
 And if we ask for the evaluation of this term, we should obtain $Just$ and a proof of equality between the two underlying numbers.
-\begin{code}[caption=Evaluation of the result, captionpos=b, label=lst1:haskell2]
-\x => \y => \z => e1_e2_testEq x y z
+\begin{code}[caption=Obtained proof, captionpos=b, label=lst1:haskell2]
+#\x => \y => \z => e1_e2_testEq x y z
 
 \x => \y => \z => Just (replace (sym (replace 
-(sym (replace (replace (plusZeroRightNeutral x)
- refl (replace (sym (plusAssociative x [] y)) 
- refl))) (replace (sym (replace (replace 
- (plusZeroRightNeutral [...] refl)))) refl)) 
-: (x : List Int) -> (y : List Int) -> 
-  (z : List Int) -> 
-  Maybe ((x + y) + x + z 
+  (sym (replace (plusZeroRightNeutral x) 
+  (plusAssociative x 0 y))) (replace 
+  (sym (replace (plusZeroRightNeutral x) 
+  (plusAssociative x 0 z))) (replace (replace 
+  (plusZeroRightNeutral (plus xy)) 
+  (plusAssociative (plus x y) 0 x)) [...]
+: (x : Nat) -> (y : Nat) -> (z : Nat) 
+  -> Maybe ((x + y) + x + z 
             = x+ (y + x) + z)
@@
 And we effectively get the proof of equality we wanted.
-As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightAssociative$).
+As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightNeutral$).
 
@@
 
-Now that we have introduced our ideas on this little example, it is time to apply them for our main goal : the implementation of a hierarchy of tactics proving equalities in algebraic structures.
+Now that we have introduced our ideas on this little example, it is time to apply them for our main goal, more general : the implementation of a hierarchy of tactics proving equalities in algebraic structures.
 Very often, the properties available on a given type are the one of a well known structure : magma, semi-group, monoid, group, commutative-group...
-We will no longer only work with the properties of associativity and neutral elements for Lists as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.
-In a Magma, we only have a $Plus$ operation, and no axioms with it.
-With a Semi-Group, we will have one axiom about this Plus : the property of associativity.
-In a Monoid, we will have one more property : the fact that we have a neutral element.
-In fact, this property will be expressed as 3 facts : there is a value $Zero$, and we've got $\forall x, Zero + x = x$ and $\forall x, x + Zero = x$.
-\\
+We will no longer only work with the properties of associativity and of right neutral element for natural numbers as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.
 
@@
 
-Note that this equality is a "week" equality : it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different.
-That's quite natural, since we want to generate proof of equality, and not to generate counter examples for dis-equalities, which is another problem.
+Note that this equality is a "week" equality : it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different -- instead, it simply produces the value $Nothing$--.
+That's quite natural, since we want to generate proof of equality, and not to generate counter examples for proving dis-equalities, which is another problem.
 
@@
 That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$.
-The second axiom, which is the most important one for a Group, is the fact that any $c1$ admits $Neg\ c1$ as an inverse, where being an inverse if the following property :
+The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :
 
@@
 -- This is just a conjunctive predicate
-hasSymmetric : (c:Type) -> (p:dataTypes.Monoid c) 
+hasSymmetric : (c:Type) -> (p:Monoid c) 
                 -> c -> c -> Type
@@
 
-class dataTypes.Monoid c => dataTypes.Group c where
+class Monoid c => Group c where
 	Minus : c -> c -> c
@@
 	Plus_inverse : (c1:c) -> 
-	             hasSymmetric c (instance) c1 (Neg c1)
+	             hasSymmetric c _ c1 (Neg c1)
 \end{code}
@@
 
-As discussed in the section above, the algorithm of normalization will not directly use the concrete value of type $T$, but a reflected term, indexed over the concrete value.
+As discussed in the section 2.1 on the smaller problem, the algorithm of normalization will not directly use the concrete value of type $T$, but a reflected term, indexed over the concrete value.
 

### 2014-07-07 newPaper/paper.tex
Again, paper continued...

@@
 \begin{code}[caption=First version of reflected numbers, captionpos=b, label=lst1:haskell2]
-data Expr : (G : Vect n Nat) -> Type where
-     Plus : {n:Nat} -> {G:Vect n Nat} -> 
-            Expr G -> Expr G -> Expr G
-     Var  : {n:Nat} -> {G:Vect n Nat} -> 
-            (i : Fin n) -> Expr G
-     Zero : {n:Nat} -> (G:Vect n Nat) -> 
-            Expr G Z
+data Expr : (gam : Vect n Nat) -> Type where
+     Plus : {n:Nat} -> {gam:Vect n Nat} -> 
+            Expr gam -> Expr gam -> Expr gam
+     Var  : {n:Nat} -> {gam:Vect n Nat} -> 
+            (i : Fin n) -> Expr gam
+     Zero : {n:Nat} -> (gam:Vect n Nat) -> 
+            Expr gam Z
 \end{code}
@@
 
-The type Expr is indexed over a vector of numbers (we will usually name it $G$), which represents our context.
+The type Expr is indexed over a vector of numbers (we will usually name it $\Gamma$), which represents our context.
 For example, in the example 1, we will have to encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are declared.
@@
 \begin{code}[caption=Reflected LHS of example 1, captionpos=b, label=lst1:haskell2]
-App (App (Var G fZ) 
-         (Var G (fS fZ))) 
-    (App (Var G (fS (fS fZ))) 
-         (Var G (fS (fS (fS fZ)))))
+App (App (Var gam fZ) 
+         (Var gam (fS fZ))) 
+    (App (Var gam (fS (fS fZ))) 
+         (Var gam (fS (fS (fS fZ)))))
 \end{code}
@@
 If, as it is probably expected, this function simply takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards, corresponding to the correctness property mentioned before : \\
-$\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
-where $interpretation$ is a function computing the interpretation of an Expr in a context $G$, that is to say, the number that this Expr is encoding.
+$\forall\ e:Expr\ \Gamma,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
+where $interpretation$ is a function computing the interpretation of an Expr in a context $\Gamma$, that is to say, the number that this Expr is encoding.
 This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
@@
 \begin{code}[caption=Second version of reflected number with embedded denotation, captionpos=b, label=lst1:haskell2]
-using (x : Nat, y : Nat, G : Vect n Nat)
+using (x : Nat, y : Nat, gam : Vect n Nat)
   data Expr : (Vect n Nat) -> Nat 
                -> Type where
-       Plus : Expr G x -> Expr G y -> 
-              Expr G (x + y)
+       Plus : Expr gam x -> Expr gam y -> 
+              Expr gam (x + y)
        Var  : (i : Fin n) -> 
-              Expr G (index i G)
-       Zero : Expr G Z
+              Expr gam (index i gam)
+       Zero : Expr gam Z
 \end{code}
-For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $G$".
-When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
+For an expression $e\ :\ Expr\ \Gamma\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $\Gamma$".
+When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
 Also, the $Zero$ expression denotes the natural number $Z$.
@@
 The reduction procedure is now supposed to be written on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
-Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ G\ x$ is the encoding of the number $x$.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
 Thus, we can write the type of $reduce$ like this : \\
-$reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
-The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
-Note that this function can't simply produce an $Expr\ G\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
+$reduce\ :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ =\ x'))$ \\
+The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
+Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
 And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
@@
 We will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
-We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
-We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ G\ y'$, and a proof of $y=y'$.
+We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
+We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
 \\
@@
 \begin{code}[caption=Syntactical equality between reflected terms, captionpos=b, label=lst1:haskell2]
-eqExpr : (e : Expr G xs) -> (e' : Expr G ys) 
+eqExpr : (e : Expr gam x) -> (e' : Expr gam y) 
          -> Maybe (e = e')
@@
 buildProof : {x : Nat} -> {y : Nat} 
-       -> Expr G x' -> Expr G y' 
+       -> Expr gam x' -> Expr gam y' 
        -> (x = x') -> (y = y') 
@@
 \begin{code}[caption=testEq, captionpos=b, label=lst1:haskell2]
-  testEq : Expr G x -> Expr G y 
+  testEq : Expr gam x -> Expr gam y 
            -> Maybe (x = y)
@@
 \begin{code}[caption=Reflected left associative numbers, captionpos=b, label=lst1:haskell2]
-data LExpr : (G : Vect n Nat) -> Nat 
+data LExpr : (gam : Vect n Nat) -> Nat 
              -> Type where
-     LPlus : LExpr G x -> (i : Fin n) 
-             -> LExpr G (x + index i G)
-     LZero : LExpr G Z
+     LPlus : LExpr gam x -> (i : Fin n) 
+             -> LExpr gam (x + index i gam)
+     LZero : LExpr gam Z
 \end{code}
@@
 
-As part of the normalization, we will write a function $expr\_l$ which converts an $Expr\ G\ x$ to a $LExpr\ G\ x'$ and which produces a proof of $x=x'$.
+As part of the normalization, we will write a function $expr\_l$ which converts an $Expr\ \Gamma\ x$ to a $LExpr\ \Gamma\ x'$ and which produces a proof of $x=x'$.
 This function will therefore use the two available properties multiple times --and especially the property of associativity--, in order to obtain the expected fully left associative form.
-Of course, because by using this new datatype $LExpr$ we've changed the representation of our encoded lists, we will need to convert back an $LExpr\ G\ x$ to an $Expr\ G\ x$.
+Of course, because by using this new datatype $LExpr$ we've changed the representation of our encoded lists, we will need to convert back an $LExpr\ \Gamma\ x$ to an $Expr\ \Gamma\ x$.
 The function $l\_expr$ will do this easy task.
@@
 \begin{code}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
-expr_l : Expr G x 
-         -> (x' ** (LExpr G x', x = x'))
+expr_l : Expr gam x 
+         -> (x' ** (LExpr gam x', x = x'))
 expr_l Zero = (_ ** (LZero, refl))
@@
       plusLExpr : (x', y' : Nat)
-            -> {G : Vect n Nat} -> {x, y : Nat} 
-            -> LExpr G x -> LExpr G y 
+            -> {gam : Vect n Nat} -> {x, y : Nat} 
+            -> LExpr gam x -> LExpr gam y 
             -> (x' = x) -> (y' = y) 
-            -> (w' ** (LExpr G w', x'+y'=w'))
+            -> (w' ** (LExpr gam w', x'+y'=w'))
       plusLExpr x' y' rx (LPlus e i) xprf yprf =
@@
 
-l_expr : LExpr G x -> Expr G x
+l_expr : LExpr gam x -> Expr gam x
 l_expr LZero = Zero
@@
 We've got two metavariables to prove.
-The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ G)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
+The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ \Gamma)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
 Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
@@
 \begin{code}[caption=Reduction function, captionpos=b, label=lst1:haskell2]
-  reduce : Expr G x -> 
-           (x' ** (Expr G x', x = x'))
+  reduce : Expr gam x -> 
+           (x' ** (Expr gam x', x = x'))
   reduce e = 
@@
 
-We will have one datatype for reflected terms in each algebraic structure. \\
-Each of these datatype is parametrised on a type $c$ : the real type on which we want to prove equalities.
-It is also indexed over an instance of the corresponding typeclass for $c$ (we usally call it $p$, because it looks like a proof saying that the structure $c$ has the desired properties), and indexed over a context (a vector $g$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the encoded value.
-A magma is only equiped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constantes, variables, and addition.
+We will have one datatype for reflecting terms in each algebraic structure. \\
+Each of these datatype is parametrised over a type $T$, which is the real type on which we want to prove equalities.
+It is also indexed over an instance of the corresponding typeclass for $T$ (we usually call it $p$, because it looks like a proof saying that the structure $T$ has the desired properties), and indexed over a context (a vector $\Gamma$ of $n$ elements of type $T$), and also indexed over a value of type $T$, which is precisely the encoded value.
+A magma is only equipped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and addition.
 
 \begin{code}[caption=Reflected terms in a Magma, captionpos=b, label=lst1:haskell2]  
-data ExprMa : Magma c -> (Vect n c) -> c -> 
+data ExprMa : Magma T -> (Vect n T) -> T -> 
               Type where
-    ConstMa : (p : Magma c) -> (g:Vect n c) 
-         -> (c1:c)  -> ExprMa p g c1 
-    PlusMa : {p : Magma c} -> {g:Vect n c}
-         -> {c1:c} -> {c2:c} 
-         -> ExprMa p g c1 
-         -> ExprMa p g c2 
-         -> ExprMa p g (Plus c1 c2) 
-    VarMa : (p:Magma c) -> {g:Vect n c}
-         -> {c1:c} 
-         -> VariableA (get_eq_magma p) g c1
-         -> ExprMa p g c1
+    ConstMa : (p : Magma T) -> (gam:Vect n T) 
+         -> (c1:T)  -> ExprMa p gam c1 
+    PlusMa : {p : Magma c} -> {gam:Vect n c}
+         -> {c1:T} -> {c2:T} 
+         -> ExprMa p gam c1 
+         -> ExprMa p gam c2 
+         -> ExprMa p gam (Plus c1 c2) 
+    VarMa : (p:Magma T) -> {gam:Vect n T}
+         -> {c1:T} 
+         -> Variable (get_eq_magma p) gam c1
+         -> ExprMa p gam c1
 \end{code}
@@
 When we encode a constant $c1$ in a context $\Gamma$, we use the constructor $ConstMa$ to produce a term of type $ExprMa\ p\ \Gamma\ c1$ : the index representing the concrete value is simply the constant $c1$.
-If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the courant instance $p$.
+If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the current instance $p$.
 
@@
                     ->Maybe(c1=c2))
-          -> (g:Vect n c) -> (i:Fin n) 
-          -> Variable c_equal g (index i g) 
+          -> (gam:Vect n c) -> (i:Fin n) 
+          -> Variable c_equal gam (index i gam) 
 \end{code}
 
-This type will be extended with some other constructors later on. For the moment, we can only express a "real variable". \\
+This type will be extended with some other constructors later on. For the moment, we can only express a "real variable", referred by its index $i$. \\
 \\
@@
     ConstG : (p : dataTypes.Group c) 
-       -> (g:Vect n c)
-       -> (c1:c) -> ExprG p g c1
-    PlusG : {p : Group c} -> {g:Vect n c} 
+      -> (gam:Vect n c)
+      -> (c1:c) -> ExprG p gam c1
+    PlusG : {p : Group c} -> {gam:Vect n c} 
       -> {c1:c} -> {c2:c} 
-       -> ExprG p g c1 
-       -> ExprG p g c2 
-       -> ExprG p g (Plus c1 c2)
-    MinusG : {p : Group c} -> {g:Vect n c} 
+      -> ExprG p gam c1 
+      -> ExprG p gam c2 
+      -> ExprG p gam (Plus c1 c2)
+    MinusG : {p : Group c} -> {gam:Vect n c} 
       -> {c1:c} -> {c2:c} 
-       -> ExprG p g c1 
-       -> ExprG p g c2 
-       -> ExprG p g (Minus c1 c2)
-    NegG : {p : Group c} -> {g:Vect n c} 
-       -> {c1:c} -> ExprG p g c1 
-       -> ExprG p g (Neg c1)
-    VarG : (p : Group c) -> {g:Vect n c} 
+      -> ExprG p gam c1 
+      -> ExprG p gam c2 
+      -> ExprG p gam (Minus c1 c2)
+    NegG : {p : Group c} -> {gam:Vect n c} 
+      -> {c1:c} -> ExprG p gam c1 
+      -> ExprG p gam (Neg c1)
+    VarG : (p : Group c) -> {gam:Vect n c} 
       -> {c1:c}
-       -> Variable (get_eq_group p) Neg g c1 
-       -> ExprG p g c1
+     -> Variable (get_eq_group p) Neg gam c1 
+      -> ExprG p gam c1
 \end{code}
 
-Of course, the index of type $c$ (the real value represented by the expression) is always expressed with the oeprations available in the typeclass $p$, and here for a Group they are$Plus$, $Minus$ and $Neg$.
+Of course, the index of type $T$ (the real value represented by the expression) is always expressed with the operations available in the typeclass $p$, and here for a Group they are$Plus$, $Minus$ and $Neg$.
 
 	\subsection { A bit of notation}
-The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (making $c$ a Semi-Group, a Monoid, a Group...).
-The fact that $c$ fulfil the specification of an algebraic structure is expressed as an instance of the corresponding typeclass, and this instance will be denoted as $p$.
-The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ g\ x$ : $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$.
+The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $T$, which  simulates a set with some properties (which make $T$ being a Semi-Group, a Monoid, a Group...).
+The fact that $T$ fulfil the specification of an algebraic structure is expressed as an instance of the corresponding typeclass, and this instance will be denoted as $p$.
+The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ \Gamma\ x$. $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$.
 We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ g\ x'$ and a proof of $p1:x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ g\ y'$ and a proof of $p2:y=y'$.
+Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p1$ of $x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p2$ of $y=y'$.
 
@@
 This is fact a nice property that all these algebraic structures admit a canonical represent for any element.
-Without this property, it would become a lot more complicated to decide the equality without bruteforcing the
+Without this property, it would become a lot more complicated to decide the equality without bruteforcing a serie of rewriting.
 These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
@@
 	groupReduce : {c:Type} -> {n:Nat} 
-	  -> (p:Group c) -> {g:Vect n c} -> {x:c}
-	  -> (ExprG p g x) 
-	  -> (x' ** (ExprG p g x', x=x'))
+	  -> (p:Group c) -> {gam:Vect n c} -> {x:c}
+	  -> (ExprG p gam x) 
+	  -> (x' ** (ExprG p gam x', x=x'))
 \end{code}
 
-This function will have more job to do in structures with multiple axioms (like Group), than it has in the easier structure underneath.
+This function will have more job to do in structures with multiple axioms (like Group), than it has to do in the easier structure underneath.
 The detail of these functions will be given in the next section.
 For the moment, we just assume that we've got such a function for each structure, and we want to decide equality between elements.
-We continue to give the details for the Group, but the principle is exactly the same with the other structures.
+We continue to give here the details for the Group structure, but the principle is exactly the same with the other structures.
 
@@
 \begin{code}[caption=Type of the function for deciding equality between elements in a Group, captionpos=b, label=lst1:haskell2]
-	groupDecideEq : (p:Group c) -> {g:Vect n c} 
-	   -> {x : c} -> {y : c} 
-	   -> (ExprG p g x) -> (ExprG p g y) 
+	groupDecideEq : (p:Group c) 
+	   -> {gam:Vect n c} -> {x : c} -> {y : c} 
+	   -> (ExprG p gam x) -> (ExprG p gam y) 
 	   -> Maybe (x = y)
@@
 The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
-Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by "composing" the two proofs $p1$ and $p2$.
+Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $p1$ and $p2$.
 
@@
 \begin{code}[caption=Composes the two proofs if the NF are the same, captionpos=b, label=lst1:haskell2]
-buildProofGroup : (p:Group c) -> {g:Vect n c} 
-  -> {x : c} -> {y : c} 
-  -> {x':c} -> {y':c} 
-  -> (ExprG p g x') 
-  -> (ExprG p g y') 
+buildProofGroup : (p:Group c) -> {gam:Vect n c} 
+  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
+  -> (ExprG p gam x') -> (ExprG p gam y') 
   -> (x = x') -> (y = y') 
@@
 
-The proof for the metavariable MbuildProofGroup is exactly the same as the corresponding one (called $bp$) in the smallest example showed in section 2.
+The proof for the metavariable MbuildProofGroup is exactly the same as the corresponding one (called $bp$) in the smallest example showed in section 2.2.
 
@@
 
+The normalization in a Magma nearly has nothing to do.
+It simply need to use the operation $Plus$ between constants under brackets.
+Indeed, we don't have any property of rewriting here, so we can only compute constants.
+
+\begin{code}[caption=Normalization function for Magma, captionpos=b, label=lst1:haskell2]
+magmaReduce : {c:Type} -> {n:Nat} 
+              -> {p:Magma c} -> {gam:Vect n c} 
+              -> {c1:c} -> (ExprMa p gam c1) 
+              -> (c2 ** (ExprMa p gam c2, c1=c2))
+magmaReduce (PlusMa
+      (ConstMa p _ const1) 
+      (ConstMa p _ const2)) = 
+         (_ ** (ConstMa p g 
+               (Plus const1 const2), refl))
+magmaReduce (PlusMa neg e1 e2) = 
+    let (r_ih1 ** (e_ih1, p_ih1)) 
+        = (magmaReduce e1) in
+    let (r_ih2 ** (e_ih2, p_ih2)) 
+        = (magmaReduce e2) in
+    ((Plus r_ih1 r_ih2) ** 
+        (PlusMa neg e_ih1 e_ih2, ?MmagmaReduce1))                    
+magmaReduce e = (_ ** (e, refl))
+\end{code}
+
+This function returns the same value for constant or for a variable.
+For a $Plus$, we apply the treatment recursively, except if it's a $Plus$ between two constants, in which case we compute the addition of these two constants.
+
 	\subsection {Normalization of terms in a Semi-Group}
+
+In a semigroup, we have to deal with the new property of associativity.
+We have to use the $Plus$ on constants computable thanks to associativity.
+Three possible patterns need this treatment : $(x+c1)+(c2+y)$ , $(x + c1) + c2$  and $c1 + (c2 + x)$ when $c1$ and $c2$ denote constants.
+We simply rearrange the parentheses, and call the normalization procedure for Magma on the constants, which is denoted [c1+c2].
+The first pattern is rewritten into $(x + [c1+c2]) + y$, the second into $x + [c1+c2]$ and the third one into $[c1 + c2] + x$.
+We show the code for the first pattern.
+\begin{code}[caption=Computing with associativity in a Semi-Group, captionpos=b, label=lst1:haskell2]
+assoc : (p:SemiGroup c) -> (gam:Vect n c) 
+        -> {c1:c} -> (ExprSG p gam c1) 
+        -> (c2 ** (ExprSG p gam c2, c1=c2))
+assoc p gam (PlusSG 
+             (PlusSG e1 
+                        (ConstSG _ _ c1)) 
+             (PlusSG (ConstSG _ _ c2) e2)) =
+    let (r_ih1 ** (e_ih1, p_ih1)) 
+       = (assoc p gam e1) in
+    let (r_ih2 ** (e_ih2, p_ih2)) 
+       = (assoc p gam e2) in
+    let (r_3 ** (e_3, p_3)) 
+       = magmaReduce (semiGroup_to_magma 
+            (PlusSG (ConstSG _ _ c1) 
+            (ConstSG _ _ c2))) in
+    let e_3' = magma_to_semiGroup p e_3 in
+    (_ ** ((PlusSG (PlusSG e_ih1 e_3') e_ih2), 
+          ?Massoc1))
+[...]
+\end{code}
+
+Note that we use function of conversions between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$.
+
+After simplifying the constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way.
+The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described on section 2.2.
 
@@
 
+In a monoid, we simply have to eliminate zeros, thanks to the two properties of left and right neutrality.
+Because zeros will be computed with constants automatically at the level underneath, we only have to simplify the addition between the constant zero and a variable.
+There is in fact two cases, since the variable can come first, or the $Zero$ can come first.
+
+\begin{code}[caption=Eliminating zeros in a Monoid, captionpos=b, label=lst1:haskell2]
+elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
+            -> (ExprMo p gam c1) 
+            -> (c2 ** (ExprMo p gam c2, c1=c2))
+elimZero c p gam 
+ (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
+        with (eq_on_set p Zero const1)
+   elimZero c p gam (PlusMo (ConstMo _ _ Zero) 
+                 (VarMo _ v)) | (Just refl) 
+          = (_ ** (VarMo _ v, ?MelimZero1))
+   elimZero c p gam (PlusMo (ConstMo _ _ const1) 
+                 (VarMo _ v)) | _ 
+          = (_ ** (PlusMo (ConstMo _ _ const1) 
+               (VarMo _ v), refl)) 
+elimZero c p gam 
+ (PlusMo (VarMo _ v) (ConstMo _ _ const2)) 
+      with (eq_as_elem_of_set p Zero const2) 
+   elimZero c p gam (PlusMo (VarMo _ v) 
+            (ConstMo _ _ Zero)) | (Just refl) 
+          = (_ ** (VarMo _ _ v, ?MelimZero2))
+   elimZero c p gam (PlusMo (VarMo _ v) 
+            (ConstMo _ _ const2)) | _ 
+          = (_ ** (PlusMo (VarMo _ v) 
+                 (ConstMo _ _ const2), refl))
+[...]
+\end{code}
+
+Now the normalization of reflected expressions in a monoid is simply made of a call to the function reducing terms on a semigroup, followed by a call to elimZero which will eliminate the remaining zeros.
+
 	\subsection {Normalization of terms in a Group}
 
-\section {Auxiliary problems : Automatic metaification and coating (1/2 pages)}
+TO BE WRITTEN
+
+\section{Completeness}
+
+Mentionning that we hope to be complete, and giving intuitive arguments briefly, but explaining that this can't be proven inside the language.
+
+\section {Auxiliary problems : Automatic metaification and coating (a bit less than a page!)}
 
@@
 
+How to compute automatically the metaified expression, by using the reflection stuff (0,5 page)
+
 	\subsection {Coating : transforming this stuff into real tactics}
 
-\section {Related work (1 page)}
+What we have to add to Idris in order to invoke this stuff as a "real tactic" (which can fail). (0,25 page)
 
-\section {Conlusion (0,5/1 page)}
+\section {Related work and comparison with other work(less than 1 page)}
+
+TO BE CONTINUED
+
+There exists implementation of ring solver for the Coq proof assistant and for Agda.
+The first one, implemented by Benjamin Grégoire and Assia Mahboubi and described in their paper "Proving equalities in a commutative ring done right in coq", is also implemented by reflection.
+The main difference is that they write a function of normalization for their reflected terms, and latter on prove that this normalization preserves the interpretation, ie, the concrete value reflected.
+This approach is quite normal for the Coq proof assistant, since it does not have a good support for dependent types.
+For example, there is no dependent pattern matching in Coq.
+However, in Idris is designed as a programming language with full support for dependent types, and that's why the temptation was high to use this good support in order to write correct by construction tactics.
+Also, in most implementations, only one algebraic level is treated.
+In the best case, they deal with rings and commutative rings.
+But is a given structure only has the properties of a group, their tactic is not usable.
+One could of course write a solver for group, but it will have a big part similar to the ring solver, but this part will have to be duplicated.
+However, in our work, each solver uses the solver of the algebraic structures underneath, increasing drastically the re-usability of code.
+
+\section {Conlusion and futher work (0,5 page)}
+TO BE CONTINUED
+
+We've constructed a correct by construction hierarchy of tactics.
+This hierarchy is going to be extended  with structures like ring and commutative ring.
+The main contribution of our work was to intensively use dependent types in order to capture the concrete value reflected by an expression.
+This enabled us to write the implementation of the tactics on a "correct by construction" style, where the proof is almost trivial and computed at the same time as the result.
+It is true that dependent types often bring more complexity, and for example, the lemmas that we have to proven as soon as one is playing with dependent types.
+Pleasantly, if the problem comes with dependent types, the solution is also accompanied with the use dependent types which helped us to write a collection of tactics without pain.
 
@@
 \appendix
-\section{Appendix Title}
+\section{Appendix : A little overview of Idris}
 
-This is the text of the appendix, if you need one.
+\input{./content/idrisOverview.tex}
 

### 2014-12-13 ICFP15paper/content/introduction.tex
for Edwin : Compiling version (with Idris 0.9.15.1) for automatic reflection of terms

--- 2014-07-03 newPaper/content/introduction.tex
@@
 These proofs are the every day routine when we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
-The indices can be for example natural numbers, and in this case the lemmas that we have to prove about these indices are arithmetic statements which use the properties of natural numbers and their usual operations : symmetry and commutativity for +, existence of neutral element for +, etc.
+For example, If a type $T$ is indexed over natural numbers, then the definition of a function which has co-domain $T$ will often require to prove that the index of the produced term is equal to the expected index.
+This proof will use properties of natural numbers and their usual operations : symmetry and commutativity for +, existence of neutral element for +, etc.
 
 \subsection{A motivating little example}
-We introduce a little example which shows how these proofs appear when we index a type by values like natural numbers.
+We introduce a little example which shows how these lemmas appear when we index a type by values like natural numbers.
 Let's define the types of $Bit$ and $Binary$ number, both indexed over the value they represent, expressed in base ten as a natural number.
@@
 \end{code}
-Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition (the $value$ is still $n$).
+Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition since the $value$ is still $n$.
 
@@
 The first component is the value $bX$ represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
-This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing the two bits produced and the proof that these two bits are effectively encoding the result of the addition we had to perform.
+This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing finally the two bits produced and the proof that these two bits are effectively encoding the result of the addition we had to perform.
 For example, with the line corresponding to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
@@
 Unfortunately, the types are mismatching for the two cases expressed by this pattern matching.
-The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] and we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\] [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
+The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\] [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
 The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
-For this kind of situation, Idris provides the possibility to write a provisional definition --provisional definitions are described with the brief overview of the language in page XX-- with the syntax "?=".
-Using a provisional definitions here will make the definition being accepted, but we will have to prove afterwards that the term provided can be converted into a term which has the right type.
+For this kind of situation, Idris provides the possibility to write a provisional definition\footnote{provisional definitions are described with the brief overview of the language in page XX} with the syntax "?=".
+Using a provisional definitions here will make the definition being accepted, but we will have to prove afterwards that in both patterns, the term provided can be converted into a term which has the right type.
 However, for both of these two patterns, it is possible to prove the equality between the expected and the provided types, and this is even stronger than we need.
-Using this equality of types it will be trivial to produce a term of the expected type by using the provided term and by simply rewriting its type with this proof.
-These proofs will be made of a series of rewriting, and each rewriting will use a property about natural numbers and the usual operations.
+These proofs will be made of a series of rewriting, and each rewriting will use a property about the operation $+$ on natural numbers.
 For example, the proof for the second pattern look like this :
-\begin{code}[caption=Proof by hand of the correspondence between expected and provided types for the second line of the definition of adc, captionpos=b, label=lst1:haskell2]
+\begin{code}[caption=Proof by hand of the correspondence between expected and provided types for the second pattern of the definition of adc, captionpos=b, label=lst1:haskell2]
 adc_lemma_2 = proof {
@@
      -> Binary (S w) (c + x + y)
-adc zero zero carry = zero # carry 
+adc zero zero carry ?= zero # carry 
         by CommutativeGroupSolver
 adc (numx # bX) (numy # bY) carry
-   = let (v1 ** (v2 ** (carry0, lsb, _))) = 
+   ?= let (v1 ** (v2 ** (carry0, lsb, _))) = 
       addBit bX bY carry in
@@
 \end{code}
-We would like to let a solver for commutative groups to find the proof on its own, and thereby, there will be no extra lemmas to prove afterwards.
+We would like to let a solver for commutative groups to find the proof on its own.
 This is a something doable because algebraic structures like monoid, group, commutative group and ring have all a very nice property : every element in them has a canonical represent.
@@
 In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
-The interesting thing that we're going to present in this paper lies in the way we construct it, in a "correct by construction" style.
+The interesting thing that we're going to present in this paper lies in the way we construct it, which is a "correct by construction" style, with the same spirit as the functions we've written above.
 
 In this paper, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group and Abelian group.
-The user will just have to prove that his structure satisfies the properties of an algebraic structure, and he will be allowed to invoke the corresponding tactic.
 

### 2014-12-13 Old/OldPapers/paper/content/method.tex
for Edwin : Compiling version (with Idris 0.9.15.1) for automatic reflection of terms

\section{An overview of the method}

Our goal is to implement the collection of tactics in Idris itself, instead of writing them in Haskell (the host language of Idris). The main interest of writing the tactic in Idris is that it becomes possible to certify the tactic by using the logical framework provided by the Idris language. Because we want to implement and certify the tactic in the language itself, we will need to work by reflection.
% Ici on donne l'idée générale. Nom de la section peut etre à revoir

%The user of one of our tactic -say the group solver- will want to use the tactic for his own specific structure. He will just have to prove that his structure respects the axioms of a group, and the solver will try to solve the equality for himself.

\subsection{why working by reflection}

As we mentioned, our decision procedure for a specific algebraic structure -say, a group- will essentially works by reducing the left and the right hand side of the equality $t1 = t2$ to their normal forms. These two terms $t1$ and $t2$ belongs to a specific type $A$ -usually defined by the user of the tactic-, and the only thing we know about this type is that it satisfies the axioms of the concerned structure -the group for example-.
\begin{lstlisting}[caption=An instance of the type class Group, captionpos=b, label=lstAGroup:haskell2]
-- A is a group
instance Group A where
	Zero = ... 
	plus x y = ... 
	neg x = ...
	-- proof of associativity for +
	-- proof that zero is a neutral for +
	-- proof that every element has an inverse
\end{lstlisting}

The decision procedure should works for any type satisfying the axioms of a group, i.e, for any type $T$ given with an instance of Group for $T$. Therefore, we can imagine the decidable equality to have the following type :
\begin{lstlisting}[captionpos=b, label=groupDecideEq:haskell2, breaklines=true]
groupDecideEq : {T:Type} -> (Group T) -> (x : T) -> (y : T) -> Maybe (x = y)
\end{lstlisting}

As we said before, this decision procedure will use a function of normalization of terms in a group, in order to reduce $x$ and $y$ before comparing them with the syntactical equality. If we continue in the same direction, this function will have the following type :
\begin{lstlisting}[captionpos=b, label=groupReduce:haskell2, breaklines=true]
groupReduce : {T:Type} -> (Group T) -> T -> T
\end{lstlisting}

Of course, when trying to solve an equality $t1 = t2$, the terms $t1$ and $t2$ can potentially contain variables (declared in the context), because we want to solve equalities with literals. For example, when the user wants to prove $ \forall x y z:A, (x+y) + z = (z + y) + x $, the equality appears in a context were $x$, $y$ and $z$ have been previously added.

One main problem of the reduction procedure $groupReduce$, is that we would like to pattern match the term of type $T$ against the different possible forms that a term can have. In our example, in a group, a term can have the form of a constant, of a variable, or of an addition between two terms in the group. And of course, we would like to recognize these different forms, and to have a treatment depending on how the term in constructed. However, this is impossible to pattern match a term of this type $T$, since we don't know anything about its internal structure, except that it satisfies the axiom of a group.
The biggest part of the problem is that it is impossible to know if a term of type $T$ is a variable or not, since this is a meta information : when working in a specific context where $x$, $y$ and $z$ of type T are declared, the expression $(x+y)+z$ is an ordinary expression of type $T$, and nothing tells us that it contains variables.

Therefore, this naive approach of trying to work directly with the effective terms doesn't work when we try to write the tactic in the language itself. We will need to reflect the terms in an abstract syntax, which encodes the different constructors that we expect for a specific algebraic structure. With these reflected terms, it will become possible to pattern match them, and then to reduce them properly.

\subsection{a set of correct by construction tactics}

As we said before, because we are implementing the tactics in Idris, it is possible to reason about them in the language. That would not have been possible if the tactics were written in Haskell, since Haskell is a usual programming language without a logical side -mainly because Haskell is not dependently typed-.

The usual approach of formal certification is to implement the method -here the reduction for every algebraic structure-, and later to prove the correctness of the implementation. This usual way is naturally good for dealing with a lot of properties to ensure, one proved after the other, and all are proved after the definition of the function. The correct by construction approach is not so good for dealing with lot of properties to ensure, but one of its merits is that the writing of the program and the proof is simultaneously -forcing the programmer to think about correctness while programming-, and every bit of code has its "proof part", making therefore a strong link between the function and its proof of correctness. If the code of the function changes, that will be much easier to change the corresponding bit in the proof. The usual approach usually fails for this purpose, since a script of proof done afterwards in a proof mode is not easy to maintain, mainly because these kind of scripts are not easy to read.

Also, working with the correct by "construction approach" means that we are working with strong specification, instead of having a type that just states about data types. Because the type of the function expresses now logical properties, we can avoid the desperate situation where someone would try to prove the correctness of an invalid function. With a strong specification, your program will not compile until it fulfil its type, which means here until the program is correct.

### 2015-09-30 NotesOnCorrectnessAndCompleteness/notesOnCorrectnessAndCompleteness.tex
added various files lying around

% Franck's notes on correctness and completeness !
% last revision : 30 Sept 2015
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage[literate]{idrislang} 
%\usepackage{multicol}

%
\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%

\mainmatter              % start of the contributions
%
\title{Notes on correctness and completeness of the approach we've followed and comparison with Coq} 
%
\titlerunning{Notes on correctness and completeness}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Franck Slama}
%
\authorrunning{Franck Slama} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%%%\tocauthor{}
%
\institute{University of St Andrews, St Andrews, United Kingdom\\
\email{fs39@st-andrews.ac.uk}
%,\\ WWW home page:
%\texttt{http://users/\homedir iekeland/web/welcome.html}
}

\maketitle              % typeset the title of the contribution

Our correctness lemma :
$\forall x y,\ norm (T(x)) = T(y) \rightarrow x = y$

If the normalisation of the reflection of x gives the reflection of y, then x=y

\end{document}

### 2015-10-01 NotesOnCorrectnessAndCompleteness/notesOnCorrectnessAndCompleteness.tex
few notes added on the correctness of the approach

@@
 
-Our correctness lemma :
-$\forall x y,\ norm (T(x)) = T(y) \rightarrow x = y$
+\section{Correctness}
+
+We note $C$ the concrete type on which we aim to prove things like $a=b$ (ie, both $a$ and $b$ have type $C$).
+\\
+\\
+There is a type for reflected terms of type $C$, noted $ExprC$.
+There could be a specialised type for normalised reflected elements $ExprNC$, but in our implementation, we still use $ExprC$ for that, which means that we've got $ExprNC = ExprC$ (this is not the case in Coq's implementation).
+\\
+\\
+Now, something important and new in our approach, is that $ExprC$ is indexed over $C$ (i.e., $ExprC$ has type $C\ \rightarrow\ Type$).
+This index represents the concrete term (of type $C$) which is encoded by this encoding.
+For each element $e$ of type $ExprC\ c$, we will note $e_c$ to explicitly show the index of this term.
+\\
+\\
+In the formalism presented in [1], which describes Coq's implementation of the Ring tactic, they use a function $\phi_{C} : ExprC\ \rightarrow\ C$ to get back the concrete element.
+They've got the same thing for the type of normalised polynomials, with the function $\phi_{NC} : ExprNC\ \rightarrow\ C$. 
+In our formalism, we don't need such functions, but we could write them.
+This function would simply return the index.
+If we don't want to cheat by simply returning the index (which is not good in term of erasure, even if it doesn't really matter here), we can easily recompute this index, and show that $\forall (c:C) (e:ExprC\ c), \phi_C (e) = c$ which really means that this function has in fact just returned the index.
+\\
+\\
+In their formalism, they also use a metaification function, also called the function of automatic reflection.
+They define it in LTac, and we define it with  by using $\%syntax$ in Idris, which allows to pattern match on syntax rather that on constructors.
+The idea of this function is simply to compute the reflected term (of type ExprC c) representing a concrete value $c$ of type $C$.
+In their formalism, this function isn't typed.
+Is our formalism, where the abstract values are index over the concrete one, and thanks to the fact that $\%syntax$ allows to produce typed terms, we define it with this type : $\tau : (c:C)\ \rightarrow\ ExprC\ c$.
+In this sense, we can talk about type safe reflection : the term produced by the automatic reflection is guarantee to be a faithful representation of the concrete term we've given in input.
+\\
+\\
+Now, what we want to do is to try to prove $a=b$.
+For that, we will compute the encoding of $a$ and $b$, namely, $\tau\ a$ and $\tau\ b$.
+The first thing to notice here is that $\tau\ a$ has for index $a$, as described by the type of the function $\tau$.
+We can thus write $(\tau\ a)_a$ in order to emphasize on the fact that $\tau\ a$ has for index $a$.
+We've got the same for $(\tau\ b)$ that we will write $(\tau\ b)_b$.
+\\
+\\
+Now, we're going to run the normalisation algorithm on both $(\tau\ a)_a$ and $(\tau\ b)_b$. The normalisation procedure has this profile: \\
+$norm : \forall\ c,\ ExprC\ c\ \rightarrow\ (c'\ **\ (Expr\ c',\ c=c'))$. \\
+That means that running the normalisation algorithm on $(\tau\ a)_a$ will produce a dependent pair composed of an element $(a':C)$ and a term of type $ExprC \ a'$, that we write $(norm((\tau\ a)_a))_{a'}$, which is a term indexed over the new index $a'$.
+This function also produces a proof of $a=a'$.
+\\
+\\
+The same thing happens when we run the normalisation algorithm on $(\tau\ b)_b$ : that produces a value $(b':C)$, a term $(norm((\tau\ b)_b))_{b'}$ which has for index the $b'$ which has just been produced, and a proof that $b=b'$.
+\\
+\\
+Now, we're going to compare syntactically $(norm((\tau\ a)_a))_{a'}$ and $(norm((\tau\ b)_b))_{b'}$.
+What is really important is that if they're syntactically the same, then they're forced to be indexed over the same index : if an element of $ExprC\ a'$ equals (syntactically!) en element of $ExprC\ b'$, then this implies $a'=b'$\footnote{In fact, in our implementation, we don't even bother and this function of syntactical comparison directly produces the proof of equality of the indices if the two reflected terms are syntactically the same : it has for type $(ExprC\ a') \rightarrow (ExprC\ b')\ \rightarrow\ Maybe(a'=b')$ and not $(e1:ExprC\ a') \rightarrow (e2:ExprC\ b')\ \rightarrow\ Maybe(e1=e2)$.
+This is simply because we're not interested in the proof of equality between the reflected terms, so instead of producing it, and having an external lemma $extLemma:forall (a'\ b':C)\ (e1:ExprC\ a')\ (e2:ExprC\ b'),\ (e1=e2)\ \rightarrow\ (a'=b')$, which says that if two reflected terms are equal, then their index are also equal, we simply just produces the proof we need.}.
+\\
+\\
+So, if they're syntactically exactly equivalent, then we've also got $a'=b'$.
+But since we already had $a=a'$ and $b=b'$, we can compose these three proofs to produce the produce a proof of the desired property $a=b$.
+\\
+In fact, everything has been made possible thanks to the following property, which doesn't even need to be expressed because it is already in the type of the function $norm$ :
+$\forall (x\ y\ :\ C),\ norm ((T(x))_x) = Y_y \rightarrow x = y$
+
+In their approach, because they don't follow this kind of "type-safe reflection", they do need a correctness lemma, which says : \\
+$\forall (e1:ExprC)(e2:ExprC),\ \phi_{NC}\ (norm\ e1)\ =\ \phi_C\ e1$.
+So, what will happen is that they will compute (norm\ e1) and (norm\ e2).
 

### 2015-10-27 paper-JFP-DTP/paper.tex
paper

% paper.tex

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfp1}

\usepackage{float}

%%% Macros for the guide only %%%
\providecommand\AMSLaTeX{AMS\,\LaTeX}
\newcommand\eg{\emph{e.g.}\ }
\newcommand\etc{\emph{etc.}}
\newcommand\bcmdtab{\noindent\bgroup\tabcolsep=0pt%
  \begin{tabular}{@{}p{10pc}@{}p{20pc}@{}}}
\newcommand\ecmdtab{\end{tabular}\egroup}
\newcommand\rch[1]{$\longrightarrow\rlap{$#1$}$\hspace{1em}}
\newcommand\lra{\ensuremath{\quad\longrightarrow\quad}}

\title[Automatically proving equivalence by type-safe reflection]
      {Automatically proving equivalence by type-safe reflection}

 \author[Franck Slama and Edwin Brady]
        {FRANCK SLAMA and EDWIN BRADY\\
         University of St Andrews, United Kingdom\\
         \email{fs39@st-andrews.ac.uk, ecb10@st-andrews.ac.uk}}

\jdate{December 2015}
\pubyear{2016}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\doi{S0956796801004857}

\newtheorem{lemma}{Lemma}[section]

\begin{document}

\label{firstpage}

\maketitle

\begin{abstract}
% The abstract should be at most 300 words for the Journal of Functional Programming
Idris is a general purpose purely functional programming language with dependent types, aiming to bring type-based program verification techniques to functional programmers. One common difficulty with programming with dependent types is that proof obligations arise naturally once programs become even moderately sized. For example, implementing an adder for binary numbers indexed over their natural number equivalents will naturally lead to proof obligations for equalities of expressions over natural numbers. 
As far as possible, we would like to solve such proof obligations automatically. In this paper, we show one way to automate such proofs by reflection. We will show how representing Idris expressions in a reflected form, indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.
With this type-safe reflection technique, the resulting expressions are guaranteed to be faithful representations of the corresponding inputs and any generated proof is guaranteed to be a proof of the required property. 
We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified lists and the associative concatenation operation, and later show how this idea is generalised to various kind of properties (or axioms) that might be available (for example : commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for Monoids, Commutative Monoids, Groups, Commutative Groups, Rings, and so on, written in Idris, for proving different kind of equivalence. We will also show how each tactic reuses the other ones from the simplest structures, thus avoiding as much as possible the duplication of code.
\end{abstract}

\tableofcontents

\section{Introduction}

Proving that one term is equal to another one within some theories is something quite common in formal certification. Of course, the two things are usually not syntactically equal --this case would be trivial--, but they are equal according to a set of properties. These properties are either obtained by implementation (the implementation of $+$ for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.

If we consider a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care bout for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
The equality symbol $=$ here means "propositionally equal", but it could also be some kind of equivalence relation on $SEM$. Even in the case of the propositional equality, proving this kind of lemma is not trivial, because of the nature of the propositional equality in dependent types theories.

\subsection{The propositional equality}
In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
Since in type theory, propositions are seen as types, the proposition that two elements $a$ and $b$ are equal must correspond to a type. Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities. This is done by HMeq which follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$". If this type is inhabited, then $a$ is said to be propositionally equal to $b$.
Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A : $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$. Often, for convenience, we write $a\ =_A\ b$ instead of $Id_A\ a\ b$, but that denotes the same type.

Of course, the judgemental equality was needed to be included in the propositional equal. That is to say, if two terms are definitionally equal, then they should be provably equal.
Definitional equality basically contains unfolding of functions and (alpha, beta, iota...) reductions, until we can't reduce any more.
For example, if $f\ :\ \mathbb{N}\ \rightarrow\ \mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is definitionally equal to $7$.
Definitional equality is noted by $\equiv$.

The judgemental equality is something which can't be negated or assumed : we can't talk about this primitive equality inside the theory. Whether or not two expressions are equal by definition is a just a matter of expanding out the definitions.

In these type theories, the judgemental equality is captured by the propositional equality simply be giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
We can see it in the very common definition of Id :
\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
Id (A:Type) : A -> A -> Type :=
  Refl : forall a:A, Id a a
\end{verbatim}
\end{center}
%\caption{The type Id}
\figrule
\end{figure}

The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be $b$ (by definition!). In this case, the constructor $Refl$ helps to create a proof of this equality : $Refl\ a$ is precisely the proof which says that $a=_Aa$.

When reading these definitions, one could incorrectly think that the propositional equality is a simple wrapper of the judgemental equality. This is not the case : the propositional equality doesn't only contains the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type. This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the other constructors, then this proposition will hold for any term of this type.

Because this proposition can be any proposition, it can also be the propositional equality. That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before, and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions. The activity of proving equality therefore becomes in these theories a human activity, and is not decidable automatically in the general case by the type-checker.

\subsection{Using dependent types increases the

\subsection{A motivating little example}

\begin{thebibliography}{}
 \bibitem[\protect\citename{Augustsson and Johnsson, }1987]{AJ187}
   Augustsson,~L. and Johnsson,~T. (1987) LML users' manual. PMG
   Report, Department of Computer Science, Chalmers University of
   Technology, Goteborg, Sweden.
 \bibitem[\protect\citename{Butcher, }1981]{Butcher}
   Butcher,~J. (1981) Copy-editing: the Cambridge handbook.
   Cambridge University Press.
 \bibitem[\protect\citename{Chicago, }1982]{Chicago}
   The Chicago manual of style. University of Chicago Press.
 \bibitem[\protect\citename{Conklin, }1987]{JC87}
   Conklin,~J. (1987) Hypertext: an introduction and survey.
   \emph{IEEE Computer}, 20~(9): pp.~17--41.
 \bibitem[\protect\citename{Dijkstra, }1976]{EWD76}
   Dijkstra,~E.~W. (1976) \emph{A Discipline of Programming}.
   Prentice-Hall.
 \bibitem[\protect\citename{Knuth, }1984]{DEK84}
   Knuth,~D.~E. (1984) Literate programming. \emph{BCS Comput. J.}
   27~(2): 97--111 (May).
 \bibitem[\protect\citename{Lamport, }1986]{LaTeX}
   Lamport,~L. (1986) \LaTeX: a document preparation system
   (2nd edition). Addison-Wesley, New York.
 \bibitem[\protect\citename{Reynolds, }1969]{JCR69}
   Reynolds,~J.~C. (1969) Transformation systems and the
   algebraic structure of atomic formulas. In B. Meltzer and
   D. Michie (editors), \emph{Machine Intelligence 5}, pp.~135--151.
   Edinburgh University Press.
 \bibitem[\protect\citename{Toyn \emph{et al.}, }1987]{TDR87}
   Toyn,~I., Dix,~A. and Runciman,~C. (1987) Performance
   polymorphism. In \emph{Functional Programming Languages and
   Computer Architecture, Lecture Notes in Computer Science, 274},
   pp.~325--346. Springer-Verlag.
\end{thebibliography}

\label{lastpage}

\end{document}

% end of JFP2egui.tex

### 2015-10-27 paper-JFP-DTP/paper.tex
paper continued

@@
 
-\tableofcontents
+%\tableofcontents
 
-\section{Introduction}
-
-Proving that one term is equal to another one within some theories is something quite common in formal certification.
-Of course, the two things are usually not syntactically equal --this case would be trivial--, but they are equal according to a set of properties.
-These properties are either obtained by implementation (the implementation of $+$ for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.
-
-If we consider a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care bout for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
-The equality symbol $=$ here means "propositionally equal", but it could also be some kind of equivalence relation on $SEM$.
-Even in the case of the propositional equality, proving this kind of lemma is not trivial, because of the nature of the propositional equality in dependent types theories.
-
-\subsection{The propositional equality}
-In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
-Since in type theory, propositions are seen as types, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
-Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities.
-This is done by HMeq which follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
-If this type is inhabited, then $a$ is said to be propositionally equal to $b$.
-Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A : $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$.
-Often, for convenience, we write $a\ =_A\ b$ instead of $Id_A\ a\ b$, but that denotes the same type.
-
-Of course, the judgemental equality was needed to be included in the propositional equal. That is to say, if two terms are definitionally equal, then they should be provably equal.
-Definitional equality basically contains unfolding of functions and (alpha, beta, iota...) reductions, until we can't reduce any more.
-For example, if $f\ :\ \mathbb{N}\ \rightarrow\ \mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is definitionally equal to $7$.
-Definitional equality is noted by $\equiv$.
-
-The judgemental equality is something which can't be negated or assumed : we can't talk about this primitive equality inside the theory.
-Whether or not two expressions are equal by definition is a just a matter of expanding out the definitions.
-
-In these type theories, the judgemental equality is captured by the propositional equality simply be giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
-We can see it in the very common definition of Id :
-\begin{figure}[H]
-\figrule
-\begin{center}
-\begin{verbatim}
-Id (A:Type) : A -> A -> Type :=
-  Refl : forall a:A, Id a a
-\end{verbatim}
-\end{center}
-%\caption{The type Id}
-\figrule
-\end{figure}
-
-The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be $b$ (by definition!).
-In this case, the constructor $Refl$ helps to create a proof of this equality : $Refl\ a$ is precisely the proof which says that $a=_Aa$.
-
-When reading these definitions, one could incorrectly think that the propositional equality is a simple wrapper of the judgemental equality.
-This is not the case : the propositional equality doesn't only contains the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the other constructors, then this proposition will hold for any term of this type.
-
-Because this proposition can be any proposition, it can also be the propositional equality.
-That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before, and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
-The activity of proving equality therefore becomes in these theories a human activity, and is not decidable automatically in the general case by the type-checker.
-
-\subsection{Using dependent types increases the
-
-\subsection{A motivating little example}
+\input{./content/introduction.tex}
 

### 2015-10-27 paper-JFP-DTP/content/introduction.tex
introduction file added

--- 2014-12-13 ICFP15paper/content/introduction.tex
@@
 
-Proving that one term is equal to another within some theories is something quite common in formal certification.
-Of course, the two things are usually not syntactically equal --this case would be trivial--, but they are equal according to a set of properties.
-These properties are either obtained by implementation (the implementation of + for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.
+Proving that one term is equal to another one within some theories is something quite common in formal certification.
+The two terms are usually not syntactically equal --this case would be trivial--, but they are equal according to some properties.
+These properties are either obtained by implementation (the implementation of $+$ for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.
+
+If we consider a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care bout for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
+The equality symbol $=$ here is the propositionally equality, but it could also be some kind of equivalence relation on $SEM$.
+Even in the case of the propositional equality, proving this kind of lemma is not trivial, because of the nature of the propositional equality in Intentional Type Theories.
+
+\subsection{The propositional equality and the jugmental equality in Intentional Type Theories}
+In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
+Since in type theory, propositions are seen as types, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
+Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities.
+This is done by HMeq which follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
+If this type is inhabited, then $a$ is said to be provably equal to $b$.
+Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A : $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$.
+Often, for convenience, $Id_A\ a\ b$ is written $a\ =_A\ b$.
+This equality is the equality which can be manipulated in the language.
+There is another --more primitive-- notion of equality in Intentional Type Theories, called judgemental (or definitional equality).
+This second equality means "equal by definition".
+The judgemental equality is something which can't be negated or assumed : we can't talk about this primitive equality inside the theory.
+Whether or not two expressions are equal by definition is a just a matter of expanding out the definitions.
+For example, if $f\ :\ \mathbb{N}\ \rightarrow\ \mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is definitionally equal to $7$.
+Definitional equality basically contains unfolding of functions and (beta, iota...) reductions, until no more reduction can be performed.
+Definitional equality is noted by $\equiv$.
+
+The judgemental equality was needed to be included in the propositional equal, because what is equal by definition should be provably equal.
+This is accomplished by giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+Id (A:Type) : A -> A -> Type :=
+  Refl : forall a:A, Id a a
+\end{verbatim}
+\end{center}
+%\caption{The type Id}
+\figrule
+\end{figure}
+
+The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be $b$ (by definition!).
+In this case, the constructor $Refl$ helps to create a proof of this equality : $Refl\ a$ is precisely the proof which says that $a=_Aa$.
+
+When reading these definitions, one could incorrectly think that the propositional equality is a simple wrapper of the judgemental equality.
+This is not the case : the propositional equality doesn't only contains the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
+This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the other constructors, then this proposition will hold for any term of this type.
+
+Because this proposition can be any proposition, it can also be the propositional equality.
+That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before, and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
+The activity of proving equality therefore becomes in these theories a human activity, and is not decidable automatically in the general case by the type-checker.
+
+For example, we can prove that $n+0 = n$ by induction, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
+
 These proofs are usually omitted "on the paper" for the ordinary every day maths, but they are required by proof assistants, no matter how obvious they are.
-Currently, in the Idris\footnote{Idris is a relatively new, purely functional, and dependently typed programming language with full dependent types, which also supports tactic based proofs  (Idris will be briefly defined the the [Section X] (at the end of the paper))} programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
+Currently, in the Idris\footnote{Idris is a relatively new, purely functional, and dependently typed programming language with full dependent types} programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
 These proofs are the every day routine when we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
@@
 Let's define the types of $Bit$ and $Binary$ number, both indexed over the value they represent, expressed in base ten as a natural number.
-\begin{code}[caption=Bit and binary number, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 data Bit : Nat -> Type where
@@
            -> Binary (S w) (bit + 2 * v)
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Bit and binary number}
+\figrule
+\end{figure}
 
@@
 
-It will be useful to be able to pad a binary number expressed with $w$ bits to a binary number with $S\ w$ digits : 
-\begin{code}[caption=Padding, captionpos=b, label=lst1:haskell2]
-pad : Binary w n -> Binary (S w) n
-pad zero = zero # b0
-pad (num # x) = pad num # x
-\end{code}
-Of course, this operation preserves the value represented by the binary number : this function takes a $Binary\ w\ n$ and produces a $Binary\ (S\ w)\ n$, so there will be no need to prove the correctness of this function after its definition since the $value$ is still $n$.
-
 Now, we're tempted to write the addition between two binary numbers.
@@
 
-\begin{code}[caption=Addition of three bits, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 addBit : Bit x -> Bit y -> Bit c ->
@@
 addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, _)))
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Addition of three bits}
+\figrule
+\end{figure}
 
@@
 Thus, we're tempted to write :
-\begin{code}[caption=Addition of two binary numbers, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 adc : Binary w x -> Binary w y -> Bit c 
@@
           adc numx numy carry0 # lsb
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Addition of two binary numbers}
+\figrule
+\end{figure}
+
 Unfortunately, the types are mismatching for the two cases expressed by this pattern matching.
@@
 For example, the proof for the second pattern look like this :
-\begin{code}[caption=Proof by hand of the correspondence between expected and provided types for the second pattern of the definition of adc, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 adc_lemma_2 = proof {
@@
     trivial;
-}
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Proof by hand of the correspondence between expected and provided types for the second pattern of the definition of adc}
+\figrule
+\end{figure}
 
@@
 Thus, we're rewriting a term by using the properties of a commutative group.
-If we've already proven that natural numbers follow the structure of commutative group, this is the code that we would like to be able to write :
 
-\begin{code}[caption=Code we'd like to write for the addition of two binary numbers, captionpos=b, label=lst1:haskell2]
-adc : Binary w x -> Binary w y -> Bit c 
-     -> Binary (S w) (c + x + y)
-adc zero zero carry ?= zero # carry 
-        by CommutativeGroupSolver
-adc (numx # bX) (numy # bY) carry
-   ?= let (v1 ** (v2 ** (carry0, lsb, _))) = 
-      addBit bX bY carry in
-          adc numx numy carry0 # lsb 
-        by CommutativeGroupSolver
-\end{code}
 We would like to let a solver for commutative groups to find the proof on its own.

### 2015-10-31 paper-JFP-DTP/content/introduction.tex
paper...

@@
 
-If we consider a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care bout for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
-The equality symbol $=$ here is the propositionally equality, but it could also be some kind of equivalence relation on $SEM$.
-Even in the case of the propositional equality, proving this kind of lemma is not trivial, because of the nature of the propositional equality in Intentional Type Theories.
+Example of these goals are $\forall\ xy:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$ for natural numbers with the standard addition and multiplication, or $\forall\ l1\ l2\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\ (l3++l4)$ for lists and the standard concatenation operation.
+Proving this kind of auxiliary lemmas is time consuming, is not general enough for being reusable, and isn't the interesting part of formal certification.
+Therefore, we would like to be able to automatize a far as possible the constructions of proofs for thins kind of lemmas for the Idris language, a dependently typed language which doesn't offer many proof automations at the moment.
+
+A common situation in formal certification is the following.
+We have a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care about for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
+The equality symbol $=$ is the propositional equality, as it was the case in the two previous examples with natural numbers and list.
+We remind the readers in the next subsection how the jugmental equality is defined in Idris, and more generally in Intensional Type Theories.
 
@@
 This equality is the equality which can be manipulated in the language.
-There is another --more primitive-- notion of equality in Intentional Type Theories, called judgemental (or definitional equality).
+There is another --more primitive-- notion of equality in Intentional Type Theories, called judgemental equality, or definitional equality.
 This second equality means "equal by definition".
@@
 \end{center}
-%\caption{The type Id}
+\caption{The type Id}
 \figrule
@@
 
-The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be $b$ (by definition!).
+The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be (by definition!) $b$.
 In this case, the constructor $Refl$ helps to create a proof of this equality : $Refl\ a$ is precisely the proof which says that $a=_Aa$.
@@
 Because this proposition can be any proposition, it can also be the propositional equality.
-That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before, and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
-The activity of proving equality therefore becomes in these theories a human activity, and is not decidable automatically in the general case by the type-checker.
+That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before (in the definition 1), and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
+The activity of proving equalities is therefore in these theories something which isn't automatically decidable by the type-checker in the general case.
 
@@
 
-In fact, this function produces a dependent pair.
-The first component is the value $bX$ represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
+This function produces a dependent pair.
+The first component is the value $bX$ (expressed as a natural number) represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
 This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing finally the two bits produced and the proof that these two bits are effectively encoding the result of the addition we had to perform.
@@
 
-Again, we note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : that the two bits produced effectively represent the addition of the three bits given in input.
+We note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : the two bits produced effectively represent the addition of the three bits given in input.
 
-Now, we want to define a function adding two binary numbers and a bit (used as a carry).
+Now, we want define a function adding two binary numbers and a bit (used as a carry).
 We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit, and this result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
@@
 Using a provisional definitions here will make the definition being accepted, but we will have to prove afterwards that in both patterns, the term provided can be converted into a term which has the right type.
-However, for both of these two patterns, it is possible to prove the equality between the expected and the provided types, and this is even stronger than we need.
+However, for both of these two patterns, it is possible to prove the equality between the expected and the provided types, which is even more than we need.
 These proofs will be made of a series of rewriting, and each rewriting will use a property about the operation $+$ on natural numbers.
@@
 
-Thus, this kind of proofs consists of a potentially long sequence of rewriting, every rewriting step using one property of the theory.
+Thus, this kind of proofs consists of a potentially long sequence of rewriting, where every rewriting step uses one of the available properties.
 Without some specific automation, this sequence of rewriting is done by the user of the proof assistant.
@@
 
-Algebraic solvers, like ring solvers, are already implemented for various proof assistants, including Coq and Agda.
-In this paper, we describe a certified implementation of an automatic solver for equalities on algebraic structures, for the Idris language.
-The interesting thing that we're going to present in this paper lies in the way we construct it, which is a "correct by construction" style, with the same spirit as the functions we've written above.
-
-In this paper, we focus on some specific theories, which are the following algebraic structures : magmas, semi-group, monoid, group and Abelian group.
-
-Each tactic will be implemented by reflection, and the reason for this will be discussed in section [Blabla].
-Working by reflection for implementing tactics has been done several times, as in [paper Coq ring] for the implementation of a ring solver for the Coq proof assistant.
+Algebraic provers, like ring provers, are already implemented for various proof assistants, including Coq and Agda.
+In this paper, we describe a certified implementation of an automatic prover for equalities in various algebraic structures, including Monoid, Groups and Rings (potentially commutative), for the Idris language.
+The novelty is in the approach that we follow, using a new kind of type-safe reflection.
+Working by reflection for implementing tactics has been done several times, as in [paper Coq ring] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
 We will compare our approach with other implementation of this kind of tactics in section [Blabla].
@@
 \begin{itemize}
-	\item 1) We follow a "correct by construction" approach for implementing the reduction procedures, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
-The approach will be presented in section 2 on a smaller problem, and on section 3 for the complete hierarchy of algebraic structures.
-Details of the normalization procedures are given in section 4.
-	\item 2) For achieving the first goal, we are extensively using dependent types in order to capture the interesting properties that matter for assuring the correctness of the method.
-We use an original representation of reflected terms which embeds the denoted --concrete-- value represented.
-The reflected terms become indexed over the concrete ones.
-These representations are presented in section 2 for the toy example, and in section 3 for the real problem.
+	\item 1) We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete ones, thus providing an easy way to pull out the proofs easily, and providing two guarantees.
+The first one is that the reflection of a term is guaranteed to be a faithful representation of the term.
+The second is that any generated proof is guaranteed to be a proof of the required property, thus increasing the totality of the prover.
+The basic ideas of the technique is presented in the section 2 on a smaller problem, and is represented for the real problem in the section 3 and 4.
+Discussion about the totality is provided in section 7.
+	\item 2) The normalisation procedures are implemented by following a correct by construction approach, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
+Again, this approach will be presented in section 2 on a smaller problem, and on section 4 for the complete hierarchy of algebraic structures.
 	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.

### 2015-11-01 paper-JFP-DTP/content/basicIdeas.tex
paper continued

--- 2014-07-07 newPaper/paper.tex
@@
-%-----------------------------------------------------------------------------
-%
-%			Edwin's and Franck's Paper !
-%  uses the template for sigplanconf LaTeX Class
-%
-%-----------------------------------------------------------------------------
+\section{A simplified problem : type safe reflection for proving equalities with universally quantified natural numbers and additions}
 
-\documentclass{sigplanconf}
-
-% The following \documentclass options may be useful:
-
-% preprint      Remove this option only once the paper is in final form.
-% 10pt          To set in 10-point type instead of 9-point.
-% 11pt          To set in 11-point type instead of 9-point.
-% authoryear    To obtain author/year citation style instead of numeric.
-
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{listings}
-\usepackage[literate]{idrislang} 
-\newcommand{\cL}{{\cal L}}
-
-%\lstset {captionpos=b}
-
-\begin{document}
-
-\special{papersize=8.5in,11in}
-\setlength{\pdfpageheight}{\paperheight}
-\setlength{\pdfpagewidth}{\paperwidth}
-
-\conferenceinfo{POPL ’15}{January 15-17, 2015, Mumbai, India.}
-\copyrightyear{2014} 
-\copyrightdata{978-1-nnnn-nnnn-n/yy/mm} 
-\preprintfooter{Our paper for POPL '15}
-\doi{nnnnnnn.nnnnnnn}
-
-% Uncomment one of the following two, if you are not going for the 
-% traditional copyright transfer agreement.
-
-%\exclusivelicense                % ACM gets exclusive license to publish, 
-                                  % you retain copyright
-
-%\permissiontopublish             % ACM gets nonexclusive license to publish
-                                  % (paid open-access papers, 
-                                  % short abstracts)
-
-\titlebanner{banner above paper title}        % These are ignored unless
-\preprintfooter{short description of paper}   % 'preprint' option specified.
-
-\title{A correct by construction hierarchy of reflective tactics for proving equalities in algebraic structures}
-%\subtitle{An example for a collection of tactics solving equalities in algebraic structures}
-
-%\authorinfo{Edwin Brady\and Franck Slama}
-%           {University of St Andrews, United-Kingdom}
-%           {ecb10@st-andrews.ac.uk \and fs39@st-andrews.ac.uk}
-\authorinfo{Author's name omitted for submission}
-           {University of ..., Country}
-           {email address omitted for submission}
-
-\bibliographystyle{abbrvnat}
-
-\maketitle
-
-\nocite{*}
-
-\input{./content/abstract.tex}
-
-\input{./content/introduction.tex}
-
-\section {A smaller problem : a correct by construction tactic for the associativity of lists concatenation}
-
-In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universally quantified natural numbers, the the properties of associativity and neutral element for the operation plus.
-For example, we would like to be able to automatically generate proofs of facts like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
+In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universally quantified natural numbers, the properties of associativity and neutral element for the operation plus.
+For example, we would like to be able to automatically generate proofs of goals like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
 For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a neutral element on the right side of the plus : $\forall x, x + Z = x$.
+Zero is also a left neutral element, but this property is not needed, as we have this behavirous by reduction as $+$ is defined recursively on its first argument.
 Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to associativity or thanks to the neutrality of Z".
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to the associative and neutral properties".
 \\
@@
 The goal of the normalization is of course to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
-That's why it will be possible to decide the equality by simply comparing the two normal forms with a week syntactical equality.
+For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normalisaiton form will be the complete left associative form.
+It will then be possible to decide the equality by simply comparing the normalised left and right hand sides with a simple syntactical equality.
 \\
 \\
-Of course, the normalization will have to fulfil the following correctness property :
-$\forall x, norm\ x = x$ [1], which means that after normalization we obtain a number which is provably equal to the input number (potentially using our properties multiple times). \\
-And after normalization, we will compare the two resulting terms with a standard syntactical equality, and if they are equal, we will get the desired proof.  \\
-Indeed :
-$norm\ x = norm\ y$ will imply $x=y$ by using the correctness lemma [1] of $norm$ two times. \\
-\\
-In fact, this is not exactly how it will work.
-This way would not be easily feasible, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
+In fact, such a normalisation function can't be written, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
 And the normalization function needs to do different treatments for a "variable natural number" (a number which has been universally quantified) and for the constant $Z$.
-This is not possible yet, because once the variables are abstracted, nothing tells us where the variables are in the left and right hand side of the equality : this is a meta information of the ASTs representing the two terms that we can't access.
+This is not possible yet, because once the variables are abstracted, they are just normal values of type $Nat$, and nothing tells us where they are in the left and right hand side of the equality : this information only exists at the level of the ASTs representing the two terms.
 For this reason, we will work by reflection, as we will do for the hierarchy of tactics in the next section.
-In this smaller example, it means that means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
+In this smaller example, it means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
 This datatype will allow us to know the internal structure of a number, ie, where the variables and constants are.
 Indeed, this datatype will allow pattern matching.
-Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't interesting for our goal.
-With the first approximation of the datatype $Expr$ presented in the Listing 1, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
+Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't what we needed.
+With the first approximation of the datatype $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
 
-\begin{code}[caption=First version of reflected numbers, captionpos=b, label=lst1:haskell2]
-data Expr : (G : Vect n Nat) -> Type where
-     Plus : {n:Nat} -> {G:Vect n Nat} -> 
-            Expr G -> Expr G -> Expr G
-     Var  : {n:Nat} -> {G:Vect n Nat} -> 
-            (i : Fin n) -> Expr G
-     Zero : {n:Nat} -> (G:Vect n Nat) -> 
-            Expr G Z
-\end{code}
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+data Expr : (gam : Vect n Nat) -> Type where
+     Plus : {n:Nat} -> {gam:Vect n Nat} -> 
+            Expr gam -> Expr gam -> Expr gam
+     Var  : {n:Nat} -> {gam:Vect n Nat} -> 
+            (i : Fin n) -> Expr gam
+     Zero : {n:Nat} -> (gam:Vect n Nat) -> 
+            Expr gam Z
+\end{verbatim}
+\end{center}
+\caption{First version of reflected numbers}
+\figrule
+\end{figure}
 
-For this smaller problem, we want to prove equalities in which numbers are always universally quantified, so we only need to represent variables (universally quantified natural numbers), the constant Z, and additions.
 Variables are represent using a De Brujin-like index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
 
-The type Expr is indexed over a vector of numbers (we will usually name it $G$), which represents our context.
-For example, in the example 1, we will have to encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are declared.
+The type Expr is indexed over a vector of numbers $\Gamma$, which is the context of all universally quantified variables.
+In the example 1, we will encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are present.
 The first element of this context denotes the variable $x1$, the second denotes $x2$, and so on.
 Thus, the left hand side will be encoded by :
-\begin{code}[caption=Reflected LHS of example 1, captionpos=b, label=lst1:haskell2]
-App (App (Var G fZ) 
-         (Var G (fS fZ))) 
-    (App (Var G (fS (fS fZ))) 
-         (Var G (fS (fS (fS fZ)))))
-\end{code}
 
-We now need to define what will be the normalization procedure.
-If, as it is probably expected, this function simply takes an Expr and produces another Expr, then we will need to prove the following lemma afterwards, corresponding to the correctness property mentioned before : \\
-$\forall\ e:Expr\ G,\ interpretation\ (reduce\ e)\ =\ interpretation\ e$ \\
-where $interpretation$ is a function computing the interpretation of an Expr in a context $G$, that is to say, the number that this Expr is encoding.
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+Plus (Plus (Var gam fZ) 
+           (Var gam (fS fZ))) 
+     (Plus (Var gam (fS (fS fZ))) 
+           (Var gam (fS (fS (fS fZ)))))
+\end{verbatim}
+\end{center}
+\caption{Reflected LHS of example 1}
+\figrule
+\end{figure}
+
+\subsection{Type safe reflection}
+
+If we continue with this first definition of $Expr$, the normalisation function will certainly take an Expr and produce another Expr, and we will need to prove the following lemma afterwards : \\
+$\forall\ e:Expr\ \Gamma,\ reify\ (reduce\ e)\ =\ reify\ e$ \\
+where $reify$ is a function computing the interpretation of an Expr in a context $\Gamma$, that is to say, the number that this Expr is encoding.
 This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
-For this little example, the reduction will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with, it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+For this little example, the reduction procedure will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with and it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
 \\
 \\
-To avoid these two sources of complexity, we will : \\
-- Use dependent types in order to directly capture and embed the concrete number that an Expr is encoding. Thus, it won't be necessary to define the $interpretation$ function. \\
-- Write a "correct by construction" reduction procedure.
-The proof of preservation of the interpretation will be computed bit by bit at the same time as the function will produce the normalized expression (by rewriting the term by using the two properties).
+To avoid these two sources of complexity, we add an index to the type $Expr$ which is the concrete number that an Expr is encoding.
+Thus, it won't be necessary to define the $interpretation$ function.
+\\
 
-Thus, the type for reflected numbers becomes :
-\begin{code}[caption=Second version of reflected number with embedded denotation, captionpos=b, label=lst1:haskell2]
-using (x : Nat, y : Nat, G : Vect n Nat)
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+using (x : Nat, y : Nat, gam : Vect n Nat)
   data Expr : (Vect n Nat) -> Nat 
                -> Type where
-       Plus : Expr G x -> Expr G y -> 
-              Expr G (x + y)
+       Plus : Expr gam x -> Expr gam y -> 
+              Expr gam (x + y)
        Var  : (i : Fin n) -> 
-              Expr G (index i G)
-       Zero : Expr G Z
-\end{code}
-For an expression $e\ :\ Expr\ G\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $G$".
-When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ G)$.
+              Expr gam (index i gam)
+       Zero : Expr gam Z
+\end{verbatim}
+\end{center}
+\caption{Second version of reflected number with embedded denotation}
+\figrule
+\end{figure}
+
+For an expression $e\ :\ Expr\ \Gamma\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $\Gamma$".
+When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
 Also, the $Zero$ expression denotes the natural number $Z$.
@@
 
-The reduction procedure is now supposed to be written on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
-Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ G\ x$ is the encoding of the number $x$.
+We want to write the reduction function on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
+Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
 Thus, we can write the type of $reduce$ like this : \\
-$reduce\ :\ Expr\ G\ x\ \rightarrow\ (x'\ **\ (Expr\ G\ x',\ x\ =\ x'))$ \\
-The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ G\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
-Note that this function can't simply produce an $Expr\ G\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
+$reduce\ :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ =\ x'))$ \\
+The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
+Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
+Said differently, even if we can prove $x=x'$ (if the function is correctly defined), we do not have $x \equiv x'$.
 And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
-The reason is simple : when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
+The reason is that when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
 \\
-We will have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
-We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ G\ x'$, and a proof of $x=x'$.
-We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ G\ y'$, and a proof of $y=y'$.
+We have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
+We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
+We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
 \\
@@
 
-\begin{code}[caption=Syntactical equality between reflected terms, captionpos=b, label=lst1:haskell2]
-eqExpr : (e : Expr G xs) -> (e' : Expr G ys) 
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+eqExpr : (e : Expr gam x) -> (e' : Expr gam y) 
          -> Maybe (e = e')
@@
 eqExpr _ _ = Nothing
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Syntactical equality between reflected terms}
+\figrule
+\end{figure}
 
-Now, if the two normalised expressions $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
+Now, if the two normalised expressions $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
 By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$.
 
-\begin{code}[caption=Building the desired proof with the two proofs of equality, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 buildProof : {x : Nat} -> {y : Nat} 
-       -> Expr G x' -> Expr G y' 
+       -> Expr gam x' -> Expr gam y' 
        -> (x = x') -> (y = y') 
@@
   buildProof e e' lp rp | Nothing = Nothing
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Building the desired proof with the two proofs of equality}
+\figrule
+\end{figure}
 
@@
 
-\begin{code}[caption=buildProof metavariable, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
   bp1 = proof {
@@
 }  
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{buildProof metavariable}
+\figrule
+\end{figure}
+
 Finally, the main function which tries to prove the equality $x=y$ simply has to reduce the two metaified terms reflecting the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs we just obtained :
-\begin{code}[caption=testEq, captionpos=b, label=lst1:haskell2]
-  testEq : Expr G x -> Expr G y 
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+  testEq : Expr gam x -> Expr gam y 
            -> Maybe (x = y)
@@
         buildProof l' r' p1 p2 
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{testEq}
+\figrule
+\end{figure}
+
 Now, we need to define the function reduce.
-To do that, we have to fix a canonical representation of associative natural numbers.
+To do that, we have to decide a canonical representation of associative natural numbers.
 We decide that the left associative form will be the canonical representation.
-Thus, the $reduce$ function will have to rewrite the metaified term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
+Thus, the $reduce$ function has to rewrite the metaified term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
 + xn)$.
@@
 Thus it will be easier to be certain that we are effectively computing the normal form : forcing properties to hold by the shape of a datatype is a good usage of dependent types when, like here, it doesn't introduce more complications.
-\begin{code}[caption=Reflected left associative numbers, captionpos=b, label=lst1:haskell2]
-data LExpr : (G : Vect n Nat) -> Nat 
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+data LExpr : (gam : Vect n Nat) -> Nat 
              -> Type where
-     LPlus : LExpr G x -> (i : Fin n) 
-             -> LExpr G (x + index i G)
-     LZero : LExpr G Z
-\end{code}
+     LPlus : LExpr gam x -> (i : Fin n) 
+             -> LExpr gam (x + index i gam)
+     LZero : LExpr gam Z
+\end{verbatim}
+\end{center}
+\caption{Reflected left associative numbers}
+\figrule
+\end{figure}
+
 This datatype has only two constructors. In fact, it combines the previous $Var$ and $Plus$ constructors so that it becomes impossible to write an expression which isn't left associative.
 
-As part of the normalization, we will write a function $expr\_l$ which converts an $Expr\ G\ x$ to a $LExpr\ G\ x'$ and which produces a proof of $x=x'$.
+As part of the normalization, we write a function $expr\_l$ which converts an $Expr\ \Gamma\ x$ to a $LExpr\ \Gamma\ x'$ and which produces a proof of $x=x'$.
 This function will therefore use the two available properties multiple times --and especially the property of associativity--, in order to obtain the expected fully left associative form.
-Of course, because by using this new datatype $LExpr$ we've changed the representation of our encoded lists, we will need to convert back an $LExpr\ G\ x$ to an $Expr\ G\ x$.
-The function $l\_expr$ will do this easy task.
 
-\begin{code}[caption=Production of the left associative form, captionpos=b, label=lst1:haskell2]
-expr_l : Expr G x 
-         -> (x' ** (LExpr G x', x = x'))
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+expr_l : Expr gam x 
+         -> (x' ** (LExpr gam x', x = x'))
 expr_l Zero = (_ ** (LZero, refl))
@@
         (_ ** (rx, ?plusLExpr2))
+\end{verbatim}
+\end{center}
+\caption{Production of the left associative form}
+\figrule
+\end{figure}
 
-l_expr : LExpr G x -> Expr G x
+Using this new datatype $LExpr$ has changed the representation of our encoded lists, so we need to convert back an $LExpr\ \Gamma\ x$ to an $Expr\ \Gamma\ x$.
+The function $l\_expr$ does this easy task.
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+l_expr : LExpr gam x -> Expr gam x
 l_expr LZero = Zero
 l_expr (LPlus x i) = Plus (l_expr x) (Var i)
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Going back from LExpr to Expr}
+\figrule
+\end{figure}
 
@@
 We've got two metavariables to prove.
-The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ G$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ G)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
+The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ \Gamma)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
 Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
 
-\begin{code}[caption=Proof of the metavariable plusLExpr1, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 plusLExpr1 = proof {
@@
 }
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Proof of the metavariable plusLExpr1}
+\figrule
+\end{figure}
 
@@
 
-\begin{code}[caption=Proof of the metavariable plusLExpr2, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 plusLExpr2 = proof
@@
   rewrite (sym yprf)
-  mrefine a_plus_zero
-\end{code}
+  mrefine plusZeroRightNeutral
+\end{verbatim}
+\end{center}
+\caption{Proof of the metavariable plusLExpr2}
+\figrule
+\end{figure}
 
@@
 
-\begin{code}[caption=Reduction function, captionpos=b, label=lst1:haskell2]
-  reduce : Expr G x -> 
-           (x' ** (Expr G x', x = x'))
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+  reduce : Expr gam x -> 
+           (x' ** (Expr gam x', x = x'))
   reduce e = 
@@
          (x' ** (l_expr e', prf))
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Reduction function}
+\figrule
+\end{figure}
 
@@
 
-\begin{code}[caption=Two test expressions, captionpos=b, label=lst1:haskell2]
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 e1 : (x, y, z : Nat) 
@@
                       (Var (fS (fS fZ))))
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Two test expressions}
+\figrule
+\end{figure}
 
 The numbers denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this by using $testEq$.
-\begin{code}[caption=Test of equality betwen e1 and e2, captionpos=b, label=lst1:haskell2]
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 e1_e2_testEq : (x, y, z : Nat) 
@@
                             (e2 x y z)
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Test of equality betwen e1 and e2}
+\figrule
+\end{figure}
 
 And if we ask for the evaluation of this term, we should obtain $Just$ and a proof of equality between the two underlying numbers.
-\begin{code}[caption=Obtained proof, captionpos=b, label=lst1:haskell2]
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
 #\x => \y => \z => e1_e2_testEq x y z
@@
             = x+ (y + x) + z)
-\end{code}
+\end{verbatim}
+\end{center}
+\caption{Obtained proof}
+\figrule
+\end{figure}
 
@@
 As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightNeutral$).
-
-\section {Back to the general problem : A hierarchy of tactics (2/3 pages)}
-
-Now that we have introduced our ideas on this little example, it is time to apply them for our main goal, more general : the implementation of a hierarchy of tactics proving equalities in algebraic structures.
-Very often, the properties available on a given type are the one of a well known structure : magma, semi-group, monoid, group, commutative-group...
-We will no longer only work with the properties of associativity and of right neutral element for natural numbers as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.
-
-We will construct a hierarchy of typeclasses.
-All our tactics (the group solver, the monoid solver...) will be able to work on any type, being given that an instance of the corresponding typeclass is provided.
-
-\subsection {Hierarchy of typeclass}
-
-All our tactics will require to have a way of testing the equality between elements of the underlying set, that is to say, a way to test equality between constants.
-For this reason, we define a notion of "set", which only requires this equality.
-All the algebraic structures above will extend this typeclass :
-
-\begin{code}[caption=Set, captionpos=b, label=lst1:haskell2]
-class Set c where
-    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
-\end{code}
-
-Note that this equality is a "week" equality : it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different -- instead, it simply produces the value $Nothing$--.
-That's quite natural, since we want to generate proof of equality, and not to generate counter examples for proving dis-equalities, which is another problem.
-
-Of course, there will be no tactic associated to $Set$, since we have no operations and not properties associated to this structure.
-Therefore, the equality in a $Set$ is just the syntactical equality, and can simply be proven with $refl$.
-
-The first real structure, almost trivial, is the magma.
-A magma has just a $Plus$ operation which let us computing constants, and has no axioms about this operation.
-Thus, as we will see later, the normalization will be simply a computation between constants inside parenthesis.
-
-\begin{code}[caption=Magma, captionpos=b, label=lst1:haskell2]   
-class Set c => Magma c where
-    Plus : c -> c -> c
-\end{code}
-
-A bit more interesting is the Semi-Group typeclass. A semi-Group is a Magma (ie, it still has a $plus$ operation), but moreover it has the property of associativity for this operation.
-
-\begin{code}[caption=Semi-Group, captionpos=b, label=lst1:haskell2]   
-class Magma c => SemiGroup c where
-    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
-           -> (Plus (Plus c1 c2) c3 
-                = Plus c1 (Plus c2 c3))
-\end{code}
-
-The Monoid, for its part, is a Semi-Group with the property of neutral element.
-
-\begin{code}[caption=Monoid, captionpos=b, label=lst1:haskell2]   
-class SemiGroup c => Monoid c where
-    Zero : c    
-    Plus_neutral_1 : (c1:c) 
-            -> (Plus Zero c1 = c1)    
-    Plus_neutral_2 : (c1:c) 
-             -> (Plus c1 Zero = c1)
-\end{code}
-
-A Group is a Monoid, with two new operations.
-The binary operation $Minus$, and the unary operation $Neg$.
-One axiom says that the $Minus$ can always be simplified with the $Neg$ and the $Plus$.
-That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$.
-The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :
-
-\begin{code}[caption=Group, captionpos=b, label=lst1:haskell2]   
--- This is just a conjunctive predicate
-hasSymmetric : (c:Type) -> (p:Monoid c) 
-                -> c -> c -> Type
-hasSymmetric c p a b = 
-         (Plus a b = Zero, Plus b a = Zero)
-
-class Monoid c => Group c where
-	Minus : c -> c -> c
-	Neg : c -> c
-	Minus_simpl : (c1:c) -> (c2:c) -> 
-	             Minus c1 c2 = Plus c1 (Neg c2) 
-	Plus_inverse : (c1:c) -> 
-	             hasSymmetric c _ c1 (Neg c1)
-\end{code}
-
-This hierarchy can be extended without difficulty with Abelian groups (with the axiom of commutativity), Ring, Commutative Rings...
-
-The important thing to remember is that these typeclass will somehow be used as predicates on types.
-In order to call a solver on a type $T$, the user of the system will have to satisfy the corresponding typeclass by providing an instance of it for his type $T$, ie, he will have to prove that the corresponding properties effectively hold for $T$.
-
-Note that the properties required by these typeclass for a specific type $T$ will either be proven if the operations on $T$ are real computable definitions, or axiomatised if the user is working on an axiomatised theory where the operations ($Plus$, $Neg$) are just axioms.
-
-As discussed in the section 2.1 on the smaller problem, the algorithm of normalization will not directly use the concrete value of type $T$, but a reflected term, indexed over the concrete value.
-
-	\subsection {Reflected terms}
-
-We will have one datatype for reflected terms in each algebraic structure. \\
-Each of these datatype is parametrised on a type $c$ : the real type on which we want to prove equalities.
-It is also indexed over an instance of the corresponding typeclass for $c$ (we usally call it $p$, because it looks like a proof saying that the structure $c$ has the desired properties), and indexed over a context (a vector $g$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the encoded value.
-A magma is only equiped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constantes, variables, and addition.
-
-\begin{code}[caption=Reflected terms in a Magma, captionpos=b, label=lst1:haskell2]  
-data ExprMa : Magma c -> (Vect n c) -> c -> 
-              Type where
-    ConstMa : (p : Magma c) -> (g:Vect n c) 
-         -> (c1:c)  -> ExprMa p g c1 
-    PlusMa : {p : Magma c} -> {g:Vect n c}
-         -> {c1:c} -> {c2:c} 
-         -> ExprMa p g c1 
-         -> ExprMa p g c2 
-         -> ExprMa p g (Plus c1 c2) 
-    VarMa : (p:Magma c) -> {g:Vect n c}
-         -> {c1:c} 
-         -> VariableA (get_eq_magma p) g c1
-         -> ExprMa p g c1
-\end{code}
-
-When we encode a constant $c1$ in a context $\Gamma$, we use the constructor $ConstMa$ to produce a term of type $ExprMa\ p\ \Gamma\ c1$ : the index representing the concrete value is simply the constant $c1$.
-If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the courant instance $p$.
-
-We've used $get\_eq\_magma$ in the constructor $VarMa$, and this is just a function which returns the equality test of the $Set$ level.
-For the moment, the definition of $Variable$ is just the following:
-
-\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
-data Variable : {c:Type} -> {n:Nat}
-  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
-  -> (Vect n c) -> c -> Type where
-    RealVariable : (c_equal:(c1:c)->(c2:c)
-                    ->Maybe(c1=c2))
-          -> (g:Vect n c) -> (i:Fin n) 
-          -> Variable c_equal g (index i g) 
-\end{code}
-
-This type will be extended with some other constructors later on. For the moment, we can only express a "real variable". \\
-\\
-Because there is no more operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
-However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :
-
-\begin{code}[caption=Reflected terms in a Group, captionpos=b, label=lst1:haskell2]
-data ExprG :  Group c -> (Vect n c) -> c -> 
-              Type where
-    ConstG : (p : dataTypes.Group c) 
-       -> (g:Vect n c)
-       -> (c1:c) -> ExprG p g c1
-    PlusG : {p : Group c} -> {g:Vect n c} 
-       -> {c1:c} -> {c2:c} 
-       -> ExprG p g c1 
-       -> ExprG p g c2 
-       -> ExprG p g (Plus c1 c2)
-    MinusG : {p : Group c} -> {g:Vect n c} 
-       -> {c1:c} -> {c2:c} 
-       -> ExprG p g c1 
-       -> ExprG p g c2 
-       -> ExprG p g (Minus c1 c2)
-    NegG : {p : Group c} -> {g:Vect n c} 
-       -> {c1:c} -> ExprG p g c1 
-       -> ExprG p g (Neg c1)
-    VarG : (p : Group c) -> {g:Vect n c} 
-       -> {c1:c}
-       -> Variable (get_eq_group p) Neg g c1 
-       -> ExprG p g c1
-\end{code}
-
-Of course, the index of type $c$ (the real value represented by the expression) is always expressed with the oeprations available in the typeclass $p$, and here for a Group they are$Plus$, $Minus$ and $Neg$.
-
-	\subsection { A bit of notation}
-The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (making $c$ a Semi-Group, a Monoid, a Group...).
-The fact that $c$ fulfil the specification of an algebraic structure is expressed as an instance of the corresponding typeclass, and this instance will be denoted as $p$.
-The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ g\ x$ : $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$.
-We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ g\ x'$ and a proof of $p1:x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ g\ y'$ and a proof of $p2:y=y'$.
-
-	\subsection {Deciding equality}
-
-Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
-This is fact a nice property that all these algebraic structures admit a canonical represent for any element.
-Without this property, it would become a lot more complicated to decide the equality without bruteforcing the
-These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
-
-The profile of these function will always be the same :
-\begin{code}[caption=Type of the reduction function for terms reflecting elements in a Group, captionpos=b, label=lst1:haskell2]
-	groupReduce : {c:Type} -> {n:Nat} 
-	  -> (p:Group c) -> {g:Vect n c} -> {x:c}
-	  -> (ExprG p g x) 
-	  -> (x' ** (ExprG p g x', x=x'))
-\end{code}
-
-This function will have more job to do in structures with multiple axioms (like Group), than it has in the easier structure underneath.
-The detail of these functions will be given in the next section.
-For the moment, we just assume that we've got such a function for each structure, and we want to decide equality between elements.
-We continue to give the details for the Group, but the principle is exactly the same with the other structures.
-
-We want to write the following function :
-\begin{code}[caption=Type of the function for deciding equality between elements in a Group, captionpos=b, label=lst1:haskell2]
-	groupDecideEq : (p:Group c) -> {g:Vect n c} 
-	   -> {x : c} -> {y : c} 
-	   -> (ExprG p g x) -> (ExprG p g y) 
-	   -> Maybe (x = y)
-\end{code}
-
-The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
-Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by "composing" the two proofs $p1$ and $p2$.
-
-\begin{code}[caption=Decides if two elements of a group are equal, captionpos=b, label=lst1:haskell2]
-groupDecideEq p e1 e2 =
-  let (x' ** (e1', p1)) = 
-    groupReduce p e1 in
-  let (y' ** (e2', p2)) = 
-    groupReduce p e2 in
-	     buildProofGroup p e1' e2' p1 p2
-\end{code}
-
-The syntactical test of equality between $e1'$ and $e2'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$.
-
-\begin{code}[caption=Composes the two proofs if the NF are the same, captionpos=b, label=lst1:haskell2]
-buildProofGroup : (p:Group c) -> {g:Vect n c} 
-  -> {x : c} -> {y : c} 
-  -> {x':c} -> {y':c} 
-  -> (ExprG p g x') 
-  -> (ExprG p g y') 
-  -> (x = x') -> (y = y') 
-  -> (Maybe (x = y))
-buildProofGroup p e1 e2 p1 p2 
-       with (exprG_eq p _ e1 e2)
-	buildProofGroup p e1 e1 p1 p2 | Just refl = 
-	         ?MbuildProofGroup
-	buildProofGroup p e1 e2 p1 p2 | Nothing =
-	          Nothing
-\end{code}
-
-The proof for the metavariable MbuildProofGroup is exactly the same as the corresponding one (called $bp$) in the smallest example showed in section 2.
-
-\section {The details : normalization functions (4/5 pages)}
-
-	\subsection {Normalization of terms in a Magma}
-
-	\subsection {Normalization of terms in a Semi-Group}
-
-	\subsection {Normalization of terms in a Monoid}
-
-	\subsection {Normalization of terms in a Group}
-
-\section {Auxiliary problems : Automatic metaification and coating (1/2 pages)}
-
-	\subsection {Automatic metaification}
-
-	\subsection {Coating : transforming this stuff into real tactics}
-
-\section {Related work (1 page)}
-
-\section {Conlusion (0,5/1 page)}
-
-\acks
-
-Acknowledgments : Chris, Mattus for discussions. Jan for help in latex.
-
-% We recommend abbrvnat bibliography style.
-
-\bibliography{biblio}
-
-\appendix
-\section{Appendix Title}
-
-This is the text of the appendix, if you need one.
-
-\end{document}
-
-%                       Revision History
-%                       -------- -------
-%  Date         Person  Ver.    Change
-%  ----         ------  ----    ------
-
-%  2013.06.29   TU      0.1--4  comments on permission/copyright notices

### 2015-11-01 paper-JFP-DTP/content/introduction.tex
paper continued

@@
 Example of these goals are $\forall\ xy:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$ for natural numbers with the standard addition and multiplication, or $\forall\ l1\ l2\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\ (l3++l4)$ for lists and the standard concatenation operation.
-Proving this kind of auxiliary lemmas is time consuming, is not general enough for being reusable, and isn't the interesting part of formal certification.
-Therefore, we would like to be able to automatize a far as possible the constructions of proofs for thins kind of lemmas for the Idris language, a dependently typed language which doesn't offer many proof automations at the moment.
+Proving this kind of auxiliary lemmas is time consuming, is much too specific for being reusable, and isn't the interesting part of formal certification.
+Therefore, we would like to be able to automatize a far as possible the constructions of proofs for this kind of lemmas for the Idris language, a dependently typed language which doesn't offer many proof automations at the moment.
 
@@
 Since in type theory, propositions are seen as types, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
-Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities.
-This is done by HMeq which follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
+Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities, and in this case, HMeq (which is the standard equality in Idris) can be used.
+It follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
 If this type is inhabited, then $a$ is said to be provably equal to $b$.
@@
 These proofs are usually omitted "on the paper" for the ordinary every day maths, but they are required by proof assistants, no matter how obvious they are.
-Currently, in the Idris\footnote{Idris is a relatively new, purely functional, and dependently typed programming language with full dependent types} programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
+Currently, in the Idris programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
 These proofs are the every day routine when we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
@@
 \begin{itemize}
-	\item 1) We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete ones, thus providing an easy way to pull out the proofs easily, and providing two guarantees.
+	\item 1) We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing an easy way to pull out the proofs easily, and providing two guarantees.
 The first one is that the reflection of a term is guaranteed to be a faithful representation of the term.
 The second is that any generated proof is guaranteed to be a proof of the required property, thus increasing the totality of the prover.
-The basic ideas of the technique is presented in the section 2 on a smaller problem, and is represented for the real problem in the section 3 and 4.
+The basic ideas of the technique is presented in the section 2 on a smaller problem with only natural numbers and the addition, and is represented for the real problem in the section 3 and 4.
 Discussion about the totality is provided in section 7.
 	\item 2) The normalisation procedures are implemented by following a correct by construction approach, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
-Again, this approach will be presented in section 2 on a smaller problem, and on section 4 for the complete hierarchy of algebraic structures.
+This approach is much more suitable for programming languages like Idris.
+Again, this approach will be presented in section 2 on the smaller problem, and on section 4 for the complete hierarchy of algebraic structures.
 	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.

### 2015-11-02 paper-JFP-DTP/content/basicIdeas.tex
JFP (special issue on DTP) paper continued

@@
 The general idea --which will also apply for the more general and bigger problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
-The goal of the normalization is of course to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
-For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normalisaiton form will be the complete left associative form.
+The goal of the normalization is to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
+For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normalisaiton form will be the complete left associative form (and every addition between an element $a$ and zero should be simplified to $a$).
 It will then be possible to decide the equality by simply comparing the normalised left and right hand sides with a simple syntactical equality.
+This is in fact what we do all the time that we have to decide if two things, written differently, are equal or not.
+When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a normal representation of polynomials, and to put both ponynomials in this form.
+If the normalised form are the same, then the two original polynomials are equal, otherwise they do not represent the same computation.\\
 \\
-\\
-In fact, such a normalisation function can't be written, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
+In fact, such a normalisation function can't be written directly, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
 And the normalization function needs to do different treatments for a "variable natural number" (a number which has been universally quantified) and for the constant $Z$.
-This is not possible yet, because once the variables are abstracted, they are just normal values of type $Nat$, and nothing tells us where they are in the left and right hand side of the equality : this information only exists at the level of the ASTs representing the two terms.
-For this reason, we will work by reflection, as we will do for the hierarchy of tactics in the next section.
-In this smaller example, it means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these.
+This is not possible yet, because once the variables are abstracted, they are just normal values of type $Nat$, and nothing tells us where they are in the left and right hand side of the equality.
+Indeed, this information only exists at the level of the ASTs representing the two terms.
+For this reason, we will work by reflection.
+In this little example, it means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
 This datatype will allow us to know the internal structure of a number, ie, where the variables and constants are.
@@
 \end{center}
-\caption{First version of reflected numbers}
+\caption{First version of reflected natural numbers}
 \figrule
@@
 $\forall\ e:Expr\ \Gamma,\ reify\ (reduce\ e)\ =\ reify\ e$ \\
-where $reify$ is a function computing the interpretation of an Expr in a context $\Gamma$, that is to say, the number that this Expr is encoding.
+where $reify$ is a function computing the interpretation of an Expr in a context $\Gamma$, that is to say, the natural number that this Expr is encoding.
 This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
@@
 \\
-To avoid these two sources of complexity, we add an index to the type $Expr$ which is the concrete number that an Expr is encoding.
-Thus, it won't be necessary to define the $interpretation$ function.
+To avoid these two sources of complexity, we add an index to the type $Expr$, and this index is the concrete number that an Expr is encoding.
+Thus, it won't be necessary to define the $reify$ function, as we will know directly the concrete element reflected by a term of type $Expr$ just by looking at its index.
 \\
@@
 using (x : Nat, y : Nat, gam : Vect n Nat)
-  data Expr : (Vect n Nat) -> Nat 
-               -> Type where
+  data Expr : (Vect n Nat) -> Nat -> Type where
        Plus : Expr gam x -> Expr gam y -> 
               Expr gam (x + y)
-       Var  : (i : Fin n) -> 
-              Expr gam (index i gam)
+       Var  : (i : Fin n) -> Expr gam (index i gam)
        Zero : Expr gam Z
@@
 \\
-We have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this small simple problem, but for the real collection of tactics, we will program an automatic metaification}.
+We have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this little tool example, but for the real collection of tactics, we will program an automatic reflection mechanism}.
 We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
@@
 \begin{verbatim}
-eqExpr : (e : Expr gam x) -> (e' : Expr gam y) 
-         -> Maybe (e = e')
-eqExpr (Plus x y) (Plus x' y') 
-        with (eqExpr x x', eqExpr y y')
-  eqExpr (Plus x y) (Plus x y) 
-        | (Just refl, Just refl) = Just refl
-  eqExpr (Plus x y) (Plus x' y') 
-        | _ = Nothing
+eqExpr : (e : Expr gam x) -> (e' : Expr gam y) -> Maybe (e = e')
+eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
+  eqExpr (Plus x y) (Plus x y) | (Just refl, Just refl) = Just refl
+  eqExpr (Plus x y) (Plus x' y') | _ = Nothing
 eqExpr (Var i) (Var j) with (decEq i j)
-  eqExpr (Var i) (Var i) 
-        | (Yes refl) = Just refl
-  eqExpr (Var i) (Var j) 
-        | _ = Nothing
+  eqExpr (Var i) (Var i) | (Yes refl) = Just refl
+  eqExpr (Var i) (Var j) | _ = Nothing
 eqExpr Zero Zero = Just refl
@@
 \begin{verbatim}
-buildProof : {x : Nat} -> {y : Nat} 
-       -> Expr gam x' -> Expr gam y' 
-       -> (x = x') -> (y = y') 
-       -> Maybe (x = y)
+buildProof : {x : Nat} -> {y : Nat} -> Expr gam x' -> Expr gam y' 
+           -> (x = x') -> (y = y') -> Maybe (x = y)
 buildProof e e' lp rp with (eqExpr e e')
-  buildProof e e lp rp | Just refl = ?bp1
+  buildProof e e lp rp | Just refl = ?MbuildProof
   buildProof e e' lp rp | Nothing = Nothing
@@
 
-As we mentioned, the proof for the metavariable is just a rewriting of the two equalities :
+As we mentioned, the proof for the metavariable $MbuildProof$ is just a rewriting of the two equalities :
 
@@
 \begin{verbatim}
-  bp1 = proof {
-  intros;
-  refine Just;
-  rewrite sym p1;
-  rewrite sym p2;
-  exact refl;
+  MbuildProof = proof {
+  intros; refine Just; rewrite sym p1; rewrite sym p2; exact refl;
 }  
@@
 \begin{verbatim}
-data LExpr : (gam : Vect n Nat) -> Nat 
-             -> Type where
+data LExpr : (gam : Vect n Nat) -> Nat -> Type where
      LPlus : LExpr gam x -> (i : Fin n) 
@@
 plusLExpr1 = proof {
-  intros;
-  rewrite sym xprf;
-  rewrite sym yprf;
-  rewrite prf;
+  intros; rewrite sym xprf; rewrite sym yprf; rewrite prf;
   mrefine plusAssociative;
@@
 plusLExpr2 = proof
-  intros
-  rewrite xprf 
-  rewrite (sym yprf)
-  mrefine plusZeroRightNeutral
+  intros; rewrite xprf; rewrite (sym yprf); mrefine plusZeroRightNeutral
 \end{verbatim}
@@
 \begin{verbatim}
-  reduce : Expr gam x -> 
-           (x' ** (Expr gam x', x = x'))
+  reduce : Expr gam x -> (x' ** (Expr gam x', x = x'))
   reduce e = 
@@
 At the moment, what we've got is not exactly a real tactic, in the sense that we only have a function which produces a value of type $Maybe (x = y)$.
-A real tactic would be a coating of this function, which could properly fail when the two terms are not equal.
+A real tactic would be a wrapper of this function that could properly fail with an error message when the two terms are not equal.
 However, here, when $x\ne y$, the function $testEq$ will simply produce the value $Nothing$.
@@
 e1_e2_testEq : (x, y, z : Nat) 
-       -> Maybe (((x + y) + (x + z)) 
-                  = (x + ((y + x) + z)))
-e1_e2_testEq x y z = testEq (e1 x y z)
-                            (e2 x y z)
+             -> Maybe (((x + y) + (x + z)) = (x + ((y + x) + z)))
+e1_e2_testEq x y z = testEq (e1 x y z) (e2 x y z)
 \end{verbatim}
@@
 
-\x => \y => \z => Just (replace (sym (replace 
-  (sym (replace (plusZeroRightNeutral x) 
-  (plusAssociative x 0 y))) (replace 
-  (sym (replace (plusZeroRightNeutral x) 
-  (plusAssociative x 0 z))) (replace (replace 
-  (plusZeroRightNeutral (plus xy)) 
-  (plusAssociative (plus x y) 0 x)) [...]
+\x => \y => \z => Just (replace (sym (replace (sym (replace 
+(plusZeroRightNeutral x) (plusAssociative x 0 y))) (replace 
+(sym (replace (plusZeroRightNeutral x) (plusAssociative x 0 z))) (replace 
+(replace (plusZeroRightNeutral (plus xy)) (plusAssociative (plus x y) 0 x)) 
+[...]
 : (x : Nat) -> (y : Nat) -> (z : Nat) 

### 2015-11-02 paper-JFP-DTP/content/introduction.tex
JFP (special issue on DTP) paper continued

@@
 If this type is inhabited, then $a$ is said to be provably equal to $b$.
-Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A : $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$.
+Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A.
+That gives $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$.
 Often, for convenience, $Id_A\ a\ b$ is written $a\ =_A\ b$.
@@
 \subsection{A motivating little example}
-We introduce a little example which shows how these lemmas appear when we index a type by values like natural numbers.
+We introduce a little example which shows how these proof obligations appear when we index a type by values like natural numbers.
 Let's define the types of $Bit$ and $Binary$ number, both indexed over the value they represent, expressed in base ten as a natural number.
@@
 
-data Binary : (width : Nat) -> (value : Nat) 
-              -> Type where
+data Binary : (width : Nat) -> (value : Nat) -> Type where
      zero : Binary Z Z
-     (#) : Binary w v -> Bit bit 
-           -> Binary (S w) (bit + 2 * v)
+     (#) : Binary w v -> Bit bit -> Binary (S w) (bit + 2 * v)
 \end{verbatim}
@@
 \begin{verbatim}
-addBit : Bit x -> Bit y -> Bit c ->
-          (bX ** (bY ** 
-              (Bit bX, Bit bY, 
-                  c + x + y = bY + 2 * bX)))
+addBit : Bit x -> Bit y -> Bit c -> (bX ** (bY ** 
+              (Bit bX, Bit bY, c + x + y = bY + 2 * bX)))
 addBit b0 b0 b0 = (_ ** (_ ** (b0, b0, _)))
@@
 
-Now, we want define a function adding two binary numbers and a bit (used as a carry).
+Now, we want to define a function adding two binary numbers and a bit (used as a carry).
 We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit, and this result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
@@
 \begin{verbatim}
-adc : Binary w x -> Binary w y -> Bit c 
-     -> Binary (S w) (c + x + y)
+adc : Binary w x -> Binary w y -> Bit c -> Binary (S w) (c + x + y)
 adc zero zero carry = zero # carry
 adc (numx # bX) (numy # bY) carry
-   = let (v1 ** (v2 ** (carry0, lsb, _))) = 
-      addBit bX bY carry in
+   = let (v1 ** (v2 ** (carry0, lsb, _))) = addBit bX bY carry in
           adc numx numy carry0 # lsb
@@
 The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
-For this kind of situation, Idris provides the possibility to write a provisional definition\footnote{provisional definitions are described with the brief overview of the language in page XX} with the syntax "?=".
-Using a provisional definitions here will make the definition being accepted, but we will have to prove afterwards that in both patterns, the term provided can be converted into a term which has the right type.
-However, for both of these two patterns, it is possible to prove the equality between the expected and the provided types, which is even more than we need.
-These proofs will be made of a series of rewriting, and each rewriting will use a property about the operation $+$ on natural numbers.
+For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
+Using a provisional definitions here will make the definition being accepted, but it also generated two proof obligations, which say that for both patterns, the provided term can be converted into a term which has the expected type.
+For both of these two patterns, it is possible to prove the equality between the expected and the provided types, because this function is correct.
+These proofs are made of a series of rewriting, and each rewriting uses a property about the operation $+$ on natural numbers.
 For example, the proof for the second pattern look like this :
@@
 This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
-This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will requires us to replace this proof by another one doing the same job : assuring the correspondance betwen the expected index and the given index.
+This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will requires us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
 The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
@@
 The novelty is in the approach that we follow, using a new kind of type-safe reflection.
-Working by reflection for implementing tactics has been done several times, as in [paper Coq ring] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
-We will compare our approach with other implementation of this kind of tactics in section [Blabla].
+Working by reflection for implementing tactics has been done several times, as in [ref 1] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
+We will compare our approach with other implementation of this kind of tactics in section 6.
 \\
@@
 The second is that any generated proof is guaranteed to be a proof of the required property, thus increasing the totality of the prover.
-The basic ideas of the technique is presented in the section 2 on a smaller problem with only natural numbers and the addition, and is represented for the real problem in the section 3 and 4.
+The basic ideas of the technique are first presented in the section 2 on a smaller problem with only natural numbers and the addition, and are adapted for the hierarchy of tactics in the section 3 and 4.
 Discussion about the totality is provided in section 7.
@@
 	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.
-For example simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuse it.
+For example simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuses it.
 As it will be discussed in section 4, reusability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.

### 2015-11-02 paper-JFP-DTP/paper.tex
paper continued
JFP (special issue on DTP) paper continued

@@
 
+\input{./content/basicIdeas.tex}
+
+\input{./content/generalProblem.tex}
+
+\input{./content/automaticReflection.tex}
+
+\input{./content/discussion.tex}
+
 \begin{thebibliography}{}
- \bibitem[\protect\citename{Augustsson and Johnsson, }1987]{AJ187}
-   Augustsson,~L. and Johnsson,~T. (1987) LML users' manual. PMG
-   Report, Department of Computer Science, Chalmers University of
-   Technology, Goteborg, Sweden.
- \bibitem[\protect\citename{Butcher, }1981]{Butcher}
-   Butcher,~J. (1981) Copy-editing: the Cambridge handbook.
-   Cambridge University Press.
- \bibitem[\protect\citename{Chicago, }1982]{Chicago}
-   The Chicago manual of style. University of Chicago Press.
- \bibitem[\protect\citename{Conklin, }1987]{JC87}
-   Conklin,~J. (1987) Hypertext: an introduction and survey.
-   \emph{IEEE Computer}, 20~(9): pp.~17--41.
- \bibitem[\protect\citename{Dijkstra, }1976]{EWD76}
-   Dijkstra,~E.~W. (1976) \emph{A Discipline of Programming}.
-   Prentice-Hall.
- \bibitem[\protect\citename{Knuth, }1984]{DEK84}
-   Knuth,~D.~E. (1984) Literate programming. \emph{BCS Comput. J.}
-   27~(2): 97--111 (May).
- \bibitem[\protect\citename{Lamport, }1986]{LaTeX}
-   Lamport,~L. (1986) \LaTeX: a document preparation system
-   (2nd edition). Addison-Wesley, New York.
- \bibitem[\protect\citename{Reynolds, }1969]{JCR69}
-   Reynolds,~J.~C. (1969) Transformation systems and the
-   algebraic structure of atomic formulas. In B. Meltzer and
-   D. Michie (editors), \emph{Machine Intelligence 5}, pp.~135--151.
-   Edinburgh University Press.
- \bibitem[\protect\citename{Toyn \emph{et al.}, }1987]{TDR87}
-   Toyn,~I., Dix,~A. and Runciman,~C. (1987) Performance
-   polymorphism. In \emph{Functional Programming Languages and
-   Computer Architecture, Lecture Notes in Computer Science, 274},
-   pp.~325--346. Springer-Verlag.
+ \bibitem[\protect\citename{B. Gregoire and A. Mahboubi}2005]{Coq2005}
+   Gregoire,~B. and Mahboubi,~A. (2005) Proving equalities in a commutative ring done right in Coq, TPHOLs 2005, Oxford, UK, August 22-25, 2005, Proceedings
+
 \end{thebibliography}

### 2015-11-22 Old/OldPapers/ICFP15paper/content/idrisOverview.tex
Problem with reflection to Nat worked

TO BE CONTINUED

Idris is a purely functional programming language fairly recent. Its main purpose is to bring formal certification into the scope of programming : instead of Coq or Isabelle, Idris is not intended to be a proof assistant, but as a real programming language usable by non-logicians. One of its more interesting features -and the one that enables to effectively manipulate logical statements and proofs in the language- is its full and effective support for dependent types. Especially, Idris has a particularly efficient dependent pattern matching, which is something mandatory for writing code correct by construction.

\subsection{Some common datatypes}

Maybe et Dec

\subsection{Dependent pattern maching}

Idris supports dependent pattern maching. Pattern maching a term can gives information about the shape of another term. For example, when pattern maching a $n:nat$ and a $v:Vect\ n\ T$ simultaneously, the case $Z$ for $n$ forces $v$ to be the empty vector $Nil$.
The following function $foo$ is therefore spotted as total by the Idris compiler.

\begin{code}[caption=A function known to be total thanks to dependent pattern maching, captionpos=b, label=lst1:haskell2]
foo : (n:Nat) -> (Vect n T) -> aType
foo Z Nil = something
foo (S pn) (h::t) = somethingElse
\end{code}

Also, computing a value or deciding an equality can potentially gives us information about multiple parameters of a function. For his purpose, Idris provides the "with" clause.

ADD EXAMPLE OF WITH CLAUSE HERE

\subsection{Provisional definitions}

blabla

\begin{code}[caption=Using decidable equality to infer the shape of parameters, captionpos=b, label=lst1:haskell2]
data Parity : Nat -> Set where
    even : Parity (plus n n)
    odd  : Parity (S (plus n n));

parity : (n:Nat) -> Parity n
parity Z = even {n=Z}
parity (S Z) = odd {n=Z}
parity (S (S k)) with parity k {
  parity (S (S (plus j j)))     
        | even ?= even {n=S j}  [paritySSe]
  parity (S (S (S (plus j j)))) 
        | odd  ?= odd {n=S j}   [paritySSo]
}
\end{code}

bla

### 2016-01-05 paper-JFP-DTP/content/basicIdeas.tex
paper...

@@
 
-In order to better explain the way we are trying to solve our problem, we will first present it on a simplified version, in which we aim to deal with universally quantified natural numbers, the properties of associativity and neutral element for the operation plus.
+We will first present the basic ideas of our technique on a simplified version, in which we aim to deal with universally quantified natural numbers, the properties of associativity and neutral element for the operation plus.
 For example, we would like to be able to automatically generate proofs of goals like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
-For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a neutral element on the right side of the plus : $\forall x, x + Z = x$.
-Zero is also a left neutral element, but this property is not needed, as we have this behavirous by reduction as $+$ is defined recursively on its first argument.
+For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a right neutral element for the $+$ operation : $\forall x, x + Z = x$.
+Note that $Z$ is also a left neutral element, but this property is not needed, as we have this behaviour by reduction as $+$ is defined recursively on its first argument.
 Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce of proof of this equality if appropriate, when "equal" has the meaning "syntactically equal or equal thanks to the associative and neutral properties".
-\\
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce a proof of this equality if possible, when "equal" has the meaning "syntactically equal or equal thanks to the associative and neutral properties".
 
-\subsection{Working by reflexion}
+\subsection{Working by reflection}
 
@@
 On the example 1, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
-The general idea --which will also apply for the more general and bigger problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
+The general idea --that will also apply for the more general and bigger problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
 The goal of the normalization is to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
-For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normalisaiton form will be the complete left associative form (and every addition between an element $a$ and zero should be simplified to $a$).
+For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normal form will be completely left associative, and every addition between an element $a$ and zero should be simplified to $a$.
 It will then be possible to decide the equality by simply comparing the normalised left and right hand sides with a simple syntactical equality.
 This is in fact what we do all the time that we have to decide if two things, written differently, are equal or not.
-When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a normal representation of polynomials, and to put both ponynomials in this form.
+When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a canonical representation of polynomials, and to put both ponynomials in this form.
 If the normalised form are the same, then the two original polynomials are equal, otherwise they do not represent the same computation.\\
@@
 And the normalization function needs to do different treatments for a "variable natural number" (a number which has been universally quantified) and for the constant $Z$.
-This is not possible yet, because once the variables are abstracted, they are just normal values of type $Nat$, and nothing tells us where they are in the left and right hand side of the equality.
-Indeed, this information only exists at the level of the ASTs representing the two terms.
+This is not possible yet, because once the variables are abstracted, they are just ordinary values of type $Nat$, and there is no way to distinguish them.
+Indeed, this information only exists at the level of the ASTs representing the two terms, and we do not have a direct access to these syntax trees.
 For this reason, we will work by reflection.
-In this little example, it means that we will define a datatype which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
-This datatype will allow us to know the internal structure of a number, ie, where the variables and constants are.
-Indeed, this datatype will allow pattern matching.
+In this little example, it means that we will define a data type which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
+This data type will allow us to know the internal structure of a number, ie, where the variables and constants are.
+Indeed, this data type will allow pattern matching.
 Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't what we needed.
-With the first approximation of the datatype $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
+With the first approximation of the data type $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
 
@@
 where $reify$ is a function computing the interpretation of an Expr in a context $\Gamma$, that is to say, the natural number that this Expr is encoding.
-This proof can be quite tricky to make because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
+This proof can be quite tricky to build because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
 For this little example, the reduction procedure will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with and it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
+As much as possible, we want to avoid the pain of proving the correctness of the tactic, and for this reason, we don't want to have to write huge proofs.
 \\
 \\
-To avoid these two sources of complexity, we add an index to the type $Expr$, and this index is the concrete number that an Expr is encoding.
-Thus, it won't be necessary to define the $reify$ function, as we will know directly the concrete element reflected by a term of type $Expr$ just by looking at its index.
-\\
+To avoid this source of complexity, we add an index to the type $Expr$, and this index is precisely the concrete number that the Expr is encoding.
+This is the first brick to our type-safe reflection mechanism.
+Thus, it won't be necessary to define the $reify$ function, as we will know directly the concrete element reflected by a term of type $Expr\ \Gamma\ x$ just by looking at its index $x$.
+Something more important is that we will also have a guarantee that the reflection of an expression $e$ is indeed a faithful representation of $e$ \\
 
@@
 We want to write the reduction function on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
-Thus, $reduce$ will produce the proof that the new Expr produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
+Thus, $reduce$ will produce the proof that the new Expr freshly produced has the same interpretation as the original Expr, and this will be made easier by the fact that the data type Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
 Thus, we can write the type of $reduce$ like this : \\
@@
 We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
-We will do the same with $e2$, and we will get a new number $y$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
+We will do the same with $e2$, and we will get a new number $y'$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
 \\
-Now, we can compare $e1'$ and $e2'$ using a standard syntactical equality because these two expressions are in normal form :
+It is now enough to simply compare $e1'$ and $e2'$ using a standard syntactical equality because these two expressions are in normal form :
 
@@
 By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$.
+This is what the function $buildProof$ is doing.
 
@@
 \end{figure}
+
+The $Expr\ \Gamma\ x'$ will be used as the normalised left hand size of the equality, which represents the value $x'$.
+Before the normalisation, the reflected LHS was reflecting the value $x$.
+The $Expr\ \Gamma\ y'$ is the normalised right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation.
+The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function also produces the proof of equality between the previous and the new concrete values.
 
@@
 
-This datatype has only two constructors. In fact, it combines the previous $Var$ and $Plus$ constructors so that it becomes impossible to write an expression which isn't left associative.
+This datatype has only two constructors.
+In fact, it combines the previous $Var$ and $Plus$ constructors so that it becomes impossible to write an expression which isn't left associative (because LPlus is only recursive on its first argument).
 
@@
 As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightNeutral$).
+
+\subsection{Construction of the reflected terms}
+
+Note that for the moment, even if what we have is perfectly usable and works, we had to create these encodings $e1$ and $e2$ by hand, which is easy but time consuming.
+We have replaced the (potentially hard) problem of proving something by the simplest problem of simply creating some encodings.
+This is already a huge simplification, because as it can be seen in the definitions of $e1$ and $e2$, the reflected terms completely follow the structure of the expression to encode : there's absolutely no creativity needed for this task, unlike the proving activity.
+The way to create the encodings is in fact so systematic that, of course, we would like to automatize it in order to get a real and completely automatic tactic.
+
+However, we can also note that even when done by hand, there is no room for making mistakes in this simple task of encoding : we simply can't generate a wrong encoding : if $e1$ and $e2$ are not respectively reflecting $((x+y) + (x+z))$ and $(x + ((y + x) + z))$ then these definitions won't typecheck because the expected and real index won't match.
+[perhaps giving example of rejected wrong definition for $e1$ here]
+
+We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively.

### 2016-01-05 paper-JFP-DTP/content/introduction.tex
paper...

@@
 
-Example of these goals are $\forall\ xy:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$ for natural numbers with the standard addition and multiplication, or $\forall\ l1\ l2\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\ (l3++l4)$ for lists and the standard concatenation operation.
+For example, when working with natural numbers, one might need to prove the lemma $\forall\ xy:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$, and when working with lists and concatenation of lists, one might need to prove something like $\forall\ l1\ l2\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\ (l3++l4)$.
 Proving this kind of auxiliary lemmas is time consuming, is much too specific for being reusable, and isn't the interesting part of formal certification.
-Therefore, we would like to be able to automatize a far as possible the constructions of proofs for this kind of lemmas for the Idris language, a dependently typed language which doesn't offer many proof automations at the moment.
+Therefore, we would like to be able to automatize a far as possible the constructions of proofs for this kind of lemmas for the Idris language, a dependently typed language which doesn't contain many proof automations at the moment.
 
 A common situation in formal certification is the following.
-We have a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B, and if this semantic says everything we care about for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
-The equality symbol $=$ is the propositional equality, as it was the case in the two previous examples with natural numbers and list.
-We remind the readers in the next subsection how the jugmental equality is defined in Idris, and more generally in Intensional Type Theories.
+We have a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B.
+If this semantic says everything we care about for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
+All the equality symbols $=$ we have encountered so far refer to the propositional equality.
+We remind the readers in the next subsection how this equality is defined in Idris, and more generally in Intentional Type Theories, and how it differs from the --more primitive-- jugmental equality.
 
@@
 In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
-Since in type theory, propositions are seen as types, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
-Thus, if $a$ and $b$ are of type\footnote{It can be usefull sometimes to allow the writing of heterogeneous equalities, and in this case, HMeq (which is the standard equality in Idris) can be used.
+Since in type theory, propositions are seen as types \cite{How80}, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
+Thus, if $a$ and $b$ are of type\footnote{It can be useful sometimes to allow the writing of heterogeneous equalities, and in this case, HMeq (which is the standard equality in Idris) can be used.
 It follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
@@
 The judgemental equality is something which can't be negated or assumed : we can't talk about this primitive equality inside the theory.
-Whether or not two expressions are equal by definition is a just a matter of expanding out the definitions.
+Whether or not two expressions are equal by definition is just a matter of expanding out the definitions.
 For example, if $f\ :\ \mathbb{N}\ \rightarrow\ \mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is definitionally equal to $7$.
@@
 
-The judgemental equality was needed to be included in the propositional equal, because what is equal by definition should be provably equal.
+The judgemental equality has to be included in the propositional equal, because what is equal by definition should be provably equal.
 This is accomplished by giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
@@
 This is not the case : the propositional equality doesn't only contains the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the other constructors, then this proposition will hold for any term of this type.
+This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the recursive constructors, then this proposition will hold for any term of this type.
 
-Because this proposition can be any proposition, it can also be the propositional equality.
-That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it was before (in the definition 1), and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
+The induction principles work for any proposition, so it also works with the propositional equality.
+That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it looks like (in the definition 1), and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
 The activity of proving equalities is therefore in these theories something which isn't automatically decidable by the type-checker in the general case.
@@
 This type of binary numbers\footnote{This representation allows to represent a binary number of width zero with the constructor $zero$, but instead we could have decided to replace this constructor with a constructor taking one bit as parameter and producing a binary number of width one.
-The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value represented by $bin$ plus the value represented by the added digit.
+The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value represented by $bin$ ($v$) plus the value represented by the added digit ($bit$).
 
 Now, we're tempted to write the addition between two binary numbers.
-We start with an auxiliary function which adds three bits (the third one will latter play the role of the carry), and produces the two bits of the result.
+We start with an auxiliary function which adds three bits (the third one will latter play the role of the carry), and produces the two bits of the result : the first one is the most significant bit and the second one is the least significant bit .
 
@@
 For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
-Using a provisional definitions here will make the definition being accepted, but it also generated two proof obligations, which say that for both patterns, the provided term can be converted into a term which has the expected type.
+Using a provisional definitions here will make the definition being accepted, but it also generates two proof obligations, one for each pattern, and both of them are saying that the provided term can be converted into a term which has the expected type.
 For both of these two patterns, it is possible to prove the equality between the expected and the provided types, because this function is correct.
 These proofs are made of a series of rewriting, and each rewriting uses a property about the operation $+$ on natural numbers.
-For example, the proof for the second pattern look like this :
+For example, the proof for the second pattern looks like this :
 \begin{figure}[H]
@@
 
-Thus, this kind of proofs consists of a potentially long sequence of rewriting, where every rewriting step uses one of the available properties.
+Thus, this kind of proofs consists of a potentially long sequence of rewriting, where each rewriting step uses one of the available properties.
 Without some specific automation, this sequence of rewriting is done by the user of the proof assistant.
@@
 
-We would like to let a solver for commutative groups to find the proof on its own.
-This is a something doable because algebraic structures like monoid, group, commutative group and ring have all a very nice property : every element in them has a canonical represent.
-Thus these kind of proofs can be automatically produced by a tactic by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality).
+We would like to let a prover for commutative groups find a proof on its own.
+This is something doable because algebraic structures like monoid, group, commutative group and ring have all a very nice property : every element in them can be normalised to a canonical representation.
+Thus these kind of proofs can be automatically produced by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality), because now that they are in normal form, being equal is just a matter of being syntactically the same entity.
 

### 2016-01-07 paper-JFP-DTP/content/basicIdeas.tex
paper JFP DTP continued

@@
 We decide that the left associative form will be the canonical representation.
-Thus, the $reduce$ function has to rewrite the metaified term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
+Thus, the $reduce$ function has to rewrite the reflected term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
 + xn)$.
@@
 
-We notice that for transforming the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
+We notice that in order to transform the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
 \\
@@
 We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively. 
+By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values.
+That implies that when applying this function to some arguments, the arguments aren't reduce before the evaluation of this function.
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+
+%reflection
+reflectList : (G : List Nat) ->
+          (x : Nat) -> (G' ** Expr (G ++ G') x)
+reflectList G Z = ([] ** Zero)
+
+...
+
+\end{verbatim}
+\end{center}
+\caption{Reflection for natural numbers}
+\figrule
+\end{figure}

### 2016-01-07 paper-JFP-DTP/content/1-introduction.tex
Renamed content files

--- 2016-01-05 paper-JFP-DTP/content/introduction.tex
@@
 The basic ideas of the technique are first presented in the section 2 on a smaller problem with only natural numbers and the addition, and are adapted for the hierarchy of tactics in the section 3 and 4.
-Discussion about the totality is provided in section 7.
+Discussion about the totality is provided in section 5.2.
 	\item 2) The normalisation procedures are implemented by following a correct by construction approach, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
 This approach is much more suitable for programming languages like Idris.
-Again, this approach will be presented in section 2 on the smaller problem, and on section 4 for the complete hierarchy of algebraic structures.
+Again, this approach will be presented in section 2 on the smaller problem, and on sections 3 and 4 for the complete hierarchy of algebraic structures.
 	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.

### 2016-01-07 paper-JFP-DTP/content/3-generalProblem.tex
Renamed content files

\section {Back to the general problem : A hierarchy of tactics}

Now that we have introduced our ideas on this little example, it is time to apply them for our main goal, more general : the implementation of a hierarchy of tactics proving equalities in algebraic structures. Very often, the properties available on a given type are the one of a well known algebraic structure : magma, semi-group, monoid, group, commutative-group...  We will no longer only work with the properties of associativity and of right neutral element for natural numbers as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.

We will construct a hierarchy of type classes and will write one tactic for each of these type class. All our tactics (the group prover, the monoid prover...) will be able to work on any type, being given that an instance of the corresponding type class is provided.

\subsection {Hierarchy of type classes}

All our tactics will require to have a way of testing the equality between elements of the underlying set, that is to say, a way to test equality between constants. For this reason, we define a notion of "set", which only requires this equality. All the algebraic structures above will extend this type class :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c where
    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
\end{verbatim}
\end{center}
\caption{Set}
\figrule
\end{figure}

Note that this decidable equality is weak in the sense that it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different -- instead, it simply produces the value $Nothing$--. That's quite natural, since we want to generate proof of equality, and not to generate counter examples for proving dis-equalities, which is another problem.

Obvisouly, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure. Therefore, equalities in a $Set$ are just the syntactical equality, and they can simply be proven with $Refl$.

The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no axioms about this operation.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Set c => Magma c where
    Plus : c -> c -> c
\end{verbatim}
\end{center}
\caption{Magma}
\figrule
\end{figure}

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with a $set\_eq$ operation), and if it has a $Plus$ operation.

A bit more interesting is the Semi-Group type class. A semi-Group is a Magma (ie, it still has a $Plus$ operation), but moreover it has the property of associativity for this operation.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class Magma c => SemiGroup c where
    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
               -> (Plus (Plus c1 c2) c3 = Plus c1 (Plus c2 c3))
\end{verbatim}
\end{center}
\caption{Semi-Group}
\figrule
\end{figure}

And the Monoid structure is a Semi-Group with the property of neutral element.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
class SemiGroup c => Monoid c where
    Zero : c    
    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 = c1)    
    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero = c1)
\end{verbatim}
\end{center}
\caption{Monoid}
\figrule
\end{figure}

A Group is a Monoid, with two new operations. The binary operation $Minus$, and the unary operation $Neg$. We must have the property that $Minus$ can always be simplified with the $Neg$ and the $Plus$. That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$. We let the possibility to express a $Minus$ just for convenience for the user, as he won't have to rewrite by hand his $Minus$ operations with sums of negations.
The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
-- This is just a conjunctive predicate
hasSymmetric : (c:Type) -> (p:Monoid c) -> c -> c -> Type
hasSymmetric c p a b = (Plus a b = Zero, Plus b a = Zero)

class Monoid c => Group c where
    Minus : c -> c -> c
    Neg : c -> c
    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) 
    Plus_inverse : (c1:c) -> hasSymmetric c _ c1 (Neg c1)
\end{verbatim}
\end{center}
\caption{Group}
\figrule
\end{figure}

This hierarchy can be extended without difficulty with Abelian groups (with the axiom of commutativity), Ring, Commutative Rings...

These type classes will somehow be used as predicates on types. In order to call a prover on a type $c$, the user of the system will have to satisfy the corresponding type class by providing an instance of it for his type $c$, ie, he will have to prove that the corresponding properties effectively hold for $c$.

Note that the properties will either be obtained "by implementation" if the operations are real --computable-- functions, or by axioms if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$, ...) are defined as axioms.

As discussed in the section 2.1 on the smaller problem, the algorithm of normalization will not directly use the concrete value of type $c$, but a reflected term, indexed over the concrete value. That still holds here.

	\subsection {Reflected terms}

We need to define a datatype for reflecting terms in each algebraic structure.
Each of these datatype is parametrised over a type $c$, which is the real type on which we want to prove equalities (the $carrier$ type). It is also indexed over an instance of the corresponding type class for $c$ (we usually call it $p$, because it behaves as a $proof$ telling that the structure $c$ has the desired properties), and indexed over a context (a vector $\Gamma$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the concrete value being encoded.
A magma is only equipped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
data ExprMa : Magma c -> (Vect n c) -> c -> Type where
    ConstMa : (p : Magma c) -> (gam:Vect n c) -> (c1:c)  -> ExprMa p gam c1 
    PlusMa : {p : Magma c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
         -> ExprMa p gam c1 -> ExprMa p gam c2 
         -> ExprMa p gam (Plus c1 c2) 
    VarMa : (p:Magma c) -> {gam:Vect n c}
         -> (i:Fin n) -> ExprMa p gam (index i gam)
\end{verbatim}
\end{center}
\caption{Reflected terms for a Magma}
\figrule
\end{figure}

When we encode a constant $c1$ in a context $\Gamma$, we use the constructor $ConstMa$ to produce a term of type $ExprMa\ p\ \Gamma\ c1$ : the index representing the concrete value is precisely this constant $c1$.
If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (ie, a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the current instance $p$.

%For the moment, the definition of $Variable$ is just the following:
%
%\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
%data Variable : {c:Type} -> {n:Nat}
%  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
%  -> (Vect n c) -> c -> Type where
%    RealVariable : (c_equal:(c1:c)->(c2:c)
%                    ->Maybe(c1=c2))
%          -> (gam:Vect n c) -> (i:Fin n) 
%          -> Variable c_equal gam (index i gam) 
%\end{code}	
%
%This type will be extended with some other constructors later on. For the moment, we can only %express a "real variable", referred by its index $i$. \\
%\\
Because there is no additional operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
data ExprG :  Group c -> (Vect n c) -> c -> Type where
    ConstG : (p : dataTypes.Group c) -> (gam:Vect n c) 
      -> (c1:c) -> ExprG p gam c1
    PlusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
      -> ExprG p gam c1 -> ExprG p gam c2 
      -> ExprG p gam (Plus c1 c2)
    MinusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
      -> ExprG p gam c1 -> ExprG p gam c2 
      -> ExprG p gam (Minus c1 c2)
    NegG : {p : Group c} -> {gam:Vect n c} -> {c1:c} 
      -> ExprG p gam c1 -> ExprG p gam (Neg c1)
    VarG : (p : Group c) -> {gam:Vect n c} 
      -> (i:Fin n) -> ExprG p gam c1
\end{verbatim}
\end{center}
\caption{Reflected terms in a Group}
\figrule
\end{figure}

The index of type $c$ (the real value represented by an expression) is always expressed by using the available operations in the type class $p$, which for a group Group are $Plus$, $Minus$ and $Neg$.

	\subsection {A bit of notation}
The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...). The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ \Gamma\ x$. The term $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p1$ of $x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p2$ of $y=y'$.

	\subsection {Deciding equality}

Each algebraic structure will have a function for reducing the reflected terms into their normal forms. The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities. Without this property, it would become a lot more complicated to decide the equality without bruteforcing a serie of rewriting, that might even not terminate.
These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
	groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {gam:Vect n c} 
	  -> {x:c} -> (ExprG p gam x) -> (x' ** (ExprG p gam x', x=x'))
\end{verbatim}
\end{center}
\caption{Type of the reduction function for terms reflecting elements in a Group}
\figrule
\end{figure}

This function has more work to do in structures with multiple axioms (like Group), than for the easier structure underneath.
The details of these functions will be given in the next section. For the moment, we just assume that we've got such a function for each structure. We continue to give here the details for the Group structure, but the principle is exactly the same with the other structures, and the exact equivalent of what we did in the previous section with natural numbers and additions.

We want to write the following function :

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
	groupDecideEq : (p:Group c) -> {gam:Vect n c} -> {x : c} -> {y : c} 
	   -> (ExprG p gam x) -> (ExprG p gam y) 
	   -> Maybe (x = y)
\end{verbatim}
\end{center}
\caption{Type of the function for deciding equality between elements in a Group}
\figrule
\end{figure}

The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $p1$ and $p2$.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
groupDecideEq p e1 e2 =
  let (x' ** (e1', p1)) = groupReduce p e1 in
  let (y' ** (e2', p2)) = groupReduce p e2 in
	     buildProofGroup p e1' e2' p1 p2
\end{verbatim}
\end{center}
\caption{Decides if two elements of a group are equal}
\figrule
\end{figure}

The syntactical test of equality between $e1'$ and $e2'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$, similarly to what we've done on the smaller example in the previous section.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
buildProofGroup : (p:Group c) -> {gam:Vect n c} 
  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
  -> (ExprG p gam x') -> (ExprG p gam y') 
  -> (x = x') -> (y = y') 
  -> (Maybe (x = y))
buildProofGroup p e1 e2 p1 p2 with (exprG_eq p _ e1 e2)
    buildProofGroup p e1 e1 p1 p2 | Just refl = ?MbuildProofGroup
    buildProofGroup p e1 e2 p1 p2 | Nothing = Nothing
\end{verbatim}
\end{center}
\caption{Composes the two proofs if the normal forms are the same}
\figrule
\end{figure}

The proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) in the smallest example showed in section 2.3.

\subsection {Automatic reflection}
$reflect : \{n:Nat\} \rightarrow (gamma : Vect\ n\ c) \rightarrow (c1:c) \rightarrow Expr\ gamma\ c1$.
The reflected term is guaranteed to be a faithful representation of the input $c1$. We can write this function because Idris has a $\%syntax$ command which let us do pattern maching on syntax (ie, it let us access the AST of a term), and this $\%syntax$ produces functions which are typed ! (unlike Coq's $Ltac$).

### 2016-01-07 paper-JFP-DTP/content/4-detailsNormalisation.tex
Renamed content files

\section {Normalisation functions and re-usability of the provers}

We now describe briefly how the normalisation functions work for some of the structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures.

	\subsection {Normalization of terms in a Semi-Group}
In a semigroup, we only have to deal with the property of associativity. As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematical way, either left associative, or right associative. But this is not the only thing that we have to do. If $x$, $y$ and $z$ denote variables, we don't only want to transform $x+((y+4)+(5+z))$ into $(((x+y)+4)+5)+z$ if we've chosen the complete left associative form. We want first to do the addition of the constants $4$ and $5$ together (and we have the right to do so because we can rearrange the parenthesis thanks to the associativity, when constants are, like in this case, near to each other). Three possible patterns need this treatment : $(x+c1)+(c2+y)$ , $(x + c1) + c2$  and $c1 + (c2 + x)$ when $c1$ and $c2$ denote constants. The simplification between the constants $c1$ and $c2$ is denoted here as $[c1+c2]$, and this computation is done by the normalisation function of the underneath $Magma$ level, called $magmaReduce$.
The first pattern is rewritten into $(x + [c1+c2]) + y$, the second into $x + [c1+c2]$ and the third one into $[c1 + c2] + x$.
We show the code for the first pattern.

\begin{figure}[H]
\figrule
\begin{center}
\begin{verbatim}
assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1=c2))
assoc p gam (PlusSG 
             (PlusSG e1 (ConstSG _ _ c1)) 
             (PlusSG (ConstSG _ _ c2) e2)) =
    let (r_ih1 ** (e_ih1, p_ih1)) = (assoc p gam e1) in
    let (r_ih2 ** (e_ih2, p_ih2)) = (assoc p gam e2) in
    let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
                             (PlusSG (ConstSG _ _ c1) (ConstSG _ _ c2))) in
    let e_3' = magma_to_semiGroup p e_3 in
        (_ ** ((PlusSG (PlusSG e_ih1 e_3') e_ih2), ?Massoc1))
[...]
\end{verbatim}
\end{center}
\caption{Computing with associativity in a Semi-Group, first pattern}
\figrule
\end{figure}

Note that we use function of conversions between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$.

After simplifying the constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way. The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described on section 2.3.

\subsection {Normalization of terms in a Monoid}

In a monoid, we simply have to eliminate zeros, thanks to the two properties of left and right neutrality. Because the simplification of additions between a zero and a constant has be done at the underneath level of semigroup (by computing the addition of these two constants), we only have to simplify the addition between the constant zero and a variable. There is in fact two cases, since the variable can come first, or the $Zero$ can come first. We can write a function $elimZero$ which rewrites these two patterns $x+Z$ and $Z+x$ as $x$.

%\begin{figure}[H]
%\figrule
%\begin{center}
%\begin{verbatim}
%elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
%            -> (ExprMo p gam c1) 
%            -> (c2 ** (ExprMo p gam c2, c1=c2))
%elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        with (eq_on_set p Zero const1)
%   elimZero c p gam (PlusMo (ConstMo _ _ Zero) (VarMo _ v)) 
%        | (Just refl) = (_ ** (VarMo _ v, ?MelimZero1))
%   elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
%        | _ = (_ ** (PlusMo (ConstMo _ _ const1) (VarMo _ v), refl)) 
%elimZero c p gam 
% (PlusMo (VarMo _ v) (ConstMo _ _ const2)) 
%      with (eq_as_elem_of_set p Zero const2) 
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ Zero)) | (Just refl) 
%          = (_ ** (VarMo _ _ v, ?MelimZero2))
%   elimZero c p gam (PlusMo (VarMo _ v) 
%            (ConstMo _ _ const2)) | _ 
%          = (_ ** (PlusMo (VarMo _ v) 
%                 (ConstMo _ _ const2), refl))
%[...]
%\end{verbatim}
%\end{center}
%\caption{Eliminating zeros in a Monoid, first pattern}
%\figrule
%\end{figure}

The normalization of reflected expressions in a monoid is simply made of a call to the function reducing terms on a semigroup, followed by a call to elimZero which will eliminates the remaining additions with zeros.

\subsection {Reusing the monoid prover for building the group prover}

The first thing that the reduction function of the group prover will have to do is simply to transform every subtraction $a-b$ into $a + -b$, and we are entitled to do so precisely because we have available the property $Minus\_simpl$ contained in the instance of the type class.
After that, the reduction continues with the propagation of the $Neg$ operation inside the parenthesis and $-(a+b)$ is transformed into $(-b) + (-a)$. Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that we're having a commutative monoid.
Then, because we've pushed some negations inside the parenthesis, we might have sequences of two or more consecutive negations. We simplify them by removing two consecutive negations every time that we find such a sequence. Once this is done, there is a last major step specific of groups to accomplish, which is the simplification of sums of symmetric elements.

This might be a direct sum of symmetric elements, like $-e1 + e2$ or $e1 + -e2$ and in these two cases, we need to simplify them when $e1$ can be reduced\footnote{It is not enough to simply check their syntactical equality, as they might not be immediately equal but could perhaps be normalised to the same term. Thus, this step --which is part of the reduction function-- uses the entire reduction function, but it calls it on a smaller term} to $e2$ , and if so the result should be $Zero$. We are entitled to do this treatment because of the property $Plus\_inverse$.

The sum to simplify might be more complex, with two or three levels of of $Plus$. In these cases, there might be some simplifications to do thanks to the associativity property. With two levels of sums, we will transform $(e1 + (-e2 + e3))$ and $(-e1 + (e2+e3))$ into $e3$ when $e1$ can be normalised to $e2$. Also, $((e1+e2) + -e3))$ and $((e1+(-e2)) + e3)$ are simplified into $e1$ when $e2$ can be reduced to $e3$. Finally, in presence of three level of sums, we will transform $(a+b) + ((-c)+d)$ and $(a+(-b)) + (c+d)$ into $a+d$ when $b$ can be reduced to $c$.

The idea : negations $-x$ are encoded as variables.

\subsection {Reusing the commutative group prover for building the ring prover}

The idea : monomials (containing products), like $3*(x*y)$ are encoded as variables when passed to the commutative group prover.

### 2016-01-07 paper-JFP-DTP/content/5-discussion.tex
Renamed content files

\section {Discussion}

	\subsection{Comparison with a traditional approach}

What we would do without type-safe reflection, and when the normalisation functions are not correct by construction : defining a (weekly typed) normalisation a function, and latter on, a lemma about this normalisation function. Building the tactic with this lemma.
cf extra slides 33-34-35-36 of my STP talk.

	\subsection {Comparison with Coq's approach}

Exactly as above, but let the typechecker do more work.
cf extra slides 37-38 of my STP talk.

	\subsection {Correctness and completeness}

		\subsubsection{Correctness}

Idea : obtained by construction. In fact, what we really needed to produce was the proof of equality, and not the normalised terms. The only two things we have to ensure are :
\begin{itemize}
\item not having introduces any axioms
\item not having introduced any non total function (is it really needed in fact if all this stuff only exists at typecheck and not at runtime? I guess it's fine as long as we haven't used non-total functions to illegitimately extract proof of lemmas)
\end{itemize}

If we meet the two conditions above, we simply can't produce a false proof of a statement, because the typechecker would complain. Like in Coq's approach where they just see if the typecheck is ok to apply $f\_correct\ (reflect\ a)\ (reflect\ b)$ in order to prove $a=b$. The typecheck only agrees if $reify\ (reflect\ a) = reify(reflect\ b)$ can be unified with $a=b$, which is the case if the $reify$ and $reflect$ functions have been correctly defined.

		\subsubsection{Completeness}

Idea 1 : Having a complete tactic is important, because otherwise the tactic is not usable often, and thus is not really valuable. However, having a formal proof of the completeness in the system itself isn't really interesting. As long as we do not meet an equality that should be provable, and which isn't proved by our tactic, there is no problem. And if this situation does happen, nothing really bad really happened : no inconsistencies have been introduced, etc. The tactic just refuses to do it automatically for the programmer. The human can still try to attempt it. \\
\\
Idea 2 : Completeness is not, as one could imagine : \\
$a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$
The reason is that we could just rewrite the hypothesis in the goal, without having to inspect the behaviour of the $norm$ function. That would prove nothing interesting.
Completeness is in fact $norm\ (reflect\ a) \neq norm(reflect\ b) \rightarrow a \neq b$, which is in fact :
$decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \neq b$.
This thing is terrible to prove in the proof assistant. It has never been done for this problem (not for Coq's ring prover, not for Agda ring prover). Can I do it for the little example with natural numbers though ? \\
\\
Idea 3 : We have the guarantee that we aren't losing any totality because of the $reflect$ function because of our index ! This is another extra advantage of our approach. We know that $reify\ (reflect\ a)$ is always unifiable with $a$ in our case. Coq's approach doesn't have this guarantee, and could potentially lose some completeness if this property is not always true.

### 2016-01-08 paper-JFP-DTP/content/5-discussion.tex
paper : traditional approach written

@@
 
-What we would do without type-safe reflection, and when the normalisation functions are not correct by construction : defining a (weekly typed) normalisation a function, and latter on, a lemma about this normalisation function.
-Building the tactic with this lemma.
-cf extra slides 33-34-35-36 of my STP talk.
+Without our type-safe reflection mechanism, the traditional way to go for this problem would be to start with the definition of the function $reify\ :\ Expr\ \rightarrow\ c$.
+As we did in our approach, there is still a function for reflecting terms $reflect\ :\ c\ \rightarrow\ Expr$.
+Because $reflect$ is a function based on syntax,  the actual way to define depends on the facilities of the proof assistant / programming language used, but there need to be a way to write functions based on syntax.
+We need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
+
+Then comes the normalisation function : $normalize\ :\ Expr\ \rightarrow\ Expr$.
+Because this function is weekly typed compared to our approach (we no longer have an index giving a guarantee about the concrete values), we need to prove a lemma of correctness about it after its definition.
+This lemma of correctness must say that the interpretation of the original (reflected) term is equal to the interpretation of the normalised (reflected) term.
+$normalize\_correct\ :\ \forall\ e:Expr,\ reify\ (normalize\ e)\ =\ reify\ e$.
+We have avoided the need of such a --complicated-- proof in our correct by construction approach enabled by our type-safe reflection mechanism.
+
+Now, similarly to what we've done, we need a syntactical equality test, checking whether two reflected expression are syntactically equal.
+The difference here is that, as often with this kind of "traditional" approach, we produce something that belongs to the world of computations, usually a (pretty uninformative) boolean.
+$beq_{Expr}\ :\ Expr\ \rightarrow\ Expr\ \rightarrow\ bool$.
+Because this boolean on it own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation are also (syntactically) equal.
+$beq_{Expr}\_correct\ :\ \forall (e1\ e2:Expr), beq_{Expr}\ e1\ e2\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+We did not have to do this proof either with our approach.
+
+We build now the kernel of the tactics, which is a decision procedure. 
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+decideEq : c -> c -> bool 
+decideEq x y =
+      let xRefl = reflect x in 
+      let yRefl = reflect y in
+      beq_Expr (normalise xRefl) (normalise yRefl)
+\end{verbatim}
+\end{center}
+\caption{Decision procedure}
+\figrule
+\end{figure}
+This function on its own is not going to be enough. We need to add it a correctness theorem, and this is precisely this theorem which is going to help us build the tactic we want.	
+$decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
+This proof can be made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times), and $reflect\_and\_reify\_correct$.
+
+Now, the tactic can be build by using this correctness theorem.
+When trying to prove a goal $a\ =\ b$, the tactic will apply this theorem $decideEq\_correct$.
+The desired proof is produced if the premise $decideEq\ a\ b\ =\ true$ holds.
+But whether this premise holds or not can simply be tested by running the function $decideEq$ on the two arguments $a$ and $b$.
+The activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
 

### 2016-01-08 paper-JFP-DTP/content/5-discussion.tex
paper : Coq approach written

@@
 
-Without our type-safe reflection mechanism, the traditional way to go for this problem would be to start with the definition of the function $reify\ :\ Expr\ \rightarrow\ c$.
+Without our type-safe reflection mechanism, the traditional way to go for this problem would be to have functions doing only computations, and some external lemmas about them.
+It would start with the definition of the function $reify\ :\ Expr\ \rightarrow\ c$.
 As we did in our approach, there is still a function for reflecting terms $reflect\ :\ c\ \rightarrow\ Expr$.
@@
 Then comes the normalisation function : $normalize\ :\ Expr\ \rightarrow\ Expr$.
-Because this function is weekly typed compared to our approach (we no longer have an index giving a guarantee about the concrete values), we need to prove a lemma of correctness about it after its definition.
+Because this function is weekly typed compared to our approach (we no longer have an index giving a guarantee about the concrete values), we need to provide a lemma of correctness about it after its definition.
 This lemma of correctness must say that the interpretation of the original (reflected) term is equal to the interpretation of the normalised (reflected) term.
@@
 
-Now, similarly to what we've done, we need a syntactical equality test, checking whether two reflected expression are syntactically equal.
+Now, similarly to what we've done, we need a syntactical equality test, checking whether two reflected expressions are syntactically equal.
 The difference here is that, as often with this kind of "traditional" approach, we produce something that belongs to the world of computations, usually a (pretty uninformative) boolean.
 $beq_{Expr}\ :\ Expr\ \rightarrow\ Expr\ \rightarrow\ bool$.
-Because this boolean on it own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation are also (syntactically) equal.
+Because this boolean on its own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation are also (syntactically) equal.
 $beq_{Expr}\_correct\ :\ \forall (e1\ e2:Expr), beq_{Expr}\ e1\ e2\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
@@
 
-We build now the kernel of the tactics, which is a decision procedure. 
+We now build the kernel of the tactics, which is the following decision procedure. 
 \begin{figure}[H]
@@
 \end{center}
-\caption{Decision procedure}
+\caption{Decision procedure of the tactic with a "traditional" approach}
 \figrule
 \end{figure}
-This function on its own is not going to be enough. We need to add it a correctness theorem, and this is precisely this theorem which is going to help us build the tactic we want.	
+This function on its own is not going to be enough. We need to add it a correctness theorem, and this is precisely this main theorem which is going to help us build the tactic we want.	
 $decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
-This proof can be made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times), and $reflect\_and\_reify\_correct$.
+This proof can be made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times, one on each side of the assumed equality), and finally $reflect\_and\_reify\_correct$ (again two times).
 
 Now, the tactic can be build by using this correctness theorem.
-When trying to prove a goal $a\ =\ b$, the tactic will apply this theorem $decideEq\_correct$.
-The desired proof is produced if the premise $decideEq\ a\ b\ =\ true$ holds.
+When trying to prove a goal $a\ =\ b$, the tactic will apply this theorem $decideEq\_correct$, which means that $x$ will be unified to $a$ and $y$ to $b$.
+The desired proof is thus produced if the premise $decideEq\ a\ b\ =\ true$ holds.
 But whether this premise holds or not can simply be tested by running the function $decideEq$ on the two arguments $a$ and $b$.
-The activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
+
+Both with this "traditional" approach or with our type-safe reflection, the activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
 
@@
 
-Exactly as above, but let the typechecker do more work.
-cf extra slides 37-38 of my STP talk.
+In Coq's latest implementation of a Ring solver, described in [1], they almost followed what has just been desribed as a "traditional" approach with the use of many auxiliary lemmas.
+The few differences essentially lie in the way the function $reflect$ is defined.
+In Coq, they use the language LTac to program this automatic reflection\footnote{This function --that they call mkPolexpr-- can be found in $plugins/ring/Ring\_tac.v$}.
+Ltac is an untyped tactic language.
+It is untyped in the sense that an Ltac "function" produces something which might not have a valid  --and unique-- type.
+Because of this, Ltac definitions can only be used in the context of goals.
+Thus, applying a tactic defined with LTac might work, and then it makes progress to the current goal, or might fail, and in this case the goal in kept unchanged.
+And anyway, the safety of the proof done is going to be checked during the final QED.
+So, Ltac functions can't be used in the statement of a lemma, and it is therefore not possible to reason about them.
+That means that it is not possible to write the lemma $reflect\_and\_reify\_correct$ of the "traditional" approach described in the previous subsection, because it is even impossible to state it as it uses a function defined in LTac.
+Because they can't even state this property, they can't finish the proof of the main theorem $decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
+Indeed, the last step of this proof was to apply two times --one for the LHS, one for the RHS-- the fact that $\forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
+One possibility would have been to add this axiom, but this is particularly unsightly, and potentially harmful.
+Also, it would imply that anyone using the ring prover would have this axiom added to his development, and would be forced to believe in it.
+Some proofs would be done automatically for the user of the system, but at the overly-expensive price of adding some uncertainty.
+Instead, what they did for Coq's implementation of the Ring prover is to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$ can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$. 
+\itemize
+\item
+If it can unify the goal $a=b$ with $reify\ (reflect\ a)\ =\ reify\ (reflect\ b)$ then it only has to check the validity of the premise by running the decision procedure.
+More precisely, if $beq_{Expr}\ (norm\ (reflect\ a))\ (norm\ (reflect\ b))$ is evaluated to $true$, then it has built the premise of $f\_correct$ and there's nothing left to prove.
+However, if the result was false, then it means that the automatic ring prover hasn't been able to prove this goal, because the LHS and RHS don't reduce to the same thing.
+If this ring prover is complete, it means that $a$ is not\footnote{Note that in this case, it hasn't produced such a formal proof of disequality, which isn't the goal of a ring prover} equal to $b$, or at least, they aren't equal only with the properties of a ring.
+\item
+if it couldn't unify the goal $a=b$ with $reify\ (reflect\ a)\ =\ reify\ (reflect\ b)$, then it means that there is something wrong in the definitions of the functions $reflect$ and $reify$, and it could print something like "The ring prover has failed unexpectedly.
+There's something wrong with the implementations of the $reflect$ and $reify$ functions".
+Nothing really bad happened, as no inconsistent proof has been produced, nor no axioms added.
+That would just mean that the implementation of the ring prover is slightly broken, and that some goals that should be automatically provable aren't automatically proved.
+Thus, it would only decrease the completeness of the prover.
+That would be of course bad --because in the extreme case, the prover never succeed in a generating a proof and is thus completely useless--, but that would only limit the scope of usage of the prover.
 

### 2016-01-10 paper-JFP-DTP/content/5-discussion.tex
paper : correctness and completeness

@@
 $beq_{Expr}\ :\ Expr\ \rightarrow\ Expr\ \rightarrow\ bool$.
-Because this boolean on its own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation are also (syntactically) equal.
+Because this boolean on its own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation should also be (syntactically) equal.
 $beq_{Expr}\_correct\ :\ \forall (e1\ e2:Expr), beq_{Expr}\ e1\ e2\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
@@
 
-Both with this "traditional" approach or with our type-safe reflection, the activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
+Both with this "traditional" approach and with our type-safe reflection, the activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
+The main advantage of our approach compared to this kind of traditional approach is that we don't need external lemmas, and more generally that the proof of correctness is obtained a lot more easily.
+Let's now see the differences more precisely with an actual implementation that almost follows the paradigm of the traditional approach.
 
@@
 Because of this, Ltac definitions can only be used in the context of goals.
-Thus, applying a tactic defined with LTac might work, and then it makes progress to the current goal, or might fail, and in this case the goal in kept unchanged.
+Thus, applying a tactic defined with LTac might work, and then it makes progress to the current goal, or it might fail, and in this case the goal is kept unchanged.
 And anyway, the safety of the proof done is going to be checked during the final QED.
@@
 Some proofs would be done automatically for the user of the system, but at the overly-expensive price of adding some uncertainty.
-Instead, what they did for Coq's implementation of the Ring prover is to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$ can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
-Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$. 
+Instead, what they did for Coq's implementation of the Ring prover is to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$ can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$ to the goal. 
 \itemize
@@
 
-Idea : obtained by construction. In fact, what we really needed to produce was the proof of equality, and not the normalised terms. The only two things we have to ensure are :
+The correctness of the tactic is in fact obtained by construction.
+Actually, what we really needed was to produce the proof of the equality $a=b$, and not the normalised terms.
+The normalised term --and more precisely their construction-- was just a support for building the desired proof of $a=b$, which is precisely the proof of correctness.
+Thus, there isn't any possibility that we've generated a wrong proof as long as we ensure that :
 \begin{itemize}
-\item not having introduces any axioms
-\item not having introduced any non total function (is it really needed in fact if all this stuff only exists at typecheck and not at runtime?
-I guess it's fine as long as we haven't used non-total functions to illegitimately extract proof of lemmas)
+\item We haven't introduced any axioms
+\item We haven't introduced any non total\footnote{Contrary to Coq, Idris allows the definition of non-total functions for helping the development of non-finished real-world projects} function or proof used for the building of the main proof $a=b$.
+Note that the normalization itself could contain some non total functions without breaking the correctness.
+That would just mean that when using this tactic, the tactic could loop infinitely and the user would never get his answer.
+That would of course be bad, but that would not produce any inconsistency --as long as none uses these definitions maliciously to prove something false--, because this non total function would be only used in the world of computations and not in the world of logic.
 \end{itemize}
 
-If we meet the two conditions above, we simply can't produce a false proof of a statement, because the typechecker would complain.
-Like in Coq's approach where they just see if the typecheck is ok to apply $f\_correct\ (reflect\ a)\ (reflect\ b)$ in order to prove $a=b$.
-The typecheck only agrees if $reify\ (reflect\ a) = reify(reflect\ b)$ can be unified with $a=b$, which is the case if the $reify$ and $reflect$ functions have been correctly defined.
+Because we've meet these two conditions, we simply can't produce a false proof of a statement, because the typechecker would complain if that would be the case.
+If Idris internal type tyeory is consistent, then one can't produce a false lemma without either introducing an axiom nor introducing a non total function and latter using it to create inconsistencies in the logic.
 
@@
 
-Idea 1 : Having a complete tactic is important, because otherwise the tactic is not usable often, and thus is not really valuable.
+Completeness means that when given an equality which is provable by using the properties of the concerned algebraic structure, the prover should be able to generate a proof of this equality.
+Having a complete tactic is important, because otherwise the tactic is not always usuable -- and in the worst case it is never usable--, and thus is not valuable.
 However, having a formal proof of the completeness in the system itself isn't really interesting.
 As long as we do not meet an equality that should be provable, and which isn't proved by our tactic, there is no problem.
-And if this situation does happen, nothing really bad really happened : no inconsistencies have been introduced, etc.
+And if this situation does happen, nothing really bad really happened : no inconsistencies have been introduced in the system.
 The tactic just refuses to do it automatically for the programmer.
-The human can still try to attempt it.
+The user can still attempt it.
 \\
-\\
-Idea 2 : Completeness is not, as one could imagine : \\
-$a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$
-The reason is that we could just rewrite the hypothesis in the goal, without having to inspect the behaviour of the $norm$ function. That would prove nothing interesting.
-Completeness is in fact $norm\ (reflect\ a) \neq norm(reflect\ b) \rightarrow a \neq b$, which is in fact :
-$decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \neq b$.
-This thing is terrible to prove in the proof assistant.
-It has never been done for this problem (not for Coq's ring prover, not for Agda ring prover).
-Can I do it for the little example with natural numbers though ?
-\\
-\\
-Idea 3 : We have the guarantee that we aren't losing any totality because of the $reflect$ function because of our index !
-This is another extra advantage of our approach.
-We know that $reify\ (reflect\ a)$ is always unifiable with $a$ in our case.
-Coq's approach doesn't have this guarantee, and could potentially lose some completeness if this property is not always true.
+
+If we believe that the normal form computed by the reduction function is indeed a normal form --which means that if two terms are equal according to the properties of an algebraic structure, then their normal form should be equal-- then the tactic is complete.
+Indeed, 
+$meta\_propetry\ :\ \forall\ a\ b,\ a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is equivalent to
+
+$completeness : \ \forall\ a\ b,\ norm\ (reflect\ a) \neq norm(reflect\ b) \rightarrow a \neq b$, which is in fact :
+$\ \forall\ a\ b,\ decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \neq b$.
+The equivalence between $meta\_property$ and $completeness$ can only be done in the meta theory because it is a proof by contradiction, which is equivalent to the law of excluded middle, that we do not have in constructive logics, on which Idris is based.
+
+Question : Does $completeness$ is provable in the system ?
+Would it be doable for the little example with natural numbers for example ?
+I want to know if it's something improvable in the theory itself (and in this case why, precisely!), or just something really hard ?
+Anyway, it has never been done for this problem (not for Coq's ring prover, nor for Agda's ring prover). Can I do it for the little example with natural numbers though ? \\
+
+Another little advantage of our approach, compared to Coq's approach, is that we have the guarantee that we can't lose any totality because of the $reflect$ function.
+The reason is that the produced reflected term is indexed over the original --concrete-- value.
+By this fact, we don't need a $reify$ function.
+If we want to go from the reflected term to the original --concrete-- value, we can simply return the index.
+Thus, we have the guarantee that $(reflect\ a)$ is always a faithful representation of $a$ with our approach.
+Coq's approach doesn't have this guarantee, and could potentially lose some completeness if this property is not always true, as mentioned earlier during the comparison with Coq's implementation.

### 2016-01-12 paper-JFP-DTP/content/2-basicIdeas.tex
paper continued and corrected. Still roughly 2 pages to write

--- 2016-01-07 paper-JFP-DTP/content/basicIdeas.tex
@@
 
-For an expression $e\ :\ Expr\ \Gamma\ x$, we will say that "$e$ denotes (or encodes) the number $x$ in the context $\Gamma$".
+For an expression $ex\ :\ Expr\ \Gamma\ x$, we will say that "$ex$ denotes (or encodes) the number $x$ in the context $\Gamma$".
 When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
 Also, the $Zero$ expression denotes the natural number $Z$.
-Finally, if $e1$ is an expression encoding the number $x$, and $e2$ is an expression encoding the number $y$, then the expression $Plus\ e1\ e2$ denotes the number $(x + y)$.
+Finally, if $ex$ is an expression encoding the number $x$, and $ey$ is an expression encoding the number $y$, then the expression $Plus\ ex\ ey$ denotes the concrete number $(x + y)$.
 
@@
 \\
-We have an expression $e1$ encoding $x$, and an expression $e2$ encoding $y$\footnote{These encodings have to be produced by hand by the user for this little tool example, but for the real collection of tactics, we will program an automatic reflection mechanism}.
-We will normalize $e1$, and this will give a new number $x'$, a new expression $e1':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
-We will do the same with $e2$, and we will get a new number $y'$, an expression $e2':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
+We have an expression $ex$ encoding $x$, and an expression $ey$ encoding $y$.
+We will normalize $ex$, and this will give a new concrete number $x'$, a new expression $ex':Expr\ \Gamma\ x'$, and a proof of $x=x'$.
+We will do the same with $ey$, and we will get a new concrete number $y'$, an expression $ey':Expr\ \Gamma\ y'$, and a proof of $y=y'$.
 \\
-It is now enough to simply compare $e1'$ and $e2'$ using a standard syntactical equality because these two expressions are in normal form :
+It is now enough to simply compare $ex'$ and $ey'$ using a standard syntactical equality because these two expressions are now supposed to be in normal form :
 
@@
 
-Now, if the two normalised expressions $e1'$ and $e2'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
+Now, if the two normalised expressions $ex'$ and $ey'$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
 By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$.
@@
            -> (x = x') -> (y = y') -> Maybe (x = y)
-buildProof e e' lp rp with (eqExpr e e')
-  buildProof e e lp rp | Just refl = ?MbuildProof
-  buildProof e e' lp rp | Nothing = Nothing
+buildProof ex' ey' lp rp with (eqExpr ex' ey')
+  buildProof ex' ex' lp rp | Just refl = ?MbuildProof
+  buildProof ex' ey' lp rp | Nothing = Nothing
 \end{verbatim}
@@
 
-The $Expr\ \Gamma\ x'$ will be used as the normalised left hand size of the equality, which represents the value $x'$.
+The $Expr\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
 Before the normalisation, the reflected LHS was reflecting the value $x$.
-The $Expr\ \Gamma\ y'$ is the normalised right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation.
+The $Expr\ \Gamma\ y'$ is the normalised reflected right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation.
 The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function also produces the proof of equality between the previous and the new concrete values.
@@
            -> Maybe (x = y)
-  testEq l r = 
-     let (x' ** (l', p1)) = reduce l in 
-     let (y' ** (r', p2)) = reduce r in
-        buildProof l' r' p1 p2 
+  testEq ex ey = 
+     let (x' ** (ex', px)) = reduce ex in 
+     let (y' ** (ey', py)) = reduce ey in
+        buildProof ex' ey' px py 
 \end{verbatim}

### 2016-01-12 paper-JFP-DTP/content/3-generalProblem.tex
paper continued and corrected. Still roughly 2 pages to write

@@
 
-All our tactics will require to have a way of testing the equality between elements of the underlying set, that is to say, a way to test equality between constants.
-For this reason, we define a notion of "set", which only requires this equality.
-All the algebraic structures above will extend this type class :
+In fact, we don't only want to be able to prove equalities, but more generally equivalence.
+We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use.
+This equivalence relation on $c$ has the following profile $(\sim=) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+
+All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants.
+For this reason, we define a notion of "set", which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$.
+All the algebraic structures will later extend this type class :
 
@@
 class Set c where
+    (~=) : (x:c) -> (y:c) -> Type
+
+    refl : (x:c) -> x~=x
+    sym : {x:c} -> {y:c} -> (x~=y) -> (y~=x)
+    trans : {x:c} -> {y:c} -> {z:c} -> (x~=y) -> (y~=z) -> (x~=z)
+
     set_eq : (x:c) -> (y:c) -> Maybe (x=y)
@@
 \end{figure}
+If the user wants to prove propositional equalities, he will simply instantiate $(\sim=)$ with the built-in $(=)$ during the definition of the instance of $Set$.
+Note that $(\sim=)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--.
+That's quite natural, since we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
 
-Note that this decidable equality is weak in the sense that it only produces a proof when the two elements are equal, but it doesn't produce a proof of dis-equality when they are different -- instead, it simply produces the value $Nothing$--.
-That's quite natural, since we want to generate proof of equality, and not to generate counter examples for proving dis-equalities, which is another problem.
+Obviously, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure.
+Therefore, equivalence in a $Set$ are just the "syntactical equivalence", and they can simply be proven with $refl$.
 
-Obvisouly, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure.
-Therefore, equalities in a $Set$ are just the syntactical equality, and they can simply be proven with $Refl$.
-
-The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no axioms about this operation.
+The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no properties about this operation.
 
@@
 
-This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with a $set\_eq$ operation), and if it has a $Plus$ operation.
+This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\sim=$ and the test $set\_eq$), and if it has a $Plus$ operation.
 
@@
 The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
-The reflected term for $x$ will be denoted $e1$, and this term will have the type $ExprG\ p\ \Gamma\ x$.
-The term $e1$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
-We've got similar thing for $y$, which is encoded by $e2$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $e1$ will produce the normal form $e1'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p1$ of $x=x'$ while running the normalisation procedure on $e2$ will produce the normal form $e2'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p2$ of $y=y'$.
+The reflected term for $x$ will be denoted $ex$, and this term will have the type $ExprG\ p\ \Gamma\ x$.
+The term $ex$ is the encoding of $x$ and its type is precisely indexed over the real value $x$, as the reflected terms embed the concrete values.
+We've got similar thing for $y$, which is encoded by $ey$, and its type is indexed over the real value $y$.
+Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x=x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y=y'$.
 
@@
 The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities.
-Without this property, it would become a lot more complicated to decide the equality without bruteforcing a serie of rewriting, that might even not terminate.
+Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
 These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
@@
 
-The first thing this function has to do is to compute the normal form of the two expressions $e1$ and $e2$ respectively encoding $x$ and $y$.
-Then, the only thing remaining to do will be to compare syntactically $e1'$ and $e2'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $p1$ and $p2$.
+The first thing this function has to do is to compute the normal form of the two expressions $ex$ and $ey$ respectively encoding $x$ and $y$.
+Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $px$ and $py$.
 
@@
 \begin{verbatim}
-groupDecideEq p e1 e2 =
-  let (x' ** (e1', p1)) = groupReduce p e1 in
-  let (y' ** (e2', p2)) = groupReduce p e2 in
-	     buildProofGroup p e1' e2' p1 p2
+groupDecideEq p ex ey =
+  let (x' ** (ex', px)) = groupReduce p ex in
+  let (y' ** (ey', py)) = groupReduce p ey in
+	     buildProofGroup p ex' ey' px py
 \end{verbatim}
@@
 
-The syntactical test of equality between $e1'$ and $e2'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$, similarly to what we've done on the smaller example in the previous section.
+The syntactical test of equality between $ex'$ and $ey'$ and the composition of the two proofs is done in the auxiliary function $buildProofGroup$, similarly to what we've done on the smaller example in the previous section.
 
@@
   -> (Maybe (x = y))
-buildProofGroup p e1 e2 p1 p2 with (exprG_eq p _ e1 e2)
-    buildProofGroup p e1 e1 p1 p2 | Just refl = ?MbuildProofGroup
-    buildProofGroup p e1 e2 p1 p2 | Nothing = Nothing
+buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
+    buildProofGroup p ex ex px py | Just refl = ?MbuildProofGroup
+    buildProofGroup p ex ey px py | Nothing = Nothing
 \end{verbatim}

### 2016-01-12 paper-JFP-DTP/content/4-detailsNormalisation.tex
paper continued and corrected. Still roughly 2 pages to write

@@
 
-We now describe briefly how the normalisation functions work for some of the structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures.
+We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures.
+By doing so, we show that the construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
+        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1=c2))
+simpleTreatment p gam e = (new expression, justification)
+\end{verbatim}
+\end{center}
+\caption{General pattern of a function doing a part of the normalisation}
+\figrule
+\end{figure}
 
@@
 In a semigroup, we only have to deal with the property of associativity.
-As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematical way, either left associative, or right associative.
+As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematic way, either left associative, or right associative.
 But this is not the only thing that we have to do.
-If $x$, $y$ and $z$ denote variables, we don't only want to transform $x+((y+4)+(5+z))$ into $(((x+y)+4)+5)+z$ if we've chosen the complete left associative form.
-We want first to do the addition of the constants $4$ and $5$ together (and we have the right to do so because we can rearrange the parenthesis thanks to the associativity, when constants are, like in this case, near to each other).
-Three possible patterns need this treatment : $(x+c1)+(c2+y)$ , $(x + c1) + c2$  and $c1 + (c2 + x)$ when $c1$ and $c2$ denote constants.
-The simplification between the constants $c1$ and $c2$ is denoted here as $[c1+c2]$, and this computation is done by the normalisation function of the underneath $Magma$ level, called $magmaReduce$.
-The first pattern is rewritten into $(x + [c1+c2]) + y$, the second into $x + [c1+c2]$ and the third one into $[c1 + c2] + x$.
+If $x$, $y$ and $z$ denote variables, we don't only want to transform $x+((y+4)+(5+z))$ into $(((x+y)+4)+5)+z$ --we assume we've chosen the complete left associative form--.
+We also want the constants $4$ and $5$ to be simplified together, because they are close to each other (and we have the right to do so because we can rearrange the parenthesis thanks to the associativity), and the result should be $((x+y)+9)+z$ on this example.
+Three possible patterns need this treatment : $(x+const1)+(const2+y)$, $(x + const1) + const2$  and $const1 + (const2 + x)$ where $const1$ and $const2$ denote constants.
+The simplification between the constants $const1$ and $const2$ is denoted here as $[const1+const2]$, and this computation can be done at the underneath $Magma$ level, with its normalisation function called $magmaReduce$.
+The first pattern is rewritten into $(x + [const1+const2]) + y$, the second into $x + [const1+const2]$ and the third one into $[const1 + const2] + x$.
 We show the code for the first pattern.
@@
 assoc p gam (PlusSG 
-             (PlusSG e1 (ConstSG _ _ c1)) 
-             (PlusSG (ConstSG _ _ c2) e2)) =
-    let (r_ih1 ** (e_ih1, p_ih1)) = (assoc p gam e1) in
-    let (r_ih2 ** (e_ih2, p_ih2)) = (assoc p gam e2) in
+             (PlusSG ex (ConstSG _ _ const1)) 
+             (PlusSG (ConstSG _ _ const2) ey)) =
+    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p gam ex) in
+    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p gam ey) in
     let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
-                             (PlusSG (ConstSG _ _ c1) (ConstSG _ _ c2))) in
+                             (PlusSG (ConstSG _ _ const1) (ConstSG _ _ const2))) in
     let e_3' = magma_to_semiGroup p e_3 in
-        (_ ** ((PlusSG (PlusSG e_ih1 e_3') e_ih2), ?Massoc1))
+        (_ ** ((PlusSG (PlusSG e_ihx e_3') e_ihy), ?Massoc1))
 [...]
@@
 \end{figure}
-
-Note that we use function of conversions between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$.
-
-After simplifying the constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way.
-The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described on section 2.3.
+The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values.
+Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here.
+We wanted to show how a prover can be reused for building another prover on an easy example.
+The functions of conversion between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$ are helping for this task.
+\\
+A meta-variable Massoc1 has been introduced. This metavariable corresponds to a proof obligation for the following property :
+$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ ~=\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ ~=\ r\_ihx$, $p\_ihy : y\ ~=\ r\_ihy$ and $p_3 : Plus\ const1\ const2 ~= r\_3$.
+This lemma says that the treatment done by this function $assoc$ on this case was sound because the previous and the new concrete values are equal.
+It can easily be proven with a few rewritings and the use of the associativity, available from the semi-group structure.
+\\
+\\
+After this simplification of constants computable thanks to associativity, the normalization for SemiGroup also has to rearrange the brackets on a systematic way.
+The fully right or fully left associative forms can both be chosen for this purpose, and the implementation is very similar to the function $expr_l$ previously described for the smaller example on section 2.3.
 

### 2016-01-12 paper-JFP-DTP/content/5-discussion.tex
paper continued and corrected. Still roughly 2 pages to write

@@
 It would start with the definition of the function $reify\ :\ Expr\ \rightarrow\ c$.
-As we did in our approach, there is still a function for reflecting terms $reflect\ :\ c\ \rightarrow\ Expr$.
-Because $reflect$ is a function based on syntax,  the actual way to define depends on the facilities of the proof assistant / programming language used, but there need to be a way to write functions based on syntax.
-We need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
+As we did in our approach, there would still be a function for reflecting terms $reflect\ :\ c\ \rightarrow\ Expr$.
+Because $reflect$ is a function based on syntax,  the actual way to define it would depend on the facilities of the proof assistant / programming language used, but there need to be a way to write functions based on syntax.
+We would need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
 
 Then comes the normalisation function : $normalize\ :\ Expr\ \rightarrow\ Expr$.
-Because this function is weekly typed compared to our approach (we no longer have an index giving a guarantee about the concrete values), we need to provide a lemma of correctness about it after its definition.
+Because this function is weekly typed compared to our approach (we no longer have an index giving a guarantee about the concrete values), we would need to provide a lemma of correctness about it after its definition.
 This lemma of correctness must say that the interpretation of the original (reflected) term is equal to the interpretation of the normalised (reflected) term.
@@
 
-Now, similarly to what we've done, we need a syntactical equality test, checking whether two reflected expressions are syntactically equal.
-The difference here is that, as often with this kind of "traditional" approach, we produce something that belongs to the world of computations, usually a (pretty uninformative) boolean.
+Now, similarly to what we've done, we would need a syntactical equality test, checking whether two reflected expressions are syntactically equal.
+The difference here is that, as often with this kind of "traditional" approach, we would produce something that belongs to the world of computations, usually a (pretty uninformative) boolean.
 $beq_{Expr}\ :\ Expr\ \rightarrow\ Expr\ \rightarrow\ bool$.
@@
 
-We now build the kernel of the tactics, which is the following decision procedure. 
+We could now build the kernel of the tactic, which is the following decision procedure. 
 \begin{figure}[H]
@@
 decideEq x y =
-      let xRefl = reflect x in 
-      let yRefl = reflect y in
-      beq_Expr (normalise xRefl) (normalise yRefl)
+      let ex = reflect x in 
+      let ey = reflect y in
+      beq_Expr (normalise ex) (normalise ey)
 \end{verbatim}
@@
 \end{figure}
-This function on its own is not going to be enough. We need to add it a correctness theorem, and this is precisely this main theorem which is going to help us build the tactic we want.	
+This function on its own would not  be enough. We would need to add a correctness theorem, and this is precisely this main theorem that would help us to build the tactic we want.	
 $decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
-This proof can be made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times, one on each side of the assumed equality), and finally $reflect\_and\_reify\_correct$ (again two times).
+This proof can can made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times, one on each side of the assumed equality), and finally $reflect\_and\_reify\_correct$ (again two times).
 
-Now, the tactic can be build by using this correctness theorem.
-When trying to prove a goal $a\ =\ b$, the tactic will apply this theorem $decideEq\_correct$, which means that $x$ will be unified to $a$ and $y$ to $b$.
+Now, the tactic could be built by using this correctness theorem.
+When trying to prove a goal $a\ =\ b$, the tactic would apply this theorem $decideEq\_correct$, which means that $x$ will be unified to $a$ and $y$ to $b$.
 The desired proof is thus produced if the premise $decideEq\ a\ b\ =\ true$ holds.
@@
 	\subsection {Comparison with Coq's approach}
+
+		\subsubsection{Description of Coq's implementation}
 
@@
 Some proofs would be done automatically for the user of the system, but at the overly-expensive price of adding some uncertainty.
-Instead, what they did for Coq's implementation of the Ring prover is to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$ can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+Instead, what they did for Coq's implementation of the Ring prover was to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$-- can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
 Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$ to the goal. 
@@
 Thus, it would only decrease the completeness of the prover.
-That would be of course bad --because in the extreme case, the prover never succeed in a generating a proof and is thus completely useless--, but that would only limit the scope of usage of the prover.
+That would be of course bad --because in the extreme case, the prover never succeeds in a generating a proof and is thus completely useless--, but that would only limit the scope of usage of the prover.
+\\
+
+		\subsubsection{Differences with Coq's implementation}
+
+Coq's ring prover follows the traditional approach of defining functions, and latter on proving many auxiliary lemmas about them.
+This is particularly adapted to Coq, which has many facilities for the construction of proofs, and an adapted proof mode.
+However, this approach kind of duplicate the work : they first tell the machine how it works, and latter, why it works.
+This separation between the world of computations and the world of logic becomes smaller in our approach with a good use of dependent types, that allows to write more specific types, and thus to capture logical properties.
+The writing of functions can therefore be guided by this more precise type, and this is one of the main benefit of the approach we've followed.
+Moreover, we almost get the proof of correctness for free, because every little bit of rewriting done for the normalisation of the reflected terms is accompanied by the logic justification which tells why this rewriting can be done --and locally, this justification is always simple !--.
+In the end, all the proofs we have to do are systematically trivial, since they only contain trivial rewritings with the available hypothesis and the use of the properties of the corresponding algebraic structure.
+\\
+
+Another difference with our implementation is that we've implemented a hierachy of tactics for many algebraic structures, but Coq's implementation only deals with ring and semi ring.
+If someone wants to prove equalities in a commutative group that isn't a ring, he simply can't use their prover.
+A dedicated prover for commutative groups would be needed, and Coq currently doesn't has one.
+The worst is that such a prover for commutative group would do very similar treatments, which means a lot of code could have been factorised.
+This is what we've obtained by taking this in account from the start.
+Our monoid prover used the underneath semi-group prover, our group prover uses the underneath monoid prover, and so on.
+\\
+
+[Mentioning performances, that might be in favour of Coq]
 
@@
 
-Completeness means that when given an equality which is provable by using the properties of the concerned algebraic structure, the prover should be able to generate a proof of this equality.
-Having a complete tactic is important, because otherwise the tactic is not always usuable -- and in the worst case it is never usable--, and thus is not valuable.
+Completeness means that when given an equality which is provable by using the properties of the concerned algebraic structure, the adapted prover should be able to generate a proof of this equality.
+Having a complete tactic is important, because otherwise the tactic is not always usuable -- and in the worst case it is never usable--, and thus is not really valuable.
 However, having a formal proof of the completeness in the system itself isn't really interesting.
@@
 Indeed, 
-$meta\_propetry\ :\ \forall\ a\ b,\ a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is equivalent to
+$meta\_propetry\ :\ \forall\ a\ b,\ a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is logically equivalent to
 

### 2016-01-12 paper-JFP-DTP/paper.tex
paper...
paper JFP DTP continued
Renamed content files
paper continued and corrected. Still roughly 2 pages to write

@@
 With this type-safe reflection technique, the resulting expressions are guaranteed to be faithful representations of the corresponding inputs and any generated proof is guaranteed to be a proof of the required property.
-We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified lists and the associative concatenation operation, and later show how this idea is generalised to various kind of properties (or axioms) that might be available (for example : commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for Monoids, Commutative Monoids, Groups, Commutative Groups, Rings, and so on, written in Idris, for proving different kind of equivalence.
+We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified natural numbers and the associative addition operation, and later show how this idea can be generalised to various kind of properties (or axioms) that might be available (for example : commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for Monoids, Commutative Monoids, Groups, Commutative Groups, Rings, and so on, written in Idris, for proving different kind of equivalence.
 We will also show how each tactic reuses the other ones from the simplest structures, thus avoiding as much as possible the duplication of code.
@@
 
-\input{./content/introduction.tex}
+\input{./content/1-introduction.tex}
 
-\input{./content/basicIdeas.tex}
+\input{./content/2-basicIdeas.tex}
 
-\input{./content/generalProblem.tex}
+\input{./content/3-generalProblem.tex}
 
-\input{./content/automaticReflection.tex}
+\input{./content/4-detailsNormalisation.tex}
 
-\input{./content/discussion.tex}
+\input{./content/5-discussion.tex}
 
@@
    Gregoire,~B. and Mahboubi,~A. (2005) Proving equalities in a commutative ring done right in Coq, TPHOLs 2005, Oxford, UK, August 22-25, 2005, Proceedings
+
+ \bibitem[\protect\citename{W.A Howard}1980]{How80}
+   Howard W.A.
+The formulae-as-types notion of construction.
+In J.
+R.
+Seldin and J.
+P.
+Hindley, editors, To H.
+B.
+Curry: Essays on Combinaory Logic, Lambda Calculus, and Formalism.
+Academic Press, 1980.
 

### 2016-01-13 paper-JFP-DTP/content/1-introduction.tex
paper continued, section 1&2, typos...

@@
 All the equality symbols $=$ we have encountered so far refer to the propositional equality.
-We remind the readers in the next subsection how this equality is defined in Idris, and more generally in Intentional Type Theories, and how it differs from the --more primitive-- jugmental equality.
+We remind the readers in the next subsection how this equality is defined in Idris, and more generally in Intentional Type Theories, and how it differs from the --more primitive-- judgemental equality.
 
-\subsection{The propositional equality and the jugmental equality in Intentional Type Theories}
+\subsection{The propositional equality and the judgemental equality in Intentional Type Theories}
 In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
 Since in type theory, propositions are seen as types \cite{How80}, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
-Thus, if $a$ and $b$ are of type\footnote{It can be useful sometimes to allow the writing of heterogeneous equalities, and in this case, HMeq (which is the standard equality in Idris) can be used.
+Thus, if $a$ and $b$ are of type\footnote{It can be useful sometimes to allow the writing of heterogeneous equalities, and in this case, JMeq (which is the standard equality in Idris) can be used.
 It follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
@@
 The first component is the value $bX$ (expressed as a natural number) represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
-This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing finally the two bits produced and the proof that these two bits are effectively encoding the result of the addition we had to perform.
+This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing finally the two bits produced and a proof that these two bits are effectively encoding the result of the addition we had to perform.
 For example, with the line corresponding to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
@@
 Now, we want to define a function adding two binary numbers and a bit (used as a carry).
-We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit, and this result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
+We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit.
+This result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
 Thus, we're tempted to write :
@@
 Using a provisional definitions here will make the definition being accepted, but it also generates two proof obligations, one for each pattern, and both of them are saying that the provided term can be converted into a term which has the expected type.
-For both of these two patterns, it is possible to prove the equality between the expected and the provided types, because this function is correct.
+For both of these two patterns, it is in fact possible to prove the equality between the expected and the provided types, because this function is correct.
 These proofs are made of a series of rewriting, and each rewriting uses a property about the operation $+$ on natural numbers.
@@
 This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
-This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will requires us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
+This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will require us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
 The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
 
-If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), and the associativity and commutativity of $+$.
-Thus, we're rewriting a term by using the properties of a commutative group.
+If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), the associativity and commutativity of $+$.
+Thus, we're rewriting a term by using the properties of a commutative monoid.
 
-We would like to let a prover for commutative groups find a proof on its own.
-This is something doable because algebraic structures like monoid, group, commutative group and ring have all a very nice property : every element in them can be normalised to a canonical representation.
+We would like to let a prover for commutative monoid find a proof on its own.
+This is something doable because algebraic structures like monoid, monoid, group and ring (all commutative or not) have all a very nice property, which is that every element in them can be normalised to a canonical representation.
 Thus these kind of proofs can be automatically produced by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality), because now that they are in normal form, being equal is just a matter of being syntactically the same entity.
@@
 Algebraic provers, like ring provers, are already implemented for various proof assistants, including Coq and Agda.
-In this paper, we describe a certified implementation of an automatic prover for equalities in various algebraic structures, including Monoid, Groups and Rings (potentially commutative), for the Idris language.
+In this paper, we describe a certified implementation of an automatic prover for equalities in various algebraic structures, including Monoid, Groups and Rings (all potentially commutative), for the Idris language.
 The novelty is in the approach that we follow, using a new kind of type-safe reflection.
 Working by reflection for implementing tactics has been done several times, as in [ref 1] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
-We will compare our approach with other implementation of this kind of tactics in section 6.
+We will compare our approach with other implementation of this kind of tactics in section 5.
 \\
@@
 The first one is that the reflection of a term is guaranteed to be a faithful representation of the term.
-The second is that any generated proof is guaranteed to be a proof of the required property, thus increasing the totality of the prover.
-The basic ideas of the technique are first presented in the section 2 on a smaller problem with only natural numbers and the addition, and are adapted for the hierarchy of tactics in the section 3 and 4.
-Discussion about the totality is provided in section 5.2.
+The second is that any generated proof is guaranteed to be a proof of the required property.
+The basic ideas of the technique are first presented in the section 2 on a smaller problem with only natural numbers and additions, and are later adapted for the hierarchy of tactics in section 3 (and section 4 contains some implementation details).
 	\item 2) The normalisation procedures are implemented by following a correct by construction approach, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
@@
 Again, this approach will be presented in section 2 on the smaller problem, and on sections 3 and 4 for the complete hierarchy of algebraic structures.
-	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each solver : we reuse as much as possible each solver in the implementation of the other solvers for the above structures.
+	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each prover : we reuse as much as possible each prover in the implementation of the other solvers for the above structures.
 For example simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuses it.
-As it will be discussed in section 4, reusability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.
-For example, reusing the monoid solver for the group solver is not trivial, since we lose the possibility to express the negations $-x$ and minus $x-y$ operations.
+As it will be discussed in section 4, re-usability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.
+For example, reusing the monoid solver for building the group solver is not trivial, since we lose the possibility to express negations ($-x$) and minus operations ($x-y$).
 \end{itemize}

### 2016-01-13 paper-JFP-DTP/content/2-basicIdeas.tex
paper continued, section 1&2, typos...

@@
 
-We will first present the basic ideas of our technique on a simplified version, in which we aim to deal with universally quantified natural numbers, the properties of associativity and neutral element for the operation plus.
+We will first present the basic ideas of our technique on a simplified problem, in which we aim to deal with universally quantified natural numbers and additions.
 For example, we would like to be able to automatically generate proofs of goals like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
-For this smaller problem, we have decided to only work with the associativity of plus : $\forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a right neutral element for the $+$ operation : $\forall x, x + Z = x$.
-Note that $Z$ is also a left neutral element, but this property is not needed, as we have this behaviour by reduction as $+$ is defined recursively on its first argument.
-Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of solvers for different algebraic structures.
+For this smaller problem, we have decided to only work with the associativity of $+$ : $plusAssociative : \forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a right neutral element for the $+$ operation : $plusZeroRightNeutral : \forall x, x + Z = x$.
+Note that $Z$ is also a left neutral element, but this property is not needed because we have this behaviour by reduction, as $+$ is defined recursively on its first argument.
+Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of provers for different algebraic structures.
 \\
@@
 This is in fact what we do all the time that we have to decide if two things, written differently, are equal or not.
-When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a canonical representation of polynomials, and to put both ponynomials in this form.
-If the normalised form are the same, then the two original polynomials are equal, otherwise they do not represent the same computation.\\
+When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a canonical representation of polynomials, and to put both polynomials in this form.
+If the normalised forms are the same, then the two original polynomials are equal, otherwise they do not represent the same computation.\\
 \\
@@
 For this little example, the reduction procedure will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with and it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
-As much as possible, we want to avoid the pain of proving the correctness of the tactic, and for this reason, we don't want to have to write huge proofs.
+We want to avoid pain as much as possible, and for this reason, we don't want to have to write huge proofs.
 \\
@@
 The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- numbers are equal.
-Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
+Note that this function can't simply produce an $Expr\ \Gamma\ x$, because the concrete number on which the resulting expression will be indexed is not necessary syntactically equal to the original number since this equality can use the two available properties.
 Said differently, even if we can prove $x=x'$ (if the function is correctly defined), we do not have $x \equiv x'$.
-And in fact, what really interest us in this function is precisely the proof of $x\ =\ x'$.
-The reason is that when we try to automatically prove $x=y$, these proofs $x=x'$ and $y=y'$ will be the crucial part for the construction of the desired proof. \\
+And in fact, what really interests us in this function is precisely this proof of $x\ =\ x'$.
+The reason is that when we try to automatically prove $x=y$, these proofs will be the crucial part for the construction of the desired proof. \\
 \\
+The tactic will work as follow.
 We have an expression $ex$ encoding $x$, and an expression $ey$ encoding $y$.
@@
 eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
-  eqExpr (Plus x y) (Plus x y) | (Just refl, Just refl) = Just refl
+  eqExpr (Plus x y) (Plus x y) | (Just Refl, Just Refl) = Just Refl
   eqExpr (Plus x y) (Plus x' y') | _ = Nothing
 eqExpr (Var i) (Var j) with (decEq i j)
-  eqExpr (Var i) (Var i) | (Yes refl) = Just refl
+  eqExpr (Var i) (Var i) | (Yes Refl) = Just Refl
   eqExpr (Var i) (Var j) | _ = Nothing
-eqExpr Zero Zero = Just refl
+eqExpr Zero Zero = Just Refl
 eqExpr _ _ = Nothing
@@
 buildProof ex' ey' lp rp with (eqExpr ex' ey')
-  buildProof ex' ex' lp rp | Just refl = ?MbuildProof
+  buildProof ex' ex' lp rp | Just Refl = ?MbuildProof
   buildProof ex' ey' lp rp | Nothing = Nothing
@@
 
-The $Expr\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
+The argument of type $Expr\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
 Before the normalisation, the reflected LHS was reflecting the value $x$.
 The $Expr\ \Gamma\ y'$ is the normalised reflected right hand size, which now represents the value $y'$, but which was encoding $y$ before the normalisation.
-The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function also produces the proof of equality between the previous and the new concrete values.
+The function also expects the proofs that $x=x'$ and $y=y'$, and we should be able to pass them because the normalisation function has also produced the proof of equality between the original and the new concrete values.
 
-As we mentioned, the proof for the metavariable $MbuildProof$ is just a rewriting of the two equalities :
+As mentioned, the proof for the metavariable $MbuildProof$ is just a rewriting of the two equalities :
 
@@
   MbuildProof = proof {
-  intros; refine Just; rewrite sym p1; rewrite sym p2; exact refl;
+  intros; refine Just; rewrite sym p1; rewrite sym p2; exact Refl;
 }  
@@
 
-Finally, the main function which tries to prove the equality $x=y$ simply has to reduce the two metaified terms reflecting the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs we just obtained :
+Finally, the main function which tries to prove the equality $x=y$ simply has to reduce the two reflected terms encoding the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs that we just obtained :
 
@@
 
-Now, we need to define the function reduce.
+We now need to define the function reduce.
 To do that, we have to decide a canonical representation of associative natural numbers.
 We decide that the left associative form will be the canonical representation.
-Thus, the $reduce$ function has to rewrite the reflected term by rearranging the parentheses in order to transform the underlying number in the form $(...((x1 + x2) + x3) ...
+Thus, the $reduce$ function has to rewrite the reflected terms by rearranging the parentheses in order to transform the underlying concrete number in the form $(...((x1 + x2) + x3) ...
 + xn)$.
@@
 As part of the normalization, we write a function $expr\_l$ which converts an $Expr\ \Gamma\ x$ to a $LExpr\ \Gamma\ x'$ and which produces a proof of $x=x'$.
-This function will therefore use the two available properties multiple times --and especially the property of associativity--, in order to obtain the expected fully left associative form.
+This function will therefore use the two available properties multiple times, while rewriting the term until the fully left associative desired form is obtained.
 
@@
          -> (x' ** (LExpr gam x', x = x'))
-expr_l Zero = (_ ** (LZero, refl))
-expr_l (Var i) = (_ ** (LPlus LZero i, refl))
+expr_l Zero = (_ ** (LZero, Refl))
+expr_l (Var i) = (_ ** (LPlus LZero i, Refl))
 expr_l (Plus ex ey) = 
-  let (xl ** (xr, xprf)) = expr_l ex in
-  let (yl ** (yr, yprf)) = expr_l ey in
-    plusLExpr _ _ xr yr xprf yprf
+  let (x' ** (ex', px)) = expr_l ex in
+  let (y' ** (ey', py)) = expr_l ey in
+  let (res ** (normRes, Pres)) = plusLExpr ex' ey' in
+    (res ** (normRes, rewrite px in (rewrite py in Pres)))
       where 
-      plusLExpr : (x', y' : Nat)
-            -> {gam : Vect n Nat} -> {x, y : Nat} 
+      plusLExpr : {gam : Vect n Nat} -> {x, y : Nat} 
             -> LExpr gam x -> LExpr gam y  
-            -> (x' = x) -> (y' = y) 
-            -> (w' ** (LExpr gam w', x'+y'=w'))
-      plusLExpr x' y' rx (LPlus e i) xprf yprf =
-        let (xrec ** (rec, prf)) = 
-          plusLExpr _ _ rx e refl refl in
-          (_ ** (LPlus rec i, ?plusLExpr1))
-      plusLExpr x' y' rx LZero xprf yprf =
-        (_ ** (rx, ?plusLExpr2))
+            -> (z ** (LExpr gam z, x+y=z))
+      plusLExpr {x=x} ex LZero =
+        (_ ** (ex, rewrite (plusZeroRightNeutral x) in Refl))            
+      plusLExpr ex (LPlus e i) =
+        let (xRec ** (rec, prfRec)) = plusLExpr ex e in
+            (_ ** (LPlus rec i, ?MplusLExpr))
+
 \end{verbatim}
@@
 \end{figure}
+In the case of an addition $Plus\ ex\ ey$, the function $expr\_l$ does the job of normalisation recursively on $ex$ and on $ey$, and then it uses the sub-function $plusLExpr$ to normalise the addition of these two --already normalised-- terms.
+This sub-function $plusLExpr$ has two kind of simplifications to do.
+When the second argument is an $LZero$, it simply returns its first arguments along with the justification for this rewriting, which obviously uses $plusZeroRightNeutral$.
+However, when the second argument is an $LPlus\ e\ i$, it continues recursively by computing $plusLExpr\ ex\ e$, and it finally adds $i$ to it.
+That had for effect to move the parenthesis on the left, and this treatment is going to be justified during the proof of the meta variable $MplusLExpr$ by the use of $plusAssociative$.
 
+This metavariable $MplusLExpr$ requires us to prove the goal : $x1 + (x2 + index\ i\ \Gamma) = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things, $prfRec : x1 + x2 = xrec$.
+By using the property of associativity on the goal, we now need to prove $(x1 + x2) + index\ i\ \Gamma$, which can be done by rewriting the proof $prfRec$ obtained recursively.
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{verbatim}
+MplusLExpr = proof {
+  intros
+  rewrite (sym (plusAssociative x1 x2 (index i gam))); rewrite prf; 
+  exact Refl;
+}
+\end{verbatim}
+\end{center}
+\caption{Proof of the metavariable plusLExpr1}
+\figrule
+\end{figure}
+
+It is really important to understand that the root of the automatic building of the desired proof of $x=y$ happens precisely in these nested use of $plusZeroRightNeutral$ and $plusAssociative$ that we've seen in the definition of $expr\_l$ and of its meta-variable $MplusLExpr$.
+These proofs replace the arithmetical proofs that we were doing previously by hand, like in section 1 with the lemma $adc\_lemma\_2$.
+\\
+\\
 Using this new datatype $LExpr$ has changed the representation of our encoded lists, so we need to convert back an $LExpr\ \Gamma\ x$ to an $Expr\ \Gamma\ x$.
@@
 
-We notice that in order to transform the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these different possibilities.
+We notice that in order to transform the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these two possibilities.
 \\
 \\
-We've got two metavariables to prove.
-The metavariable $plusLExpr1$ requires us to prove the goal : $x' + y' = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things,  $(xprf\ :\ x'\ =\ x)$, $(yprf\ :\ y'\ =\ x1\ +\ index\ i\ \Gamma)$ and $(prf\ :\ x\ +\ x1\ =\ xrec)$.
-Proving this goal uses the property of associativity after rewriting the goal with these three proof of equality $xprf$, $yprf$ and $prf$.
-
-\begin{figure}[H]
-\figrule
-\begin{center}
-\begin{verbatim}
-plusLExpr1 = proof {
-  intros; rewrite sym xprf; rewrite sym yprf; rewrite prf;
-  mrefine plusAssociative;
-}
-\end{verbatim}
-\end{center}
-\caption{Proof of the metavariable plusLExpr1}
-\figrule
-\end{figure}
-
-And the metavariable plusLExpr2 uses the fact that $Z$ is a neutral element for the addition.
-
-\begin{figure}[H]
-\figrule
-\begin{center}
-\begin{verbatim}
-plusLExpr2 = proof
-  intros; rewrite xprf; rewrite (sym yprf); mrefine plusZeroRightNeutral
-\end{verbatim}
-\end{center}
-\caption{Proof of the metavariable plusLExpr2}
-\figrule
-\end{figure}
 

### 2016-01-13 paper-JFP-DTP/paper.tex
paper continued, section 1&2, typos...

@@
 With this type-safe reflection technique, the resulting expressions are guaranteed to be faithful representations of the corresponding inputs and any generated proof is guaranteed to be a proof of the required property.
-We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified natural numbers and the associative addition operation, and later show how this idea can be generalised to various kind of properties (or axioms) that might be available (for example : commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for Monoids, Commutative Monoids, Groups, Commutative Groups, Rings, and so on, written in Idris, for proving different kind of equivalence.
+We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified natural numbers and the associative addition operation, and will later show how this idea can be generalised to various kind of properties that might be available (for example, commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for Monoids, Commutative Monoids, Groups, Commutative Groups, Rings, and so on, written in Idris, for proving different kind of equivalence.
 We will also show how each tactic reuses the other ones from the simplest structures, thus avoiding as much as possible the duplication of code.

### 2016-01-13 paper-JFP-DTP/content/2-basicIdeas.tex
Added the source file for the example with Nat

@@
 \x => \y => \z => Just (replace (sym (replace (sym (replace 
-(plusZeroRightNeutral x) (plusAssociative x 0 y))) (replace 
-(sym (replace (plusZeroRightNeutral x) (plusAssociative x 0 z))) (replace 
-(replace (plusZeroRightNeutral (plus xy)) (plusAssociative (plus x y) 0 x)) 
-[...]
+(sym (plusAssociative x 0 y)) (replace (replace (sym 
+(plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
+(replace (sym (plusAssociative x 0 z)) (replace (replace 
+(sym (plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
+(plusAssociative (x+y) x z)) [...]
 : (x : Nat) -> (y : Nat) -> (z : Nat) 
   -> Maybe ((x + y) + x + z 
-            = x+ (y + x) + z)
+            = x + (y + x) + z)
 \end{verbatim}
@@
 And we effectively get the proof of equality we wanted.
-As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $plus$ ($plusZeroRightNeutral$).
+As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $+$ ($plusZeroRightNeutral$).
 
@@
 We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively.
+We want to program an automatic way of going from the concrete values (of type $Nat$), to the reflected terms (of type $Expr$).
+The only way to do that is to inspect the abstract syntax tree of the concrete value.
 By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values.

### 2016-01-13 paper-JFP-DTP/content/5-discussion.tex
subtitle, Alternative approach

@@
 
-	\subsection{Comparison with a traditional approach}
+	\subsection{An alternative approach, more traditional}
 
@@
 As we did in our approach, there would still be a function for reflecting terms $reflect\ :\ c\ \rightarrow\ Expr$.
-Because $reflect$ is a function based on syntax,  the actual way to define it would depend on the facilities of the proof assistant / programming language used, but there need to be a way to write functions based on syntax.
+Because $reflect$ is a function based on syntax, it needs to use Idris' reflection mechanism.
 We would need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.

### 2016-01-15 paper-JFP-DTP/content/1-introduction.tex
Added the source file for the example with Nat
paper : use of listings and corrections

@@
 Proving this kind of auxiliary lemmas is time consuming, is much too specific for being reusable, and isn't the interesting part of formal certification.
-Therefore, we would like to be able to automatize a far as possible the constructions of proofs for this kind of lemmas for the Idris language, a dependently typed language which doesn't contain many proof automations at the moment.
-
-A common situation in formal certification is the following.
-We have a function $f\ :\ A \rightarrow B$, and two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B.
-If this semantic says everything we care about for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
-All the equality symbols $=$ we have encountered so far refer to the propositional equality.
+Therefore, we would like to be able to automatize a far as possible the constructions of this kind of proofs for the Idris language, a dependently typed language which doesn't contain many proof automations at the moment.
+\\
+\\
+When doing formal certification, many functions have a lemma of correctness that talks about equality of terms.
+More precisely, the following situation is very familiar.
+The function that we wants to certify is $f\ :\ A \rightarrow B$.
+We have two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B.
+If this semantic says everything that matters for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
+The interesting thing is that, very often, the type $SEM$ has the structure of some nice algebraic structure, and it becomes thus possible to decide equalities in it automatically.
+All the equality symbols we have encountered so far refer to the propositional equality.
 We remind the readers in the next subsection how this equality is defined in Idris, and more generally in Intentional Type Theories, and how it differs from the --more primitive-- judgemental equality.
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 Id (A:Type) : A -> A -> Type :=
   Refl : forall a:A, Id a a
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 data Bit : Nat -> Type where
@@
      (#) : Binary w v -> Bit bit -> Binary (S w) (bit + 2 * v)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 addBit : Bit x -> Bit y -> Bit c -> (bX ** (bY ** 
@@
 addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, _)))
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 adc : Binary w x -> Binary w y -> Bit c -> Binary (S w) (c + x + y)
@@
           adc numx numy carry0 # lsb
-\end{verbatim}
+\end{lstlisting}
 \end{center}
-\caption{Addition of two binary numbers}
+\caption{Addition of two binary numbers and a carry}
 \figrule
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 adc_lemma_2 = proof {
@@
     trivial;
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 Working by reflection for implementing tactics has been done several times, as in [ref 1] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
-We will compare our approach with other implementation of this kind of tactics in section 5.
+We will compare our approach with other implementations of this kind of tactics in section 5.
 \\

### 2016-01-15 paper-JFP-DTP/content/2-basicIdeas.tex
paper : use of listings and corrections

@@
 \begin{center}
-\begin{verbatim}
-data Expr : (gam : Vect n Nat) -> Type where
-     Plus : {n:Nat} -> {gam:Vect n Nat} -> 
-            Expr gam -> Expr gam -> Expr gam
-     Var  : {n:Nat} -> {gam:Vect n Nat} -> 
-            (i : Fin n) -> Expr gam
-     Zero : {n:Nat} -> (gam:Vect n Nat) -> 
-            Expr gam Z
-\end{verbatim}
+\begin{lstlisting}
+data Expr : ($\Gamma$ : Vect n Nat) -> Type where
+     Plus : {n:Nat} -> {$\Gamma$:Vect n Nat} -> 
+            Expr $\Gamma$ -> Expr $\Gamma$ -> Expr $\Gamma$
+     Var  : {n:Nat} -> {$\Gamma$:Vect n Nat} -> 
+            (i : Fin n) -> Expr $\Gamma$
+     Zero : {n:Nat} -> ($\Gamma$:Vect n Nat) -> 
+            Expr $\Gamma$ Z
+\end{lstlisting}
 \end{center}
@@
 
-Variables are represent using a De Brujin-like index : (Var fZ) denotes a variable, (Var (fS fZ)) another one, and so on.
+Variables are represent using a De Brujin-like index : (Var FZ) denotes a variable, (Var (FS FZ)) another one, and so on.
 
@@
 \begin{center}
-\begin{verbatim}
-Plus (Plus (Var gam fZ) 
-           (Var gam (fS fZ))) 
-     (Plus (Var gam (fS (fS fZ))) 
-           (Var gam (fS (fS (fS fZ)))))
-\end{verbatim}
+\begin{lstlisting}
+e1 : (x, y, z : Nat) -> 
+           Expr [x, y, z] ((x + y) + (x + z))
+e1 x y z = Plus (Plus (Var FZ) (Var (FS FZ))) 
+                (Plus (Var FZ) (Var (FS (FS FZ))))
+\end{lstlisting}
 \end{center}
-\caption{Reflected LHS of example 1}
+\caption{Reflected LHS of example 1 (with the first version of reflected natural numbers)}
 \figrule
@@
 \begin{center}
-\begin{verbatim}
-using (x : Nat, y : Nat, gam : Vect n Nat)
+\begin{lstlisting}
+using (x : Nat, y : Nat, $\Gamma$ : Vect n Nat)
   data Expr : (Vect n Nat) -> Nat -> Type where
-       Plus : Expr gam x -> Expr gam y -> 
-              Expr gam (x + y)
-       Var  : (i : Fin n) -> Expr gam (index i gam)
-       Zero : Expr gam Z
-\end{verbatim}
+       Plus : Expr $\Gamma$ x -> Expr $\Gamma$ y -> 
+              Expr $\Gamma$ (x + y)
+       Var  : (i : Fin n) -> Expr $\Gamma$ (index i $\Gamma$)
+       Zero : Expr $\Gamma$ Z
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-eqExpr : (e : Expr gam x) -> (e' : Expr gam y) -> Maybe (e = e')
+\begin{lstlisting}
+eqExpr : (e : Expr $\Gamma$ x) -> (e' : Expr $\Gamma$ y) -> Maybe (e = e')
 eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
@@
 eqExpr _ _ = Nothing
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-buildProof : {x : Nat} -> {y : Nat} -> Expr gam x' -> Expr gam y' 
+\begin{lstlisting}
+buildProof : {x : Nat} -> {y : Nat} -> Expr $\Gamma$ x' -> Expr $\Gamma$ y' 
            -> (x = x') -> (y = y') -> Maybe (x = y)
@@
   buildProof ex' ey' lp rp | Nothing = Nothing
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
   MbuildProof = proof {
@@
 }  
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-  testEq : Expr gam x -> Expr gam y 
+\begin{lstlisting}
+  testEq : Expr $\Gamma$ x -> Expr $\Gamma$ y 
            -> Maybe (x = y)
@@
         buildProof ex' ey' px py 
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-data LExpr : (gam : Vect n Nat) -> Nat -> Type where
-     LPlus : LExpr gam x -> (i : Fin n) 
-             -> LExpr gam (x + index i gam)
-     LZero : LExpr gam Z
-\end{verbatim}
+\begin{lstlisting}
+data LExpr : ($\Gamma$ : Vect n Nat) -> Nat -> Type where
+     LPlus : LExpr $\Gamma$ x -> (i : Fin n) 
+             -> LExpr $\Gamma$ (x + index i $\Gamma$)
+     LZero : LExpr $\Gamma$ Z
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-expr_l : Expr gam x 
-         -> (x' ** (LExpr gam x', x = x'))
+\begin{lstlisting}
+expr_l : Expr $\Gamma$ x 
+         -> (x' ** (LExpr $\Gamma$ x', x = x'))
 expr_l Zero = (_ ** (LZero, Refl))
@@
       where 
-      plusLExpr : {gam : Vect n Nat} -> {x, y : Nat} 
-            -> LExpr gam x -> LExpr gam y  
-            -> (z ** (LExpr gam z, x+y=z))
+      plusLExpr : {$\Gamma$ : Vect n Nat} -> {x, y : Nat} 
+            -> LExpr $\Gamma$ x -> LExpr $\Gamma$ y  
+            -> (z ** (LExpr $\Gamma$ z, x+y=z))
       plusLExpr {x=x} ex LZero =
@@
 
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 MplusLExpr = proof {
   intros
-  rewrite (sym (plusAssociative x1 x2 (index i gam))); rewrite prf; 
+  rewrite (sym (plusAssociative x1 x2 (index i $\Gamma$))); rewrite prf; 
   exact Refl;
 }
-\end{verbatim}
+\end{lstlisting}
 \end{center}
-\caption{Proof of the metavariable plusLExpr1}
+\caption{Proof of the metavariable MplusLExpr}
 \figrule
@@
 \begin{center}
-\begin{verbatim}
-l_expr : LExpr gam x -> Expr gam x
+\begin{lstlisting}
+l_expr : LExpr $\Gamma$ x -> Expr $\Gamma$ x
 l_expr LZero = Zero
 l_expr (LPlus x i) = Plus (l_expr x) (Var i)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-  reduce : Expr gam x -> (x' ** (Expr gam x', x = x'))
+\begin{lstlisting}
+  reduce : Expr $\Gamma$ x -> (x' ** (Expr $\Gamma$ x', x = x'))
   reduce e = 
@@
          (x' ** (l_expr e', prf))
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 e1 : (x, y, z : Nat) 
     -> Expr [x, y, z] ((x+y) + (x+z))
-e1 x y z = Plus (Plus (Var fZ) 
-                      (Var (fS fZ))) 
-                (Plus (Var fZ) 
-                      (Var (fS (fS fZ))))
+e1 x y z = Plus (Plus (Var FZ) 
+                      (Var (FS FZ))) 
+                (Plus (Var FZ) 
+                      (Var (FS (FS FZ))))
 
@@
      -> Expr [x, y, z] (x + ((y + x) + z))
-e2 x y z = Plus (Var fZ) 
-                (Plus (Plus (Var (fS fZ)) 
-                            (Var fZ)) 
-                      (Var (fS (fS fZ))))
-\end{verbatim}
+e2 x y z = Plus (Var FZ) 
+                (Plus (Plus (Var (FS FZ)) 
+                            (Var FZ)) 
+                      (Var (FS (FS FZ))))
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 e1_e2_testEq : (x, y, z : Nat) 
@@
 e1_e2_testEq x y z = testEq (e1 x y z) (e2 x y z)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 #\x => \y => \z => e1_e2_testEq x y z
@@
             = x + (y + x) + z)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 By using Idris's reflection mechanism, we can tag a function with the keyword "\%reflection", which means that this function runs on syntax instead of values.
-That implies that when applying this function to some arguments, the arguments aren't reduce before the evaluation of this function.
 
@@
 \begin{center}
-\begin{verbatim}
-
+\begin{lstlisting} [mathescape]
 %reflection
-reflectList : (G : List Nat) ->
-          (x : Nat) -> (G' ** Expr (G ++ G') x)
-reflectList G Z = ([] ** Zero)
-
-...
-
-\end{verbatim}
+total
+reflectNat : {n:Nat} -> ($\Gamma$ : Vect n Nat) -> (x:Nat) -> (m ** ($\Gamma'$ : Vect m Nat ** (Expr ($\Gamma$ ++ ') x)))
+\end{lstlisting}
 \end{center}
-\caption{Reflection for natural numbers}
+\caption{Type of the function for reflecting natural numbers}
 \figrule
 \end{figure}
+
+This function takes

### 2016-01-15 paper-JFP-DTP/content/3-generalProblem.tex
paper : use of listings and corrections

@@
 We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use.
-This equivalence relation on $c$ has the following profile $(\sim=) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+This equivalence relation on $c$ has the following profile $(\simeq) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
 
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 class Set c where
-    (~=) : (x:c) -> (y:c) -> Type
+    ($\simeq$) : (x:c) -> (y:c) -> Type
 
-    refl : (x:c) -> x~=x
-    sym : {x:c} -> {y:c} -> (x~=y) -> (y~=x)
-    trans : {x:c} -> {y:c} -> {z:c} -> (x~=y) -> (y~=z) -> (x~=z)
+    refl : (x:c) -> x $\simeq$ x
+    sym : {x:c} -> {y:c} -> (x $\simeq$ y) -> (y $\simeq$ x)
+    trans : {x:c} -> {y:c} -> {z:c} -> (x $\simeq$ y) -> (y $\simeq$ z) -> (x $\simeq$ z)
 
-    set_eq : (x:c) -> (y:c) -> Maybe (x=y)
-\end{verbatim}
+    set_eq : (x:c) -> (y:c) -> Maybe (x $\simeq$ y)
+\end{lstlisting}
 \end{center}
@@
 \end{figure}
-If the user wants to prove propositional equalities, he will simply instantiate $(\sim=)$ with the built-in $(=)$ during the definition of the instance of $Set$.
-Note that $(\sim=)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--.
+If the user wants to prove propositional equalities, he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the instance of $Set$.
+Note that $(\simeq)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--.
 That's quite natural, since we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 class Set c => Magma c where
     Plus : c -> c -> c
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 
-This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\sim=$ and the test $set\_eq$), and if it has a $Plus$ operation.
+This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.
 
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 class Magma c => SemiGroup c where
     Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
-               -> (Plus (Plus c1 c2) c3 = Plus c1 (Plus c2 c3))
-\end{verbatim}
+               -> (Plus (Plus c1 c2) c3 $\simeq$ Plus c1 (Plus c2 c3))
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 class SemiGroup c => Monoid c where
     Zero : c    
-    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 = c1)    
-    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero = c1)
-\end{verbatim}
+    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 $\simeq$ c1)    
+    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero $\simeq$ c1)
+\end{lstlisting}
 \end{center}
@@
 \end{figure}
-
-A Group is a Monoid, with two new operations.
+We can continue the hierarchy with groups, which are a more interesting.
+A group is a monoid, with two new operations.
 The binary operation $Minus$, and the unary operation $Neg$.
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 -- This is just a conjunctive predicate
 hasSymmetric : (c:Type) -> (p:Monoid c) -> c -> c -> Type
-hasSymmetric c p a b = (Plus a b = Zero, Plus b a = Zero)
+hasSymmetric c p a b $\simeq$ (Plus a b $\simeq$ Zero, Plus b a $\simeq$ Zero)
 
@@
     Neg : c -> c
-    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 = Plus c1 (Neg c2) 
+    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
     Plus_inverse : (c1:c) -> hasSymmetric c _ c1 (Neg c1)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 data ExprMa : Magma c -> (Vect n c) -> c -> Type where
-    ConstMa : (p : Magma c) -> (gam:Vect n c) -> (c1:c)  -> ExprMa p gam c1 
-    PlusMa : {p : Magma c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
-         -> ExprMa p gam c1 -> ExprMa p gam c2 
-         -> ExprMa p gam (Plus c1 c2) 
-    VarMa : (p:Magma c) -> {gam:Vect n c}
-         -> (i:Fin n) -> ExprMa p gam (index i gam)
-\end{verbatim}
+    ConstMa : (p : Magma c) -> ($\Gamma$:Vect n c) -> (c1:c)  -> ExprMa p $\Gamma$ c1 
+    PlusMa : {p : Magma c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
+         -> ExprMa p $\Gamma$ c1 -> ExprMa p $\Gamma$ c2 
+         -> ExprMa p $\Gamma$ (Plus c1 c2) 
+    VarMa : (p:Magma c) -> {$\Gamma$:Vect n c}
+         -> (i:Fin n) -> ExprMa p $\Gamma$ (index i $\Gamma$)
+\end{lstlisting}
 \end{center}
@@
 %                    ->Maybe(c1=c2))
-%          -> (gam:Vect n c) -> (i:Fin n) 
-%          -> Variable c_equal gam (index i gam) 
+%          -> ($\Gamma$:Vect n c) -> (i:Fin n) 
+%          -> Variable c_equal $\Gamma$ (index i $\Gamma$) 
 %\end{code}	
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 data ExprG :  Group c -> (Vect n c) -> c -> Type where
-    ConstG : (p : dataTypes.Group c) -> (gam:Vect n c) 
-      -> (c1:c) -> ExprG p gam c1
-    PlusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
-      -> ExprG p gam c1 -> ExprG p gam c2 
-      -> ExprG p gam (Plus c1 c2)
-    MinusG : {p : Group c} -> {gam:Vect n c} -> {c1:c} -> {c2:c} 
-      -> ExprG p gam c1 -> ExprG p gam c2 
-      -> ExprG p gam (Minus c1 c2)
-    NegG : {p : Group c} -> {gam:Vect n c} -> {c1:c} 
-      -> ExprG p gam c1 -> ExprG p gam (Neg c1)
-    VarG : (p : Group c) -> {gam:Vect n c} 
-      -> (i:Fin n) -> ExprG p gam c1
-\end{verbatim}
+    ConstG : (p : dataTypes.Group c) -> ($\Gamma$:Vect n c) 
+         -> (c1:c) -> ExprG p $\Gamma$ c1
+    PlusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
+         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
+         -> ExprG p $\Gamma$ (Plus c1 c2)
+    MinusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
+         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
+         -> ExprG p $\Gamma$ (Minus c1 c2)
+    NegG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} 
+         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ (Neg c1)
+    VarG : (p : Group c) -> {$\Gamma$:Vect n c} 
+         -> (i:Fin n) -> ExprG p $\Gamma$ c1
+\end{lstlisting}
 \end{center}
@@
 	\subsection {A bit of notation}
-The equality we are trying to prove is $x=y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...).
+The equality we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a set with some properties (which make $c$ being a Semi-Group, or a Monoid, or a Group...).
 The fact that $c$ fulfils the specification of an algebraic structure is expressed as an instance of the corresponding type class, and this instance will be denoted as $p$.
@@
 We've got similar thing for $y$, which is encoded by $ey$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x=x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y=y'$.
+Running the normalisation procedure on $ex$ will produce the normal form $ex'$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $px$ of $x \simeq x'$ while running the normalisation procedure on $ey$ will produce the normal form $ey'$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $py$ of $y \simeq y'$.
 
-	\subsection {Deciding equality}
+	\subsection {Deciding equivalence}
 
@@
 \begin{center}
-\begin{verbatim}
-	groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {gam:Vect n c} 
-	  -> {x:c} -> (ExprG p gam x) -> (x' ** (ExprG p gam x', x=x'))
-\end{verbatim}
+\begin{lstlisting}
+groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {$\Gamma$:Vect n c} 
+	   -> {x:c} -> (ExprG p $\Gamma$ x) -> (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-	groupDecideEq : (p:Group c) -> {gam:Vect n c} -> {x : c} -> {y : c} 
-	   -> (ExprG p gam x) -> (ExprG p gam y) 
-	   -> Maybe (x = y)
-\end{verbatim}
+\begin{lstlisting}
+groupDecideEq : (p:Group c) -> {$\Gamma$:Vect n c} -> {x : c} -> {y : c} 
+	     -> (ExprG p $\Gamma$ x) -> (ExprG p $\Gamma$ y) 
+	     -> Maybe (x $\simeq$ y)
+\end{lstlisting}
 \end{center}
@@
 The first thing this function has to do is to compute the normal form of the two expressions $ex$ and $ey$ respectively encoding $x$ and $y$.
-Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x=y$ by using the two proofs $px$ and $py$.
+Then, the only thing remaining to do will be to compare syntactically $ex'$ and $ey'$, and if they are equal, then we will get $x'=y'$, and we will be able to produce the desired proof of $x \simeq y$ by using the two proofs $px : x \simeq x'$ and $py : y \simeq y'$.
 
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 groupDecideEq p ex ey =
@@
 	     buildProofGroup p ex' ey' px py
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
-buildProofGroup : (p:Group c) -> {gam:Vect n c} 
+\begin{lstlisting}
+buildProofGroup : (p:Group c) -> {$\Gamma$:Vect n c} 
   -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
-  -> (ExprG p gam x') -> (ExprG p gam y') 
-  -> (x = x') -> (y = y') 
-  -> (Maybe (x = y))
+  -> (ExprG p $\Gamma$ x') -> (ExprG p $\Gamma$ y') 
+  -> (x $\simeq$ x') -> (y $\simeq$ y') 
+  -> (Maybe (x $\simeq$ y))
 buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
@@
     buildProofGroup p ex ey px py | Nothing = Nothing
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \end{figure}
-
-The proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) in the smallest example showed in section 2.3.
+We've used here the function $exprG\_eq$ to decide the syntactical equality between the reflected terms.
+This function plays the same role as $eqExpr$, presented on section 2.3 for the smaller example.
+Also, the proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) for the smaller example.
 
 \subsection {Automatic reflection}
-$reflect : \{n:Nat\} \rightarrow (gamma : Vect\ n\ c) \rightarrow (c1:c) \rightarrow Expr\ gamma\ c1$.
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+reflect : 
+\end{lstlisting}
+\end{center}
+\caption{Type of the reflecting function for ...}
+\figrule
+\end{figure}
+
 The reflected term is guaranteed to be a faithful representation of the input $c1$.

### 2016-01-15 paper-JFP-DTP/content/4-detailsNormalisation.tex
paper : use of listings and corrections

@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
-        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1=c2))
+        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1 $\simeq$ c2))
 simpleTreatment p gam e = (new expression, justification)
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 \begin{center}
-\begin{verbatim}
+\begin{lstlisting}
 assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
-        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1=c2))
+        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1 $\simeq$ c2))
 assoc p gam (PlusSG 
@@
 [...]
-\end{verbatim}
+\end{lstlisting}
 \end{center}
@@
 The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values.
-Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here.
-We wanted to show how a prover can be reused for building another prover on an easy example.
+Note that the simplification $[const1+const2]$ is done by calling the magma prover, which gives a new expression $e3$, even thought it could perfectly be done directly here, but we wanted to show how a prover can be reused for building another prover on an easy example.
 The functions of conversion between the different levels : $semiGroup\_to\_magma$ and $magma\_to\_semiGroup$ are helping for this task.
@@
 A meta-variable Massoc1 has been introduced. This metavariable corresponds to a proof obligation for the following property :
-$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ ~=\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ ~=\ r\_ihx$, $p\_ihy : y\ ~=\ r\_ihy$ and $p_3 : Plus\ const1\ const2 ~= r\_3$.
-This lemma says that the treatment done by this function $assoc$ on this case was sound because the previous and the new concrete values are equal.
+$Massoc1 : Plus\ (Plus\ x\ const1)\ (Plus\ const2\ y)\ \simeq\ Plus\ (Plus\ r\_ihx\ r_3)\ r\_ihy$ in a context where there is, amongst other things, $p\_ihx : x\ \simeq\ r\_ihx$, $p\_ihy : y\ \simeq\ r\_ihy$ and $p_3 : Plus\ const1\ const2 \simeq r\_3$.
+This lemma says that the treatment done by this function $assoc$ on this pattern was sound because the previous and the new concrete values are equal.
 It can easily be proven with a few rewritings and the use of the associativity, available from the semi-group structure.
@@
 %\begin{center}
-%\begin{verbatim}
+%\begin{lstlisting}
 %elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
 %            -> (ExprMo p gam c1) 
-%            -> (c2 ** (ExprMo p gam c2, c1=c2))
+%            -> (c2 ** (ExprMo p gam c2, c1 $\simeq$ c2))
 %elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
@@
 %[...]
-%\end{verbatim}
+%\end{lstlisting}
 %\end{center}

### 2016-01-15 paper-JFP-DTP/content/5-discussion.tex
paper : use of listings and corrections

@@
 Because $reflect$ is a function based on syntax, it needs to use Idris' reflection mechanism.
-We would need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
+We would need to prove the correctness of these two functions, and it turns out that one lemma is enough for completely specifying the expected behaviour of these two functions at the same time : $reflect\_and\_reify\_correct\ :\ \forall\ x:c,\ reify\ (reflect\ x)\  \simeq \ x$.
 
@@
 This lemma of correctness must say that the interpretation of the original (reflected) term is equal to the interpretation of the normalised (reflected) term.
-$normalize\_correct\ :\ \forall\ e:Expr,\ reify\ (normalize\ e)\ =\ reify\ e$.
+$normalize\_correct\ :\ \forall\ e:Expr,\ reify\ (normalize\ e)\  \simeq\ reify\ e$.
 We have avoided the need of such a --complicated-- proof in our correct by construction approach enabled by our type-safe reflection mechanism.
@@
 Because this boolean on its own is pretty uninformative, we now need a proof of correctness for this function as well, which says that if this function decides that two given terms are syntactically equal, then their interpretation should also be (syntactically) equal.
-$beq_{Expr}\_correct\ :\ \forall (e1\ e2:Expr), beq_{Expr}\ e1\ e2\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+$beq_{Expr}\_correct\ :\ \forall (e1\ e2:Expr), beq_{Expr}\ e1\ e2\ =\ true\ \rightarrow\ reify\ e1\  \simeq\ reify\ e2$.
 We did not have to do this proof either with our approach.
@@
 This function on its own would not  be enough. We would need to add a correctness theorem, and this is precisely this main theorem that would help us to build the tactic we want.	
-$decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
+$decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\  \simeq\ y$.
 This proof can can made by combining, in this order, the three lemmas $beq_{Expr}\_correct$, $normalize\_correct$ (two times, one on each side of the assumed equality), and finally $reflect\_and\_reify\_correct$ (again two times).
@@
 Now, the tactic could be built by using this correctness theorem.
-When trying to prove a goal $a\ =\ b$, the tactic would apply this theorem $decideEq\_correct$, which means that $x$ will be unified to $a$ and $y$ to $b$.
+When trying to prove a goal $a\ \simeq \ b$, the tactic would apply this theorem $decideEq\_correct$, which means that $x$ will be unified to $a$ and $y$ to $b$.
 The desired proof is thus produced if the premise $decideEq\ a\ b\ =\ true$ holds.
@@
 That means that it is not possible to write the lemma $reflect\_and\_reify\_correct$ of the "traditional" approach described in the previous subsection, because it is even impossible to state it as it uses a function defined in LTac.
-Because they can't even state this property, they can't finish the proof of the main theorem $decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$.
-Indeed, the last step of this proof was to apply two times --one for the LHS, one for the RHS-- the fact that $\forall\ x:c,\ reify\ (reflect\ x)\ =\ x$.
+Because they can't even state this property, they can't finish the proof of the main theorem $decideEq\_correct\ :\ \forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x \simeq\ y$.
+Indeed, the last step of this proof was to apply two times --one for the LHS, one for the RHS-- the fact that $\forall\ x:c,\ reify\ (reflect\ x)\ \simeq\ x$.
 One possibility would have been to add this axiom, but this is particularly unsightly, and potentially harmful.
@@
 Some proofs would be done automatically for the user of the system, but at the overly-expensive price of adding some uncertainty.
-Instead, what they did for Coq's implementation of the Ring prover was to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ =\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$-- can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ =\ reify\ e2$.
+Instead, what they did for Coq's implementation of the Ring prover was to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ \simeq\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$-- can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ \simeq\ reify\ e2$.
 Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$ to the goal. 
@@
 \item
-If it can unify the goal $a=b$ with $reify\ (reflect\ a)\ =\ reify\ (reflect\ b)$ then it only has to check the validity of the premise by running the decision procedure.
+If it can unify the goal $a \simeq b$ with $reify\ (reflect\ a)\ \simeq\ reify\ (reflect\ b)$ then it only has to check the validity of the premise by running the decision procedure.
 More precisely, if $beq_{Expr}\ (norm\ (reflect\ a))\ (norm\ (reflect\ b))$ is evaluated to $true$, then it has built the premise of $f\_correct$ and there's nothing left to prove.
@@
 \item
-if it couldn't unify the goal $a=b$ with $reify\ (reflect\ a)\ =\ reify\ (reflect\ b)$, then it means that there is something wrong in the definitions of the functions $reflect$ and $reify$, and it could print something like "The ring prover has failed unexpectedly.
+if it couldn't unify the goal $a \simeq b$ with $reify\ (reflect\ a)\ \simeq\ reify\ (reflect\ b)$, then it means that there is something wrong in the definitions of the functions $reflect$ and $reify$, and it could print something like "The ring prover has failed unexpectedly.
 There's something wrong with the implementations of the $reflect$ and $reify$ functions".
@@
 The correctness of the tactic is in fact obtained by construction.
-Actually, what we really needed was to produce the proof of the equality $a=b$, and not the normalised terms.
-The normalised term --and more precisely their construction-- was just a support for building the desired proof of $a=b$, which is precisely the proof of correctness.
+Actually, what we really needed was to produce the proof of the equality $a \simeq b$, and not the normalised terms.
+The normalised term --and more precisely their construction-- was just a support for building the desired proof of $a \simeq b$, which is precisely the proof of correctness.
 Thus, there isn't any possibility that we've generated a wrong proof as long as we ensure that :
@@
 \item We haven't introduced any axioms
-\item We haven't introduced any non total\footnote{Contrary to Coq, Idris allows the definition of non-total functions for helping the development of non-finished real-world projects} function or proof used for the building of the main proof $a=b$.
+\item We haven't introduced any non total\footnote{Contrary to Coq, Idris allows the definition of non-total functions for helping the development of non-finished real-world projects} function or proof used for the building of the main proof $a \simeq b$.
 Note that the normalization itself could contain some non total functions without breaking the correctness.
@@
 Because we've meet these two conditions, we simply can't produce a false proof of a statement, because the typechecker would complain if that would be the case.
-If Idris internal type tyeory is consistent, then one can't produce a false lemma without either introducing an axiom nor introducing a non total function and latter using it to create inconsistencies in the logic.
+If Idris internal type theory is consistent, then one can't produce a false lemma without either introducing an axiom nor introducing a non total function and latter using it to create inconsistencies in the logic.
 
@@
 Indeed, 
-$meta\_propetry\ :\ \forall\ a\ b,\ a=b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is logically equivalent to
+$meta\_propetry\ :\ \forall\ a\ b,\ a \simeq b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is logically equivalent to
 
-$completeness : \ \forall\ a\ b,\ norm\ (reflect\ a) \neq norm(reflect\ b) \rightarrow a \neq b$, which is in fact :
-$\ \forall\ a\ b,\ decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \neq b$.
+$completeness : \ \forall\ a\ b,\ norm\ (reflect\ a) \neq norm(reflect\ b) \rightarrow a \not{\simeq} b$, which is in fact :
+$\ \forall\ a\ b,\ decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \not{\simeq} b$.
 The equivalence between $meta\_property$ and $completeness$ can only be done in the meta theory because it is a proof by contradiction, which is equivalent to the law of excluded middle, that we do not have in constructive logics, on which Idris is based.

### 2016-01-15 paper-JFP-DTP/content/2-basicIdeas.tex
Fix2 test expression e1 with the use of 1rst datatype
paper : various fix

@@
 e1 : (x, y, z : Nat) -> 
-           Expr [x, y, z] ((x + y) + (x + z))
+           Expr [x, y, z]
 e1 x y z = Plus (Plus (Var FZ) (Var (FS FZ))) 
@@
   intros
-  rewrite (sym (plusAssociative x1 x2 (index i $\Gamma$))); rewrite prf; 
+  rewrite (sym (plusAssociative x1 x2 (index i $\Gamma$))); rewrite prfRec; 
   exact Refl;
@@
 
-The numbers denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this by using $testEq$.
+The numbers denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this fact by using $testEq$.
 
@@
 \end{center}
-\caption{Test of equality betwen e1 and e2}
+\caption{Generating automatically the desired proof}
 \figrule

### 2016-01-15 paper-JFP-DTP/content/3-generalProblem.tex
paper : various fix

@@
 Now that we have introduced our ideas on this little example, it is time to apply them for our main goal, more general : the implementation of a hierarchy of tactics proving equalities in algebraic structures.
-Very often, the properties available on a given type are the one of a well known algebraic structure : magma, semi-group, monoid, group, commutative-group...
-We will no longer only work with the properties of associativity and of right neutral element for natural numbers as we did in the previous section, but we will have available the properties of different structures, and these tactics will usable for any type that satisfies these properties.
+Very often, the properties available on a given type are the ones of a well known algebraic structure, like semi-group, monoid, group, ring...
+We will no longer only work with the properties of associativity and of right neutral element for natural numbers as it was the case in the previous section, but we will have available the properties of different structures, and these tactics will be usable for any type that satisfies these properties.
 
 We will construct a hierarchy of type classes and will write one tactic for each of these type class.
-All our tactics (the group prover, the monoid prover...) will be able to work on any type, being given that an instance of the corresponding type class is provided.
+All our tactics (the ring prover, the group prover...) will be able to work on any type, being given that an instance of the corresponding type class is provided.
 
@@
 In fact, we don't only want to be able to prove equalities, but more generally equivalence.
-We want to allow the user the possibility to define his own equivalence notion that interests him, assuming he's able to provide the properties that belongs to the algebraic structure that he wants to use.
-This equivalence relation on $c$ has the following profile $(\simeq) : (x:c) \rightarrow (y:c) \rightarrow Type$\footnote{This Type would be a Prop in system like Coq that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+We want to give the user the possibility to define his own equivalence notion, assuming he's able to provide the properties of the structure he wants to use.
+This equivalence relation on $c$ has the following profile $(\simeq) : c \rightarrow y \rightarrow Type$\footnote{This 'Type' would be a 'Prop' in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
 
@@
 class Set c where
-    ($\simeq$) : (x:c) -> (y:c) -> Type
+    ($\simeq$) : c -> c -> Type
 
@@
 \end{figure}
-If the user wants to prove propositional equalities, he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the instance of $Set$.
+If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the instance of $Set$.
 Note that $(\simeq)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--.
@@
 
-Obviously, there will be no tactic associated to $Set$, since we have no operations and no properties associated to this structure.
+Obviously, there is no tactic associated to $Set$, since we have no operations and no properties associated to this structure.
 Therefore, equivalence in a $Set$ are just the "syntactical equivalence", and they can simply be proven with $refl$.
-
-The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation which let us computing constants, and has no properties about this operation.
+The first real structure, almost trivial, is the magma. A magma has just a $Plus$ operation, and no properties about it.
 
@@
 This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.
+In fact, all operations, like the Plus here, need to be "compatible" with the equality : $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow (Plus\ c1\ c2) \simeq (Plus\ c1'\ c2')$.
+This need only comes from the fact that we're dealing with any equivalence relation
 

### 2016-01-15 paper-JFP-DTP/content/3-generalProblem.tex
paper fix

@@
   -> (Maybe (x $\simeq$ y))
-buildProofGroup p ex ey px py with (exprG_eq p _ ex ey)
-    buildProofGroup p ex ex px py | Just refl = ?MbuildProofGroup
-    buildProofGroup p ex ey px py | Nothing = Nothing
+buildProofGroup p ex' ey' px py with (exprG_eq p ex' ey')
+    buildProofGroup p ex' ex' px py | Just refl = ?MbuildProofGroup
+    buildProofGroup p ex' ey' px py | Nothing = Nothing
 \end{lstlisting}
@@
 \end{figure}
-We've used here the function $exprG\_eq$ to decide the syntactical equality between the reflected terms.
+We've used here the function $exprG\_eq$ to decide the syntactical equality between the normalised reflected terms $ex'$ and $ey'$.
 This function plays the same role as $eqExpr$, presented on section 2.3 for the smaller example.
-Also, the proof for the metavariable $MbuildProofGroup$ is exactly the same as the corresponding one (called $MbuildProof$) for the smaller example.
+Also, the proof for the metavariable $MbuildProofGroup$ corresponds exactly tto what was done for the metariable $MbuildProof$ in the smaller example.
 

### 2016-01-15 paper-JFP-DTP/content/3-generalProblem.tex
Added new section
paper fix

@@
 This function plays the same role as $eqExpr$, presented on section 2.3 for the smaller example.
-Also, the proof for the metavariable $MbuildProofGroup$ corresponds exactly tto what was done for the metariable $MbuildProof$ in the smaller example.
+Also, the proof for the metavariable $MbuildProofGroup$ corresponds exactly to what was done for the metariable $MbuildProof$ in the smaller example.
 
@@
 (unlike Coq's $Ltac$).
+
+TO DO : Body and description of this function

### 2016-01-17 paper-JFP-DTP/content/1-introduction.tex
Added new section
various fix

@@
 Currently, in the Idris programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
-These proofs are the every day routine when we are working in a dependently typed language, as Idris, and especially when we're intensively indexing types over values in order to capture certain logical properties.
+These proofs are the every day routine when working in a dependently typed language, as Idris, and especially when intensively indexing types over values in order to capture certain logical properties.
 For example, If a type $T$ is indexed over natural numbers, then the definition of a function which has co-domain $T$ will often require to prove that the index of the produced term is equal to the expected index.
@@
 This type of binary numbers\footnote{This representation allows to represent a binary number of width zero with the constructor $zero$, but instead we could have decided to replace this constructor with a constructor taking one bit as parameter and producing a binary number of width one.
-The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value represented by $bin$ ($v$) plus the value represented by the added digit ($bit$).
+The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value ($v$) represented by $bin$ plus the value ($bit$) represented by the added digit.
 
-Now, we're tempted to write the addition between two binary numbers.
+We want to write a function performing the addition between two binary numbers.
 We start with an auxiliary function which adds three bits (the third one will latter play the role of the carry), and produces the two bits of the result : the first one is the most significant bit and the second one is the least significant bit .
@@
 
-We note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : the two bits produced effectively represent the addition of the three bits given in input.
+It should be pointed that the proof component of this function is established without any effort, as we always give the "\_" symbol for the proof which means "infer it!".
+For each of the eight patterns, a simple computation provides the desired proof, and thus the typechecker can deal with them.
+We also note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : the two bits produced effectively represent the addition of the three bits given in input.
 
-Now, we want to define a function adding two binary numbers and a bit (used as a carry).
-We decide that this addition will only work for two binary numbers expressed with the same number of bits, and we produce a result with one more bit.
+We finally define the function adding two binary numbers and a bit (used as a carry).
+We decide that this addition will only work for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
 This result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
@@
 Using a provisional definitions here will make the definition being accepted, but it also generates two proof obligations, one for each pattern, and both of them are saying that the provided term can be converted into a term which has the expected type.
-For both of these two patterns, it is in fact possible to prove the equality between the expected and the provided types, because this function is correct.
+For both of these two patterns, it is in fact possible to prove the equality between the expected and the provided types, precisely because this function is correct.
 These proofs are made of a series of rewriting, and each rewriting uses a property about the operation $+$ on natural numbers.
@@
 This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will require us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
-The re-usability of this kind of proofs is indeed very low, since they are performing rewritings for a very specific term.
+The re-usability of this kind of proofs is therefore null, because they are performing rewritings for very specific terms.
 
@@
 We would like to let a prover for commutative monoid find a proof on its own.
-This is something doable because algebraic structures like monoid, monoid, group and ring (all commutative or not) have all a very nice property, which is that every element in them can be normalised to a canonical representation.
-Thus these kind of proofs can be automatically produced by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality (Leibniz equality), because now that they are in normal form, being equal is just a matter of being syntactically the same entity.
+This is something doable because algebraic structures like monoids, groups and rings (all commutative or not) have all a very nice property, which is that every element in them can be normalised to a canonical representation.
+Thus these kind of proofs can be automatically produced by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality --Leibniz equality--, because when they will be in normal form, being equal will just be a matter of being syntactically the same entity.
 
@@
 Working by reflection for implementing tactics has been done several times, as in [ref 1] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
-We will compare our approach with other implementations of this kind of tactics in section 5.
+We will compare our approach with other implementations of this kind of tactics in section 6.
 \\
@@
 Again, this approach will be presented in section 2 on the smaller problem, and on sections 3 and 4 for the complete hierarchy of algebraic structures.
-	\item 3) We develop a hierarchy of tactics which enables to increase a lot the reusability of each prover : we reuse as much as possible each prover in the implementation of the other solvers for the above structures.
-For example simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuses it.
-As it will be discussed in section 4, re-usability can be quite tricky to get when we want to reuse a solver from a structure above which is less expressive.
-For example, reusing the monoid solver for building the group solver is not trivial, since we lose the possibility to express negations ($-x$) and minus operations ($x-y$).
+	\item 3) We develop a hierarchy of tactics where each tactic reuses the one from the immediately above and simplest structure.
+For example, simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuses it.
+Re-usability can be quite tricky to obtain when we want to reuse a prover from a structure which is less expressive.
+For example, reusing the monoid prover for building the group prover is not trivial, since we lose the possibility to express negations ($-x$) and minus operations ($x-y$).
+Some specific encodings will be presented in section 4 for solving this problem.
 \end{itemize}

### 2016-01-17 paper-JFP-DTP/content/2-basicIdeas.tex
Added new section
various fix

@@
 Note that $Z$ is also a left neutral element, but this property is not needed because we have this behaviour by reduction, as $+$ is defined recursively on its first argument.
-Of course, some structures can have more or less properties than these two, and this work will be extended in the next sections when we will write a general hierarchy of provers for different algebraic structures.
+Of course, some datatypes can have more or less properties than these two, and this work will be extended in sections 3 and 4 where we will present a very general hierarchy of provers for multiple algebraic structures.
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce a proof of this equality if possible, when "equal" has the meaning "syntactically equal or equal thanks to the associative and neutral properties".
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce a proof of this equality if appropriate, where "equal" has the meaning "syntactically equal or equal thanks to the associativity and right neutral properties".
 
@@
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-On the example 1, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
-The general idea --that will also apply for the more general and bigger problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
+On the example 1 just above, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
+The general idea --that will also apply for the more general problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
 The goal of the normalization is to compute a canonical representation for a number $x$, such that any other number provably equal to $x$ (by using the two available properties) will have the same canonical representation.
-For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normal form will be completely left associative, and every addition between an element $a$ and zero should be simplified to $a$.
-It will then be possible to decide the equality by simply comparing the normalised left and right hand sides with a simple syntactical equality.
-This is in fact what we do all the time that we have to decide if two things, written differently, are equal or not.
+For example, the normalisation might transform $x+((y+Z)+z)$ into $(x+y)+z$ if we decide that the normal form will be completely left associative, and that every addition between an element $a$ and zero should be simplified to $a$.
+It will then be possible to decide the equality by simply comparing the normalised left and right hand sides with a simple and strict syntactical equality.
+This is in fact what we do all the time when we have to decide if two things, written differently, are equal or not.
 When a human is given two mathematical polynomials and has to decide the equality of these functions, a technique that always works is to decide once and for all a canonical representation of polynomials, and to put both polynomials in this form.
@@
 total
-reflectNat : {n:Nat} -> ($\Gamma$ : Vect n Nat) -> (x:Nat) -> (m ** ($\Gamma'$ : Vect m Nat ** (Expr ($\Gamma$ ++ ') x)))
+reflectNat : {n:Nat} -> ($\Gamma$ : Vect n Nat) -> (x:Nat) 
+                   -> (m ** ($\Gamma'$ : Vect m Nat ** (Expr ($\Gamma$ ++ ') x)))
 \end{lstlisting}
@@
 
-This function takes
+This function takes ...
+TO DO : body of this function and brief description

### 2016-01-17 paper-JFP-DTP/content/1-introduction.tex
writing the part about the auto reflection for Nat

@@
 \begin{lstlisting}
-Id (A:Type) : A -> A -> Type :=
+Id (A:Type) : A $\rightarrow$ A $\rightarrow$ Type :=
   Refl : forall a:A, Id a a
@@
 \begin{lstlisting}
-data Bit : Nat -> Type where
+data Bit : Nat $\rightarrow$ Type where
      b0 : Bit Z
@@
 
-data Binary : (width : Nat) -> (value : Nat) -> Type where
+data Binary : (width : Nat) $\rightarrow$ (value : Nat) $\rightarrow$ Type where
      zero : Binary Z Z
-     (#) : Binary w v -> Bit bit -> Binary (S w) (bit + 2 * v)
+     (#) : Binary w v $\rightarrow$ Bit bit $\rightarrow$ Binary (S w) (bit + 2 * v)
 \end{lstlisting}
@@
 \begin{lstlisting}
-addBit : Bit x -> Bit y -> Bit c -> (bX ** (bY ** 
+addBit : Bit x $\rightarrow$ Bit y $\rightarrow$ Bit c $\rightarrow$ (bX ** (bY ** 
               (Bit bX, Bit bY, c + x + y = bY + 2 * bX)))
@@
 \begin{lstlisting}
-adc : Binary w x -> Binary w y -> Bit c -> Binary (S w) (c + x + y)
+adc : Binary w x $\rightarrow$ Binary w y $\rightarrow$ Bit c $\rightarrow$ Binary (S w) (c + x + y)
 adc zero zero carry = zero # carry

### 2016-01-17 paper-JFP-DTP/content/2-basicIdeas.tex
writing the part about the auto reflection for Nat

@@
 \begin{lstlisting}
-data Expr : ($\Gamma$ : Vect n Nat) -> Type where
-     Plus : {n:Nat} -> {$\Gamma$:Vect n Nat} -> 
-            Expr $\Gamma$ -> Expr $\Gamma$ -> Expr $\Gamma$
-     Var  : {n:Nat} -> {$\Gamma$:Vect n Nat} -> 
-            (i : Fin n) -> Expr $\Gamma$
-     Zero : {n:Nat} -> ($\Gamma$:Vect n Nat) -> 
+data Expr : ($\Gamma$ : Vect n Nat) $\rightarrow$ Type where
+     Plus : {n:Nat} $\rightarrow$ {$\Gamma$:Vect n Nat} $\rightarrow$ 
+            Expr $\Gamma$ $\rightarrow$ Expr $\Gamma$ $\rightarrow$ Expr $\Gamma$
+     Var  : {n:Nat} $\rightarrow$ {$\Gamma$:Vect n Nat} $\rightarrow$ 
+            (i : Fin n) $\rightarrow$ Expr $\Gamma$
+     Zero : {n:Nat} $\rightarrow$ ($\Gamma$:Vect n Nat) $\rightarrow$ 
             Expr $\Gamma$ Z
@@
 \begin{lstlisting}
-e1 : (x, y, z : Nat) -> 
+e1 : (x, y, z : Nat) $\rightarrow$ 
            Expr [x, y, z]
@@
 using (x : Nat, y : Nat, $\Gamma$ : Vect n Nat)
-  data Expr : (Vect n Nat) -> Nat -> Type where
-       Plus : Expr $\Gamma$ x -> Expr $\Gamma$ y -> 
+  data Expr : (Vect n Nat) $\rightarrow$ Nat $\rightarrow$ Type where
+       Plus : Expr $\Gamma$ x $\rightarrow$ Expr $\Gamma$ y $\rightarrow$ 
               Expr $\Gamma$ (x + y)
-       Var  : (i : Fin n) -> Expr $\Gamma$ (index i $\Gamma$)
+       Var  : (i : Fin n) $\rightarrow$ Expr $\Gamma$ (index i $\Gamma$)
        Zero : Expr $\Gamma$ Z
@@
 \begin{lstlisting}
-eqExpr : (e : Expr $\Gamma$ x) -> (e' : Expr $\Gamma$ y) -> Maybe (e = e')
+eqExpr : (e : Expr $\Gamma$ x) $\rightarrow$ (e' : Expr $\Gamma$ y) $\rightarrow$ Maybe (e = e')
 eqExpr (Plus x y) (Plus x' y') with (eqExpr x x', eqExpr y y')
@@
 \begin{lstlisting}
-buildProof : {x : Nat} -> {y : Nat} -> Expr $\Gamma$ x' -> Expr $\Gamma$ y' 
-           -> (x = x') -> (y = y') -> Maybe (x = y)
+buildProof : {x : Nat} $\rightarrow$ {y : Nat} $\rightarrow$ Expr $\Gamma$ x' $\rightarrow$ Expr $\Gamma$ y' 
+           $\rightarrow$ (x = x') $\rightarrow$ (y = y') $\rightarrow$ Maybe (x = y)
 buildProof ex' ey' lp rp with (eqExpr ex' ey')
@@
 \begin{lstlisting}
-  testEq : Expr $\Gamma$ x -> Expr $\Gamma$ y 
-           -> Maybe (x = y)
+  testEq : Expr $\Gamma$ x $\rightarrow$ Expr $\Gamma$ y $\rightarrow$ Maybe (x = y)
   testEq ex ey = 
@@
 \begin{lstlisting}
-data LExpr : ($\Gamma$ : Vect n Nat) -> Nat -> Type where
-     LPlus : LExpr $\Gamma$ x -> (i : Fin n) 
-             -> LExpr $\Gamma$ (x + index i $\Gamma$)
+data LExpr : ($\Gamma$ : Vect n Nat) $\rightarrow$ Nat $\rightarrow$ Type where
+     LPlus : LExpr $\Gamma$ x $\rightarrow$ (i : Fin n) 
+             $\rightarrow$ LExpr $\Gamma$ (x + index i $\Gamma$)
      LZero : LExpr $\Gamma$ Z
@@
 expr_l : Expr $\Gamma$ x 
-         -> (x' ** (LExpr $\Gamma$ x', x = x'))
+         $\rightarrow$ (x' ** (LExpr $\Gamma$ x', x = x'))
 expr_l Zero = (_ ** (LZero, Refl))
@@
       where 
-      plusLExpr : {$\Gamma$ : Vect n Nat} -> {x, y : Nat} 
-            -> LExpr $\Gamma$ x -> LExpr $\Gamma$ y  
-            -> (z ** (LExpr $\Gamma$ z, x+y=z))
+      plusLExpr : {$\Gamma$ : Vect n Nat} $\rightarrow$ {x, y : Nat} 
+            $\rightarrow$ LExpr $\Gamma$ x -> LExpr $\Gamma$ y  
+            $\rightarrow$ (z ** (LExpr $\Gamma$ z, x+y=z))
       plusLExpr {x=x} ex LZero =
@@
 \begin{lstlisting}
-l_expr : LExpr $\Gamma$ x -> Expr $\Gamma$ x
+l_expr : LExpr $\Gamma$ x $\rightarrow$ Expr $\Gamma$ x
 l_expr LZero = Zero
@@
 \begin{lstlisting}
-  reduce : Expr $\Gamma$ x -> (x' ** (Expr $\Gamma$ x', x = x'))
+  reduce : Expr $\Gamma$ x $\rightarrow$ (x' ** (Expr $\Gamma$ x', x = x'))
   reduce e = 
@@
 e1 : (x, y, z : Nat) 
-    -> Expr [x, y, z] ((x+y) + (x+z))
+    $\rightarrow$ Expr [x, y, z] ((x+y) + (x+z))
 e1 x y z = Plus (Plus (Var FZ) 
@@
 e2 : (x, y, z : Nat) 
-     -> Expr [x, y, z] (x + ((y + x) + z))
+     $\rightarrow$ Expr [x, y, z] (x + ((y + x) + z))
 e2 x y z = Plus (Var FZ) 
@@
 e1_e2_testEq : (x, y, z : Nat) 
-             -> Maybe (((x + y) + (x + z)) = (x + ((y + x) + z)))
+             $\rightarrow$ Maybe (((x + y) + (x + z)) = (x + ((y + x) + z)))
 e1_e2_testEq x y z = testEq (e1 x y z) (e2 x y z)
@@
 total
-reflectNat : {n:Nat} -> ($\Gamma$ : Vect n Nat) -> (x:Nat) 
-                   -> (m ** ($\Gamma'$ : Vect m Nat ** (Expr ($\Gamma$ ++ ') x)))
+reflectNat : {n:Nat} $\rightarrow$ ($\Gamma$ : Vect n Nat) $\rightarrow$ (x:Nat) 
+             $\rightarrow$ (m ** ($\Gamma'$ : Vect m Nat ** (Expr ($\Gamma$ ++ $\Gamma'$) x)))
 \end{lstlisting}
@@
 
-This function takes ...
-TO DO : body of this function and brief description
+This function will reflect the natural number $x$ in the context of $\Gamma$, which contains $n$ already abstracted variables.
+This function will compute an extension --of arbitrary size $m$-- to the context, called $\Gamma'$.
+This extension will contain the variables used in $x$ that weren't already present in $\Gamma$.
+It will also produce the reflected term, which is expressed in the complete context $\Gamma ++\ \Gamma'$, and which is, of course, indexed over the concrete value $x$.
+
+If $x$ is the natural number $Z$, then we don't have any variable to add to $\Gamma$, so the extension will be the empty vector, and the reflected expression is simply $Zero$.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+reflectNat {n=n} G Z = (Z ** ([] ** (Zero {n=n+0} {G=G++[]})))
+\end{lstlisting}
+\end{center}
+\caption{Reflecting natural numbers - first pattern}
+\figrule
+\end{figure}
+
+If the natural number to reflect is an addition $x+y$, then we will start by reflecting $x$ in the context of $\Gamma$.
+That will give us an extension $\Gamma'$ and an expression $ex$ of type $Expr\ (\Gamma ++\ \Gamma')\ x$.
+We continue by reflecting $y$, but this time in the context $(\Gamma ++\ \Gamma')$ of $n+m$ already abstracted variables, because the reflection of $x$ has potentially abstracted some new variables, and we don't want to abstract them a second time.
+That will give us a second extension $\Gamma''$ and an expression $ey$ of type $Expr\ ((\Gamma ++\ \Gamma')\ ++\ \Gamma'')\ y$.
+We now simply want to return the $Plus$ of $ex$ and $ey$, but we can't immediately, because these encodings aren't defined on the same context.
+The context in which $ex$ makes sense is $(\Gamma ++\ \Gamma')$, but the context in which $ey$ makes sense is $((\Gamma ++\ \Gamma')\ ++\ \Gamma'')$.
+We will therefore need a function of weakening that takes a reflected expression $ex$, and returns the same expression, but expressed in an augmented context.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+reflectNat $\Gamma$ (x + y) =
+     let (_ ** ($\Gamma'$ ** ex)) = (reflectNat $\Gamma$ x) in
+     let (_ ** ($\Gamma''$ ** ey)) = (reflectNat ($\Gamma$ ++ $\Gamma'$) y) in
+     let result = Plus (weaken $\Gamma''$ ex) ey in 
+        (_ ** (($\Gamma'$ ++ $\Gamma''$) ** ?MreflectNat_1))
+\end{lstlisting}
+\end{center}
+\caption{Reflecting natural numbers - second pattern}
+\figrule
+\end{figure}
+
+The total extension that has been computed is $\Gamma' ++\ \Gamma''$, and the metavariable $MreflectNat\_1$ simply uses the associativity of the append operation to prove that the provided context $((\Gamma ++\ \Gamma')\ ++\ \Gamma'')$ is equal to the expected one $(\Gamma ++\ (\Gamma'\ ++\ \Gamma''))$.
+\\
+\\
+The function $weaken$ is easy to write because we're adding the extension $\Gamma'$ on the right of $\Gamma$.
+For example, extending the context $[v,\ w,\ x]$ with $[y,\ z]$ produces the context $[v,\ w,\ x,\ y,\ z]$.
+The variables $v$, $w$ and $x$, which were refereed to as the first, second and third variables in $\Gamma$, are still the first, second and third variables in the complete context.
+That means that a variable $Var\ i$ will still be a $Var\ i$, the position $i$ hasn't changed by augmenting the context.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+weaken : {n:Nat} $\rightarrow$ {m:Nat} $\rightarrow$ {$\Gamma$:Vect n Nat} $\rightarrow$ {x:Nat} 
+          $\rightarrow$ ($\Gamma'$:Vect m Nat) $\rightarrow$ (Expr $\Gamma$ x) $\rightarrow$ (Expr ($\Gamma$ ++ $\Gamma'$) x)
+weaken $\Gamma'$ Zero = Zero
+weaken $\Gamma'$ (Plus e1 e2) = Plus (weaken $\Gamma'$ e1) (weaken $\Gamma'$ e2)
+weaken {n=n} {m=m} {$\Gamma$=$\Gamma$} $\Gamma'$ (Var i) = Var (convertFin _ i m)    
+\end{lstlisting}
+\end{center}
+\caption{Weakening function}
+\figrule
+\end{figure}
+
+The position $i$ hasn't changed but however, the original $i$ had type $Fin\ n$, but the new $i$ must have type $Fin\ (n+m)$.
+This is why we has to use $convertFin$, which returns the same element, but seen in a bigger $Fin$.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+convertFin : (n:Nat) $\rightarrow$ (i:Fin n) $\rightarrow$ (x:Nat) $\rightarrow$ Fin (n+x)
+\end{lstlisting}
+\end{center}
+\caption{Conversion of Fin}
+\figrule
+\end{figure}

### 2016-01-17 paper-JFP-DTP/content/2-basicIdeas.tex
Section 2.5 about reflection for the small example finished

@@
 
-Note that for the moment, even if what we have is perfectly usable and works, we had to create these encodings $e1$ and $e2$ by hand, which is easy but time consuming.
+For the moment, even if what we have is perfectly usable and works, we had to create these encodings $e1$ and $e2$ by hand, which is easy but time consuming.
 We have replaced the (potentially hard) problem of proving something by the simplest problem of simply creating some encodings.
@@
 However, we can also note that even when done by hand, there is no room for making mistakes in this simple task of encoding : we simply can't generate a wrong encoding : if $e1$ and $e2$ are not respectively reflecting $((x+y) + (x+z))$ and $(x + ((y + x) + z))$ then these definitions won't typecheck because the expected and real index won't match.
-[perhaps giving example of rejected wrong definition for $e1$ here]
 
@@
 \end{center}
-\caption{Reflecting natural numbers - first pattern}
+\caption{Reflecting natural numbers - pattern for zero}
 \figrule
@@
 \end{center}
-\caption{Reflecting natural numbers - second pattern}
+\caption{Reflecting natural numbers - pattern for a plus}
 \figrule
@@
 \end{figure}
+
+We've treated the case of the constant $Z$ and the case of the addition.
+We now have to deal with the last possibility of a variable.
+For encoding a variable, we must see if this variable is already present in the context $\Gamma$ of already abstracted variables.
+If it is already present there, then there is no extension to build, and the reflected term is simple $Var\ (convertFin\ n\ i\ Z)$, where $i$ is the position of the variable in this context $\Gamma$.
+The conversion was needed because the original context was $\Gamma$, and the new one is $(\Gamma ++\ [\ ])$, which aren't automatically unifiable.
+However, if if this variable is missing, then we should add it, which means that we will return an extension containing this single variable.
+The original context $\Gamma$ had a size of $n$, and we've built an extension of size one, so the complete context has therefore size $(n+1)$.
+The variable that we've added is currently the last one of these  $(n+1)$ variables.
+If we use a function\footnote{This function doesn't have the type $(n:Nat)\ \rightarrow Fin\ n$ because there is no last element of a $Fin$ of size zero as there's no element at all}  $lastElement'\ :\ (pn:Nat)\ \rightarrow\ Fin\ (pn+1)$ to construct the last element of a $Fin$ of size $pn+1$, then the reflected term that we need to produce is simply $Var\ (lastElement'\ n)$.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+reflectNat {n=n} $\Gamma$ t with (isElement t $\Gamma$)
+  | Just (i ** p) = let result = Var {$\Gamma$=$\Gamma$++[]} (convertFin n i Z) in
+                        (Z ** ([] ** ?MreflectNat_2))
+  | Nothing ?= (((S Z) ** ([t] ** Var {$\Gamma$=$\Gamma$++[t]} (lastElement' n))))
+
+\end{lstlisting}
+\end{center}
+\caption{Reflecting natural numbers - pattern for a variable}
+\figrule
+\end{figure}
+
+$isElement$ is the function which checks whether an element $x$ belongs to a vector $\Gamma$, and if so, returns $Just$ and a dependent pair containing the index of $x$ in this vector, and a proof that $index\ i\ \Gamma\ =\ x$.
+
+ \begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting} [mathescape]
+isElement : {n:Nat} -> (x : a) -> ($\Gamma$ : Vect n a) 
+            -> Maybe (i:Fin n ** (index i $\Gamma$ = x))
+\end{lstlisting}
+\end{center}
+\caption{Returns the index and the proof that this index is the right one}
+\figrule
+\end{figure}
+
+The metavariable $MreflectNat\_2$ will require us to prove that the provided term of type $Expr\ (\Gamma ++\ [\ ])\ (index\ (convertFin\ n\ i\ Z)\ (\Gamma ++\ [\ ]))$ is transformable into a term of the expected type $Expr\ (\Gamma ++\ [\ ])\ t$.
+This is doable by using the proof $p$ --returned by $isElement$-- which says that $index\ i\ \Gamma\ =\ t$, together with the fact that $convertFin$ does not change the index, but only converts its type.
+
+As for the case where $t$ wasn't in the original context (the $Nothing$ case), we will need to prove that we can convert $(index\ (lastElement'\ n)\ (\Gamma ++\ [t]))$ into $t$.
+This is doable because we can prove the lemma $elemInBigerVect\ :\ \{T:Type\}\ \rightarrow \{v:Vect\ n\ T\} \rightarrow (i:Fin\ n) \rightarrow (elem:T) \rightarrow (proofInside : index\ i\ v\ =\ elem) \rightarrow (head:T) \rightarrow (index\ (FS\ i)\ (head::v)\ =\ elem)$.
+
+That finishes the automatic reflection for this very specific prover.

### 2016-01-17 paper-JFP-DTP/content/3-generalProblem.tex
Before describing the automatic reflection for Ring

@@
 class Set c where
-    ($\simeq$) : c -> c -> Type
+    ($\simeq$) : c $\rightarrow$ c $\rightarrow$ Type
 
-    refl : (x:c) -> x $\simeq$ x
-    sym : {x:c} -> {y:c} -> (x $\simeq$ y) -> (y $\simeq$ x)
-    trans : {x:c} -> {y:c} -> {z:c} -> (x $\simeq$ y) -> (y $\simeq$ z) -> (x $\simeq$ z)
+    refl : (x:c) $\rightarrow$ x $\simeq$ x
+    sym : {x:c} $\rightarrow$ {y:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ x)
+    trans : {x:c} $\rightarrow$ {y:c} $\rightarrow$ {z:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)
 
-    set_eq : (x:c) -> (y:c) -> Maybe (x $\simeq$ y)
+    set_eq : (x:c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
 \end{lstlisting}
@@
 class Set c => Magma c where
-    Plus : c -> c -> c
+    Plus : c $\rightarrow$ c $\rightarrow$ c
 \end{lstlisting}
@@
 class Magma c => SemiGroup c where
-    Plus_assoc : (c1:c) -> (c2:c) -> (c3:c) 
+    Plus_assoc : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ (c3:c) 
                -> (Plus (Plus c1 c2) c3 $\simeq$ Plus c1 (Plus c2 c3))
@@
     Zero : c    
-    Plus_neutral_1 : (c1:c) -> (Plus Zero c1 $\simeq$ c1)    
-    Plus_neutral_2 : (c1:c) -> (Plus c1 Zero $\simeq$ c1)
+    Plus_neutral_1 : (c1:c) $\rightarrow$ (Plus Zero c1 $\simeq$ c1)    
+    Plus_neutral_2 : (c1:c) $\rightarrow$ (Plus c1 Zero $\simeq$ c1)
 \end{lstlisting}
@@
 \begin{lstlisting}
--- This is just a conjunctive predicate
-hasSymmetric : (c:Type) -> (p:Monoid c) -> c -> c -> Type
+-- This is a conjunctive predicate
+hasSymmetric : (c:Type) $\rightarrow$ (p:Monoid c) $\rightarrow$ c $\rightarrow$ c $\rightarrow$ Type
 hasSymmetric c p a b $\simeq$ (Plus a b $\simeq$ Zero, Plus b a $\simeq$ Zero)
@@
 class Monoid c => Group c where
-    Minus : c -> c -> c
-    Neg : c -> c
-    Minus_simpl : (c1:c) -> (c2:c) -> Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
-    Plus_inverse : (c1:c) -> hasSymmetric c _ c1 (Neg c1)
+    Minus : c $\rightarrow$ c $\rightarrow$ c
+    Neg : c $\rightarrow$ c
+    Minus_simpl : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
+    Plus_inverse : (c1:c) $\rightarrow$ hasSymmetric c _ c1 (Neg c1)
 \end{lstlisting}
@@
 \begin{lstlisting}
-data ExprMa : Magma c -> (Vect n c) -> c -> Type where
-    ConstMa : (p : Magma c) -> ($\Gamma$:Vect n c) -> (c1:c)  -> ExprMa p $\Gamma$ c1 
-    PlusMa : {p : Magma c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
-         -> ExprMa p $\Gamma$ c1 -> ExprMa p $\Gamma$ c2 
-         -> ExprMa p $\Gamma$ (Plus c1 c2) 
-    VarMa : (p:Magma c) -> {$\Gamma$:Vect n c}
-         -> (i:Fin n) -> ExprMa p $\Gamma$ (index i $\Gamma$)
+data ExprMa : Magma c $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
+    ConstMa : (p : Magma c) $\rightarrow$ ($\Gamma$:Vect n c) $\rightarrow$ (c1:c)  $\rightarrow$ ExprMa p $\Gamma$ c1 
+    PlusMa : {p : Magma c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprMa p $\Gamma$ c1 $\rightarrow$ ExprMa p $\Gamma$ c2 
+         $\rightarrow$ ExprMa p $\Gamma$ (Plus c1 c2) 
+    VarMa : (p:Magma c) $\rightarrow$ {$\Gamma$:Vect n c}
+         $\rightarrow$ (i:Fin n) $\rightarrow$ ExprMa p $\Gamma$ (index i $\Gamma$)
 \end{lstlisting}
@@
 \begin{lstlisting}
-data ExprG :  Group c -> (Vect n c) -> c -> Type where
-    ConstG : (p : dataTypes.Group c) -> ($\Gamma$:Vect n c) 
-         -> (c1:c) -> ExprG p $\Gamma$ c1
-    PlusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
-         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
-         -> ExprG p $\Gamma$ (Plus c1 c2)
-    MinusG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} -> {c2:c} 
-         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ c2 
-         -> ExprG p $\Gamma$ (Minus c1 c2)
-    NegG : {p : Group c} -> {$\Gamma$:Vect n c} -> {c1:c} 
-         -> ExprG p $\Gamma$ c1 -> ExprG p $\Gamma$ (Neg c1)
-    VarG : (p : Group c) -> {$\Gamma$:Vect n c} 
-         -> (i:Fin n) -> ExprG p $\Gamma$ c1
+data ExprG :  Group c $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
+    ConstG : (p : dataTypes.Group c) $\rightarrow$ ($\Gamma$:Vect n c) 
+         -> (c1:c) $\rightarrow$ ExprG p $\Gamma$ c1
+    PlusG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ c2 
+         $\rightarrow$ ExprG p $\Gamma$ (Plus c1 c2)
+    MinusG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ c2 
+         $\rightarrow$ ExprG p $\Gamma$ (Minus c1 c2)
+    NegG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ (Neg c1)
+    VarG : (p : Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+         $\rightarrow$ (i:Fin n) $\rightarrow$ ExprG p $\Gamma$ c1
 \end{lstlisting}
@@
 \begin{lstlisting}
-groupReduce : {c:Type} -> {n:Nat} -> (p:Group c) -> {$\Gamma$:Vect n c} 
-	   -> {x:c} -> (ExprG p $\Gamma$ x) -> (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
+groupReduce : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+	   $\rightarrow$ {x:c} $\rightarrow$ (ExprG p $\Gamma$ x) $\rightarrow$ (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
 \end{lstlisting}
@@
 \begin{lstlisting}
-groupDecideEq : (p:Group c) -> {$\Gamma$:Vect n c} -> {x : c} -> {y : c} 
-	     -> (ExprG p $\Gamma$ x) -> (ExprG p $\Gamma$ y) 
-	     -> Maybe (x $\simeq$ y)
+groupDecideEq : (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {x : c} $\rightarrow$ {y : c} 
+	     $\rightarrow$ (ExprG p $\Gamma$ x) $\rightarrow$ (ExprG p $\Gamma$ y) 
+	     $\rightarrow$ Maybe (x $\simeq$ y)
 \end{lstlisting}
@@
 \begin{lstlisting}
-buildProofGroup : (p:Group c) -> {$\Gamma$:Vect n c} 
-  -> {x : c} -> {y : c} -> {x':c} -> {y':c} 
-  -> (ExprG p $\Gamma$ x') -> (ExprG p $\Gamma$ y') 
-  -> (x $\simeq$ x') -> (y $\simeq$ y') 
-  -> (Maybe (x $\simeq$ y))
+buildProofGroup : (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+  $\rightarrow$ {x : c} $\rightarrow$ {y : c} $\rightarrow$ {x':c} $\rightarrow$ {y':c} 
+  $\rightarrow$ (ExprG p $\Gamma$ x') $\rightarrow$ (ExprG p $\Gamma$ y') 
+  $\rightarrow$ (x $\simeq$ x') $\rightarrow$ (y $\simeq$ y') 
+  $\rightarrow$ (Maybe (x $\simeq$ y))
 buildProofGroup p ex' ey' px py with (exprG_eq p ex' ey')

### 2016-01-17 paper-JFP-DTP/content/4-detailsNormalisation.tex
Before describing the automatic reflection for Ring

@@
 \begin{lstlisting}
-littleRewriting : (p:algebraic structure on c) -> (gam:Vect n c) -> {c1:c} 
-        -> (Expr p gam c1) -> (c2 ** (Expr p gam c2, c1 $\simeq$ c2))
-simpleTreatment p gam e = (new expression, justification)
+littleRewriting : (p:algebraic structure on c) $\rightarrow$ ($\Gamma$:Vect n c) 
+       $\rightarrow$ {c1:c} $\rightarrow$ (Expr p $\Gamma$ c1) $\rightarrow$ (c2 ** (Expr p $\Gamma$ c2, c1 $\simeq$ c2))
+littleRewriting p $\Gamma$ e = (new expression, justification)
 \end{lstlisting}
@@
 \begin{lstlisting}
-assoc : (p:SemiGroup c) -> (gam:Vect n c) -> {c1:c} 
-        -> (ExprSG p gam c1) -> (c2 ** (ExprSG p gam c2, c1 $\simeq$ c2))
+assoc : (p:SemiGroup c) $\rightarrow$ (gam:Vect n c) $\rightarrow$ {c1:c} 
+        $\rightarrow$ (ExprSG p gam c1) $\rightarrow$ (c2 ** (ExprSG p gam c2, c1 $\simeq$ c2))
 assoc p gam (PlusSG 
@@
         (_ ** ((PlusSG (PlusSG e_ihx e_3') e_ihy), ?Massoc1))
-[...]
 \end{lstlisting}

### 2016-01-17 paper-JFP-DTP/content/6-discussion.tex
Before describing the automatic reflection for Ring

--- 2016-01-15 paper-JFP-DTP/content/5-discussion.tex
@@
 \begin{center}
-\begin{verbatim}
-decideEq : c -> c -> bool 
+\begin{lstlisting}
+decideEq : c $\rightarrow$ c $\rightarrow$ bool 
 decideEq x y =
@@
       let ey = reflect y in
-      beq_Expr (normalise ex) (normalise ey)
-\end{verbatim}
+      $beq_{Expr}$ (normalise ex) (normalise ey)
+\end{lstlisting}
 \end{center}
@@
 Instead, what they did for Coq's implementation of the Ring prover was to replace the main theorem $\forall\  x\ y:c,\ decideEq\ x\ y\ =\ true\ \rightarrow\ x\ \simeq\ y$ by the following weaker --but still powerful enough to build the desired tactic-- lemma\footnote{This lemma --that they call $setpolynomial\_simplify\_ok$-- can be found in $plugins/ring/Setoid\_ring\_normalize.v$}  : $f\_correct\ :\ \forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ \simeq\ reify\ e2$.
-Now, the tactic works like this. When trying to prove the goal $a=b$, it computes $(reflect\ a)$ and $(reflect\ b)$. It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$ to the goal. 
+Now, the tactic works like this.
+When trying to prove the goal $a \simeq b$, it computes $(reflect\ a)$ and $(reflect\ b)$.
+It then tries to apply $(f\_correct\ (reflect\ a)\ (reflect\ b))$ to the goal.
 \itemize

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
Section 3.5 about the general auto reflection written

@@
 Unfortunately, the types are mismatching for the two cases expressed by this pattern matching.
-The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\] [I AM SURPRISED HERE BY THE TYPE OF THE PROVIDED TERM].
+The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\]
 The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]

### 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex
perfs tests
Section 3.5 about the general auto reflection written

@@
 It is really important to understand that the root of the automatic building of the desired proof of $x=y$ happens precisely in these nested use of $plusZeroRightNeutral$ and $plusAssociative$ that we've seen in the definition of $expr\_l$ and of its meta-variable $MplusLExpr$.
-These proofs replace the arithmetical proofs that we were doing previously by hand, like in section 1 with the lemma $adc\_lemma\_2$.
+These proofs replace the arithmetical proofs that we were doing previously by hand, as it was the case in section 1 with the lemma $adc\_lemma\_2$.
 \\
@@
 
-That finishes the automatic reflection for this very specific prover.
+That finishes the automatic reflection for this specific prover. The encoding of $((x+y) + (x+z))$ can now be produced automatically with $reflectNat\ [\ ]\ ((x+y)\ +\ (x+z))$.

### 2016-01-18 paper-JFP-DTP/content/3-generalProblem.tex
Section 3.5 about the general auto reflection written

@@
 
+As we did for the very specific problem in section 2, we want to program an automatic reflection mechanism, in order to let the machine build automatically the encodings for us.
+This time, we want to write a function able to compute the ring encoding for any type that behaves as a ring.
+Said differently, we've got a carrier type $(c:Type)$, an instance $(p:Ring\ c)$ that says that $c$ behaves as a ring, a context $\Gamma$ of already abstracted variables, and an element $(x:c)$ to encode, and we want to return an $ExprR\ p\ \Gamma\ x$.
+We are doing this reflection for any type $c$, so we don't know which constants are inhabiting this type.
+For this reason, we will need to expect an extra argument --$funReflectCsf$-- of type $(ConstantsRingReflector\ c\ p)$ which is a function able to do the reflection for constants only.
+We will be able to provide this function for various datatypes, like $Nat$, $Z$ and $List$.
+But if the user wants to use the reflection mechanism for a new datatype of this own, then he will only have to provide this function of encoding for the constants of his type.
+He won't have to rebuild the entire reflection function.
+A $ConstantsRingReflector$ for a given type $c$ and a given instance $p$ is a function that takes any context $\Gamma$, any element of $c$, and which tries to produce an $(ExprR\ p\ \Gamma\ x)$.
+
 \begin{figure}[H]
@@
 \begin{lstlisting}
-reflect : 
+ConstantsRingReflector : {c:Type} $\rightarrow$ (p:Ring c) $\rightarrow$ Type
+ConstantsRingReflector {c=c} p = 
+     ({n:Nat} $\rightarrow$ ($\Gamma$:Vect n c) $\rightarrow$ (x:c) $\rightarrow$ Maybe (ExprR p $\Gamma$ x))
 \end{lstlisting}
 \end{center}
-\caption{Type of the reflecting function for ...}
+\caption{Type of the function doing the reflection for constants}
 \figrule
@@
 
-The reflected term is guaranteed to be a faithful representation of the input $c1$.
-We can write this function because Idris has a $\%syntax$ command which let us do pattern maching on syntax (ie, it let us access the AST of a term), and this $\%syntax$ produces functions which are typed !
-(unlike Coq's $Ltac$).
+If $x$ is indeed a constant, then it is supposed to return $Just$ and a term reflecting this constant.
+If $x$ was something else (a variable, a sum $a+b$, a product $a*b$, ...) then it simply returns $Nothing$.
+\\
+\\
+We can now present the type of the reflection function for ring.
 
-TO DO : Body and description of this function
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+reflectRingTerm : {c:Type} $\rightarrow$ {p:Ring c} $\rightarrow$ {n:Nat} $\rightarrow$ ($\Gamma$ : Vect n c) 
+     $\rightarrow$ (funReflectCsf:ConstantsRingReflector {c=c} p) $\rightarrow$ (x:c) 
+     $\rightarrow$ (n' ** ($\Gamma'$:Vect n' c ** (ExprR {c=c} {n=n+n'} p ($\Gamma$ ++ $\Gamma'$) x)))
+\end{lstlisting}
+\end{center}
+\caption{Automatic reflection to ring}
+\figrule
+\end{figure}
+
+This function will work very similarly to the reflection function for $Nat$ that we've writen in section 2.5.
+Here, there are just more constructions to deal with, because $c$ is equipped with negations, subtractions and multiplications.
+We just show the pattern for the multiplication.
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+reflectRingTerm {p} {n=n} $\Gamma$ funReflectCsf (a*b) = 
+   let (n' ** ($\Gamma'$ ** a')) = reflectRingTerm $\Gamma$ funReflectCsf a in
+   let (n'' ** ($\Gamma''$ ** b')) = reflectRingTerm ($\Gamma$ ++ $\Gamma'$) funReflectCsf b in
+   let this = MultR (weakenR $\Gamma''$ a') b' in
+	((n' + n'') ** (($\Gamma'$++$\Gamma''$) 
+	  ** (convertVectInExprR (plusAssociative n n' n'')
+	                   (vectAppendAssociative $\Gamma$ $\Gamma'$ $\Gamma''$) this)))
+\end{lstlisting}
+\end{center}
+\caption{Automatic reflection - case of a multiplication}
+\figrule
+\end{figure}
+
+After having done the same thing with additions, negations and subtractions, we will just have the case of constants and variable missing.
+But since we've got $funReflectCsf$ --the function that tries to do the reflection for constants only--, we can easily finish the definition of $reflectRingTerm$.
+If the dedicated reflector $funReflectCsf$ has returned something, then we know the input is a variable, and we have its encoding.
+Otherwise, we just have to treat it as a variable.
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+reflectRingTerm {p} {n=n} $\Gamma$ funReflectCsf t =
+    case funReflectCsf $\Gamma$ t of
+      -- If 'funReflectCsf' has decided that 't' is a constant, then we trust it
+      Just this => Z ** ([] ** this)
+      -- If not, then it should be considered as a variable
+      Nothing => case (isElement t $\Gamma$) of
+          Just (i ** pr) => let res = VarR p (RealVariable i) in
+			        (Z ** ([ ] ** res))
+          Nothing => let res = VarR p (RealVariable (lastElement' n)) in
+			 (S Z ** ([t] ** res))
+
+\end{lstlisting}
+\end{center}
+\caption{Automatic reflection - case of constants and variables}
+\figrule
+\end{figure}
+
+We've omitted to show a few conversions of type in order to keep this part of the definition more readable. \\
+\\
+It is remarkable that the reflected term is guaranteed to be a faithful representation of the input.
+We've been able to define this function with a very strong and precise type because Idris has a $\%syntax$ command which let us do arbitrary pattern matching on syntax, and this $\%syntax$ produces functions which are typed, unlike Ltac terms in Coq.

### 2016-01-18 paper-JFP-DTP/content/4-detailsNormalisation.tex
Section 3.5 about the general auto reflection written

@@
 
-We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised underneath structures.
-By doing so, we show that the construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
+We now describe briefly how the normalisation functions work for some of the different algebraic structures, and we also explain how we've been able to reuse each prover for the more specialised structures.
+The construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
 \begin{figure}[H]

### 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex
toDo list updated

@@
 The position $i$ hasn't changed but however, the original $i$ had type $Fin\ n$, but the new $i$ must have type $Fin\ (n+m)$.
-This is why we has to use $convertFin$, which returns the same element, but seen in a bigger $Fin$.
+This is why we've used $convertFin$, which returns the same element, but seen in a bigger $Fin$.
 
@@
 \begin{lstlisting} [mathescape]
-isElement : {n:Nat} -> (x : a) -> ($\Gamma$ : Vect n a) 
-            -> Maybe (i:Fin n ** (index i $\Gamma$ = x))
+isElement : {n:Nat} $\rightarrow$ (x : a) $\rightarrow$ ($\Gamma$ : Vect n a) 
+            $\rightarrow$ Maybe (i:Fin n ** (index i $\Gamma$ = x))
 \end{lstlisting}

### 2016-01-18 paper-JFP-DTP/content/4-detailsNormalisation.tex
toDo list updated

@@
 In a semigroup, we only have to deal with the property of associativity.
-As it was the case with the toy example in the previous section, we will have to rearrange the parenthesis on a systematic way, either left associative, or right associative.
+As it was the case with the toy example in section 2, we will have to rearrange the parenthesis on a systematic way, either left associative, or right associative.
 But this is not the only thing that we have to do.
@@
 The simplification between the constants $const1$ and $const2$ is denoted here as $[const1+const2]$, and this computation can be done at the underneath $Magma$ level, with its normalisation function called $magmaReduce$.
-The first pattern is rewritten into $(x + [const1+const2]) + y$, the second into $x + [const1+const2]$ and the third one into $[const1 + const2] + x$.
+The first pattern is rewritten into $(x + [const1+const2]) + y$, the second into $(x + [const1+const2])$ and the third one into $([const1 + const2] + x)$.
 We show the code for the first pattern.
@@
 \begin{lstlisting}
-assoc : (p:SemiGroup c) $\rightarrow$ (gam:Vect n c) $\rightarrow$ {c1:c} 
-        $\rightarrow$ (ExprSG p gam c1) $\rightarrow$ (c2 ** (ExprSG p gam c2, c1 $\simeq$ c2))
-assoc p gam (PlusSG 
+assoc : (p:SemiGroup c) $\rightarrow$ ($\Gamma$:Vect n c) $\rightarrow$ {c1:c} 
+        $\rightarrow$ (ExprSG p $\Gamma$ c1) $\rightarrow$ (c2 ** (ExprSG p $\Gamma$ c2, c1 $\simeq$ c2))
+assoc p $\Gamma$ (PlusSG 
              (PlusSG ex (ConstSG _ _ const1)) 
              (PlusSG (ConstSG _ _ const2) ey)) =
-    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p gam ex) in
-    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p gam ey) in
+    let (r_ihx ** (e_ihx, p_ihx)) = (assoc p $\Gamma$ ex) in
+    let (r_ihy ** (e_ihy, p_ihy)) = (assoc p $\Gamma$ ey) in
     let (r_3 ** (e_3, p_3)) = magmaReduce (semiGroup_to_magma 
@@
 In a monoid, we simply have to eliminate zeros, thanks to the two properties of left and right neutrality.
-Because the simplification of additions between a zero and a constant has be done at the underneath level of semigroup (by computing the addition of these two constants), we only have to simplify the addition between the constant zero and a variable.
+Because the simplification of additions between a zero and another constant has already be done at the level of semigroup (by computing the addition of these two constants), we only have to simplify the addition between the constant zero and a variable.
 There is in fact two cases, since the variable can come first, or the $Zero$ can come first.
-We can write a function $elimZero$ which rewrites these two patterns $x+Z$ and $Z+x$ as $x$.
+We can write a function $elimZero$ which rewrites these two patterns $x+Z$ and $Z+x$ into $x$.
 
@@
 
-The normalization of reflected expressions in a monoid is simply made of a call to the function reducing terms on a semigroup, followed by a call to elimZero which will eliminates the remaining additions with zeros.
+The normalization of reflected expressions in a monoid is simply made of a call to the function reducing terms on a semigroup, followed by a call to elimZero that will eliminates the remaining additions with zeros.
 
@@
 
-The first thing that the reduction function of the group prover will have to do is simply to transform every subtraction $a-b$ into $a + -b$, and we are entitled to do so precisely because we have available the property $Minus\_simpl$ contained in the instance of the type class.
+The first thing that the reduction function of the group prover will have to do is simply to transform every subtraction $(a-b)$ into $(a + (-b))$, and we are entitled to do so precisely because we have available the property $Minus\_simpl$ contained in the instance of the type class.
 After that, the reduction continues with the propagation of the $Neg$ operation inside the parenthesis and $-(a+b)$ is transformed into $(-b) + (-a)$.
@@
 In these cases, there might be some simplifications to do thanks to the associativity property.
-With two levels of sums, we will transform $(e1 + (-e2 + e3))$ and $(-e1 + (e2+e3))$ into $e3$ when $e1$ can be normalised to $e2$.
+With two levels of sums, we will transform $(e1 + (-e2 + e3))$ and $(-e1 + (e2+e3))$ into $e3$ when $e1$ can be reduced to $e2$.
 Also, $((e1+e2) + -e3))$ and $((e1+(-e2)) + e3)$ are simplified into $e1$ when $e2$ can be reduced to $e3$.

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
section 1 : updating the mismatching indices

@@
 adc (numx # bX) (numy # bY) carry
-   = let (v1 ** (v2 ** (carry0, lsb, _))) = addBit bX bY carry in
+   = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) = addBit bX bY carry in
           adc numx numy carry0 # lsb
@@
 The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\]
-The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
+The problem is even worse for the second case where the expected index is 
+\[(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\] while we're trying to provide a term index over 
+\[plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))\]
 For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
@@
 This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
-This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will require us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
+This mean that just a little change in our datatypes $Bit$ and $Binary$, or in the definitions of $addBit$ and $adc$ will require us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
 The re-usability of this kind of proofs is therefore null, because they are performing rewritings for very specific terms.
 
-If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), the associativity and commutativity of $+$.
+If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), and the associativity and commutativity of $+$.
 Thus, we're rewriting a term by using the properties of a commutative monoid.

### 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex
syntax fix

@@
 Note that $Z$ is also a left neutral element, but this property is not needed because we have this behaviour by reduction, as $+$ is defined recursively on its first argument.
-Of course, some datatypes can have more or less properties than these two, and this work will be extended in sections 3 and 4 where we will present a very general hierarchy of provers for multiple algebraic structures.
+Of course, other datatypes can have more or less properties than these two, and this work will be extended in sections 3 and 4 where we will present a very general hierarchy of provers for various algebraic structures.
 \\

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
Added bibtex

--- 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
@@
 
-Proving that one term is equal to another one within some theories is something quite common in formal certification.
-The two terms are usually not syntactically equal --this case would be trivial--, but they are equal according to some properties.
-These properties are either obtained by implementation (the implementation of $+$ for natural numbers satisfies the property of associativity for example), or by axioms if we are working within an axiomatised structure.
+Proving that one term is equal to another within some theories is quite common
+in formal certification. The two terms are usually not syntactically 
+equal---this case would be trivial---but they are equal relative to some
+equivalence properties. These properties are either obtained by lemmas
+(the implementation of $+$ for natural numbers satisfies the property of
+associativity for example), or by axioms.
 
-For example, when working with natural numbers, one might need to prove the lemma $\forall\ xy:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$, and when working with lists and concatenation of lists, one might need to prove something like $\forall\ l1\ l2\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\ (l3++l4)$.
-Proving this kind of auxiliary lemmas is time consuming, is much too specific for being reusable, and isn't the interesting part of formal certification.
-Therefore, we would like to be able to automatize a far as possible the constructions of this kind of proofs for the Idris language, a dependently typed language which doesn't contain many proof automations at the moment.
-\\
-\\
+For example, when working with natural numbers, one might need to prove the
+lemma $\forall\ x, y:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$. Or,
+when working
+with lists and concatenation of lists, one might need to prove the lemma
+$\forall\ l1,\ l2,\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\
+(l3++l4)$. Proving this kind of auxiliary lemmas is time consuming,
+typically not reusable due to being very specific, and is rarely
+the interesting part of formal
+certification. Therefore, we would like to be able to automatize as far as
+possible the constructions of this kind of proofs. We have implemented
+the machinery to do so in Idris, a
+dependently typed language~\cite{brady2013idris}.
+
 When doing formal certification, many functions have a lemma of correctness that talks about equality of terms.

### 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex

--- 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex
@@
 For this reason, we will work by reflection.
-In this little example, it means that we will define a data type which will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
-This data type will allow us to know the internal structure of a number, ie, where the variables and constants are.
-Indeed, this data type will allow pattern matching.
+In this little example, it means that we will define a datatype that will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
+This datatype will let us inspect the internal structure of a number by pattern matching.
 Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't what we needed.
-With the first approximation of the data type $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructor $Plus$, $Var$ and $Zero$, which gives us the informations we want.
+With the first approximation of the data type $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructors $Plus$, $Var$ and $Zero$.
 
@@
 
-If we continue with this first definition of $Expr$, the normalisation function will certainly take an Expr and produce another Expr, and we will need to prove the following lemma afterwards : \\
+If we continue with this first definition of $Expr$, the normalisation function will certainly take an Expr and produce another Expr, and we will need to prove the following correctness lemma afterwards : \\
 $\forall\ e:Expr\ \Gamma,\ reify\ (reduce\ e)\ =\ reify\ e$ \\
@@
 This proof can be quite tricky to build because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
-For this little example, the reduction procedure will not be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more rewriting rules to deal with and it will certainly become more problematic to "unfold" the definition of a gigantic reduction procedure.
-We want to avoid pain as much as possible, and for this reason, we don't want to have to write huge proofs.
+For this little example, the reduction procedure won't be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more properties to deal with and it will certainly become more problematic to unfold the definition of a gigantic reduction procedure.
+We want to avoid pain as much as possible, and for this reason, we don't want to have to write big proofs, and even to avoid the writing of proofs as much as possible.
 \\
@@
 Thus, it won't be necessary to define the $reify$ function, as we will know directly the concrete element reflected by a term of type $Expr\ \Gamma\ x$ just by looking at its index $x$.
-Something more important is that we will also have a guarantee that the reflection of an expression $e$ is indeed a faithful representation of $e$ \\
+We also get the guarantee that the reflection of an expression $e$ is indeed a faithful representation of $e$ \\
 

### 2016-01-18 paper-JFP-DTP/content/6-discussion.tex

@@
 
-In Coq's latest implementation of a Ring solver, described in [1], they almost followed what has just been desribed as a "traditional" approach with the use of many auxiliary lemmas.
+In Coq's latest implementation of a Ring solver, described in~\cite{Coq2005}, they almost followed what has just been desribed as a "traditional" approach with the use of many auxiliary lemmas.
 The few differences essentially lie in the way the function $reflect$ is defined.

### 2016-01-18 paper-JFP-DTP/paper.tex
Check
paper : use of listings and corrections
Added new section

@@
 \usepackage{float}
+\usepackage{listings}
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  keepspaces=true,
+  mathescape
+}
 
@@
 
-\input{./content/5-discussion.tex}
+\input{./content/5-results.tex}
 
-\begin{thebibliography}{}
- \bibitem[\protect\citename{B. Gregoire and A. Mahboubi}2005]{Coq2005}
-   Gregoire,~B. and Mahboubi,~A. (2005) Proving equalities in a commutative ring done right in Coq, TPHOLs 2005, Oxford, UK, August 22-25, 2005, Proceedings
+\input{./content/6-discussion.tex}
 
- \bibitem[\protect\citename{W.A Howard}1980]{How80}
-   Howard W.A.
+\bibliographystyle{jfp}
+\bibliography{dtp}
+
+%\begin{thebibliography}{}
+% \bibitem[\protect\citename{B. Gregoire and A. Mahboubi}2005]{Coq2005}
+%   Gregoire,~B. and Mahboubi,~A. (2005) Proving equalities in a commutative ring done right in Coq, TPHOLs 2005, Oxford, UK, August 22-25, 2005, Proceedings
+
+% \bibitem[\protect\citename{W.A Howard}1980]{How80}
+%   Howard W.A.
 The formulae-as-types notion of construction.
@@
 
-\end{thebibliography}
+%\end{thebibliography}
 

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
addBit fix with Refl insteak of joker

--- 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
@@
               (Bit bX, Bit bY, c + x + y = bY + 2 * bX)))
-addBit b0 b0 b0 = (_ ** (_ ** (b0, b0, _)))
-addBit b0 b0 b1 = (_ ** (_ ** (b0, b1, _)))
-addBit b0 b1 b0 = (_ ** (_ ** (b0, b1, _)))
-addBit b0 b1 b1 = (_ ** (_ ** (b1, b0, _)))
-addBit b1 b0 b0 = (_ ** (_ ** (b0, b1, _)))
-addBit b1 b0 b1 = (_ ** (_ ** (b1, b0, _)))
-addBit b1 b1 b0 = (_ ** (_ ** (b1, b0, _)))
-addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, _)))
+addBit b0 b0 b0 = (_ ** (_ ** (b0, b0, Refl)))
+addBit b0 b0 b1 = (_ ** (_ ** (b0, b1, Refl)))
+addBit b0 b1 b0 = (_ ** (_ ** (b0, b1, Refl)))
+addBit b0 b1 b1 = (_ ** (_ ** (b1, b0, Refl)))
+addBit b1 b0 b0 = (_ ** (_ ** (b0, b1, Refl)))
+addBit b1 b0 b1 = (_ ** (_ ** (b1, b0, Refl)))
+addBit b1 b1 b0 = (_ ** (_ ** (b1, b0, Refl)))
+addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, Refl)))
 \end{lstlisting}
@@
 
-It should be pointed that the proof component of this function is established without any effort, as we always give the "\_" symbol for the proof which means "infer it!".
-For each of the eight patterns, a simple computation provides the desired proof, and thus the typechecker can deal with them.
+It should be pointed that the proof component of this function is established without any effort, as we always give the "Refl" for the proof.
+Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and "Refl" can therefore be used.
 We also note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : the two bits produced effectively represent the addition of the three bits given in input.

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex

--- 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
@@
 
-When doing formal certification, many functions have a lemma of correctness that talks about equality of terms.
-More precisely, the following situation is very familiar.
-The function that we wants to certify is $f\ :\ A \rightarrow B$.
-We have two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B \rightarrow SEM$ of interpretation which associate a semantic to each type A and B.
-If this semantic says everything that matters for the behaviour of $f$, then the lemma of correctness of f can be expressed as $forall\ a:A,\ semA\ a = semB (f\ a)$.
-The interesting thing is that, very often, the type $SEM$ has the structure of some nice algebraic structure, and it becomes thus possible to decide equalities in it automatically.
-All the equality symbols we have encountered so far refer to the propositional equality.
-We remind the readers in the next subsection how this equality is defined in Idris, and more generally in Intentional Type Theories, and how it differs from the --more primitive-- judgemental equality.
+When doing formal certification, we often require equality proofs between
+terms.
+For example, say we want to certify the function $f\ :\ A \rightarrow
+B$. We have two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B
+\rightarrow SEM$ which associate a semantics to each type A
+and B. If this semantics describes all of the important properties of
+the behaviour of $f$,
+then the lemma of correctness of $f$ can be expressed as $forall\ a:A,\ semA\ a =
+semB (f\ a)$. Very often, the type $SEM$ has some algebraic
+algebraic structure, and it becomes thus possible to
+decide equalities in it by following some standard rewriting rules.  
+All the equality symbols we have
+encountered so far refer to the propositional equality. In the next section,
+we will see how this is defined in Idris, and how it differs from the more
+primitive judgemental equality.
 
-\subsection{The propositional equality and the judgemental equality in Intentional Type Theories}
-In Mathematics, the equality is a proposition, which can hold or not, which can be disproved, etc...
-Since in type theory, propositions are seen as types \cite{How80}, the proposition that two elements $a$ and $b$ are equal must correspond to a type.
-Thus, if $a$ and $b$ are of type\footnote{It can be useful sometimes to allow the writing of heterogeneous equalities, and in this case, JMeq (which is the standard equality in Idris) can be used.
-It follows exactly the same idea as the standard, homogeneous, equality that we've reminded} $A$, then the type $Id_A(a, b)$ represents the proposition "$a$ is equal to $b$".
-If this type is inhabited, then $a$ is said to be provably equal to $b$.
-Thus, Id is a type family (parametrised by the type $A$) indexed over two elements of A.
-That gives $Id (A:Type)\ :\ A \rightarrow A \rightarrow Type$.
-Often, for convenience, $Id_A\ a\ b$ is written $a\ =_A\ b$.
-This equality is the equality which can be manipulated in the language.
-There is another --more primitive-- notion of equality in Intentional Type Theories, called judgemental equality, or definitional equality.
-This second equality means "equal by definition".
-The judgemental equality is something which can't be negated or assumed : we can't talk about this primitive equality inside the theory.
-Whether or not two expressions are equal by definition is just a matter of expanding out the definitions.
-For example, if $f\ :\ \mathbb{N}\ \rightarrow\ \mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is definitionally equal to $7$.
-Definitional equality basically contains unfolding of functions and (beta, iota...) reductions, until no more reduction can be performed.
-Definitional equality is noted by $\equiv$.
+\subsection{Propositional and Judgemental Equalities in Intentional Type Theories}
+In Mathematics, equality is a proposition, which can hold or otherwise, or be disproved.
+Since in type theory, propositions are seen as types~\cite{How80}, the
+proposition that two elements $x$ and $y$ are equal corresponds to a type.
+Thus, if $x$ and $y$ are of type $a$, then the type
+$Id_a(x, y)$ represents the proposition "$x$ is equal to $y$". If this type is
+inhabited, then $x$ is said to be provably equal to $y$.  Thus, $Id$ is a type
+family (parametrised by the type $a$) indexed over two elements of $a$, giving
+$Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$. For
+convenience, we write $Id_a\ x\ y$ as $x\ =_a\ y$.  This equality is the
+equality which can be manipulated in the language.
 
-The judgemental equality has to be included in the propositional equal, because what is equal by definition should be provably equal.
-This is accomplished by giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
-\begin{figure}[H]
-\figrule
-\begin{center}
+There is another more
+primitive notion of equality in Intentional Type Theories, called judgemental
+equality, or definitional equality. This second equality means ``equal by
+definition'' The judgemental equality cannot be negated or
+assumed; we cannot talk about this primitive equality inside the theory.
+Whether or not two expressions are equal by definition is a matter of
+evaluating the definitions. For example, if $f\ :\ \mathbb{N}\ \rightarrow\
+\mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is
+definitionally equal to $7$. Definitional equality entails unfolding
+of functions and reductions, until no more reduction can be
+performed. We denote definitional equality by $\equiv$.
+
+The judgemental equality has to be included in the propositional equal, because
+what is equal by definition should be provably equal.  This is accomplished by
+giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
+One way to implement $Id$ in Idris is:
+
 \begin{lstlisting}
-Id (A:Type) : A $\rightarrow$ A $\rightarrow$ Type :=
-  Refl : forall a:A, Id a a
+data Id : a -> a -> Type where
+     Refl : (x : a) -> Id x x
 \end{lstlisting}
-\end{center}
-\caption{The type Id}
-\figrule
-\end{figure}
 
-The only chance for $(Id_A\ a\ b)$ to be inhabited is therefore that $a$ should be (by definition!) $b$.
-In this case, the constructor $Refl$ helps to create a proof of this equality : $Refl\ a$ is precisely the proof which says that $a=_Aa$.
+The only way for $(Id_a\ x\ y)$ to be inhabited is therefore that $x$ should be
+(by definition!) $y$. In this case, the constructor $Refl$ helps to create a
+proof of this equality : $Refl\ x$ is precisely the proof which says that
+$x=_ax$.
 
-When reading these definitions, one could incorrectly think that the propositional equality is a simple wrapper of the judgemental equality.
-This is not the case : the propositional equality doesn't only contains the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-This induction principle says that if a proposition holds for the base cases, and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the recursive constructors, then this proposition will hold for any term of this type.
+When reading these definitions, one could think that the
+propositional equality is a simple wrapper of the judgemental equality. This is
+not the case: the propositional equality does not only contain the judgemental
+equality, because in these type theories, a principle of induction is
+associated with each inductive type. This induction principle says that if a
+proposition holds for the base cases, and if it can be showed that when it
+holds for some terms then it will also hold for the bigger terms obtained by
+using the recursive constructors, then this proposition will hold for any term
+of this type.
 
-The induction principles work for any proposition, so it also works with the propositional equality.
-That means that the axiom of induction principle has made the type $Id(a,b)$ bigger than it looks like (in the definition 1), and there are therefore things which are provably equal which unfortunately aren't equal syntactically after an expanding of all the definitions.
-The activity of proving equalities is therefore in these theories something which isn't automatically decidable by the type-checker in the general case.
+Induction principles work for any proposition, so it also works with the
+propositional equality. That means that the axiom of induction principle has
+made the type $Id(a,b)$ bigger than it looks like (in the definition 1), and
+there are therefore things which are provably equal which unfortunately aren't
+equal syntactically after an expanding of all the definitions. The activity of
+proving equalities is therefore in these theories something which isn't
+automatically decidable by the type-checker in the general case.
 
-For example, we can prove that $n+0 = n$ by induction, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
+For example, we can prove that $n+0 = n$ by induction, even if $n+0 \not\equiv
+n$ with the usual definition of $+$, recursive on its first argument.
 
-These proofs are usually omitted "on the paper" for the ordinary every day maths, but they are required by proof assistants, no matter how obvious they are.
-Currently, in the Idris programming language, they are done by hand, and they are made of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
-These proofs are the every day routine when working in a dependently typed language, as Idris, and especially when intensively indexing types over values in order to capture certain logical properties.
-For example, If a type $T$ is indexed over natural numbers, then the definition of a function which has co-domain $T$ will often require to prove that the index of the produced term is equal to the expected index.
-This proof will use properties of natural numbers and their usual operations : symmetry and commutativity for +, existence of neutral element for +, etc.
+These proofs are usually omitted ``on paper'' for everyday mathematics, but
+they are required by proof assistants, no matter how obvious they are to the
+human reader.
+Currently, in the Idris programming language, proofs must be completed
+by hand, and consist
+of a potentially long sequence of rewriting, where each step of
+rewriting uses one of the available properties. These proofs can be
+an everyday
+routine when working in a dependently typed language, and especially
+when intensively indexing types over values in order to capture certain logical
+properties.  For example, If a type $T$ is indexed over natural numbers, then
+the definition of a function which has codomain $T$ will often require a
+proof that the index of the produced term is equal to the expected index. This
+proof may use properties of natural numbers and their usual operations:
+symmetry and commutativity for +, existence of neutral element for +, etc.
 
-\subsection{A motivating little example}
-We introduce a little example which shows how these proof obligations appear when we index a type by values like natural numbers.
-Let's define the types of $Bit$ and $Binary$ number, both indexed over the value they represent, expressed in base ten as a natural number.
+\subsection{Motivating example: Verified Binary Arithmetic}
+
+We revisit an example from previous work~\cite{bradytfp07} 
+which shows how proof obligations might appear
+when we index a type by natural numbers.  We define the types of
+$Bit$ and $Binary$ number in Figure~\ref{binarynums}, 
+both indexed over the value they represent, expressed as a natural number.
+
 \begin{figure}[H]
@@
 \caption{Bit and binary number}
+\label{binarynums}
 \figrule
@@
 
-This type of binary numbers\footnote{This representation allows to represent a binary number of width zero with the constructor $zero$, but instead we could have decided to replace this constructor with a constructor taking one bit as parameter and producing a binary number of width one.
-The representation we've chosen is a bit easier to manipulate.}  has two constructors, and the most interesting one is $\#$ which enables to construct bigger binary numbers : we can extend a binary number $bin$ by adding a digit to its right, and the value represented by this new binary number is two times the value ($v$) represented by $bin$ plus the value ($bit$) represented by the added digit.
+It allows representing a
+binary number of width zero with the constructor $zero$, 
+and representing larger numbers with $\#$.
+We
+can extend a binary number $bin$ by adding a digit to its right, and the value
+represented by this new binary number is two times the value ($v$) represented
+by $bin$ plus the value ($bit$) represented by the added digit.
 
-We want to write a function performing the addition between two binary numbers.
-We start with an auxiliary function which adds three bits (the third one will latter play the role of the carry), and produces the two bits of the result : the first one is the most significant bit and the second one is the least significant bit .
+To write a function performing the addition between two binary numbers,
+we start with an auxiliary function (Figure~\ref{addbit}) which adds three bits
+(the third one plays the role of a carry flag), and produces the two bits of
+the result.
 
@@
 \caption{Addition of three bits}
+\label{addbit}
 \figrule
@@
 
-This function produces a dependent pair.
-The first component is the value $bX$ (expressed as a natural number) represented by the first bit produced (of type $Bit\ bX$), and the second component is itself another dependent pair.
-This second dependent pair is composed of the value $bY$ represented by the second bit produced, and of a triple containing finally the two bits produced and a proof that these two bits are effectively encoding the result of the addition we had to perform.
-For example, with the line corresponding to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
+This function produces a dependent pair. The first component is the value $bX$
+(expressed as a natural number) represented by the first bit produced (of type
+$Bit\ bX$), and the second component is another dependent pair. This
+second dependent pair is composed of the value $bY$ represented by the second
+bit produced, and of a triple containing the two bits produced and a
+proof that these two bits correctly encode the result of the addition.
+For example, with the line corresponding to the computation
+$1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$)
+and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
 
-It should be pointed that the proof component of this function is established without any effort, as we always give the "\_" symbol for the proof which means "infer it!".
-For each of the eight patterns, a simple computation provides the desired proof, and thus the typechecker can deal with them.
-We also note that there is no need to produce any lemma of correctness about $addBit$ afterwards : the correct by construction style in which it is written already gives the property that we want : the two bits produced effectively represent the addition of the three bits given in input.
+It should be pointed that the proof component of this function is established
+without any effort, as we always give the ``\_'' symbol for the proof which means
+``infer it!''. For each of the eight patterns, a simple computation provides the
+desired proof, and thus the typechecker can deal with them.  We also note that
+there is no need to produce any lemma of correctness about $addBit$ afterwards.
+The correct by construction style in which it is written already gives the
+property that we want: the two bits produced effectively represent the
+addition of the three bits given in input.
 
-We finally define the function adding two binary numbers and a bit (used as a carry).
-We decide that this addition will only work for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
-This result should of course represent the value $c + x + y$ if $c$ is representing the value of the input bit, and if $x$ and $y$ are representing the values of the two binary numbers given in input.
-Thus, we're tempted to write :
-\begin{figure}[H]
-\figrule
-\begin{center}
+We finally define the function adding two binary numbers and a carry flag.
+This addition works for two binary numbers
+expressed with the same number of bits, and it produces a result with one more
+bit. This result represents the value $c + x + y$, where $c$ 
+representis the value of the input carry bit, and $x$ and $y$ represent
+the values of the two binary numbers given in input. Thus, we'd like to
+be able to write:
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Addition of two binary numbers and a carry}
-\figrule
-\end{figure}
 
-Unfortunately, the types are mismatching for the two cases expressed by this pattern matching.
-The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we're providing a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\]
-The problem is even worse for the second line where the result is expected to have type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0)))\] while we're trying to provide a term of type \[plus\ v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\]
-For this kind of situation, Idris provides the possibility to write a provisional definition with the syntax "?=".
-Using a provisional definitions here will make the definition being accepted, but it also generates two proof obligations, one for each pattern, and both of them are saying that the provided term can be converted into a term which has the expected type.
-For both of these two patterns, it is in fact possible to prove the equality between the expected and the provided types, precisely because this function is correct.
-These proofs are made of a series of rewriting, and each rewriting uses a property about the operation $+$ on natural numbers.
-For example, the proof for the second pattern looks like this :
+Unfortunately, the types do not match for either of the two cases. 
+The result of the first line $adc\ zero\ zero\ carry$ is
+expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we
+provide a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\]. The
+problem is even worse for the second line where the result is expected to have
+type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\
+(plus\ v\ (plus\ v\ 0)))\] while we provide a term of type \[plus\
+v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\].
+
+For
+this situation, Idris provides \emph{provisional definitions} with the 
+syntax \texttt{?=}. Using a provisional definition here will make
+the definition acceptable, but it also generates two proof obligations, one
+for each pattern, both of which say that the provided term can be
+converted into a term which has the expected type.  For both of these
+patterns, it is possible to prove the equality between the expected and
+the provided types. These proofs
+consist of a series of rewriting steps, where each step uses a property about the
+operation $+$ on natural numbers.  For example, the proof for the second
+pattern looks like this (with some detail elided):
+
 \begin{figure}[H]
@@
     [...]
-    rewrite sym (plusAssociative (plus c (plus bit0 (plus v v))) bit1 (plus v1 v1));
+    rewrite sym (plusAssociative (plus c (plus bit0 (plus v v))) 
+                 bit1 (plus v1 v1));
     rewrite (plusAssociative c (plus bit0 (plus v v)) bit1);
@@
 
-Thus, this kind of proofs consists of a potentially long sequence of rewriting, where each rewriting step uses one of the available properties.
-Without some specific automation, this sequence of rewriting is done by the user of the proof assistant.
-This is time consuming, and just a little change in the left or the right hand side of the equality will most of the time invalidate completely the proof done by hand.
-This mean that just a little change in our datatypes, or in the definitions of $addBit$ and $adc$ will require us to replace this proof by another one doing the same job, which is assuring the correspondence between the expected index and the given index.
-The re-usability of this kind of proofs is therefore null, because they are performing rewritings for very specific terms.
+Thus, this kind of proof consists of a potentially long sequence of rewriting
+steps,
+where each step uses one of the available properties. Without some
+specific automation, this sequence of rewriting must be done by the programmer.
+This is time consuming, and perhaps more importantly, even a small
+change in the left
+or the right hand side of the equality may invalidate the proof.
+This mean that a minor change in our
+datatypes, or in the definition of $addBit$ or $adc$ will require us to
+replace this proof by another one doing the same job.
+This kind
+of proof is therefore not reusable, because it performs rewritings for
+very specific terms.
 
-If we look at the proof we've written by hand, we realize that we're only using the existence of neutral elements (this is the lemma $plusZeroRightNeutral$), the associativity and commutativity of $+$.
-Thus, we're rewriting a term by using the properties of a commutative monoid.
+If we look at the proof we have written by hand, we realize that we only use
+the existence of neutral elements (this is the lemma $plusZeroRightNeutral$),
+and the associativity and commutativity of $+$. Thus, we're rewriting a term by
+using the properties of a commutative monoid.
 
-We would like to let a prover for commutative monoid find a proof on its own.
-This is something doable because algebraic structures like monoids, groups and rings (all commutative or not) have all a very nice property, which is that every element in them can be normalised to a canonical representation.
-Thus these kind of proofs can be automatically produced by simply computing the normal forms for each side of the equality, and then comparing them with the syntactical equality --Leibniz equality--, because when they will be in normal form, being equal will just be a matter of being syntactically the same entity.
+We would like to let a generic prover for commutative monoid find a proof on its own.
+This should be achievable, because algebraic structures like monoids, groups and
+rings (commutative or otherwise) have a very useful property, that
+every expression can be normalised to a canonical representation. Thus
+these kind of proofs can be automatically produced by computing the
+normal forms for each side of the equality, and then comparing them using
+syntactical equality. When they are in normal
+form, being equal is a matter of being syntactically the same entity.
 
@@
 
-Algebraic provers, like ring provers, are already implemented for various proof assistants, including Coq and Agda.
-In this paper, we describe a certified implementation of an automatic prover for equalities in various algebraic structures, including Monoid, Groups and Rings (all potentially commutative), for the Idris language.
-The novelty is in the approach that we follow, using a new kind of type-safe reflection.
-Working by reflection for implementing tactics has been done several times, as in [ref 1] for the implementation of a ring solver for the Coq proof assistant, but their reflection mechanism  doesn't provide any guarantee, contrary to our type-safe reflection.
-We will compare our approach with other implementations of this kind of tactics in section 6.
-\\
-\\
-Our contribution mainly consists of three points :
+Ring solvers are already implemented for various proof
+assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda/\%22?n=Libraries.UsingTheRingSolver}}. 
+In this paper, we describe a certified
+implementation of an automatic prover for equalities in a
+\emph{hierarchy} of algebraic
+structures, including Monoid, Groups and Rings (all potentially commutative),
+for the Idris language.
+
+The principal novelty is in the approach that we follow, using a
+new kind of \emph{type-safe reflection}.  Working by reflection for implementing
+tactics has been done several times, for the implementation of a
+ring solver for the Coq proof assistant, but without
+providing any guarantees, contrary to our type-safe reflection. We will
+compare our approach with other implementations in Section~\ref{sect:discuss}.
+
+We make the following contributions:
+
 \begin{itemize}
-	\item 1) We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing an easy way to pull out the proofs easily, and providing two guarantees.
-The first one is that the reflection of a term is guaranteed to be a faithful representation of the term.
-The second is that any generated proof is guaranteed to be a proof of the required property.
-The basic ideas of the technique are first presented in the section 2 on a smaller problem with only natural numbers and additions, and are later adapted for the hierarchy of tactics in section 3 (and section 4 contains some implementation details).
-	\item 2) The normalisation procedures are implemented by following a correct by construction approach, instead of implementing a normalization procedure, and proving afterwards that this function effectively computes a normal form.
-This approach is much more suitable for programming languages like Idris.
-Again, this approach will be presented in section 2 on the smaller problem, and on sections 3 and 4 for the complete hierarchy of algebraic structures.
-	\item 3) We develop a hierarchy of tactics where each tactic reuses the one from the immediately above and simplest structure.
-For example, simplifying neutral elements is only implemented at the monoid level, and each level above monoid reuses it.
-Re-usability can be quite tricky to obtain when we want to reuse a prover from a structure which is less expressive.
-For example, reusing the monoid prover for building the group prover is not trivial, since we lose the possibility to express negations ($-x$) and minus operations ($x-y$).
-Some specific encodings will be presented in section 4 for solving this problem.
+        \item We present a new type-safe reflection mechanism, where the
+        reflected terms are indexed over the concrete inputs, thus providing a
+        direct way to pull out the proofs, and providing two guarantees.
+        The first guarantee is that the reflection of a term is guaranteed to be a
+        faithful representation of the term. The second is that any generated
+        proof is guaranteed to be a proof of the required property. The basic
+        ideas of the technique are first presented in Section~\ref{sect:ideas} on a
+        smaller problem with only natural numbers and addition, and are later
+        adapted for the hierarchy of tactics in Section~\ref{sect:general}. 
+        Section~\ref{sect:reuse} contains some implementation details.
+
+        \item The normalisation procedures are implemented by following a
+        correct by construction approach, instead of implementing a
+        normalization procedure, and proving afterwards that this function
+        effectively computes a normal form. This approach is much more suitable
+        for programming languages like Idris. Again, this approach will be
+        presented in Section~\ref{sect:ideas} on the smaller problem, 
+        and in Sections~\ref{sect:general} and~\ref{sect:reuse}
+        for the complete hierarchy of algebraic structures.
+
+        \item We develop a hierarchy of tactics where each tactic reuses the
+        rewriting machinery from the structure immediately above in the hierarchy.
+        For example,
+        simplifying neutral elements is only implemented at the monoid level,
+        and each level above monoid reuses it. Re-usability can be quite tricky
+        to obtain when we want to reuse a prover from a structure which is less
+        expressive. For example, reusing the monoid prover for building the
+        group prover is not trivial, since we lose the possibility to express
+        negations ($-x$) and minus operations ($x-y$). Some specific encodings
+        will be presented in Section~\ref{sect:reuse} for solving this problem.
+
 \end{itemize}

### 2016-01-18 paper-JFP-DTP/content/5-results.tex
Results

\section {Results}

We've implemented a hierarchy of tactics for proving equalities in various algebraic structures. The implementation uses a type-safe reflection mechanism where reflected expressions are indexed over idris expressions, which enables to pull out the proof easily from the index and to write the reduction procedures on a correct-by-construction way. This approach removes much of the difficulty of the construction of the proof. All the proofs we had to do during our implementation were straightforward.

In section 1, we've presented a small example with binary numbers and their addition, for which the user had to prove the lemma $adc\_lemma\_2$ by hand. We're now almost ready to use our commutative monoid prover for building automatically a proof of the desired equality.

\begin{figure}[H]
\figrule
\begin{center}
\begin{lstlisting}
plus (plus c (plus bit (plus v (plus v 0)))) (plus bit (plus v (plus v 0))) 
   = plus v2 (plus (plus (plus v1 v) v) (plus (plus (plus v1 v) v) 0))
\end{lstlisting}
\end{center}
\figrule
\end{figure}

In a context we're we've got the following induction hypothesis \[ihn:(plus\ (plus\ c\ bit)\ bit1\ =\ plus\ vLsb\ (plus\ vCarry0\ (plus\ vCarry0\ 0)))\]
The goal can't be solved directly by calling commutative monoid prover, as it uses different variables in its left and right hand side. However, the induction hypothesis $ihn$ --obtained with the recursive call-- creates a link between these variables. Therefore, we can use our reduction procedure for commutative monoid four times : two for reducing the LHS and RHS of $ihn$, and two for reducing the LHS and RHS of the goal. After these four normalisations, it becomes possible to simply rewrite the new induction impothesis $ihn'$ on the current goal.

\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
     LHS_{ihn} & RHS_{ihn} \\
     LHS'_{ihn} & RHS'_{ihn} \\};
  \path
    (m-1-1) edge [-stealth] node [left] {$norm$}
    						node [right] {$=$} (m-2-1)

    (m-1-1.east|-m-1-2) edge node [above] {$=$} 
    						 node [below] {$ihn$} (m-1-2)

    (m-2-1.east|-m-2-2) edge [dashed,-] node [above] {$=$}
										node [below] {$ihn'$} (m-2-2)

    (m-1-2) edge [-stealth] node [right] {$norm$}
    						node [left] {$=$} (m-2-2);  
\end{tikzpicture}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
  {
     LHS_{goal} & RHS_{goal} \\
     LHS'_{goal} & RHS'_{goal} \\
     LHS''_{goal} \\};
  \path
    (m-1-1) edge [-stealth] node [left] {$norm$}
							node [right] {$=$} (m-2-1)

            edge [dashed,-] node [above] {$?=$} (m-1-2)

    (m-2-1.east|-m-2-2) edge [dashed,-] node [below] {$?=$} (m-2-2)
    (m-2-1) edge [-stealth] node [left] {$ihn'$}
							node [right] {$=$} (m-3-1)

    (m-1-2) edge [-stealth] node [left] {$=$}  
							node [right] {$norm$} (m-2-2)

    (m-3-1) edge [double] node [right] {$=$} (m-2-2);    
\end{tikzpicture}

This example was a bit unconventional in the sense that 4 calls to the normalisation function were needed. Usually, only two calls to the normalisation are enough, one for the left hand side and one for the right hand side, and they are done in the decision procedure $commutativeMonoidDecideEq$. These two extra calls were needed because of the known equality that we had from the induction hypothesis. This example\footnote{Which can be found in $RingIdris/Examples/binary.idr$}, which was therefore at the boundary to deduction modulo [REF], shows that our implementation can deal with concrete goals obtained during the development of real programs.

### 2016-01-18 paper-JFP-DTP/content/4-detailsNormalisation.tex
Section 4 finished

@@
-\section {Normalisation functions and re-usability of the provers}
+\section{Normalisation functions and Reusability of Provers}
+
+\label{sect:reuse}
 
@@
 
-\subsection {Normalization of terms in a Monoid}
+\subsection {Reusing the semigroup prover for building the monoid prover}
 
@@
 
-%\begin{figure}[H]
-%\figrule
-%\begin{center}
-%\begin{lstlisting}
-%elimZero : (p:Monoid c)->(gam:Vect n c)->{c1:c} 
-%            -> (ExprMo p gam c1) 
-%            -> (c2 ** (ExprMo p gam c2, c1 $\simeq$ c2))
-%elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
-%        with (eq_on_set p Zero const1)
-%   elimZero c p gam (PlusMo (ConstMo _ _ Zero) (VarMo _ v)) 
-%        | (Just refl) = (_ ** (VarMo _ v, ?MelimZero1))
-%   elimZero c p gam (PlusMo (ConstMo _ _ const1) (VarMo _ v)) 
-%        | _ = (_ ** (PlusMo (ConstMo _ _ const1) (VarMo _ v), refl)) 
-%elimZero c p gam 
-% (PlusMo (VarMo _ v) (ConstMo _ _ const2)) 
-%      with (eq_as_elem_of_set p Zero const2) 
-%   elimZero c p gam (PlusMo (VarMo _ v) 
-%            (ConstMo _ _ Zero)) | (Just refl) 
-%          = (_ ** (VarMo _ _ v, ?MelimZero2))
-%   elimZero c p gam (PlusMo (VarMo _ v) 
-%            (ConstMo _ _ const2)) | _ 
-%          = (_ ** (PlusMo (VarMo _ v) 
-%                 (ConstMo _ _ const2), refl))
-%[...]
-%\end{lstlisting}
-%\end{center}
-%\caption{Eliminating zeros in a Monoid, first pattern}
-%\figrule
-%\end{figure}
-
 The normalization of reflected expressions in a monoid is simply made of a call to the function reducing terms on a semigroup, followed by a call to elimZero that will eliminates the remaining additions with zeros.
-
-\subsection {Reusing the monoid prover for building the group prover}
 
@@
 
-The idea : negations $-x$ are encoded as variables.
+The last remaining step is to call the monoid prover.
+Unfortunately, we can't directly write a function of conversion from group to monoid, because in a group we have the possibility to express negations and subtractions, that we do not have in a monoid.
+The idea is that we will encode negations as variables, and we will let the monoid prover deal with them as ordinary variables.
+In order to achieve this goal, we need the following datatype :
+
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+data Variable : {c:Type} -> {n:Nat} -> (Vect n c) -> c -> Type where
+    RealVariable : ($\Gamma$:Vect n c) -> (i:Fin n) -> Variable $\Gamma$ (index i $\Gamma$)
+    EncodingNegOfVar : ($\Gamma$:Vect n c) -> (i:Fin n) -> Variable $\Gamma$ (Neg (index i $\Gamma$))
+\end{lstlisting}
+\end{center}
+\caption{Real variables and encodings}
+\figrule
+\end{figure}
+
+We only need to encode negations of variables, as negations of constants can be simplified into a constant.
+Also, there can't be a negation of something more complicated than an atom (a variable or a constant), because of the treatments we've done beforehand, especially because we've propagated all the negations symbols inward the parenthesis.
+
+The constructor for variable now takes a Variable, instead of taking directly a $(Fin\ n)$.
+\begin{figure}[H]
+\figrule
+\begin{center}
+\begin{lstlisting}
+    VarG : (p:Group c) -> {$\Gamma$:Vect n c} -> {val:c} 
+          -> (Variable $\Gamma$ val) -> ExprG p $\Gamma$ val
+\end{lstlisting}
+\end{center}
+\caption{new Var constructor}
+\figrule
+\end{figure}
+
+Thanks to this encoding, we can now transform an $ExprG$ to an $ExprMo$.
+A constant $(ConstG\ p\ \Gamma\ c1)$ will be transformed into the corresponding constant $(ConstMo\ p\ \Gamma\ c1)$, a $PlusG$ into the corresponding $PlusMo$, a variable $(VarG\ p\ var)$ into the corresponding variable $(VarMo\ p\ var)$, the negation of a constant into the resulting constant, and finally the negation of a variable $i$ into a $(Var\ p\ \Gamma\ (EncodingNegOfVar\ \Gamma\ i))$.
+Once this encoding is computed, the antepenultimate step is to call the monoid prover and to let it finish the normalisation.
+Finally, we need to convert back the result into an $ExprG$ by interpreting the $EncodingNegOfVar$ into a $Neg$ of the corresponding variable.
+
+\subsection {Reusing the group prover for building the commutative group prover}
+
+The normalisation of the commutative group prover starts by calling the group prover.
+Here, the two algebraic structures have the same power of expressivity, and thus there's no specific encoding to design.
+Then, we simply need to decide an ordering for variables and constants.
+We arbitrary decide that $Var\ FZ\ <\ Var\ (FS\ FZ)\ <\ Var\ (FS\ (FS\ FZ))\ <\ ...\ <\ Constant$.
+The normalisation function will therefore simply sort the expression with this ordering and will do basic simplifications.
 
@@
 
-The idea : monomials (containing products), like $3*(x*y)$ are encoded as variables when passed to the commutative group prover.
+The most important specific task of the ring reduction is to develop entirely the expression with the use of distributivity.
+After that, it needs to call the commutative group prover, but again, we're losing some expressivity as it won't be possible to express a product at the commutative group level.
+We follow the same idea that we've implemented for the conversion going from group to monoid, and we will use some specific encoding.
+More precisely, all monomials (containing products), like $3*(x*y)$ are encoded as variables when passed to the commutative group prover.
+This is possible because we've fully developed the expression, which had for effect to ensure that a product can't contain a sum.
+We will add a coonstructor to the type $Variable$ in order to encode these monomials.

### 2016-01-18 paper-JFP-DTP/content/5-results.tex
Section 4 finished

@@
 In a context we're we've got the following induction hypothesis \[ihn:(plus\ (plus\ c\ bit)\ bit1\ =\ plus\ vLsb\ (plus\ vCarry0\ (plus\ vCarry0\ 0)))\]
-The goal can't be solved directly by calling commutative monoid prover, as it uses different variables in its left and right hand side.
+The goal can't be solved directly by calling the commutative monoid prover, as it uses different variables in its left and right hand side.
 However, the induction hypothesis $ihn$ --obtained with the recursive call-- creates a link between these variables.
-Therefore, we can use our reduction procedure for commutative monoid four times : two for reducing the LHS and RHS of $ihn$, and two for reducing the LHS and RHS of the goal.
+Therefore, the solution is simply to use our reduction procedure for commutative monoid four times : two for reducing the LHS and RHS of $ihn$ --that gives a new proof of equality $ihn'$--, and two for reducing the LHS and RHS of the goal.
 After these four normalisations, it becomes possible to simply rewrite the new induction impothesis $ihn'$ on the current goal.
@@
 This example was a bit unconventional in the sense that 4 calls to the normalisation function were needed.
-Usually, only two calls to the normalisation are enough, one for the left hand side and one for the right hand side, and they are done in the decision procedure $commutativeMonoidDecideEq$.
+Usually, only two calls to the normalisation are enough, one for the left hand side and one for the right hand side, and they are done automatically in the decision procedure $commutativeMonoidDecideEq$.
 These two extra calls were needed because of the known equality that we had from the induction hypothesis.
-This example\footnote{Which can be found in $RingIdris/Examples/binary.idr$}, which was therefore at the boundary to deduction modulo [REF], shows that our implementation can deal with concrete goals obtained during the development of real programs.
+This example\footnote{Which can be found in $RingIdris/Examples/binary.idr$}, which was therefore at the boundary to deduction modulo~\cite{Dowek03}, shows that our implementation can deal with concrete goals obtained during the development of real programs.

### 2016-01-18 paper-JFP-DTP/content/1-introduction.tex
update

@@
 
-It should be pointed that the proof component of this function is established
-without any effort, as we always give the ``\_'' symbol for the proof which means
-``infer it!''. For each of the eight patterns, a simple computation provides the
-desired proof, and thus the typechecker can deal with them.  We also note that
+It should be pointed that the proof component of this function is established without any effort, as we always give the "Refl" for the proof.
+Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and "Refl" can therefore be used.
+We also note that
 there is no need to produce any lemma of correctness about $addBit$ afterwards.
@@
 expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we
-provide a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))\]. The
-problem is even worse for the second line where the result is expected to have
-type \[plus\ (plus\ c\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\ (plus\ bit\
-(plus\ v\ (plus\ v\ 0)))\] while we provide a term of type \[plus\
-v2\ (plus\ (plus\ (plus\ v1\ v)\ v)\ (plus\ (plus (plus\ v1\ v)\ v)\ 0))\].
+provide a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0))).\]
+The problem is even worse for the second case where the expected index is 
+\[(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\] while we're trying to provide a term indexed over 
+\[plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))\]
 

### 2016-01-18 paper-JFP-DTP/content/6-discussion.tex
Edit introduction
Remaining to do : perfs test & more related work

@@
 \section {Discussion}
+
+\label{sect:discuss}
 
@@
 
-Coq's ring prover follows the traditional approach of defining functions, and latter on proving many auxiliary lemmas about them.
+Coq's ring prover follows the traditional\footnote{By ``conventional" we have in mind the proof engineering style developed in the Coq'Art ~\cite{BertotC04}}  approach of defining functions, and latter on proving many auxiliary lemmas about them.
 This is particularly adapted to Coq, which has many facilities for the construction of proofs, and an adapted proof mode.
 However, this approach kind of duplicate the work : they first tell the machine how it works, and latter, why it works.
-This separation between the world of computations and the world of logic becomes smaller in our approach with a good use of dependent types, that allows to write more specific types, and thus to capture logical properties.
+This separation between the world of computations and the world of logic becomes smaller in our approach with a fine use of dependent types, that allows to write more specific types, and thus to capture logical properties.
 The writing of functions can therefore be guided by this more precise type, and this is one of the main benefit of the approach we've followed.
 Moreover, we almost get the proof of correctness for free, because every little bit of rewriting done for the normalisation of the reflected terms is accompanied by the logic justification which tells why this rewriting can be done --and locally, this justification is always simple !--.
-In the end, all the proofs we have to do are systematically trivial, since they only contain trivial rewritings with the available hypothesis and the use of the properties of the corresponding algebraic structure.
+In the end, all the proofs we have to do are systematically straightforward, since they only contain rewritings with the available hypothesis and the use of the properties of the corresponding algebraic structure.
 \\
@@
 The correctness of the tactic is in fact obtained by construction.
-Actually, what we really needed was to produce the proof of the equality $a \simeq b$, and not the normalised terms.
+Actually, what we really needed to produce was the proof of the equality $a \simeq b$, and not the normalised terms.
 The normalised term --and more precisely their construction-- was just a support for building the desired proof of $a \simeq b$, which is precisely the proof of correctness.
@@
 Note that the normalization itself could contain some non total functions without breaking the correctness.
-That would just mean that when using this tactic, the tactic could loop infinitely and the user would never get his answer.
+That would just mean that when using this tactic, the tactic could potentially loop infinitely and the user would never get his answer.
 That would of course be bad, but that would not produce any inconsistency --as long as none uses these definitions maliciously to prove something false--, because this non total function would be only used in the world of computations and not in the world of logic.
@@
 Because we've meet these two conditions, we simply can't produce a false proof of a statement, because the typechecker would complain if that would be the case.
-If Idris internal type theory is consistent, then one can't produce a false lemma without either introducing an axiom nor introducing a non total function and latter using it to create inconsistencies in the logic.
+If Idris internal type theory is consistent and if Idris' internal implementation follows it (and we believe in both!), then one can't produce a false lemma without either introducing an axiom nor introducing a non total function and latter using it maliciously to create an inconsistency in the logic.
 
@@
 As long as we do not meet an equality that should be provable, and which isn't proved by our tactic, there is no problem.
-And if this situation does happen, nothing really bad really happened : no inconsistencies have been introduced in the system.
-The tactic just refuses to do it automatically for the programmer.
-The user can still attempt it.
+And if this situation does happen one day, nothing really bad happened : no inconsistencies would have been introduced in the system.
+The tactic would just refuse to do it automatically, but the user could still attempt it.
 \\
 
-If we believe that the normal form computed by the reduction function is indeed a normal form --which means that if two terms are equal according to the properties of an algebraic structure, then their normal form should be equal-- then the tactic is complete.
+In definitive, we want the tactic to be complete, without requiring a formal proof of it.
+If we believe that the normal form computed by the reduction function is indeed a normal form --which means that if two terms are equal according to the properties of the algebraic structure, then their normal form should be equal-- then the tactic is complete.
 Indeed, 
-$meta\_propetry\ :\ \forall\ a\ b,\ a \simeq b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is logically equivalent to
+$isNormalForm\ :\ \forall\ a\ b,\ a \simeq b \rightarrow norm\ (reflect\ a) = norm\ (reflect\ b)$ is logically equivalent to
 
@@
 $\ \forall\ a\ b,\ decideEq\ (reflect\ a)\ (reflect\ b) = Nothing \rightarrow a \not{\simeq} b$.
-The equivalence between $meta\_property$ and $completeness$ can only be done in the meta theory because it is a proof by contradiction, which is equivalent to the law of excluded middle, that we do not have in constructive logics, on which Idris is based.
+This equivalence between $isNormalForm$ and $completeness$ can only be done in the meta theory because it is a proof by contradiction, which is equivalent to the law of excluded middle, that we do not have in constructive logics, on which Idris is based.
+\\
 
-Question : Does $completeness$ is provable in the system ?
-Would it be doable for the little example with natural numbers for example ?
-I want to know if it's something improvable in the theory itself (and in this case why, precisely!), or just something really hard ?
-Anyway, it has never been done for this problem (not for Coq's ring prover, nor for Agda's ring prover). Can I do it for the little example with natural numbers though ? \\
+So far, we haven't encountered any goal that should be provable but which failed to be proven automatically.
+However, if this situation would appear one day, it would certainly come from a small kind of simplification that would have been forgotten somewhere in the hierarchy, and it would be easy to fix it by adding a sub-function that deals with this specific simplification at the appropriated level.
+And all the more specialised provers that reuse this prover will also be fixed at the same time thanks to the compositionality of our provers.
+\\
 
@@
 If we want to go from the reflected term to the original --concrete-- value, we can simply return the index.
-Thus, we have the guarantee that $(reflect\ a)$ is always a faithful representation of $a$ with our approach.
-Coq's approach doesn't have this guarantee, and could potentially lose some completeness if this property is not always true, as mentioned earlier during the comparison with Coq's implementation.
+Thus, unlike Coq, we have the formal guarantee that $(reflect\ a)$ is always a faithful representation of $a$ with our approach, which means that we can't lose any totality because of the $reflect$ function.

### 2016-01-19 paper-JFP-DTP/content/1-introduction.tex
corrections with Edwin advices on latex

@@
 Thus, if $x$ and $y$ are of type $a$, then the type
-$Id_a(x, y)$ represents the proposition "$x$ is equal to $y$". If this type is
+$Id_a(x, y)$ represents the proposition ``$x$ is equal to $y$". If this type is
 inhabited, then $x$ is said to be provably equal to $y$.  Thus, $Id$ is a type
@@
 what is equal by definition should be provably equal.  This is accomplished by
-giving constructors to the type $Id(a,a)$ and nothing when "$a$ is not $b$".
+giving constructors to the type $Id(a,a)$ and nothing when ``$a$ is not $b$".
 One way to implement $Id$ in Idris is:
@@
 propositional equality. That means that the axiom of induction principle has
-made the type $Id(a,b)$ bigger than it looks like (in the definition 1), and
+made the type $Id(a,b)$ bigger than it looks like in the above definition, and
 there are therefore things which are provably equal which unfortunately aren't
@@
 
-It should be pointed that the proof component of this function is established without any effort, as we always give the "Refl" for the proof.
-Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and "Refl" can therefore be used.
+It should be pointed that the proof component of this function is established without any effort, as we always give the ``Refl" for the proof.
+Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and ``Refl" can therefore be used.
 We also note that
@@
 \caption{Proof by hand of the correspondence between expected and provided types for the second pattern of the definition of adc}
+\label{adc_lemma_2}
 \figrule

### 2016-01-19 paper-JFP-DTP/content/2-basicIdeas.tex
corrections with Edwin advices on latex

--- 2016-01-18 paper-JFP-DTP/content/2-basicIdeas.tex
@@
-\section{A simplified problem : type safe reflection for proving equalities with universally quantified natural numbers and additions}
+\section{A Simplified Problem: Proving Equalities with Natural Numbers and Additions}
+
+\label{sect:ideas}
 
@@
 \\
-Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce a proof of this equality if appropriate, where "equal" has the meaning "syntactically equal or equal thanks to the associativity and right neutral properties".
+Thus, in definitive, in this section, we want to write a decision procedure, able to tell if two expressions composed of universally quantified natural numbers and additions of these numbers are equal, and to produce a proof of this equality if appropriate, where ``equal" has the meaning syntactically equal or equal thanks to the associativity and right neutral properties.
 
@@
 In fact, such a normalisation function can't be written directly, because in the LHS and RHS of $x=y$, we potentially have variables which have been universally quantified.
-And the normalization function needs to do different treatments for a "variable natural number" (a number which has been universally quantified) and for the constant $Z$.
+And the normalization function needs to do different treatments for a ``variable natural number" (a number which has been universally quantified) and for the constant $Z$.
 This is not possible yet, because once the variables are abstracted, they are just ordinary values of type $Nat$, and there is no way to distinguish them.
@@
 For this reason, we will work by reflection.
-In this little example, it means that we will define a datatype that will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of "variable numbers", $Z$, and additions of these things.
+In this little example, it means that we will define a datatype that will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of ``variable numbers", $Z$, and additions of these things.
 This datatype will let us inspect the internal structure of a number by pattern matching.
 Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't what we needed.
-With the first approximation of the data type $Expr$ presented in the Fig 6, we will be allowed to pattern match an encoding of number against the constructors $Plus$, $Var$ and $Zero$.
+With the first approximation of the data type $Expr$ presented in the figure~\ref{reflectedNaturalNumbers0}, we will be allowed to pattern match an encoding of number against the constructors $Plus$, $Var$ and $Zero$.
 
@@
 \caption{First version of reflected natural numbers}
+\label{reflectedNaturalNumbers0}
 \figrule
@@
 \end{center}
-\caption{Reflected LHS of example 1 (with the first version of reflected natural numbers)}
+\caption{Reflected LHS with the first version of reflected numbers showed in \ref{reflectedNaturalNumbers0}}
 \figrule
@@
 \caption{Second version of reflected number with embedded denotation}
+\label{reflectedNaturalNumbers}
 \figrule
@@
 
-For an expression $ex\ :\ Expr\ \Gamma\ x$, we will say that "$ex$ denotes (or encodes) the number $x$ in the context $\Gamma$".
+For an expression $ex\ :\ Expr\ \Gamma\ x$, we will say that ``$ex$ denotes (or encodes) the number $x$ in the context $\Gamma$".
 When an expression is a variable, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
@@
 
-We want to write the reduction function on a "correct by construction" way, which means that no additional proof should be required after the definition of the function.
+We want to write the reduction function on a ``correct by construction" way, which means that no additional proof should be required after the definition of the function.
 Thus, $reduce$ will produce the proof that the new Expr freshly produced has the same interpretation as the original Expr, and this will be made easier by the fact that the data type Expr is now indexed over the real --concrete-- number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
@@
 \caption{Syntactical equality between reflected terms}
+\label{syntacticalEq}
 \figrule
@@
 \caption{Building the desired proof with the two proofs of equality}
+\label{buildProof}
 \figrule
@@
 \caption{buildProof metavariable}
+\label{MbuildProof}
 \figrule
@@
 \caption{testEq}
+\label{testEq}
 \figrule
@@
 \caption{Reflected left associative numbers}
+\label{LExpr}
 \figrule
@@
 \caption{Production of the left associative form}
+\label{expr_l}
 \figrule
@@
 \caption{Proof of the metavariable MplusLExpr}
+\label{MplusExpr}
 \figrule
@@
 \caption{Going back from LExpr to Expr}
+\label{l_expr}
 \figrule
@@
 \caption{Reduction function}
+\label{reduce}
 \figrule
@@
 
-\subsection{Usage of the "tactic"}
+\subsection{Usage of the ``tactic"}
 
-It's now time to see how to use this minimalist "tactic".
+It's now time to see how to use this minimalist ``tactic".
 Let's define two expressions $e1$ and $e2$, respectively representing the numbers $((x + y) + (x + z))$ and $(x + ((y + x) + z))$ in the context $[x, y, z]$ of three abstracted variables.
@@
 \caption{Two test expressions}
+\label{e1_e2}
 \figrule
@@
 \caption{Generating automatically the desired proof}
+\label{e1_e2_testEq}
 \figrule
@@
 \caption{Obtained proof}
+\label{obtainedProof}
 \figrule
@@
 \subsection{Construction of the reflected terms}
+\label{sect:ReflectNat}
 
@@
 
-We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our "prover" works, but that's not enough for being used intensively.
+We still want an automatic way of constructing these reflected terms because what we have so far is enough to demonstrate that our prover works, but that's not yet enough for being used intensively.
 We want to program an automatic way of going from the concrete values (of type $Nat$), to the reflected terms (of type $Expr$).
@@
 \caption{Type of the function for reflecting natural numbers}
+\label{reflectNatType}
 \figrule
@@
 \caption{Reflecting natural numbers - pattern for zero}
+\label{reflectNat_pattern1}
 \figrule
@@
 \caption{Reflecting natural numbers - pattern for a plus}
+\label{reflectNat_pattern2}
 \figrule
@@
 \caption{Weakening function}
+\label{weaken}
 \figrule
@@
 \end{center}
-\caption{Conversion of Fin}
+\caption{Conversion of a Fin}
+\label{convertFin}
 \figrule
@@
 \caption{Reflecting natural numbers - pattern for a variable}
+\label{reflectNat_pattern3}
 \figrule
@@
 \caption{Returns the index and the proof that this index is the right one}
+\label{isElement}
 \figrule

### 2016-01-19 paper-JFP-DTP/content/3-generalProblem.tex
corrections with Edwin advices on latex

@@
-\section {Back to the general problem : A hierarchy of tactics}
+\section{The General Problem: A Hierarchy of Tactics}
+
+\label{sect:general}
 
@@
 All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants.
-For this reason, we define a notion of "set", which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$.
+For this reason, we define a notion of ``set\footnote{This notion of ``Set" isn't a formalisation of sets and their elements, but is only the Setoid that we need} ", which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$.
 All the algebraic structures will later extend this type class :
@@
 \caption{Set}
+\label{Set}
 \figrule
@@
 \caption{Magma}
+\label{Magma}
 \figrule
@@
 In fact, all operations, like the Plus here, need to be "compatible" with the equality : $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow (Plus\ c1\ c2) \simeq (Plus\ c1'\ c2')$.
-This need only comes from the fact that we're dealing with any equivalence relation
+This need only comes from the fact that we're dealing with any equivalence relation.
 
-A bit more interesting is the Semi-Group type class. A semi-Group is a Magma (ie, it still has a $Plus$ operation), but moreover it has the property of associativity for this operation.
+A bit more interesting is the Semi-Group typeclass. A Semi-Group is a Magma (ie, it still has a $Plus$ operation), but moreover it has the property of associativity for this operation.
 
@@
 \caption{Semi-Group}
+\label{SemiGroup}
 \figrule
@@
 \caption{Monoid}
+\label{Monoid}
 \figrule
@@
 We must have the property that $Minus$ can always be simplified with the $Neg$ and the $Plus$.
-That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ -b)$.
+That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ (-b))$.
 We let the possibility to express a $Minus$ just for convenience for the user, as he won't have to rewrite by hand his $Minus$ operations with sums of negations.
@@
 \caption{Group}
+\label{Group}
 \figrule
@@
 
-Note that the properties will either be obtained "by implementation" if the operations are real --computable-- functions, or by axioms if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$, ...) are defined as axioms.
+Note that the properties will either be obtained by implementation if the operations are real --computable-- functions, or by axioms if the user is working in an axiomatised theory where the operations ($Plus$, $Neg$, ...) are defined as axioms.
 
@@
 \caption{Reflected terms for a Magma}
+\label{ExprMa}
 \figrule
@@
 
-%For the moment, the definition of $Variable$ is just the following:
-%
-%\begin{code}[caption=Reflected variables, captionpos=b, label=lst1:haskell2]  
-%data Variable : {c:Type} -> {n:Nat}
-%  -> (c_equal : (c1:c)->(c2:c)->Maybe(c1=c2)) 
-%  -> (Vect n c) -> c -> Type where
-%    RealVariable : (c_equal:(c1:c)->(c2:c)
-%                    ->Maybe(c1=c2))
-%          -> ($\Gamma$:Vect n c) -> (i:Fin n) 
-%          -> Variable c_equal $\Gamma$ (index i $\Gamma$) 
-%\end{code}	
-%
-%This type will be extended with some other constructors later on. For the moment, we can only %express a "real variable", referred by its index $i$. \\
-%\\
 Because there is no additional operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
@@
 data ExprG :  Group c $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
-    ConstG : (p : dataTypes.Group c) $\rightarrow$ ($\Gamma$:Vect n c) 
+    ConstG : (p : Group c) $\rightarrow$ ($\Gamma$:Vect n c) 
          -> (c1:c) $\rightarrow$ ExprG p $\Gamma$ c1
@@
 \caption{Reflected terms in a Group}
+\label{ExprG}
 \figrule
@@
 \caption{Type of the reduction function for terms reflecting elements in a Group}
+\label{groupReduce}
 \figrule
@@
 \caption{Type of the function for deciding equality between elements in a Group}
+\label{groupDecideEqType}
 \figrule
@@
 \caption{Decides if two elements of a group are equal}
+\label{groupDecideEq}
 \figrule
@@
 \caption{Composes the two proofs if the normal forms are the same}
+\label{buildProofGroup}
 \figrule
@@
 \caption{Type of the function doing the reflection for constants}
+\label{ConstantsRingReflector}
 \figrule
@@
 \caption{Automatic reflection to ring}
+\label{reflectRingTermType}
 \figrule
@@
 
-This function will work very similarly to the reflection function for $Nat$ that we've writen in section 2.5.
+This function will work very similarly to the reflection function for $Nat$ that we've writen in section~\ref{sect:ReflectNat}.
 Here, there are just more constructions to deal with, because $c$ is equipped with negations, subtractions and multiplications.
@@
 \caption{Automatic reflection - case of a multiplication}
+\label{reflectRingTerm_pattern1}
 \figrule
@@
 \caption{Automatic reflection - case of constants and variables}
+\label{reflectRingTerm_pattern2}
 \figrule
@@
 It is remarkable that the reflected term is guaranteed to be a faithful representation of the input.
-We've been able to define this function with a very strong and precise type because Idris has a $\%syntax$ command which let us do arbitrary pattern matching on syntax, and this $\%syntax$ produces functions which are typed, unlike Ltac terms in Coq.
+We've been able to define this function with a very strong and precise type because Idris has a ``$\%syntax$" command which let us do arbitrary pattern matching on syntax, and the functions defined with ``$\%syntax$" are typed, unlike Ltac terms in Coq.

### 2016-01-19 paper-JFP-DTP/content/6-discussion.tex
little fix
corrections with Edwin advices on latex

@@
 \caption{Decision procedure of the tactic with a "traditional" approach}
+\label{decideEqTrad}
 \figrule
@@
 
-Both with this "traditional" approach and with our type-safe reflection, the activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
+Both with this ``traditional" approach and with our type-safe reflection, the activity of proving has effectively been replaced by the task of running a function, and a simple evaluation is now enough to produce the desired proof.
 The main advantage of our approach compared to this kind of traditional approach is that we don't need external lemmas, and more generally that the proof of correctness is obtained a lot more easily.
@@
 
-In Coq's latest implementation of a Ring solver, described in~\cite{Coq2005}, they almost followed what has just been desribed as a "traditional" approach with the use of many auxiliary lemmas.
+In Coq's latest implementation of a Ring solver, described in~\cite{Coq2005}, they almost followed what has just been desribed as a ``traditional" approach with the use of many auxiliary lemmas.
 The few differences essentially lie in the way the function $reflect$ is defined.
@@
 \item
-if it couldn't unify the goal $a \simeq b$ with $reify\ (reflect\ a)\ \simeq\ reify\ (reflect\ b)$, then it means that there is something wrong in the definitions of the functions $reflect$ and $reify$, and it could print something like "The ring prover has failed unexpectedly.
+if it couldn't unify the goal $a \simeq b$ with $reify\ (reflect\ a)\ \simeq\ reify\ (reflect\ b)$, then it means that there is something wrong in the definitions of the functions $reflect$ and $reify$, and it could print something like ``The ring prover has failed unexpectedly.
 There's something wrong with the implementations of the $reflect$ and $reify$ functions".
@@
 
-Coq's ring prover follows the traditional\footnote{By ``conventional" we have in mind the proof engineering style developed in the Coq'Art ~\cite{BertotC04}}  approach of defining functions, and latter on proving many auxiliary lemmas about them.
-This is particularly adapted to Coq, which has many facilities for the construction of proofs, and an adapted proof mode.
+Coq's ring prover follows the traditional\footnote{By ``traditional" we have in mind the proof engineering style developed in the Coq'Art ~\cite{BertotC04}}  approach of defining functions, and latter on proving many auxiliary lemmas about them.
+This is particularly adapted to Coq, which has many facilities for the construction of proofs, and a powerful proof mode.
 However, this approach kind of duplicate the work : they first tell the machine how it works, and latter, why it works.

### 2016-01-19 paper-JFP-DTP/content/1-introduction.tex
more fix

@@
         normalization procedure, and proving afterwards that this function
-        effectively computes a normal form. This approach is much more suitable
+        is correct. This approach is much more suitable
         for programming languages like Idris. Again, this approach will be
@@
         group prover is not trivial, since we lose the possibility to express
-        negations ($-x$) and minus operations ($x-y$). Some specific encodings
+        negations ($-x$) and subtractions ($x-y$). Some specific encodings
         will be presented in Section~\ref{sect:reuse} for solving this problem.

### 2016-01-19 paper-JFP-DTP/content/1-introduction.tex
new fix

@@
 \subsection{Propositional and Judgemental Equalities in Intentional Type Theories}
-In Mathematics, equality is a proposition, which can hold or otherwise, or be disproved.
+In Mathematics, equality is a proposition, e.g. we can disprove an equality or 
+assume an equality as a hypothesis.
 Since in type theory, propositions are seen as types~\cite{How80}, the
@@
 $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$. For
-convenience, we write $Id_a\ x\ y$ as $x\ =_a\ y$.  This equality is the
+convenience, we write $(Id_a\ x\ y)$ as $(x\ =_a\ y)$. This equality is the
 equality which can be manipulated in the language.
 
-There is another more
-primitive notion of equality in Intentional Type Theories, called judgemental
+There is another, more
+primitive, notion of equality in Intentional Type Theories, called judgemental
 equality, or definitional equality. This second equality means ``equal by
@@
 of functions and reductions, until no more reduction can be
-performed. We denote definitional equality by $\equiv$.
+performed. We denote the definitional equality by $\equiv$.
 
@@
 (by definition!) $y$. In this case, the constructor $Refl$ helps to create a
-proof of this equality : $Refl\ x$ is precisely the proof which says that
+proof of this equality : $(Refl\ x)$ is precisely the proof which says that
 $x=_ax$.
@@
 
-It should be pointed that the proof component of this function is established without any effort, as we always give the ``Refl" for the proof.
+It should be pointed that the proof component of this function is established without any effort, as we always give ``Refl" for the proof.
 Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and ``Refl" can therefore be used.
@@
 bit. This result represents the value $c + x + y$, where $c$ 
-representis the value of the input carry bit, and $x$ and $y$ represent
+represents the value of the input carry bit, and $x$ and $y$ represent
 the values of the two binary numbers given in input. Thus, we'd like to
@@
 new kind of \emph{type-safe reflection}.  Working by reflection for implementing
-tactics has been done several times, for the implementation of a
+tactics has been done several times, including for the implementation of a
 ring solver for the Coq proof assistant, but without
@@
         \item We develop a hierarchy of tactics where each tactic reuses the
-        rewriting machinery from the structure immediately above in the hierarchy.
+        rewriting machinery from the structure immediately below in the hierarchy.
         For example,

### 2016-01-19 paper-JFP-DTP/content/2-basicIdeas.tex
new fix

@@
 For this little example, the reduction procedure won't be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more properties to deal with and it will certainly become more problematic to unfold the definition of a gigantic reduction procedure.
-We want to avoid pain as much as possible, and for this reason, we don't want to have to write big proofs, and even to avoid the writing of proofs as much as possible.
+We want to avoid pain as much as possible, and for this reason, we want to avoid the writing of big proofs, and even to avoid the writing of proofs as much as possible.
 \\

### 2016-01-19 paper-JFP-DTP/content/2-basicIdeas.tex
Almost done. futur work to write

@@
 For example, extending the context $[v,\ w,\ x]$ with $[y,\ z]$ produces the context $[v,\ w,\ x,\ y,\ z]$.
-The variables $v$, $w$ and $x$, which were refereed to as the first, second and third variables in $\Gamma$, are still the first, second and third variables in the complete context.
-That means that a variable $Var\ i$ will still be a $Var\ i$, the position $i$ hasn't changed by augmenting the context.
+The variables $v$, $w$ and $x$, which were respectively refereed to as the first, second and third variables in $\Gamma$, are still the first, second and third variables in the complete context.
+That means that a variable $Var\ i$ will still be a $Var\ i$ after the weakening, the position $i$ hasn't changed by augmenting the context.
 
@@
 As for the case where $t$ wasn't in the original context (the $Nothing$ case), we will need to prove that we can convert $(index\ (lastElement'\ n)\ (\Gamma ++\ [t]))$ into $t$.
-This is doable because we can prove the lemma $elemInBigerVect\ :\ \{T:Type\}\ \rightarrow \{v:Vect\ n\ T\} \rightarrow (i:Fin\ n) \rightarrow (elem:T) \rightarrow (proofInside : index\ i\ v\ =\ elem) \rightarrow (head:T) \rightarrow (index\ (FS\ i)\ (head::v)\ =\ elem)$.
+This is doable because we can prove the lemma $indexOfLastElem\ :\ \{T:Type\}\ \rightarrow\ \{n:Nat\}\ \rightarrow\ (v:Vect\ n\ T)\ \rightarrow\ (x:T)\ \rightarrow\ index\ (lastElement'\ n)\ (v++[x])\ =\ x$
 

### 2016-01-19 paper-JFP-DTP/content/3-generalProblem.tex
Almost done. futur work to write

@@
 We want to give the user the possibility to define his own equivalence notion, assuming he's able to provide the properties of the structure he wants to use.
-This equivalence relation on $c$ has the following profile $(\simeq) : c \rightarrow y \rightarrow Type$\footnote{This 'Type' would be a 'Prop' in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+This equivalence relation on $c$ has the following profile $(\simeq) : c \rightarrow c \rightarrow Type$\footnote{This 'Type' would be a 'Prop' in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
 
@@
 For this reason, we define a notion of ``set\footnote{This notion of ``Set" isn't a formalisation of sets and their elements, but is only the Setoid that we need} ", which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$.
-All the algebraic structures will later extend this type class :
+All the algebraic structures will later extend this typeclass :
 
@@
 This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equiped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.
-In fact, all operations, like the Plus here, need to be "compatible" with the equality : $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow (Plus\ c1\ c2) \simeq (Plus\ c1'\ c2')$.
-This need only comes from the fact that we're dealing with any equivalence relation.
+In fact, there is an additional requirement which is that all operations, like the Plus here, need to be "compatible" with the equality : $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((Plus\ c1\ c2) \simeq (Plus\ c1'\ c2'))$.
+This requirement comes from the fact that we're dealing with any equivalence relation.
+The user is free to use his own equivalence relation, but it should be compatible with the operations that he is using.
 
@@
 \end{figure}
-We can continue the hierarchy with groups, which are a more interesting.
+We can continue the hierarchy with the more interesting structure of Group.
 A group is a monoid, with two new operations.
@@
 That means that $Minus$ is not a "primitive" operation of a Group, since we can always rewrite $(a-b)$ into $(a\ +\ (-b))$.
-We let the possibility to express a $Minus$ just for convenience for the user, as he won't have to rewrite by hand his $Minus$ operations with sums of negations.
-The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an inverse is the following property :
+We let the possibility to express a $Minus$ just for convenience for the user, so that he won't have to rewrite by hand his $Minus$ operations with sums of negations.
+The second axiom, which is the most important one for a Group, is the fact that any value $c1$ admits $Neg\ c1$ for inverse, where being an additive inverse is the following property :
 
@@
 
-These type classes will somehow be used as predicates on types.
-In order to call a prover on a type $c$, the user of the system will have to satisfy the corresponding type class by providing an instance of it for his type $c$, ie, he will have to prove that the corresponding properties effectively hold for $c$.
+These typeclasses will somehow be used as predicates which classify types.
+In order to call a prover on a type $c$, the user of the system will have to satisfy the corresponding typeclass by providing an instance of it for his type $c$, ie, he will have to prove that the corresponding properties effectively hold for $c$.
 
@@
 Each of these datatype is parametrised over a type $c$, which is the real type on which we want to prove equalities (the $carrier$ type).
-It is also indexed over an instance of the corresponding type class for $c$ (we usually call it $p$, because it behaves as a $proof$ telling that the structure $c$ has the desired properties), and indexed over a context (a vector $\Gamma$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the concrete value being encoded.
+It is also indexed over an instance of the corresponding typeclass for $c$ (we usually call it $p$, because it behaves as a $proof$ telling that the structure $c$ has the desired properties), and indexed over a context (a vector $\Gamma$ of $n$ elements of type $c$), and also indexed over a value of type $c$, which is precisely the concrete value being encoded.
 A magma is only equipped with one operation $Plus$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.
@@
 When we encode a constant $c1$ in a context $\Gamma$, we use the constructor $ConstMa$ to produce a term of type $ExprMa\ p\ \Gamma\ c1$ : the index representing the concrete value is precisely this constant $c1$.
-If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (ie, a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation defined in the current instance $p$.
+If e1 is an expression of type $ExprMa\ p\ \Gamma\ c1$ (ie, a term encoding the value $c1$), and $e2$ is an expression of type $ExprMa\ p\ \Gamma\ c2$ (ie, a term encoding the value $c2$), then the term $PlusMa\ e1\ e2$ will have the type $ExprMa\ p\ \Gamma\ (Plus\ c1\ c2)$, ie, this term will encode the value $(Plus\ c1\ c2)$, where $Plus$ is the operation available in the current instance $p$.
 
@@
 Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
-These functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
+These normalisation functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
 

### 2016-01-19 paper-JFP-DTP/content/5-results.tex
corrections with Edwin advices on latex
Almost done. futur work to write

@@
 \section {Results}
+\label{sect:results}
 
@@
 \end{center}
+\label{goal}
 \figrule
@@
 These two extra calls were needed because of the known equality that we had from the induction hypothesis.
-This example\footnote{Which can be found in $RingIdris/Examples/binary.idr$}, which was therefore at the boundary to deduction modulo~\cite{Dowek03}, shows that our implementation can deal with concrete goals obtained during the development of real programs.
+This example\footnote{Which can be found in $RingIdris/Provers/binary.idr$}, which was therefore at the boundary to deduction modulo~\cite{Dowek03}, shows that our implementation can deal with concrete goals obtained during the development of real programs.

### 2016-01-19 paper-JFP-DTP/content/6-discussion.tex
Almost done. futur work to write

@@
 
-[Mentioning performances, that might be in favour of Coq]
+In terms of performances, our implementation is fast enough for providing answers in a decent amont of time --usually a few seconds--.
+The automatic proof of the lemma $adc\_lemma\_2$ (presented in section~\ref{sect:motivatingExample} and proven automatically in section ~\ref{sect:results}) is constantly generated and printed in less than 8 seconds on a dual core i5 processor @ 2.4ghz and is therefore perfectly usable, even for this kind of moderate size goals that come from a real application.
+Knowing that the display of the proof takes most of the time, it would be interesting to evaluate precisely the time needed for its generation only, without outputting the result.
+However, we suspect that our implementation will be slower than Coq, because they've used a more efficient representation of normalised polynomials.
+It might be possible that we've traded some efficiency against some simplicity and compositionality.
+
+	\subsection{Other related work}
+
+As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism.
+Their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
+Because $Ltac$ is untyped, it would not permit to define --as we did-- a dependently typed function $reflect$ that is guaranteed to return a faithful encoding of the given input value.
+$Mtac$~\cite{Ziliani13} is an extension to Coq that supports dependently-typed tactic programming, which uses monad to avoid the need to touch the trusted kernel typechecker of Coq.
+\\
+
+Doing proofs by reflection has been intensively investigated by many authors, including Adam Chlipala, in particular in his book~\cite{ChlipalaBook} and in~\cite{Malecha14}.
+However, we haven't found anything similar to the type-safe reflection technique that we've presented in this paper.
+\\
+
+At the connexion bewteen proof automation and proof by reflection, we can cite SSreflect~\cite{GonthierTuto}, which is an extension of Coq which aims to provide additional formalisation and tactics suitable for long mathematical proofs.
+\\
+
+Proof automation has been investigated for various proof assistants and programming languages.
+In Coq, appart from the $ring$ prover~\cite{Coq2005}, there's also the $omega$ solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
+a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
+Various automations have also been done in Agda~\cite{Lindblad04}.
+\\
+
+Finally, when we leave the ground of nice mathematical structures (like groups, rings and fields), one can decide to work with various rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} starts.
 
@@
 
-The correctness of the tactic is in fact obtained by construction.
+The correctness of all of our tactics is in fact obtained by construction.
 Actually, what we really needed to produce was the proof of the equality $a \simeq b$, and not the normalised terms.
@@
 However, if this situation would appear one day, it would certainly come from a small kind of simplification that would have been forgotten somewhere in the hierarchy, and it would be easy to fix it by adding a sub-function that deals with this specific simplification at the appropriated level.
-And all the more specialised provers that reuse this prover will also be fixed at the same time thanks to the compositionality of our provers.
+All the more specialised provers that reuse this prover would also be fixed at the same time thanks to the compositionality of our provers.
 \\
 
-Another little advantage of our approach, compared to Coq's approach, is that we have the guarantee that we can't lose any totality because of the $reflect$ function.
+Another little advantage of our approach is that we have the guarantee that we can't lose any totality because of the $reflect$ function.
 The reason is that the produced reflected term is indexed over the original --concrete-- value.
@@
 Thus, unlike Coq, we have the formal guarantee that $(reflect\ a)$ is always a faithful representation of $a$ with our approach, which means that we can't lose any totality because of the $reflect$ function.
+
+	\subsection {Conclusion and related work}
+
+XXX

### 2016-01-19 paper-JFP-DTP/content/6-discussion.tex
Finished

@@
 In terms of performances, our implementation is fast enough for providing answers in a decent amont of time --usually a few seconds--.
-The automatic proof of the lemma $adc\_lemma\_2$ (presented in section~\ref{sect:motivatingExample} and proven automatically in section ~\ref{sect:results}) is constantly generated and printed in less than 8 seconds on a dual core i5 processor @ 2.4ghz and is therefore perfectly usable, even for this kind of moderate size goals that come from a real application.
+For example, the automatic proof of the lemma $adc\_lemma\_2$ (presented in section~\ref{sect:motivatingExample} and proven automatically in section ~\ref{sect:results}) is constantly generated and printed in less than 8 seconds on a dual core i5 processor @ 2.4ghz and is therefore perfectly usable, even for this kind of moderate to large size proofs that come from a real application.
 Knowing that the display of the proof takes most of the time, it would be interesting to evaluate precisely the time needed for its generation only, without outputting the result.
 However, we suspect that our implementation will be slower than Coq, because they've used a more efficient representation of normalised polynomials.
-It might be possible that we've traded some efficiency against some simplicity and compositionality.
+It might be possible that we've traded some efficiency against some simplicity and compositionality of the various provers.
 
@@
 
-	\subsection {Conclusion and related work}
+	\subsection {Conclusion and futur work}
 
-XXX
+This paper has shown that type-safe reflection techniques can enable to manipulate proofs terms easily.
+We've been able to build an entire hierarchy of tactics for algebraic structures precisely because the construction of the proof terms has been greatly facilitated by the use this technique.
+We end up with a hierarchy of reflexive tactics for proving equivalence in semigroups, monoids, commutative monoids, groups, commutative groups, pre semi-ring, semi-ring, rings and commutative rings, without having had to repeat our efforts.
+
+It is even possible to extend a prover in order to build a dedicated one for a specific structure that isn't exactly a well known algebraic structure.
+We are at the moment investigating the extention of the pre semi-ring prover in order to build a prover for regular expressions.
+Regular expressions have a union of language playing the role of $+$ (which is associative, commutative, and which has a neutral element $\emptyset$) and a product of language playing the role of $*$, which is associative, distributive over the union, and which admits the empty word $\epsilon$ for neutral element.
+However, it has also a Kleene $star$ construction and a few properties about it.
+We currently aim to extend the pre semi-ring prover in order to build a prover for regular expression.
+Such a prover would be useful for doing formally verified transformations of regular grammars and final states machines for free, which could go up to a formally verified (for free!) parser generator for regular languages.

### 2016-01-19 paper-JFP-DTP/content/2-basicIdeas.tex
Various fix

@@
 
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Reflected LHS with the first version of reflected numbers showed in \ref{reflectedNaturalNumbers0}}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Production of the left associative form}
-\label{expr_l}
-\figrule
-\end{figure}
+
 In the case of an addition $Plus\ ex\ ey$, the function $expr\_l$ does the job of normalisation recursively on $ex$ and on $ey$, and then it uses the sub-function $plusLExpr$ to normalise the addition of these two --already normalised-- terms.

### 2016-01-19 paper-JFP-DTP/content/3-generalProblem.tex
Various fix

@@
 As we did for the very specific problem in section 2, we want to program an automatic reflection mechanism, in order to let the machine build automatically the encodings for us.
-This time, we want to write a function able to compute the ring encoding for any type that behaves as a ring.
-Said differently, we've got a carrier type $(c:Type)$, an instance $(p:Ring\ c)$ that says that $c$ behaves as a ring, a context $\Gamma$ of already abstracted variables, and an element $(x:c)$ to encode, and we want to return an $ExprR\ p\ \Gamma\ x$.
+This time, we want to write a function able to compute encodings for any type that behaves as one of our algebraic structure.
+We will show here the details for Ring, which is one of the most sophisticated structures.
+So, we've got a carrier type $(c:Type)$, an instance $(p:Ring\ c)$ which says that $c$ behaves as a ring, a context $\Gamma$ of already abstracted variables, and an element $(x:c)$ to encode, and we want to return an $ExprR\ p\ \Gamma\ x$.
 We are doing this reflection for any type $c$, so we don't know which constants are inhabiting this type.
@@
 But if the user wants to use the reflection mechanism for a new datatype of this own, then he will only have to provide this function of encoding for the constants of his type.
-He won't have to rebuild the entire reflection function.
+He won't have to rebuild the entire reflection mechanism.
 A $ConstantsRingReflector$ for a given type $c$ and a given instance $p$ is a function that takes any context $\Gamma$, any element of $c$, and which tries to produce an $(ExprR\ p\ \Gamma\ x)$.

### 2016-01-19 paper-JFP-DTP/content/5-results.tex
Various fix

@@
 \end{lstlisting}
-\end{center}
-\label{goal}
-\figrule
-\end{figure}
 
-In a context we're we've got the following induction hypothesis \[ihn:(plus\ (plus\ c\ bit)\ bit1\ =\ plus\ vLsb\ (plus\ vCarry0\ (plus\ vCarry0\ 0)))\]
-The goal can't be solved directly by calling the commutative monoid prover, as it uses different variables in its left and right hand side.
+In a context we're we've got the following induction hypothesis
+
+\begin{lstlisting}
+ihn:(plus (plus c bit) bit1 = plus vLsb (plus vCarry0 (plus vCarry0 0)))
+\end{lstlisting}
+
+This goal can't be solved directly by calling the commutative monoid prover, as it uses different variables in its left and right hand side.
 However, the induction hypothesis $ihn$ --obtained with the recursive call-- creates a link between these variables.

### 2016-01-19 paper-JFP-DTP/content/6-discussion.tex
Various fix

@@
 
-At the connexion bewteen proof automation and proof by reflection, we can cite SSreflect~\cite{GonthierTuto}, which is an extension of Coq which aims to provide additional formalisation and tactics suitable for long mathematical proofs.
+At the connexion bewteen proof automation and proof by reflection, we can cite SSreflect~\cite{GonthierTuto}, which is an extension of Coq which aims to provide additional formalisations and tactics suitable for long mathematical proofs.
 \\
@@
 
-Finally, when we leave the ground of nice mathematical structures (like groups, rings and fields), one can decide to work with various rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+Finally, when we leave the ground of nice mathematical structures (like groups, rings and fields), one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
 This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} starts.
@@
 This paper has shown that type-safe reflection techniques can enable to manipulate proofs terms easily.
-We've been able to build an entire hierarchy of tactics for algebraic structures precisely because the construction of the proof terms has been greatly facilitated by the use this technique.
+We've been able to build an entire hierarchy of tactics for algebraic structures precisely because the construction of the proof terms has been greatly facilitated by the use of this technique.
 We end up with a hierarchy of reflexive tactics for proving equivalence in semigroups, monoids, commutative monoids, groups, commutative groups, pre semi-ring, semi-ring, rings and commutative rings, without having had to repeat our efforts.
@@
 We currently aim to extend the pre semi-ring prover in order to build a prover for regular expression.
-Such a prover would be useful for doing formally verified transformations of regular grammars and final states machines for free, which could go up to a formally verified (for free!) parser generator for regular languages.
+Such a prover would be useful for doing formally-verified-for-free transformations of regular grammars and final states machines, which could go up to a formally verified (for free!) parser generator for regular languages.

### 2016-01-19 paper-JFP-DTP/content/1-introduction.tex
Almost done. futur work to write
Final fix before second submission

@@
 \subsection{Motivating example: Verified Binary Arithmetic}
+\label{sect:motivatingExample}
 
@@
 In this paper, we describe a certified
-implementation of an automatic prover for equalities in a
+implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris}} of an automatic prover for equalities in a
 \emph{hierarchy} of algebraic
@@
         simplifying neutral elements is only implemented at the monoid level,
-        and each level above monoid reuses it. Re-usability can be quite tricky
+        and each level inheriting from it will reuse it. Re-usability can be quite tricky
         to obtain when we want to reuse a prover from a structure which is less

### 2016-01-19 paper-JFP-DTP/content/4-detailsNormalisation.tex
corrections with Edwin advices on latex
Final fix before second submission

@@
 The construction of the desired proof is always done bit by bit, because every rewriting of the reflected term is immediately justified by a little accompanying proof which shows the equivalence of the previous and new index.
-\begin{figure}[H]
-\figrule
-\begin{center}
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{General pattern of a function doing a part of the normalisation}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Computing with associativity in a Semi-Group, first pattern}
-\figrule
-\end{figure}
+
 The treatment is applied recursively on the sub-expressions $ex$ and $ey$, which produces the new expressions $e\_ihx$ and $e\_ihy$ together with the new concrete values they represent ($r\_ihx$ and $r\_ihy$) and the proofs ($p\_ihx$ and $p\_ihy$) justifying these transformation on the concrete values.
@@
 
-The first thing that the reduction function of the group prover will have to do is simply to transform every subtraction $(a-b)$ into $(a + (-b))$, and we are entitled to do so precisely because we have available the property $Minus\_simpl$ contained in the instance of the type class.
+\subsection {Reusing the monoid prover for building the group prover}
+
+The first thing that the reduction function of the group prover has to do is to transform every subtraction $(a-b)$ into $(a + (-b))$, and we are entitled to do so precisely because we have available the property $Minus\_simpl$ contained in the instance of the typeclass.
 After that, the reduction continues with the propagation of the $Neg$ operation inside the parenthesis and $-(a+b)$ is transformed into $(-b) + (-a)$.
@@
 \caption{Real variables and encodings}
+\label{Variable}
 \figrule
@@
 \caption{new Var constructor}
+\label{VarG}
 \figrule
@@
 This is possible because we've fully developed the expression, which had for effect to ensure that a product can't contain a sum.
-We will add a coonstructor to the type $Variable$ in order to encode these monomials.
+We add a coonstructor to the type $Variable$ in order to encode these monomials.

### 2016-01-19 paper-JFP-DTP/content/2-basicIdeas.tex
Final fix before second submission
Minor edits to the paper

@@
 We will first present the basic ideas of our technique on a simplified problem, in which we aim to deal with universally quantified natural numbers and additions.
-For example, we would like to be able to automatically generate proofs of goals like $\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$ [Example 1]. \\
+For example, we would like to be able to automatically generate proofs of goals like:
+
+\noindent
+\textbf{Example:}
+$\forall x1\ x2\ x3:Nat,\ (x1 + l2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$  \\
 For this smaller problem, we have decided to only work with the associativity of $+$ : $plusAssociative : \forall x1\ x2\ x3,\ (x1 + x2) + x3 = x1 + (x2 + x3)$ and with the fact that Z is a right neutral element for the $+$ operation : $plusZeroRightNeutral : \forall x, x + Z = x$.
@@
 When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context.
-On the example 1 just above, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
+In the example above, after abstraction of the variables, the goal becomes simply $(x1 + x2) + (x3 + x4) = (x1 + (x2 + x3)) + x4$, which is something of the general form $x=y$.
 The general idea --that will also apply for the more general problem detailed in the next sections-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them using Leibniz syntactical equality.
@@
 For this reason, we will work by reflection.
-In this little example, it means that we will define a datatype that will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of ``variable numbers", $Z$, and additions of these things.
+In this example, it means that we will define a datatype that will be used as an encoding of natural numbers, or more precisely, an encoding of natural numbers composed of ``variable numbers", $Z$, and additions of these things.
 This datatype will let us inspect the internal structure of a number by pattern matching.
 Previously, we were only able to pattern match a natural number against the constructors $Z$ and $S$, which wasn't what we needed.
-With the first approximation of the data type $Expr$ presented in the figure~\ref{reflectedNaturalNumbers0}, we will be allowed to pattern match an encoding of number against the constructors $Plus$, $Var$ and $Zero$.
+With the first approximation of the data type $Expr$ presented in Figure~\ref{reflectedNaturalNumbers0}, we will be allowed to pattern match an encoding of number against the constructors $Plus$, $Var$ and $Zero$.
 
@@
 
-Variables are represent using a De Brujin-like index : (Var FZ) denotes a variable, (Var (FS FZ)) another one, and so on.
+Variables are represent using a De Brujin-like index: $Var\:FZ$ denotes a variable, $Var\:(FS\:FZ)$ another one, and so on.
 
 The type Expr is indexed over a vector of numbers $\Gamma$, which is the context of all universally quantified variables.
-In the example 1, we will encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are present.
+In the example, we will encode $(x1 + x2) + (x3 + x4)$ and $(x1 + (x2 + x3)) + x4$ in a context where four elements are present.
 The first element of this context denotes the variable $x1$, the second denotes $x2$, and so on.
@@
 \begin{lstlisting}
-e1 : (x, y, z : Nat) $\rightarrow$ 
-           Expr [x, y, z]
+e1 : (x, y, z : Nat) $\rightarrow$ Expr [x, y, z]
 e1 x y z = Plus (Plus (Var FZ) (Var (FS FZ))) 
@@
 This proof can be quite tricky to build because it relies on the complete behaviour of the reduction and on the way the interpretation is computed.
-For this little example, the reduction procedure won't be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more properties to deal with and it will certainly become more problematic to unfold the definition of a gigantic reduction procedure.
+For this example, the reduction procedure won't be too heavy, but in the next sections, with more sophisticated algebraic structures, we will have more properties to deal with and it will certainly become more problematic to unfold the definition of a gigantic reduction procedure.
 We want to avoid pain as much as possible, and for this reason, we want to avoid the writing of big proofs, and even to avoid the writing of proofs as much as possible.
@@
 \end{lstlisting}
-\end{center}
-\caption{testEq}
-\label{testEq}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Two test expressions}
-\label{e1_e2}
-\figrule
-\end{figure}
 

### 2016-01-20 paper-JFP-DTP/content/2-basicIdeas.tex
Inlining some of the code which was previously in figures

@@
 \end{lstlisting}
-\end{center}
-\caption{Syntactical equality between reflected terms}
-\label{syntacticalEq}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Building the desired proof with the two proofs of equality}
-\label{buildProof}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{buildProof metavariable}
-\label{MbuildProof}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Proof of the metavariable MplusLExpr}
-\label{MplusExpr}
-\figrule
-\end{figure}
 
@@
 The function $l\_expr$ does this easy task.
-\begin{figure}[H]
-\figrule
-\begin{center}
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Going back from LExpr to Expr}
-\label{l_expr}
-\figrule
-\end{figure}
 
 We notice that in order to transform the expression into its left associative equivalent representation, we've effectively needed to know where the variables and the $Z$ constants are : the functions $expr\_l$ and $l\_expr$ are doing different treatments for these two possibilities.
-\\
 \\
@@
 \end{lstlisting}
-\end{center}
-\caption{Reduction function}
-\label{reduce}
-\figrule
-\end{figure}
 
@@
 
+\begin{figure}[H]
+\figrule
+\begin{center}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{center}
+\caption{Two test expressions}
+\label{e1_e2}
+\figrule
+\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Generating automatically the desired proof}
-\label{e1_e2_testEq}
-\figrule
-\end{figure}
 
@@
 \end{center}
-\caption{Obtained proof}
+\caption{Automatically generated proof}
 \label{obtainedProof}
@@
 
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting} [mathescape]
@@
 \end{lstlisting}
-\end{center}
-\caption{Type of the function for reflecting natural numbers}
-\label{reflectNatType}
-\figrule
-\end{figure}
 
@@
 \\
-The function $weaken$ is easy to write because we're adding the extension $\Gamma'$ on the right of $\Gamma$.
+The function $weaken$ (Figure ~\ref{weaken}) is easy to write because we're adding the extension $\Gamma'$ on the right of $\Gamma$.
 For example, extending the context $[v,\ w,\ x]$ with $[y,\ z]$ produces the context $[v,\ w,\ x,\ y,\ z]$.
@@
 
- \begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting} [mathescape]
@@
 \end{lstlisting}
-\end{center}
-\caption{Conversion of a Fin}
-\label{convertFin}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Returns the index and the proof that this index is the right one}
-\label{isElement}
-\figrule
-\end{figure}
 

### 2016-01-20 paper-JFP-DTP/content/6-discussion.tex
Inlining some of the code which was previously in figures

@@
 We could now build the kernel of the tactic, which is the following decision procedure.
-\begin{figure}[H]
-\figrule
-\begin{center}
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Decision procedure of the tactic with a "traditional" approach}
-\label{decideEqTrad}
-\figrule
-\end{figure}
+
 This function on its own would not  be enough. We would need to add a correctness theorem, and this is precisely this main theorem that would help us to build the tactic we want.	
@@
 Because $Ltac$ is untyped, it would not permit to define --as we did-- a dependently typed function $reflect$ that is guaranteed to return a faithful encoding of the given input value.
-$Mtac$~\cite{Ziliani13} is an extension to Coq that supports dependently-typed tactic programming, which uses monad to avoid the need to touch the trusted kernel typechecker of Coq.
+$Mtac$~\cite{Ziliani13} is an extension to Coq that supports dependently-typed tactic programming that uses monad to avoid the need to touch the trusted kernel typechecker of Coq.
 \\
@@
 
-Finally, when we leave the ground of nice mathematical structures (like groups, rings and fields), one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
 This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} starts.

### 2016-02-03 paper-JFP-DTP/content/1-introduction.tex
Minor edits to the paper
Added a reference to the JFP DTP paper

@@
 then the lemma of correctness of $f$ can be expressed as $forall\ a:A,\ semA\ a =
-semB (f\ a)$. Very often, the type $SEM$ has some algebraic
+semB (f\ a)$.
+This situation happens very often in model transformations where the behaviour of an object has been preserved by some transformations, as for example in the CompCert certified compiler where the original C program and the generated assembly code are proven to have the same operational semantic \cite{Leroy09}.
+Very often, the type $SEM$ has some algebraic
 algebraic structure, and it becomes thus possible to
@@
 operation $+$ on natural numbers.  For example, the proof for the second
-pattern looks like this (with some detail elided):
+pattern is shown in Figure \ref{adc_lemma_2} (with some detail elided):
 

### 2016-02-15 ProofsAndTests/paper-TAP/paper-TAP.tex
start

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Automatic predicate testing in formal certification}
\subtitle{You've only proven what you've said, not what you meant!}

% a short form should be given in case it is too long for the running head
%\titlerunning{Automatic predicate testing in formal certification}

\author{Franck Slama\\
         \email{fs39@st-andrews.ac.uk}}

\institute{University of St Andrews, United Kingdom}

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
The use of formal methods helps to increase the confidence towards critical software. However, a formal proof is only a guarantee relative to a formal specification, and not about the real requirements. These real requirements are expressed in a natural language like english, and there is always a jump when going from an informal certification to a formal certification, expressed in a specific logic. Thus, proving the correctness about a piece of software always makes the implicit assumption that there is adequacy between the formalised specification and the real requirements. Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the specification says exactly and completely what it should say. Why should we trust more these predicates than the code that we've first refused to trust, leading to these proofs? We show in this paper that the proving activity hasn't replaced the testing activity. Instead, it has changed the object being tested. We show recent ideas about how to deal with this new kind of tests, including how to automate them as far as possible.
\keywords{Formal certification, predicate testing, specification, tests, requirements}
\end{abstract}

\input{./content/1-introduction.tex}

\end{document}

### 2016-02-15 ProofsAndTests/paper-TAP/content/1-introduction.tex
paper, abstract

\section{Introduction}

[This paper is just a placeholder for the final pdf]

### 2016-02-15 ProofsAndTests/paper-TAP/paper-TAP.tex
paper, abstract

@@
 \begin{abstract}
-The use of formal methods helps to increase the confidence towards critical software.
-However, a formal proof is only a guarantee relative to a formal specification, and not about the real requirements.
-These real requirements are expressed in a natural language like english, and there is always a jump when going from an informal certification to a formal certification, expressed in a specific logic.
-Thus, proving the correctness about a piece of software always makes the implicit assumption that there is adequacy between the formalised specification and the real requirements.
-Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the specification says exactly and completely what it should say.
-Why should we trust more these predicates than the code that we've first refused to trust, leading to these proofs?
+The use of formal methods helps to increase the confidence in critical software.
+However, a formal proof is only a guarantee relative to a formal specification, and not necessary about the real requirements.
+These real requirements are expressed in a natural language like english, and there is always a jump when going from an informal specification to a formal specification, expressed in a logical theory.
+Thus, proving the correctness about a piece of software always makes the implicit assumption that there is adequacy between the formalised specification, written with predicates, and the real requirements.
+Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the formal specification says exactly and completely what it should say.
+Why should we trust more these predicates than the code that we've first refused to trust blindly, leading to these proofs?
 We show in this paper that the proving activity hasn't replaced the testing activity.
-Instead, it has changed the object being tested.
-We show recent ideas about how to deal with this new kind of tests, including how to automate them as far as possible.
-\keywords{Formal certification, predicate testing, specification, tests, requirements}
+Instead, it has changed the object which requires to be tested.
+We show recent ideas about how to deal with this new kind of tests, which includes how to automate them as far as possible.
+\keywords{Formal certification, specification, predicate testing, proofs, tests, requirements}
 \end{abstract}

### 2016-02-15 ProofsAndTests/paper-TAP/paper-TAP.tex
abstract finished

@@
 However, a formal proof is only a guarantee relative to a formal specification, and not necessary about the real requirements.
-These real requirements are expressed in a natural language like english, and there is always a jump when going from an informal specification to a formal specification, expressed in a logical theory.
-Thus, proving the correctness about a piece of software always makes the implicit assumption that there is adequacy between the formalised specification, written with predicates, and the real requirements.
+These real requirements are expressed in a natural language, like english, and there is always a jump when going from an informal specification to a formal specification expressed in a logical theory.
+Thus, proving the correctness of a piece of software always makes the implicit assumption that there is adequacy between the formalised specification --written with logical statements and predicates-- and the real requirements.
 Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the formal specification says exactly and completely what it should say.
 Why should we trust more these predicates than the code that we've first refused to trust blindly, leading to these proofs?
-We show in this paper that the proving activity hasn't replaced the testing activity.
-Instead, it has changed the object which requires to be tested.
-We show recent ideas about how to deal with this new kind of tests, which includes how to automate them as far as possible.
-\keywords{Formal certification, specification, predicate testing, proofs, tests, requirements}
+We show in this paper that the proving activity hasn't replaced the testing activity but has only changed the object which requires to be tested.
+Instead of testing code, we now need to test predicates.
+We will show how to identify what needs to be tested in this new context.
+We also present recent ideas about how to conduct these tests on a few examples, and how to automate them as far as possible.
+
+\keywords{Formal certification, specification, predicate testing, formal proof, tests, requirements}
 \end{abstract}

### 2016-02-16 ProofsAndTests/paper-TAP/content/1-introduction.tex
Abstract submitted
intro written

@@
 
-[This paper is just a placeholder for the final pdf]
+There are multiple ways for increasing confidence in the software we build.
+One of them is to formally prove its correctness, using a proof assistant.
+These proofs assistants enable to write code, logical statements and proof in the same language, and offer the guarantee that every proof will be automatically checked.
+The appearance of these languages, and the way to write programs in them, varies from one to another.
+Many of them are functional programming languages, like Coq, Idris and Agda, and others, like the B-Toolkit belong to the imperative paradigm.
+These different paradigm are internally supported by different logic.
+Systems like Coq, Idris and Agda are based on various higher order logic (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic.
+These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software.
+However, these guarantee tend to be too often consider as perfect, when they are in fact far from it.
+Knuth was --certainly ironically-- saying "Beware of bugs in the above code; I have only proved it correct, not tried it".
+The reality is precisely that a proof isn't enough.
+When we've proved the correctness of a software, we've only gained the guarantee expressed by the proven lemma, and nothing more.
+Thus, if we want to trust the proven software, we're now forced to believe that there is adequacy between the formal specification, written in some formal logic, and the informal requirements.
+A switch has occurred.
+We used to have to trust code, but we now have to trust logical statements and predicates.
+But when the specification is too often as complicated as the code, why should we blindly believe in it, when we've first refused to blindly trust the code?
+Too few people realize that the proving activity has created this silent assumption of adequacy between the formalised specification and the original real requirements, usually written in english.
+Of course, the proving activity has undeniably increased the confidence in the produced software, and the aim of this paper isn't to say otherwise.
+Many successful applications of these formal methods [CompCert ref, etc] have showed the value of developing a machined-checked proof of correctness.
+Instead, the primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approach, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
+More precisely, we :
+\begin{itemize}
+\item Show how a formal specification can be wrong or under specified, leading to a proof that does not convey the expected guarantee, and the usual and naive approaches to this problem (section 2)
+\item Show the distinction between structural properties, and the properties on the content, and how to identify what needs to be tested for the former category (section 3).
+\item Show how to conduct these tests in the proof assistant itself, by automatically generating terms, and how to automate these tests as far as possible (section 4)
+\end{itemize}
+
+\
+
+To add in the intro :
+
+- Many companies doing critical software don't believe in 
+(one sentence)
+
+2) How a formal specification can go wrong
+
+- How a predicate can be insufficient, leading to an under-specified behaviour
+(it is then unfortunately very easy to produce a "formally-verified" wrong implementation)
+[a quarter of a page : from 1.75 to 2.00]
+
+- Different known approach for enhancing the confidence in a predicate : defining many of them and proving their equivalence, doing small tests on some data that we know belong or not to the predicate, etc...
+[half a page : from 2.00 to 2.30]
+
+3) "Structural" VS "on content" properties
+
+- differences between "structural" properties, and properties about the "content".
+[half a page : from 2.30 to 3.00]
+
+4) Testing the predicates inside the proof assistant
+
+- it is often easier to generate things rather than to characterise them 
+[a quarter of a page : from 3.00 to 3.15]
+
+- testing the structural properties by automatic generation of the terms that should satisfy them, and running a decision procedure on them.
+[half a page : from 3.15 to 3.45]
+
+- Going a step further : "test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
+[half a page : from 3.45 to 4.15]
+
+5) Conclusions, related work, and future work (a quarter of a page) 
+[from 4.15 to 4.30]
+
+6) References 
+[a quarter of a page : from 4.30 to 4.45]

### 2016-02-16 ProofsAndTests/paper-TAP/paper-TAP.tex
Abstract submitted
intro written

@@
 % first the title is needed
-\title{Automatic predicate testing in formal certification}
+\title{Short paper : Automatic predicate testing in formal certification}
 \subtitle{You've only proven what you've said, not what you meant!}
@@
 \begin{abstract}
-The use of formal methods helps to increase the confidence in critical software.
+The use of formal methods and proof assistants helps to increase the confidence in critical software.
 However, a formal proof is only a guarantee relative to a formal specification, and not necessary about the real requirements.
@@
 We will show how to identify what needs to be tested in this new context.
-We also present recent ideas about how to conduct these tests on a few examples, and how to automate them as far as possible.
+We also present recent ideas about how to conduct these tests inside the proof assistant on a few examples, and how to automate them as far as possible.
 
-\keywords{Formal certification, specification, predicate testing, formal proof, tests, requirements}
+\keywords{Formal certification, specification, predicate testing, proof assistant, tests, requirements}
 \end{abstract}

### 2016-02-16 ProofsAndTests/paper-TAP/content/1-introduction.tex
Structure of the paper

@@
 However, these guarantee tend to be too often consider as perfect, when they are in fact far from it.
-Knuth was --certainly ironically-- saying "Beware of bugs in the above code; I have only proved it correct, not tried it".
+Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
 The reality is precisely that a proof isn't enough.
@@
 
+We will use Idris~\cite{brady2013idris}, a dependently typed programming language that can also be seen as a proof assistant, but all the ideas that we present here can be applied to any proof assistant with dependent types.
+
 \
@@
 (one sentence)
-
-2) How a formal specification can go wrong
-
-- How a predicate can be insufficient, leading to an under-specified behaviour
-(it is then unfortunately very easy to produce a "formally-verified" wrong implementation)
-[a quarter of a page : from 1.75 to 2.00]
-
-- Different known approach for enhancing the confidence in a predicate : defining many of them and proving their equivalence, doing small tests on some data that we know belong or not to the predicate, etc...
-[half a page : from 2.00 to 2.30]
-
-3) "Structural" VS "on content" properties
-
-- differences between "structural" properties, and properties about the "content".
-[half a page : from 2.30 to 3.00]
-
-4) Testing the predicates inside the proof assistant
-
-- it is often easier to generate things rather than to characterise them 
-[a quarter of a page : from 3.00 to 3.15]
-
-- testing the structural properties by automatic generation of the terms that should satisfy them, and running a decision procedure on them.
-[half a page : from 3.15 to 3.45]
-
-- Going a step further : "test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
-[half a page : from 3.45 to 4.15]
-
-5) Conclusions, related work, and future work (a quarter of a page) 
-[from 4.15 to 4.30]
-
-6) References 
-[a quarter of a page : from 4.30 to 4.45]

### 2016-02-16 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
Structure of the paper

\section{How a formal specification can go wrong}

- How a predicate can be insufficient, leading to an under-specified behaviour
(it is then unfortunately very easy to produce a ``formally-verified" wrong implementation)
[a quarter of a page : from 1.75 to 2.00]

- Different known approach for enhancing the confidence in a predicate : defining many of them and proving their equivalence, doing small tests on some data that we know belong or not to the predicate, etc... 
[half a page : from 2.00 to 2.30]

### 2016-02-16 ProofsAndTests/paper-TAP/content/4-testingInside.tex
Structure of the paper

\section{Testing the predicates inside the proof assistant}

- it is often easier to generate things rather than to characterise them 
[a quarter of a page : from 3.00 to 3.15]

- testing the structural properties by automatic generation of the terms that should satisfy them, and running a decision procedure on them.
[half a page : from 3.15 to 3.45]

- Going a step further : ``test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
[half a page : from 3.45 to 4.15]

### 2016-02-16 ProofsAndTests/paper-TAP/content/5-conclusion.tex
Structure of the paper

\section{Conclusions, related work, and future work (a quarter of a page)}
[from 4.15 to 4.30]

### 2016-02-18 ProofsAndTests/paper-TAP/content/1-introduction.tex
paper continued

@@
 The reality is precisely that a proof isn't enough.
-When we've proved the correctness of a software, we've only gained the guarantee expressed by the proven lemma, and nothing more.
-Thus, if we want to trust the proven software, we're now forced to believe that there is adequacy between the formal specification, written in some formal logic, and the informal requirements.
+When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.
+
+Say we want to implement a formally verified sorting function that works for list of elements of type $T$, where $T$ is ordered by a relation $\leq$.
+We can decide to put more or less information inside the type of the sorting function $sort$, in order to make it more or less explicit that this function is precisely a sorting function.
+We can decide to only use a ``weak" type, like $sort\ :\ \{T:Type\}\ \rightarrow\ List\ T\ \rightarrow\ List\ T$, and to use an external lemma to ensure the correctness of the function.
+Which property does this function has to respect ?
+First, the output has to be sorted.
+So first, we need to define this notion of being sorted.
+For this, we can define an inductive predicate.
+
+\begin{lstlisting}
+data isSorted : {T:Type} -> (Tord : Order T) 
+                 -> (l:List T) -> Type where
+    NilIsSorted : (Tord : Order T) -> isSorted Tord []
+    SingletonIsSorted : (Tord : Order T) -> (x:T) -> isSorted Tord [x]
+    ConsSorted : {Tord : Order T} -> (h1:T) -> (h2:T) -> (l:List T) 
+                 -> (isSorted Tord (h2::t)) -> (h1 $\leq$ h2) 
+                 -> (isSorted Tord (h1::(h2::t)))
+\end{lstlisting}
+
+The first and second constructor of this predicate says that $[\ ]$ and $[x]$ are sorted according to any order order, and for any $x$.
+The third one says that a list of two or more elements is sorted if the head $h1$ is lower or equal than the next element $h2$, and if the list deprived from its head is also sorted.
+
+In order to express that the result produced by the function $sort$ is sorted, we can write and prove the following lemma:
+$sort\_correct : \forall\ (T:Type)\ (l:List\ T),\ isSorted\ (sort\ l)$
+
+The problem with this specification is that it doesn't says anything about the content of the output.
+The function $sort$ could just return the empty list $[\ ]$ all the time, it would still be possible to prove this correctness lemma.
+Indeed, this formal specification only says that the result should be sorted according to $isSorted$, and since $[\ ]$ is sorted (by $NilIsSorted$), this wrong implementation would be seen as ``correct".
+Here, the problem is that the function is underspecified, and it is therefore possible to write a senseless implementation, which is unfortunately seen as correct.
+Only a careful reader could realize that the lemma $sort\_correct$ forgets to mention that the input and output list should be in bijection, meaning that everything which was originally in the input list should still be in the output, and that nothing else should have been added.
+
+Things more nasty can happen.
+Imagine that the third constructor of $isSorted$, called $consSorted$ would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.
+Then, the function $sort$ could do whatever it wants, its result would be seen as ``sorted".
+For example, if the bijection condition has been added to the formal specification --and we can't therefore return $[\ ]$--, it would for example be possible to return the exact input, unchanged.
+Such a function would also unintentionally be seen as ``correct".
+One could object that the person doing the proof of correctness should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than the next element in the list.
+The reality is that currently, many effort are going in the direction of proof automation, where the machine builds the proof automatically.
+Many proof assistants are enhanced regularly with new automations that solves automatically some kind of goals.
+For example, Coq has already a Ring prover~\cite{coq2005}, a solver for Presburger arithmetic, and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures [ref].
+There are even extensions to languages, such as Ltac~\cite{DelahayeLTac} and Mtac~\cite{Ziliani13} that aim to help the automation of tactics.
+Thus, the general tendency is precisely to let the machine find the proof on its own, and to only interact with the system when needed.
+And in the imperative world, the formal B-method also follows the same spirit, with many routine proofs being proven automatically.
+The problem is that the machine is never going to find a proof ``too easy", and will never report that something seems weird with the specification given by the user.
+We can't reasonably demand that the user of the proof assistant should read all of the automatically generated proof, looking for proofs that seem ``too easy" in order to find weaknesses in the specifications.
+
+Thus, if we want to trust the proven software, we're now forced to believe that there is adequacy between the formal specification --written in some formal logic-- and the informal requirements.
 A switch has occurred.
@@
 Of course, the proving activity has undeniably increased the confidence in the produced software, and the aim of this paper isn't to say otherwise.
-Many successful applications of these formal methods [CompCert ref, etc] have showed the value of developing a machined-checked proof of correctness.
-Instead, the primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approach, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
+Many successful applications of these formal methods, such as a formally verified C compiler \cite{Leroy09} have shown the value of developing a machined-checked proof of correctness.
+Instead, the primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approaches, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
 More precisely, we :
 \begin{itemize}
-\item Show how a formal specification can be wrong or under specified, leading to a proof that does not convey the expected guarantee, and the usual and naive approaches to this problem (section 2)
+\item Show the usual and naive approaches to the problem of a badly specified function which leads to a proof that does not convey the expected guarantee (section 2)
 \item Show the distinction between structural properties, and the properties on the content, and how to identify what needs to be tested for the former category (section 3).
-\item Show how to conduct these tests in the proof assistant itself, by automatically generating terms, and how to automate these tests as far as possible (section 4)
+\item Show a way to conduct these tests in the proof assistant itself, by automatically generating terms, and we present a technique to automate these tests as far as possible (section 4)
 \end{itemize}
 
-We will use Idris~\cite{brady2013idris}, a dependently typed programming language that can also be seen as a proof assistant, but all the ideas that we present here can be applied to any proof assistant with dependent types.
-
-\
-
-To add in the intro :
-
-- Many companies doing critical software don't believe in 
-(one sentence)
+We use Idris~\cite{brady2013idris}, a dependently typed programming language that can also be seen as a proof assistant, but all the ideas that we present here can be applied to any proof assistant based on type theory.

### 2016-02-18 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
paper continued

@@
-\section{How a formal specification can go wrong}
-
-- How a predicate can be insufficient, leading to an under-specified behaviour
-(it is then unfortunately very easy to produce a ``formally-verified" wrong implementation)
-[a quarter of a page : from 1.75 to 2.00]
+\section{Naive and usual approaches to the adequacy problem}
 

### 2016-02-18 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
paper...

@@
 
-- Different known approach for enhancing the confidence in a predicate : defining many of them and proving their equivalence, doing small tests on some data that we know belong or not to the predicate, etc...
-[half a page : from 2.00 to 2.30]
+In order to gain confidence in the formal specifications we write, an easy an naive approach is to test the predicate itself.
+That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
+
+With the predicate $isSorted$ defined in section 1, we can try to prove that it holds on the list [3, 5, 7] that we know sorted according to our intentions.
+
+\begin{lstlisting}
+isSorted_test1 : isSorted natIsOrdered [3, 5, 7]
+isSorted_test1 = 
+  let p1 : (3 <= 5) = ?MA in
+  let p2 : (5 <= 7) = ?MB in
+    ConsSorted 3 5 [7] (ConsSorted 5 7 [] (SingletonIsSorted _ 7) p2) p1
+
+\end{lstlisting}
+
+- Other usual approach to mention : defining many of them and proving their equivalence,
+[quarter of a page]

### 2016-02-19 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
save

@@
 isSorted_test1 = 
-  let p1 : (3 <= 5) = ?MA in
-  let p2 : (5 <= 7) = ?MB in
+  let p1 : (3 <= 5) = tryDec (lowerEqDec nat_partialOrder 3 5) in
+  let p2 : (5 <= 7) = tryDec (lowerEqDec nat_partialOrder 5 7) in
     ConsSorted 3 5 [7] (ConsSorted 5 7 [] (SingletonIsSorted _ 7) p2) p1
+\end{lstlisting}
 
+It is important to realize that this kind of tests is very particular in the sense that is is a \emph{test by proof} : we show that the predicate holds by doing the proof on a specific value, here $[3,\ 5,\ 7]$.
+
+We could go a step forward, and remove the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decided if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
+
+\begin{lstlisting}
+decideIsSorted : (Tord : Order T) -> (l:List T) -> Dec(isSorted Tord v)
+decideIsSorted Tord [] = Yes (NilIsSorted Tord)
+decideIsSorted Tord [x] = Yes (SingletonIsSorted Tord x)
+decideIsSorted Tord (h1::(h2::t)) with (lowerEqDec Tord h1 h2)
+  decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
+          with (decideIsSorted Tord (h2::t))
+    decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
+          | (Yes h2_tail_sorted) = Yes (ConsSorted h1 h2 t h2_tail_sorted h1_lower_h2)
+    decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
+          | (No h2_tail_not_sorted) = No [...]
+  decideIsSorted Tord (h1::(h2::t)) | No h1_not_lower_h2 = No [...]
 \end{lstlisting}
+
+When the list has two or more elements, we start by checking if the first two elements are ordered.
+If so, we continue to check recursively with the list deprived from its head $h1$, and otherwise we already know that the list isn't sorted.
 

### 2016-02-19 ProofsAndTests/paper-TAP/content/1-introduction.tex
paper continued

@@
 Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
-The reality is precisely that a proof isn't enough.
+The reality is precisely that a proof isn't enough\footnote{That's without taking in account that when the proof is only done on the paper (and not machine-checked), the proof can be wrong.
+The paper never complains about a false proof}.
 When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.
@@
 We can decide to only use a ``weak" type, like $sort\ :\ \{T:Type\}\ \rightarrow\ List\ T\ \rightarrow\ List\ T$, and to use an external lemma to ensure the correctness of the function.
-Which property does this function has to respect ?
+Which property does this function has to respect?
 First, the output has to be sorted.
@@
 Then, the function $sort$ could do whatever it wants, its result would be seen as ``sorted".
-For example, if the bijection condition has been added to the formal specification --and we can't therefore return $[\ ]$--, it would for example be possible to return the exact input, unchanged.
+Said differently, any list is seen as ``sorted" according to this predicate, just because of this single typo.
+For example, if the bijection condition has been added to the formal specification --and therefore the function can no longer return $[\ ]$ all the time--, it would for example be possible to return the exact input, unchanged.
 Such a function would also unintentionally be seen as ``correct".
@@
 \begin{itemize}
-\item Show the usual and naive approaches to the problem of a badly specified function which leads to a proof that does not convey the expected guarantee (section 2)
-\item Show the distinction between structural properties, and the properties on the content, and how to identify what needs to be tested for the former category (section 3).
-\item Show a way to conduct these tests in the proof assistant itself, by automatically generating terms, and we present a technique to automate these tests as far as possible (section 4)
+	\item Show the usual and naive approaches to the problem of a badly specified function which leads to a proof that does not convey the expected guarantee (section~\ref{sect:naiveApproaches})
+	\item Show the distinction between structural properties, and the properties on the content, and how to identify what needs to be tested for the former category (section~\ref{sect:differentKindOfProperties}).
+	\item Present a way to conduct these tests in the proof assistant itself, by automatically generating terms, and we present a technique to automate these tests as far as possible (section~\ref{sect:testingInside})
 \end{itemize}

### 2016-02-19 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
paper continued

@@
 
-In order to gain confidence in the formal specifications we write, an easy an naive approach is to test the predicate itself.
+\label{sect:naiveApproaches}
+
+When confronted to this problem of adequacy between the intuitive notion and the formalised one, a first possibility that usually requires quite a lot of effort is to formalise the notion multiple times, with different predicates, and to prove that they are equivalent.
+With our example, that means that we need to find another formalisation $isSorted'$ of being sorted, and to prove the following lemma.
+$pred\_equiv : \forall\ (T:Type)\ (l:List\ T), (isSorted\ l) \leftrightarrow (isSorted'\ l)$.
+This approach aims at increasing the confidence in our formal definitions by assuming that if we've managed to define multiple times the same notion, then we've surely succeeded to define the notion we wanted.
+The biggest problem with this approach is to be able to find some alternative formalisations that are sufficiently different from the original one.
+Obviously, if the new normalisations are too similar to the original one (and in the worst case the new ones are just syntactical variants of the first one), then we won't gain any guarantee.
+The ideal is to be able to capture the same notion by using very different point of view, but this task can be incredibly difficult, and the more different they are, the harder the equivalence proofs will be.
+
+\
+
+In order to gain confidence in the formal specifications we write, another traditional approach is to test the predicate on some values.
 That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
@@
 isSorted_test1 = 
-  let p1 : (3 <= 5) = tryDec (lowerEqDec nat_partialOrder 3 5) in
-  let p2 : (5 <= 7) = tryDec (lowerEqDec nat_partialOrder 5 7) in
+  let p1 : (3 <= 5) = tryDec (lowerEqDec natIsOrdered 3 5) in
+  let p2 : (5 <= 7) = tryDec (lowerEqDec natIsOrdered 5 7) in
     ConsSorted 3 5 [7] (ConsSorted 5 7 [] (SingletonIsSorted _ 7) p2) p1
@@
 
-It is important to realize that this kind of tests is very particular in the sense that is is a \emph{test by proof} : we show that the predicate holds by doing the proof on a specific value, here $[3,\ 5,\ 7]$.
+It is important to realize that this kind of tests is very particular in the sense that it is a \emph{test done by proof} : we show that the predicate holds on some specific value, here $[3,\ 5,\ 7]$, by doing the proof.
 
-We could go a step forward, and remove the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decided if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
+We could go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decided if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
 
@@
 
-- Other usual approach to mention : defining many of them and proving their equivalence,
-[quarter of a page]
+Now, in order to do \emph{tests by proof} on the predicate $isSorted$, we can simply run the above decision procedure.
+
+\begin{lstlisting}
+isSorted_test1' : Dec (isSorted natIsOrdered [3,5,7])
+isSorted_test1' = decideIsSorted natIsOrdered [3,5,7] 
+\end{lstlisting}
+
+And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and the automatically generated proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
+
+With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.
+It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted and unsorted --according to our intention-- but we can let the machine produce the proof that the predicate holds or not on these specific values.
+This isn't bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.
+We would like to make sure that our intuitive notion and the formalised one coincide for every term.

### 2016-02-19 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
paper continued

\section{``Structural" VS ``on content" properties}

\label{sect:differentKindOfProperties}

- differences between ``structural" properties, and properties about the ``content".
[half a page : from 2.30 to 3.00]

### 2016-02-19 ProofsAndTests/paper-TAP/content/3-testingInside.tex
paper continued
save before change from vect to list

--- 2016-02-16 ProofsAndTests/paper-TAP/content/4-testingInside.tex
@@
-\section{Testing the predicates inside the proof assistant}
+\section{Testing the predicate by automatic generation of terms}
+
+\label{sect:testingInside}
+
+It is often easier to generate examples of something than it is to precisely define it. Thus, we will define a function that is supposed to generate the lists that we know sorted.
+
+in order to ensure that the predicate captures the notion that we want for \emph{any} term, we write a function that will
 
@@
 [a quarter of a page : from 3.00 to 3.15]
+
+- New : a very nice aspect is that we can now automatically generate \emph{all} the terms that satisfy the predicate.
+it is therefore possible to see some speciffic terms and to check weather they should have been captures or not.
 

### 2016-02-20 ProofsAndTests/paper-TAP/content/3-testingInside.tex
sanity save 2
tests done

@@
 
-It is often easier to generate examples of something than it is to precisely define it. Thus, we will define a function that is supposed to generate the lists that we know sorted.
+It is often easier to generate examples of something than it is to precisely define it.
+We could operate an interesting change of point of view by generating the set of terms that we precisely wanted to describe with the predicate.
+With our example of sorted lists, that means that we could try to define the same notion of being sorted, but this time with a definition example-based.
+Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different to the predicate that if the two notions agree, then we will know we've gained a great confidence on the predicate.
+
+We decide to use coinduction and the type LList (which stands for ``Lazy List", that's a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$.
+
+$generateSortedList : (T:Type)\ \rightarrow\ (recEnu:RecursivelyEnumerable\ T) \rightarrow\ (Tord : PartialOrder\ T) \rightarrow\ (n:Nat) \rightarrow LList (List T)$.
+
+To do so, the type $T$ needs to be $RecursivelyEnumerable$, which means that there should there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
+
+We want to check that this function and the predicate agree, which means that any list generated by $generateSortedList$ should be a yes instance of the predicate $isSorted$, and reciprocally.
+
+Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted list of size $n$ are automatically sorted or not by running the decision procedure.
+But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generative function and the predicate coincide on a finite unfolding of the resulting lazy list.
 
@@
 - it is often easier to generate things rather than to characterise them 
-[a quarter of a page : from 3.00 to 3.15]
+[a quarter of a page : from 3.00 to 3.15
 
 - New : a very nice aspect is that we can now automatically generate \emph{all} the terms that satisfy the predicate.
-it is therefore possible to see some speciffic terms and to check weather they should have been captures or not.
+it is therefore possible to see some specific terms and to check weather they should have been captures or not.
 

### 2016-02-20 ProofsAndTests/paper-TAP/content/1-introduction.tex
paper almost finished

@@
 \begin{itemize}
-	\item Show the usual and naive approaches to the problem of a badly specified function which leads to a proof that does not convey the expected guarantee (section~\ref{sect:naiveApproaches})
-	\item Show the distinction between structural properties, and the properties on the content, and how to identify what needs to be tested for the former category (section~\ref{sect:differentKindOfProperties}).
-	\item Present a way to conduct these tests in the proof assistant itself, by automatically generating terms, and we present a technique to automate these tests as far as possible (section~\ref{sect:testingInside})
+	\item Show some basic approaches to the problem of underspecification (section~\ref{sect:naiveApproaches})
+	\item Present a way to test in the proof assistant if the predicate effectively captures the notion that we wanted to define, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
+	\item Present how we can go a step forward by replacing these tests about the predicate by some proofs (section~\ref{sect:aStepForward})
 \end{itemize}

### 2016-02-20 ProofsAndTests/paper-TAP/content/3-testingInside.tex
paper almost finished

@@
 It is often easier to generate examples of something than it is to precisely define it.
-We could operate an interesting change of point of view by generating the set of terms that we precisely wanted to describe with the predicate.
-With our example of sorted lists, that means that we could try to define the same notion of being sorted, but this time with a definition example-based.
-Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different to the predicate that if the two notions agree, then we will know we've gained a great confidence on the predicate.
+We can operate an interesting change of point of view by generating the set of terms that we precisely wanted to describe with the predicate.
+With our example of sorted lists, that means that we need to define the same notion of being sorted, but this time with a definition example-based.
+Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different from the predicate that if the two notions agree, then we will have gained a great confidence on the predicate.
 
@@
 
-Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted list of size $n$ are automatically sorted or not by running the decision procedure.
+Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running the decision procedure.
 But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generative function and the predicate coincide on a finite unfolding of the resulting lazy list.
+The biggest this unfolding is, the better the guarantee is about $isSorted$.
 
-in order to ensure that the predicate captures the notion that we want for \emph{any} term, we write a function that will
+Let $T$ be a type that only contains three values $A$, $B$ and $C$. Since $T$ is finite, it is therefore obviously recursively enumerable. We define on it the strict order $A < B < C$.
+Let's automatically generate the first $m$ sorted lists of $T$, of size $n$, by unfolding $m$ times the result of $generateSortedList$.
 
-- it is often easier to generate things rather than to characterise them 
-[a quarter of a page : from 3.00 to 3.15
+\begin{lstlisting}
+testGenSorted : (m:Nat) -> (n:Nat) -> Maybe(Vect m (List T))
+testGenSorted m n = let x = generateSortedList T TisRecEnu TisOrdered n in
+		      	unfold_n_times x m
+\end{lstlisting}
 
-- New : a very nice aspect is that we can now automatically generate \emph{all} the terms that satisfy the predicate.
-it is therefore possible to see some specific terms and to check weather they should have been captures or not.
+We can obtain the first 8 sorted lists of size 5 by evaluating $testGenSorted\ 8\ 5$:
 
-- testing the structural properties by automatic generation of the terms that should satisfy them, and running a decision procedure on them.
-[half a page : from 3.15 to 3.45]
+\begin{lstlisting}
+Just [[A, A, A, A], [A, A, A, B], [A, A, A, C], [A, A, B, B],
+      [A, A, B, C], [A, A, C, C], [A, B, B, B], [A, B, B, C]] : Maybe (Vect 8 (List T))
+\end{lstlisting}
 
-- Going a step further : ``test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
-[half a page : from 3.45 to 4.15]
+Now, instead of simply generating the first $m$ sorted lists, we want to run the decision procedure on all of these $m$ tests in order to know if the predicate and the generating function agree on this portion.
+The result will be a vector $m$ booleans.
+
+\begin{lstlisting}
+testSorted : (m:Nat) -> (n:Nat) -> Vect m Bool
+testSorted m n = 
+  let x = generateSortedList T TisRecEnu TisOrdered n in
+    let y = LLmap (\l => let res = decideIsSorted TisOrdered l in
+			    case res of
+			      Yes _ => True
+			      No _ => False) x in
+	unfold_n_times_with_padding y m True
+\end{lstlisting}
+
+And we can inspect the result of running the first 8 tests by evaluating $testSorted\ 8\ 5$.
+
+\begin{lstlisting}
+[True, True, True, True, True, True, True, True] : Vect 8 Bool
+\end{lstlisting}
+
+When we want to test $isSorted$ on a big number of tests, we might not want to inspect by hand the result of each test.
+We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the $and$ on each element of the resulting vector.
+Now, we can test the predicate on the first 50 sorted lists of size by running $testSorted\_result\ 50\ 8$ and we get the overall result $True$ which means that the predicate agree with the generator.

### 2016-02-20 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
paper almost finished

@@
-\section{``Structural" VS ``on content" properties}
+\section{Replacing tests on the predicate by proofs}
 
-\label{sect:differentKindOfProperties}
+\label{sect:aStepForward}
 
-- differences between ``structural" properties, and properties about the ``content".
-[half a page : from 2.30 to 3.00]
+- Going a step further : ``test automation by proof", which automatises all of them at once, by doing (another) formal proof, which shows that there is equivalence between all the terms generated and the original predicate that we wanted to certify correct according to the original specification.
+[half a page : from 3.45 to 4.15]

### 2016-02-20 ProofsAndTests/paper-TAP/content/3-testingInside.tex
paper needs to be cut

@@
 
-It is often easier to generate examples of something than it is to precisely define it.
+It is often easier to generate examples of a notion than it is to precisely define it.
 We can operate an interesting change of point of view by generating the set of terms that we precisely wanted to describe with the predicate.
@@
 
-We decide to use coinduction and the type LList (which stands for ``Lazy List", that's a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$.
+We decide to use coinduction and the type Stream (that's a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$ with what we call a generator.
 
-$generateSortedList : (T:Type)\ \rightarrow\ (recEnu:RecursivelyEnumerable\ T) \rightarrow\ (Tord : PartialOrder\ T) \rightarrow\ (n:Nat) \rightarrow LList (List T)$.
+$generateSortedList : (T:Type)\ \rightarrow\ (recEnu:RecursivelyEnumerable\ T) \rightarrow\ (Tord : Order\ T) \rightarrow\ (n:Nat) \rightarrow Stream\ (List\ T)$.
 
@@
 Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running the decision procedure.
-But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generative function and the predicate coincide on a finite unfolding of the resulting lazy list.
-The biggest this unfolding is, the better the guarantee is about $isSorted$.
+But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} of the resulting stream.
+The biggest this observation is, the better the guarantee is about $isSorted$.
 
-Let $T$ be a type that only contains three values $A$, $B$ and $C$. Since $T$ is finite, it is therefore obviously recursively enumerable. We define on it the strict order $A < B < C$.
+We show how the observation is made on an example.
+Let $T$ be a type that only contains three values $A$, $B$ and $C$.
+Since $T$ is finite, it is therefore obviously recursively enumerable.
+We define on it the strict order $A < B < C$.
 Let's automatically generate the first $m$ sorted lists of $T$, of size $n$, by unfolding $m$ times the result of $generateSortedList$.
@@
 \end{lstlisting}
-
-We can obtain the first 8 sorted lists of size 5 by evaluating $testGenSorted\ 8\ 5$:
+We can ask for the first 8 sorted lists of size 5 by evaluating $testGenSorted\ 8\ 5$:
 
@@
 \end{lstlisting}
-
-Now, instead of simply generating the first $m$ sorted lists, we want to run the decision procedure on all of these $m$ tests in order to know if the predicate and the generating function agree on this portion.
+Now, instead of simply generating the first $m$ sorted lists, we want to run the decision procedure on all of these $m$ tests in order to know if the predicate and the generator agree on this portion.
 The result will be a vector $m$ booleans.
@@
 \end{lstlisting}
-
 And we can inspect the result of running the first 8 tests by evaluating $testSorted\ 8\ 5$.
@@
 \end{lstlisting}
-
-When we want to test $isSorted$ on a big number of tests, we might not want to inspect by hand the result of each test.
-We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the $and$ on each element of the resulting vector.
-Now, we can test the predicate on the first 50 sorted lists of size by running $testSorted\_result\ 50\ 8$ and we get the overall result $True$ which means that the predicate agree with the generator.
+When we want to test $isSorted$ on a big number of tests, we might not want to inspect manually the result of each test.
+We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the boolean $And$ on each element of the resulting vector.
+Now, we can for example test the predicate on the first 50 sorted lists of size 8 by running $testSorted\_result\ 50\ 8$ and if we do so we get the overall result $True$ which means that the predicate agree with the generator on all these 50 tests.

### 2016-02-20 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
paper needs to be cut

--- 2016-02-19 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
@@
-\section{``Structural" VS ``on content" properties}
+\section{Replacing tests on the predicate by proofs}
 
-\label{sect:differentKindOfProperties}
+\label{sect:aStepForward}
 
-- differences between ``structural" properties, and properties about the ``content".
-[half a page : from 2.30 to 3.00]
+In order to go a step forward, we can decide to replace the tests on the predicate by proofs.
+Instead of testing the predicate on a finite subset of the infinitely many sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
+
+\begin{lstlisting}
+generated_implies_pred_holds : {T:Type} -> (recEnu:RecursivelyEnumerable T) 
+        -> (Tord : Order T) -> (n:Nat) 
+        -> (All (generateSortedList T recEnu Tord n) (\l => isSorted Tord l))
+\end{lstlisting}
+
+The lemma has the advantage of not requiring the predicate to be decidable, when this was needed in the previous section where we tested the predicate on a finite observation.
+
+\begin{proof}
+By induction on $n$.
+When $n$ is zero, there is only one sorted list generated, which is the empty list, and we know that the empty list is sorted thanks to the constructor $NilIsSorted$.
+When n is some successor $(S\ pn)$, we know by using recursively the lemma on the smaller value $pn$ that all sorted lists of size $pn$ are sorted according to $isSorted$.
+Since the $Stream$ of all sorted lists of size $(S\ pn)$ has been made from the $Stream$ of all sorted lists of size $pn$ by adding to all of them --on the head position-- an element lower or equal to the current heads, we know that the property has been preserved at the higher rank.
+\end{proof}
+
+One could object that this lemma is itself built by using a predicate, $All$, and that we can't necessary trust blindly such a specification.
+The answer is that no guarantee is perfect, and all we can do is to add more guarantee, but there is always something to trust.
+Moreover, this new kind of specification and proofs --about the predicate itself-- uses more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
+If they are part of some standard library being used intensively, there is very low risk that they don't capture the desired semantic.
+
+In the other direction, we can prove that when the predicate $isSorted$ holds on some list, then this list is one of the list produced by the generator.
+The proof is done by induction on the list.
+The case of an empty list is trivial.
+For a singleton $[x]$, the proof uses the fact that the $T$ is recursively enumerable, and more precisely that the provided map is surjective, which means that it has hit any value, and thus in particular $x$.
+When there are two or more elements, we know that the smaller list deprived from its head is sorted according to $isSorted$ by construction.
+By using the lemma recursively on this smaller list together with the fact that the head is lower than the next element and the fact that the map is surjective, we get what we need.

### 2016-02-20 ProofsAndTests/paper-TAP/content/5-conclusion.tex
paper needs to be cut

@@
-\section{Conclusions, related work, and future work (a quarter of a page)}
-[from 4.15 to 4.30]
+\section{Conclusions and future work}
+
+We've presented some new ideas and also reintroduced some known techniques that aim to gain confidence in a predicate inside a proof assistant.
+Our new technique is based on the automatic generation of terms that should have the desired property.
+Thus, our goal was to make sure that a generator of terms and the predicate coincide.
+This adequacy between the predicate and the generator helps to gain confidence in the predicate, because generating examples of something and giving a definition of the same thing are tasks which are sufficiently different that it is very unlikely that they could coincide by luck.
+The technique presented on section \ref{sect:testingInside} was based on a finite observation of the automatically generated terms processed by a decision procedure.
+The limitation of this approach is that the predicate has to be decidable.
+This isn't a big restriction, as a huge portion of the predicates that are used for writing specifications are decidable.
+We've also shown on an example how these tests can be completely automated.
+In section \ref{sect:aStepForward} we've presented ideas for going a step forward and replacing these tests on the predicate by some new proofs.
+These new proofs have the advantage of not requiring the predicate to be decidable.
+However, they imply that we have to trust these new specifications, which are this time about the predicate itself.
+
+We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the correctness of a function relatively to a bad specification.
+
+A natural extension of the work presented here is to make it more generic, and to study how proof assistants could help --by providing primitive bricks-- to gain confidence in the formal definitions that we write.
+A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own notion.
+However, as soon as dependent types are concerned, and when types like lists become indexed by elements like natural numbers, it becomes usually impossible to reuse immediately definitions that were working for the non indexed version.

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
still a bit more of cutting to do

@@
 One of them is to formally prove its correctness, using a proof assistant.
-These proofs assistants enable to write code, logical statements and proof in the same language, and offer the guarantee that every proof will be automatically checked.
+These proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
 The appearance of these languages, and the way to write programs in them, varies from one to another.
@@
 Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
-The reality is precisely that a proof isn't enough\footnote{That's without taking in account that when the proof is only done on the paper (and not machine-checked), the proof can be wrong.
-The paper never complains about a false proof}.
+The reality is precisely that a proof isn't enough.
 When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.
@@
 Say we want to implement a formally verified sorting function that works for list of elements of type $T$, where $T$ is ordered by a relation $\leq$.
-We can decide to put more or less information inside the type of the sorting function $sort$, in order to make it more or less explicit that this function is precisely a sorting function.
 We can decide to only use a ``weak" type, like $sort\ :\ \{T:Type\}\ \rightarrow\ List\ T\ \rightarrow\ List\ T$, and to use an external lemma to ensure the correctness of the function.
@@
 
-The first and second constructor of this predicate says that $[\ ]$ and $[x]$ are sorted according to any order order, and for any $x$.
+The first and second constructor of this predicate says that $[\ ]$ and $[x]$ are sorted according to any order, and for any $x$.
 The third one says that a list of two or more elements is sorted if the head $h1$ is lower or equal than the next element $h2$, and if the list deprived from its head is also sorted.
@@
 Then, the function $sort$ could do whatever it wants, its result would be seen as ``sorted".
-Said differently, any list is seen as ``sorted" according to this predicate, just because of this single typo.
+Said differently, any list would be seen as ``sorted" according to this predicate, just because of this single typo.
 For example, if the bijection condition has been added to the formal specification --and therefore the function can no longer return $[\ ]$ all the time--, it would for example be possible to return the exact input, unchanged.
@@
 Thus, the general tendency is precisely to let the machine find the proof on its own, and to only interact with the system when needed.
-And in the imperative world, the formal B-method also follows the same spirit, with many routine proofs being proven automatically.
 The problem is that the machine is never going to find a proof ``too easy", and will never report that something seems weird with the specification given by the user.
@@
 But when the specification is too often as complicated as the code, why should we blindly believe in it, when we've first refused to blindly trust the code?
-Too few people realize that the proving activity has created this silent assumption of adequacy between the formalised specification and the original real requirements, usually written in english.
+The proving activity has created this silent assumption of adequacy between the formalised specification and the original real requirements, usually written in english.
 Of course, the proving activity has undeniably increased the confidence in the produced software, and the aim of this paper isn't to say otherwise.

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
still a bit more of cutting to do

@@
 Obviously, if the new normalisations are too similar to the original one (and in the worst case the new ones are just syntactical variants of the first one), then we won't gain any guarantee.
-The ideal is to be able to capture the same notion by using very different point of view, but this task can be incredibly difficult, and the more different they are, the harder the equivalence proofs will be.
+The ideal would be to capture the same notion by using very different point of view, and we will show in section~\ref{sect:testingInside} an original approach for doing so.
 
@@
 
-We could go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decided if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
+We could go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decides if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
 
@@
 \end{lstlisting}
-
-When the list has two or more elements, we start by checking if the first two elements are ordered.
-If so, we continue to check recursively with the list deprived from its head $h1$, and otherwise we already know that the list isn't sorted.
-
-Now, in order to do \emph{tests by proof} on the predicate $isSorted$, we can simply run the above decision procedure.
+Now, in order to do \emph{tests by proof} on the predicate $isSorted$, we can simply run the decision procedure.
 
@@
 \end{lstlisting}
-
 And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and the automatically generated proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
@@
 This isn't bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.
-We would like to make sure that our intuitive notion and the formalised one coincide for every term.

### 2016-02-21 ProofsAndTests/paper-TAP/content/3-testingInside.tex
still a bit more of cutting to do

@@
 
-We want to check that this function and the predicate agree, which means that any list generated by $generateSortedList$ should be a yes instance of the predicate $isSorted$, and reciprocally.
-
+We want to check that this function and the predicate coincide.
 Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running the decision procedure.
-But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} of the resulting stream.
+But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} (arbitrary big) of the resulting stream.
 The biggest this observation is, the better the guarantee is about $isSorted$.
@@
 We show how the observation is made on an example.
-Let $T$ be a type that only contains three values $A$, $B$ and $C$.
+Let $T$ be a type that only contains three constant values $A$, $B$ and $C$.
 Since $T$ is finite, it is therefore obviously recursively enumerable.

### 2016-02-21 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
still a bit more of cutting to do

@@
 In order to go a step forward, we can decide to replace the tests on the predicate by proofs.
-Instead of testing the predicate on a finite subset of the infinitely many sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
+Instead of testing the predicate on a finite subset of the generated sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
 

### 2016-02-21 ProofsAndTests/paper-TAP/content/5-conclusion.tex
still a bit more of cutting to do

@@
 Our new technique is based on the automatic generation of terms that should have the desired property.
-Thus, our goal was to make sure that a generator of terms and the predicate coincide.
+Thus, our goal was to make sure that the predicate and a generator of terms coincide.
 This adequacy between the predicate and the generator helps to gain confidence in the predicate, because generating examples of something and giving a definition of the same thing are tasks which are sufficiently different that it is very unlikely that they could coincide by luck.
@@
 A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own notion.
-However, as soon as dependent types are concerned, and when types like lists become indexed by elements like natural numbers, it becomes usually impossible to reuse immediately definitions that were working for the non indexed version.

### 2016-02-21 ProofsAndTests/paper-TAP/paper-TAP.tex
Structure of the paper
paper continued
still a bit more of cutting to do

@@
 \usepackage{url}
+
+\usepackage{listings}
+
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  keepspaces=true,
+  mathescape
+}
 
@@
 However, a formal proof is only a guarantee relative to a formal specification, and not necessary about the real requirements.
-These real requirements are expressed in a natural language, like english, and there is always a jump when going from an informal specification to a formal specification expressed in a logical theory.
-Thus, proving the correctness of a piece of software always makes the implicit assumption that there is adequacy between the formalised specification --written with logical statements and predicates-- and the real requirements.
+There is always a jump when going from an informal specification to a formal specification expressed in a logical theory.
+Thus, proving the correctness of a piece of software always makes the implicit assumption that there is adequacy between the formalised specification --written with logical statements and predicates-- and the real requirements --often written in english--.
 Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the formal specification says exactly and completely what it should say.
@@
 Instead of testing code, we now need to test predicates.
-We will show how to identify what needs to be tested in this new context.
-We also present recent ideas about how to conduct these tests inside the proof assistant on a few examples, and how to automate them as far as possible.
+We present recent ideas about how to conduct these tests inside the proof assistant on a few examples, and how to automate them as far as possible.
 
@@
 
+\input{./content/2-specCanGoWrong.tex}
+
+\input{./content/3-testingInside.tex}
+
+\input{./content/4-differentKindOfProperties.tex}
+
+\input{./content/5-conclusion.tex}
+
+\bibliographystyle{splncs03}
+\bibliography{bib-TAP}
+
 \end{document}

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
Save

@@
 First, the output has to be sorted.
-So first, we need to define this notion of being sorted.
-For this, we can define an inductive predicate.
+So first, we need to define this notion of being sorted, here as an inductive predicate :
 
@@
 There are even extensions to languages, such as Ltac~\cite{DelahayeLTac} and Mtac~\cite{Ziliani13} that aim to help the automation of tactics.
-Thus, the general tendency is precisely to let the machine find the proof on its own, and to only interact with the system when needed.
 The problem is that the machine is never going to find a proof ``too easy", and will never report that something seems weird with the specification given by the user.
-We can't reasonably demand that the user of the proof assistant should read all of the automatically generated proof, looking for proofs that seem ``too easy" in order to find weaknesses in the specifications.
 

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
Save

@@
 That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
-
-With the predicate $isSorted$ defined in section 1, we can try to prove that it holds on the list [3, 5, 7] that we know sorted according to our intentions.
+For example, with the predicate $isSorted$ defined in section 1, we can prove that it holds on the list [3, 5, 7] that we know sorted according to our intentions.
 
@@
 \end{lstlisting}
-
 It is important to realize that this kind of tests is very particular in the sense that it is a \emph{test done by proof} : we show that the predicate holds on some specific value, here $[3,\ 5,\ 7]$, by doing the proof.
 
-We could go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that decides if a list $l$ is sorted --according to isSorted-- and that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
+We can go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
 

### 2016-02-21 ProofsAndTests/paper-TAP/content/3-testingInside.tex
Save

@@
 Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different from the predicate that if the two notions agree, then we will have gained a great confidence on the predicate.
-
-We decide to use coinduction and the type Stream (that's a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$ with what we call a generator.
+We decide to use coinduction and the type Stream (a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$ with what we call a generator.
 
@@
 Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running the decision procedure.
-But since there can be an infinite amount of generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} (arbitrary big) of the resulting stream.
-The biggest this observation is, the better the guarantee is about $isSorted$.
+But since there can be infinitely many generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} of the resulting stream.
+The key point is that this observation can be arbitrary big, and the bigger it is, the better the guarantee is about $isSorted$.
 
@@
 testGenSorted : (m:Nat) -> (n:Nat) -> Maybe(Vect m (List T))
-testGenSorted m n = let x = generateSortedList T TisRecEnu TisOrdered n in
-		      	unfold_n_times x m
+testGenSorted m n = let x = generateSortedList T TisRecEnu TisOrdered n 
+                        in unfold_n_times x m
 \end{lstlisting}
@@
 We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the boolean $And$ on each element of the resulting vector.
-Now, we can for example test the predicate on the first 50 sorted lists of size 8 by running $testSorted\_result\ 50\ 8$ and if we do so we get the overall result $True$ which means that the predicate agree with the generator on all these 50 tests.
+Now, we can for example test the predicate on the first 50 sorted lists of size 8 by running $testSorted\_result\ 50\ 8$ and if we do so we get the overall result $True$ which means that the predicate agrees with the generator on all these 50 tests.

### 2016-02-21 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
Save

@@
 \end{lstlisting}
-
-The lemma has the advantage of not requiring the predicate to be decidable, when this was needed in the previous section where we tested the predicate on a finite observation.
-
 \begin{proof}
@@
 Since the $Stream$ of all sorted lists of size $(S\ pn)$ has been made from the $Stream$ of all sorted lists of size $pn$ by adding to all of them --on the head position-- an element lower or equal to the current heads, we know that the property has been preserved at the higher rank.
+\qed
 \end{proof}
 
-One could object that this lemma is itself built by using a predicate, $All$, and that we can't necessary trust blindly such a specification.
+This lemma has the advantage of not requiring the predicate to be decidable, when this was needed in the previous section where we tested the predicate on a finite observation.
+However, one could object that this lemma is itself built by using a predicate, $All$, and that we can't necessary trust blindly such a specification.
 The answer is that no guarantee is perfect, and all we can do is to add more guarantee, but there is always something to trust.
-Moreover, this new kind of specification and proofs --about the predicate itself-- uses more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
+Moreover, this new kind of specifications and proofs --about the predicate itself-- uses more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
 If they are part of some standard library being used intensively, there is very low risk that they don't capture the desired semantic.
@@
 In the other direction, we can prove that when the predicate $isSorted$ holds on some list, then this list is one of the list produced by the generator. 
-The proof is done by induction on the list.
+\begin{proof}
+By induction on the list.
 The case of an empty list is trivial.
@@
 By using the lemma recursively on this smaller list together with the fact that the head is lower than the next element and the fact that the map is surjective, we get what we need.
+\qed
+\end{proof}

### 2016-02-21 ProofsAndTests/paper-TAP/content/5-conclusion.tex
Save

@@
 The technique presented on section \ref{sect:testingInside} was based on a finite observation of the automatically generated terms processed by a decision procedure.
-The limitation of this approach is that the predicate has to be decidable.
-This isn't a big restriction, as a huge portion of the predicates that are used for writing specifications are decidable.
 We've also shown on an example how these tests can be completely automated.
@@
 
-A natural extension of the work presented here is to make it more generic, and to study how proof assistants could help --by providing primitive bricks-- to gain confidence in the formal definitions that we write.
+A natural extension of the work presented here is to make it more generic, and to study how proof assistants could help --by providing some primitive bricks-- to gain confidence in the formal definitions that we write.
 A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own notion.

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
Finished

@@
 
-There are multiple ways for increasing confidence in the software we build.
-One of them is to formally prove its correctness, using a proof assistant.
+One way to increase the confidence in the software we build is to formally prove its correctness using a proof assistant.
 These proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
@@
 \end{lstlisting}
-
-The first and second constructor of this predicate says that $[\ ]$ and $[x]$ are sorted according to any order, and for any $x$.
+The first and second constructor of this predicate say that $[\ ]$ and $[x]$ are sorted according to any order, and for any $x$.
 The third one says that a list of two or more elements is sorted if the head $h1$ is lower or equal than the next element $h2$, and if the list deprived from its head is also sorted.
-
 In order to express that the result produced by the function $sort$ is sorted, we can write and prove the following lemma:
-$sort\_correct : \forall\ (T:Type)\ (l:List\ T),\ isSorted\ (sort\ l)$
-
+$sort\_correct : \forall\ (T:Type)\ (l:List\ T),\ isSorted\ (sort\ l)$.
 The problem with this specification is that it doesn't says anything about the content of the output.
 The function $sort$ could just return the empty list $[\ ]$ all the time, it would still be possible to prove this correctness lemma.
-Indeed, this formal specification only says that the result should be sorted according to $isSorted$, and since $[\ ]$ is sorted (by $NilIsSorted$), this wrong implementation would be seen as ``correct".
-Here, the problem is that the function is underspecified, and it is therefore possible to write a senseless implementation, which is unfortunately seen as correct.
+Here, the problem is that the function is underspecified, and it is therefore possible to write a senseless implementation, which is unfortunately provably correct.
 Only a careful reader could realize that the lemma $sort\_correct$ forgets to mention that the input and output list should be in bijection, meaning that everything which was originally in the input list should still be in the output, and that nothing else should have been added.
@@
 Then, the function $sort$ could do whatever it wants, its result would be seen as ``sorted".
-Said differently, any list would be seen as ``sorted" according to this predicate, just because of this single typo.
-For example, if the bijection condition has been added to the formal specification --and therefore the function can no longer return $[\ ]$ all the time--, it would for example be possible to return the exact input, unchanged.
-Such a function would also unintentionally be seen as ``correct".
+Said differently, any list would be seen as ``sorted" according to this predicate, just because of this single typo, and the algorithm could for example return its input unchanged.
 One could object that the person doing the proof of correctness should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than the next element in the list.
@@
 Many proof assistants are enhanced regularly with new automations that solves automatically some kind of goals.
-For example, Coq has already a Ring prover~\cite{coq2005}, a solver for Presburger arithmetic, and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures [ref].
+For example, Coq has already a Ring prover~\cite{coq2005}, a solver for Presburger arithmetic, and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures~\cite{Slama2016}.
 There are even extensions to languages, such as Ltac~\cite{DelahayeLTac} and Mtac~\cite{Ziliani13} that aim to help the automation of tactics.
@@
 But when the specification is too often as complicated as the code, why should we blindly believe in it, when we've first refused to blindly trust the code?
-The proving activity has created this silent assumption of adequacy between the formalised specification and the original real requirements, usually written in english.
-Of course, the proving activity has undeniably increased the confidence in the produced software, and the aim of this paper isn't to say otherwise.
-Many successful applications of these formal methods, such as a formally verified C compiler \cite{Leroy09} have shown the value of developing a machined-checked proof of correctness.
-Instead, the primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approaches, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
+The primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approaches, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
 More precisely, we :
@@
 	\item Show some basic approaches to the problem of underspecification (section~\ref{sect:naiveApproaches})
-	\item Present a way to test in the proof assistant if the predicate effectively captures the notion that we wanted to define, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
+	\item Present a way to test in the proof assistant whether the predicate effectively captures the notion that we wanted to define, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
 	\item Present how we can go a step forward by replacing these tests about the predicate by some proofs (section~\ref{sect:aStepForward})

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
Finished

@@
 That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
-For example, with the predicate $isSorted$ defined in section 1, we can prove that it holds on the list [3, 5, 7] that we know sorted according to our intentions.
+For example, with the predicate $isSorted$ defined in section 1, we can prove that it holds on the list [3, 5, 7] that we know sorted according to our informal sense of what being sorted is.
 
@@
 \end{lstlisting}
-It is important to realize that this kind of tests is very particular in the sense that it is a \emph{test done by proof} : we show that the predicate holds on some specific value, here $[3,\ 5,\ 7]$, by doing the proof.
-
+This test is a \emph{test done by proof} : we show that the predicate holds on some specific value, here $[3,\ 5,\ 7]$, by doing the proof.
 We can go a step forward by removing the need of doing these specific proofs by hand, because in this case, the predicate isSorted can be decided : there exists an algorithm that produces a proof of $(isSorted\ l)$ if appropriate, or a proof of $(not\ (isSorted\ l))$ otherwise.
@@
 \end{lstlisting}
-Now, in order to do \emph{tests by proof} on the predicate $isSorted$, we can simply run the decision procedure.
+Now, in order to do \emph{tests by proof}, we can simply run the decision procedure.
 

### 2016-02-21 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
Finished

@@
 The answer is that no guarantee is perfect, and all we can do is to add more guarantee, but there is always something to trust.
-Moreover, this new kind of specifications and proofs --about the predicate itself-- uses more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
+Moreover, this new kind of specifications and proofs --about the predicate itself-- use more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
 If they are part of some standard library being used intensively, there is very low risk that they don't capture the desired semantic.
@@
 In the other direction, we can prove that when the predicate $isSorted$ holds on some list, then this list is one of the list produced by the generator.
-\begin{proof}
-By induction on the list.
+The proof is done by induction on the list.
 The case of an empty list is trivial.
@@
 \qed
-\end{proof}

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
save

@@
 One way to increase the confidence in the software we build is to formally prove its correctness using a proof assistant.
-These proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
-The appearance of these languages, and the way to write programs in them, varies from one to another.
+Proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
 Many of them are functional programming languages, like Coq, Idris and Agda, and others, like the B-Toolkit belong to the imperative paradigm.
@@
 These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software.
-However, these guarantee tend to be too often consider as perfect, when they are in fact far from it.
+However, these guarantees tend to be too often consider as perfect, when they are in fact far from it.
 Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
@@
 
-Say we want to implement a formally verified sorting function that works for list of elements of type $T$, where $T$ is ordered by a relation $\leq$.
-We can decide to only use a ``weak" type, like $sort\ :\ \{T:Type\}\ \rightarrow\ List\ T\ \rightarrow\ List\ T$, and to use an external lemma to ensure the correctness of the function.
+Say we want to implement a formally verified sorting function for list of elements of type $T$, where $T$ is ordered by a relation $\leq$.
+We can decide to define the sorting function with a ``weak" type, like $sort\ :\ List\ T\ \rightarrow\ List\ T$, and to use an external lemma to ensure the correctness of the function.
 Which property does this function has to respect?
-First, the output has to be sorted.
-So first, we need to define this notion of being sorted, here as an inductive predicate :
+First, the output has to be sorted, so we need to define this notion of being sorted, here as an inductive predicate :
 
@@
     SingletonIsSorted : (Tord : Order T) -> (x:T) -> isSorted Tord [x]
-    ConsSorted : {Tord : Order T} -> (h1:T) -> (h2:T) -> (l:List T) 
+    ConsSorted : {Tord : Order T} -> (h1:T) -> (h2:T) -> (t:List T) 
                  -> (isSorted Tord (h2::t)) -> (h1 $\leq$ h2) 
@@
 The first and second constructor of this predicate say that $[\ ]$ and $[x]$ are sorted according to any order, and for any $x$.
-The third one says that a list of two or more elements is sorted if the head $h1$ is lower or equal than the next element $h2$, and if the list deprived from its head is also sorted.
+The third one says that a list of two or more elements is sorted if $h1 \le h2$, and if the list deprived from its head is also sorted.
 In order to express that the result produced by the function $sort$ is sorted, we can write and prove the following lemma:
-$sort\_correct : \forall\ (T:Type)\ (l:List\ T),\ isSorted\ (sort\ l)$.
+$sort\_correct : \forall\ (T:Type)\ (Tord:Order\ T)\ (l:List\ T),\ isSorted\ Tord\ (sort\ l)$.
 The problem with this specification is that it doesn't says anything about the content of the output.
@@
 Here, the problem is that the function is underspecified, and it is therefore possible to write a senseless implementation, which is unfortunately provably correct.
-Only a careful reader could realize that the lemma $sort\_correct$ forgets to mention that the input and output list should be in bijection, meaning that everything which was originally in the input list should still be in the output, and that nothing else should have been added.
+Only a careful reader could realise that the lemma $sort\_correct$ forgets to mention that the input and output list should be in bijection, meaning that everything which was originally in the input list should still be in the output, and that nothing else should have been added.
 
@@
 Imagine that the third constructor of $isSorted$, called $consSorted$ would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.
-Then, the function $sort$ could do whatever it wants, its result would be seen as ``sorted".
-Said differently, any list would be seen as ``sorted" according to this predicate, just because of this single typo, and the algorithm could for example return its input unchanged.
+With this mistake, any list would be seen as ``sorted", just because of this single typo, and the algorithm could for example return its input unchanged.
 One could object that the person doing the proof of correctness should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than the next element in the list.
-The reality is that currently, many effort are going in the direction of proof automation, where the machine builds the proof automatically.
-Many proof assistants are enhanced regularly with new automations that solves automatically some kind of goals.
+The reality is quite different because many effort are going in the direction of proof automation, which aims to let the machine generate automatically the proof for some kind of goals.
 For example, Coq has already a Ring prover~\cite{coq2005}, a solver for Presburger arithmetic, and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures~\cite{Slama2016}.
@@
 
-Thus, if we want to trust the proven software, we're now forced to believe that there is adequacy between the formal specification --written in some formal logic-- and the informal requirements.
+Thus, if we want to trust the proven software, we're now forced to believe that there is adequacy between the formal specification and the informal requirements.
 A switch has occurred.
@@
 	\item Show some basic approaches to the problem of underspecification (section~\ref{sect:naiveApproaches})
-	\item Present a way to test in the proof assistant whether the predicate effectively captures the notion that we wanted to define, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
+	\item Present a way to test the predicate in the proof assistant, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
 	\item Present how we can go a step forward by replacing these tests about the predicate by some proofs (section~\ref{sect:aStepForward})

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
save

@@
 
-When confronted to this problem of adequacy between the intuitive notion and the formalised one, a first possibility that usually requires quite a lot of effort is to formalise the notion multiple times, with different predicates, and to prove that they are equivalent.
+When confronted to this problem of adequacy between the intuitive notion and the formalised one, a first possibility is to formalise the notion multiple times, with different predicates, and to prove that they are equivalent.
 With our example, that means that we need to find another formalisation $isSorted'$ of being sorted, and to prove the following lemma.
@@
 \begin{lstlisting}
-decideIsSorted : (Tord : Order T) -> (l:List T) -> Dec(isSorted Tord v)
+decideIsSorted : (Tord : Order T) -> (l:List T) -> Dec(isSorted Tord l)
 decideIsSorted Tord [] = Yes (NilIsSorted Tord)
@@
 And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and the automatically generated proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
-
 With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.
-It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted and unsorted --according to our intention-- but we can let the machine produce the proof that the predicate holds or not on these specific values.
+It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted and unsorted but we can let the machine produce the proof that the predicate holds or not on these specific values.
 This isn't bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.

### 2016-02-21 ProofsAndTests/paper-TAP/content/3-testingInside.tex
save

@@
 
-It is often easier to generate examples of a notion than it is to precisely define it.
+The key idea that this paper wants to convey is that it is often easier to generate examples of a notion than it is to precisely define it.
 We can operate an interesting change of point of view by generating the set of terms that we precisely wanted to describe with the predicate.
@@
 Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different from the predicate that if the two notions agree, then we will have gained a great confidence on the predicate.
-We decide to use coinduction and the type Stream (a coinductive version of lists, potentially infinite) in order to generate all the sorted lists of size $n$ with what we call a generator.
+We decide to use coinduction and the type Stream (a coinductive version of lists, potentially infinite) in order to generate --with what we call a generator-- all the sorted lists of size $n$.
 
@@
 
-To do so, the type $T$ needs to be $RecursivelyEnumerable$, which means that there should there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
+To do so, the type $T$ needs to be $RecursivelyEnumerable$, which means that there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
 
@@
   let x = generateSortedList T TisRecEnu TisOrdered n in
-    let y = LLmap (\l => let res = decideIsSorted TisOrdered l in
+    let y = Smap (\l => let res = decideIsSorted TisOrdered l in
 			    case res of

### 2016-02-21 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
save

@@
 In order to go a step forward, we can decide to replace the tests on the predicate by proofs.
-Instead of testing the predicate on a finite subset of the generated sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
+Instead of testing the predicate on a finite subset of all the generated sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
 
@@
 
-This lemma has the advantage of not requiring the predicate to be decidable, when this was needed in the previous section where we tested the predicate on a finite observation.
+This lemma has the advantage of not requiring the predicate to be decidable, when this was needed in the previous section where we automatically tested the predicate on a finite observation.
 However, one could object that this lemma is itself built by using a predicate, $All$, and that we can't necessary trust blindly such a specification.
@@
 The proof is done by induction on the list.
-The case of an empty list is trivial.
-For a singleton $[x]$, the proof uses the fact that the $T$ is recursively enumerable, and more precisely that the provided map is surjective, which means that it has hit any value, and thus in particular $x$.
-When there are two or more elements, we know that the smaller list deprived from its head is sorted according to $isSorted$ by construction.
-By using the lemma recursively on this smaller list together with the fact that the head is lower than the next element and the fact that the map is surjective, we get what we need.
-\qed
+For a singleton $[x]$, the proof uses the fact that $T$ is recursively enumerable, which means that the map has hit any value, and thus in particular $x$.
+When there are two or more elements, we know that $isSorted$ holds on the smaller list deprived from its head by construction.
+We finish the proof by using the lemma recursively on this smaller list together with the fact that the head is lower than the next element and the fact that the map is surjective.\qed

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
save 2

@@
 \begin{lstlisting}
-data isSorted : {T:Type} -> (Tord : Order T) 
-                 -> (l:List T) -> Type where
+data isSorted : {T:Type} -> (Tord : Order T) -> (l:List T) -> Type where
     NilIsSorted : (Tord : Order T) -> isSorted Tord []
@@
 With this mistake, any list would be seen as ``sorted", just because of this single typo, and the algorithm could for example return its input unchanged.
-One could object that the person doing the proof of correctness should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than the next element in the list.
+One could object that when doing the proof of correctness, we should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than its next element.
 The reality is quite different because many effort are going in the direction of proof automation, which aims to let the machine generate automatically the proof for some kind of goals.
@@
 	\item Show some basic approaches to the problem of underspecification (section~\ref{sect:naiveApproaches})
-	\item Present a way to test the predicate in the proof assistant, by automatically generating terms, and we automate these tests as far as possible (section~\ref{sect:testingInside})
+	\item Present a new way to test the predicate in the proof assistant, by automatically generating terms, and we completely automate these tests (section~\ref{sect:testingInside})
 	\item Present how we can go a step forward by replacing these tests about the predicate by some proofs (section~\ref{sect:aStepForward})
@@
 
-We use Idris~\cite{brady2013idris}, a dependently typed programming language that can also be seen as a proof assistant, but all the ideas that we present here can be applied to any proof assistant based on type theory.
+We use Idris~\cite{brady2013idris}, a dependently typed programming language, but all the ideas that we present here can be applied to any proof assistant based on type theory.

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
save 2

@@
 That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
-For example, with the predicate $isSorted$ defined in section 1, we can prove that it holds on the list [3, 5, 7] that we know sorted according to our informal sense of what being sorted is.
+For example, with the predicate $isSorted$ defined above, we can prove that it holds on the list [3, 5, 7] that we know sorted.
 
@@
 \end{lstlisting}
-And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and the automatically generated proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
+And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and a proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
 With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.

### 2016-02-21 ProofsAndTests/paper-TAP/content/5-conclusion.tex
Finished
save 2

@@
 Our new technique is based on the automatic generation of terms that should have the desired property.
-Thus, our goal was to make sure that the predicate and a generator of terms coincide.
-This adequacy between the predicate and the generator helps to gain confidence in the predicate, because generating examples of something and giving a definition of the same thing are tasks which are sufficiently different that it is very unlikely that they could coincide by luck.
-The technique presented on section \ref{sect:testingInside} was based on a finite observation of the automatically generated terms processed by a decision procedure.
+This adequacy between the predicate and the generator helps to gain confidence in the predicate.
+The technique presented on section \ref{sect:testingInside} was based on a finite observation of the infinitely many generated terms processed by a decision procedure.
 We've also shown on an example how these tests can be completely automated.
@@
 We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the correctness of a function relatively to a bad specification.
-
-A natural extension of the work presented here is to make it more generic, and to study how proof assistants could help --by providing some primitive bricks-- to gain confidence in the formal definitions that we write.
-A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own notion.
+A natural extension of the work presented here is to make it more generic, and to study how proof assistants could provide some primitive bricks that could help to gain confidence in the formal definitions that we write.
+A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own properties.

### 2016-02-21 ProofsAndTests/paper-TAP/content/1-introduction.tex
TAP 06 paper submitted !

@@
 Many of them are functional programming languages, like Coq, Idris and Agda, and others, like the B-Toolkit belong to the imperative paradigm.
-These different paradigm are internally supported by different logic.
-Systems like Coq, Idris and Agda are based on various higher order logic (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic.
+These different paradigms are internally supported by different logic.
+Systems like Coq, Idris and Agda are based on various higher order logics (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic.
 These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software.
-However, these guarantees tend to be too often consider as perfect, when they are in fact far from it.
+However, these guarantees tend to be too often considered as perfect, when they are in fact far from it.
 Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
@@
 The third one says that a list of two or more elements is sorted if $h1 \le h2$, and if the list deprived from its head is also sorted.
-In order to express that the result produced by the function $sort$ is sorted, we can write and prove the following lemma:
+In order to express that the result of $sort$ is sorted, we can prove the following lemma:
 $sort\_correct : \forall\ (T:Type)\ (Tord:Order\ T)\ (l:List\ T),\ isSorted\ Tord\ (sort\ l)$.
@@
 Imagine that the third constructor of $isSorted$, called $consSorted$ would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.
-With this mistake, any list would be seen as ``sorted", just because of this single typo, and the algorithm could for example return its input unchanged.
+Any list would be seen as ``sorted", just because of this single typo, and the algorithm could for example return its input unchanged.
 One could object that when doing the proof of correctness, we should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than its next element.
 The reality is quite different because many effort are going in the direction of proof automation, which aims to let the machine generate automatically the proof for some kind of goals.
-For example, Coq has already a Ring prover~\cite{coq2005}, a solver for Presburger arithmetic, and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures~\cite{Slama2016}.
+For example, Coq has already a Ring prover~\cite{coq2005} and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures~\cite{Slama2016}.
 There are even extensions to languages, such as Ltac~\cite{DelahayeLTac} and Mtac~\cite{Ziliani13} that aim to help the automation of tactics.

### 2016-02-21 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
TAP 06 paper submitted !

@@
 The biggest problem with this approach is to be able to find some alternative formalisations that are sufficiently different from the original one.
-Obviously, if the new normalisations are too similar to the original one (and in the worst case the new ones are just syntactical variants of the first one), then we won't gain any guarantee.
-The ideal would be to capture the same notion by using very different point of view, and we will show in section~\ref{sect:testingInside} an original approach for doing so.
+Obviously, if the new formalisations are too similar to the original one (and in the worst case the new ones are just syntactical variants of the first one), then we won't gain any guarantee.
+The ideal would be to capture the same notion by using very different points of view, and we will show in section~\ref{sect:testingInside} an original approach for doing so.
 
@@
 In order to gain confidence in the formal specifications we write, another traditional approach is to test the predicate on some values.
-That consists of defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
+That consists in defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
 For example, with the predicate $isSorted$ defined above, we can prove that it holds on the list [3, 5, 7] that we know sorted.
@@
 \end{lstlisting}
-And if we ask to evaluate $isSorted\_test1'$, the system will answer $Yes$ and a proof of $isSorted\ [3,\ 5,\ 7]$, which is what we want, and we therefore know that the predicate has passed this test.
+And if evaluate $isSorted\_test1'$, the system will answer $Yes$ and a proof of $isSorted\ [3,\ 5,\ 7]$, which means that the predicate has passed this test.
 With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.

### 2016-02-21 ProofsAndTests/paper-TAP/content/3-testingInside.tex
TAP 06 paper submitted !

@@
 \end{lstlisting}
-Now, instead of simply generating the first $m$ sorted lists, we want to run the decision procedure on all of these $m$ tests in order to know if the predicate and the generator agree on this portion.
+Now, instead of simply generating the first $m$ sorted lists, we run the decision procedure on all of these $m$ tests in order to know if the predicate and the generator agree on this portion.
 The result will be a vector $m$ booleans.

### 2016-02-21 ProofsAndTests/paper-TAP/content/5-conclusion.tex
TAP 06 paper submitted !

@@
 
-We've presented some new ideas and also reintroduced some known techniques that aim to gain confidence in a predicate inside a proof assistant.
-Our new technique is based on the automatic generation of terms that should have the desired property.
+We've presented a new way to test a predicate based on an automatic generation of terms that should have the desired property.
 This adequacy between the predicate and the generator helps to gain confidence in the predicate.
@@
 We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the correctness of a function relatively to a bad specification.
-A natural extension of the work presented here is to make it more generic, and to study how proof assistants could provide some primitive bricks that could help to gain confidence in the formal definitions that we write.
+A natural extension of the work presented here is to make it more generic, and to study how proof assistants could provide some primitive bricks that could help to gain confidence in logical properties.
 A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own properties.

### 2016-04-25 ProofsAndTests/paper-TAP/content/3-testingInside.tex
[TAP 2016] Making the code match exactly the paper

@@
 \end{lstlisting}
+
+However, if the predicate $isSorted$ has been incorrectly written, for example with the broken condition $(h1 \leq h1)$ instead of $(h1 \leq h2)$ in its third constructor $consSorted$, then the result of this test will be :
+
+\begin{lstlisting}
+[True, True, True, True, True, True, True, True] : Vect 8 Bool
+\end{lstlisting}
+
 When we want to test $isSorted$ on a big number of tests, we might not want to inspect manually the result of each test.

### 2016-04-25 ProofsAndTests/paper-TAP/content/3-testingInside.tex
continuing...

@@
 \begin{lstlisting}
-testGenSorted : (m:Nat) -> (n:Nat) -> Maybe(Vect m (List T))
-testGenSorted m n = let x = generateSortedList T TisRecEnu TisOrdered n 
+testGenerator : (m:Nat) -> (n:Nat) -> Maybe(Vect m (List T))
+testGenerator m n = 
+	let x = generateSortedList T TisRecEnu TisOrdered n 
 	    in unfold_n_times x m
 \end{lstlisting}
-We can ask for the first 8 sorted lists of size 5 by evaluating $testGenSorted\ 8\ 5$:
+We can ask for the first 8 sorted lists of size 4 by evaluating $testGenerator\ 8\ 4$:
 
@@
 Just [[A, A, A, A], [A, A, A, B], [A, A, A, C], [A, A, B, B],
-      [A, A, B, C], [A, A, C, C], [A, B, B, B], [A, B, B, C]] : Maybe (Vect 8 (List T))
+      [A, A, B, C], [A, A, C, C], [A, B, B, B], [A, B, B, C]] 
+      : Maybe (Vect 8 (List T))
 \end{lstlisting}
@@
 \end{lstlisting}
-And we can inspect the result of running the first 8 tests by evaluating $testSorted\ 8\ 5$.
+And we can inspect the result of running the first 8 tests of size 4 by evaluating $testSorted\ 8\ 4$.
 
@@
 
-However, if the predicate $isSorted$ has been incorrectly written, for example with the broken condition $(h1 \leq h1)$ instead of $(h1 \leq h2)$ in its third constructor $consSorted$, then the result of this test will be :
-
-\begin{lstlisting}
-[True, True, True, True, True, True, True, True] : Vect 8 Bool
-\end{lstlisting}
+However, if the predicate $isSorted$ has been incorrectly written, then the result of this test might inform us that there's something wrong with the formal specification.
+For example, if we've forgotten the third constructor $consSorted$ in the definition of $isSorted$, then the result of (testSorted 8 4) will be : [False, False, False, False, False, False, False, False] because no produced list will be seen as sorted according to the predicate, and we will therefore know that the predicate $isSorted$ does not capture our intuitive notion of sortedness.
 
@@
 We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the boolean $And$ on each element of the resulting vector.
-Now, we can for example test the predicate on the first 50 sorted lists of size 8 by running $testSorted\_result\ 50\ 8$ and if we do so we get the overall result $True$ which means that the predicate agrees with the generator on all these 50 tests.
+Now, we can for example test the predicate on the first 50 sorted lists of size 9 by running $testSorted\_result\ 50\ 9$ and if we do so we get the overall result $True$ which means that the predicate agrees with the generator on all these 50 tests.

### 2016-04-27 ProofsAndTests/paper-TAP/content/3-testingInside.tex
paper

@@
 
+When we want to test $isSorted$ on a big number of tests, we might not want to inspect manually the result of each test.
+We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted\ m\ n$ and does the boolean $And$ on each element of the resulting vector.
+Now, we can for example test the predicate on the first 50 sorted lists of size 9 by running $testSorted\_result\ 50\ 9$ and if we do so we get the overall result $True$ which means that the predicate agrees with the generator on all these 50 tests.
+
 However, if the predicate $isSorted$ has been incorrectly written, then the result of this test might inform us that there's something wrong with the formal specification.
-For example, if we've forgotten the third constructor $consSorted$ in the definition of $isSorted$, then the result of (testSorted 8 4) will be : [False, False, False, False, False, False, False, False] because no produced list will be seen as sorted according to the predicate, and we will therefore know that the predicate $isSorted$ does not capture our intuitive notion of sortedness.
-
-When we want to test $isSorted$ on a big number of tests, we might not want to inspect manually the result of each test.
-We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted$ and does the boolean $And$ on each element of the resulting vector.
-Now, we can for example test the predicate on the first 50 sorted lists of size 9 by running $testSorted\_result\ 50\ 9$ and if we do so we get the overall result $True$ which means that the predicate agrees with the generator on all these 50 tests.
+For example, if we've forgotten the third constructor $consSorted$ in the definition of $isSorted$, then the result of $(testSorted\ 8\ 4)$ will be $False$, which means that at least one of the produced list is not seen as sorted according to $isSorted$, and we will therefore know that this predicate  does not capture our intuitive notion of sortedness.

### 2016-04-27 ProofsAndTests/paper-TAP/content/5-conclusion.tex
paper

@@
 This adequacy between the predicate and the generator helps to gain confidence in the predicate.
-The technique presented on section \ref{sect:testingInside} was based on a finite observation of the infinitely many generated terms processed by a decision procedure.
+The technique presented on section \ref{sect:testingInside} was based on a finite observation of the infinite quantity of generated terms processed by a decision procedure.
 We've also shown on an example how these tests can be completely automated.
@@
 We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the correctness of a function relatively to a bad specification.
-A natural extension of the work presented here is to make it more generic, and to study how proof assistants could provide some primitive bricks that could help to gain confidence in logical properties.
-A possibility is to provide many robust and generic concepts, like being sorted, so that the user won't have to define his own properties.
+One could however question why formal certification is needed at all, if after going through all the effort of interactive theorem proving we still have to test the specification itself and need tools to support it.
+We believe that the process of formalisation precisely helps to uncover things that were missing or weakly specified in the informal requirements.
+For example, the development of the formally verified compiler CompCert~\cite{Krebbers2014} has contributed to brought to light many under-specified behaviours in the specification of the C standard.
+
+A natural extension of the work presented here is to study how proof assistants themselves could help to gain confidence about predicates and logical formulae.
+One possible way could be to see how proof assistants could be equipped with an execution engine for formal specifications, that would replace the very specific machinery that we've showed in the running example of this paper.
+Another possibility, which in fact complements the first, is to have the proof assistant providing many robust and generic concepts once and for all (like the predicate being sorted), because that would save the user from the error prone activity of writing many primitive logical properties.
+Equipping proof assistants with many generic and useful concepts already available, like bricks ready to be assembled, is one of the current challenge to make proofs assistants really usable.

### 2016-04-27 ProofsAndTests/paper-TAP/content/1-introduction.tex
continuing...
almost ready for the camera-ready version

@@
 Knuth was --certainly ironically-- saying ``Beware of bugs in the above code; I have only proved it correct, not tried it".
-The reality is precisely that a proof isn't enough.
+The reality is precisely that a proof is not enough.
 When we prove the correctness of a function, we only gain the guarantee expressed by the proven lemma, and nothing more.
@@
 \begin{lstlisting}
-data isSorted : {T:Type} -> (Tord : Order T) -> (l:List T) -> Type where
+data isSorted : {T:Type} -> (Order T) -> (List T) -> Type where
     NilIsSorted : (Tord : Order T) -> isSorted Tord []
@@
 $sort\_correct : \forall\ (T:Type)\ (Tord:Order\ T)\ (l:List\ T),\ isSorted\ Tord\ (sort\ l)$.
-The problem with this specification is that it doesn't says anything about the content of the output.
+The problem with this specification is that it does not say anything about the content of the output.
 The function $sort$ could just return the empty list $[\ ]$ all the time, it would still be possible to prove this correctness lemma.
@@
 
-Things more nasty can happen.
+Another bug in the specification could have been to simply forget the third constructor $ConsSorted$.
+But things more nasty can happen.
 Imagine that the third constructor of $isSorted$, called $consSorted$ would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.

### 2016-04-27 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
almost ready for the camera-ready version

@@
 In order to gain confidence in the formal specifications we write, another traditional approach is to test the predicate on some values.
-That consists in defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it doesn't when it shouldn't.
+That consists in defining a few terms, usually by hand, for which we know if the predicate should hold or not, and to prove that the predicate effectively holds when it should, and that it does not when it should not.
 For example, with the predicate $isSorted$ defined above, we can prove that it holds on the list [3, 5, 7] that we know sorted.
@@
   let p2 : (5 <= 7) = tryDec (lowerEqDec natIsOrdered 5 7) in
-    ConsSorted 3 5 [7] (ConsSorted 5 7 [] (SingletonIsSorted _ 7) p2) p1
+    ConsSorted 3 5 [7] 
+      (ConsSorted 5 7 [] (SingletonIsSorted _ 7) p2) p1
 \end{lstlisting}
@@
 \begin{lstlisting}
-decideIsSorted : (Tord : Order T) -> (l:List T) -> Dec(isSorted Tord l)
+decideIsSorted : (Tord : Order T) -> (l:List T) 
+                 -> Dec(isSorted Tord l)
 decideIsSorted Tord [] = Yes (NilIsSorted Tord)
@@
 decideIsSorted Tord (h1::(h2::t)) with (lowerEqDec Tord h1 h2)
-  decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
-          with (decideIsSorted Tord (h2::t))
-    decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
-          | (Yes h2_tail_sorted) = Yes (ConsSorted h1 h2 t h2_tail_sorted h1_lower_h2)
-    decideIsSorted Tord (h1::(h2::t)) | (Yes h1_lower_h2) 
+  | (Yes h1_lower_h2) with (decideIsSorted Tord (h2::t))
+      | (Yes h2_tail_sorted) = Yes 
+      		(ConsSorted h1 h2 t h2_tail_sorted h1_lower_h2) 
       | (No h2_tail_not_sorted) = No [...]
-  decideIsSorted Tord (h1::(h2::t)) | No h1_not_lower_h2 = No [...]
+  | (No h1_not_lower_h2) = No [...]
 \end{lstlisting}
@@
 \end{lstlisting}
-And if evaluate $isSorted\_test1'$, the system will answer $Yes$ and a proof of $isSorted\ [3,\ 5,\ 7]$, which means that the predicate has passed this test.
+And if we evaluate $isSorted\_test1'$, the system will answer $Yes$ and a proof of $isSorted\ [3,\ 5,\ 7]$, which means that the predicate has passed this test.
 With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.
 It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted and unsorted but we can let the machine produce the proof that the predicate holds or not on these specific values.
-This isn't bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.
+This is not bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.

### 2016-04-27 ProofsAndTests/paper-TAP/content/3-testingInside.tex
almost ready for the camera-ready version

@@
 
-$generateSortedList : (T:Type)\ \rightarrow\ (recEnu:RecursivelyEnumerable\ T) \rightarrow\ (Tord : Order\ T) \rightarrow\ (n:Nat) \rightarrow Stream\ (List\ T)$.
+$generateSortedList : (T:Type)\ \rightarrow\ (recEnu:RecEnum\ T) \rightarrow\ (Tord : Order\ T) \rightarrow\ (n:Nat) \rightarrow Stream\ (List\ T)$.
 
-To do so, the type $T$ needs to be $RecursivelyEnumerable$, which means that there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
+To do so, the type $T$ needs to be recursively enumerable, which means that there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
 

### 2016-04-27 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
almost ready for the camera-ready version

@@
 \begin{lstlisting}
-generated_implies_pred_holds : {T:Type} -> (recEnu:RecursivelyEnumerable T) 
-        -> (Tord : Order T) -> (n:Nat) 
-        -> (All (generateSortedList T recEnu Tord n) (\l => isSorted Tord l))
+generated_implies_pred_holds : {T:Type} 
+   -> (recEnu:RecEnum T) -> (Tord : Order T) -> (n:Nat) 
+   -> (All (generateSortedList T recEnu Tord n) 
+           (\l => isSorted Tord l))
 \end{lstlisting}

### 2016-04-27 ProofsAndTests/paper-TAP/paper-TAP.tex
Finished
paper
almost ready for the camera-ready version

@@
 Why should we trust more these predicates than the code that we've first refused to trust blindly, leading to these proofs?
-We show in this paper that the proving activity hasn't replaced the testing activity but has only changed the object which requires to be tested.
+We show in this paper that the proving activity has not replaced the testing activity but has only changed the object which requires to be tested.
 Instead of testing code, we now need to test predicates.
@@
 
-\keywords{Formal certification, specification, predicate testing, proof assistant, tests, requirements}
+\keywords{Formal certification, predicate testing, proof assistant}
 \end{abstract}
@@
 \bibliographystyle{splncs03}
+
+\nocite{DeMillo1979}
+
 \bibliography{bib-TAP}

### 2016-04-28 ProofsAndTests/paper-TAP/content/1-introduction.tex
TAP paper

@@
 Proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
-Many of them are functional programming languages, like Coq, Idris and Agda, and others, like the B-Toolkit belong to the imperative paradigm.
+Many of them are functional programming languages, like Coq~\cite{Bertot2004}, Idris~\cite{brady2013idris} and Agda~\cite{Norell2008}, and others, like the B-Method~\cite{Abrial1991} belong to the imperative paradigm.
 These different paradigms are internally supported by different logic.
@@
 
-We use Idris~\cite{brady2013idris}, a dependently typed programming language, but all the ideas that we present here can be applied to any proof assistant based on type theory.
+We use Idris, a dependently typed programming language, but all the ideas that we present here can be applied to any proof assistant based on type theory.
+The running example that we use in this paper can be found online at \url{https://github.com/FranckS/}.

### 2016-04-28 ProofsAndTests/paper-TAP/content/4-differentKindOfProperties.tex
TAP paper

@@
 In order to go a step forward, we can decide to replace the tests on the predicate by proofs.
-Instead of testing the predicate on a finite subset of all the generated sorted lists, as we did in the previous section, we can prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
+Instead of testing the predicate on a finite subset of all the generated sorted lists, as we did in the previous section, we can try to prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
 

### 2016-04-28 ProofsAndTests/paper-TAP/content/1-introduction.tex
paper TAP almost ready

@@
 
-One way to increase the confidence in the software we build is to formally prove its correctness using a proof assistant.
+One way to increase our confidence in software is to formally prove its correctness using a proof assistant.
 Proofs assistants enable to write code, logical statements and proofs in the same language, and offer the guarantee that every proof will be automatically checked.
@@
 These different paradigms are internally supported by different logic.
-Systems like Coq, Idris and Agda are based on various higher order logics (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the formal B method is based on Hoare logic.
-These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence of the produced software.
+Systems like Coq, Idris and Agda are based on various higher order logics (CoC, a variant of ML and LUO respectively) and are realisations of the Curry-Howard correspondence, while the B-Method is based on Hoare logic.
+These different foundations lead to different philosophies and different ways to implement and verify a software, but all of them greatly increase the confidence on the produced software.
 However, these guarantees tend to be too often considered as perfect, when they are in fact far from it.
@@
 But things more nasty can happen.
-Imagine that the third constructor of $isSorted$, called $consSorted$ would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.
+Imagine that this constructor would have been written with a typo, and that the condition $(h1 \leq h2) $ would have been incorrectly written as $(h1 \leq h1)$.
 Any list would be seen as ``sorted", just because of this single typo, and the algorithm could for example return its input unchanged.
 One could object that when doing the proof of correctness, we should realize that the proof is being done too easily, without having to use the essential property that the output is being built such as any element in the list is always lower or equal than its next element.
-The reality is quite different because many effort are going in the direction of proof automation, which aims to let the machine generate automatically the proof for some kind of goals.
+The reality is quite different because many effort are going in the direction of proof automation, which aims to let the machine automatically generate the proof for some kind of goals.
 For example, Coq has already a Ring prover~\cite{coq2005} and many others automations, and Idris has been recently equipped with a hierarchy of provers for algebraics structures~\cite{Slama2016}.
@@
 But when the specification is too often as complicated as the code, why should we blindly believe in it, when we've first refused to blindly trust the code?
-The primary aim of this paper is to raise awareness of the adequacy concern, and to see how heterogeneous approaches, that mix both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
+The primary aim of this paper is to raise awareness on the adequacy concern, and to see how heterogeneous approaches, that mixes both proofs and tests, can help to go a step forward in the certification process, in the context of proof assistants based on type theory.
 More precisely, we :
@@
 	\item Show some basic approaches to the problem of underspecification (section~\ref{sect:naiveApproaches})
-	\item Present a new way to test the predicate in the proof assistant, by automatically generating terms, and we completely automate these tests (section~\ref{sect:testingInside})
-	\item Present how we can go a step forward by replacing these tests about the predicate by some proofs (section~\ref{sect:aStepForward})
+	\item Present a new way to test the predicate in the proof assistant, by automatically generating terms, and we completely automate these tests.
+We also show how we can go a step forward by replacing these tests about the predicate by some proofs.
+(section~\ref{sect:testingInside})
+	\item We discuss possible directions for making dependently typed programming languages more adapted to the testing of specifications (section~\ref{sect:Conclusion}).
 \end{itemize}

### 2016-04-28 ProofsAndTests/paper-TAP/content/2-specCanGoWrong.tex
paper TAP almost ready

@@
 With this technique, we can run semi-automatically a few tests on the predicate $isSorted$.
-It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted and unsorted but we can let the machine produce the proof that the predicate holds or not on these specific values.
-This is not bad --and this is in fact all that is done usually, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.
+It is semi-automatic in the sense that we still have to define by hand some terms that we know sorted or unsorted but we can let the machine produce the proof that the predicate holds or not on these specific values.
+This is not too bad --and this is in fact all of what is usually done, when it is actually done-- but we would like to have a stronger guarantee, and not only that the predicate will coincide with our intuitive notion on a couple of tested terms.

### 2016-04-28 ProofsAndTests/paper-TAP/content/3-testingInside.tex
paper TAP almost ready

@@
 With our example of sorted lists, that means that we need to define the same notion of being sorted, but this time with a definition example-based.
-Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different from the predicate that if the two notions agree, then we will have gained a great confidence on the predicate.
+Writing a function that produces all the sorted lists might be a bit more difficult than just giving a few examples, but this complicated function will have the main advantage of being so different from the predicate that if the two notions agree, then we will have gain a great confidence on the predicate.
 We decide to use coinduction and the type Stream (a coinductive version of lists, potentially infinite) in order to generate --with what we call a generator-- all the sorted lists of size $n$.
@@
 
-To do so, the type $T$ needs to be recursively enumerable, which means that there should exists a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that all value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
+To do so, the type $T$ needs to be recursively enumerable, which means that there must exist a computational map $Nat\ \rightarrow\ Maybe\ c$ with the condition that this map is surjective, which means that any value of type $c$ should be hit at least once by the map : $map\_is\_surjective : (y:c)\ \rightarrow\ (x:Nat\ **\ (computableMap\ x\ =\ Just\ y))$.
 
 We want to check that this function and the predicate coincide.
-Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running the decision procedure.
+Since the predicate $isSorted$ is decidable with $decideIsSorted$, we can automatically check whether the generated sorted lists of size $n$ are automatically sorted (according to $isSorted$) by running this decision procedure.
 But since there can be infinitely many generated sorted lists of size $n$ when $n > 0$, we will only check that the generator and the predicate coincide on a finite observation\footnote{A finite observation of a stream, also called approximation at rank $m$ of a stream, is a vector of size $m$ that has the same $m$ first elements than the stream} of the resulting stream.
@@
 
-When we want to test $isSorted$ on a big number of tests, we might not want to inspect manually the result of each test.
+When we want to test $isSorted$ on a large number of tests, we might not want to inspect manually the result of each test.
 We can write a function $testSorted\_result : (m:Nat)\ \rightarrow\ (n:Nat)\ \rightarrow\ Bool$ that calls $testSorted\ m\ n$ and does the boolean $And$ on each element of the resulting vector.
@@
 For example, if we've forgotten the third constructor $consSorted$ in the definition of $isSorted$, then the result of $(testSorted\ 8\ 4)$ will be $False$, which means that at least one of the produced list is not seen as sorted according to $isSorted$, and we will therefore know that this predicate  does not capture our intuitive notion of sortedness.
+
+In order to go a step forward, we can decide to replace the tests on the predicate by proofs.
+Instead of testing the predicate on a finite subset of all the generated sorted lists as we just did, we can try to prove that any of the automatically generated sorted list is provably sorted according to $isSorted$.
+
+\begin{lstlisting}
+generated_implies_pred_holds : {T:Type} 
+   -> (recEnu:RecEnum T) -> (Tord : Order T) -> (n:Nat) 
+   -> (All (generateSortedList T recEnu Tord n) 
+           (\l => isSorted Tord l))
+\end{lstlisting}
+\begin{proof}
+By induction on $n$.
+When $n$ is zero, there is only one sorted list generated, which is the empty list, and we know that the empty list is sorted thanks to the constructor $NilIsSorted$.
+When n is some successor $(S\ pn)$, we know by using recursively the lemma on the smaller value $pn$ that all sorted lists of size $pn$ are sorted according to $isSorted$.
+Since the $Stream$ of all sorted lists of size $(S\ pn)$ has been made from the $Stream$ of all sorted lists of size $pn$ by adding to all of them --on the head position-- an element lower or equal to their respective current heads, we know that the property has been preserved at the higher rank.
+\qed
+\end{proof}
+
+This lemma has the advantage of not requiring the predicate to be decidable, whereas this was needed when we automatically tested the predicate on a finite observation.
+However, one could object that this lemma is itself built by using a predicate, $All$, and that we can't necessary trust blindly such a specification.
+The answer is that no guarantee is perfect, and all we can do is to add some guarantees, but there is necessarily always something to trust.
+Moreover, this new kind of specifications and proofs --about the predicate itself-- uses more primitive components like streams and the predicate $All$, and these components can be provided once and for all.
+If they are part of some standard library being used intensively, there is very low risk that they do not capture the desired semantic.

### 2016-04-28 ProofsAndTests/paper-TAP/paper-TAP.tex
paper TAP almost ready

@@
 % first the title is needed
-\title{Short paper : Automatic predicate testing in formal certification}
+\title{Automatic predicate testing in formal certification}
 \subtitle{You've only proven what you've said, not what you meant!}
@@
 There is always a jump when going from an informal specification to a formal specification expressed in a logical theory.
-Thus, proving the correctness of a piece of software always makes the implicit assumption that there is adequacy between the formalised specification --written with logical statements and predicates-- and the real requirements --often written in english--.
+Thus, proving the correctness of a piece of software always makes the implicit assumption that there is adequacy between the formalised specification --written with logical statements and predicates-- and the real requirements --often written in English--.
 Unfortunately, a huge part of the complexity lies precisely in the specification itself, and it is far from obvious that the formal specification says exactly and completely what it should say.
@@
 
-\input{./content/4-differentKindOfProperties.tex}
-
-\input{./content/5-conclusion.tex}
+\input{./content/4-conclusion.tex}
 

### 2016-04-28 ProofsAndTests/paper-TAP/content/4-conclusion.tex
added the file with the updated name

\section{Conclusions and future work}
\label{sect:Conclusion}

In this paper we've presented a new way to test a predicate based on an automatic generation of terms that should have the desired property. This adequacy between the predicate and the generator helps to gain confidence in the predicate. The technique presented on section \ref{sect:testingInside} was based on a finite observation of the terms generated in endless amount, processed by the decision procedure. We have also shown on an example how these tests can be automated.

We haven't been able to find much work done in the direction of predicate testing in the environment of proof assistants, but we strongly believe that this aspect is crucial, as there is absolutely no point to prove the ``correctness" of a function relatively to a bad specification. One could however question why formal certification is needed at all, if after going through all the effort of interactive theorem proving we still have to test the specification itself, and also need tools to support it. We believe that the process of formalisation helps to uncover things that were missing or weakly specified in the informal requirements. For example, the development of the formally verified compiler CompCert~\cite{Krebbers2014} has contributed to brought to light many under-specified behaviours in the specification of the C standard.

The machinery developed for the running example presented in this paper is extremely specific and it is not reasonable to believe that this work should and could be done for every formal specification. What it shows is that we really need to explore how proof assistants themselves could help to gain confidence about predicates and logical formulae. A possible direction could be to build execution engines for formal specifications written in dependent type theories. Such a system would take in input a predicate and would produce some of the terms that make this predicate hold. The ideal would be to have a query system where one could ask the system to try to look if some specific terms are captured by the predicate. Since the problem of finding proof is undecidable in the general case in higher-order logics (that's also the case in first-order logic), such a system can't be complete and entirely automatic, and therefore the user would have to help the system at times.

Another possible direction is to equip proof assistants with many robust and generic concepts (like being sorted) once and for all. That would save the user from the error prone activity of writing many primitive logical properties. Equipping proof assistants with many generic and useful concepts already available, like bricks ready to be assembled, is another current challenge to make proofs assistants really usable.

### 2016-06-05 Old/NotesOnCorrectnessAndCompleteness/notesOnCorrectnessAndCompleteness.tex
Moved a folder and updated the README

--- 2015-10-01 NotesOnCorrectnessAndCompleteness/notesOnCorrectnessAndCompleteness.tex
@@
 
+\abstract
+I compare briefly Coq's ring prover and the provers for different algebraic structures we've built for Idris. I just focus on the correctness and completeness in theses notes.
+
 \section{Correctness}
@@
 This index represents the concrete term (of type $C$) which is encoded by this encoding.
-For each element $e$ of type $ExprC\ c$, we will note $e_c$ to explicitly show the index of this term.
+For each element $e$ of type $ExprC\ c$, we will note $e_c$ to explicitly show the index of this term (which can entirely be computed by the type-checker).
 \\
@@
 Now, we're going to compare syntactically $(norm((\tau\ a)_a))_{a'}$ and $(norm((\tau\ b)_b))_{b'}$.
-What is really important is that if they're syntactically the same, then they're forced to be indexed over the same index : if an element of $ExprC\ a'$ equals (syntactically!) en element of $ExprC\ b'$, then this implies $a'=b'$\footnote{In fact, in our implementation, we don't even bother and this function of syntactical comparison directly produces the proof of equality of the indices if the two reflected terms are syntactically the same : it has for type $(ExprC\ a') \rightarrow (ExprC\ b')\ \rightarrow\ Maybe(a'=b')$ and not $(e1:ExprC\ a') \rightarrow (e2:ExprC\ b')\ \rightarrow\ Maybe(e1=e2)$.
-This is simply because we're not interested in the proof of equality between the reflected terms, so instead of producing it, and having an external lemma $extLemma:forall (a'\ b':C)\ (e1:ExprC\ a')\ (e2:ExprC\ b'),\ (e1=e2)\ \rightarrow\ (a'=b')$, which says that if two reflected terms are equal, then their index are also equal, we simply just produces the proof we need.}.
+What is really important is that if they're syntactically the same, then they're forced to be indexed over the same index : if an element of $ExprC\ a'$ equals (syntactically!) en element of $ExprC\ b'$, then this implies \footnote{In fact, in our implementation, we don't even bother and this function of syntactical comparison directly produces the proof of equality of the indices if the two reflected terms are syntactically the same : it has for type $(ExprC\ a') \rightarrow (ExprC\ b')\ \rightarrow\ Maybe(a'=b')$ and not $(e1:ExprC\ a') \rightarrow (e2:ExprC\ b')\ \rightarrow\ Maybe(e1=e2)$.
+This is simply because we're not interested in the proof of equality between the reflected terms, so instead of producing it, and having later an external lemma $extLemma:forall (a'\ b':C)\ (e1:ExprC\ a')\ (e2:ExprC\ b'),\ (e1=e2)\ \rightarrow\ (a'=b')$, which says that if two reflected terms are equal, then their index are also equal, we simply directly produce the proof we need.} $a'=b'$ \footnote{Note that the default equality is heterogeneous in Idris}.
 \\
@@
 \\
-In fact, everything has been made possible thanks to the following property, which doesn't even need to be expressed because it is already in the type of the function $norm$ :
-$\forall (x\ y\ :\ C),\ norm ((T(x))_x) = Y_y \rightarrow x = y$
-
+In fact, everything has been made possible thanks to the following property (which doesn't even need to be expressed because it is already in the type of the function $norm$) : \\
+$\forall (x\ y\ :\ C),\ norm ((\tau(x))_x) = Y_y \rightarrow x = y$
+\\
+\\
 In their approach, because they don't follow this kind of "type-safe reflection", they do need a correctness lemma, which says : \\
-$\forall (e1:ExprC)(e2:ExprC),\ \phi_{NC}\ (norm\ e1)\ =\ \phi_C\ e1$.
-So, what will happen is that they will compute (norm\ e1) and (norm\ e2).
-
-If the normalisation of the reflection of x gives the reflection of y, then x=y
+$\forall (e:ExprC),\ \phi_{NC}\ (norm\ e)\ =\ \phi_C\ e$. \\
+So, what will happen is that they will compute $(norm\ \tau(a))$ and $(norm\ \tau(b))$, exactly like us. 
+Then, like us, they will compare these two values syntactically. If they are equal, then they get
+$(norm\ \tau(a)) = (norm\ \tau(b))$.
+By applying $\phi_{NC}$ to both sides of this equality, they get $(\phi_{NC} (norm\ \tau(a))) = \phi_{NC} ((norm\ \tau(b)))$.
+But since $\phi_{NC} (norm\ \tau(a)) = \phi_C\ (\tau(a))$ and $\phi_{NC} (norm\ \tau(b)) = \phi_C\ (\tau(b))$ (from their correctness lemma), they get : \\
+$\phi_C (\tau(a)) = \phi_C (\tau(b))$. \\
+Now, if the functions $\phi_C$ and $\tau$ have been correctly defined, we should have $\forall x, \phi_C (\tau(x)) = x$.
+The problem is that this lemma hasn't been proved, and can't even be expressed, simply because the function $\tau$ is defined by using LTac.
+Unlike the $\%syntax$ directive in Idris, which still produces "normal" (typed) functions on which we can reason, LTac intend to define untyped things, which lie outside the world of terms and types of Coq.
+\\
+However, that doesn't break the correctness of their approach, because the correctness is in fact ensured by Coq's typechecker.
+When you try to produce a proof of $a=b$, there's no risk to produce a proof of $\phi_C (\tau(a)) = \phi_C (\tau(b))$, which is indeed an equality proof on the type $C$.
+If the typechecker will be able to unify together the left hand sides $a$ and $\phi_C (\tau(a))$, and together the right hand sides $b$ and $\phi_C (\tau(b))$, then the system will consider that they have generated a valid proof of $a=b$.
+If not, then the type-system will shout.
+\\
+Obviously, this can affect the completeness of the approach.
+Let's imagine that the function $\tau$ (defined in Ltac), always return the encoding of the constant $O:C$.
+Let's also imagine that the function $\phi_C$ is well defined (ie, it pattern matches on the constructor of ExprC, and returns the underlying expressions of type $C$).
+Then, every time we will apply this tactic, it will produce $expr_0$ (reflecting $O$) for both $\tau(a)$ and $\tau(b)$.
+The normalisation will give $norm(expr_O)$ for both $norm(\tau(a))$ and $norm(\tau(b))$ (we don't care of what it actually computes), which are therefore equal :
+$norm(expr_O) = norm(expr_O)$. By applying $\phi_{NC}$ on both side we will get $\phi_{NC} (norm(expr_O)) = \phi_{NC} (norm(expr_O))$. By applying the correctness lemma 2 times, we get :
+$\phi_C (expr_O) = \phi_C (expr_O)$.
+Since $\phi_C$ is supposed to be correct, $\phi_C$ applied on the encoding of $O$ should give back $O$.
+We've therefore produce a proof of $0=0$.
+That means that this equality is the only one we can prove in this system where $\tau$ has been defined to always return the same term.
+We've completely lost the completeness.
+\\
+\\
+Here, we see clearly the big advantage of our type-safe reflection : we're absolutely sure that $\tau(x)$ is a faithful representation of $x$, because $\tau(x)$ has for type $\tau : (c:C)\ \rightarrow\ ExprC\ c$.
+And we don't need a function $\tau$ to get back the concrete values, as we already have them available at the type level !
+\\
+\\
+Now, one could be tempted to think that having the property $\forall x, \phi_C (\tau(x)) = x$ (which is not hard to get) is enough for ensuring the completeness.
+In fact, that's not enough, but that's necessary.
+As shown before, without this property, the prover they've build is still correct but might be completely unusable because completely incomplete.
+But that's not enough.
 

### 2017-03-22 paper-CICM-17/content/1-introduction.tex
Added paper for CICM-17

--- 2016-02-03 paper-JFP-DTP/content/1-introduction.tex
@@
 \section{Introduction}
+
+TO BE UPDATED LATER : (copied version from the older paper as a placeholder)
 
@@
 Very often, the type $SEM$ has some algebraic 
-algebraic structure, and it becomes thus possible to
+structure, and it becomes thus possible to
 decide equalities in it by following some standard rewriting rules.  

### 2017-03-22 paper-CICM-17/paper.tex
Added paper for CICM-17

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration

\usepackage{amsfonts}
\usepackage{float}
\usepackage{url}
\usepackage{listings}

\usepackage{tikz}
\usetikzlibrary{matrix}

\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
  keepspaces=true,
  mathescape
}

\newcommand{\figrule}{\begin{center}\hrule\end{center}}

%
\begin{document}
%
\pagestyle{headings}  % switches on printing of running heads

\mainmatter              % start of the contributions
%
\title{Automatically Proving Equivalence by Type-Safe Reflection}
\subtitle{Or how to use dependent types to solve some of the problems they have caused}
%
\titlerunning{xxx}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Franck Slama \and Edwin Brady}
\institute{University of St Andrews, Scotland, UK,\\
\email{fs39@st-andrews.ac.uk, ecb10@st-andrews.ac.uk}}

\maketitle              % typeset the title of the contribution

\begin{abstract}
One common difficulty with programming and reasoning with dependent types is that proof obligations arise naturally once programs become even moderately sized. For example, implementing an adder for binary numbers indexed over their natural number equivalents will naturally lead to proof obligations for equalities of expressions over universally quantified natural numbers. The need of these proofs of equalities comes, in intensional type theories, from the fact that the propositional equality enables to prove as equal terms that are not judgementally equal, which means that the typechecker can't always obtain equalities by reduction, and therefore often needs proofs before accepting definitions.
As far as possible, we would like to solve such proof obligations automatically. In this paper, we show one way to automate these proofs by reflection for the dependently-typed programming language Idris. We will show how type-safe reflection, where reflected terms are indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.

We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified natural numbers and the associative addition operation, and will later generalise these ideas for building automatic provers for various kind of properties (commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings. We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.
\keywords{proof automation, equivalence, equality, proof by reflection, correct-by-construction software}
\end{abstract}
%

\input{./content/1-introduction.tex}

\end{document}

### 2017-03-26 paper-CICM-17/content/1-introduction.tex
CICM paper : intro currently in progress

--- 2016-01-19 paper-JFP-DTP/content/1-introduction.tex
@@
 
-Proving that one term is equal to another within some theories is quite common
-in formal certification. The two terms are usually not syntactically 
-equal---this case would be trivial---but they are equal relative to some
-equivalence properties. These properties are either obtained by lemmas
-(the implementation of $+$ for natural numbers satisfies the property of
-associativity for example), or by axioms.
-
-For example, when working with natural numbers, one might need to prove the
-lemma $\forall\ x, y:Nat,\ x*x\ +\ 3*x\ +\ x*y\ =\ x*(3+y+x)$. Or,
-when working
-with lists and concatenation of lists, one might need to prove the lemma
-$\forall\ l1,\ l2,\ l3:List A,\ ((l1\ ++\ (l2 ++ l3))\ ++\ l4\ =\ (l1++l2)\ ++\
-(l3++l4)$. Proving this kind of auxiliary lemmas is time consuming,
-typically not reusable due to being very specific, and is rarely
-the interesting part of formal
-certification. Therefore, we would like to be able to automatize as far as
-possible the constructions of this kind of proofs. We have implemented
-the machinery to do so in Idris, a
-dependently typed language~\cite{brady2013idris}.
-
-When doing formal certification, we often require equality proofs between
-terms.
-For example, say we want to certify the function $f\ :\ A \rightarrow
-B$. We have two functions $sem_A\ :\ A \rightarrow SEM$ and $sem_B\ : \ B
-\rightarrow SEM$ which associate a semantics to each type A
-and B. If this semantics describes all of the important properties of
-the behaviour of $f$,
-then the lemma of correctness of $f$ can be expressed as $forall\ a:A,\ semA\ a =
-semB (f\ a)$. Very often, the type $SEM$ has some algebraic
-algebraic structure, and it becomes thus possible to
-decide equalities in it by following some standard rewriting rules.  
-All the equality symbols we have
-encountered so far refer to the propositional equality. In the next section,
-we will see how this is defined in Idris, and how it differs from the more
-primitive judgemental equality.
-
-\subsection{Propositional and Judgemental Equalities in Intentional Type Theories}
-In Mathematics, equality is a proposition, e.g. we can disprove an equality or 
-assume an equality as a hypothesis.
+Proofs assistants like Coq(ref), Agda(ref) and Idris(ref) are based on Intensional Type Theories, that contain two notions of equality.
+One is the equality that can be manipulated in the language, called the propositional equality.
+This equality corresponds to the equality found in mathematics : this is a proposition that can be assumed, negated, proved or disproved.
 Since in type theory, propositions are seen as types~\cite{How80}, the
 proposition that two elements $x$ and $y$ are equal corresponds to a type.
-Thus, if $x$ and $y$ are of type $a$, then the type
+If $x$ and $y$ are of type $a$, then the type
 $Id_a(x, y)$ represents the proposition ``$x$ is equal to $y$". If this type is
-inhabited, then $x$ is said to be provably equal to $y$.  Thus, $Id$ is a type
-family (parametrised by the type $a$) indexed over two elements of $a$, giving
-$Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$. For
-convenience, we write $(Id_a\ x\ y)$ as $(x\ =_a\ y)$. This equality is the
-equality which can be manipulated in the language.
+inhabited, then $x$ is said to be provably equal to $y$.
+Thus, $Id$ is a type family (parametrised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
+For
+convenience, we write $(Id_a\ x\ y)$ as $(x\ =_a\ y)$.
 
-There is another, more
-primitive, notion of equality in Intentional Type Theories, called judgemental
-equality, or definitional equality. This second equality means ``equal by
-definition'' The judgemental equality cannot be negated or
-assumed; we cannot talk about this primitive equality inside the theory.
+The second notion of equality in Intentional Type Theories is a primitive concept, called judgemental equality, or definitional equality.
+This second equality means ``equal by definition''.
+The judgemental equality cannot be negated or assumed; we cannot talk about this primitive equality inside the theory.
 Whether or not two expressions are equal by definition is a matter of
 evaluating the definitions. For example, if $f\ :\ \mathbb{N}\ \rightarrow\
-\mathbb{N}$ is defined by $f\ x\ \equiv\ x\ +\ 2$, then $f\ 5$ is
+\mathbb{N}$ is defined by $f\ x\ :=\ x\ +\ 2$, then $f\ 5$ is
 definitionally equal to $7$. Definitional equality entails unfolding
@@
 
-The judgemental equality has to be included in the propositional equal, because
-what is equal by definition should be provably equal.  This is accomplished by
-giving constructors to the type $Id(a,a)$ and nothing when ``$a$ is not $b$".
-One way to implement $Id$ in Idris is:
+The judgemental equality has to be included in the propositional equality, because what is equal by definition should be provably equal.
+This is accomplished by giving constructors to the type $Id(a,a)$ and nothing when ``$a$ is not $b$".
+In these theories, $Id$ is therefore implemented with the following sum type that has only one constructor :
 
 \begin{lstlisting}
-data Id : a -> a -> Type where
-     Refl : (x : a) -> Id x x
+data Id : a $\rightarrow$ a $\rightarrow$ Type where
+     Refl : (x : a) $\rightarrow$ Id x x
 \end{lstlisting}
@@
 The only way for $(Id_a\ x\ y)$ to be inhabited is therefore that $x$ should be
-(by definition!) $y$. In this case, the constructor $Refl$ helps to create a
+(by definition) $y$. In this case, the constructor $Refl$ helps to create a
 proof of this equality : $(Refl\ x)$ is precisely the proof which says that
@@
 
-When reading these definitions, one could think that the
-propositional equality is a simple wrapper of the judgemental equality. This is
-not the case: the propositional equality does not only contain the judgemental
-equality, because in these type theories, a principle of induction is
-associated with each inductive type. This induction principle says that if a
-proposition holds for the base cases, and if it can be showed that when it
-holds for some terms then it will also hold for the bigger terms obtained by
-using the recursive constructors, then this proposition will hold for any term
-of this type.
+When reading these definitions, one could wrongly think that the propositional equality is a simple wrapper of the judgemental equality.
+This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
+An inductive principle says that if a proposition holds for the base cases (ie, the constant constructors), and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the recursive constructors, then this proposition will hold for any term of this type.
+More formally, if a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$ in type-theory, defined in idris as :
 
-Induction principles work for any proposition, so it also works with the
-propositional equality. That means that the axiom of induction principle has
-made the type $Id(a,b)$ bigger than it looks like in the above definition, and
-there are therefore things which are provably equal which unfortunately aren't
-equal syntactically after an expanding of all the definitions. The activity of
-proving equalities is therefore in these theories something which isn't
-automatically decidable by the type-checker in the general case.
+\begin{lstlisting}
+data T : Type where
+  T0 : T
+  T1 : T $\rightarrow$ T          
+\end{lstlisting}
 
-For example, we can prove that $n+0 = n$ by induction, even if $n+0 \not\equiv
-n$ with the usual definition of $+$, recursive on its first argument.
+then we have : \\
+$T\_ind : \forall P:T \rightarrow Type,\ (P\ T0) \rightarrow (\forall t:T,\ P\ t \rightarrow P\ (T1\ t)) \rightarrow (\forall t:T,\ P\ t)$.
 
-These proofs are usually omitted ``on paper'' for everyday mathematics, but
-they are required by proof assistants, no matter how obvious they are to the
-human reader.
-Currently, in the Idris programming language, proofs must be completed
-by hand, and consist
-of a potentially long sequence of rewriting, where each step of
-rewriting uses one of the available properties. These proofs can be
-an everyday
-routine when working in a dependently typed language, and especially
-when intensively indexing types over values in order to capture certain logical
-properties.  For example, If a type $T$ is indexed over natural numbers, then
-the definition of a function which has codomain $T$ will often require a
-proof that the index of the produced term is equal to the expected index. This
-proof may use properties of natural numbers and their usual operations:
-symmetry and commutativity for +, existence of neutral element for +, etc.
+The principles of induction can be used for proving any proposition, so they also work for the propositional equality.
+For example, we can prove that $n+0 = n$ for all $n$ by induction on the $Nat\ n$, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
+That means that the axiom of induction has made the type $Id_{a}(x,y)$ bigger than it looks like : it contains terms that are not generated by $Refl$ (the only constructor of $Id_{a}(x,y)$), but that are added by the inductive principles.
+There are therefore things which are provably equal which unfortunately are not definitionally equal after unfolding all the definitions and after full reductions.
+The activity of proving equalities is therefore in these theories something which isn't automatically decidable by the type-checker in the very general case.
+
+Proving that one term is equal to another is quite common
+in formal certification.
+For example, when working with natural numbers, one might need to prove the lemma $\forall\ x, y:Nat,\ x*x + 3*x + x*y\ =\ x*(3+y+x)$.
+Or, when working with lists and concatenation of lists, one might need to prove the lemma $\forall\ l1,\ l2,\ l3:List A,\ (l1 ++ (l2 ++ l3)) ++ l4\ =\ (l1++l2) ++ (l3++l4)$.
+Currently, in the Idris programming language, these proofs must be completed by hand, and consist of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
+Proving this kind of auxiliary lemmas is time consuming and typically not reusable due to being very specific.
+These proofs are usually omitted ``on paper'' for everyday mathematics, but they are required by proof assistants, no matter how obvious they are to the human reader.
+These proofs can be an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
 

### 2017-03-26 paper-CICM-17/content/1-introduction.tex
CICM paper : intro continued

@@
 
-We revisit an example from previous work~\cite{bradytfp07} 
-which shows how proof obligations might appear
-when we index a type by natural numbers.  We define the types of
-$Bit$ and $Binary$ number in Figure~\ref{binarynums}, 
-both indexed over the value they represent, expressed as a natural number.
+We revisit an example from a previous work of Edwin Brady~\cite{bradytfp07} which shows how proof obligations may arise when a type is indexed by natural numbers.
+In this development, the goal is to implement a formally verified library of binary numbers.
+In order to ensure the correctness of the methods that will be developed, we define the types $Bit$ and $Binary$ as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.
 
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Bit and binary number}
-\label{binarynums}
-\figrule
-\end{figure}
 
-It allows representing a
-binary number of width zero with the constructor $zero$, 
-and representing larger numbers with $\#$.
-We
-can extend a binary number $bin$ by adding a digit to its right, and the value
-represented by this new binary number is two times the value ($v$) represented
-by $bin$ plus the value ($bit$) represented by the added digit.
+We want to write a function performing the addition between two binary numbers. In order to do so, we start with an auxiliary function, which adds three bits
+(the third one plays the role of a carry flag), and produces the two bits of
+the result, where the first one is the most significant bit.
 
-To write a function performing the addition between two binary numbers,
-we start with an auxiliary function (Figure~\ref{addbit}) which adds three bits
-(the third one plays the role of a carry flag), and produces the two bits of
-the result.
-
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Addition of three bits}
-\label{addbit}
-\figrule
-\end{figure}
 
-This function produces a dependent pair. The first component is the value $bX$
-(expressed as a natural number) represented by the first bit produced (of type
-$Bit\ bX$), and the second component is another dependent pair. This
-second dependent pair is composed of the value $bY$ represented by the second
-bit produced, and of a triple containing the two bits produced and a
-proof that these two bits correctly encode the result of the addition.
-For example, with the line corresponding to the computation
-$1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$)
-and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
+This function produces a dependent pair.
+The first component is the value $bX$ (expressed as a natural number) representing the first bit produced (therefore of type $Bit\ bX$), and the second component is another dependent pair.
+This second dependent pair is composed of the value $bY$ representing the second bit produced, and of a triple containing the two bits produced and a proof that these two bits correctly encode the result of the addition.
+For example, on the fifth line, which corresponds to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
 
-It should be pointed that the proof component of this function is established without any effort, as we always give ``Refl" for the proof.
-Indeed, for each of the eight patterns, a simple computation reduces the given and expected values to the same value, and ``Refl" can therefore be used.
-We also note that
-there is no need to produce any lemma of correctness about $addBit$ afterwards.
-The correct by construction style in which it is written already gives the
-property that we want: the two bits produced effectively represent the
-addition of the three bits given in input.
-
-We finally define the function adding two binary numbers and a carry flag.
-This addition works for two binary numbers
-expressed with the same number of bits, and it produces a result with one more
-bit. This result represents the value $c + x + y$, where $c$ 
-represents the value of the input carry bit, and $x$ and $y$ represent
-the values of the two binary numbers given in input. Thus, we'd like to
-be able to write:
+We finally define the function $adc$ that adds two binary numbers and a carry flag.
+This addition works for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
+This result represents the value $c + x + y$, where $c$ represents the value of the input carry bit, and $x$ and $y$ represent the values of the two binary numbers given in input.
+Therefore, we would like to be able to write :
 
 \begin{lstlisting}
-adc : Binary w x $\rightarrow$ Binary w y $\rightarrow$ Bit c $\rightarrow$ Binary (S w) (c + x + y)
+adc : Binary w x $\rightarrow$ Binary w y $\rightarrow$ Bit c 
+      $\rightarrow$ Binary (S w) (c + x + y)
 adc zero zero carry = zero # carry
 adc (numx # bX) (numy # bY) carry
-   = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) = addBit bX bY carry in
+   = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) 
+          = addBit bX bY carry in
           adc numx numy carry0 # lsb
@@
 
-Unfortunately, the types do not match for either of the two cases. 
-The result of the first line $adc\ zero\ zero\ carry$ is
-expected to have the type\[Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)\] but we
-provide a term of type \[Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0))).\]
-The problem is even worse for the second case where the expected index is 
-\[(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))\] while we're trying to provide a term indexed over 
-\[plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))\]
+The first pattern performs the addition between the values zero, zero and a carry.
+In this simple case, we simply have to produce the one-digit value $zero\ \#\ carry$.
+The second pattern is when the two binary numbers are respectively $(numx\ \#\ bX)$ and $(numy\ \#\ bY)$ where $numx$ and $numy$ are smaller binary numbers and $bX$ and $bY$ are bits added to the right of $numx$ and $numy$.
+In this case, we start by adding the least significant bits $bX$ and $bY$ with the $carry$, and this is performed by calling the function $addBit$.
+It will essentially produce 2 bits, the first one representing the carry of the result (called $carry0$), and the second one the least-significant bit of the result (called $lsb$).
+Once we have this intermediate result, we can call recursively $adc$ on the smaller terms $numx$ and $numy$ with $carry0$ as the carry that we need to propagate, and finally add $lsb$ to the right of this result.
 
-For
-this situation, Idris provides \emph{provisional definitions} with the 
-syntax \texttt{?=}. Using a provisional definition here will make
-the definition acceptable, but it also generates two proof obligations, one
-for each pattern, both of which say that the provided term can be
-converted into a term which has the expected type.  For both of these
-patterns, it is possible to prove the equality between the expected and
-the provided types. These proofs
-consist of a series of rewriting steps, where each step uses a property about the
-operation $+$ on natural numbers.  For example, the proof for the second
-pattern is shown in Figure \ref{adc_lemma_2} (with some detail elided):
+Unfortunately, this straightforward definition is rejected by Idris' type-checker because the types mismatch for both of the two patterns.
+The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type $Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)$ but we provide a term of type $Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))$.
+The problem is even worse for the second case where the expected index is \\
+$(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))$
+while we're trying to provide a term indexed over \\
+$plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))$.
+We know that the definition of $adc$ we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and idris can not figure out a proof of compatibility on its own.
+If we use provisional definitions[ref], the previous definition can be accepted by the system, but we now have two proof obligations $adc\_lemma\_1$ and $adc\_lemma\_2$ demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
 
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting}
@@
     trivial;
+}
 \end{lstlisting}
-\end{center}
-\caption{Proof by hand of the correspondence between expected and provided types for the second pattern of the definition of adc}
-\label{adc_lemma_2}
-\figrule
-\end{figure}
 
-Thus, this kind of proof consists of a potentially long sequence of rewriting
-steps,
-where each step uses one of the available properties. Without some
-specific automation, this sequence of rewriting must be done by the programmer.
-This is time consuming, and perhaps more importantly, even a small
-change in the left
-or the right hand side of the equality may invalidate the proof.
-This mean that a minor change in our
-datatypes, or in the definition of $addBit$ or $adc$ will require us to
-replace this proof by another one doing the same job.
-This kind
-of proof is therefore not reusable, because it performs rewritings for
-very specific terms.
+As we see, this kind of proof consists of a potentially long sequence of rewriting steps, where each step uses one of the available properties: neutral element, commutativity, associativity.
+Without some specific automation, this sequence of rewriting must be done by the programmer.
+This is time consuming and more importantly, just a small change in the left or the right hand side of the equality may invalidate the proof.
+That means that a minor change in our datatypes, or in the definition of $addBit$ or $adc$ will require us to replace this proof by another one fulfilling the same task.
+This kind of proof is therefore absolutely not reusable, because it performs rewritings for very specific terms.
+These proofs become the everyday routine in any dependently-typed language, as soon as we index types over values in order to capture some logical properties.
+As we've seen, when a type $T$ is indexed over $Nat$, the definition of a function which has codomain $T$ is likely to require a proof of equality between the produced index and the expected one.
 
-If we look at the proof we have written by hand, we realize that we only use
-the existence of neutral elements (this is the lemma $plusZeroRightNeutral$),
-and the associativity and commutativity of $+$. Thus, we're rewriting a term by
-using the properties of a commutative monoid.
-
+The proof that we have written by hand only uses the existence of neutral elements, and the associativity and commutativity of $+$ on $Nat$.
+Thus, we're rewriting a term by using the properties of a commutative monoid.
 We would like to let a generic prover for commutative monoid find a proof on its own.
-This should be achievable, because algebraic structures like monoids, groups and
-rings (commutative or otherwise) have a very useful property, that
-every expression can be normalised to a canonical representation. Thus
-these kind of proofs can be automatically produced by computing the
-normal forms for each side of the equality, and then comparing them using
-syntactical equality. When they are in normal
-form, being equal is a matter of being syntactically the same entity.
 
@@
 In this paper, we describe a certified
-implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris}} of an automatic prover for equalities in a
-\emph{hierarchy} of algebraic
+implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
 structures, including Monoid, Groups and Rings (all potentially commutative),
@@
 The principal novelty is in the approach that we follow, using a
-new kind of \emph{type-safe reflection}.  Working by reflection for implementing
-tactics has been done several times, including for the implementation of a
-ring solver for the Coq proof assistant, but without
+new kind of \emph{type-safe reflection}.
+Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without
 providing any guarantees, contrary to our type-safe reflection. We will
-compare our approach with other implementations in Section~\ref{sect:discuss}.
+compare our approach with other implementations in Section~\ref{sect:AlternativeApproaches}.
 
-We make the following contributions:
+The contributions of this paper are the following :
 
-\begin{itemize}
-        \item We present a new type-safe reflection mechanism, where the
-        reflected terms are indexed over the concrete inputs, thus providing a
-        direct way to pull out the proofs, and providing two guarantees.
-        The first guarantee is that the reflection of a term is guaranteed to be a
-        faithful representation of the term. The second is that any generated
-        proof is guaranteed to be a proof of the required property. The basic
-        ideas of the technique are first presented in Section~\ref{sect:ideas} on a
-        smaller problem with only natural numbers and addition, and are later
-        adapted for the hierarchy of tactics in Section~\ref{sect:general}. 
-        Section~\ref{sect:reuse} contains some implementation details.
+\begin{enumerate}
+
+\item We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing a
+direct way to pull out the proofs, and providing the guarantee that the reflection of a term is indeed a faithful representation of the term.
+The basic ideas of the technique are first presented in Section~\ref{sect:aSimpleProofAutomation} on a smaller problem with only natural numbers and addition, and are later adapted for a hierarchy of tactics proving equalities in algebraic strcures, in Section~\ref{sect:ProvingEquivalencesInAlgebraicStructures}.
 
@@
 correct by construction approach, instead of implementing a
-        normalization procedure, and proving afterwards that this function
-        is correct. This approach is much more suitable
+normalisation procedure, and proving afterwards that this function
+is correct with external lemmas. This approach is much more suitable
 for programming languages like Idris. Again, this approach will be
-        presented in Section~\ref{sect:ideas} on the smaller problem, 
-        and in Sections~\ref{sect:general} and~\ref{sect:reuse}
-        for the complete hierarchy of algebraic structures.
+presented in Section~\ref{sect:aSimpleProofAutomation} on the smaller problem, 
+and in Sections~\ref{sect:ProvingEquivalencesInAlgebraicStructures} for the complete hierarchy of algebraic structures.
 
 \item We develop a hierarchy of tactics where each tactic reuses the
-        rewriting machinery from the structure immediately below in the hierarchy.
-        For example,
-        simplifying neutral elements is only implemented at the monoid level,
-        and each level inheriting from it will reuse it. Re-usability can be quite tricky
-        to obtain when we want to reuse a prover from a structure which is less
-        expressive. For example, reusing the monoid prover for building the
+rewriting machinery of the structure from which it naturally inherits.
+For example, simplifying neutral elements is only implemented at the monoid level,
+and each level inheriting from it will reuse it.
+Re-usability is difficult to obtain when we want to reuse the prover of a less expressive structure.
+For example, reusing the monoid prover for building the
 group prover is not trivial, since we lose the possibility to express
 negations ($-x$) and subtractions ($x-y$). Some specific encodings
-        will be presented in Section~\ref{sect:reuse} for solving this problem.
+will be presented in Section~\ref{sect:NormalisationsFunctionsAndReusabilityOfTheProvers} for solving this problem.
 
-\end{itemize}
+\end{enumerate}

### 2017-03-26 paper-CICM-17/paper.tex
CICM paper : intro continued

@@
 
-We will first present the technique on a small example, where we aim to automatically prove equalities with universally quantified natural numbers and the associative addition operation, and will later generalise these ideas for building automatic provers for various kind of properties (commutativity, distributivity, existence of neutral elements...), leading to a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
+After a presentation of our key ideas on a small example, we will build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
 We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.

### 2017-03-27 paper-CICM-17/content/1-introduction.tex
added basic tactic for Nat and + assoc

@@
 This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-An inductive principle says that if a proposition holds for the base cases (ie, the constant constructors), and if it can be showed that when it holds for some terms then it will also hold for the bigger terms obtained by using the recursive constructors, then this proposition will hold for any term of this type.
-More formally, if a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$ in type-theory, defined in idris as :
+If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in idris as :
 
@@
 
-The principles of induction can be used for proving any proposition, so they also work for the propositional equality.
+The principles of induction can be used for proving any proposition $P$, so they also work for the propositional equality.
 For example, we can prove that $n+0 = n$ for all $n$ by induction on the $Nat\ n$, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
@@
 in formal certification.
-For example, when working with natural numbers, one might need to prove the lemma $\forall\ x, y:Nat,\ x*x + 3*x + x*y\ =\ x*(3+y+x)$.
-Or, when working with lists and concatenation of lists, one might need to prove the lemma $\forall\ l1,\ l2,\ l3:List A,\ (l1 ++ (l2 ++ l3)) ++ l4\ =\ (l1++l2) ++ (l3++l4)$.
-Currently, in the Idris programming language, these proofs must be completed by hand, and consist of a potentially long sequence of rewriting, where each step of rewriting uses one of the available properties.
-Proving this kind of auxiliary lemmas is time consuming and typically not reusable due to being very specific.
-These proofs are usually omitted ``on paper'' for everyday mathematics, but they are required by proof assistants, no matter how obvious they are to the human reader.
-These proofs can be an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
+This is an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
 
@@
 This addition works for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
-This result represents the value $c + x + y$, where $c$ represents the value of the input carry bit, and $x$ and $y$ represent the values of the two binary numbers given in input.
-Therefore, we would like to be able to write :
+This result represents the sum of the three concrete values of the inputs.
+Therefore, we want to write :
 
 \begin{lstlisting}
-adc : Binary w x $\rightarrow$ Binary w y $\rightarrow$ Bit c 
-      $\rightarrow$ Binary (S w) (c + x + y)
+adc : Binary w x $\rightarrow$ Binary w y $\rightarrow$ Bit c $\rightarrow$ Binary (S w) (c + x + y)
 adc zero zero carry = zero # carry
 adc (numx # bX) (numy # bY) carry
-   = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) 
-          = addBit bX bY carry in
+  = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) = addBit bX bY carry in
           adc numx numy carry0 # lsb
 \end{lstlisting}
-
-The first pattern performs the addition between the values zero, zero and a carry.
-In this simple case, we simply have to produce the one-digit value $zero\ \#\ carry$.
-The second pattern is when the two binary numbers are respectively $(numx\ \#\ bX)$ and $(numy\ \#\ bY)$ where $numx$ and $numy$ are smaller binary numbers and $bX$ and $bY$ are bits added to the right of $numx$ and $numy$.
-In this case, we start by adding the least significant bits $bX$ and $bY$ with the $carry$, and this is performed by calling the function $addBit$.
-It will essentially produce 2 bits, the first one representing the carry of the result (called $carry0$), and the second one the least-significant bit of the result (called $lsb$).
-Once we have this intermediate result, we can call recursively $adc$ on the smaller terms $numx$ and $numy$ with $carry0$ as the carry that we need to propagate, and finally add $lsb$ to the right of this result.
 
@@
 $plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))$.
-We know that the definition of $adc$ we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and idris can not figure out a proof of compatibility on its own.
+We know that the definition of $adc$ we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and Idris can not figure out a proof of compatibility on its own.
 If we use provisional definitions[ref], the previous definition can be accepted by the system, but we now have two proof obligations $adc\_lemma\_1$ and $adc\_lemma\_2$ demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
@@
 That means that a minor change in our datatypes, or in the definition of $addBit$ or $adc$ will require us to replace this proof by another one fulfilling the same task.
-This kind of proof is therefore absolutely not reusable, because it performs rewritings for very specific terms.
 These proofs become the everyday routine in any dependently-typed language, as soon as we index types over values in order to capture some logical properties.
-As we've seen, when a type $T$ is indexed over $Nat$, the definition of a function which has codomain $T$ is likely to require a proof of equality between the produced index and the expected one.
 
@@
 new kind of \emph{type-safe reflection}.
-Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without
-providing any guarantees, contrary to our type-safe reflection. We will
+Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without the type-safe aspect.
+We will
 compare our approach with other implementations in Section~\ref{sect:AlternativeApproaches}.

### 2017-03-27 paper-CICM-17/content/2-basicTactic.tex
added basic tactic for Nat and + assoc

\section{A simple proof automation for universally quantified natural numbers and their associative addition}
\label{sect:aSimpleProofAutomation}

We will first present the basic ideas of our technique on a simplified problem, in which we aim to deal with universally quantified natural numbers and additions. We would like to be able to automatically generate proofs of this kind of goals :
\\
\textbf{Example 1.} $\forall (x\ y\ z:Nat),\ (x + y) + (x + z) = x + ((y + x) + z)$  \\
\
In this section, we will only work with the properties of associativity and of neutral element for the $+$ operation on $Nat$.

\subsection{Type-safe reflection}

When trying to prove this kind of equalities, the variables are abstracted, and they become part of the context. The general idea --that will also apply for the more general problem detailed in the next section-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them with a syntactic equality test. The problem is that such a normalisation function will need to do different treatments for a ``variable natural number" (ie, a number which has been universally quantified) and for the constant $Z$. For this reason, we will need to work by reflection, with a datatype encoding natural numbers composed of ``variable numbers", $Z$, and additions of these things. The novelty is that we use a a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent :

\begin{lstlisting}
using (x : Nat, y : Nat, $\Gamma$ : Vect n Nat)
  data Expr : (Vect n Nat) $\rightarrow$ Nat $\rightarrow$ Type where
       Plus : Expr $\Gamma$ x $\rightarrow$ Expr $\Gamma$ y $\rightarrow$ Expr $\Gamma$ (x + y)
       Var  : (i : Fin n) $\rightarrow$ Expr $\Gamma$ (index i $\Gamma$)
       Zero : Expr $\Gamma$ Z
\end{lstlisting}

For an expression $e_x\ :\ Expr\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
When an expression is a variable $Var\ i$, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
Also, the $Zero$ expression denotes the natural number $Z$.
Finally, and fairly trivially, if $e_x$ is an expression encoding the number $x$, and $e_y$ is an expression encoding the number $y$, then the expression $Plus\ e_x\ e_y$ denotes the concrete natural number $(x + y)$.

Variables are represented using a De Brujin-like index: $Var\:FZ$ denotes the first variable abstracted, $Var\:(FS\:FZ)$ the second one, and so on.

Let's write the encoding of the left hand-side of the example 1 :
\begin{lstlisting}
e1 : (x, y, z : Nat) $\rightarrow$ Expr [x, y, z]
e1 x y z = Plus (Plus (Var FZ) (Var (FS FZ))) 
                (Plus (Var FZ) (Var (FS (FS FZ))))
\end{lstlisting}

	\subsection{A correct by construction approach}
\label{sect:aCorrectByConstructionApproach}

We want to write the reduction function on a \emph{correct by construction} way, which means that no additional proof should be required after the definition of the function. Thus, $reduce$ will produce the proof that the new Expr freshly produced has the same interpretation as the original Expr, and this will be made easier by the fact that the datatype Expr is indexed over the real --concrete-- natural number : a term of type $Expr\ \Gamma\ x$ is the encoding of the number $x$.
Thus, we can write the type of $reduce$ like this : \\
$reduce\ :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ =\ x'))$ \\
The function $reduce$ produces a dependent pair : the new concrete number $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new --concrete-- natural numbers are equal.

This function doesn't simply produce an $Expr\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x=x'$ is an essential component for building the desired proof of $x=y$, which is the main problem that we are trying to solve.
\\
More precisely, the tactic will work as follow.
We have an expression $e_x$ encoding $x$, and an expression $e_y$ encoding $y$. We will normalize $e_x$ and this will give a new concrete number $x'$, a new expression $e_{x'}:Expr\ \Gamma\ x'$, and a proof of $x=x'$. We will do the same with $e_y$ and we will get a new concrete number $y'$, an expression $e_{y'}:Expr\ \Gamma\ y'$, and a proof of $y=y'$. \\
It is now enough to simply compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test because these two expressions are now supposed to be in normal form.

\begin{lstlisting}
eqExpr : (e : Expr $\Gamma$ x) $\rightarrow$ (e' : Expr $\Gamma$ y) $\rightarrow$ Maybe (e = e')
\end{lstlisting}

Now, if the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$.
By rewriting the two equalities $x=x'$ and $y=y'$ (that we obtained during the normalisations) in the new equality $x'=y'$, we can get a proof of $x=y$. This is what the function $buildProof$ is doing.

\begin{lstlisting}
buildProof : {x : Nat} $\rightarrow$ {y : Nat} $\rightarrow$ Expr $\Gamma$ x' $\rightarrow$ Expr $\Gamma$ y' 
           $\rightarrow$ (x = x') $\rightarrow$ (y = y') $\rightarrow$ Maybe (x = y)
buildProof ex' ey' lp rp with (eqExpr ex' ey')
  buildProof ex' ex' lp rp | Just Refl = ?MbuildProof
  buildProof ex' ey' lp rp | Nothing = Nothing
\end{lstlisting}

The argument of type $Expr\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$. Before the normalisation, the reflected LHS was reflecting the value $x$. The same applies to the right-hand side. The function also expects a proof of $x=x'$ and of $y=y'$, and we will be able to provide them because the normalisation function also produces the proof of equality between the original and the new concrete values.

As mentioned, the proof-term that fills the hole represented by the metavariable $MbuildProof$ is just a rewriting of the two equalities that we have :

\begin{lstlisting}
MbuildProof = proof {
  intros; refine Just; rewrite sym p1; rewrite sym p2; exact Refl;
}  
\end{lstlisting}

Finally, the main function which tries to prove the equality $x=y$ simply has to reduce the two reflected terms encoding the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs that have been obtained :

\begin{lstlisting}
  testEq : Expr $\Gamma$ x $\rightarrow$ Expr $\Gamma$ y $\rightarrow$ Maybe (x = y)
  testEq ex ey = 
     let (x' ** (ex', px)) = reduce ex in 
     let (y' ** (ey', py)) = reduce ey in
        buildProof ex' ey' px py 
\end{lstlisting}

The remaining part is to define the function $reduce$. We decide that the left associative form will be the canonical representation. We define a new datatype which captures the property of being left-associative by its shape :

\begin{lstlisting}
data LExpr : ($\Gamma$ : Vect n Nat) $\rightarrow$ Nat $\rightarrow$ Type where
     LPlus : LExpr $\Gamma$ x $\rightarrow$ (i : Fin n) 
             $\rightarrow$ LExpr $\Gamma$ (x + index i $\Gamma$)
     LZero : LExpr $\Gamma$ Z
\end{lstlisting}

This datatype has only two constructors. In fact, it combines the previous $Var$ and $Plus$ constructors so that it becomes impossible to write an expression which isn't left associative, because the constructor LPlus is only recursive on its first argument.

As part of the normalization, we write a function $expr\_l$ which converts an $Expr\ \Gamma\ x$ to a $LExpr\ \Gamma\ x'$ and which produces a proof of $x=x'$. This function will therefore use the two available properties multiple times, while rewriting the term until the fully left associative desired form is obtained.

\begin{lstlisting}
expr_l : Expr $\Gamma$ x $\rightarrow$ (x' ** (LExpr $\Gamma$ x', x = x'))
expr_l Zero = (_ ** (LZero, Refl))
expr_l (Var i) = (_ ** (LPlus LZero i, Refl))
expr_l (Plus ex ey) = 
  let (x' ** (ex', px)) = expr_l ex in
  let (y' ** (ey', py)) = expr_l ey in
  let (res ** (normRes, Pres)) = plusLExpr ex' ey' in
    (res ** (normRes, rewrite px in (rewrite py in Pres))) where 
      plusLExpr : {$\Gamma$ : Vect n Nat} $\rightarrow$ {x, y : Nat} $\rightarrow$ LExpr $\Gamma$ x 
                  $\rightarrow$ LExpr $\Gamma$ y  $\rightarrow$ (z ** (LExpr $\Gamma$ z, x+y=z))
      plusLExpr {x=x} ex LZero =
        (_ ** (ex, rewrite (plusZeroRightNeutral x) in Refl))            
      plusLExpr ex (LPlus e i) =
        let (xRec ** (rec, prfRec)) = plusLExpr ex e in
            (_ ** (LPlus rec i, ?MplusLExpr))

\end{lstlisting}

In the case of an addition $Plus\ ex\ ey$, the function $expr\_l$ does the job of normalisation recursively on $ex$ and on $ey$, and then it uses the sub-function $plusLExpr$ to normalise the addition of these two --already normalised-- terms. This sub-function $plusLExpr$ has two kind of simplifications to do. When the second argument is an $LZero$, it simply returns its first arguments along with the justification for this rewriting, which obviously uses $plusZeroRightNeutral$. However, when the second argument is an $LPlus\ e\ i$, it continues recursively by computing $plusLExpr\ ex\ e$, and it finally adds $i$ to it. That had for effect to move the parenthesis on the left, and the correctness of this treatment is going to be justified by the use of $plusAssociative$ in the proof that corresponds to the meta variable $MplusLExpr$ .

This metavariable $MplusLExpr$ --that expressed the equality between the old and the new index-- requires us to prove the goal : $x1 + (x2 + index\ i\ \Gamma) = xrec + index\ i\ \Gamma$ in a context where we've got, amongst other things, $prfRec : x1 + x2 = xrec$.
By using the property of associativity on the goal, we now need to prove $(x1 + x2) + index\ i\ \Gamma$, which can be done by rewriting the proof $prfRec$ obtained recursively.

\begin{lstlisting}
MplusLExpr = proof {
  intros
  rewrite (sym (plusAssociative x1 x2 (index i $\Gamma$))); 
  rewrite prfRec; 
  exact Refl;
}
\end{lstlisting}

It is really important to understand that the kernel of the automatic construction of the desired proof of $x=y$ happens precisely in these nested use of $plusZeroRightNeutral$ and $plusAssociative$ that we've seen in the definition of $expr\_l$ and of its meta-variable $MplusLExpr$. These proofs replace the arithmetical proofs that we were doing previously by hand, as it was the case in section~\ref{sect:HowProofObligationsArise} with the lemma $adc\_lemma\_2$.

The normalisation function is just the composition of the functions $expr\_l$ and $l\_expr$, where the later is the conversion back to the type $Expr$ :

\begin{lstlisting}
  reduce : Expr $\Gamma$ x $\rightarrow$ (x' ** (Expr $\Gamma$ x', x = x'))
  reduce e = let (x' ** (e', prf)) = expr_l e 
                 in (x' ** (l_expr e', prf))
\end{lstlisting}

At the moment, what we've got is not exactly a real tactic, in the sense that we only have a function which produces a value of type $Maybe (x = y)$. A real tactic would be a wrapper of this function that would fail properly with an error message when the two terms are not equal. However, here, when $x\ne y$, the function $testEq$ will simply produce the value $Nothing$. \\

	\subsection{Usage of the ``tactic"}

It's now time to see how to use this minimalist ``tactic".
Let's go back to the example 1. We had defined the expression $e1$ representing the value $((x + y) + (x + z))$ and we now have to produce the encoding for $(x + ((y + x) + z))$, still in the context $[x, y, z]$ of three abstracted variables.

\begin{lstlisting}
e2 : (x, y, z : Nat) 
     $\rightarrow$ Expr [x, y, z] (x + ((y + x) + z))
e2 x y z = Plus (Var FZ) 
                (Plus (Plus (Var (FS FZ)) 
                            (Var FZ)) 
                      (Var (FS (FS FZ))))
\end{lstlisting}

The numbers denoted by the expressions $e1$ and $e2$ are equal, and we can generate a proof of this fact by using $testEq$.

\begin{lstlisting}
e1_e2_testEq : (x, y, z : Nat) 
      $\rightarrow$ Maybe (((x + y) + (x + z)) = (x + ((y + x) + z)))
e1_e2_testEq x y z = testEq (e1 x y z) (e2 x y z)
\end{lstlisting}

And if we ask for the evaluation of this term, we obtain $Just$ and a proof of equality between the two underlying concrete values :

\begin{lstlisting}
#\x => \y => \z => e1_e2_testEq x y z

\x => \y => \z => Just (replace (sym (replace (sym (replace 
(sym (plusAssociative x 0 y)) (replace (replace (sym 
(plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
(replace (sym (plusAssociative x 0 z)) (replace (replace 
(sym (plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
(plusAssociative (x+y) x z)) [...]
: (x : Nat) $\rightarrow$ (y : Nat) $\rightarrow$ (z : Nat) 
  $\rightarrow$ Maybe ((x + y) + (x + z) 
            = x + ((y + x) + z))
\end{lstlisting}

As expected, this proof uses the properties of associativity ($plusAssociative$) and the property of neutrality of $Z$ for $+$ ($plusZeroRightNeutral$).

### 2017-03-28 paper-CICM-17/content/2-basicTactic.tex
Dammit, not enough space, will have to remove section 2 on the smaller tactic

@@
 The general idea --that will also apply for the more general problem detailed in the next section-- will be to normalize both sides of the "potential equality" $x=y$, and afterwards to compare them with a syntactic equality test.
-The problem is that such a normalisation function will need to do different treatments for a ``variable natural number" (ie, a number which has been universally quantified) and for the constant $Z$.
+The problem is that such a normalisation function will need to do different treatments for a ``variable natural number" (ie, a number which has been universally quantified), for the constant $Z$ and for a sum.
 For this reason, we will need to work by reflection, with a datatype encoding natural numbers composed of ``variable numbers", $Z$, and additions of these things.
@@
 Also, the $Zero$ expression denotes the natural number $Z$.
-Finally, and fairly trivially, if $e_x$ is an expression encoding the number $x$, and $e_y$ is an expression encoding the number $y$, then the expression $Plus\ e_x\ e_y$ denotes the concrete natural number $(x + y)$.
+Finally, if $e_x$ is an expression encoding the number $x$, and $e_y$ is an expression encoding the number $y$, then the expression $Plus\ e_x\ e_y$ denotes the concrete natural number $(x + y)$.
 
-Variables are represented using a De Brujin-like index: $Var\:FZ$ denotes the first variable abstracted, $Var\:(FS\:FZ)$ the second one, and so on.
+Note that variables are represented using a De Brujin-like index: $Var\:FZ$ denotes the first variable abstracted, $Var\:(FS\:FZ)$ the second one, and so on.
 
-Let's write the encoding of the left hand-side of the example 1 :
+We can define the reflection of the left-hand side of the example 1 like this :
 \begin{lstlisting}
-e1 : (x, y, z : Nat) $\rightarrow$ Expr [x, y, z]
+e1 : (x, y, z : Nat) $\rightarrow$ Expr [x, y, z] ((x + y) + (x + z))
 e1 x y z = Plus (Plus (Var FZ) (Var (FS FZ))) 
@@
 \end{lstlisting}
+
+The two important things to remember with this type-safe reflection is that the reflection of a term is guaranteed to be a faithful encoding of it because of the index, and that this index will help us to build the construction of the desired proof of equality, as we will show in the next subsection.
 
@@
 This function doesn't simply produce an $Expr\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x=x'$ is an essential component for building the desired proof of $x=y$, which is the main problem that we are trying to solve.
-\\
 More precisely, the tactic will work as follow.
@@
 The same applies to the right-hand side.
-The function also expects a proof of $x=x'$ and of $y=y'$, and we will be able to provide them because the normalisation function also produces the proof of equality between the original and the new concrete values.
+The function also expects a proof of $x=x'$ and of $y=y'$, that we will be able to provide because the normalisation function has produced them.
 
@@
 We decide that the left associative form will be the canonical representation.
-We define a new datatype which captures the property of being left-associative by its shape :
+To help for this task, we define a new datatype which captures the property of being left-associative by its shape :
 
@@
 
-It is really important to understand that the kernel of the automatic construction of the desired proof of $x=y$ happens precisely in these nested use of $plusZeroRightNeutral$ and $plusAssociative$ that we've seen in the definition of $expr\_l$ and of its meta-variable $MplusLExpr$.
-These proofs replace the arithmetical proofs that we were doing previously by hand, as it was the case in section~\ref{sect:HowProofObligationsArise} with the lemma $adc\_lemma\_2$.
+It is really important to understand that the heart of the automatic construction of the desired proof of $x=y$ happens precisely in the repeated use of $plusZeroRightNeutral$ and $plusAssociative$ that we've seen in the definition of the fixpoint computed by $expr\_l$ and of its meta-variable $MplusLExpr$.
+The complete proof is produced during the recursive calls to $expr\_l$ that will compose the use of $plusAssociative$ for producing automatically a proof that can replace the arithmetical proofs that we were doing previously by hand with the lemma $adc\_lemma\_2$.
 
@@
 
-\x => \y => \z => Just (replace (sym (replace (sym (replace 
-(sym (plusAssociative x 0 y)) (replace (replace (sym 
-(plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
-(replace (sym (plusAssociative x 0 z)) (replace (replace 
-(sym (plusZeroRightNeutral x)) Refl) Refl))) (replace (sym 
-(plusAssociative (x+y) x z)) [...]
+\x => \y => \z => Just (replace (sym (replace (sym (replace (sym 
+(plusAssociative x 0 y)) (replace (replace (sym (plusZeroRightNeutral 
+x)) Refl) Refl))) (replace (sym (replace (sym (plusAssociative x 0 z)) 
+(replace (replace (sym (plusZeroRightNeutral x)) Refl) Refl))) 
+(replace (sym (plusAssociative (x+y) x z)) [...]
 : (x : Nat) $\rightarrow$ (y : Nat) $\rightarrow$ (z : Nat) 
-  $\rightarrow$ Maybe ((x + y) + (x + z) 
-            = x + ((y + x) + z))
+  $\rightarrow$ Maybe ((x + y) + (x + z) = x + ((y + x) + z))
 \end{lstlisting}

### 2017-03-28 paper-CICM-17/content/3-hierarchyOfTactics.tex
Dammit, not enough space, will have to remove section 2 on the smaller tactic

\section{A hierarchy of tactics}

In this section, we want to prove equalities not only for $Nat$ but for various datatypes, and for various properties. With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic prover that works for any type that is an instance of an algebraic structure. In fact, that will be a hierarchy of provers --with one prover for each algebraic structure-- that can prove equalities of terms in any instance of the corresponding structure. We will no longer only work with the properties of associativity and of right neutral element for natural numbers as it was the case in the previous section, but we will have available the properties of a given algebraic structure. These tactics will be usable on any type that satisfies these properties. The properties of a given algebraic algebraic structure will be expressed in a corresponding interface. This interface will extend the interface from which it naturally inherits. For example, the $Group$ interface will extend the $Monoid$ one. We will end up with a hierarchy of interfaces and will write one tactic for each of these interface. All the tactics that we will implement (the ring prover, the group prover, etc) will be able to work on any type, being given that an implementation of the corresponding interface is provided.

	\subsection{Proving equivalences instead of equalities}

With some reasonable additional effort, we could produce a collection of tactics for proving equivalences instead of only equality. The machinery would be very similar, and if we do the right abstraction from the start, we could gain one more degree of genericity, with the freedom of choosing the equivalence relation (which can of course be the usual equality). We want to give to the user the possibility to define his own equivalence notion, as long as he's able to provide the properties of the algebraic structure he wants to use. Let's call $c$ the \emph{carrier} type, ie, the type on which we want to prove equivalences. The equivalence relation on $c$ has the following profile $(\simeq) : c \rightarrow c \rightarrow Type$\footnote{This 'Type' would be a 'Prop' in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.

All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants. For this reason, we define a notion of Set\footnote{This notion of Set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$. All the algebraic structures will later extend this interface :

\begin{figure}[H]
\begin{center}
\begin{lstlisting}
interface Set c where
    ($\simeq$) : c $\rightarrow$ c $\rightarrow$ Type

    refl : (x:c) $\rightarrow$ x $\simeq$ x
    sym : {x:c} $\rightarrow$ {y:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ x)
    trans : {x:c} $\rightarrow$ {y:c} $\rightarrow$ {z:c} 
            $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)

    set_eq : (x:c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
\end{lstlisting}
\end{center}
\caption{Set}
\label{Set}
\end{figure}
If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the $Set$ implementation.
Note that $(\simeq)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--. That's natural, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.

Obviously, there is no tactic associated to $Set$, since we have no operations and no properties associated to this structure. Therefore, equivalences in a $Set$ are just the "syntactic equivalences", and they can simply be proven with $refl$\footnote{$refl$ should not be confused with $Refl$, the only constructor of $=$. But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of $refl$ will be $Refl$. Therefore, $refl$ of the interface $Set$ is a generalisation of the most specialised $Refl$ for $=$}.

As it was the case in section~\ref{sect:aSimpleProofAutomation}, the kernel of our machinery will be a function $reduce$ that will be composed of multiples functions composed together, that each do a part of the normalisation and that generate the proof of equivalence between old and new concrete values for this part of the normalisation. That was the two functions $expr\_l$ and $l\_expr$ on the smaller problem presented in chapter~\ref{sect:aSimpleProofAutomation}.

Working with equivalences instead of equalities will bring one complication. These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use idris' "rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal. More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use "rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$. This is a classical problem of working within a setoid, and this problem can be slightly mitigated when the programming language offers a good support for rewriting terms in setoids. However, Idris isn't equipped with any automation for setoids, ans everything will have to be done by hand. For this purpose, we will use the following lemma :

$eq\_preserves\_eq : \{c:Type\} \rightarrow (Set\ c) \rightarrow (x:c) \rightarrow (y:c) \rightarrow (c1:c) \rightarrow (c2:c) \rightarrow (x \simeq c1) \rightarrow (y \simeq c2) \rightarrow (c1 \simeq c2) \rightarrow (x \simeq y)$.

This lemma says that the equivalence preserves the equivalence, which means that in order to prove $x \simeq y$, we can prove the (hopefully) simplest problem $c1 \simeq c2$, provided that $x \simeq c1$ and that $y \simeq c2$. This lemma will be intensively used for building the proofs of equivalence (between the old and new indices) that are required by the development of our machinery.

\begin{proof}
We have $x \simeq c1$ and we have $c1 \simeq c2$, therefore we have $x \simeq c2$ by the use of the axiom of transitivity (of the interface $Set$). \

We also have $y \simeq c2$ and therefore we have $c2 \simeq y$ by the use of the symmetry axiom (of the interface $Set$). \

Now that we have these fresh proofs of $x \simeq c2$ and of $c2 \simeq y$, we can use one last time the property of transitivity, in order to get a proof of $x \simeq y$.

\end{proof}

This lemma will be basically used as a replacement to the "rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.

		\subsection{Hierarchy of interfaces}

Each interface will contain the axioms of the corresponding abstract structure. The first structure, completely trivial, is the magma. A magma is a structure built on top of $Set$, that just adds a $Plus$ operation, and no specific properties about it :

\begin{lstlisting}
interface Set c => Magma c where
    Plus : c $\rightarrow$ c $\rightarrow$ c
\end{lstlisting}

This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equipped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.
In fact, there is an additional requirement that will apply for all operators (in this case, the Plus operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $Plus$ : \\ $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((Plus\ c1\ c2) \simeq (Plus\ c1'\ c2'))$ \\ 
This requirement comes from the fact that we're dealing with any equivalence relation. The user is free to use his own equivalence relation, but it should be compatible with the operations that he is using.
As it was the case for $Set$, there is no tactic to write for $Magma$, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by $refl$\footnote{$refl$ should not be confused with $Refl$, the only constructor of $=$. But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of $refl$ will be $Refl$. Therefore, $refl$ of the interface $Set$ is a generalisation of the most specialised $Refl$ for $=$}.

A semi-group is a magma (ie, it still has a $Plus$ operation), but moreover it has the property of associativity for this operation.

\begin{lstlisting}
interface Magma c => SemiGroup c where
    Plus_assoc : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ (c3:c) 
               $\rightarrow$ (Plus (Plus c1 c2) c3 $\simeq$ Plus c1 (Plus c2 c3))
\end{lstlisting}

The Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called $Zero$.

\begin{lstlisting}
interface SemiGroup c => Monoid c where
    Zero : c    
    Plus_neutral_1 : (c1:c) $\rightarrow$ (Plus Zero c1 $\simeq$ c1)    
    Plus_neutral_2 : (c1:c) $\rightarrow$ (Plus c1 Zero $\simeq$ c1)
\end{lstlisting}

The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring, Ring, but we only show Group for reasons of space :

\begin{lstlisting}
interface Monoid c => Group c where
    Minus : c $\rightarrow$ c $\rightarrow$ c
    Neg : c $\rightarrow$ c
    Minus_simpl : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
    Plus_inverse : (c1:c) $\rightarrow$ (Plus c1 (Neg c1) $\simeq$ Zero, Plus (Neg c1) c1 $\simeq$ Zero)
\end{lstlisting}

As discussed on the smaller problem, the algorithm of normalisation will not directly use the concrete values of type $c$, but reflected terms instead, indexed over the concrete values. That still holds here.

### 2017-03-28 paper-CICM-17/paper.tex
added basic tactic for Nat and + assoc
Dammit, not enough space, will have to remove section 2 on the smaller tactic

@@
 We will show how type-safe reflection, where reflected terms are indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.
-
-After a presentation of our key ideas on a small example, we will build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
+We start with the presentation of our key ideas on a small example, and then we build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
 We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.
@@
 
+\input{./content/2-basicTactic.tex}
+
+\input{./content/3-hierarchyOfTactics.tex}
+
 \end{document}

### 2017-03-29 paper-CICM-17/content/1-introduction.tex
Related work to write + minor modifs to do

@@
 
-\item We present a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing a
-direct way to pull out the proofs, and providing the guarantee that the reflection of a term is indeed a faithful representation of the term.
-The basic ideas of the technique are first presented in Section~\ref{sect:aSimpleProofAutomation} on a smaller problem with only natural numbers and addition, and are later adapted for a hierarchy of tactics proving equalities in algebraic strcures, in Section~\ref{sect:ProvingEquivalencesInAlgebraicStructures}.
+\item We present in section~\ref{sect:typeSafeReflection} a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing a direct way to pull out the proofs, and providing the guarantee that the reflection of a term is indeed a faithful representation of the term.
 
-\item The normalisation procedures are implemented by following a
-correct by construction approach, instead of implementing a
-normalisation procedure, and proving afterwards that this function
-is correct with external lemmas. This approach is much more suitable
-for programming languages like Idris. Again, this approach will be
-presented in Section~\ref{sect:aSimpleProofAutomation} on the smaller problem, 
-and in Sections~\ref{sect:ProvingEquivalencesInAlgebraicStructures} for the complete hierarchy of algebraic structures.
+\item The normalisation procedures are implemented by following a correct by construction approach, instead of proving the correctness afterwards with auxiliary lemmas.
+This approach is much more suitable for programming languages like Idris.
+This correct-by-construction aspect is developed in section~\ref{sect:correctByConstruction}
 
@@
 rewriting machinery of the structure from which it naturally inherits.
-For example, simplifying neutral elements is only implemented at the monoid level,
-and each level inheriting from it will reuse it.
-Re-usability is difficult to obtain when we want to reuse the prover of a less expressive structure.
+For example, simplifying neutral elements is only implemented at the monoid level, and each level inheriting from it will reuse it.
+Reusability is difficult to obtain when we want to reuse the prover of a less expressive structure.
 For example, reusing the monoid prover for building the
@@
 negations ($-x$) and subtractions ($x-y$). Some specific encodings
-will be presented in Section~\ref{sect:NormalisationsFunctionsAndReusabilityOfTheProvers} for solving this problem.
+will be presented in Section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
 

### 2017-03-29 paper-CICM-17/content/3-hierarchyOfTactics.tex
Related work to write + minor modifs to do

@@
-\section{A hierarchy of tactics}
+\section{A hierarchy of provers}
 
-In this section, we want to prove equalities not only for $Nat$ but for various datatypes, and for various properties.
-With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic prover that works for any type that is an instance of an algebraic structure.
-In fact, that will be a hierarchy of provers --with one prover for each algebraic structure-- that can prove equalities of terms in any instance of the corresponding structure.
-We will no longer only work with the properties of associativity and of right neutral element for natural numbers as it was the case in the previous section, but we will have available the properties of a given algebraic structure.
-These tactics will be usable on any type that satisfies these properties.
-The properties of a given algebraic algebraic structure will be expressed in a corresponding interface.
+What we want to build is a prover able to prover not only equalities for $Nat$, $List$,or nay specific type but for various datatypes, and for various properties.
+With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for Semi-Groups, Monoids, Commutative-Monoids, Groups, Commutative-Groups, Semi-Rings and Rings.
+These tactics will be usable on any type that satisfies the properties of one of these algebraic structure.
+The properties of a given algebraic algebraic structure will be expressed in a corresponding interface (sometimes called typeclass in other progamming languages).
 This interface will extend the interface from which it naturally inherits.
@@
 
-\begin{figure}[H]
-\begin{center}
 \begin{lstlisting}
@@
     ($\simeq$) : c $\rightarrow$ c $\rightarrow$ Type
-
     refl : (x:c) $\rightarrow$ x $\simeq$ x
     sym : {x:c} $\rightarrow$ {y:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ x)
-    trans : {x:c} $\rightarrow$ {y:c} $\rightarrow$ {z:c} 
-            $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)
-
+    trans : {x:c} $\rightarrow$ {y:c} $\rightarrow$ {z:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)    
     set_eq : (x:c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
 \end{lstlisting}
-\end{center}
-\caption{Set}
-\label{Set}
-\end{figure}
+
 If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the $Set$ implementation.
@@
 
-As it was the case in section~\ref{sect:aSimpleProofAutomation}, the kernel of our machinery will be a function $reduce$ that will be composed of multiples functions composed together, that each do a part of the normalisation and that generate the proof of equivalence between old and new concrete values for this part of the normalisation.
-That was the two functions $expr\_l$ and $l\_expr$ on the smaller problem presented in chapter~\ref{sect:aSimpleProofAutomation}.
+%The kernel of our machinery will be a function $norm$ that will be composed of %multiples functions composed together, that each do a part of the normalisation %and that generate the proof of equivalence between old and new concrete values %for this part of the normalisation.
 
@@
 
-This lemma says that the equivalence preserves the equivalence, which means that in order to prove $x \simeq y$, we can prove the (hopefully) simplest problem $c1 \simeq c2$, provided that $x \simeq c1$ and that $y \simeq c2$.
-This lemma will be intensively used for building the proofs of equivalence (between the old and new indices) that are required by the development of our machinery.
-
 \begin{proof}
@@
 Now that we have these fresh proofs of $x \simeq c2$ and of $c2 \simeq y$, we can use one last time the property of transitivity, in order to get a proof of $x \simeq y$.
-
+\qed
 \end{proof}
 
-This lemma will be basically used as a replacement to the "rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.
+This lemma says that the equivalence preserves the equivalence, which means that in order to prove $x \simeq y$, we can prove the (hopefully) simplest problem $c1 \simeq c2$, provided that $x \simeq c1$ and that $y \simeq c2$.
+This lemma will be intensively used for building the proofs of correctness of the normalisation functions that will help to build the desired proof of $x \simeq y$.
+It will be basically used as a replacement to the "rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.
 
@@
 
-Each interface will contain the axioms of the corresponding abstract structure.
+The operations, constants and properties of each algebraic structure will be described in a dedicated interface.
 The first structure, completely trivial, is the magma.
@@
 interface Set c => Magma c where
-    Plus : c $\rightarrow$ c $\rightarrow$ c
+    + : c $\rightarrow$ c $\rightarrow$ c
 \end{lstlisting}
 
-This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equipped with the equivalence relation $\simeq$ and the test $set\_eq$), and if it has a $Plus$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the Plus operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $Plus$ : \\ $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((Plus\ c1\ c2) \simeq (Plus\ c1'\ c2'))$ \\
+This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test $set\_eq$), and if it has a $+$ operation.
+In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.
 The user is free to use his own equivalence relation, but it should be compatible with the operations that he is using.
-As it was the case for $Set$, there is no tactic to write for $Magma$, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by $refl$\footnote{$refl$ should not be confused with $Refl$, the only constructor of $=$.
-But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of $refl$ will be $Refl$.
-Therefore, $refl$ of the interface $Set$ is a generalisation of the most specialised $Refl$ for $=$}.
+As it was the case for $Set$, there is no tactic to write for $Magma$, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by $refl$.
 
@@
 interface Magma c => SemiGroup c where
-    Plus_assoc : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ (c3:c) 
-               $\rightarrow$ (Plus (Plus c1 c2) c3 $\simeq$ Plus c1 (Plus c2 c3))
+    Plus_assoc : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ (c3:c) $\rightarrow$ ((c1+c2)+c3 $\simeq$ c1+(c2+c3))
 \end{lstlisting}
 
-The Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called $Zero$.
+As example of magma, we can cite $Nat$ equipped with its operation of addition, $List$ and the operation of concatenation.
+
+Now, the Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called $Zero$.
 
@@
     Zero : c    
-    Plus_neutral_1 : (c1:c) $\rightarrow$ (Plus Zero c1 $\simeq$ c1)    
-    Plus_neutral_2 : (c1:c) $\rightarrow$ (Plus c1 Zero $\simeq$ c1)
+    Plus_neutral_1 : (c1:c) $\rightarrow$ (Zero+c1 $\simeq$ c1)    
+    Plus_neutral_2 : (c1:c) $\rightarrow$ (c1+Zero $\simeq$ c1)
 \end{lstlisting}
@@
     Neg : c $\rightarrow$ c
-    Minus_simpl : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ Minus c1 c2 $\simeq$ Plus c1 (Neg c2) 
-    Plus_inverse : (c1:c) $\rightarrow$ (Plus c1 (Neg c1) $\simeq$ Zero, Plus (Neg c1) c1 $\simeq$ Zero)
+    Minus_simpl : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ Minus c1 c2 $\simeq$ c1 + (Neg c2) 
+    Plus_inverse : (c1:c) $\rightarrow$ (c1 + (Neg c1) $\simeq$ Zero, (Neg c1) + c1 $\simeq$ Zero)
 \end{lstlisting}
 
-As discussed on the smaller problem, the algorithm of normalisation will not directly use the concrete values of type $c$, but reflected terms instead, indexed over the concrete values.
-That still holds here.
+The notion of group uses two new operations ($Neg$ and $Minus$), but $Minus$ is not primitive and can be simplified with $+$ and $Neg$.
+The important property of a group is that every element $c1$ must admit $Neg\ c1$ as inverse element for the $+$ operation.
+
+\subsection{Type-safe reflection}
+\label{sect:typeSafeReflection}
+
+When trying to prove an equivalence $x \simeq y$, the universally-quantified variables have been abstracted, and they are now part of the context.
+Our tatics will work by normalising both sides of the "potential equivalence" $x \simeq y$, and afterwards comparing them with a syntactic equality test.
+The problem is that such a normalisation function will need to do different treatments for a a variable, for a constant and for an operator.
+For this reason, we will need to work by reflection.
+There will be one type of reflected terms for each algebraic structure.
+The novelty is not the use of reflected terms (this has already been done, as discussed in~\ref{references}), but it is the use of a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent.
+Each of these datatype is parametrised over a type $c$, which is the real type on which we want to prove equalities (the $carrier$ type).
+It is also indexed over an implementation of the corresponding interface for $c$ (we usually call it $p$, because it behaves as a $proof$ telling that the structure $c$ has the desired properties), indexed over a context of abstracted variables (a vector $\Gamma$ of $n$ elements of type $c$), and most important it is also indexed over a value of type $c$, which is precisely the concrete value being encoded.
+
+A magma is only equipped with one operation $+$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.
+
+\begin{lstlisting}
+data ExprMa : Magma c $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
+    ConstMa : (p : Magma c) $\rightarrow$ ($\Gamma$:Vect n c) $\rightarrow$ (c1:c)  $\rightarrow$ ExprMa p $\Gamma$ c1 
+    PlusMa : {p : Magma c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprMa p $\Gamma$ c1 $\rightarrow$ ExprMa p $\Gamma$ c2 $\rightarrow$ ExprMa p $\Gamma$ (c1+c2) 
+    VarMa : (p:Magma c) $\rightarrow$ ($\Gamma$:Vect n c)
+         $\rightarrow$ (i:Fin n) $\rightarrow$ ExprMa p $\Gamma$ (index i $\Gamma$)
+\end{lstlisting}
+
+For an expression $e_x\ :\ ExprMa\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
+When an expression is a variable $(VarMa\ \_\ \Gamma\ i)$, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
+The expression $(ConstMa\ \_\ \Gamma\ k)$ denotes the constant $k$ of type $c$.
+Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression $PlusMa\ e_x\ e_y$ denotes the concrete value $(x + y)$.
+Because the reflected terms embed their corresponding inputs, they are guarantee to be faithful representations of them.
+
+There is no additional operations in SemiGroup and Monoid, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for Magma that we have given above.
+However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :
+
+\begin{lstlisting}
+data ExprG :  Group c $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
+    ConstG : (p : Group c) $\rightarrow$ ($\Gamma$:Vect n c) $\rightarrow$ (c1:c) $\rightarrow$ ExprG p $\Gamma$ c1
+    PlusG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ c2 $\rightarrow$ ExprG p $\Gamma$ (c1+c2)
+    MinusG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} $\rightarrow$ {c2:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ c2 $\rightarrow$ ExprG p $\Gamma$ (Minus c1 c2)
+    NegG : {p : Group c} $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {c1:c} 
+         $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ (Neg c1)
+    VarG : (p : Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+         $\rightarrow$ (i:Fin n) $\rightarrow$ ExprG p $\Gamma$ (index i $\Gamma$)
+\end{lstlisting}
+
+The index of type $c$ (the real value encoded by an expression) is always expressed by using the lookup function $index$ and the available operations in the implementation $p$, which for a group are $+$, $Minus$ and $Neg$.
+
+	\subsection{A correct by construction approach}
+\label{sect:correctByConstruction}
+
+We want to write the reduction function on a \emph{correct by construction} way, which means that no additional proof should be required after the definition of the function.
+Thus, $norm$ will produce the proof that the new expression freshly produced has the same interpretation as the original expression, and this will be made easier by the fact that all the datatypes for reflecting terms ($ExprSG$, $ExprG$, $ExprR$) are indexed over the real --concrete-- value : a term of type $Expr\ \Gamma\ x$ is the encoding of the concrete value $x$ in the context $\Gamma$.
+Thus, for each algebraic structure, the type of the function $norm$ will look like this : \\
+$norm :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ \simeq\ x'))$ \\
+Every function of normalisation $norm$ will produce a dependent pair : the new concrete value $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
+
+This function doesn't simply produce an $Expr\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
+
+We will explain how the tactic works by commenting the construction of the tactic for Group, but the other algebraic structures will be equipped similarly.
+The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a group.
+The reflected term for $x$ will be denoted $e_x$, and this term will have the type $ExprG\ p\ \Gamma\ x$, which means that $e_x$ is guaranteed to be the encoding of $x$ because $e_x$ is precisely indexed over the real value $x$.
+We've got similar things for $y$, which is encoded by $e_y$, and its type is indexed over the real value $y$.
+Running the normalisation procedure on $e_x$ will produce the normal form $e_{x'}$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p_x$ of $x \simeq x'$ while running the normalisation procedure on $e_y$ will produce the normal form $e_{y'}$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p_y$ of $y \simeq y'$.
+It is now enough to simply compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test $exprG\_eq$ because these two expressions are now supposed to be in normal form.
+
+Now, if the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$, which implies $x' \simeq y'$ because the propositional equality is more restrictive than an equivalence relation.
+By using the two equivalences $x \simeq x'$ and $y \simeq y'$ that we obtained during the normalisations in the new equivalence $x' \simeq y'$ with the use of $eq\_preserves\_eq$, we can get a proof of $x \simeq y$.
+This is what the function $buildProof$ is doing :
+
+\begin{lstlisting}
+buildProofGroup : (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {x : c} $\rightarrow$ {y : c} 
+  $\rightarrow$ {x':c} $\rightarrow$ {y':c} $\rightarrow$ (ExprG p $\Gamma$ x') $\rightarrow$ (ExprG p $\Gamma$ y') 
+  $\rightarrow$ (x $\simeq$ x') $\rightarrow$ (y $\simeq$ y') $\rightarrow$ (Maybe (x $\simeq$ y))
+buildProofGroup p ex' ey' px py with (exprG_eq p ex' ey')
+    buildProofGroup p ex' ex' px py | Just refl = ?MbuildProofGroup
+    buildProofGroup p ex' ey' px py | Nothing = Nothing
+\end{lstlisting}
+
+The argument of type $ExprG\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
+Before the normalisation, the reflected LHS was reflecting the value $x$.
+The same applies to the right-hand side.
+The function also expects proofs of $x \simeq x'$ and of $y \simeq y'$, that we will be able to provide because the normalisation function has produced them while it was building the normalised reflected terms.
+
+% TALK ABOUT THE METAVARIABLE MbuildProodGroup to add here !
+
+Finally, the main function which tries to prove the equality $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs that have been obtained :
+
+\begin{lstlisting}
+groupDecideEq : (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {x : c} $\rightarrow$ {y : c} 
+               $\rightarrow$ (ExprG p $\Gamma$ x) $\rightarrow$ (ExprG p $\Gamma$ y) $\rightarrow$ Maybe (x $\simeq$ y)
+groupDecideEq p ex ey =
+  let (x' ** (ex', px)) = groupNormalise p ex in
+  let (y' ** (ey', py)) = groupNormalise p ey in
+	     buildProofGroup p ex' ey' px py
+\end{lstlisting}
+
+The remaining part is to define the function $norm$.
+Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
+The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities.
+Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
+These normalisation functions will compute the canonical representation of the reflected input term, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the original real value (indexing the input term) and the produced real value (indexing the produced term).
+Thus, the type of the reduction function for group is :
+
+\begin{lstlisting}
+groupNormalise : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (p:Group c) 	   	   	   
+	   $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {x:c} $\rightarrow$ (ExprG p $\Gamma$ x) 
+	   $\rightarrow$ (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
+\end{lstlisting}
+
+The functions of normalisation have more work to do in structures with multiple axioms (like Group and Ring), than for the easier structure (like Semi-Group and Monoid).
+The details of these functions will be given in the next section.
+
+\subsection{Normalisation functions}
+\label{sect:normalFormShape}
+
+For the sake of completeness, we will describe the normal form for the ring structure, as it has all the operations and all the properties about these operations.
+The normalisation function takes in input an expression expressed with sums, products, constants (zero, one...) and variables that belong to an ordered set $\mathcal{V}$ of variables.
+In short, the normalisation function takes in input a polynomial of multiple variables.
+In output, it must produce a normal form representing the same polynomial.
+Therefore, we have to decide a canonical representation of polynomials.
+Many canonical representations can be used.
+We decide to stick with classical mathematical conventions.
+The first one, is that the polynomial will be completely developed, ie, the distributivity of $*$ over $+$ will be applied until it can't be applied anymore.
+The advantage is the simplicity, as factorising would be significantly more complicated and we're looking for simplicity as the normalisation functions will also have to produce the proof of equivalence between the old and the new index.
+Because the polynomial is completely developed, at the toplevel, it is a sum :
+
+\[
+P = \sum_{i=1}^{a}\ (\prod_{j=1}^{b} Monomial_{i}^j)
+\text{ where } 
+Monomial_{i}^j = C_{i}^j * \prod_{k=1}^{c} Var_{i,k}^{j}
+\]
+with $C_{i}^j$ a constant, and $Var_{i,k}^{j}$ one of the variable that belong to $\mathcal{V}$.
+
+One could be surprised by the fact that the normal form is a sum of product of monomial, and not directly a sum of monomials.
+The reason is the following.
+A monomial is a product of a constant $C_{i}^j$ (like $4$) and of a product of variables (like $x*y*z$).
+For example,  $5*(x*(y*z))$ is a monomial.
+Now let's consider the term $(5*(x*(y*z))) * (4*(z*z))$.
+This term is not a monomial, but we could be tempted to simplify it into the monomial $20*(x*(y*(z*(z*z))))$.
+However, that would assume that the product is always commutative, ie, that we can re-organize the subterms of a product as we want.
+This is the case in a commutative ring, but this does not hold for any ring.
+Therefore, after the full development, the polynomial is a sum of \emph{product of monomials}, and not directly a sum of monomials.
+The only rearrangement that can be done with the multiplication is to check if two constants are consecutive in a product, and if so, to replace them by the constant that represents the product of them.
+
+However, because $+$ is always a commutative operator in a ring, the different products of monomials themselves can be rearranged in different ways in this sum.
+That will be done at the level of the commutative group prover if we can provide an ordering between products of monomials.
+
+The ordering on product of monomials will be defined by using an ordering on monomials, that we will call $isBefore\_mon$, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
+\[
+Monomial_{i}^{j} = C_{i}^{j} * (Var_{i,1}^{j} * \prod_{k=2}^{c} Var_{i,k}^{j})
+\text{ and }
+Monomial_{i'}^{j'} = C_{i'}^{j'} * (Var_{i',1}^{j'} * \prod_{k=2}^{c'} Var_{i',k}^{j'})
+\]
+
+The order between these two monomials will be decided by looking at the order between the first variables $Var_{i,1}^{j}$ and $Var_{i',1}^{j'}$.
+In the case where both monomials start with the same variable, we continue by inspecting the remaining variables.
+If we can't continue because one of the two monomials has less variables than the other, then we decide that the one with fewer variables comes first (exactly like the word "house" comes before the word "housemate" with the lexicographic order).
+
+We can now build the order on \emph{product of monomials} that we needed.
+We name it $isBefore$.
+Given two product of monomials $Prod_{i}$ and $Prod_{i'}$ we need to decide which one comes first.
+We will obviously use the order $isBefore\_mon$ on the first monomials of these two products.
+If it says that $Monomial_{i}^{1}$ comes before $Monomial_{i'}^{1}$, then we decide that $Prod_{i}$ comes before $Prod_{i'}$.
+Conversely, if that's $Monomial_{i'}^{1}$ that comes first, then we decide that $Prod_{i'}$ comes first.
+However, if $Monomial_{i}^{1}$ and $Monomial_{i'}^{1}$ have exactly the same position in the order, then we continue by inspecting the remaining monomials with a recursive call on $isBefore$.
+As previously, if we can't continue because one of the two products has less monomials than the other, then the one with fewer monomials will come first.
+
+A few additional conventions had to be decided about the normal form :
+\begin{itemize}
+
+\item The top-level sum of the polynomial will be written on the completely right-associative form : \\
+$prodMon_1\ +\ (prodMon_2\ + (prodMon_3\ +\ (...\ +\ prodMon_a)))$
+\item The two levels of product that we have, namely the products of monomials (at the main level) and the products of variables (at the monomial level), will also be written on the completely right-associative form.
+\item We always simplify as much as possible with constants.
+That includes doing the simplification for the addition with zero and for the multiplication with one, doing the computation between two constants that are nearby in a sum or in a product, etc...
+\item We always replace the binary operator $Minus$ by a sum and a $Neg$ (following $a-b=a+(-b)$) since the binary operator $Minus$ are just allowed for the user's convenience, but internally we simplify them all.
+\item We always simplify the sum of an expression $e$ and its inverse $-e$ into the constant zero.
+
+\end{itemize}
+
+\subsection{Reusing the provers}
+\label{sect:reusabilityOfTheProvers}
+
+The normalisation function of each structure will use as much as possible the normalisation function of the structure from which it inherits.
+For example, the normalisation on monoids will reuse the normalisation on semi-groups so that it won't have to deal with the associativity.
+For this case, that's not a problem since the datatype reflecting terms in semi-groups has the same expressivity power than the one for monoids.
+However, that's not the case with groups and monoids: if the function of normalisation for group needs to call the normalisation for monoid, it will have to encode the negations that can't be directly expressed in a monoid.
+For this reason, we will develop some specific encodings.
+
+The idea is that we will encode negations as variables, and we will let the monoid prover deal with them as ordinary variables.
+In order to achieve this goal, we need the following datatype that will help us to distinguish between a real variable and the encoding of a negation :
+
+\begin{lstlisting}
+data Variable : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
+  RealVariable : ($\Gamma$:Vect n c) $\rightarrow$ (i:Fin n) $\rightarrow$ Variable $\Gamma$ (index i $\Gamma$)
+  EncodingNeg : ($\Gamma$:Vect n c) $\rightarrow$ (i:Fin n) $\rightarrow$ Variable $\Gamma$ (Neg (index i $\Gamma$))
+\end{lstlisting}
+
+We only need to encode negations of variables, as the negation of a constant can be simplified into a constant.
+Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in group will have first systematically propagated $Neg$ inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
+
+The constructor for variables now takes a Variable as parameter, instead of taking directly an element of $(Fin\ n)$ :
+
+\begin{lstlisting}
+    VarG : (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {val:c} 
+           $\rightarrow$ (Variable $\Gamma$ val) $\rightarrow$ ExprG p $\Gamma$ val
+\end{lstlisting}
+
+Thanks to this encoding, we can now transform an $ExprG$ to an $ExprMo$.
+A constant $(ConstG\ p\ \Gamma\ c1)$ will be transformed into the corresponding constant $(ConstMo\ p\ \Gamma\ c1)$, a $PlusG$ into the corresponding $PlusMo$, a real variable into the same real variable, the negation of a constant into the resulting constant, and finally the negation of a variable $i$ into a $(VarMo\ p\ (EncodingNegOfVar\ \Gamma\ i))$.
+
+We will use a similar technique for converting an expression from the ring level to the commutative-group level, where we will encode the product of monomials because the product is not an operation define at the commutative-group level.
+That will enable the function of normalisation for ring to benefit from the normalisation function for commutative-group.
+
+\subsection{Automatic reflection}
+
+We've built an automatic reflection mechanism, in order to let the machine build automatically the reflected terms for us.
+This is not something absolutely vital, as it only simplifies the usage of the tactics, avoiding to write long encodings by hand.
+To do so, we've used Idris' reflection mechanism, which enables to do pattern matching on syntax, rather than on constructors.

### 2017-03-29 paper-CICM-17/content/4-relatedWork.tex
Related work to write + minor modifs to do

\section{Related work}
\label{sect:relatedWork}

blablabla

### 2017-03-29 paper-CICM-17/content/5-conclusions.tex
Related work to write + minor modifs to do

\section{Results, conclusions and futur work}

We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were defying functions that use dependent types. This solution takes the form of a hierarchy of provers for equivalence in algebraic structures. These provers are generic in multiple ways: they work for many algebraic structures (Semi-Group, Monoid, Commutative Monoid, Group, Commutative Group, Ring and Semi-Ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality). The implementation is modular and each prover reuses the prover of the structure from which it directly inherits. \\

These provers can automatically prove equivalence between terms, which discharges the user of providing proofs for all the proof obligations that appear naturally when using dependent types. Thefore, these provers enable the user to focus on the interesting and main proofs of the development that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated. We think that the human time is best used when it is devoted for proving things that can't be automated. \\

The correct-by-construction method that we have followed has involved the design of a type-safe reflection mechanism where reflected terms are indexed over the concrete expressions, and from which we are able to pull out the proofs easily. The construction of the proof is done step by step, and it follows the construction of the normalised terms. Therefore, unlike other implementations --like Coq's implementation of a ring prover-- we do not have the traditional duplication that happens when separating the computational content from the proof of correctness. Instead, in our approach, the construction of the proof is guided by the construction of the normalised terms, which, in addition of avoiding redundancy, simplifies the proof generation considerably. This development shows that if dependent types effectively bring some new problems, especially the index mismatch, they are also very expressive tools for building correct-by-construction software where the development is driven by the types. \\

The work that we have presented here can be extended to build new provers for less common algebraic structures and for more specific structures. For example, regular expressions form a Pre Semi-Ring with some extra axioms. We want to refactor the Semi-Ring level with the creation of the intermediate structure of Pre Semi-Ring, that wouldn't have the property that $0$ is an annihilator element for the product. Then, we could build a specific prover for regular expressions, that would use the normalisation function of the Pre Semi-Ring level and that would deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new star operation.

### 2017-03-29 paper-CICM-17/paper.tex
Related work to write + minor modifs to do

--- 2017-03-26 paper-CICM-17/paper.tex
@@
 
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{amsfonts}
@@
 We will show how type-safe reflection, where reflected terms are indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.
-
-After a presentation of our key ideas on a small example, we will build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
+We build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
 We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.
@@
 
+% The chap 2 on the smaller tactic for Nat has been removed for space constraint
+
+\input{./content/3-hierarchyOfTactics.tex}
+
+\input{./content/4-relatedWork.tex}
+
+\input{./content/5-conclusions.tex}
+
+\bibliographystyle{splncs03}
+\bibliography{dtp}
+
 \end{document}

### 2017-03-29 paper-CICM-17/content/4-relatedWork.tex
CICM 17 : first draft for related work

@@
 
-blablabla
+As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism.
+Their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
+\\
+
+Doing proofs by reflection has been intensively investigated by many authors, including Adam Chlipala, in particular in his book~\cite{ChlipalaBook} and in~\cite{Malecha14}.
+However, we haven't found anything similar to the type-safe reflection technique that we've presented in this paper.
+\\
+
+At the connexion bewteen proof automation and proof by reflection, we can cite SSreflect~\cite{GonthierTuto}, which is an extension of Coq which aims to provide additional formalisations and tactics suitable for long mathematical proofs.
+\\
+
+Proof automations have been investigated for various proof assistants and programming languages.
+In Coq, appart from the $ring$ prover~\cite{Coq2005}, there's also the $omega$ solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
+a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
+Various automations have also been done in Agda~\cite{Lindblad04}.
+\\
+
+$Mtac$~\cite{Ziliani13} is an extension to Coq that supports dependently-typed tactic programming that uses monad to avoid the need to touch the trusted kernel typechecker of Coq. \\
+
+Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts.

### 2017-03-29 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM 17 : biblio to continue

@@
 
-The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring, Ring, but we only show Group for reasons of space :
+The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring and Ring, but we only show the Group interface for reasons of space :
 
@@
 There will be one type of reflected terms for each algebraic structure.
-The novelty is not the use of reflected terms (this has already been done, as discussed in~\ref{references}), but it is the use of a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent.
+The novelty is not the use of reflected terms (this has already been done, as discussed in section~\ref{sect:relatedWork}), but is the use of a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent.
 Each of these datatype is parametrised over a type $c$, which is the real type on which we want to prove equalities (the $carrier$ type).

### 2017-03-29 paper-CICM-17/content/4-relatedWork.tex
CICM 17 : biblio to continue

@@
 
-As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism.
-Their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
+As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism, and without the correct-by-construction approach : they have first defined the normalisation of terms, and afterwards they've proven the correctness of the normalisation with the external lemma :
+$\forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ \simeq\ reify\ e2$ which therefore needs a $reify$ function that we do not need in our development.
+Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on ssyntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
+\\
+
+Proof automations have been investigated for various proof assistants and programming languages.
+In Coq, appart from the $ring$ prover~\cite{Coq2005}, there's also the $omega$ solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
+a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
+Various automations have also been done in Agda~\cite{Lindblad04}.
 \\
@@
 
-Proof automations have been investigated for various proof assistants and programming languages.
-In Coq, appart from the $ring$ prover~\cite{Coq2005}, there's also the $omega$ solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
-a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
-Various automations have also been done in Agda~\cite{Lindblad04}.
+Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts.
 \\
 
-$Mtac$~\cite{Ziliani13} is an extension to Coq that supports dependently-typed tactic programming that uses monad to avoid the need to touch the trusted kernel typechecker of Coq. \\
-
-Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
-This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts.
+We also need to mention $Mtac$~\cite{Ziliani13}, which is an extension to Coq that supports dependently-typed tactic programming that uses monad to avoid the need to touch the trusted kernel typechecker of Coq.

### 2017-03-29 paper-CICM-17/content/1-introduction.tex
paper almost ready, still need to proofread a lot

@@
 
-Proofs assistants like Coq(ref), Agda(ref) and Idris(ref) are based on Intensional Type Theories, that contain two notions of equality.
+Proofs assistants like Coq~\cite{BertotC04}, Agda(ref) and Idris~\cite{brady2013idris} are based on Intensional Type Theories, that contain two notions of equality.
 One is the equality that can be manipulated in the language, called the propositional equality.
@@
 in formal certification.
-This is an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
+This is even an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
 
@@
 
-We revisit an example from a previous work of Edwin Brady~\cite{bradytfp07} which shows how proof obligations may arise when a type is indexed by natural numbers.
+We revisit an example from a previous work of Edwin Brady~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
 In this development, the goal is to implement a formally verified library of binary numbers.
@@
 
-The proof that we have written by hand only uses the existence of neutral elements, and the associativity and commutativity of $+$ on $Nat$.
+The proof that we have written by hand only uses the existence of a neutral element, and the associativity and commutativity of $+$ on $Nat$.
 Thus, we're rewriting a term by using the properties of a commutative monoid.
@@
 
-Ring solvers are already implemented for various proof
-assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda/\%22?n=Libraries.UsingTheRingSolver}}. 
+Prover for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda/\%22?n=Libraries.UsingTheRingSolver}} ring prover.
 In this paper, we describe a certified
@@
 new kind of \emph{type-safe reflection}.
-Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without the type-safe aspect.
+Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without the type-safe and correct-by-construction aspects.
 We will
-compare our approach with other implementations in Section~\ref{sect:AlternativeApproaches}.
+compare our approach with other implementations in Section~\ref{sect:relatedWork}.
 

### 2017-03-29 paper-CICM-17/content/3-hierarchyOfTactics.tex
paper almost ready, still need to proofread a lot

@@
 
-What we want to build is a prover able to prover not only equalities for $Nat$, $List$,or nay specific type but for various datatypes, and for various properties.
+What we want to build is a prover able to prover not only equalities for $Nat$, $List$, or any specific type but for various datatypes, and for various properties.
 With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for Semi-Groups, Monoids, Commutative-Monoids, Groups, Commutative-Groups, Semi-Rings and Rings.

### 2017-03-29 paper-CICM-17/content/4-relatedWork.tex
paper almost ready, still need to proofread a lot

@@
 $\forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ \simeq\ reify\ e2$ which therefore needs a $reify$ function that we do not need in our development.
-Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on ssyntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
+Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on syntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
 \\
@@
 a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
-Various automations have also been done in Agda~\cite{Lindblad04}.
+Various automations have also been done in Agda~\cite{DBLP:conf/mpc/KokkeS15} ~\cite{Lindblad04}.
 \\

### 2017-03-30 paper-CICM-17/content/1-introduction.tex
some fixes

@@
 
-Proofs assistants like Coq~\cite{BertotC04}, Agda(ref) and Idris~\cite{brady2013idris} are based on Intensional Type Theories, that contain two notions of equality.
+Proofs assistants like Coq~\cite{BertotC04}, Agda~\cite{DBLP:conf/tldi/Norell09} and Idris~\cite{brady2013idris} are based on Intensional Type Theories, that contain two notions of equality.
 One is the equality that can be manipulated in the language, called the propositional equality.
@@
 The only way for $(Id_a\ x\ y)$ to be inhabited is therefore that $x$ should be
-(by definition) $y$. In this case, the constructor $Refl$ helps to create a
-proof of this equality : $(Refl\ x)$ is precisely the proof which says that
+(by definition) $y$. In this case, the constructor \code{Refl} helps to create a
+proof of this equality : \code{(Refl\ x)} is precisely the proof which says that
 $x=_ax$.
@@
 
-then we have : \\
+then we have the following induction principle for $T$: \\
 $T\_ind : \forall P:T \rightarrow Type,\ (P\ T0) \rightarrow (\forall t:T,\ P\ t \rightarrow P\ (T1\ t)) \rightarrow (\forall t:T,\ P\ t)$.
@@
 For example, we can prove that $n+0 = n$ for all $n$ by induction on the $Nat\ n$, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
-That means that the axiom of induction has made the type $Id_{a}(x,y)$ bigger than it looks like : it contains terms that are not generated by $Refl$ (the only constructor of $Id_{a}(x,y)$), but that are added by the inductive principles.
+That means that the axiom of induction has made the type $Id_{a}(x,y)$ bigger than it looks like : it contains terms that are not generated by \code{Refl} (the only constructor of $Id_{a}(x,y)$), but that are added by the inductive principles.
 There are therefore things which are provably equal which unfortunately are not definitionally equal after unfolding all the definitions and after full reductions.
@@
 In this development, the goal is to implement a formally verified library of binary numbers.
-In order to ensure the correctness of the methods that will be developed, we define the types $Bit$ and $Binary$ as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.
+In order to ensure the correctness of the methods that will be developed, we define the types \code{Bit} and \code{Binary} as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.
 
@@
 This function produces a dependent pair.
-The first component is the value $bX$ (expressed as a natural number) representing the first bit produced (therefore of type $Bit\ bX$), and the second component is another dependent pair.
-This second dependent pair is composed of the value $bY$ representing the second bit produced, and of a triple containing the two bits produced and a proof that these two bits correctly encode the result of the addition.
-For example, on the fifth line, which corresponds to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit $b0$ (encoding $0_2$) and the bit $b1$ (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
+The first component is the value \code{bX} (expressed as a natural number) representing the first bit produced (therefore of type \code{Bit\ bX}), and the second component is another dependent pair.
+This second dependent pair is composed of the value \code{bY} representing the second bit produced, and of a triple containing the two bits produced and a proof that these two bits correctly encode the result of the addition.
+For example, on the fifth line, which corresponds to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit \code{b0} (encoding $0_2$) and the bit \code{b1} (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
 
-We finally define the function $adc$ that adds two binary numbers and a carry flag.
+We finally define the function \code{adc} that adds two binary numbers and a carry flag.
 This addition works for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
@@
 Unfortunately, this straightforward definition is rejected by Idris' type-checker because the types mismatch for both of the two patterns.
-The result of the first line $adc\ zero\ zero\ carry$ is expected to have the type $Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)$ but we provide a term of type $Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))$.
-The problem is even worse for the second case where the expected index is \\
-$(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))$
-while we're trying to provide a term indexed over \\
-$plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))$.
-We know that the definition of $adc$ we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and Idris can not figure out a proof of compatibility on its own.
-If we use provisional definitions[ref], the previous definition can be accepted by the system, but we now have two proof obligations $adc\_lemma\_1$ and $adc\_lemma\_2$ demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
+The result of the first line \code{adc zero zero carry} is expected to have the type \codeL{Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)} but we provide a term of type \code{Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))}.
+The problem is even worse for the second case where the expected index is \codeL{(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))}
+while we're trying to provide a term indexed over
+\codeL{plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))}.
+We know that the definition of \code{adc} we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and Idris can not figure out a proof of compatibility on its own.
+If we use provisional definitions[ref], the previous definition can be accepted by the system, but we now have two proof obligations \code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
 
@@
 This is time consuming and more importantly, just a small change in the left or the right hand side of the equality may invalidate the proof.
-That means that a minor change in our datatypes, or in the definition of $addBit$ or $adc$ will require us to replace this proof by another one fulfilling the same task.
+That means that a minor change in our datatypes, or in the definition of \code{addBit} or \code{adc} will require us to replace this proof by another one fulfilling the same task.
 These proofs become the everyday routine in any dependently-typed language, as soon as we index types over values in order to capture some logical properties.
 
-The proof that we have written by hand only uses the existence of a neutral element, and the associativity and commutativity of $+$ on $Nat$.
+The proof that we have written by hand only uses the existence of a neutral element, and the associativity and commutativity of $+$ on \code{Nat}.
 Thus, we're rewriting a term by using the properties of a commutative monoid.

### 2017-03-30 paper-CICM-17/content/3-hierarchyOfTactics.tex
some fixes

--- 2017-03-29 paper-CICM-17/content/3-hierarchyOfTactics.tex
@@
 
-What we want to build is a prover able to prover not only equalities for $Nat$, $List$,or nay specific type but for various datatypes, and for various properties.
+What we want to build is a prover able to prover not only equalities for \code{Nat}, \code{List}, or any specific type but for various datatypes, and for various properties.
 With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for Semi-Groups, Monoids, Commutative-Monoids, Groups, Commutative-Groups, Semi-Rings and Rings.
@@
 This interface will extend the interface from which it naturally inherits.
-For example, the $Group$ interface will extend the $Monoid$ one.
+For example, the \code{Group} interface will extend the \code{Monoid} one.
 We will end up with a hierarchy of interfaces and will write one tactic for each of these interface.
@@
 Let's call $c$ the \emph{carrier} type, ie, the type on which we want to prove equivalences.
-The equivalence relation on $c$ has the following profile $(\simeq) : c \rightarrow c \rightarrow Type$\footnote{This 'Type' would be a 'Prop' in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+The equivalence relation on $c$ has the following profile $(\simeq) : \code{c} \rightarrow \code{c} \rightarrow \code{Type}$\footnote{This \code{Type} would be a \code{Prop} in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
 
 All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants.
-For this reason, we define a notion of Set\footnote{This notion of Set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test $set\_eq$.
-All the algebraic structures will later extend this interface :
+For this reason, we define a notion of \code{Set}\footnote{This notion of set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test \code{set\_eq}.
+All the interfaces representing the algebraic structures will later extend \code{Set} :
 
@@
 
-If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the $Set$ implementation.
-Note that $(\simeq)$ is only weakly decidable in the sense that $set\_eq$ only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value $Nothing$--.
-That's natural, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
+If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
+Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--.
+That is normal, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
 
-Obviously, there is no tactic associated to $Set$, since we have no operations and no properties associated to this structure.
-Therefore, equivalences in a $Set$ are just the "syntactic equivalences", and they can simply be proven with $refl$\footnote{$refl$ should not be confused with $Refl$, the only constructor of $=$.
-But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of $refl$ will be $Refl$.
-Therefore, $refl$ of the interface $Set$ is a generalisation of the most specialised $Refl$ for $=$}.
+Obviously, there is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
+Therefore, equivalences in a \code{Set} are just the "syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
+But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}.
+Therefore, \code{refl} of the interface \code{Set} is a generalisation of the most specialised \code{Refl} for \code{=}}.
 
@@
 This is a classical problem of working within a setoid, and this problem can be slightly mitigated when the programming language offers a good support for rewriting terms in setoids.
-However, Idris isn't equipped with any automation for setoids, ans everything will have to be done by hand.
+However, Idris isn't equipped with any automation for setoids, and everything will have to be done by hand.
 For this purpose, we will use the following lemma :
 
-$eq\_preserves\_eq : \{c:Type\} \rightarrow (Set\ c) \rightarrow (x:c) \rightarrow (y:c) \rightarrow (c1:c) \rightarrow (c2:c) \rightarrow (x \simeq c1) \rightarrow (y \simeq c2) \rightarrow (c1 \simeq c2) \rightarrow (x \simeq y)$.
+$\code{eq\_preserves\_eq} : \{c:Type\} \rightarrow (Set\ c) \rightarrow (x:c) \rightarrow (y:c) \rightarrow (c1:c) \rightarrow (c2:c) \rightarrow (x \simeq c1) \rightarrow (y \simeq c2) \rightarrow (c1 \simeq c2) \rightarrow (x \simeq y)$.
 
 \begin{proof}
-We have $x \simeq c1$ and we have $c1 \simeq c2$, therefore we have $x \simeq c2$ by the use of the axiom of transitivity (of the interface $Set$). \
+We have $x \simeq c1$ and we have $c1 \simeq c2$, therefore we have $x \simeq c2$ by the use of the axiom of transitivity (of the interface \code{Set}). \
 
-We also have $y \simeq c2$ and therefore we have $c2 \simeq y$ by the use of the symmetry axiom (of the interface $Set$). \
+We also have $y \simeq c2$ and therefore we have $c2 \simeq y$ by the use of the symmetry axiom (of the interface \code{Set}). \
 
@@
 The first structure, completely trivial, is the magma.
-A magma is a structure built on top of $Set$, that just adds a $Plus$ operation, and no specific properties about it :
+A magma is a structure built on top of \code{Set}, that just adds a \code{Plus} operation, and no specific properties about it :
 
@@
 
-This code means that a type $c$ (for $carrier$) is a Magma if it is already a Set (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test $set\_eq$), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $Plus\_preserves\_equiv : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
+This code means that a type $c$ (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
+In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.
 The user is free to use his own equivalence relation, but it should be compatible with the operations that he is using.
-As it was the case for $Set$, there is no tactic to write for $Magma$, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by $refl$.
+As it was the case for \code{Set}, there is no tactic to write for \code{Magma}, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by \code{refl}.
 
-A semi-group is a magma (ie, it still has a $Plus$ operation), but moreover it has the property of associativity for this operation.
+A semi-group is a magma (ie, it still has a \code{Plus} operation), but moreover it has the property of associativity for this operation.
 
@@
 
-As example of magma, we can cite $Nat$ equipped with its operation of addition, $List$ and the operation of concatenation.
+As example of magma, we can cite \code{Nat} equipped with its operation of addition and \code{List} and the operation of concatenation.
 
-Now, the Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called $Zero$.
+The Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called \code{Zero}.
 
@@
 
-The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring and Ring, but we only show the Group interface for reasons of space :
+The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring and Ring, but we only show the \code{Group} interface for reasons of space :
 
@@
 
-The notion of group uses two new operations ($Neg$ and $Minus$), but $Minus$ is not primitive and can be simplified with $+$ and $Neg$.
-The important property of a group is that every element $c1$ must admit $Neg\ c1$ as inverse element for the $+$ operation.
+The notion of group uses two new operations (\code{Neg} and \code{Minus}), but \code{Minus} is not primitive and can be simplified with $+$ and \code{Neg}.
+The important property of a group is that every element \code{c1} must admit \code{Neg\ c1} as inverse element for the $+$ operation.
 
@@
 The novelty is not the use of reflected terms (this has already been done, as discussed in section~\ref{sect:relatedWork}), but is the use of a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent.
-Each of these datatype is parametrised over a type $c$, which is the real type on which we want to prove equalities (the $carrier$ type).
-It is also indexed over an implementation of the corresponding interface for $c$ (we usually call it $p$, because it behaves as a $proof$ telling that the structure $c$ has the desired properties), indexed over a context of abstracted variables (a vector $\Gamma$ of $n$ elements of type $c$), and most important it is also indexed over a value of type $c$, which is precisely the concrete value being encoded.
+Each of these datatype is parametrised over a type \code{c}, which is the real type on which we want to prove equalities (the \emph{carrier} type).
+It is also indexed over an implementation of the corresponding interface for \code{c} (we usually call it \code{p}, because it behaves as a \emph{proof} telling that the structure \code{c} has the desired properties), indexed over a context of abstracted variables (a vector $\Gamma$ of \code{n} elements of type \code{c}), and most important it is also indexed over a value of type \code{c}, which is precisely the concrete value being encoded.
 
-A magma is only equipped with one operation $+$. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.
+A magma is only equipped with the operation of addition. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.
 
@@
 
-For an expression $e_x\ :\ ExprMa\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
-When an expression is a variable $(VarMa\ \_\ \Gamma\ i)$, the denoted number is simply the corresponding variable in the context, ie, $(index\ i\ \Gamma)$.
-The expression $(ConstMa\ \_\ \Gamma\ k)$ denotes the constant $k$ of type $c$.
-Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression $PlusMa\ e_x\ e_y$ denotes the concrete value $(x + y)$.
+For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
+When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is simply the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
+The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$.
+Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression \code{PlusMa\ $e_x$\ $e_y$} denotes the concrete value $(x + y)$.
 Because the reflected terms embed their corresponding inputs, they are guarantee to be faithful representations of them.
@@
 There is no additional operations in SemiGroup and Monoid, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for Magma that we have given above.
-However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :
+However, the \code{Group} interface will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
 
@@
 
-The index of type $c$ (the real value encoded by an expression) is always expressed by using the lookup function $index$ and the available operations in the implementation $p$, which for a group are $+$, $Minus$ and $Neg$.
+The index of type $c$ (the real value encoded by an expression) is always expressed by using the lookup function \code{index} and the available operations in the implementation $p$, which for a group are $+$, \code{Minus} and \code{Neg}.
 
@@
 
-We want to write the reduction function on a \emph{correct by construction} way, which means that no additional proof should be required after the definition of the function.
-Thus, $norm$ will produce the proof that the new expression freshly produced has the same interpretation as the original expression, and this will be made easier by the fact that all the datatypes for reflecting terms ($ExprSG$, $ExprG$, $ExprR$) are indexed over the real --concrete-- value : a term of type $Expr\ \Gamma\ x$ is the encoding of the concrete value $x$ in the context $\Gamma$.
-Thus, for each algebraic structure, the type of the function $norm$ will look like this : \\
-$norm :\ Expr\ \Gamma\ x\ \rightarrow\ (x'\ **\ (Expr\ \Gamma\ x',\ x\ \simeq\ x'))$ \\
-Every function of normalisation $norm$ will produce a dependent pair : the new concrete value $x'$, and a pair made of an $Expr\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
+We want to write the normalisation function on a \emph{correct by construction} way, which means that no additional proof should be required after the definition of the function.
+Thus, \code{norm} will produce the proof that the new expression freshly produced has the same interpretation as the original expression, and this will be made easier by the fact that all the datatypes for reflecting terms (\code{ExprMa}, \code{ExprG}, \code{ExprR}, etc) are indexed over the real --concrete-- value : a term of type $\code{Expr}\ \Gamma\ x$ is the encoding of the concrete value $x$ in the context $\Gamma$.
+Thus, for each algebraic structure, the type of the function \code{norm} will look like this : \\
+$\code{norm} :\ \code{Expr}\ \Gamma\ x\ \rightarrow\ (x'\ \code{**}\ (\code{Expr}\ \Gamma\ x',\ x\ \simeq\ x'))$ \\
+Every function of normalisation \code{norm} will produce a dependent pair : the new concrete value $x'$, and a pair made of an $\code{Expr}\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
 
-This function doesn't simply produce an $Expr\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
+This function doesn't simply produce an $\code{Expr}\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
 
@@
 The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a group.
-The reflected term for $x$ will be denoted $e_x$, and this term will have the type $ExprG\ p\ \Gamma\ x$, which means that $e_x$ is guaranteed to be the encoding of $x$ because $e_x$ is precisely indexed over the real value $x$.
+The reflected term for $x$ will be denoted $e_x$, and this term will have the type $\code{ExprG}\ p\ \Gamma\ x$, which means that $e_x$ is guaranteed to be the encoding of $x$ because $e_x$ is precisely indexed over the real value $x$.
 We've got similar things for $y$, which is encoded by $e_y$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $e_x$ will produce the normal form $e_{x'}$ of type $ExprG\ p\ \Gamma\ x'$ and a proof $p_x$ of $x \simeq x'$ while running the normalisation procedure on $e_y$ will produce the normal form $e_{y'}$ of type $ExprG\ p\ \Gamma\ y'$ and a proof $p_y$ of $y \simeq y'$.
-It is now enough to simply compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test $exprG\_eq$ because these two expressions are now supposed to be in normal form.
+Running the normalisation procedure on $e_x$ will produce the normal form $e_{x'}$ of type $\code{ExprG}\ p\ \Gamma\ x'$ and a proof $p_x$ of $x \simeq x'$ while running the normalisation procedure on $e_y$ will produce the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof $p_y$ of $y \simeq y'$.
+It is now enough to simply compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test \code{exprG\_eq} because these two expressions are now supposed to be in normal form.
 

### 2017-03-30 paper-CICM-17/paper.tex
paper CICM-17 : fixed subtitle
some fixes

@@
 
+\def\code#1{{\mbox{\texttt{#1}}}} % For short element of code in the test that should not be hyphenated
+\def\codeL#1{{\texttt{#1}}} % For Long element of code in the text that might need to be hyphenated
+
 \newcommand{\figrule}{\begin{center}\hrule\end{center}}
@@
 \title{Automatically Proving Equivalence by Type-Safe Reflection}
-\subtitle{Or how to use dependent types to solve some of the problems they have caused}
+\subtitle{Or how to use dependent types to address some of the problems they have caused}
 %

### 2017-03-30 paper-CICM-17/content/1-introduction.tex
more fixes

@@
 
-\item We develop a hierarchy of tactics where each tactic reuses the
-rewriting machinery of the structure from which it naturally inherits.
+\item We develop a hierarchy of tactics (and not just a tactic for, say, Ring) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
 For example, simplifying neutral elements is only implemented at the monoid level, and each level inheriting from it will reuse it.

### 2017-03-30 paper-CICM-17/content/3-hierarchyOfTactics.tex
more fixes

@@
 Now, if the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$, which implies $x' \simeq y'$ because the propositional equality is more restrictive than an equivalence relation.
-By using the two equivalences $x \simeq x'$ and $y \simeq y'$ that we obtained during the normalisations in the new equivalence $x' \simeq y'$ with the use of $eq\_preserves\_eq$, we can get a proof of $x \simeq y$.
-This is what the function $buildProof$ is doing :
+By using the two equivalences $x \simeq x'$ and $y \simeq y'$ that we obtained during the normalisations in the new equivalence $x' \simeq y'$ with the use of \code{eq\_preserves\_eq}, we can get a proof of $x \simeq y$.
+This is what the function \code{buildProof} is doing :
 
@@
 
-The argument of type $ExprG\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
+The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
 Before the normalisation, the reflected LHS was reflecting the value $x$.
@@
 
-Finally, the main function which tries to prove the equality $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function $buildProof$ in order to compose the two proofs that have been obtained :
+Finally, the main function which tries to prove the equality $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
 
@@
 
-The remaining part is to define the function $norm$.
+The remaining part is to define the function \code{norm}.
 Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
 The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities.
-Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
-These normalisation functions will compute the canonical representation of the reflected input term, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the original real value (indexing the input term) and the produced real value (indexing the produced term).
+Without this property, it would become a lot more complicated to decide the equality without brute-forcing a series of rewritings, that would have no guarantee to terminate.
+As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the initial real value (indexing the input term) and the final real value (indexing the produced term).
 Thus, the type of the reduction function for group is :
@@
 
-The functions of normalisation have more work to do in structures with multiple axioms (like Group and Ring), than for the easier structure (like Semi-Group and Monoid).
-The details of these functions will be given in the next section.
+The function of normalisation has more work to do for structures with many axioms (Commutative-Monoid, Group, Commutative-Group, Semi-Ring and Ring), than for the easier structure (Semi-Group and Monoid).
+The operation performed by these normalisation functions will be detailed in the next subsection.
 
@@
 Therefore, after the full development, the polynomial is a sum of \emph{product of monomials}, and not directly a sum of monomials.
-The only rearrangement that can be done with the multiplication is to check if two constants are consecutive in a product, and if so, to replace them by the constant that represents the product of them.
+The only rearrangement that can be done towards the multiplication is to check if two constants are consecutive in a product, and if so, to replace them by the constant that represents the product of them.
 
 However, because $+$ is always a commutative operator in a ring, the different products of monomials themselves can be rearranged in different ways in this sum.
-That will be done at the level of the commutative group prover if we can provide an ordering between products of monomials.
+That will be done at the level of the commutative group prover if we can provide an ordering on products of monomials.
 
-The ordering on product of monomials will be defined by using an ordering on monomials, that we will call $isBefore\_mon$, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
+This ordering on product of monomials will be defined by using an ordering on monomials, that we will call \code{isBefore\_mon}, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
 \[
@@
 We can now build the order on \emph{product of monomials} that we needed.
-We name it $isBefore$.
+We name it \code{isBefore}.
 Given two product of monomials $Prod_{i}$ and $Prod_{i'}$ we need to decide which one comes first.
-We will obviously use the order $isBefore\_mon$ on the first monomials of these two products.
+We will obviously use the order \code{isBefore\_mon} on the first monomials of these two products.
 If it says that $Monomial_{i}^{1}$ comes before $Monomial_{i'}^{1}$, then we decide that $Prod_{i}$ comes before $Prod_{i'}$.
 Conversely, if that's $Monomial_{i'}^{1}$ that comes first, then we decide that $Prod_{i'}$ comes first.
-However, if $Monomial_{i}^{1}$ and $Monomial_{i'}^{1}$ have exactly the same position in the order, then we continue by inspecting the remaining monomials with a recursive call on $isBefore$.
+However, if $Monomial_{i}^{1}$ and $Monomial_{i'}^{1}$ have exactly the same position in the order, then we continue by inspecting the remaining monomials with a recursive call.
 As previously, if we can't continue because one of the two products has less monomials than the other, then the one with fewer monomials will come first.
@@
 \item We always simplify as much as possible with constants.
-That includes doing the simplification for the addition with zero and for the multiplication with one, doing the computation between two constants that are nearby in a sum or in a product, etc...
-\item We always replace the binary operator $Minus$ by a sum and a $Neg$ (following $a-b=a+(-b)$) since the binary operator $Minus$ are just allowed for the user's convenience, but internally we simplify them all.
+That includes performing the simplification for the addition with zero and for the multiplication with zero and one, doing the computation between two constants that are nearby in a sum or in a product, etc...
+\item We always replace the binary operator \code{Minus} by a sum and a \code{Neg} (following $a-b=a+(-b)$) since the binary operator \code{Minus} is just allowed for the user's convenience, but internally we simplify them all.
 \item We always simplify the sum of an expression $e$ and its inverse $-e$ into the constant zero.
@@
 
-The normalisation function of each structure will use as much as possible the normalisation function of the structure from which it inherits.
-For example, the normalisation on monoids will reuse the normalisation on semi-groups so that it won't have to deal with the associativity.
-For this case, that's not a problem since the datatype reflecting terms in semi-groups has the same expressivity power than the one for monoids.
-However, that's not the case with groups and monoids: if the function of normalisation for group needs to call the normalisation for monoid, it will have to encode the negations that can't be directly expressed in a monoid.
+One of the novelty of our work is that we haven't built just a prover for a specific algebraic structure, but we have instead built a hierarchy of provers for many algebraic structures.
+Morover, each prover is not isolated from the others, and the various simplifications are not duplicated at the different levels : the normalisation function of each structure uses as much as possible the normalisation function of the structure from which it inherits.
+For example, the normalisation on monoids reuses the normalisation on semi-groups in such a way that it does not have to deal with the associativity.
+For this case, it's not a problem since the datatype reflecting terms in semi-groups has the same expressivity power than the one for monoids.
+However, that's not the case with groups and monoids: the function of normalisation for groups that will need to call the normalisation for monoids will have to encode the negations, because they can't be directly expressed in a monoid.
 For this reason, we will develop some specific encodings.
@@
 We only need to encode negations of variables, as the negation of a constant can be simplified into a constant.
-Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in group will have first systematically propagated $Neg$ inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
+Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in group has first systematically propagated \code{Neg} inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
 
-The constructor for variables now takes a Variable as parameter, instead of taking directly an element of $(Fin\ n)$ :
+All the constructors for variables now take a \code{Variable} as parameter, instead of taking directly an element of $(\code{Fin}\ n)$. That gives the following, for groups :
 
@@
 
-Thanks to this encoding, we can now transform an $ExprG$ to an $ExprMo$.
-A constant $(ConstG\ p\ \Gamma\ c1)$ will be transformed into the corresponding constant $(ConstMo\ p\ \Gamma\ c1)$, a $PlusG$ into the corresponding $PlusMo$, a real variable into the same real variable, the negation of a constant into the resulting constant, and finally the negation of a variable $i$ into a $(VarMo\ p\ (EncodingNegOfVar\ \Gamma\ i))$.
+Thanks to this encoding, we can now transform an \code{ExprG} from the group level to an \code{ExprMo} at the monoid level.
+A constant $(\code{ConstG}\ p\ \Gamma\ c1)$ will be transformed into the corresponding constant $(\code{ConstMo}\ p\ \Gamma\ c1)$, a \code{PlusG} into the corresponding \code{PlusMo}, a real variable into the same real variable, the negation of a constant into the resulting constant, and finally the negation of a variable $i$ into a $(\code{VarMo}\ p\ (\code{EncodingNegOfVar}\ \Gamma\ i))$.
 
-We will use a similar technique for converting an expression from the ring level to the commutative-group level, where we will encode the product of monomials because the product is not an operation define at the commutative-group level.
-That will enable the function of normalisation for ring to benefit from the normalisation function for commutative-group.
+We've used a similar technique for converting an expression from the ring level to the commutative-group level, where we have encoded the product of monomials, because the product is not an operation defined at the commutative-group level.
+That enables the function of normalisation for ring to benefit from the normalisation function for commutative-group.
 

### 2017-03-30 paper-CICM-17/content/4-relatedWork.tex
more fixes

@@
 As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism, and without the correct-by-construction approach : they have first defined the normalisation of terms, and afterwards they've proven the correctness of the normalisation with the external lemma :
-$\forall\ (e1\ e2\ :\ Expr),\ beq_{Expr}\ (norm\ e1)\ (norm\ e2)\ =\ true\ \rightarrow\ reify\ e1\ \simeq\ reify\ e2$ which therefore needs a $reify$ function that we do not need in our development.
+$\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\ (\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\ \code{reify}\ e2$ which therefore needs a \code{reify} function that we do not need in our development.
 Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on syntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
@@
 
-Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there's usually no normal form.
+Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there is usually no normal form.
 This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts.

### 2017-03-30 paper-CICM-17/content/5-conclusions.tex
more fixes

@@
 Instead, in our approach, the construction of the proof is guided by the construction of the normalised terms, which, in addition of avoiding redundancy, simplifies the proof generation considerably.
-This development shows that if dependent types effectively bring some new problems, especially the index mismatch, they are also very expressive tools for building correct-by-construction software where the development is driven by the types.
+This development shows that if dependent types effectively bring some new problems, they are also very expressive tools for building correct-by-construction software where the development is \emph{driven by the types}.
 \\
@@
 We want to refactor the Semi-Ring level with the creation of the intermediate structure of Pre Semi-Ring, that wouldn't have the property that $0$ is an annihilator element for the product.
-Then, we could build a specific prover for regular expressions, that would use the normalisation function of the Pre Semi-Ring level and that would deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new star operation.
+Then, we could build a specific prover for regular expressions, that would use the normalisation function of the Pre Semi-Ring level and that would deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
again, more fixes

@@
 Obviously, there is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
-Therefore, equivalences in a \code{Set} are just the "syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
+Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
 But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}.
@@
 Working with equivalences instead of equalities will bring one complication.
-These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use idris' "rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal.
-More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use "rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$.
+These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use Idris' ``rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal.
+More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$.
 This is a classical problem of working within a setoid, and this problem can be slightly mitigated when the programming language offers a good support for rewriting terms in setoids.
@@
 This lemma will be intensively used for building the proofs of correctness of the normalisation functions that will help to build the desired proof of $x \simeq y$.
-It will be basically used as a replacement to the "rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.
+It will be basically used as a replacement to the ``rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.
 
@@
 This code means that a type $c$ (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be "compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
+In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.
@@
 When trying to prove an equivalence $x \simeq y$, the universally-quantified variables have been abstracted, and they are now part of the context.
-Our tatics will work by normalising both sides of the "potential equivalence" $x \simeq y$, and afterwards comparing them with a syntactic equality test.
+Our tatics will work by normalising both sides of the ``potential equivalence" $x \simeq y$, and afterwards comparing them with a syntactic equality test.
 The problem is that such a normalisation function will need to do different treatments for a a variable, for a constant and for an operator.
@@
 In the case where both monomials start with the same variable, we continue by inspecting the remaining variables.
-If we can't continue because one of the two monomials has less variables than the other, then we decide that the one with fewer variables comes first (exactly like the word "house" comes before the word "housemate" with the lexicographic order).
+If we can't continue because one of the two monomials has less variables than the other, then we decide that the one with fewer variables comes first (exactly like the word ``house" comes before the word ``housemate" with the lexicographic order).
 

### 2017-03-31 paper-CICM-17/paper.tex
again, more fixes

@@
 For example, implementing an adder for binary numbers indexed over their natural number equivalents will naturally lead to proof obligations for equalities of expressions over universally quantified natural numbers.
-The need of these proofs of equalities comes, in intensional type theories, from the fact that the propositional equality enables to prove as equal terms that are not judgementally equal, which means that the typechecker can't always obtain equalities by reduction, and therefore often needs proofs before accepting definitions.
+The need of these proofs of equalities comes, in intensional type theories, from the fact that the propositional equality enables to prove as equal terms that are not judgmentally equal, which means that the typechecker can't always obtain equalities by reduction, and therefore often needs proofs before accepting definitions.
 As far as possible, we would like to solve such proof obligations automatically.
@@
 We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.
+
 \keywords{proof automation, equivalence, equality, proof by reflection, correct-by-construction software}

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Needs more proofreading and if possible english check by Edwin

@@
 
-What we want to build is a prover able to prover not only equalities for \code{Nat}, \code{List}, or any specific type but for various datatypes, and for various properties.
+What we want to build is a prover able to prove not only equalities for \code{Nat}, \code{List}, or any specific type, but for various datatypes, and for various properties.
 With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for Semi-Groups, Monoids, Commutative-Monoids, Groups, Commutative-Groups, Semi-Rings and Rings.
@@
 We will end up with a hierarchy of interfaces and will write one tactic for each of these interface.
-All the tactics that we will implement (the ring prover, the group prover, etc) will be able to work on any type, being given that an implementation of the corresponding interface is provided.
+At every level of the hierarchy, we will be able to work on any type, being given that an implementation of the corresponding interface is provided.
 
@@
 
-The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equality, which represents the value $x'$.
+The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equivalence, which represents the value $x'$.
 Before the normalisation, the reflected LHS was reflecting the value $x$.
@@
 
-Finally, the main function which tries to prove the equality $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
+Finally, the main function which tries to prove the equivalence $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
 
@@
 The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities.
-Without this property, it would become a lot more complicated to decide the equality without brute-forcing a series of rewritings, that would have no guarantee to terminate.
-As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the initial real value (indexing the input term) and the final real value (indexing the produced term).
+Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
+As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equivalence between the initial real value (indexing the input term) and the final real value (indexing the produced term).
 Thus, the type of the reduction function for group is :

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
again, more fixes
Minor paper edits

@@
 inhabited, then $x$ is said to be provably equal to $y$.
-Thus, $Id$ is a type family (parametrised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
+Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
 For
@@
 This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in idris as :
+If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in Idris as :
 
@@
 
-Prover for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda/\%22?n=Libraries.UsingTheRingSolver}} ring prover.
+Prover for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda\%5C?n=Libraries.UsingTheRingSolver}} ring prover.
 In this paper, we describe a certified
-implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
+implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris/Provers}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
 structures, including Monoid, Groups and Rings (all potentially commutative),
@@
 We will
-compare our approach with other implementations in Section~\ref{sect:relatedWork}.
+compare our approach with other implementations in section~\ref{sect:relatedWork}.
 
@@
 negations ($-x$) and subtractions ($x-y$). Some specific encodings
-will be presented in Section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
+will be presented in section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
 

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Minor paper edits

@@
 
-If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
-Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--.
+If the user wants to prove propositional equalities, they will instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
+Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it produces the value \code{Nothing}--.
 That is normal, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
 
-Obviously, there is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
-Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
+There is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
+Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
 But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}.
@@
 
-$\code{eq\_preserves\_eq} : \{c:Type\} \rightarrow (Set\ c) \rightarrow (x:c) \rightarrow (y:c) \rightarrow (c1:c) \rightarrow (c2:c) \rightarrow (x \simeq c1) \rightarrow (y \simeq c2) \rightarrow (c1 \simeq c2) \rightarrow (x \simeq y)$.
+$\code{eq\_preserves\_eq} : \{c:\code{Type}\} \rightarrow \{\code{Set}\ c\} \rightarrow (x:c) \rightarrow (y:c) \rightarrow (c1:c) \rightarrow (c2:c) \rightarrow (x \simeq c1) \rightarrow (y \simeq c2) \rightarrow (c1 \simeq c2) \rightarrow (x \simeq y)$.
 
@@
 
-This code means that a type $c$ (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
+This code means that a type \code{c} (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
+In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.
@@
 
-As example of magma, we can cite \code{Nat} equipped with its operation of addition and \code{List} and the operation of concatenation.
+As example of magma, we can cite \code{Nat} equipped with its operation of addition and \code{List} with the operation of concatenation.
 
@@
 For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
-When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is simply the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
+When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
 The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$.
@@
 There is no additional operations in SemiGroup and Monoid, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for Magma that we have given above.
-However, the \code{Group} interface will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
+However, the datatype for reflected terms in Groups will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
 
@@
 
-This function doesn't simply produce an $\code{Expr}\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
+This function doesn't only produce an $\code{Expr}\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
 
@@
 Running the normalisation procedure on $e_x$ will produce the normal form $e_{x'}$ of type $\code{ExprG}\ p\ \Gamma\ x'$ and a proof $p_x$ of $x \simeq x'$ while running the normalisation procedure on $e_y$ will produce the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof $p_y$ of $y \simeq y'$.
-It is now enough to simply compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test \code{exprG\_eq} because these two expressions are now supposed to be in normal form.
+It is now enough to compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test \code{exprG\_eq} because these two expressions are now supposed to be in normal form.
 
@@
 buildProofGroup p ex' ey' px py with (exprG_eq p ex' ey')
-    buildProofGroup p ex' ex' px py | Just refl = ?MbuildProofGroup
+    buildProofGroup p ex' ex' px py | Just refl = 
+                Just(eq_preserves_eq x y c1 c2 lp rp e1_equiv_e2)
     buildProofGroup p ex' ey' px py | Nothing = Nothing
@@
 
-% TALK ABOUT THE METAVARIABLE MbuildProodGroup to add here !
-
-Finally, the main function which tries to prove the equivalence $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
+Finally, the main function which tries to prove the equivalence $x \simeq y$ has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
 
@@
 Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
-As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equivalence between the initial real value (indexing the input term) and the final real value (indexing the produced term).
+As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term).
 Thus, the type of the reduction function for group is :
@@
 
-We've built an automatic reflection mechanism, in order to let the machine build automatically the reflected terms for us.
+We've built an automatic reflection mechanism which enables to let the machine build automatically the reflected terms for us.
 This is not something absolutely vital, as it only simplifies the usage of the tactics, avoiding to write long encodings by hand.

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
More small edits

@@
 
-When reading these definitions, one could wrongly think that the propositional equality is a simple wrapper of the judgemental equality.
+When reading these definitions, one could think that the propositional equality is a wrapper of the judgemental equality.
 This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
@@
 
-We revisit an example from a previous work of Edwin Brady~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
-In this development, the goal is to implement a formally verified library of binary numbers.
+We revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
+The goal is to implement a formally verified library of binary numbers.
 In order to ensure the correctness of the methods that will be developed, we define the types \code{Bit} and \code{Binary} as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
CICM-17 : final fixes

--- 2017-03-31 paper-CICM-17/content/1-introduction.tex
@@
 inhabited, then $x$ is said to be provably equal to $y$.
-Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
+Thus, $Id$ is a type family (parametrised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
 For
@@
 We will
-compare our approach with other implementations in section~\ref{sect:relatedWork}.
+compare our approach with other implementations in Section~\ref{sect:relatedWork}.
 
@@
 negations ($-x$) and subtractions ($x-y$). Some specific encodings
-will be presented in section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
+will be presented in Section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
 

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM-17 : final fixes

--- 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
@@
 The function also expects proofs of $x \simeq x'$ and of $y \simeq y'$, that we will be able to provide because the normalisation function has produced them while it was building the normalised reflected terms.
-
-% TALK ABOUT THE METAVARIABLE MbuildProodGroup to add here !
 

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM-17 : again, final fixes

@@
 This code means that a type $c$ (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
+In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM-17 : again2, final fixes

@@
 
-This code means that a type $c$ (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
+This code means that a type \code{c} (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
 In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\

### 2017-03-31 paper-CICM-17/paper.tex
Needs more proofreading and if possible english check by Edwin
CICM-17 : final fixes
CICM-17 : removed comma

@@
-% This is LLNCS.DEM the demonstration file of
-% the LaTeX macro package from Springer-Verlag
+% Franck Slama's and Edwin Brady's paper for CICM-2017
+% That follows the LLNCS style from Springer-Verlag
 % for Lecture Notes in Computer Science,
@@
 \author{Franck Slama \and Edwin Brady}
-\institute{University of St Andrews, Scotland, UK,\\
+\institute{University of St Andrews, Scotland, UK\\
 \email{fs39@st-andrews.ac.uk, ecb10@st-andrews.ac.uk}}
@@
 
-\keywords{proof automation, equivalence, equality, proof by reflection, correct-by-construction software}
+\keywords{proof automation, equivalence, equality, proof by reflection, correct-by-construction software, type-driven development}
 \end{abstract}

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
CICM-17 : more final fixes

@@
 $\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\ (\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\ \code{reify}\ e2$ which therefore needs a \code{reify} function that we do not need in our development.
+Another difference is that Coq's prover only deals with rings and semi-rings (commutative or not), but not with any of intermediate structure (semi-group, monoid, group, etc), unlike us.
+However, their implementation has better performances due to the use of sparse normal form and more optimised algorithms.
 Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on syntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
More fixes

@@
 As we see, this kind of proof consists of a potentially long sequence of rewriting steps, where each step uses one of the available properties: neutral element, commutativity, associativity.
-Without some specific automation, this sequence of rewriting must be done by the programmer.
+Without some specific automation, this sequence of rewritings must be done by the programmer.
 This is time consuming and more importantly, just a small change in the left or the right hand side of the equality may invalidate the proof.
@@
 implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris/Provers}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
-structures, including Monoid, Groups and Rings (all potentially commutative),
+structures, including monoid, groups and rings (all potentially commutative),
 for the Idris language.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM-17 : automatic reflection
More fixes

@@
 What we want to build is a prover able to prove not only equalities for \code{Nat}, \code{List}, or any specific type, but for various datatypes, and for various properties.
-With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for Semi-Groups, Monoids, Commutative-Monoids, Groups, Commutative-Groups, Semi-Rings and Rings.
+With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for semi-groups, monoids, commutative-monoids, groups, commutative-groups, semi-rings and rings.
 These tactics will be usable on any type that satisfies the properties of one of these algebraic structure.

### 2017-03-31 paper-CICM-17/paper.tex
More fixes

@@
 We will show how type-safe reflection, where reflected terms are indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.
-We build a hierarchy of reflexive tactics for proving equivalences in Semi-Groups, Monoids, Commutative Monoids, Groups, Commutative Groups, Semi-Rings and Rings.
+We build a hierarchy of reflexive tactics for proving equivalences in semi-groups, monoids, commutative monoids, groups, commutative groups, semi-rings and rings.
 We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
Restored the changes from Edwins two commits

@@
 
-When reading these definitions, one could wrongly think that the propositional equality is a simple wrapper of the judgemental equality.
+When reading these definitions, one could think that the propositional equality is a simple wrapper of the judgemental equality.
 This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
@@
 
-We revisit an example from a previous work of Edwin Brady~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
+We revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
 In this development, the goal is to implement a formally verified library of binary numbers.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Restored the changes from Edwins two commits

@@
 
-If the user wants to prove propositional equalities, then he will simply instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
+If the user wants to prove propositional equalities, they will instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
 Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--.
@@
 
-Obviously, there is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
+There is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
 Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
@@
 
-Finally, the main function which tries to prove the equivalence $x \simeq y$ simply has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
+Finally, the main function which tries to prove the equivalence $x \simeq y$ has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
 

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
Agaaaaain more fixes

@@
 We know that the definition of \code{adc} we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and Idris can not figure out a proof of compatibility on its own.
-If we use provisional definitions[ref], the previous definition can be accepted by the system, but we now have two proof obligations \code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
+If we use provisional definitions~\cite{DBLP:conf/plpv/Brady13}, the previous definition can be accepted by the system, but we now have two proof obligations \code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
 

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
Agaaaaain2 more fixes

@@
 Proof automations have been investigated for various proof assistants and programming languages.
-In Coq, appart from the $ring$ prover~\cite{Coq2005}, there's also the $omega$ solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
-a universally quantified formula made of equations and inequations), and a Field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
+In Coq, appart from the ring prover~\cite{Coq2005}, there's also the omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
+a universally quantified formula made of equations and inequations), and a field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
 Various automations have also been done in Agda~\cite{DBLP:conf/mpc/KokkeS15} ~\cite{Lindblad04}.
@@
 \\
-
-We also need to mention $Mtac$~\cite{Ziliani13}, which is an extension to Coq that supports dependently-typed tactic programming that uses monad to avoid the need to touch the trusted kernel typechecker of Coq.

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
Minor paper edits
Agaaaaain2 more fixes

@@
-\section{Results, conclusions and futur work}
+\section{Results, conclusions and future work}
 
@@
 This solution takes the form of a hierarchy of provers for equivalence in algebraic structures.
-These provers are generic in multiple ways: they work for many algebraic structures (Semi-Group, Monoid, Commutative Monoid, Group, Commutative Group, Ring and Semi-Ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality).
+These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality).
 The implementation is modular and each prover reuses the prover of the structure from which it directly inherits.
@@
 The work that we have presented here can be extended to build new provers for less common algebraic structures and for more specific structures.
-For example, regular expressions form a Pre Semi-Ring with some extra axioms.
-We want to refactor the Semi-Ring level with the creation of the intermediate structure of Pre Semi-Ring, that wouldn't have the property that $0$ is an annihilator element for the product.
-Then, we could build a specific prover for regular expressions, that would use the normalisation function of the Pre Semi-Ring level and that would deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.
+For example, regular expressions form a ``pre semi-ring" with some extra axioms.
+We want to refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that wouldn't have the property that $0$ is an annihilator element for the product.
+Then, we could build a specific prover for regular expressions, that would use the normalisation function of the pre semi-ring level and that would only have to deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
Agaaaaain3 more fixes

@@
 
-\item We develop a hierarchy of tactics (and not just a tactic for, say, Ring) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
+\item We develop a hierarchy of tactics (and not just a tactic for, say, ring) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
 For example, simplifying neutral elements is only implemented at the monoid level, and each level inheriting from it will reuse it.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Agaaaaain3 more fixes

@@
 
-The Monoid structure is a Semi-Group with the property of neutral element for a distinguished element called \code{Zero}.
+The monoid structure is a semi-group with the property of neutral element for a distinguished element called \code{Zero}.
 
@@
 
-The hierarchy of interfaces continues with Commutative-Monoid, Group, Commutative Group, Semi-Ring and Ring, but we only show the \code{Group} interface for reasons of space :
+The hierarchy of interfaces continues with commutative-monoids, groups, commutative groups, semi-rings and rings, but we only show the \code{Group} interface for reasons of space :
 
@@
 
-A magma is only equipped with the operation of addition. Thus, we only have three concepts to express in order to reflect terms in a Magma : constants, variables, and additions.
+A magma is only equipped with the operation of addition. Thus, we only have three concepts to express in order to reflect terms in a magma : constants, variables, and additions.
 
@@
 
-There is no additional operations in SemiGroup and Monoid, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for Magma that we have given above.
-However, the datatype for reflected terms in Groups will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
+There is no additional operations in semi-groups and monoids, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for magma that we have given above.
+However, the datatype for reflected terms in groups will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
 
@@
 
-We will explain how the tactic works by commenting the construction of the tactic for Group, but the other algebraic structures will be equipped similarly.
+We will explain how the tactic works by commenting the construction of the tactic for groups, but the other algebraic structures will be equipped similarly.
 The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a group.
@@
 
-The function of normalisation has more work to do for structures with many axioms (Commutative-Monoid, Group, Commutative-Group, Semi-Ring and Ring), than for the easier structure (Semi-Group and Monoid).
-The operation performed by these normalisation functions will be detailed in the next subsection.
+The function of normalisation has more work to do for structures with many axioms (commutative-monoids, groups, commutative-groups, semi-rings and rings), than for the easier structure (semi-groups and monoids).
+The operations performed by these normalisation functions will be detailed in the next subsection.
 

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
AgaaaaainX more fixes

@@
 
-With some reasonable additional effort, we could produce a collection of tactics for proving equivalences instead of only equality.
+With some reasonable additional effort, we could produce a collection of tactics for proving equivalences instead of only equalities.
 The machinery would be very similar, and if we do the right abstraction from the start, we could gain one more degree of genericity, with the freedom of choosing the equivalence relation (which can of course be the usual equality).
@@
 But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}.
-Therefore, \code{refl} of the interface \code{Set} is a generalisation of the most specialised \code{Refl} for \code{=}}.
+Therefore, \code{refl} of the interface \code{Set} is a generalisation of \code{Refl}}.
 
@@
 This code means that a type \code{c} (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply for all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
+In fact, there is an additional requirement that will apply to all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
 This requirement comes from the fact that we're dealing with any equivalence relation.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Just a few things to remove to make it 15 pages

@@
 \begin{lstlisting}
-groupNormalise : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (p:Group c) 	   	   	   
-	   $\rightarrow$ {$\Gamma$:Vect n c} $\rightarrow$ {x:c} $\rightarrow$ (ExprG p $\Gamma$ x) 
-	   $\rightarrow$ (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
+groupNormalise : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (p:Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+       $\rightarrow$ {x:c} $\rightarrow$ (ExprG p $\Gamma$ x) $\rightarrow$ (x' ** (ExprG p $\Gamma$ x', x $\simeq$ x'))
 \end{lstlisting}
@@
 The function of normalisation has more work to do for structures with many axioms (commutative-monoids, groups, commutative-groups, semi-rings and rings), than for the easier structure (semi-groups and monoids).
-The operations performed by these normalisation functions will be detailed in the next subsection.
+The operations performed by the normalisation functions are detailed in the next section.
 
@@
 
-For the sake of completeness, we will describe the normal form for the ring structure, as it has all the operations and all the properties about these operations.
+For the sake of completeness, we describe the normal form for rings which is our most sophisticated structure.
 The normalisation function takes in input an expression expressed with sums, products, constants (zero, one...) and variables that belong to an ordered set $\mathcal{V}$ of variables.
@@
 However, because $+$ is always a commutative operator in a ring, the different products of monomials themselves can be rearranged in different ways in this sum.
-That will be done at the level of the commutative group prover if we can provide an ordering on products of monomials.
-
-This ordering on product of monomials will be defined by using an ordering on monomials, that we will call \code{isBefore\_mon}, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
+That will be done at the underneath level for commutative groups if we can provide an ordering on products of monomials.
+This ordering will be defined by using an ordering on monomials, that we will call \code{isBefore\_mon}, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
 \[
@@
 
-\item The top-level sum of the polynomial will be written on the completely right-associative form : \\
+\item The top-level sum of the polynomial will be in right-associative form : \\
 $prodMon_1\ +\ (prodMon_2\ + (prodMon_3\ +\ (...\ +\ prodMon_a)))$
-\item The two levels of product that we have, namely the products of monomials (at the main level) and the products of variables (at the monomial level), will also be written on the completely right-associative form.
+\item All the products that we have at the two levels, namely the products of monomials (at the main level) and the products of variables (at the monomial level), will also be written on the completely right-associative form.
 \item We always simplify as much as possible with constants.
-That includes performing the simplification for the addition with zero and for the multiplication with zero and one, doing the computation between two constants that are nearby in a sum or in a product, etc...
-\item We always replace the binary operator \code{Minus} by a sum and a \code{Neg} (following $a-b=a+(-b)$) since the binary operator \code{Minus} is just allowed for the user's convenience, but internally we simplify them all.
+That includes simplifying the addition with zero and the multiplication with the constants zero and one, doing the computations between two nearby constants, etc...
 \item We always simplify the sum of an expression $e$ and its inverse $-e$ into the constant zero.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
almost ready

@@
 
-The hierarchy of interfaces continues with commutative-monoids, groups, commutative groups, semi-rings and rings, but we only show the \code{Group} interface for reasons of space :
+The hierarchy of interfaces continues, but we only show the \code{Group} interface for reasons of space :
 
@@
 
-\item The top-level sum of the polynomial will be in right-associative form : \\
+\item The top-level sum of the polynomial will be put in right-associative form : \\
 $prodMon_1\ +\ (prodMon_2\ + (prodMon_3\ +\ (...\ +\ prodMon_a)))$
-\item All the products that we have at the two levels, namely the products of monomials (at the main level) and the products of variables (at the monomial level), will also be written on the completely right-associative form.
+\item All the products that we have (the products of monomials and the products of variables), will also be written on the completely right-associative form.
 \item We always simplify as much as possible with constants.
@@
 We've used a similar technique for converting an expression from the ring level to the commutative-group level, where we have encoded the product of monomials, because the product is not an operation defined at the commutative-group level.
-That enables the function of normalisation for ring to benefit from the normalisation function for commutative-group.
+That enables the function of normalisation for rings to benefit from the normalisation function for commutative-groups.
 

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
almost ready

@@
 $\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\ (\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\ \code{reify}\ e2$ which therefore needs a \code{reify} function that we do not need in our development.
-Another difference is that Coq's prover only deals with rings and semi-rings (commutative or not), but not with any of intermediate structure (semi-group, monoid, group, etc), unlike us.
+Another difference is that Coq's prover only deals with rings and semi-rings (commutative or not), but not with any of the intermediate structures (semi-group, monoid, group, etc), unlike us.
 However, their implementation has better performances due to the use of sparse normal form and more optimised algorithms.
@@
 
-Doing proofs by reflection has been intensively investigated by many authors, including Adam Chlipala, in particular in his book~\cite{ChlipalaBook} and in~\cite{Malecha14}.
+Doing proofs by reflection has been intensively investigated as in~\cite{ChlipalaBook} and~\cite{Malecha14}.
 However, we haven't found anything similar to the type-safe reflection technique that we've presented in this paper.
-\\
-
-At the connexion bewteen proof automation and proof by reflection, we can cite SSreflect~\cite{GonthierTuto}, which is an extension of Coq which aims to provide additional formalisations and tactics suitable for long mathematical proofs.
 \\

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
almost ready

@@
 The implementation is modular and each prover reuses the prover of the structure from which it directly inherits.
+These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations as \code{adc\_lemma\_2} encountered in~\ref{sect:motivatingExample}.
+Therefore, these provers enable the user to focus on the interesting proofs of the development that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.
+The human time is best used when it is devoted for proving things that can't be automated.
 \\
 
-These provers can automatically prove equivalence between terms, which discharges the user of providing proofs for all the proof obligations that appear naturally when using dependent types.
-Thefore, these provers enable the user to focus on the interesting and main proofs of the development that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.
-We think that the human time is best used when it is devoted for proving things that can't be automated.
-\\
-
-The correct-by-construction method that we have followed has involved the design of a type-safe reflection mechanism where reflected terms are indexed over the concrete expressions, and from which we are able to pull out the proofs easily.
+The correct-by-construction method that we have followed has involved the design of a type-safe reflection mechanism where reflected terms are indexed over the concrete inputs, and from which we are able to pull out the proofs easily.
 The construction of the proof is done step by step, and it follows the construction of the normalised terms.
-Therefore, unlike other implementations --like Coq's implementation of a ring prover-- we do not have the traditional duplication that happens when separating the computational content from the proof of correctness.
+Therefore, unlike Coq's and Agda's ring provers, we do not have the traditional duplication that happens when separating the computational content from the proof of correctness.
 Instead, in our approach, the construction of the proof is guided by the construction of the normalised terms, which, in addition of avoiding redundancy, simplifies the proof generation considerably.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
almost ready to submit the final PDF !

@@
 The machinery would be very similar, and if we do the right abstraction from the start, we could gain one more degree of genericity, with the freedom of choosing the equivalence relation (which can of course be the usual equality).
-We want to give to the user the possibility to define his own equivalence notion, as long as he's able to provide the properties of the algebraic structure he wants to use.
+The user could define his own equivalence notion, as long as he's able to provide the proofs of the properties of the algebraic structure he wants to use.
 Let's call $c$ the \emph{carrier} type, ie, the type on which we want to prove equivalences.
@@
 All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants.
-For this reason, we define a notion of \code{Set}\footnote{This notion of set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation --accompanied with the proofs that this is indeed an equivalence relation-- and this equivalence test \code{set\_eq}.
+For this reason, we define a notion of \code{Set}\footnote{This notion of set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation and this equivalence test \code{set\_eq}.
 All the interfaces representing the algebraic structures will later extend \code{Set} :
@@
 We only need to encode negations of variables, as the negation of a constant can be simplified into a constant.
-Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in group has first systematically propagated \code{Neg} inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
+Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in groups has first systematically propagated \code{Neg} inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
 

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
almost ready to submit the final PDF !

@@
 
-As mentioned, Coq's implementation of a ring prover was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism, and without the correct-by-construction approach : they have first defined the normalisation of terms, and afterwards they've proven the correctness of the normalisation with the external lemma :
+As mentioned, Coq's implementation of a ring prover~\cite{Coq2005} was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism, and without the correct-by-construction approach : they have first defined the normalisation of terms, and afterwards they've proven the correctness of the normalisation with the external lemma :
 $\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\ (\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\ \code{reify}\ e2$ which therefore needs a \code{reify} function that we do not need in our development.
@@
 Proof automations have been investigated for various proof assistants and programming languages.
-In Coq, appart from the ring prover~\cite{Coq2005}, there's also the omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
+In Coq, appart from the ring prover, there's also the omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
 a universally quantified formula made of equations and inequations), and a field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
@@
 
-Doing proofs by reflection has been intensively investigated as in~\cite{ChlipalaBook} and~\cite{Malecha14}.
-However, we haven't found anything similar to the type-safe reflection technique that we've presented in this paper.
-\\
+Proofs by reflection has been intensively studied as in~\cite{ChlipalaBook} and~\cite{Malecha14}, but without anything similar to the type-safe reflection that we have presented here. \\
 

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
almost ready to submit the final PDF !

@@
 
-We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were defying functions that use dependent types.
+We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were using indexed types.
 This solution takes the form of a hierarchy of provers for equivalence in algebraic structures.
@@
 The implementation is modular and each prover reuses the prover of the structure from which it directly inherits.
-These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations as \code{adc\_lemma\_2} encountered in~\ref{sect:motivatingExample}.
-Therefore, these provers enable the user to focus on the interesting proofs of the development that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.
-The human time is best used when it is devoted for proving things that can't be automated.
-\\
+These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations as \code{adc\_lemma\_2}.
+Therefore, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.\\
 
@@
 
-The work that we have presented here can be extended to build new provers for less common algebraic structures and for more specific structures.
+This work can be extended to build new provers for less common algebraic structures and for more specific structures.
 For example, regular expressions form a ``pre semi-ring" with some extra axioms.
-We want to refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that wouldn't have the property that $0$ is an annihilator element for the product.
+We will refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that will not have the property that $0$ is an annihilator element for the product.
 Then, we could build a specific prover for regular expressions, that would use the normalisation function of the pre semi-ring level and that would only have to deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
Certainly the final version

@@
 More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$.
-This is a classical problem of working within a setoid, and this problem can be slightly mitigated when the programming language offers a good support for rewriting terms in setoids.
-However, Idris isn't equipped with any automation for setoids, and everything will have to be done by hand.
-For this purpose, we will use the following lemma :
+This is a classical problem of working within a setoid, which is slightly mitigated when the programming language offers a good support for rewriting terms in setoids.
+However, Idris isn't equipped with any support for setoids, and everything will have to be done by hand.
+For this reason, we define the lemma :
 
@@
 Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
-As said before, these normalisation functions will compute the canonical representation of the reflected input term by rewriting the given term multiple times, and at the same time, it will produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term).
+As said before, the normalisation functions also produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term).
 Thus, the type of the reduction function for group is :

### 2017-03-31 paper-CICM-17/content/4-relatedWork.tex
Certainly the final version

@@
 
-Finally, when we leave the ground of nice mathematical structures like groups, rings and fields, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such problems, because there is usually no normal form.
-This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and heuristics for proof search starts.
+Finally, if we leave the ground of nice mathematical structures, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such systems, because there is usually no normal form.
+This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and proof search heuristics starts.
 \\

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
Certainly the final version

@@
 These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality).
-The implementation is modular and each prover reuses the prover of the structure from which it directly inherits.
-These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations as \code{adc\_lemma\_2}.
-Therefore, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time and energy for proving routine lemmas that can be automated.\\
+The implementation is modular and each prover reuses the prover of the structure from which it inherits.
+These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations, as \code{adc\_lemma\_2}.
+Thus, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time for proving routine lemmas that can be automated.\\
 

### 2017-03-31 paper-CICM-17/content/1-introduction.tex
CICM-17 : Final version

@@
 inhabited, then $x$ is said to be provably equal to $y$.
-Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $Id (a:Type)\ :\ a \rightarrow a \rightarrow Type$.
+Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $\code{Id} (a:\code{Type}):a \rightarrow a \rightarrow \code{Type}$.
 For
-convenience, we write $(Id_a\ x\ y)$ as $(x\ =_a\ y)$.
+convenience, we write $(\code{Id}_a\ x\ y)$ as $(x\ =_a\ y)$.
 
@@
 
-The only way for $(Id_a\ x\ y)$ to be inhabited is therefore that $x$ should be
+The only way for $(\code{Id}_a\ x\ y)$ to be inhabited is therefore that $x$ should be
 (by definition) $y$. In this case, the constructor \code{Refl} helps to create a
@@
 This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-If a type $T$ is an inductive type with a constant constructor and a recursive constructor, ie, $T = 1 + T$, defined in Idris as :
+If \code{T} is an inductive type with a constant constructor and a recursive constructor, ie, \code{T = 1 + T}, defined in Idris as :
 
@@
 
-then we have the following induction principle for $T$: \\
-$T\_ind : \forall P:T \rightarrow Type,\ (P\ T0) \rightarrow (\forall t:T,\ P\ t \rightarrow P\ (T1\ t)) \rightarrow (\forall t:T,\ P\ t)$.
+then we have the following induction principle for \code{T}: \\
+$\code{T\_ind} : \forall P:\code{T} \rightarrow \code{Type},\ (P\ \code{T0}) \rightarrow (\forall t:\code{T},\ P\ t \rightarrow P\ (\code{T1}\ t)) \rightarrow (\forall t:\code{T},\ P\ t)$.
 

### 2017-03-31 paper-CICM-17/content/3-hierarchyOfTactics.tex
CICM-17 : Final version

@@
 The user could define his own equivalence notion, as long as he's able to provide the proofs of the properties of the algebraic structure he wants to use.
-Let's call $c$ the \emph{carrier} type, ie, the type on which we want to prove equivalences.
-The equivalence relation on $c$ has the following profile $(\simeq) : \code{c} \rightarrow \code{c} \rightarrow \code{Type}$\footnote{This \code{Type} would be a \code{Prop} in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+Let's call \code{c} the \emph{carrier} type, ie, the type on which we want to prove equivalences.
+The equivalence relation on \code{c} has the following profile $(\simeq) : \code{c} \rightarrow \code{c} \rightarrow \code{Type}$\footnote{This \code{Type} would be a \code{Prop} in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
 
@@
 
-For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the number $x$ in the context $\Gamma$.
-When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted number is the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
+For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the value $x$ in the context $\Gamma$.
+When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted value is the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
 The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$.
@@
 $\code{norm} :\ \code{Expr}\ \Gamma\ x\ \rightarrow\ (x'\ \code{**}\ (\code{Expr}\ \Gamma\ x',\ x\ \simeq\ x'))$ \\
-Every function of normalisation \code{norm} will produce a dependent pair : the new concrete value $x'$, and a pair made of an $\code{Expr}\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete number we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
+Every function of normalisation \code{norm} will produce a dependent pair : the new concrete value $x'$, and a pair made of an $\code{Expr}\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete value we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
 

### 2017-03-31 paper-CICM-17/content/5-conclusions.tex
CICM-17 : Final version

@@
 We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were using indexed types.
-This solution takes the form of a hierarchy of provers for equivalence in algebraic structures.
+This solution takes the form of a hierarchy of provers for equivalences in algebraic structures.
 These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality).
@@
 For example, regular expressions form a ``pre semi-ring" with some extra axioms.
-We will refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that will not have the property that $0$ is an annihilator element for the product.
+We will refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that will not necessary have the property that $0$ is an annihilator element for the product.
 Then, we could build a specific prover for regular expressions, that would use the normalisation function of the pre semi-ring level and that would only have to deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.

### 2017-05-02 paper-CICM-17/rebuttal_reviews_Brady_Slama_CICM17.txt
starting the rebuttal

> ----------------------- REVIEW 1 ---------------------
> PAPER: 42
> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
> AUTHORS: Franck Slama and Edwin Brady
> 
> Overall evaluation: 2 (accept)
> 
> ----------- Overall evaluation -----------
> --- Summary ---
> 
> The article presents a technique to automate proofs of equivalence over some algebraic structures, in the context of proof assistants based on dependent types, in particular Idris.
> The motivation behind this work is to relieve users of the burden of writing trivial but tedious proofs, in particular proofs of type equality that are required for some definitions over dependent types.
> The proofs are done by rewriting terms to a canonical form. The rewriting functions work on reflected terms. The main novel idea of the paper is the use of dependent types to index the reflected terms with the concrete term that they represent. This enforces a "correct-by-construction" approach, where the normalizing functions must also provide a proof that they preserve equivalence, thus facilitating the rest of the proof.

Agreed.

> --- Content ---
> 
> The use of indexed reflected terms seems to be a good approach. The modularity of the tactics to work on different structures, as well as the automation within Indris, are nice additions.

Agreed.

> The example in the introduction provides a good motivation, although proofs of type equivalences are not the only use case for this work.

You're entierly right : the motivating example is an illustration of how these proofs obligations can naturally occur, but there are also many other possible utilisations of our proof automation.

> One question that comes to mind is whether it easy to prove properties of the normalization functions (e.g. given a concrete equivalence relation, proving that each equivalence class indeed has a unique canonical representative). Does the use of reflected terms have an impact on such proofs?

Things become very difficult when we leave the ground of the algebraic structures in which we are working (monoids, groups and rings essentially) and move to an arbitrary equivalence relation with arbitrary axioms. Such a system can be seen as a rewriting system of 
dedution modulo. Determining if such a rewriting system admits a normal form requires to determine if the rewriting system formed is confluent and terminating. Both questions are very difficult to answer in the general case. Currently, systems like Dedukti 
(an implementation of lambda-Pi-calculus modulo) can plug to an auxiliary confluence checker such as CSI^ho ( http://cl-informatik.uibk.ac.at/software/csi/ho/ ) or ACPH ( http://coco.nue.riec.tohoku.ac.jp/2015/papers/acph.pdf ).

It is interesting to note that the study of the properties of an alleged normalisation function can only be meta-theoretic : it's not possible to prove it within the proof assistant. Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also notice it for their implementation
of their Ring prover for Coq. It's not an issue as it does not affect the correctness of the decision procedure.

> An interesting question for future work would be to exploit external provers (first-order theorem provers, rewriting engines...) to find equivalences on structures with arbitrary rewriting rules, and later reconstruct the proofs within the proof assistant.

Yes, being able to use an external prover is an interesting future work to investigate, as it would allow to deal with arbitrary rewriting rules.

> --- Presentation ---
> The paper is generally well-written. A lot of effort has been spent on providing motivation and context, and the content is easy to follow even without being an expert in the domain. The various aspects involved are each presented in sufficient detail.

Thanks.

> Subsection 2.1 seems a bit out of order. It essentially amounts to an extended footnote: for the use case described in the introduction, it seems that one really wants to prove type equalities rather than arbitrary equivalences. The extension to arbitrary equivalence relations is interesting, but not central. It is therefore a bit confusing to see that subsection so early in the text, and I would suggest moving it to a later point.

You are entierely right. We will just mention at the start of Section 2 that we want to be general and to work with any equivalence relation (~=), and we will move this section 2.1 just after what is currently section 2.6. 
We still need to say two words about the equivalence relation (~=) at the start of section 2 as we use it everywhere and it needs to be introduced briefly.

> Textual errors found:
> p.11: "admit a canonical [representative]"
> p.15: "deduction modulo and proof search heuristics [start]."

Sorry about them. Thanks for finding them. We will correct.

> ----------------------- REVIEW 2 ---------------------
> PAPER: 42
> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
> AUTHORS: Franck Slama and Edwin Brady
> 
> Overall evaluation: 1 (weak accept)
> 
> ----------- Overall evaluation -----------
> [This is a signed review]
> 
> First two items:
> 1. Paper is definitely in-scope for Calculemus
> 2. As written, it is too long by 2 pages [but see below].
> 
> This paper describes one method (type-safe reflection) to obtain methods of proving
> equivalence which is (somewhat) modular.
> 
> The explanations are clear -- but written in an extremely verbose style.  Cutting
> two pages would be very easy; cutting 3, feasible.
> 
> Saving such space would allow a bit of room to detail some features of Idris a bit
> more for the average CICM reader, such as
> a. unbound variables (like 'a' in the definition of Id being implicitly quantified
>   over Type)
> b. that ** is dependent pairing
> c. why no implicits are used (such as in the signature of addBit)
> 
> The start of section 2 talks about 'natural inheritance' of interfaces, such as that
> between Group and Monoid.  When such algebraic hierarchies are done "naively", yes
> this appears to occur.  But as soon as the hierarchy grows large enough, such
> 'natural' inheritance becomes very brittle.  What replaces it are explicit views, aka
> theory morphisms [see MMT [3] and IMPS].  One can still use transport along such morphisms
> to move results from one place to another.  Some morphisms are still special (the
> display morphisms of the opposite category]; see "Theory Presentation Combinators" [1]
> by Carette & O'Connor for the details. (All papers linked have been published)
> 
> Furthermore, there seems to be an implicit assumption that algebraic theories
> (such as Group) have a canonical presentation -- this is not true.  And that is
> exactly what "Realms" [2] are all about.
> 
> The largest flaw though seems to be the implicit assumption that lots (most? all?)
> interesting structures have normal forms.  And that these can be found using a
> modular algorithm.  Without ever talking about Knuth-Bendix, no less!  If one is going
> to talk about a "correct by construction" approach, then the *definition* of what a
> normal form is (i.e. syntax where intensional equality is a decision procedure for
> equality).  This is implicit everywhere, but never explicit.  And it is not an
> invariant kept by the types!
> 
> For example, the word problem for groups is undecidable.  This implies that there are
> theories with no normal forms.
> 
> In section 2.5, an *order* is snuck in, onto the set of variables.  This is crucial!
> It should not be buried.  Hint: Groebner bases.  Term orders are important.
> But normal forms are extremely sensitive to term orders (as bad as exponential).
> Furthermore, the discussion in this section implicitly assumes associativity.  There
> are many useful, interesting, non-associative algebras...
> 
> Overall, the point is that given a structure X, then the induced language-of-X
> has a nice structure as well.  Most of it is fairly automatically derivable from the
> definition of X.
> 
> Various further comments:
> - a citation in the last paragraph of the introduction ("This is even an everyday
> routine when ...") would be appreciated.  I index types with values all the time,
> and yet have managed to avoid such proofs (as well as 'subst') completely.
> - in the example for 'adc' on p.4, why doesn't (plus 0 0) reduce to 0 ?  That seems odd.
> - it should be mentioned that with the right combinators, proofs such as those of
>    adc_lemma_2 really are not bad.  [This doesn't take anything away from the core
>     contributions of the paper!]
> - There is a huge buried assumption, that all terms over algebraic structures
>   have decidable equality!  [set_eq of the Set interface].  This clashes quite a bit
>   with the naive interpretation of lifting constants into the Expression domains
>   (ConstMa, ConstG, etc).  This means, for example, that the 'reals' cannot be
>   seen as a Ring, unless one forbids representations of constants other than those
>   from Z or Q.
> - Why on earth would there be a 'hand proof' (eq_preserves_eq) on p.7 in a paper
>   about Idris?!?
> - What happened to Pointed Magma?  Left-unital semigroup?  Right-unital semigroup?
>   Quasigroup? left and right loops?
>   All sorts of odd things happen when one starts being fine-grained.
> - The interface to a Group should have *either* unary or binary -, and then use a
>   *definitional* (and thus conservative) extensio to add the other.  This is again
>   what Realms are useful for.
> - It is worth mentioning at the start of section 2.3 that "analysing syntax" is the
>   heart of reflection.
> - ExprMa: who writes this?  Humans?  The structure of this is SO regular, it screams
>   automation.  It is done that way in MathScheme.
> - It should also be noted that this paper uses the 'local method' of reasoning
>   with syntax; it should be constrasted with the 'global method' as well. [4]
> - What is the reason for lifting constants?
> - Given that there is a 'hierarchy' of algebraic structures, why are the syntax
>   types hand-built and non-hierarchical?  [I know why, but not all readers will
>   know!]
> - buildPRoofGroup:
>   a) is there a need for one of these per structure? [Group, Magma, etc]?
>   b) what is 'exprG_eq' ?
>   c) who binds 'c1' and 'c2'?
> - p.13 "We always simplify as much as possible with constants".
>   a) what is 'simplification'?
>   b) this isn't something you have a choice to do / not do.  If you don't, you won't
>   get a normal form.
> - The hacks involved in Variable (i.e. treat Neg as a pseudo-variable) don't scale
>   well.  Once you get to > 100 structures in your hierarchy, that will become quite
>   obvious.  And life gets really fun when you get above 1000 (as we now have).
> - section 2.7 really REALLY deserves to be longer.  There are interesting things to
>   be said here.
> - You really should also compare to Agda's reflection in the related work.  Possibly
>   to Lean as well.
> 
> [1] https://arxiv.org/abs/1204.0053
> [2] https://arxiv.org/abs/1405.5956
> [3] https://arxiv.org/abs/1105.0548
> [4] https://arxiv.org/abs/1305.6052
> 
> ===
> 
> Jacques Carette
> 
> ----------------------- REVIEW 3 ---------------------
> PAPER: 42
> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
> AUTHORS: Franck Slama and Edwin Brady
> 
> Overall evaluation: 0 (borderline paper)
> 
> ----------- Overall evaluation -----------
> The paper presents an approach to automation of equalities in Idris.
> Idris is a dependently typed programming language with propositional
> equality, which is not automatically decidable. In this setting
> proving equalities automatically for a wide set of common cases is
> certainly very desirable.
> 
> The paper starts with a simple example of a bit-vector adder
> illustrating the (well known) issues with proving equality and
> motivates the need for a generic prover for typical structures – one
> would like to have a generic prover that can handle various datatypes
> with a similar structures.
> 
> The design and implementation of the provers are then described. The
> provers are based on reflection that is type-safe and correct by
> construction (Agda and Coq don’t have this property). Additionally,
> the tactics are developed in a hierarchical way so that the
> normalization is generic and can be inherited from the underlying
> structures. The developed provers cover many useful algebraic cases
> such as monoids, semi-groups, groups, semi-rings and rings.
> 
> Overall the paper is a nice presentation of a new and useful
> development in Idris, with some novel features compared to similar
> tools.
> 
> ------------------------------------------------------
> 
> Best wishes,

### 2017-05-02 paper-CICM-17/rebuttal_reviews_Brady_Slama_CICM17.txt
updated rebuttal

@@
 
-You're entierly right : the motivating example is an illustration of how these proofs obligations can naturally occur, but there are also many other possible utilisations of our proof automation.
+You're entierly right : the motivating example is an illustration of how these proofs obligations can naturally occur, but there are also many other possible utilisations of our proof automations.
 
@@
 
-It is interesting to note that the study of the properties of an alleged normalisation function can only be meta-theoretic : it's not possible to prove it within the proof assistant.
-Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also notice it for their implementation
+It is interesting to note that the study of the properties of an alleged normalisation function can only be meta-theoretic : it's not possible to prove completeness within the proof assistant.
+Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also noticed it for their implementation
 of their Ring prover for Coq. It's not an issue as it does not affect the correctness of the decision procedure.
@@
 
-Sorry about them. Thanks for finding them. We will correct.
+Sorry about them, and thanks for finding them. We will correct.
 
@@
 > equivalence which is (somewhat) modular.
-> 
 > The explanations are clear -- but written in an extremely verbose style.  Cutting
 > two pages would be very easy; cutting 3, feasible.
-> 
 > Saving such space would allow a bit of room to detail some features of Idris a bit
@@
 > c. why no implicits are used (such as in the signature of addBit)
-> 
+
+We believe it will be possible to condense the paper to make space for your suggestions.
+
 > The start of section 2 talks about 'natural inheritance' of interfaces, such as that
@@
 > by Carette & O'Connor for the details. (All papers linked have been published)
-> 
+
+Agreed, but this is not a problem for us, as the hierarchy is not supposed to deal with more
+than the 7 to 10 useful albebraic structures. Coq for example only has a Ring prover (commutative
+or not), but we've obtained, almost for free, provers for semigroups, monoids, and groups thanks
+to the modularity with which we compute the normal form.
+
 > Furthermore, there seems to be an implicit assumption that algebraic theories
@@
 > exactly what "Realms" [2] are all about.
-> 
+
+Sorry, we haven't been very clear on this point. When we were saying that groups admit a 
+normal form, we were talking about structures that *only* have the axioms of a group,
+without additional axioms. Additional axioms can of course lead to a rewriting system
+that is not confluent (or that is not guarantee to terminate), thus destroying the existence
+of the normal form. This is not particular to group and this also applies to the other algebraic structures.
+
+In our paper, we only focus in these "pures" algebraic structures (semi-groups, monoids,
+groups, rings...) that do not have any additional axiom and which are guarantee to have a normal form.
+If the user uses a structure that does have additional axioms, then there is no guarantee that
+one of our prover will be able produce the proofs that he/she needs (it will depend on the equation
+he needs to prove and the necessity to use the extra axioms). If it fails, he can try to 
+extend the closest algebraic structure of the hierarchy by dealing on his own with his 
+additional axioms (in the case where his concrete structure effectively admits a normal form).
+
 > The largest flaw though seems to be the implicit assumption that lots (most? all?)
 > interesting structures have normal forms.  And that these can be found using a
-> modular algorithm.  Without ever talking about Knuth-Bendix, no less!  If one is going
+> modular algorithm. Without ever talking about Knuth-Bendix, no less!
+
+Agreed, there are indeed many structures which do not have normal norms. But deduction modulo, 
+although very interesting, was somehow out of scope for our goals. We initially wanted Idris 
+to be equiped with some automations, as Coq is with its ring tactic, and we've done it on a modular way
+which made it easy to also implement at the same time (and almost for free) a prover for semi-groups, 
+monoids, and groups (commutative or not). We do not want to clame that every interesting structure 
+can be treated by one prover of our hierachy of tactics. However, our hierarchy of provers is more general
+than the automations available in Coq or Agda for example.
+
+> If one is going
 > to talk about a "correct by construction" approach, then the *definition* of what a
@@
 > invariant kept by the types!
-> 
+
+Yes, the types only say that the original reflected term and the normalised one both represent the
+same concrete value, and this is in fact the proof of *correctness* of the approach. 
+The *completeness* is a meta property that we cannot prove in the language itself. 
+Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also noticed it for their implementation
+of their Ring prover for Coq. It's not an issue as it does not affect the correctness of the decision procedure.
+
 > For example, the word problem for groups is undecidable.  This implies that there are
 > theories with no normal forms.
-> 
+
+Yes, we agree, see our answer to the two previous questions.
+
 > In section 2.5, an *order* is snuck in, onto the set of variables.  This is crucial!
@@
 > But normal forms are extremely sensitive to term orders (as bad as exponential).
+
+The set of variables V is said to be ordered in section 2.5 (page 12). Note that this is not a limitation
+as it is always possible to build an adhoc ordering by using the order of appearance of the variables in the goal A = B
+This is in fact exactly what our automatically reflection does (mentionned briefly in section 2.7).
+
 > Furthermore, the discussion in this section implicitly assumes associativity.  There
 > are many useful, interesting, non-associative algebras...
-> 
-> Overall, the point is that given a structure X, then the induced language-of-X
-> has a nice structure as well.  Most of it is fairly automatically derivable from the
-> definition of X.
-> 
+
+We agree that there are many useful non-associative algebras. However, our work was focused on rings
+and all of the underneath structures. This is why, at the very root, we have semi-groups. Therefore,
+the associativity is the first axiom that we deal with, and our machinery is not intended for non-associative algebras.
+
 > Various further comments:
@@
 > and yet have managed to avoid such proofs (as well as 'subst') completely.
+
+Edwin, what would you answer here  I could perhaps say that anyway, our automation is *not only* intended
+for problems related to the use of dependent types (as the first reviewer has kindly mentionned). Proving
+equalities happens more often that proving type equalities...
+
 > - in the example for 'adc' on p.4, why doesn't (plus 0 0) reduce to 0 ?  That seems odd.
+
+Edwin : I don't know if it is still the case. It might have been due to an older version of Idris.
+
 > - it should be mentioned that with the right combinators, proofs such as those of
@@
 >     contributions of the paper!]
+
+Agreed.
+
 > - There is a huge buried assumption, that all terms over algebraic structures
@@
 >   with the naive interpretation of lifting constants into the Expression domains
->   (ConstMa, ConstG, etc).  This means, for example, that the 'reals' cannot be
->   seen as a Ring, unless one forbids representations of constants other than those
->   from Z or Q.
+>   (ConstMa, ConstG, etc).
+
+We think that there is a misunderstanding on this point : the required decidable equality
+is only a decidable equality for the set of constants of the polynomials. For example, when 
+trying to prove equalities over Nat, the needed decidable equality is the one that decides 
+the equality between two natural numbers (with no variables and no operations), called decEq in Idris.
+
 > - Why on earth would there be a 'hand proof' (eq_preserves_eq) on p.7 in a paper
 >   about Idris?!?
+
+We thought such a 'hand proof' would be easier to read to a general audience.
+
 > - What happened to Pointed Magma?  Left-unital semigroup?  Right-unital semigroup?
@@
 >   All sorts of odd things happen when one starts being fine-grained.
+
+As said before, we did not aim to implement all algebraic structures that exist.
+Our machinery only deals with the intermediate algebraic structures that lead to rings.
+That certainly wasn't said clearly in the paper. We will make it clearer.
+
 > - The interface to a Group should have *either* unary or binary -, and then use a
@@
 >   what Realms are useful for.
+
+Having both is useful if the user has something to prove with both the unary Neg and the binary Minus. 
+Expressing in the interface something like "there is either a unary or binary -" 
+would have been a hassle. However, this is not a limitation : it is always possible 
+for the user to define one with the other (and the proof Minus_simpl that he/she 
+has to provide then becomes trivial).
+
 > - It is worth mentioning at the start of section 2.3 that "analysing syntax" is the
 >   heart of reflection.
+
+Agreed, we will mention it.
+
 > - ExprMa: who writes this?  Humans?  The structure of this is SO regular, it screams
 >   automation.  It is done that way in MathScheme.
+
+The datatypes themselves are handwritten. However, building the terms of type ExprMa, ExprG, ExprR,
+ie, building the encoding of the concrete values is done automatically by our automatic reflection mecanism (described briefly in 2.7).
+
 > - It should also be noted that this paper uses the 'local method' of reasoning
 >   with syntax; it should be constrasted with the 'global method' as well. [4]
+
+I don't know what to answer to this. Edwin ? :)
+
 > - What is the reason for lifting constants?
+
+This is so trivial that I don't know how to answer to this point. Any idea Edwin ?
+
 > - Given that there is a 'hierarchy' of algebraic structures, why are the syntax
@@
 >   know!]
+
+This is so trivial that I don't know how to answer to this point. Any idea Edwin ?
+
 > - buildPRoofGroup:
@@
 >   b) what is 'exprG_eq' ?
+
+This is a standard equality test (mentionned on the fly page 7) that checks equality of syntax.
+
 >   c) who binds 'c1' and 'c2'?
+
+Sorry, it's a typo. c1 should be x' and c2 should be y'. We will fix it. Thanks for finding it.
+
 > - p.13 "We always simplify as much as possible with constants".
@@
 >   get a normal form.
+
+Indeed, we *have* to do it. We still wanted to describe what the normalisation function has to do,
+and not to only mention the arbitrary decisions that we had to take (such as writing everything on
+the complete left or complete right associative form).
+
+These simplifications entail simplifying the addition with zero, and the multiplication 
+with the constants zero and with one, doing the computations between two nearby constants, etc.
+(mentionned briefly on page 13)
+
 > - The hacks involved in Variable (i.e. treat Neg as a pseudo-variable) don't scale
@@
 >   obvious.  And life gets really fun when you get above 1000 (as we now have).
+
+We completely agree, but we won't have more than 10 structures. The goal was to implement automatic provers
+for the albebraic structures that are underneath rings, not to deal with all possible algebraic structure.
+Having a modular computation of the normal form as the one we have would be very difficult if not impossible to maintain for
+more than 10, or perhaps 20 algebraic structures. We would encounter a lot of other issues before we encounter the
+one you mentionned about the hacks in Variable. Our modularity was useful for getting almost for free some
+provers for the undernath structures of rings, but this approach would not necessarly would be well suited for dealing with
+100 or 1000 structures, nor it would for dealing with arbitrary rewriting rules.
+
 > - section 2.7 really REALLY deserves to be longer.  There are interesting things to
 >   be said here.
+
+Ok, we will say more about it in the final draft.
+
 > - You really should also compare to Agda's reflection in the related work.  Possibly
 >   to Lean as well.
-> 
-> [1] https://arxiv.org/abs/1204.0053
-> [2] https://arxiv.org/abs/1405.5956
-> [3] https://arxiv.org/abs/1105.0548
-> [4] https://arxiv.org/abs/1305.6052
-> 
-> ===
-> 
-> Jacques Carette
-> 
+
+Ok.
+
 > ----------------------- REVIEW 3 ---------------------
@@
 > certainly very desirable.
-> 
+
+Agreed.
+
 > The paper starts with a simple example of a bit-vector adder
@@
 > with a similar structures.
-> 
+
+Agreed.
+
 > The design and implementation of the provers are then described. The
@@
 > such as monoids, semi-groups, groups, semi-rings and rings.
-> 
+
+Agreed.
+
 > Overall the paper is a nice presentation of a new and useful
@@
 > tools.
-> 
-> ------------------------------------------------------
-> 
-> Best wishes,
+
+Thanks, we agree.

### 2017-05-02 paper-CICM-17/rebuttal_reviews_Brady_Slama_CICM17.txt
Brutal pruning of rebuttal!

--- 2017-05-02 paper-CICM-17/rebuttal_reviews_Brady_Slama_CICM17.txt
@@
-> ----------------------- REVIEW 1 ---------------------
-> PAPER: 42
-> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
-> AUTHORS: Franck Slama and Edwin Brady
-> 
-> Overall evaluation: 2 (accept)
-> 
-> ----------- Overall evaluation -----------
-> --- Summary ---
-> 
-> The article presents a technique to automate proofs of equivalence over some algebraic structures, in the context of proof assistants based on dependent types, in particular Idris.
-> The motivation behind this work is to relieve users of the burden of writing trivial but tedious proofs, in particular proofs of type equality that are required for some definitions over dependent types.
-> The proofs are done by rewriting terms to a canonical form.
-The rewriting functions work on reflected terms.
-The main novel idea of the paper is the use of dependent types to index the reflected terms with the concrete term that they represent.
-This enforces a "correct-by-construction" approach, where the normalizing functions must also provide a proof that they preserve equivalence, thus facilitating the rest of the proof.
+We thank the reviews for their careful reading of the paper and many helpful
+comments, which we will address in a revision. In particular, we are sure we
+can condense the paper according to Jacques' useful suggestions.
 
-Agreed.
+> although proofs of type equivalences are not the only use case
 
-> --- Content ---
-> 
-> The use of indexed reflected terms seems to be a good approach. The modularity of the tactics to work on different structures, as well as the automation within Indris, are nice additions.
+Indeed; the motivating example is an illustration of how these proofs
+obligations occur naturally, but there are many other uses.
 
-Agreed.
+> One question that comes to mind is whether it easy to prove properties of the
+> normalization functions
 
-> The example in the introduction provides a good motivation, although proofs of type equivalences are not the only use case for this work.
+Things become difficult when we leave the ground of the algebraic
+structures in which we work (monoids, groups and rings essentially) and
+move to an arbitrary equivalence relation with arbitrary axioms. Such a system
+can be seen as a rewriting system of deduction modulo. Determining if such a
+rewriting system admits a normal form requires determining whether the rewriting
+system is confluent and terminating. Both questions are difficult
+to answer in the general case. Currently, systems like Dedukti (an
+implementation of lambda-Pi-calculus modulo) can plug to an auxiliary
+confluence checker such as CSI^ho
+(http://cl-informatik.uibk.ac.at/software/csi/ho/) or ACPH
+(http://coco.nue.riec.tohoku.ac.jp/2015/papers/acph.pdf).
 
-You're entierly right : the motivating example is an illustration of how these proofs obligations can naturally occur, but there are also many other possible utilisations of our proof automation.
+It is interesting to note that the study of the properties of a proposed
+normalisation function can only be meta-theoretic: it's not possible to prove
+completeness within the language. Benjamin Grégoire and Assia Mahboubi
+[cited in our paper] have also noticed it for their implementation of their
+Ring prover for Coq. However, this does not affect the correctness of the
+decision procedure.
 
-> One question that comes to mind is whether it easy to prove properties of the normalization functions (e.g.
-given a concrete equivalence relation, proving that each equivalence class indeed has a unique canonical representative).
-Does the use of reflected terms have an impact on such proofs?
+> An interesting question for future work would be to exploit external provers
 
-Things become very difficult when we leave the ground of the algebraic structures in which we are working (monoids, groups and rings essentially) and move to an arbitrary equivalence relation with arbitrary axioms.
-Such a system can be seen as a rewriting system of
-dedution modulo.
-Determining if such a rewriting system admits a normal form requires to determine if the rewriting system formed is confluent and terminating.
-Both questions are very difficult to answer in the general case.
-Currently, systems like Dedukti
-(an implementation of lambda-Pi-calculus modulo) can plug to an auxiliary confluence checker such as CSI^ho ( http://cl-informatik.uibk.ac.at/software/csi/ho/ ) or ACPH ( http://coco.nue.riec.tohoku.ac.jp/2015/papers/acph.pdf ).
+Agreed, e.g. this would allow dealing with arbitrary rewriting rules.
 
-It is interesting to note that the study of the properties of an alleged normalisation function can only be meta-theoretic : it's not possible to prove it within the proof assistant.
-Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also notice it for their implementation
-of their Ring prover for Coq. It's not an issue as it does not affect the correctness of the decision procedure.
+> ... as soon as the hierarchy grows large enough, such
+> 'natural' inheritance becomes very brittle ....
 
-> An interesting question for future work would be to exploit external provers (first-order theorem provers, rewriting engines...) to find equivalences on structures with arbitrary rewriting rules, and later reconstruct the proofs within the proof assistant.
+Agreed, but our hierarchy is not supposed to deal with more than the most
+useful algebraic structures which arise in Idris programming. Coq for example
+only has a Ring solver but we have also provers for semigroups, monoids, and
+groups thanks to the modularity with which we compute the normal form.
 
-Yes, being able to use an external prover is an interesting future work to investigate, as it would allow to deal with arbitrary rewriting rules.
+> there seems to be an implicit assumption that algebraic theories
+> have a canonical presentation
 
-> --- Presentation ---
-> The paper is generally well-written.
-A lot of effort has been spent on providing motivation and context, and the content is easy to follow even without being an expert in the domain.
-The various aspects involved are each presented in sufficient detail.
+To clarify: when we say groups admit a normal form, we mean structures that
+*only* have the axioms of a group.  We focus on these "pure" algebraic
+structures (semi-groups, monoids, groups, rings...) that do not have
+additional axioms and which do have a normal form.
 
-Thanks.
+> The largest flaw though seems to be the implicit assumption interesting structures have normal forms.
 
-> Subsection 2.1 seems a bit out of order.
-It essentially amounts to an extended footnote: for the use case described in the introduction, it seems that one really wants to prove type equalities rather than arbitrary equivalences.
-The extension to arbitrary equivalence relations is interesting, but not central.
-It is therefore a bit confusing to see that subsection so early in the text, and I would suggest moving it to a later point.
+Agreed, there are indeed many structures which do not have normal norms. But deduction modulo, 
+although very interesting, was somehow out of scope for our goals. We initially wanted Idris 
+to be equiped with some automations, which we do in a modular way which makes it easy to implement
+provers for other structures.
 
-You are entierely right.
-We will just mention at the start of Section 2 that we want to be general and to work with any equivalence relation (~=), and we will move this section 2.1 just after what is currently section 2.6.
-We still need to say two words about the equivalence relation (~=) at the start of section 2 as we use it everywhere and it needs to be introduced briefly.
+We don't claim that every interesting structure can be treated by one prover of
+our hierachy of tactics, although our hierarchy of provers is more general than
+the automations available in e.g. Coq or Agda.
 
-> Textual errors found:
-> p.11: "admit a canonical [representative]"
-> p.15: "deduction modulo and proof search heuristics [start]."
-
-Sorry about them. Thanks for finding them. We will correct.
-
-> ----------------------- REVIEW 2 ---------------------
-> PAPER: 42
-> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
-> AUTHORS: Franck Slama and Edwin Brady
-> 
-> Overall evaluation: 1 (weak accept)
-> 
-> ----------- Overall evaluation -----------
-> [This is a signed review]
-> 
-> First two items:
-> 1. Paper is definitely in-scope for Calculemus
-> 2. As written, it is too long by 2 pages [but see below].
-> 
-> This paper describes one method (type-safe reflection) to obtain methods of proving
-> equivalence which is (somewhat) modular.
-> 
-> The explanations are clear -- but written in an extremely verbose style.  Cutting
-> two pages would be very easy; cutting 3, feasible.
-> 
-> Saving such space would allow a bit of room to detail some features of Idris a bit
-> more for the average CICM reader, such as
-> a. unbound variables (like 'a' in the definition of Id being implicitly quantified
->   over Type)
-> b. that ** is dependent pairing
-> c. why no implicits are used (such as in the signature of addBit)
-> 
-> The start of section 2 talks about 'natural inheritance' of interfaces, such as that
-> between Group and Monoid.  When such algebraic hierarchies are done "naively", yes
-> this appears to occur.  But as soon as the hierarchy grows large enough, such
-> 'natural' inheritance becomes very brittle.  What replaces it are explicit views, aka
-> theory morphisms [see MMT [3] and IMPS].  One can still use transport along such morphisms
-> to move results from one place to another.  Some morphisms are still special (the
-> display morphisms of the opposite category]; see "Theory Presentation Combinators" [1]
-> by Carette & O'Connor for the details. (All papers linked have been published)
-> 
-> Furthermore, there seems to be an implicit assumption that algebraic theories
-> (such as Group) have a canonical presentation -- this is not true.  And that is
-> exactly what "Realms" [2] are all about.
-> 
-> The largest flaw though seems to be the implicit assumption that lots (most? all?)
-> interesting structures have normal forms.  And that these can be found using a
-> modular algorithm.  Without ever talking about Knuth-Bendix, no less!  If one is going
+> If one is going
 > to talk about a "correct by construction" approach, then the *definition* of what a
@@
 > But normal forms are extremely sensitive to term orders (as bad as exponential).
+
+The set of variables V is said to be ordered in section 2.5 (page 12). Note that this is not a limitation
+as it is always possible to build an adhoc ordering by using the order of appearance of the variables in the goal A = B
+This is in fact exactly what our automatically reflection does (mentionned briefly in section 2.7).
+
 > Furthermore, the discussion in this section implicitly assumes associativity.  There
 > are many useful, interesting, non-associative algebras...
-> 
-> Overall, the point is that given a structure X, then the induced language-of-X
-> has a nice structure as well.  Most of it is fairly automatically derivable from the
-> definition of X.
-> 
-> Various further comments:
-> - a citation in the last paragraph of the introduction ("This is even an everyday
-> routine when ...") would be appreciated.  I index types with values all the time,
-> and yet have managed to avoid such proofs (as well as 'subst') completely.
-> - in the example for 'adc' on p.4, why doesn't (plus 0 0) reduce to 0 ?  That seems odd.
-> - it should be mentioned that with the right combinators, proofs such as those of
->    adc_lemma_2 really are not bad.  [This doesn't take anything away from the core
->     contributions of the paper!]
-> - There is a huge buried assumption, that all terms over algebraic structures
->   have decidable equality!  [set_eq of the Set interface].  This clashes quite a bit
->   with the naive interpretation of lifting constants into the Expression domains
->   (ConstMa, ConstG, etc).  This means, for example, that the 'reals' cannot be
->   seen as a Ring, unless one forbids representations of constants other than those
->   from Z or Q.
-> - Why on earth would there be a 'hand proof' (eq_preserves_eq) on p.7 in a paper
->   about Idris?!?
-> - What happened to Pointed Magma?  Left-unital semigroup?  Right-unital semigroup?
->   Quasigroup? left and right loops?
->   All sorts of odd things happen when one starts being fine-grained.
+
+We agree that there are many useful non-associative algebras. However, our work
+was focused on rings and all of the underneath structures. This is why, at the
+very root, we have semi-groups. Therefore, the associativity is the first axiom
+that we deal with, and our machinery is not intended for non-associative
+algebras.
+
+> buried assumption, that all terms over algebraic structures have decidable equality!
+
+To clarify: the required decidable equality is only a decidable equality for
+the constants, e.g, when proving equalities over Nat, the needed decidable
+equality is the one that decides the equality between two natural numbers (with
+no variables and no operations), called decEq in Idris.
+
 > - The interface to a Group should have *either* unary or binary -, and then use a
@@
 >   what Realms are useful for.
-> - It is worth mentioning at the start of section 2.3 that "analysing syntax" is the
->   heart of reflection.
-> - ExprMa: who writes this?  Humans?  The structure of this is SO regular, it screams
->   automation.  It is done that way in MathScheme.
-> - It should also be noted that this paper uses the 'local method' of reasoning
->   with syntax; it should be constrasted with the 'global method' as well. [4]
-> - What is the reason for lifting constants?
-> - Given that there is a 'hierarchy' of algebraic structures, why are the syntax
->   types hand-built and non-hierarchical?  [I know why, but not all readers will
->   know!]
-> - buildPRoofGroup:
->   a) is there a need for one of these per structure? [Group, Magma, etc]?
->   b) what is 'exprG_eq' ?
->   c) who binds 'c1' and 'c2'?
-> - p.13 "We always simplify as much as possible with constants".
->   a) what is 'simplification'?
->   b) this isn't something you have a choice to do / not do.  If you don't, you won't
->   get a normal form.
-> - The hacks involved in Variable (i.e. treat Neg as a pseudo-variable) don't scale
->   well.  Once you get to > 100 structures in your hierarchy, that will become quite
->   obvious.  And life gets really fun when you get above 1000 (as we now have).
-> - section 2.7 really REALLY deserves to be longer.  There are interesting things to
->   be said here.
-> - You really should also compare to Agda's reflection in the related work.  Possibly
->   to Lean as well.
-> 
-> [1] https://arxiv.org/abs/1204.0053
-> [2] https://arxiv.org/abs/1405.5956
-> [3] https://arxiv.org/abs/1105.0548
-> [4] https://arxiv.org/abs/1305.6052
-> 
-> ===
-> 
-> Jacques Carette
-> 
-> ----------------------- REVIEW 3 ---------------------
-> PAPER: 42
-> TITLE: Automatically Proving Equivalence by Type-Safe Reflection
-> AUTHORS: Franck Slama and Edwin Brady
-> 
-> Overall evaluation: 0 (borderline paper)
-> 
-> ----------- Overall evaluation -----------
-> The paper presents an approach to automation of equalities in Idris.
-> Idris is a dependently typed programming language with propositional
-> equality, which is not automatically decidable. In this setting
-> proving equalities automatically for a wide set of common cases is
-> certainly very desirable.
-> 
-> The paper starts with a simple example of a bit-vector adder
-> illustrating the (well known) issues with proving equality and
-> motivates the need for a generic prover for typical structures – one
-> would like to have a generic prover that can handle various datatypes
-> with a similar structures.
-> 
-> The design and implementation of the provers are then described. The
-> provers are based on reflection that is type-safe and correct by
-> construction (Agda and Coq don’t have this property). Additionally,
-> the tactics are developed in a hierarchical way so that the
-> normalization is generic and can be inherited from the underlying
-> structures. The developed provers cover many useful algebraic cases
-> such as monoids, semi-groups, groups, semi-rings and rings.
-> 
-> Overall the paper is a nice presentation of a new and useful
-> development in Idris, with some novel features compared to similar
-> tools.
-> 
-> ------------------------------------------------------
-> 
-> Best wishes,
+
+Having both is useful if the user has something to prove with both the unary Neg and the binary Minus. 
+Expressing in the interface something like "there is either a unary or binary -" 
+would have been a hassle. However, this is not a limitation : it is always possible 
+for the user to define one with the other (and the proof Minus_simpl that he/she 
+has to provide then becomes trivial).
+
+>   this isn't something you have a choice to do / not do.
+
+Indeed, we *must* do it. We still wanted to describe what the normalisation function has to do,
+and not to only mention the arbitrary decisions that we had to take (such as writing everything on
+the complete left or complete right associative form).
+
+These simplifications entail simplifying the addition with zero, and the multiplication 
+with the constants zero and with one, doing the computations between two nearby constants, etc.
+(mentionned briefly on page 13)
+
+> The hacks involved in Variable (i.e. treat Neg as a pseudo-variable) don't scale
+> well.
+
+We agree, but our goal was to implement automatic provers for the albebraic
+structures that are underneath rings, not to deal with all possible algebraic
+structure. We'll make this clearer in the paper.

### 2017-05-03 paper-CICM-17/rebuttal-proposed.txt
Added proposed rebuttal text
There's about 30 words to spare...

--- 2017-05-02 paper-CICM-17/rebuttal_reviews_Brady_Slama_CICM17.txt
@@
 comments, which we will address in a revision. In particular, we are sure we
-can condense the paper according to Jacques' useful suggestions.
+can condense the paper according to Jacques' suggestions.
 
-> although proofs of type equivalences are not the only use case
-
-Indeed; the motivating example is an illustration of how these proofs
-obligations occur naturally, but there are many other uses.
-
-> One question that comes to mind is whether it easy to prove properties of the
-> normalization functions
-
-Things become difficult when we leave the ground of the algebraic
-structures in which we work (monoids, groups and rings essentially) and
-move to an arbitrary equivalence relation with arbitrary axioms. Such a system
-can be seen as a rewriting system of deduction modulo. Determining if such a
-rewriting system admits a normal form requires determining whether the rewriting
-system is confluent and terminating. Both questions are difficult
-to answer in the general case. Currently, systems like Dedukti (an
-implementation of lambda-Pi-calculus modulo) can plug to an auxiliary
-confluence checker such as CSI^ho
-(http://cl-informatik.uibk.ac.at/software/csi/ho/) or ACPH
-(http://coco.nue.riec.tohoku.ac.jp/2015/papers/acph.pdf).
-
-It is interesting to note that the study of the properties of a proposed
-normalisation function can only be meta-theoretic: it's not possible to prove
-completeness within the language. Benjamin Grégoire and Assia Mahboubi
-[cited in our paper] have also noticed it for their implementation of their
-Ring prover for Coq. However, this does not affect the correctness of the
-decision procedure.
-
-> An interesting question for future work would be to exploit external provers
-
-Agreed, e.g. this would allow dealing with arbitrary rewriting rules.
-
-> ... as soon as the hierarchy grows large enough, such
-> 'natural' inheritance becomes very brittle ....
-
-Agreed, but our hierarchy is not supposed to deal with more than the most
-useful algebraic structures which arise in Idris programming. Coq for example
-only has a Ring solver but we have also provers for semigroups, monoids, and
-groups thanks to the modularity with which we compute the normal form.
-
-> there seems to be an implicit assumption that algebraic theories
-> have a canonical presentation
-
-To clarify: when we say groups admit a normal form, we mean structures that
-*only* have the axioms of a group.  We focus on these "pure" algebraic
-structures (semi-groups, monoids, groups, rings...) that do not have
-additional axioms and which do have a normal form.
-
-> The largest flaw though seems to be the implicit assumption interesting structures have normal forms.
-
-Agreed, there are indeed many structures which do not have normal norms. But deduction modulo, 
-although very interesting, was somehow out of scope for our goals. We initially wanted Idris 
-to be equiped with some automations, which we do in a modular way which makes it easy to implement
-provers for other structures.
+Regarding the choice of structures and normal forms: our hierarchy is not
+supposed to deal with more than the most useful algebraic structures which
+arise in Idris programming. Coq for example only has a Ring solver but we have
+also provers for semigroups, monoids, and groups thanks to the modularity with
+which we compute the normal form.  We focus on "pure" algebraic structures that
+do not have additional axioms and which do have a normal form.
 
@@
 
-> If one is going
-> to talk about a "correct by construction" approach, then the *definition* of what a
-> normal form is (i.e. syntax where intensional equality is a decision procedure for
-> equality).  This is implicit everywhere, but never explicit.  And it is not an
-> invariant kept by the types!
+On proving properties of normalisation: Things become difficult when we leave
+the ground of the algebraic structures in which we work (monoids, groups and
+rings essentially) and move to an arbitrary equivalence relation with arbitrary
+axioms. Such a system can be seen as a rewriting system of deduction modulo.
+Determining if such a rewriting system admits a normal form requires
+determining whether the rewriting system is confluent and terminating. Both
+questions are difficult to answer in the general case. Currently, systems like
+Dedukti (an implementation of lambda-Pi-calculus modulo) can plug to an
+auxiliary confluence checker such as CSI^ho or ACPH.
 
-Yes, the types only say that the original reflected term and the normalised one both represent the
-same concrete value, and this is in fact the proof of *correctness* of the approach. 
-The *completeness* is a meta property that we cannot prove in the language itself. 
-Benjamin Grégoire and Assia Mahboubi [cited in our paper] have also noticed it for their implementation
-of their Ring prover for Coq. It's not an issue as it does not affect the correctness of the decision procedure.
+It is interesting to note that the study of the properties of a proposed
+normalisation function can only be meta-theoretic. Indeed, Grégoire and
+Mahboubi notice it for their Ring solver in Coq.  However, this does not affect
+the correctness of the decision procedure.
 
-> For example, the word problem for groups is undecidable.  This implies that there are
-> theories with no normal forms.
+there are indeed many structures which do not have normal norms. But deduction
+modulo, although very interesting, was somehow out of scope for our goals. We
+initially wanted Idris to be equiped with some automations, which we do in a
+modular way which makes it easy to implement provers for other structures.
 
-Yes, we agree, see our answer to the two previous questions.
-
-> In section 2.5, an *order* is snuck in, onto the set of variables.  This is crucial!
-> It should not be buried.  Hint: Groebner bases.  Term orders are important.
-> But normal forms are extremely sensitive to term orders (as bad as exponential).
-
-The set of variables V is said to be ordered in section 2.5 (page 12). Note that this is not a limitation
-as it is always possible to build an adhoc ordering by using the order of appearance of the variables in the goal A = B
-This is in fact exactly what our automatically reflection does (mentionned briefly in section 2.7).
-
-> Furthermore, the discussion in this section implicitly assumes associativity.  There
-> are many useful, interesting, non-associative algebras...
-
-We agree that there are many useful non-associative algebras. However, our work
-was focused on rings and all of the underneath structures. This is why, at the
-very root, we have semi-groups. Therefore, the associativity is the first axiom
-that we deal with, and our machinery is not intended for non-associative
-algebras.
-
-> buried assumption, that all terms over algebraic structures have decidable equality!
-
-To clarify: the required decidable equality is only a decidable equality for
-the constants, e.g, when proving equalities over Nat, the needed decidable
-equality is the one that decides the equality between two natural numbers (with
-no variables and no operations), called decEq in Idris.
-
-> - The interface to a Group should have *either* unary or binary -, and then use a
->   *definitional* (and thus conservative) extensio to add the other.  This is again
->   what Realms are useful for.
-
-Having both is useful if the user has something to prove with both the unary Neg and the binary Minus. 
-Expressing in the interface something like "there is either a unary or binary -" 
-would have been a hassle. However, this is not a limitation : it is always possible 
-for the user to define one with the other (and the proof Minus_simpl that he/she 
-has to provide then becomes trivial).
-
->   this isn't something you have a choice to do / not do.
-
-Indeed, we *must* do it. We still wanted to describe what the normalisation function has to do,
-and not to only mention the arbitrary decisions that we had to take (such as writing everything on
-the complete left or complete right associative form).
-
-These simplifications entail simplifying the addition with zero, and the multiplication 
-with the constants zero and with one, doing the computations between two nearby constants, etc.
-(mentionned briefly on page 13)
-
-> The hacks involved in Variable (i.e. treat Neg as a pseudo-variable) don't scale
-> well.
-
-We agree, but our goal was to implement automatic provers for the albebraic
-structures that are underneath rings, not to deal with all possible algebraic
-structure. We'll make this clearer in the paper.
+The motivating example is an illustration of how these proofs obligations occur
+naturally, but there are many other uses.

### 2017-05-03 paper-CICM-17/rebuttal-proposed.txt
rebuttal updated

@@
-We thank the reviews for their careful reading of the paper and many helpful
-comments, which we will address in a revision. In particular, we are sure we
-can condense the paper according to Jacques' suggestions.
+We thank the reviewers for their careful reading of the paper and many helpful
+comments, which we will incorporate in the final draft. In particular, we are 
+sure we can condense the paper according to Jacques' suggestions.
 
@@
 On proving properties of normalisation: Things become difficult when we leave
-the ground of the algebraic structures in which we work (monoids, groups and
-rings essentially) and move to an arbitrary equivalence relation with arbitrary
-axioms. Such a system can be seen as a rewriting system of deduction modulo.
-Determining if such a rewriting system admits a normal form requires
+the ground of the "pure" algebraic structures in which we work (semigroups, monoids, 
+groups and rings essentially) and move to an arbitrary equivalence relation with 
+arbitrary axioms. Such a system can be seen as a rewriting system of deduction 
+modulo. Determining if such a rewriting system admits a normal form requires
 determining whether the rewriting system is confluent and terminating. Both
@@
 Dedukti (an implementation of lambda-Pi-calculus modulo) can plug to an
-auxiliary confluence checker such as CSI^ho or ACPH.
+auxiliary confluence checker such as CSI^ho or ACPH. That would be an interesting
+future work for Idris, as suggested by the first reviewer.
+
+There are indeed many structures which do not have normal norms. But deduction
+modulo, although very interesting, was somehow out of scope for our goals. We
+initially wanted Idris to be equipped with some automations, which we did in a
+modular way which made it easy to implement provers for other structures as well.
 
@@
 normalisation function can only be meta-theoretic. Indeed, Grégoire and
-Mahboubi notice it for their Ring solver in Coq.  However, this does not affect
+Mahboubi noticed it for their Ring solver in Coq. However, this does not affect
 the correctness of the decision procedure.
-
-there are indeed many structures which do not have normal norms. But deduction
-modulo, although very interesting, was somehow out of scope for our goals. We
-initially wanted Idris to be equiped with some automations, which we do in a
-modular way which makes it easy to implement provers for other structures.
 

### 2017-05-16 paper-CICM-17/rebuttal-proposed.txt
final version of the rebuttal certainly
minor changes before Edwin shortens the paper

@@
 We don't claim that every interesting structure can be treated by one prover of
-our hierachy of tactics, although our hierarchy of provers is more general than
+our hierarchy of tactics, although our hierarchy of provers is more general than
 the automations available in e.g. Coq or Agda.
@@
 questions are difficult to answer in the general case. Currently, systems like
-Dedukti (an implementation of lambda-Pi-calculus modulo) can plug to an
-auxiliary confluence checker such as CSI^ho or ACPH. That would be an interesting
-future work for Idris, as suggested by the first reviewer.
+Dedukti (an implementation of lambda-Pi-calculus modulo) can plug to an auxiliary 
+confluence checker such as CSI^ho or ACPH. That would be an interesting future 
+work for Idris, as suggested by the first reviewer.
 
-There are indeed many structures which do not have normal norms. But deduction
-modulo, although very interesting, was somehow out of scope for our goals. We
-initially wanted Idris to be equipped with some automations, which we did in a
-modular way which made it easy to implement provers for other structures as well.
+There are indeed many structures which do not have normal norms. But this problem, 
+although very interesting, was somehow out of scope for our goals. We initially 
+wanted Idris to be equipped with some automations, which we did in a modular way 
+and with a new type-safe reflection approach, which made it easy to implement 
+provers for other structures as well.
 

### 2017-05-16 paper-CICM-17/content/1-introduction.tex
minor changes before Edwin shortens the paper

@@
 
-Prover for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda\%5C?n=Libraries.UsingTheRingSolver}} ring prover.
+Provers for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda\%5C?n=Libraries.UsingTheRingSolver}} ring prover.
 In this paper, we describe a certified
 implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris/Provers}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
-structures, including monoid, groups and rings (all potentially commutative),
+structures, including monoids, groups and rings (all potentially commutative),
 for the Idris language.
@@
 
-\item We develop a hierarchy of tactics (and not just a tactic for, say, ring) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
+\item We develop a hierarchy of tactics (and not just a tactic for, say, rings) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
 For example, simplifying neutral elements is only implemented at the monoid level, and each level inheriting from it will reuse it.

### 2017-05-16 paper-CICM-17/content/3-hierarchyOfTactics.tex
minor changes before Edwin shortens the paper

@@
 What we want to build is a prover able to prove not only equalities for \code{Nat}, \code{List}, or any specific type, but for various datatypes, and for various properties.
-With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrix, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for semi-groups, monoids, commutative-monoids, groups, commutative-groups, semi-rings and rings.
+With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrices, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for semi-groups, monoids, commutative-monoids, groups, commutative-groups, semi-rings and rings.
 These tactics will be usable on any type that satisfies the properties of one of these algebraic structure.
@@
     buildProofGroup p ex' ex' px py | Just refl = 
-                Just(eq_preserves_eq x y c1 c2 lp rp e1_equiv_e2)
+                Just(eq_preserves_eq x y x' y' lp rp e1_equiv_e2)
     buildProofGroup p ex' ey' px py | Nothing = Nothing
@@
 Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
-The fact that all of these algebraic structures admit a canonical represent for any element is in fact a very nice property that we are using in order to decide equalities.
+The fact that all of these algebraic structures admit a canonical representative for any element is in fact a very nice property that we are using in order to decide equalities.
 Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.

### 2017-05-16 paper-CICM-17/content/1-introduction.tex
Prune abstract and section 1

--- 2017-03-31 paper-CICM-17/content/1-introduction.tex
@@
 
-Proofs assistants like Coq~\cite{BertotC04}, Agda~\cite{DBLP:conf/tldi/Norell09} and Idris~\cite{brady2013idris} are based on Intensional Type Theories, that contain two notions of equality.
-One is the equality that can be manipulated in the language, called the propositional equality.
-This equality corresponds to the equality found in mathematics : this is a proposition that can be assumed, negated, proved or disproved.
-Since in type theory, propositions are seen as types~\cite{How80}, the
-proposition that two elements $x$ and $y$ are equal corresponds to a type.
+Proofs assistants like Coq~\cite{BertotC04} 
+and programming languages like 
+Agda~\cite{DBLP:conf/tldi/Norell09} and
+Idris~\cite{brady2013idris} are based on Intensional Type Theories that contain
+two notions of equality: propositional equality, that can be
+manipulated in the language, and judgemental (or definition) equality.
+Propositional equality corresponds to the mathematical notion: this is
+a proposition that can be assumed, negated, proved or disproved. Since in type
+theory, propositions are types~\cite{How80}, the proposition that two
+elements $x$ and $y$ are equal corresponds to a type.
 If $x$ and $y$ are of type $a$, then the type
@@
 
-The second notion of equality in Intentional Type Theories is a primitive concept, called judgemental equality, or definitional equality.
-This second equality means ``equal by definition''.
-The judgemental equality cannot be negated or assumed; we cannot talk about this primitive equality inside the theory.
-Whether or not two expressions are equal by definition is a matter of
+Judgemental equality, on the other hand, is a primitive concept of
+the type theory.
+Whether or not two expressions are judgementally equal is a matter of
 evaluating the definitions. For example, if $f\ :\ \mathbb{N}\ \rightarrow\
@@
 
-The judgemental equality has to be included in the propositional equality, because what is equal by definition should be provably equal.
-This is accomplished by giving constructors to the type $Id(a,a)$ and nothing when ``$a$ is not $b$".
-In these theories, $Id$ is therefore implemented with the following sum type that has only one constructor :
+Judgemental equality is included in propositional equality
+because what is equal by definition is provably equal. This is
+accomplished by giving a constructor for the type $Id(a,a)$ and no
+constructor when ``$a$ is not $b$".  In these theories, $Id$ is therefore
+implemented with the following type with one constructor:
 
@@
 
-The only way for $(\code{Id}_a\ x\ y)$ to be inhabited is therefore that $x$ should be
-(by definition) $y$. In this case, the constructor \code{Refl} helps to create a
-proof of this equality : \code{(Refl\ x)} is precisely the proof which says that
-$x=_ax$.
+The only way for $(\code{Id}_a\ x\ y)$ to be inhabited is therefore that $x$
+equals $y$. In this case, the constructor \code{Refl} helps
+to create a proof of this equality: \code{(Refl\ x)} is precisely the proof
+which says that $x=_ax$.
 
-When reading these definitions, one could think that the propositional equality is a simple wrapper of the judgemental equality.
-This is not the case : the propositional equality does not only contain the judgemental equality, because in these type theories, a principle of induction is associated with each inductive type.
-If \code{T} is an inductive type with a constant constructor and a recursive constructor, ie, \code{T = 1 + T}, defined in Idris as :
+The propositional equality does not only contain the judgemental equality,
+however, because a principle of induction is associated with each inductive
+type. If \code{T} is an inductive type with a constant constructor and a
+recursive constructor, ie, \code{T = 1 + T}, defined in Idris as:
 
@@
 
-The principles of induction can be used for proving any proposition $P$, so they also work for the propositional equality.
-For example, we can prove that $n+0 = n$ for all $n$ by induction on the $Nat\ n$, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on its first argument.
-That means that the axiom of induction has made the type $Id_{a}(x,y)$ bigger than it looks like : it contains terms that are not generated by \code{Refl} (the only constructor of $Id_{a}(x,y)$), but that are added by the inductive principles.
-There are therefore things which are provably equal which unfortunately are not definitionally equal after unfolding all the definitions and after full reductions.
-The activity of proving equalities is therefore in these theories something which isn't automatically decidable by the type-checker in the very general case.
-
-Proving that one term is equal to another is quite common
-in formal certification.
-This is even an everyday routine when working in a dependently typed language, and especially when intensively indexing types over values in order to capture some logical properties, as we will show in the following subsection.
+For example, we can prove that $n+0 = n$ for all $n$ by induction on the $Nat\
+n$, even if $n+0 \not\equiv n$ with the usual definition of $+$, recursive on
+its first argument. So, the axiom of induction means the type $Id_{a}(x,y)$
+contains not only the canonical form \code{Refl}, but also those added by
+inductive principles.  There are therefore things which are \emph{provably}
+equal, but not \emph{definitionally} equal.  Proving equalities is therefore in
+these theories something which isn't automatically decidable by the
+type-checker in the general case.
 
@@
 
-We revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which shows how proof obligations may arise when a type is indexed by natural numbers.
-In this development, the goal is to implement a formally verified library of binary numbers.
-In order to ensure the correctness of the methods that will be developed, we define the types \code{Bit} and \code{Binary} as indexed over the value they represent (expressed as a natural number), following a correct-by-construction approach.
+Proving that one term is equal to another is common in formal verification, and
+proof obligations arise naturally in dependently typed programming when
+indexing types over values in order to capture some logical properties.
+To demonstrate this,
+we revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which
+shows how proof obligations arise when a type is indexed by natural
+numbers. Our goal is to implement a verified
+library of binary numbers. To ensure functional correctness,
+we define types \code{Bit} and \code{Binary},
+indexed over the value they represent (expressed as a natural number):
 
@@
 
-We want to write a function performing the addition between two binary numbers. In order to do so, we start with an auxiliary function, which adds three bits
-(the third one plays the role of a carry flag), and produces the two bits of
-the result, where the first one is the most significant bit.
+We will write a function to add two binary numbers.
+To do so, we begin with an auxiliary function, which adds three bits
+(the third is a carry bit), and produces the two bits of
+the result, where the first is the more significant bit:
 
@@
 addBit b0 b0 b1 = (_ ** (_ ** (b0, b1, Refl)))
-addBit b0 b1 b0 = (_ ** (_ ** (b0, b1, Refl)))
-addBit b0 b1 b1 = (_ ** (_ ** (b1, b0, Refl)))
-addBit b1 b0 b0 = (_ ** (_ ** (b0, b1, Refl)))
-addBit b1 b0 b1 = (_ ** (_ ** (b1, b0, Refl)))
-addBit b1 b1 b0 = (_ ** (_ ** (b1, b0, Refl)))
-addBit b1 b1 b1 = (_ ** (_ ** (b1, b1, Refl)))
+{- ... remaining cases follow the same pattern ... -}
 \end{lstlisting}
 
-This function produces a dependent pair.
-The first component is the value \code{bX} (expressed as a natural number) representing the first bit produced (therefore of type \code{Bit\ bX}), and the second component is another dependent pair.
-This second dependent pair is composed of the value \code{bY} representing the second bit produced, and of a triple containing the two bits produced and a proof that these two bits correctly encode the result of the addition.
-For example, on the fifth line, which corresponds to the computation $1_2 + 0_2 + 0_2 = (01)_2$, the function produces the bit \code{b0} (encoding $0_2$) and the bit \code{b1} (encoding $1_2$), and a proof that $1 + 0 + 0 = 1 + (2*0)$.
+The syntax \code{(n ** t)} denotes a \emph{dependent pair}, where the 
+type of the second argument \texttt{t} can refer to the first argument
+\code{n}. So, we can read this type as: ``there exists a number \code{bX},
+and a number \code{bY}, such that we have two bits \code{Bit bX} and
+\code{Bit bY} and the sum of the input bits \code{c}, \code{x} and
+\code{y} equals \code{bY + 2 * bX}.''
+For example, on the second line,
+which corresponds to the computation $0_2 + 0_2 + 1_2 = (01)_2$,
+the function produces this bits \code{b0} and \code{b1}, and
+a proof that $0 + 0 + 1 = 1 + (2 \times 0)$.
 
-We finally define the function \code{adc} that adds two binary numbers and a carry flag.
-This addition works for two binary numbers expressed with the same number of bits, and it produces a result with one more bit.
-This result represents the sum of the three concrete values of the inputs.
-Therefore, we want to write :
+We then define the function \code{adc} that adds two binary numbers and a carry
+bit. This works for two binary numbers with the same number
+of bits, and produces a result with one more bit. We would like to write:
 
@@
 adc (numx # bX) (numy # bY) carry
-  = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) = addBit bX bY carry in
+  = let (vCarry0 ** (vLsb ** (carry0, lsb, _))) 
+            = addBit bX bY carry in
         adc numx numy carry0 # lsb
@@
 
-Unfortunately, this straightforward definition is rejected by Idris' type-checker because the types mismatch for both of the two patterns.
-The result of the first line \code{adc zero zero carry} is expected to have the type \codeL{Binary\ 1\ (plus\ (plus\ c\ 0)\ 0)} but we provide a term of type \code{Binary\ 1\ (plus\ c\ (plus\ 0\ (plus\ 0\ 0)))}.
-The problem is even worse for the second case where the expected index is \codeL{(plus\ (plus\ c\ (plus\ bit2\ (plus\ v1\ (plus\ v1\ 0))))\ (plus\ bit\ (plus\ v\ (plus\ v\ 0))))}
+Unfortunately, this definition is rejected because the types of each side do
+not match.  The result of the first line \code{adc zero zero carry} is expected
+to have the type:
+
+\codeL{Binary\ 1\ ((c\ + 0)\ + 0)} but we provide a
+term of type:
+
+\code{Binary\ 1\ (c\ + (0\ + (0\ + 0)))}.
+
+The problem is even worse for the second case where the expected index is:
+
+\codeL{((c\ + (bit2\ + (v1\ + (v1\ + 0)))) + (bit\ + (v\ + (v\ + 0))))}
 while we're trying to provide a term indexed over
-\codeL{plus\ vLsb\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ (plus\ (plus\ (plus\ vCarry0\ v1)\ v)\ 0))}.
-We know that the definition of \code{adc} we have given is correct, and it has \emph{provably} the expected type, but it does not have it \emph{immediately} or \emph{by reduction} : after full reductions\footnote{In dependently typed theories, the presence of types predicated on values creates the need of using reductions also inside of types}, the expected and provided types are still different, and Idris can not figure out a proof of compatibility on its own.
-If we use provisional definitions~\cite{DBLP:conf/plpv/Brady13}, the previous definition can be accepted by the system, but we now have two proof obligations \code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality between the two types, and they have to be provided by hand --here with a proof script-- :
+
+\codeL{vLsb\ + (((vCarry0\ + v1) + v)\ + (((vCarry0\ + v1)\ + v)\ + 0))}.
+
+The definition of \code{adc} we have given would behave correctly, and it has
+\emph{provably} the expected type, but it does not have it \emph{immediately}
+or \emph{judgementally}: after full reductions the expected and provided types are still
+different.
+If we use provisional definitions~\cite{DBLP:conf/plpv/Brady13}, the previous
+definition can be accepted by the system, but we now have two proof obligations
+\code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality
+between the two types, and they have to be provided by hand.
+For example, using a proof script:
 
@@
 
-As we see, this kind of proof consists of a potentially long sequence of rewriting steps, where each step uses one of the available properties: neutral element, commutativity, associativity.
-Without some specific automation, this sequence of rewritings must be done by the programmer.
-This is time consuming and more importantly, just a small change in the left or the right hand side of the equality may invalidate the proof.
-That means that a minor change in our datatypes, or in the definition of \code{addBit} or \code{adc} will require us to replace this proof by another one fulfilling the same task.
-These proofs become the everyday routine in any dependently-typed language, as soon as we index types over values in order to capture some logical properties.
+As we see, this kind of proof consists of a potentially long sequence of
+rewriting steps, each using one of the available properties: neutral element,
+commutativity, associativity. Without some automation, this sequence of
+rewritings must be done by the programmer.  Not only is this time consuming,
+just a small change in the definition may lead to a different proof obligation,
+thus invalidating the proof. 
+A minor change in the datatype, or 
+the definition of \code{addBit} or \code{adc} will require us
+to do a new proof, and thus, without support from the machine, these
+proofs
+could become the everyday routine in any dependently-typed language.
 
-The proof that we have written by hand only uses the existence of a neutral element, and the associativity and commutativity of $+$ on \code{Nat}.
-Thus, we're rewriting a term by using the properties of a commutative monoid.
-We would like to let a generic prover for commutative monoid find a proof on its own.
+Our handwritten proof uses only the existence of a neutral
+element, and the associativity and commutativity of $+$ on \code{Nat}. Thus,
+we're rewriting a term by using the properties of a commutative monoid. We
+would like a generic prover for commutative monoids to find a proof 
+automatically.
 
@@
 
-Prover for some algebraic structures have already been implemented for various proof assistants, including Coq~\cite{Coq2005} and Agda\footnote{\url{http://wiki.portal.chalmers.se/agda\%5C?n=Libraries.UsingTheRingSolver}} ring prover.
-In this paper, we describe a certified
-implementation\footnote{The implementation of our hierarchy of tactics can be found online at \url{https://github.com/FranckS/RingIdris/Provers}} of an automatic prover for equalities in a \emph{hierarchy} of algebraic
-structures, including monoid, groups and rings (all potentially commutative),
-for the Idris language.
-
-The principal novelty is in the approach that we follow, using a
-new kind of \emph{type-safe reflection}.
-Working by reflection for implementing tactics has been done several times, including for the implementation of a ring solver for the Coq proof assistant, but without the type-safe and correct-by-construction aspects.
-We will
-compare our approach with other implementations in section~\ref{sect:relatedWork}.
-
-The contributions of this paper are the following :
+Provers for some algebraic structures have already been implemented for various
+proof assistants, including Coq~\cite{Coq2005} and
+Agda\footnote{\url{http://wiki.portal.chalmers.se/agda\%5C?n=Libraries.UsingTheRingSolver}}. 
+In this paper, we describe an implementation\footnote{The
+implementation of our hierarchy of tactics can be found online at
+\url{https://github.com/FranckS/RingIdris/Provers}} of an automatic prover for
+equalities in a \emph{hierarchy} of algebraic structures, including monoids,
+groups and rings (all potentially commutative), for the Idris language.,
+making the following contributions:
 
@@
 
-\item We present in section~\ref{sect:typeSafeReflection} a new type-safe reflection mechanism, where the reflected terms are indexed over the concrete inputs, thus providing a direct way to pull out the proofs, and providing the guarantee that the reflection of a term is indeed a faithful representation of the term.
+\item We present a type-safe reflection mechanism
+(section~\ref{sect:typeSafeReflection}), where the reflected terms are indexed
+over the concrete terms, providing a direct way to extract proofs and
+guaranteeing that the reflected term is a sound representation.
 
-\item The normalisation procedures are implemented by following a correct by construction approach, instead of proving the correctness afterwards with auxiliary lemmas.
-This approach is much more suitable for programming languages like Idris.
-This correct-by-construction aspect is developed in section~\ref{sect:correctByConstruction}
+\item The normalisation procedures are implemented by following a correct by
+construction approach (section~\ref{sect:correctByConstruction}), instead of proving the correctness afterwards with
+auxiliary lemmas. This approach is much more suitable for programming languages
+like Idris.
 
-\item We develop a hierarchy of tactics (and not just a tactic for, say, ring) where each tactic reuses the rewriting machinery of the structure from which it naturally inherits.
-For example, simplifying neutral elements is only implemented at the monoid level, and each level inheriting from it will reuse it.
-Reusability is difficult to obtain when we want to reuse the prover of a less expressive structure.
-For example, reusing the monoid prover for building the
-group prover is not trivial, since we lose the possibility to express
-negations ($-x$) and subtractions ($x-y$). Some specific encodings
-will be presented in section~\ref{sect:reusabilityOfTheProvers} to overcome this problem.
+\item We develop a \emph{hierarchy} of tactics where each tactic reuses the
+rewriting machinery of the structure from which it inherits.  For
+example, simplifying neutral elements is implemented only at the monoid level,
+and reused at other levels.
+It is challenging
+to reuse the prover of a less expressive structure; for
+example, using the monoid prover to build the group prover is 
+tricky because
+we lose the ability to express negations ($-x$) and
+subtractions ($x-y$). We present encodings
+(section~\ref{sect:reusabilityOfTheProvers}) to overcome this problem.
 
 \end{enumerate}
+
+The principal novelty is in using 
+\emph{type-safe reflection}.  Working by reflection for implementing tactics
+has been done several times, including the implementation of a ring solver
+for Coq, but without the type-safety and
+correctness-by-construction. We compare our approach with other
+implementations in section~\ref{sect:relatedWork}.

### 2017-05-16 paper-CICM-17/paper.tex
Prune abstract and section 1

@@
 \title{Automatically Proving Equivalence by Type-Safe Reflection}
-\subtitle{Or how to use dependent types to address some of the problems they have caused}
+%\subtitle{Or how to use dependent types to address some of the problems they have caused}
 %
@@
 \begin{abstract}
-One common difficulty with programming and reasoning with dependent types is that proof obligations arise naturally once programs become even moderately sized.
-For example, implementing an adder for binary numbers indexed over their natural number equivalents will naturally lead to proof obligations for equalities of expressions over universally quantified natural numbers.
-The need of these proofs of equalities comes, in intensional type theories, from the fact that the propositional equality enables to prove as equal terms that are not judgmentally equal, which means that the typechecker can't always obtain equalities by reduction, and therefore often needs proofs before accepting definitions.
-As far as possible, we would like to solve such proof obligations automatically.
-In this paper, we show one way to automate these proofs by reflection for the dependently-typed programming language Idris.
-We will show how type-safe reflection, where reflected terms are indexed by the original Idris expression, leads to straightforward construction and manipulation of proofs.
-We build a hierarchy of reflexive tactics for proving equivalences in semi-groups, monoids, commutative monoids, groups, commutative groups, semi-rings and rings.
-We will also show how each tactic reuses the ones from the simplest structures, thus avoiding as much as possible the duplication of code and proofs.
+One difficulty with programming with dependent types is that proof obligations
+arise naturally once programs become even moderately sized. For example,
+implementing an adder for binary numbers indexed over their natural number
+equivalents naturally leads to proof obligations for equalities of expressions
+over natural numbers. The need for these equality proofs comes, in intensional
+type theories, from the fact that the propositional equality enables us to
+prove as equal terms that are not judgementally equal, which means that the
+typechecker can't always obtain equalities by reduction.  As far as possible,
+we would like to solve such proof obligations automatically. In this paper, we
+show one way to automate these proofs by reflection in the dependently typed
+programming language Idris.  We show how defining reflected terms indexed by
+the original Idris expression allows us to construct and manipulate proofs. We
+build a hierarchy of tactics for proving equivalences in semi-groups, monoids,
+commutative monoids, groups, commutative groups, semi-rings and rings. We also
+show how each tactic reuses those from simpler structures, thus avoiding
+duplication of code and proofs.
 

### 2017-05-16 paper-CICM-17/content/1-introduction.tex
Pruning up to 2.3; add acknowledgements

@@
 
+% EB - I took out a step because without the whole proof, all we really
+% need to see is some representative examples of rewritings
 \begin{lstlisting}
@@
     [...]
-    rewrite sym (plusAssociative (plus c (plus bit0 (plus v v))) 
-                 bit1 (plus v1 v1));
     rewrite (plusAssociative c (plus bit0 (plus v v)) bit1);
@@
 
-As we see, this kind of proof consists of a potentially long sequence of
-rewriting steps, each using one of the available properties: neutral element,
+Such proofs consist of a potentially long sequence of
+rewriting steps, each using one of the properties: neutral element,
 commutativity, associativity. Without some automation, this sequence of
 rewritings must be done by the programmer.  Not only is this time consuming,
-just a small change in the definition may lead to a different proof obligation,
+but a small change in the definition may lead to a different proof obligation,
 thus invalidating the proof. 
@@
 construction approach (section~\ref{sect:correctByConstruction}), instead of proving the correctness afterwards with
-auxiliary lemmas. This approach is much more suitable for programming languages
-like Idris.
+auxiliary lemmas. 
+% This approach is much more suitable for programming languages like Idris.
 

### 2017-05-16 paper-CICM-17/content/3-hierarchyOfTactics.tex
Pruning up to 2.3; add acknowledgements

@@
 
-What we want to build is a prover able to prove not only equalities for \code{Nat}, \code{List}, or any specific type, but for various datatypes, and for various properties.
-With some abstraction, we can solve the problem of generating proof of equality for many datatypes like natural numbers, lists, matrices, and many others \emph{all at once} by implementing a very generic hierarchy of provers that can prove equalities of terms for semi-groups, monoids, commutative-monoids, groups, commutative-groups, semi-rings and rings.
-These tactics will be usable on any type that satisfies the properties of one of these algebraic structure.
-The properties of a given algebraic algebraic structure will be expressed in a corresponding interface (sometimes called typeclass in other progamming languages).
-This interface will extend the interface from which it naturally inherits.
-For example, the \code{Group} interface will extend the \code{Monoid} one.
-We will end up with a hierarchy of interfaces and will write one tactic for each of these interface.
-At every level of the hierarchy, we will be able to work on any type, being given that an implementation of the corresponding interface is provided.
+We aim to build a prover not only for equalities on \code{Nat}, \code{List}, or
+any specific type, but for generic datatypes and properties. Using
+the right
+abstraction, we can generate proofs of equalities for
+many datatypes \emph{at once} by implementing a generic hierarchy of provers
+for semi-groups, monoids,
+commutative-monoids, groups, commutative-groups, semi-rings and rings. 
+The
+properties of an algebraic structure are expressed in an
+interface (an interface in Idris is similar to a type class
+in Haskell). This interface will extend the interface from which it
+inherits; for example, \code{Group} extends \code{Monoid}.
+This leads to a hierarchy of interfaces, with one tactic
+for each. At every level of the hierarchy, we will be able
+to work on any type, as long as there is a corresponding implementation of
+the interface.
 
@@
 
-With some reasonable additional effort, we could produce a collection of tactics for proving equivalences instead of only equalities.
-The machinery would be very similar, and if we do the right abstraction from the start, we could gain one more degree of genericity, with the freedom of choosing the equivalence relation (which can of course be the usual equality).
-The user could define his own equivalence notion, as long as he's able to provide the proofs of the properties of the algebraic structure he wants to use.
-Let's call \code{c} the \emph{carrier} type, ie, the type on which we want to prove equivalences.
-The equivalence relation on \code{c} has the following profile $(\simeq) : \code{c} \rightarrow \code{c} \rightarrow \code{Type}$\footnote{This \code{Type} would be a \code{Prop} in systems --like Coq-- that make a distinction between the world of computations and the world of logical statements}, and has to be accompanied by the usual properties of reflexivity, symmetry and transitivity.
+With some additional effort, we can produce a collection of tactics for proving
+\emph{equivalences}, rather than only equalities. 
+The machinery is very similar and
+we gain another
+degree of genericity, with the freedom of choosing the equivalence relation
+(which can be the usual equality). 
+The user can define their own
+notion of equivalence, as long as they provide the proofs of the
+properties of the relevant algebraic structure. Let's call \code{c} the
+\emph{carrier} type, i.e., the type on which we want to prove equivalences. 
+The
+equivalence relation on \code{c} has the following profile $(\simeq) : \code{c}
+\rightarrow \code{c} \rightarrow \code{Type}$\footnote{This \code{Type} would
+be a \code{Prop} in systems, like Coq, that make a distinction between the
+world of computations and the world of logical statements}, and must be
+be reflexive, symmetric and transitive.
 
-All of our tactics will require to have a way of testing this equivalence between elements of the underlying set, that is to say, a way to test equivalence between constants.
-For this reason, we define a notion of \code{Set}\footnote{This notion of set isn't a formalisation of sets and their elements, but is only a way to talk about the carrier type and an equivalence relation, sometimes called Setoid}, which only requires the definition of the equivalence relation and this equivalence test \code{set\_eq}.
-All the interfaces representing the algebraic structures will later extend \code{Set} :
+Our tactics need to be able to test this equivalence
+between elements of the underlying set, that is a way of testing
+equivalence of constants. We therefore define a notion of
+\code{Set}\footnote{This notion of set 
+is a way to talk about the carrier type and an equivalence
+relation, sometimes called Setoid}, which requires the definition of the
+equivalence relation and an equivalence test \code{set\_eq}. All the interfaces
+representing the algebraic structures will later extend \code{Set}:
 
@@
     ($\simeq$) : c $\rightarrow$ c $\rightarrow$ Type
-    refl : (x:c) $\rightarrow$ x $\simeq$ x
-    sym : {x:c} $\rightarrow$ {y:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ x)
-    trans : {x:c} $\rightarrow$ {y:c} $\rightarrow$ {z:c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)    
-    set_eq : (x:c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
+    refl : (x : c) $\rightarrow$ x $\simeq$ x
+    sym : {x, y : c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ x)
+    trans : {x, y, z : c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)    
+    set_eq : (x : c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
 \end{lstlisting}
 
-If the user wants to prove propositional equalities, they will instantiate $(\simeq)$ with the built-in $(=)$ during the definition of the \code{Set} implementation.
-Note that $(\simeq)$ is only weakly decidable in the sense that \code{set\_eq} only produces a proof when the two elements are equivalent, but it doesn't produce a proof of dis-equivalence when they are different -- instead, it simply produces the value \code{Nothing}--.
-That is normal, as we want to generate proof of equivalence, and not to generate counter examples for proving dis-equivalence, which is another problem.
-
-There is no tactic associated to \code{Set}, since we have no operations and no properties associated to this structure.
-Therefore, equivalences in a \code{Set} are just the ``syntactic equivalences", and they can simply be proven with \code{refl}\footnote{\code{refl} should not be confused with \code{Refl}, the only constructor of $=$.
-But of course, when $(\simeq)$ is instantiated with the equality $=$, the instantiation of \code{refl} will be \code{Refl}.
-Therefore, \code{refl} of the interface \code{Set} is a generalisation of \code{Refl}}.
+To prove propositional equalities, the user instantiates
+$(\simeq)$ with the built-in $(=)$ when implementing \code{Set}.
+Note that $(\simeq)$ is only weakly decidable in the sense
+that \code{set\_eq} only produces a proof when the two elements are equivalent,
+but it doesn't produce a proof of dis-equivalence when they are different,
+instead producing the value \code{Nothing}. Our goal is only to generate
+proofs of equivalance, not to produce counter-examples.
+%
+There is no tactic associated with \code{Set}, since we have no operations or
+properties associated to this structure. Equivalences in a
+\code{Set} are ``syntactic equivalences'' and can be
+proven with \code{refl}\footnote{\code{refl} is not to be confused with
+\code{Refl}, the constructor of $=$, but when $(\simeq)$ is
+instantiated with the equality $=$, \code{refl} is implemented by
+\code{Refl}. Therefore, \code{refl} of the interface \code{Set} is a
+generalisation of \code{Refl}}.
 
@@
 
-Working with equivalences instead of equalities will bring one complication.
-These proofs of correctness that we will have to produce by hand will be slightly more complicated, as we won't be able to use Idris' ``rewrite" command, that enables to rewrite a subterm of a term by another one, provided that the two subterms are equal.
-More precisely, if we need to prove $P\ x$, and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in order to transform the goal into $P\ x'$, as we would normally do if the available proof would be an equality $x=x'$.
-This is a classical problem of working within a setoid, which is slightly mitigated when the programming language offers a good support for rewriting terms in setoids.
-However, Idris isn't equipped with any support for setoids, and everything will have to be done by hand.
-For this reason, we define the lemma :
+Working with equivalences instead of equalities brings one complication: Proofs
+of correctness that we produce by hand cannot use Idris' ``rewrite'' mechanism,
+which enables rewriting of a subterm by another one, provided that the two
+subterms are propositionally equal. 
+%
+% EB - I think we can skip this detail
+%If we need to prove $P\ x$,
+%and have a proof of equivalence $pr:x \simeq x'$, we can't use ``rewrite pr" in
+%order to transform the goal into $P\ x'$, as we would normally do if the
+%available proof would be an equality $x=x'$. 
+%
+This is a classical problem of
+working within a setoid, which can be mitigated by programming
+language support for rewriting terms in setoids. However, Idris
+is not equipped with any such support. For this reason, we define the 
+following lemma, using the methods of the \code{Set} interface:
 
@@
 
-\begin{proof}
-We have $x \simeq c1$ and we have $c1 \simeq c2$, therefore we have $x \simeq c2$ by the use of the axiom of transitivity (of the interface \code{Set}). \
+% EB - I took this out because, as the reviewer says, it's odd to have a
+% hand proof here, and I don't think the code itself is interesting enough to
+% include since it's a fairly routine use of the interface
 
-We also have $y \simeq c2$ and therefore we have $c2 \simeq y$ by the use of the symmetry axiom (of the interface \code{Set}). \
+%\begin{proof}
+%We have $x \simeq c1$ and we have $c1 \simeq c2$, therefore we have $x \simeq c2$ by the use of the axiom of transitivity (of the interface \code{Set}). \
 
-Now that we have these fresh proofs of $x \simeq c2$ and of $c2 \simeq y$, we can use one last time the property of transitivity, in order to get a proof of $x \simeq y$.
-\qed
-\end{proof}
+%We also have $y \simeq c2$ and therefore we have $c2 \simeq y$ by the use of the symmetry axiom (of the interface \code{Set}). \
 
-This lemma says that the equivalence preserves the equivalence, which means that in order to prove $x \simeq y$, we can prove the (hopefully) simplest problem $c1 \simeq c2$, provided that $x \simeq c1$ and that $y \simeq c2$.
-This lemma will be intensively used for building the proofs of correctness of the normalisation functions that will help to build the desired proof of $x \simeq y$.
-It will be basically used as a replacement to the ``rewrite" tactic that we can no longer use since we are dealing with equivalence and not equality.
+%Now that we have these fresh proofs of $x \simeq c2$ and of $c2 \simeq y$, we can use one last time the property of transitivity, in order to get a proof of $x \simeq y$.
+%\qed
+%\end{proof}
+
+This lemma says that the equivalence preserves the equivalence, which means
+that in order to prove $x \simeq y$, we can prove a smaller
+problem $c1 \simeq c2$, provided that $x \simeq c1$ and that $y \simeq c2$.
+We will use this lemma extensively.
 
@@
 
-The operations, constants and properties of each algebraic structure will be described in a dedicated interface.
-The first structure, completely trivial, is the magma.
-A magma is a structure built on top of \code{Set}, that just adds a \code{Plus} operation, and no specific properties about it :
+We describe operations, constants and properties of each algebraic structure
+in an interface. The first structure is
+the magma. which is a structure built on top of \code{Set} that adds
+\code{Plus} operation, and no specific properties:
 
@@
 
-This code means that a type \code{c} (for $carrier$) is a \code{Magma} if it is already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$ and the equivalence test \code{set\_eq}), and if it has a $+$ operation.
-In fact, there is an additional requirement that will apply to all operators (in this case, the $+$ operation), which is that they need to be ``compatible" with the equivalence relation, which is expressed by the following axiom for $+$ : \\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow \{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow \{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2 \simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$ \\
-This requirement comes from the fact that we're dealing with any equivalence relation.
-The user is free to use his own equivalence relation, but it should be compatible with the operations that he is using.
-As it was the case for \code{Set}, there is no tactic to write for \code{Magma}, because there is no property at all : all equivalences are again syntactic equivalences, and they can all be proven by \code{refl}.
+This code means that a type \code{c} (for $carrier$) is a \code{Magma} if it is
+already a \code{Set} (ie, it is equipped with the equivalence relation $\simeq$
+and the equivalence test \code{set\_eq}), and if it has a $+$ operation.  In
+fact, there is an additional requirement that will apply to all operators (in
+this case, the $+$ operation), which is that they need to be ``compatible" with
+the equivalence relation, which is expressed by the following axiom for $+$:
+\\ $\code{Plus\_preserves\_equiv} : \{c:\code{Type}\} \rightarrow
+\{\code{Magma}\ c\} \rightarrow\{c1:c\} \rightarrow \{c2:c\} \rightarrow
+\{c1':c\} \rightarrow \{c2':c\} \rightarrow (c1 \simeq c1') \rightarrow (c2
+\simeq c2') \rightarrow ((c1+c2) \simeq (c1'+c2'))$
+
+We have this requirement because we support
+any equivalence relation.
+As with \code{Set}, there is no
+tactic for \code{Magma}, because there is no property; all
+equivalences are again syntactic equivalences, and can thus be proven by
+\code{refl}.
 
@@
 interface Magma c => SemiGroup c where
-    Plus_assoc : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ (c3:c) $\rightarrow$ ((c1+c2)+c3 $\simeq$ c1+(c2+c3))
+  Plus_assoc : (c1 : c) $\rightarrow$ (c2 : c) $\rightarrow$ (c3 : c) $\rightarrow$ 
+               ((c1 + c2) + c3 $\simeq$ c1 + (c2 + c3))
 \end{lstlisting}
 
-As example of magma, we can cite \code{Nat} equipped with its operation of addition and \code{List} with the operation of concatenation.
-
-The monoid structure is a semi-group with the property of neutral element for a distinguished element called \code{Zero}.
+\noindent
+Examples of magma are \code{Nat} equipped with addition, and \code{List} with
+concatenation.  Next, a monoid is a semi-group with a ``neutral``, or
+distinguished, element called \code{Zero}.
 
@@
   Zero : c    
-    Plus_neutral_1 : (c1:c) $\rightarrow$ (Zero+c1 $\simeq$ c1)    
-    Plus_neutral_2 : (c1:c) $\rightarrow$ (c1+Zero $\simeq$ c1)
+  Plus_neutral_1 : (c1 : c) $\rightarrow$ (Zero + c1 $\simeq$ c1)    
+  Plus_neutral_2 : (c1 : c) $\rightarrow$ (c1 + Zero $\simeq$ c1)
 \end{lstlisting}
 
-The hierarchy of interfaces continues, but we only show the \code{Group} interface for reasons of space :
+\noindent
+The hierarchy of interfaces continues with \code{Group}:
 
@@
   Neg : c $\rightarrow$ c
-    Minus_simpl : (c1:c) $\rightarrow$ (c2:c) $\rightarrow$ Minus c1 c2 $\simeq$ c1 + (Neg c2) 
-    Plus_inverse : (c1:c) $\rightarrow$ (c1 + (Neg c1) $\simeq$ Zero, (Neg c1) + c1 $\simeq$ Zero)
+  Minus_simpl : (c1 : c) $\rightarrow$ (c2 : c) $\rightarrow$ Minus c1 c2 $\simeq$ c1 + (Neg c2) 
+  Plus_inverse : (c1 : c) $\rightarrow$ (c1 + (Neg c1) $\simeq$ Zero, 
+                              (Neg c1) + c1 $\simeq$ Zero)
 \end{lstlisting}
 
-The notion of group uses two new operations (\code{Neg} and \code{Minus}), but \code{Minus} is not primitive and can be simplified with $+$ and \code{Neg}.
-The important property of a group is that every element \code{c1} must admit \code{Neg\ c1} as inverse element for the $+$ operation.
+The notion of group uses two new operations (\code{Neg} and \code{Minus}), but
+\code{Minus} can be simplified with $+$ and \code{Neg}.
+The important property of a group is that every element \code{c1} must admit
+\code{Neg\ c1} as inverse element for $+$.
+For reasons of space, we elide the remaining details of the hierarchy.
 
@@
 
-When trying to prove an equivalence $x \simeq y$, the universally-quantified variables have been abstracted, and they are now part of the context.
-Our tatics will work by normalising both sides of the ``potential equivalence" $x \simeq y$, and afterwards comparing them with a syntactic equality test.
-The problem is that such a normalisation function will need to do different treatments for a a variable, for a constant and for an operator.
-For this reason, we will need to work by reflection.
-There will be one type of reflected terms for each algebraic structure.
-The novelty is not the use of reflected terms (this has already been done, as discussed in section~\ref{sect:relatedWork}), but is the use of a type-safe reflection mechanism where we index the reflected terms by the concrete value that they represent.
-Each of these datatype is parametrised over a type \code{c}, which is the real type on which we want to prove equalities (the \emph{carrier} type).
-It is also indexed over an implementation of the corresponding interface for \code{c} (we usually call it \code{p}, because it behaves as a \emph{proof} telling that the structure \code{c} has the desired properties), indexed over a context of abstracted variables (a vector $\Gamma$ of \code{n} elements of type \code{c}), and most important it is also indexed over a value of type \code{c}, which is precisely the concrete value being encoded.
+When proving an equivalence $x \simeq y$, the universally-quantified
+variables are abstracted. Our
+tatics normalise both sides of the ``potential equivalence'' $x
+\simeq y$, and compare the results by syntactic equality. The
+difficulty is that the normalisation function needs to consider
+variables, constants and operators. For this
+reason, we work by reflection. We have one type of reflected
+terms for each algebraic structure. The novelty is not the use of reflected
+terms, but the use of a type-safe reflection
+mechanism where we index the reflected terms by the concrete value that they
+represent. Each of these datatype is parametrised over a type \code{c}, which
+is the type on which we want to prove equalities (the \emph{carrier}
+type). It is also indexed over an implementation of the corresponding interface
+for \code{c} (we usually call it \code{p}, because it behaves as a \emph{proof}
+telling that the structure \code{c} has the desired properties), indexed over a
+context of abstracted variables (a vector $\Gamma$ of \code{n} elements of type
+\code{c}). Most importantly, it is indexed over a value of type \code{c},
+which is the concrete value being encoded.
 
-A magma is only equipped with the operation of addition. Thus, we only have three concepts to express in order to reflect terms in a magma : constants, variables, and additions.
+A magma is equipped with one operation, addition. Thus, to reflect terms in a
+magma we express constants, variables, and addition:
 
@@
 
-For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we will say that $e_x$ denotes (or encodes) the value $x$ in the context $\Gamma$.
-When an expression is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted value is the corresponding variable in the context, ie, \code{(index\ i\ $\Gamma$)}.
-The expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type $c$.
-Finally, if $e_x$ is an expression encoding the concrete value $x$, and $e_y$ is an expression encoding the concrete value $y$, then the expression \code{PlusMa\ $e_x$\ $e_y$} denotes the concrete value $(x + y)$.
-Because the reflected terms embed their corresponding inputs, they are guarantee to be faithful representations of them.
+For an expression $e_x\ :\ \code{ExprMa}\ \Gamma\ x$, we say that $e_x$
+denotes the value $x$ in the context $\Gamma$.  When an expression
+is a variable \code{(VarMa\ \_\ $\Gamma$\ i)}, the denoted value is the
+corresponding variable in the context, i.e., \code{(index\ i\ $\Gamma$)}.  The
+expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type
+$c$. Finally, if $e_x$ is an expression encoding the concrete value $x$, and
+$e_y$ is an expression encoding the concrete value $y$, then the expression
+\code{PlusMa\ $e_x$\ $e_y$} denotes the concrete value $(x + y)$. Because the
+reflected terms embed their corresponding inputs, they are guaranteed to be
+sound representations.
 
-There is no additional operations in semi-groups and monoids, so the datatypes that reflect terms in these two structures will have exactly the same shape as the one for magma that we have given above.
-However, the datatype for reflected terms in groups will have to introduce two new constructors for the \code{Neg} and \code{Minus} operations :
+There are no additional operations in semi-groups or monoids, so the reflected
+datatypes have the same shape as that for magma.  However, the datatype for
+reflected terms in groups introduces two new constructors for \code{Neg}
+and \code{Minus}:
 

### 2017-05-16 paper-CICM-17/content/3-hierarchyOfTactics.tex
Finish pruning section 2

@@
 
-The index of type $c$ (the real value encoded by an expression) is always expressed by using the lookup function \code{index} and the available operations in the implementation $p$, which for a group are $+$, \code{Minus} and \code{Neg}.
+The index of type $c$ (the value encoded by an expression) is always expressed
+by using the lookup function \code{index} and the available operations in the
+implementation $p$, which for a group are $+$, \code{Minus} and \code{Neg}.
 
-	\subsection{A correct by construction approach}
+\subsection{A correct-by-construction approach}
 \label{sect:correctByConstruction}
 
-We want to write the normalisation function on a \emph{correct by construction} way, which means that no additional proof should be required after the definition of the function.
-Thus, \code{norm} will produce the proof that the new expression freshly produced has the same interpretation as the original expression, and this will be made easier by the fact that all the datatypes for reflecting terms (\code{ExprMa}, \code{ExprG}, \code{ExprR}, etc) are indexed over the real --concrete-- value : a term of type $\code{Expr}\ \Gamma\ x$ is the encoding of the concrete value $x$ in the context $\Gamma$.
-Thus, for each algebraic structure, the type of the function \code{norm} will look like this : \\
-$\code{norm} :\ \code{Expr}\ \Gamma\ x\ \rightarrow\ (x'\ \code{**}\ (\code{Expr}\ \Gamma\ x',\ x\ \simeq\ x'))$ \\
-Every function of normalisation \code{norm} will produce a dependent pair : the new concrete value $x'$, and a pair made of an $\code{Expr}\ \Gamma\ x'$ which is the new encoded term indexed over the new concrete value we have just produced, and a proof that old and new concrete values $x$ and $x'$ are equivalent.
+We take a \emph{correct-by-construction} approach to implementing
+normalisation, which means that no additional proof will be required after
+defining normalisation.
+The normalisation function \code{norm} produces a new expression,
+and a proof that it
+has the same interpretation as the original.
+This will be enforced by the fact that all the datatypes for
+reflected terms (\code{ExprMa}, \code{ExprG}, \code{ExprR}, etc) are indexed
+over the concrete value: a term of type $\code{Expr}\ \Gamma\ x$ is
+the encoding of the concrete value $x$ in the context $\Gamma$.  For each
+structure, the type of \code{norm} has
+the following form:
+\\ $\code{norm} :\ \code{Expr}\ \Gamma\ x\ \rightarrow\ (x'\ \code{**}\
+(\code{Expr}\ \Gamma\ x',\ x\ \simeq\ x'))$ \\ 
+Every instance of
+\code{norm} produces a dependent pair: the new concrete value $x'$, and a
+pair made of an $\code{Expr}\ \Gamma\ x'$ which is the new encoded term indexed
+over the new concrete value we have just produced, and a proof that old and new
+concrete values $x$ and $x'$ are equivalent.
+This proof of $x \simeq y$ is the crucial component which allows us to
+automatically produce proofs of equivalences.
 
-This function doesn't only produce an $\code{Expr}\ \Gamma\ x$ (with an internal conversion of type) because the proof of $x \simeq x'$ is an essential component for building the desired proof of $x \simeq y$, which is the main problem that we are trying to solve.
+We will explain how to implement the tactic for \code{Group} specifically,
+using a normalisation function \code{groupNormalise},
+but the other structures are implemented similarly. 
+The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are
+elements of the type $c$, which implements a group. The reflected term for $x$
+is denoted $e_x$, and has type $\code{ExprG}\ p\
+\Gamma\ x$, which means that $e_x$ is guaranteed to be the encoding of $x$.  
+Similarly,
+$y$, is encoded by $e_y$, and its type is indexed by the
+value $y$.  Evaluating \code{norm} on $e_x$ produces the
+normal form $e_{x'}$ of type $\code{ExprG}\ p\ \Gamma\ x'$ and a proof $p_x$ of
+$x \simeq x'$.
+Similarly, for $e_y$, it produces
+the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof
+$p_y$ of $y \simeq y'$.  It now suffices to compare $e_{x'}$ and $e_{y'}$
+using a syntactic equality test \code{exprG\_eq} on the normal forms.
 
-We will explain how the tactic works by commenting the construction of the tactic for groups, but the other algebraic structures will be equipped similarly.
-The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are elements of the type $c$, which  simulates a group.
-The reflected term for $x$ will be denoted $e_x$, and this term will have the type $\code{ExprG}\ p\ \Gamma\ x$, which means that $e_x$ is guaranteed to be the encoding of $x$ because $e_x$ is precisely indexed over the real value $x$.
-We've got similar things for $y$, which is encoded by $e_y$, and its type is indexed over the real value $y$.
-Running the normalisation procedure on $e_x$ will produce the normal form $e_{x'}$ of type $\code{ExprG}\ p\ \Gamma\ x'$ and a proof $p_x$ of $x \simeq x'$ while running the normalisation procedure on $e_y$ will produce the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof $p_y$ of $y \simeq y'$.
-It is now enough to compare $e_{x'}$ and $e_{y'}$ using a standard syntactic equality test \code{exprG\_eq} because these two expressions are now supposed to be in normal form.
-
-Now, if the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then they necessary have the same type\footnote{We are working with the heterogeneous equality JMeq by default in Idris, but as always, the only way to have a proof of a:A = b:B is when A=B}, and therefore $x'=y'$, which implies $x' \simeq y'$ because the propositional equality is more restrictive than an equivalence relation.
-By using the two equivalences $x \simeq x'$ and $y \simeq y'$ that we obtained during the normalisations in the new equivalence $x' \simeq y'$ with the use of \code{eq\_preserves\_eq}, we can get a proof of $x \simeq y$.
-This is what the function \code{buildProof} is doing :
+If the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then
+they necessarily have the same type\footnote{We work with the
+heterogeneous equality by default in Idris, but as always, the only way to
+have a proof of \code{a:A = b:B} is when \code{A = B}}, and therefore $x'=y'$, which implies
+$x' \simeq y'$ because the propositional equality is more restrictive than an
+equivalence relation.  By using the two equivalences $x \simeq x'$ and $y
+\simeq y'$, and the new equivalence
+$x' \simeq y'$, we can use \code{eq\_preserves\_eq}, to get the desired proof of
+$x \simeq y$. We put all of this together in a function
+\code{buildProofGroup}:
 
@@
 
-The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected left hand size of the equivalence, which represents the value $x'$.
-Before the normalisation, the reflected LHS was reflecting the value $x$.
-The same applies to the right-hand side.
-The function also expects proofs of $x \simeq x'$ and of $y \simeq y'$, that we will be able to provide because the normalisation function has produced them while it was building the normalised reflected terms.
+The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected
+left hand size of the equivalence, which represents the value $x'$. 
+The function also expects proofs of $x \simeq x'$ and
+of $y \simeq y'$, which are built by the normalisation process.
 
-Finally, the main function which tries to prove the equivalence $x \simeq y$ has to normalise the two reflected terms encoding the left and the right hand side, and to use the function \code{buildProof} in order to compose the two proofs that have been obtained :
+Finally, the main function which tries to prove the equivalence $x \simeq y$
+has to normalise the two reflected terms encoding the left and the right hand
+side, and use the function \code{buildProof} to compose the two
+proofs:
 
@@
 
-The remaining part is to define the function \code{norm}.
-Each algebraic structure will have a function for reducing the reflected terms into their normal forms.
-The fact that all of these algebraic structures admit a canonical representative for any element is in fact a very nice property that we are using in order to decide equalities.
-Without this property, it would become a lot more complicated to decide the equivalence without brute-forcing a series of rewritings, that would have no guarantee to terminate.
-As said before, the normalisation functions also produce the proof of equivalence between the initial real value (the index of the input term) and the final real value (the index of the produced term).
-Thus, the type of the reduction function for group is :
+It remains to define the function \code{groupNormalise}, which is an
+instance of \code{norm} for groups:
 
@@
 
-The function of normalisation has more work to do for structures with many axioms (commutative-monoids, groups, commutative-groups, semi-rings and rings), than for the easier structure (semi-groups and monoids).
-The operations performed by the normalisation functions are detailed in the next section.
+Each algebraic structure has a function for reducing reflected terms to normal.
+The algebraic structures which concern us admit a canonical representative for
+any element, a property which we use to decide equalities. Without this
+property, it would be more complicated to decide equivalence without
+brute-forcing a series of rewritings, that would have no termination guarantee.
+
+The normalisation function has more work to do for structures with many
+axioms (commutative-monoids, groups, commutative-groups, semi-rings and rings),
+than for the simpler structures (semi-groups and monoids). In the next
+section, we describe the normalisation process.
 
@@
 
-For the sake of completeness, we describe the normal form for rings which is our most sophisticated structure.
-The normalisation function takes in input an expression expressed with sums, products, constants (zero, one...) and variables that belong to an ordered set $\mathcal{V}$ of variables.
-In short, the normalisation function takes in input a polynomial of multiple variables.
-In output, it must produce a normal form representing the same polynomial.
-Therefore, we have to decide a canonical representation of polynomials.
-Many canonical representations can be used.
-We decide to stick with classical mathematical conventions.
-The first one, is that the polynomial will be completely developed, ie, the distributivity of $*$ over $+$ will be applied until it can't be applied anymore.
-The advantage is the simplicity, as factorising would be significantly more complicated and we're looking for simplicity as the normalisation functions will also have to produce the proof of equivalence between the old and the new index.
-Because the polynomial is completely developed, at the toplevel, it is a sum :
+We describe the normal form for rings, which is our most sophisticated
+structure. The input to normalisation is an expression 
+with sums, products, constants and variables belonging to an
+ordered set $\mathcal{V}$ of variables, which is
+a polynomial of multiple variables. As output, it produces a
+normal form representing the same polynomial. Therefore, we need a
+canonical representation of polynomials. 
+There are several possibilities, but we choose
+classical mathematical conventions: 
+the polynomial will be completely developed, i.e., the
+distributivity of $*$ over $+$ will be applied until it cannot be applied
+further.
+This is a simple but effective approach: the benefit of simplicity is that we
+can directly produce a proof of equivalence between the new and old indices
+during normalisation.
+Because the polynomial is completely developed, at the
+toplevel, it is a sum:
 
@@
 
-One could be surprised by the fact that the normal form is a sum of product of monomial, and not directly a sum of monomials.
-The reason is the following.
-A monomial is a product of a constant $C_{i}^j$ (like $4$) and of a product of variables (like $x*y*z$).
-For example,  $5*(x*(y*z))$ is a monomial.
-Now let's consider the term $(5*(x*(y*z))) * (4*(z*z))$.
-This term is not a monomial, but we could be tempted to simplify it into the monomial $20*(x*(y*(z*(z*z))))$.
-However, that would assume that the product is always commutative, ie, that we can re-organize the subterms of a product as we want.
-This is the case in a commutative ring, but this does not hold for any ring.
-Therefore, after the full development, the polynomial is a sum of \emph{product of monomials}, and not directly a sum of monomials.
-The only rearrangement that can be done towards the multiplication is to check if two constants are consecutive in a product, and if so, to replace them by the constant that represents the product of them.
+It may be surprising that the normal form is a sum of product of
+monomials, and not directly a sum of monomials. This is because a
+monomial is a product of a constant $C_{i}^j$ (e.g. $4$) and of a product of
+variables (e.g. $x*y*z$). For example, $5*(x*(y*z))$ is a monomial. Now let's
+consider the term $(5*(x*(y*z))) * (4*(z*z))$. This term is not a monomial, but
+we could be tempted to simplify it into the monomial $20*(x*(y*(z*(z*z))))$.
+However, that would assume that the product is always commutative, which
+is not the case for \emph{every} ring.
+Therefore, after 
+development, the polynomial is a sum of \emph{product of monomials}, and
+not directly a sum of monomials. The only rearrangement that can be done
+towards the multiplication is to check if two constants are consecutive in a
+product, and if so, to replace them by the constant that represents 
+their product.
 
-However, because $+$ is always a commutative operator in a ring, the different products of monomials themselves can be rearranged in different ways in this sum.
-That will be done at the underneath level for commutative groups if we can provide an ordering on products of monomials.
-This ordering will be defined by using an ordering on monomials, that we will call \code{isBefore\_mon}, which works by looking at the order of variables for comparing two monomials $Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
+However, because $+$ is always commutative in a ring, the different
+products of monomials themselves can be rearranged in different ways in this
+sum. This will be done at the lower level for commutative groups if we can
+provide an ordering on products of monomials. This ordering will be defined 
+with an ordering on monomials, called \code{isBefore\_mon}, which
+looks at the order of variables for comparing two monomials
+$Monomial_{i}^{j}$ and $Monomial_{i'}^{j'}$.
 \[
@@
 
-The order between these two monomials will be decided by looking at the order between the first variables $Var_{i,1}^{j}$ and $Var_{i',1}^{j'}$.
-In the case where both monomials start with the same variable, we continue by inspecting the remaining variables.
-If we can't continue because one of the two monomials has less variables than the other, then we decide that the one with fewer variables comes first (exactly like the word ``house" comes before the word ``housemate" with the lexicographic order).
+The order between these two monomials is decided by looking at the order
+between the variables $Var_{i,1}^{j}$ and $Var_{i',1}^{j'}$. 
+If both monomials start with the same variable, we continue by inspecting
+the remaining variables. If one of the two monomials
+has fewer variables, that one comes first.
 
-We can now build the order on \emph{product of monomials} that we needed.
-We name it \code{isBefore}.
-Given two product of monomials $Prod_{i}$ and $Prod_{i'}$ we need to decide which one comes first.
-We will obviously use the order \code{isBefore\_mon} on the first monomials of these two products.
-If it says that $Monomial_{i}^{1}$ comes before $Monomial_{i'}^{1}$, then we decide that $Prod_{i}$ comes before $Prod_{i'}$.
-Conversely, if that's $Monomial_{i'}^{1}$ that comes first, then we decide that $Prod_{i'}$ comes first.
-However, if $Monomial_{i}^{1}$ and $Monomial_{i'}^{1}$ have exactly the same position in the order, then we continue by inspecting the remaining monomials with a recursive call.
-As previously, if we can't continue because one of the two products has less monomials than the other, then the one with fewer monomials will come first.
+We can now build the order on \emph{product of monomials}, named
+\code{isBefore}. Given two products of monomials $Prod_{i}$ and
+$Prod_{i'}$ we need to decide which one comes first. 
+We will use the
+order \code{isBefore\_mon} on the first monomials of these two products. If it
+says that $Monomial_{i}^{1}$ comes before $Monomial_{i'}^{1}$, then we decide
+that $Prod_{i}$ comes before $Prod_{i'}$. Conversely, if
+$Monomial_{i'}^{1}$ comes first, then $Prod_{i'}$ comes
+first. However, if $Monomial_{i}^{1}$ and $Monomial_{i'}^{1}$ have exactly the
+same position in the order, then we continue by inspecting the remaining
+monomials recursively. As previously, if 
+one of the two products has fewer monomials than the other, then that one 
+comes first.
 
-A few additional conventions had to be decided about the normal form :
+Additionally, we use the following conventions when deciding on
+a normal form:
 \begin{itemize}
 
-\item The top-level sum of the polynomial will be put in right-associative form : \\
+\item The top-level sum of the polynomial is in right-associative form: \\
 $prodMon_1\ +\ (prodMon_2\ + (prodMon_3\ +\ (...\ +\ prodMon_a)))$
-\item All the products that we have (the products of monomials and the products of variables), will also be written on the completely right-associative form.
-\item We always simplify as much as possible with constants.
-That includes simplifying the addition with zero and the multiplication with the constants zero and one, doing the computations between two nearby constants, etc...
-\item We always simplify the sum of an expression $e$ and its inverse $-e$ into the constant zero.
+\item All the products that we have (the products of monomials and the products
+of variables), are written in right-associative form.
+\item We simplify as much as possible with constants. This includes
+simplifying addition with zero and multiplication with the constants
+zero and one, doing the computations between two nearby constants, etc\ldots
+\item We simplify the sum of an expression $e$ and its inverse $-e$ to
+zero.
 
@@
 
-One of the novelty of our work is that we haven't built just a prover for a specific algebraic structure, but we have instead built a hierarchy of provers for many algebraic structures.
-Morover, each prover is not isolated from the others, and the various simplifications are not duplicated at the different levels : the normalisation function of each structure uses as much as possible the normalisation function of the structure from which it inherits.
-For example, the normalisation on monoids reuses the normalisation on semi-groups in such a way that it does not have to deal with the associativity.
-For this case, it's not a problem since the datatype reflecting terms in semi-groups has the same expressivity power than the one for monoids.
-However, that's not the case with groups and monoids: the function of normalisation for groups that will need to call the normalisation for monoids will have to encode the negations, because they can't be directly expressed in a monoid.
-For this reason, we will develop some specific encodings.
+A novelty of our work is that not only have we built a prover for a
+specific algebraic structure, but we have also built a hierarchy of provers
+for many algebraic structures. Moreover, each prover reuses components
+of the others so that
+the various simplifications are not duplicated at the different
+levels: the normalisation function of each structure uses as much as possible
+the normalisation function of the structure from which it inherits. For
+example, the normalisation on monoids reuses the normalisation on semi-groups
+in such a way that it does not have to deal with associativity. In this
+case, the datatype reflecting terms in semi-groups has
+the same expressive power as that for monoids. However, 
+there is a difficulty with
+groups and monoids: normalisation for groups
+needs to reuse the normalisation for monoids, meaning that
+we will have to encode 
+negations somehow, which can't be directly expressed in a monoid. For this
+reason, we develop some specific encodings.
 
-The idea is that we will encode negations as variables, and we will let the monoid prover deal with them as ordinary variables.
-In order to achieve this goal, we need the following datatype that will help us to distinguish between a real variable and the encoding of a negation :
+The idea is that we encode negations as variables, and let the
+monoid prover consider them as ordinary variables. To achieve this,
+we use the following datatype that helps us distinguish between a
+variable and the encoding of a negation:
 
@@
 
-We only need to encode negations of variables, as the negation of a constant can be simplified into a constant.
-Also, there can't be a negation of something different than an atom (a variable or a constant), because the normalisation in groups has first systematically propagated \code{Neg} inside the parenthesis, following the simplification\footnote{Note that we have to be careful and not simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative} $-(a+b) = -b + -a$.
+We only need to encode negations of variables, as we can simplify 
+the negation of a constant into a constant.
+Also, there cannot be a negation of something
+different non-atomic (i.e. a variable or a constant), because normalisation 
+of
+groups has systematically propagated \code{Neg} inside the parentheses,
+following simplification\footnote{Note that we have to be careful and not
+simplify it to $(-a) + (-b)$ as it would assume that $+$ is commutative}
+$-(a+b) = -b + -a$.
 
-All the constructors for variables now take a \code{Variable} as parameter, instead of taking directly an element of $(\code{Fin}\ n)$. That gives the following, for groups :
+All the constructors for variables now take a \code{Variable} as parameter,
+instead of taking directly an element of $(\code{Fin}\ n)$. That gives the
+following, for groups:
 
@@
 
-Thanks to this encoding, we can now transform an \code{ExprG} from the group level to an \code{ExprMo} at the monoid level.
-A constant $(\code{ConstG}\ p\ \Gamma\ c1)$ will be transformed into the corresponding constant $(\code{ConstMo}\ p\ \Gamma\ c1)$, a \code{PlusG} into the corresponding \code{PlusMo}, a real variable into the same real variable, the negation of a constant into the resulting constant, and finally the negation of a variable $i$ into a $(\code{VarMo}\ p\ (\code{EncodingNegOfVar}\ \Gamma\ i))$.
+Thanks to this encoding, we can now transform an \code{ExprG} from the group
+level to an \code{ExprMo} at the monoid level. A constant $(\code{ConstG}\ p\
+\Gamma\ c1)$ is transformed into the corresponding constant
+$(\code{ConstMo}\ p\ \Gamma\ c1)$, a \code{PlusG} into the corresponding
+\code{PlusMo}, a variable into the same variable, the negation of a
+constant into the resulting constant, and finally the negation of a variable
+$i$ into a $(\code{VarMo}\ p\ (\code{EncodingNegOfVar}\ \Gamma\ i))$.
 
-We've used a similar technique for converting an expression from the ring level to the commutative-group level, where we have encoded the product of monomials, because the product is not an operation defined at the commutative-group level.
-That enables the function of normalisation for rings to benefit from the normalisation function for commutative-groups.
+We use a similar technique for converting an expression from the ring level
+to the commutative-group level, where we encode the product of monomials,
+because the product is not defined at the commutative-group level.
+That enables the function of normalisation for rings to benefit from the
+normalisation function for commutative-groups.
 
@@
 
-We've built an automatic reflection mechanism which enables to let the machine build automatically the reflected terms for us.
-This is not something absolutely vital, as it only simplifies the usage of the tactics, avoiding to write long encodings by hand.
-To do so, we've used Idris' reflection mechanism, which enables to do pattern matching on syntax, rather than on constructors.
+We have built an automatic reflection mechanism which enables the machine to
+build the reflected terms for us automatically. This is not essential to our
+approach, but it simplifies the usage of the tactics by removing the need to
+write long encodings by hand.  To do so, we used Idris' reflection mechanism,
+which enables pattern matching on syntax, rather than on constructors.

### 2017-05-19 paper-CICM-17/content/1-introduction.tex
Some additional changes

@@
 two notions of equality: propositional equality, that can be
-manipulated in the language, and judgemental (or definition) equality.
+manipulated in the language, and judgemental (or definitional) equality.
 Propositional equality corresponds to the mathematical notion: this is
@@
 
-The only way for $(\code{Id}_a\ x\ y)$ to be inhabited is therefore that $x$
-equals $y$. In this case, the constructor \code{Refl} helps
-to create a proof of this equality: \code{(Refl\ x)} is precisely the proof
+The only way for $(\code{Id}_a\ x\ y)$ to be inhabited is therefore that $x$ and $y$ are equal by definition.
+In this case, the constructor \code{Refl} helps to create a proof of this equality: \code{(Refl\ x)} is precisely the proof
 which says that $x=_ax$.
@@
 (the third is a carry bit), and produces the two bits of
-the result, where the first is the more significant bit:
+the result, where the first is the most significant bit:
 
@@
 
-Unfortunately, this definition is rejected because the types of each side do
-not match.  The result of the first line \code{adc zero zero carry} is expected
-to have the type:
-
-\codeL{Binary\ 1\ ((c\ + 0)\ + 0)} but we provide a
-term of type:
-
-\code{Binary\ 1\ (c\ + (0\ + (0\ + 0)))}.
-
-The problem is even worse for the second case where the expected index is:
+Unfortunately, this definition is rejected because the types do
+not match for both patterns.  
+%The result of the first line \code{adc zero zero carry} is expected
+%to have the type:
+%
+%\codeL{Binary\ 1\ ((c\ + 0)\ + 0)} but we provide a
+%term of type:
+%
+%\code{Binary\ 1\ (c\ + (0\ + (0\ + 0)))}.
+For the second case, the expected index is:
 
@@
 could become the everyday routine in any dependently-typed language.
+It is worth mentioning that even without using dependent types, these proof obligations for equalities happen very frequently during the formal certification of most applications.
 
@@
 equalities in a \emph{hierarchy} of algebraic structures, including monoids,
-groups and rings (all potentially commutative), for the Idris language.,
+groups and rings (all potentially commutative), for the Idris language,
 making the following contributions:

### 2017-05-19 paper-CICM-17/content/3-hierarchyOfTactics.tex
Some additional changes

@@
 commutative-monoids, groups, commutative-groups, semi-rings and rings.
-The
-properties of an algebraic structure are expressed in an
+The properties of an algebraic structure are expressed in an
 interface (an interface in Idris is similar to a type class
@@
 be a \code{Prop} in systems, like Coq, that make a distinction between the
-world of computations and the world of logical statements}, and must be
-be reflexive, symmetric and transitive.
+world of computations and the world of logical statements}, and must be reflexive, symmetric and transitive.
 
@@
     trans : {x, y, z : c} $\rightarrow$ (x $\simeq$ y) $\rightarrow$ (y $\simeq$ z) $\rightarrow$ (x $\simeq$ z)    
-    set_eq : (x : c) $\rightarrow$ (y:c) $\rightarrow$ Maybe (x $\simeq$ y)
+    set_eq : (x : c) $\rightarrow$ (y : c) $\rightarrow$ Maybe (x $\simeq$ y)
 \end{lstlisting}
@@
 
-Working with equivalences instead of equalities brings one complication: Proofs
+Working with equivalences instead of equalities brings one complication : the proofs
 of correctness that we produce by hand cannot use Idris' ``rewrite'' mechanism,
@@
 We have this requirement because we support
-any equivalence relation.
+any equivalence relation. The user is free to define the equivalence relation of their choice, but it should be compatible with the operations that he is using.
 As with \code{Set}, there is no
@@
 Examples of magma are \code{Nat} equipped with addition, and \code{List} with
-concatenation.  Next, a monoid is a semi-group with a ``neutral``, or
-distinguished, element called \code{Zero}.
+concatenation.  Next, a monoid is a semi-group with the property of neutral element for a distinguished element called \code{Zero}.
 
@@
 When proving an equivalence $x \simeq y$, the universally-quantified
-variables are abstracted. Our
+variables are abstracted and they become part of the context. Our
 tatics normalise both sides of the ``potential equivalence'' $x
@@
 variables, constants and operators. For this
-reason, we work by reflection. We have one type of reflected
+reason, we work by reflection, which allows us to work on syntax instead of values. We define one type of reflected
 terms for each algebraic structure. The novelty is not the use of reflected
@@
          $\rightarrow$ ExprG p $\Gamma$ c1 $\rightarrow$ ExprG p $\Gamma$ (Neg c1)
-    VarG : (p : Group c) $\rightarrow$ {$\Gamma$:Vect n c} 
+    VarG : (p : Group c) $\rightarrow$ ($\Gamma$:Vect n c) 
          $\rightarrow$ (i:Fin n) $\rightarrow$ ExprG p $\Gamma$ (index i $\Gamma$)
@@
 We will explain how to implement the tactic for \code{Group} specifically,
-using a normalisation function \code{groupNormalise},
 but the other structures are implemented similarly.
@@
 the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof
-$p_y$ of $y \simeq y'$.  It now suffices to compare $e_{x'}$ and $e_{y'}$
-using a syntactic equality test \code{exprG\_eq} on the normal forms.
-
-If the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then
+$p_y$ of $y \simeq y'$.  It now suffices to compare the normal forms $e_{x'}$ and $e_{y'}$
+using a syntactic equality test \code{exprG\_eq}. If the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then
 they necessarily have the same type\footnote{We work with the
@@
 buildProofGroup p ex' ey' px py with (exprG_eq p ex' ey')
-    buildProofGroup p ex' ex' px py | Just refl = 
-                Just(eq_preserves_eq x y x' y' lp rp e1_equiv_e2)
+    buildProofGroup p ex' ey' px py | Just ex'_equiv_ey' = 
+                Just(eq_preserves_eq x y x' y' px py ex'_equiv_ey')
     buildProofGroup p ex' ey' px py | Nothing = Nothing
@@
 
-The argument of type $\code{ExprG}\ p\ \Gamma\ x'$ is the normalised reflected
-left hand size of the equivalence, which represents the value $x'$. 
-The function also expects proofs of $x \simeq x'$ and
+The arguments of type $\code{ExprG}\ p\ \Gamma\ x'$ and $\code{ExprG}\ p\ \Gamma\ y'$ are the normalised reflected
+left and right hand sides of the equivalence, which respectively represent the value $x'$ and $y'$. 
+This function also expects proofs of $x \simeq x'$ and
 of $y \simeq y'$, which are built by the normalisation process.
@@
 
-Each algebraic structure has a function for reducing reflected terms to normal.
-The algebraic structures which concern us admit a canonical representative for
+Each algebraic structure is equipped with a function for reducing reflected terms to their normal form.
+The algebraic theories which concern us admit a canonical representative\footnote{This only holds in the absence of additional axioms}  for
 any element, a property which we use to decide equalities. Without this
@@
 We describe the normal form for rings, which is our most sophisticated
-structure. The input to normalisation is an expression 
+structure. The input to the normalisation function is an expression 
 with sums, products, constants and variables belonging to an
-ordered set $\mathcal{V}$ of variables, which is
-a polynomial of multiple variables. As output, it produces a
+ordered set $\mathcal{V}$ of variables. In short, the normalisation function takes in input a polynomial of multiple variables. As output, it produces a
 normal form representing the same polynomial. Therefore, we need a
@@
 This is a simple but effective approach: the benefit of simplicity is that we
-can directly produce a proof of equivalence between the new and old indices
+can directly produce a proof of equivalence between the new and old concrete values
 during normalisation.
@@
 monomials, and not directly a sum of monomials. This is because a
-monomial is a product of a constant $C_{i}^j$ (e.g. $4$) and of a product of
+monomial is a product of a constant $C_{i}^j$ (e.g. $5$) and of a product of
 variables (e.g. $x*y*z$). For example, $5*(x*(y*z))$ is a monomial. Now let's
@@
 development, the polynomial is a sum of \emph{product of monomials}, and
-not directly a sum of monomials. The only rearrangement that can be done
+not directly a sum of monomials. The only rearrangement that can and needs to be done
 towards the multiplication is to check if two constants are consecutive in a
@@
 sum. This will be done at the lower level for commutative groups if we can
-provide an ordering on products of monomials. This ordering will be defined 
-with an ordering on monomials, called \code{isBefore\_mon}, which
+provide an ordering on products of monomials. This ordering will be defined by using an ordering on monomials, called \code{isBefore\_mon}, which
 looks at the order of variables for comparing two monomials
@@
 
-A novelty of our work is that not only have we built a prover for a
-specific algebraic structure, but we have also built a hierarchy of provers
+A novelty of our work is that instead of building a prover for a
+specific algebraic structure (like ring), we have built a hierarchy of provers
 for many algebraic structures. Moreover, each prover reuses components
@@
 case, the datatype reflecting terms in semi-groups has
-the same expressive power as that for monoids. However, 
+the same expressive power as that for monoids, so a term of a monoid can easily be transformed into a term of a semi-group. However, 
 there is a difficulty with
@@
 Also, there cannot be a negation of something
-different non-atomic (i.e. a variable or a constant), because normalisation 
+different non-atomic (i.e. a term that is not a variable or a constant), because normalisation 
 of

### 2017-05-19 paper-CICM-17/paper.tex
Pruning up to 2.3; add acknowledgements
Some additional changes

@@
 \begin{abstract}
-One difficulty with programming with dependent types is that proof obligations
-arise naturally once programs become even moderately sized. For example,
+One difficulty with reasoning and programming with dependent types is that proof obligations arise naturally once programs become even moderately sized. For example,
 implementing an adder for binary numbers indexed over their natural number
@@
 
+\subsubsection*{Acknowledgements}
+We thank the anonymous reviewers and Jacques Carette for their insightful
+comments on an earlier draft. We are also grateful for the support of
+the Scottish Informatics and Computer Science Alliance (SICSA) and
+EPSRC grant EP/N024222/1.
+
 \bibliographystyle{splncs03}

### 2017-05-19 paper-CICM-17/content/3-hierarchyOfTactics.tex
Some additional changes 2
Some additional changes 3

@@
 We have this requirement because we support
-any equivalence relation. The user is free to define the equivalence relation of their choice, but it should be compatible with the operations that he is using.
+any equivalence relation. The user is free to define the equivalence relation of their choice, but it should be compatible with the operations that they are using.
 As with \code{Set}, there is no
@@
 We will explain how to implement the tactic for \code{Group} specifically,
-but the other structures are implemented similarly.
+and the other structures are implemented similarly.
 The equivalence we are trying to prove is $x \simeq y$, where $x$ and $y$ are

### 2017-05-19 paper-CICM-17/content/3-hierarchyOfTactics.tex
Some additional changes 4

@@
 case, the datatype reflecting terms in semi-groups has
-the same expressive power as that for monoids, so a term of a monoid can easily be transformed into a term of a semi-group. However, 
+the same expressive power as that for monoids, so a term valid in a monoid can easily be transformed into a corresponding term of a semi-group. However, 
 there is a difficulty with

### 2017-05-19 paper-CICM-17/content/1-introduction.tex
Final changes for me

@@
 inhabited, then $x$ is said to be provably equal to $y$.
-Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $\code{Id} (a:\code{Type}):a \rightarrow a \rightarrow \code{Type}$.
+Thus, $Id$ is a type family (parameterised by the type $a$) indexed over two elements of $a$, giving $\code{Id}\ (a:\code{Type}):a \rightarrow a \rightarrow \code{Type}$.
 For
@@
 library of binary numbers. To ensure functional correctness,
-we define types \code{Bit} and \code{Binary},
-indexed over the value they represent (expressed as a natural number):
+we define the types \code{Bit} and \code{Binary} indexed over the value they represent (expressed as a natural number):
 
@@
 
-Our handwritten proof uses only the existence of a neutral
+Our handwritten proof \code{adc\_lemma\_2} uses only the existence of a neutral
 element, and the associativity and commutativity of $+$ on \code{Nat}. Thus,

### 2017-05-19 paper-CICM-17/content/3-hierarchyOfTactics.tex
Final changes for me

@@
 We describe operations, constants and properties of each algebraic structure 
-in an interface. The first structure is
-the magma. which is a structure built on top of \code{Set} that adds
+in an interface. The first algebraic theory is \code{Magma}, which is a structure built on top of \code{Set} that adds
 \code{Plus} operation, and no specific properties:
@@
 and the equivalence test \code{set\_eq}), and if it has a $+$ operation.  In
-fact, there is an additional requirement that will apply to all operators (in
-this case, the $+$ operation), which is that they need to be ``compatible" with
+fact, there is an additional requirement that will apply to all operations (in
+this case, the $+$ operation), which is that they need to be ``compatible'' with
 the equivalence relation, which is expressed by the following axiom for $+$:
@@
 tatics normalise both sides of the ``potential equivalence'' $x
-\simeq y$, and compare the results by syntactic equality. The
+\simeq y$, and compare the results by syntactic equivalence. The
 difficulty is that the normalisation function needs to consider
@@
 corresponding variable in the context, i.e., \code{(index\ i\ $\Gamma$)}.  The
-expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ of type
-$c$. Finally, if $e_x$ is an expression encoding the concrete value $x$, and
+expression \code{(ConstMa\ \_\ $\Gamma$\ k)} denotes the constant $k$ in any context $\Gamma$. Finally, if $e_x$ is an expression encoding the concrete value $x$, and
 $e_y$ is an expression encoding the concrete value $y$, then the expression
@@
 the normal form $e_{y'}$ of type $\code{ExprG}\ p\ \Gamma\ y'$ and a proof
-$p_y$ of $y \simeq y'$.  It now suffices to compare the normal forms $e_{x'}$ and $e_{y'}$
-using a syntactic equality test \code{exprG\_eq}. If the two normalised expressions $e_{x'}$ and $e_{y'}$ are equal, then
-they necessarily have the same type\footnote{We work with the
-heterogeneous equality by default in Idris, but as always, the only way to
-have a proof of \code{a:A = b:B} is when \code{A = B}}, and therefore $x'=y'$, which implies
-$x' \simeq y'$ because the propositional equality is more restrictive than an
-equivalence relation.  By using the two equivalences $x \simeq x'$ and $y
-\simeq y'$, and the new equivalence
-$x' \simeq y'$, we can use \code{eq\_preserves\_eq}, to get the desired proof of
-$x \simeq y$. We put all of this together in a function
-\code{buildProofGroup}:
+$p_y$ of $y \simeq y'$.
+It now suffices to compare the normal forms $e_{x'}$ and $e_{y'}$ using a syntactic equivalence test \code{ExprG\_eq}, because once everything is in normal form, being equivalent is just a matter of being syntactically equivalent.
+
+\begin{lstlisting}
+exprG_eq : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (p:Group c) $\rightarrow$ ($\Gamma$:Vect n c) 
+           $\rightarrow$ {x' : c} $\rightarrow$ {y' : c} 
+           $\rightarrow$ (ex' : ExprG p $\Gamma$ x') $\rightarrow$ (ey' : ExprG p $\Gamma$ y') 
+           $\rightarrow$ Maybe(x' $\simeq$ y')
+\end{lstlisting}
+This syntactical equivalence test checks if the two input terms $ex'$ and $ey'$ are \emph{syntactically} the same, and if they do, it directly builds a proof of equivalence between their indices $x' \simeq y'$, which is what we need, because we can use it
+with the two equivalences $x \simeq x'$ and $y\simeq y'$ that we already have, in order to get the desired proof of $x \simeq y$ with \code{eq\_preserves\_eq}.
+We put all of this together in a function \code{buildProofGroup}:
 
@@
 Each algebraic structure is equipped with a function for reducing reflected terms to their normal form.
-The algebraic theories which concern us admit a canonical representative\footnote{This only holds in the absence of additional axioms}  for
+The algebraic theories which concern us admit a canonical representative\footnote{It only holds for ``pure'' algebraic structures, ie, in the absence of additional axioms}  for
 any element, a property which we use to decide equalities. Without this

### 2017-05-20 paper-CICM-17/content/1-introduction.tex
Add a bit more stuff, prune the bibliography

@@
 and programming languages like 
-Agda~\cite{DBLP:conf/tldi/Norell09} and
+Agda~\cite{norell2007thesis} and
 Idris~\cite{brady2013idris} are based on Intensional Type Theories that contain
@@
 To demonstrate this,
-we revisit an example from previous work~\cite{DBLP:conf/plpv/Brady13} which
+we revisit an example from previous work~\cite{BradyTFP07} which
 shows how proof obligations arise when a type is indexed by natural
@@
 (the third is a carry bit), and produces the two bits of
-the result, where the first is the most significant bit:
+the result, where the first is the more significant bit:
 
@@
 \codeL{((c\ + (bit2\ + (v1\ + (v1\ + 0)))) + (bit\ + (v\ + (v\ + 0))))}
-while we're trying to provide a term indexed over
+\\
+while we're trying to provide a term indexed over:
 
@@
 different.
-If we use provisional definitions~\cite{DBLP:conf/plpv/Brady13}, the previous
-definition can be accepted by the system, but we now have two proof obligations
-\code{adc\_lemma\_1} and \code{adc\_lemma\_2} demanding proofs of equality
-between the two types, and they have to be provided by hand.
+To make the previous
+definition acceptable, we need to solve two proof obligations
+\code{adc\_lemma\_1} and \code{adc\_lemma\_2} which demand proofs of equality
+between the two types.
 For example, using a proof script:
@@
 element, and the associativity and commutativity of $+$ on \code{Nat}. Thus,
-we're rewriting a term by using the properties of a commutative monoid. We
-would like a generic prover for commutative monoids to find a proof 
+we're rewriting a term by using the properties of a commutative monoid. 
+With the right choice of combinators, such proofs could be made much simpler,
+but we would like a generic prover for commutative monoids to find a proof
 automatically.

### 2017-05-20 paper-CICM-17/content/3-hierarchyOfTactics.tex
Add a bit more stuff, prune the bibliography

@@
 sound representations.
+This is a \emph{local} approach to syntax representation~\cite{Farmer13}
+in that the reflected representation will \emph{only} represent terms in
+a magma.
 
@@
 data Variable : {c:Type} $\rightarrow$ {n:Nat} $\rightarrow$ (Vect n c) $\rightarrow$ c $\rightarrow$ Type where
-  RealVariable : ($\Gamma$:Vect n c) $\rightarrow$ (i:Fin n) $\rightarrow$ Variable $\Gamma$ (index i $\Gamma$)
-  EncodingNeg : ($\Gamma$:Vect n c) $\rightarrow$ (i:Fin n) $\rightarrow$ Variable $\Gamma$ (Neg (index i $\Gamma$))
+  RealVariable : ($\Gamma$:Vect n c) $\rightarrow$ 
+                 (i:Fin n) $\rightarrow$ Variable $\Gamma$ (index i $\Gamma$)
+  EncodingNeg : ($\Gamma$:Vect n c) $\rightarrow$ 
+                (i:Fin n) $\rightarrow$ Variable $\Gamma$ (Neg (index i $\Gamma$))
 \end{lstlisting}
@@
 which enables pattern matching on syntax, rather than on constructors. 
+While we omit the full details due to space constraints, reflecting values
+involves defining functions of the following form:
+
+\begin{lstlisting}
+%reflection reflectGroup : (p : Group c) $\rightarrow$ ($\Gamma$ : Vect n c) $\rightarrow$ 
+                           (x : c) $\rightarrow$ ($\Gamma'$ ** ExprG p ($\Gamma'$ ++ $\Gamma$) x)
+\end{lstlisting}
+
+The \texttt{\%reflection} annotation means that Idris treats this as a 
+compile time function on \emph{syntax}. Given a value of type
+\texttt{c}, in some context $\Gamma$, it constructs a reflected expression 
+in an extended context \texttt{$\Gamma'$ ++ $\Gamma$}. The context
+contains references to subexpressions 
+which are not themselves representable by \texttt{ExprG}.

### 2017-05-20 paper-CICM-17/content/4-relatedWork.tex
minor changes before Edwin shortens the paper
Add a bit more stuff, prune the bibliography

@@
 
-As mentioned, Coq's implementation of a ring prover~\cite{Coq2005} was also using proof-by-reflection techniques, but without the guarantees obtained with our type-safe reflection mechanism, and without the correct-by-construction approach : they have first defined the normalisation of terms, and afterwards they've proven the correctness of the normalisation with the external lemma :
-$\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\ (\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\ \code{reify}\ e2$ which therefore needs a \code{reify} function that we do not need in our development.
-Another difference is that Coq's prover only deals with rings and semi-rings (commutative or not), but not with any of the intermediate structures (semi-group, monoid, group, etc), unlike us.
-However, their implementation has better performances due to the use of sparse normal form and more optimised algorithms.
-Our automatic reflection was written with Idris reflection mechanism which allows to do pattern matching on syntax, and their automatic reflection is programmed in Ltac~\cite{DelahayeLTac} : a proof dedicated and untyped meta-language for the writing of automations.
-\\
+Coq's ring solver~\cite{Coq2005}, like ours, is
+implemented using proof-by-reflection techniques, but without the guarantees obtained with
+our type-safe reflection mechanism, and without the correct-by-construction
+approach: first, they define the normalisation of terms, then
+they prove correctness of the normalisation with an external lemma:
+$\forall\ (e1\ e2\ :\ \code{Expr}),\ beq_{Expr}\ (\code{norm}\ e1)\
+(\code{norm}\ e2)\ =\ true\ \rightarrow\ \code{reify}\ e1\ \simeq\
+\code{reify}\ e2$.
+This needs a \code{reify} function that we do not
+in our development. Another difference is that Coq's prover only deals
+with rings and semi-rings (commutative or not), but not with any of the
+intermediate structures (semi-group, monoid, group, etc), unlike our solver. 
+However,
+their implementation has better performances due to the use of sparse normal
+form and more optimised algorithms.  Our automatic reflection was written with
+Idris reflection mechanism which allows pattern matching on syntax, and
+their automatic reflection is programmed in Ltac~\cite{DelahayeLTac}, a proof
+dedicated and untyped meta-language for the writing of automations.
+More recently, the Mtac extension~\cite{Ziliani13} provides a
+typed language for implementation proof automations.
 
-Proof automations have been investigated for various proof assistants and programming languages.
-In Coq, appart from the ring prover, there's also the omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic (i.e.
-a universally quantified formula made of equations and inequations), and a field~\cite{DelahayeField} decision procedure for real numbers, which plugs to Coq's ring prover after simplification of the multiplicative inverses.
-Various automations have also been done in Agda~\cite{DBLP:conf/mpc/KokkeS15} ~\cite{Lindblad04}.
-\\
+Proof automations have been investigated for various proof assistants and
+programming languages. In Coq, appart from the ring prover, there's also the
+Omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic
+(i.e. a universally quantified formula made of equations and inequations), and
+a field~\cite{DelahayeField} decision procedure for real numbers, which plugs
+to Coq's ring prover after simplification of the multiplicative inverses.
+Various automations have also been done in
+Agda~\cite{DBLP:conf/mpc/KokkeS15,Lindblad04}.
+%
+Proofs by reflection has been intensively
+studied~\cite{ChlipalaBook,Malecha14}, but again without anything similar to
+the type-safe reflection that we have presented here.
 
-Proofs by reflection has been intensively studied as in~\cite{ChlipalaBook} and~\cite{Malecha14}, but without anything similar to the type-safe reflection that we have presented here. \\
-
-Finally, if we leave the ground of nice mathematical structures, one can decide to work with arbitrary rewriting rules, but in the general case there isn't a complete decision procedure for such systems, because there is usually no normal form.
-This is where deduction modulo~\cite{Dowek03} ~\cite{DelahayeModulo} and proof search heuristics starts.
-\\
+If we leave the ground of nice mathematical structures, one can decide
+to work with arbitrary rewriting rules, but in the general case there isn't a
+complete decision procedure for such systems, because there is usually no
+normal form. This is where deduction modulo~\cite{Dowek03,DelahayeModulo} and
+proof search heuristics start.

### 2017-05-20 paper-CICM-17/content/5-conclusions.tex
Add a bit more stuff, prune the bibliography

@@
 
-We have implemented a very generic solution to the initial problem of index mismatch that appeared in section~\ref{sect:motivatingExample} when we were using indexed types.
-This solution takes the form of a hierarchy of provers for equivalences in algebraic structures.
-These provers are generic in multiple ways: they work for many algebraic structures (semi-group, monoid, commutative monoid, group, commutative group, ring and semi-ring), for any type that behaves as one of these structures, and for any equivalence relation on this type (and not only for the propositional equality).
-The implementation is modular and each prover reuses the prover of the structure from which it inherits.
-These provers can automatically prove equivalences between terms, which discharges the user of providing proofs for some proof obligations, as \code{adc\_lemma\_2}.
-Thus, these provers enable the user to focus on the interesting proofs that requires specific knowledge and creativity, instead of wasting time for proving routine lemmas that can be automated.\\
+We have implemented a very generic solution to the initial problem of index
+mismatch that appeared in section~\ref{sect:motivatingExample} when we were
+using indexed types. This solution takes the form of a hierarchy of provers for
+equivalences in algebraic structures. These provers are generic in multiple
+ways: they work for many algebraic structures (semi-group, monoid, commutative
+monoid, group, commutative group, ring and semi-ring), for any type that
+behaves as one of these structures, and for any equivalence relation on this
+type (and not only for the propositional equality). The implementation is
+modular and each prover reuses the prover of the structure from which it
+inherits. These provers can automatically prove equivalences between terms,
+which discharges the user of providing proofs for some proof obligations, as
+\code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the
+interesting proofs that requires specific knowledge and creativity, instead of
+wasting time for proving routine lemmas that can be automated.
 
-The correct-by-construction method that we have followed has involved the design of a type-safe reflection mechanism where reflected terms are indexed over the concrete inputs, and from which we are able to pull out the proofs easily.
-The construction of the proof is done step by step, and it follows the construction of the normalised terms.
-Therefore, unlike Coq's and Agda's ring provers, we do not have the traditional duplication that happens when separating the computational content from the proof of correctness.
-Instead, in our approach, the construction of the proof is guided by the construction of the normalised terms, which, in addition of avoiding redundancy, simplifies the proof generation considerably.
-This development shows that if dependent types effectively bring some new problems, they are also very expressive tools for building correct-by-construction software where the development is \emph{driven by the types}.
-\\
+The correct-by-construction method that we have followed has involved the
+design of a type-safe reflection mechanism where reflected terms are indexed
+over the concrete inputs, and from which we are able to pull out the proofs
+easily. The construction of the proof is done step by step, and it follows the
+construction of the normalised terms. Therefore, unlike Coq's and Agda's ring
+provers, we do not have the traditional duplication that happens when
+separating the computational content from the proof of correctness. Instead, in
+our approach, the construction of the proof is guided by the construction of
+the normalised terms, which, in addition of avoiding redundancy, simplifies the
+proof generation considerably. This development shows that if dependent types
+effectively bring some new problems, they are also very expressive tools for
+building correct-by-construction software where the development is \emph{driven
+by the types}.
 
-This work can be extended to build new provers for less common algebraic structures and for more specific structures.
-For example, regular expressions form a ``pre semi-ring" with some extra axioms.
-We will refactor the semi-ring level with the creation of the intermediate structure of pre semi-ring, that will not necessary have the property that $0$ is an annihilator element for the product.
-Then, we could build a specific prover for regular expressions, that would use the normalisation function of the pre semi-ring level and that would only have to deal with the specific properties of regular expressions : the neutral element $\varnothing$ for the concatenation of languages is also a neutral element for the product of languages, the idempotence of the addition of languages, and the rules of simplifications for the new Kleene star operation.
+This work can be extended to build new provers for less common algebraic
+structures and for more specific structures. For example, regular expressions
+form a ``pre semi-ring" with some extra axioms. We will refactor the semi-ring
+level with the creation of the intermediate structure of pre semi-ring, that
+will not necessary have the property that $0$ is an annihilator element for the
+product. Then, we could build a specific prover for regular expressions, that
+would use the normalisation function of the pre semi-ring level and that would
+only have to deal with the specific properties of regular expressions : the
+neutral element $\varnothing$ for the concatenation of languages is also a
+neutral element for the product of languages, the idempotence of the addition
+of languages, and the rules of simplifications for the new Kleene star
+operation.

### 2017-05-20 paper-CICM-17/content/1-introduction.tex
A bit more related work

@@
 which says that $x=_ax$. 
+Here, we are using the notation of Idris, where unbound variables
+like \texttt{a} in the definition of \texttt{Id} are
+\emph{implicitly} quantified, as a concise programming notation.
 
@@
 we're rewriting a term by using the properties of a commutative monoid. 
-With the right choice of combinators, such proofs could be made much simpler,
-but we would like a generic prover for commutative monoids to find a proof
-automatically.
+With the right choice of combinators~\cite{Carette12} such proofs could be made
+much simpler, but we would like a generic prover for commutative monoids to
+find a proof automatically.
 

### 2017-05-20 paper-CICM-17/content/3-hierarchyOfTactics.tex
A bit more related work

@@
 A novelty of our work is that instead of building a prover for a
-specific algebraic structure (like ring), we have built a hierarchy of provers
-for many algebraic structures. Moreover, each prover reuses components
+specific algebraic structure, we have built a hierarchy of provers.
+Each prover reuses components
 of the others so that
-the various simplifications are not duplicated at the different
-levels: the normalisation function of each structure uses as much as possible
+the simplifications are not duplicated at different
+levels: normalisation of each structure uses as much as possible
 the normalisation function of the structure from which it inherits. For
-example, the normalisation on monoids reuses the normalisation on semi-groups
-in such a way that it does not have to deal with associativity. In this
+example, normalisation on monoids reuses normalisation on semi-groups
+so that it does not have to deal with associativity. In this
 case, the datatype reflecting terms in semi-groups has
-the same expressive power as that for monoids, so a term valid in a monoid can easily be transformed into a corresponding term of a semi-group. However, 
+the same expressive power as that for monoids, so a term in a monoid can
+be transformed directly into a corresponding term of a semi-group. However, 
 there is a difficulty with
-groups and monoids: normalisation for groups
-needs to reuse the normalisation for monoids, meaning that
+groups and monoids: if normalisation for groups
+uses the normalisation for monoids, 
 we will have to encode 
-negations somehow, which can't be directly expressed in a monoid. For this
-reason, we develop some specific encodings.
+negations somehow, which can't be directly expressed in a monoid.
+Therefore, we develop some specific encodings.
 
@@
 We have built an automatic reflection mechanism which enables the machine to
-build the reflected terms for us automatically. This is not essential to our
+build reflected terms. This is not essential to our
 approach, but it simplifies the usage of the tactics by removing the need to

### 2017-05-20 paper-CICM-17/content/4-relatedWork.tex
A bit more related work

@@
 \code{reify}\ e2$.
-This needs a \code{reify} function that we do not
-in our development. Another difference is that Coq's prover only deals
+This needs a \code{reify} function, which we do not need.
+Furthermore, Coq's prover deals
 with rings and semi-rings (commutative or not), but not with any of the
-intermediate structures (semi-group, monoid, group, etc), unlike our solver. 
+intermediate structures (semi-group, monoid, group, etc).
 However,
@@
 More recently, the Mtac extension~\cite{Ziliani13} provides a
-typed language for implementation proof automations.
+typed language for implementating proof automation.
 
-Proof automations have been investigated for various proof assistants and
-programming languages. In Coq, appart from the ring prover, there's also the
+As well as the ring solver, Coq also provides the
 Omega solver~\cite{Cregut04}, which solves a goal in Presburger arithmetic
@@
 to Coq's ring prover after simplification of the multiplicative inverses.
-Various automations have also been done in
+Agda's reflection
+mechanism\footnote{\url{http://wiki.portal.chalmers.se/agda/pmwiki.php?n=ReferenceManual.Reflection}}
+gives access to a representation of the current goal (that is,
+the required type) at a particular point in a program.
+This allows various proof automations to be done in
 Agda~\cite{DBLP:conf/mpc/KokkeS15,Lindblad04}.
+
+Proofs by reflection has been intensively
+studied~\cite{ChlipalaBook,Malecha14}, but without anything similar to
+the type-safe reflection that we have presented here.
 %
-Proofs by reflection has been intensively
-studied~\cite{ChlipalaBook,Malecha14}, but again without anything similar to
-the type-safe reflection that we have presented here.
-
 If we leave the ground of nice mathematical structures, one can decide

### 2017-05-20 paper-CICM-17/content/5-conclusions.tex
A bit more related work

@@
 
-We have implemented a very generic solution to the initial problem of index
-mismatch that appeared in section~\ref{sect:motivatingExample} when we were
+We have implemented a generic solution to the initial problem of index
+mismatch (section~\ref{sect:motivatingExample}) when 
 using indexed types. This solution takes the form of a hierarchy of provers for
-equivalences in algebraic structures. These provers are generic in multiple
+equivalences in algebraic structures. These provers are generic in several
 ways: they work for many algebraic structures (semi-group, monoid, commutative
-monoid, group, commutative group, ring and semi-ring), for any type that
-behaves as one of these structures, and for any equivalence relation on this
-type (and not only for the propositional equality). The implementation is
+monoid, group, commutative group, ring and semi-ring); for any type that
+behaves as one of these structures; and, for any equivalence relation on this
+type. The implementation is
 modular and each prover reuses the prover of the structure from which it
 inherits. These provers can automatically prove equivalences between terms,
-which discharges the user of providing proofs for some proof obligations, as
-\code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the
+so the user need not prove obligations by hand,
+like \code{adc\_lemma\_2}. Thus, these provers enable the user to focus on the
 interesting proofs that requires specific knowledge and creativity, instead of
-wasting time for proving routine lemmas that can be automated.
+routine, automatable, lemmas.
 
-The correct-by-construction method that we have followed has involved the
-design of a type-safe reflection mechanism where reflected terms are indexed
-over the concrete inputs, and from which we are able to pull out the proofs
-easily. The construction of the proof is done step by step, and it follows the
-construction of the normalised terms. Therefore, unlike Coq's and Agda's ring
-provers, we do not have the traditional duplication that happens when
-separating the computational content from the proof of correctness. Instead, in
-our approach, the construction of the proof is guided by the construction of
-the normalised terms, which, in addition of avoiding redundancy, simplifies the
-proof generation considerably. This development shows that if dependent types
+Our correct-by-construction method involved the design of a type-safe
+reflection mechanism where reflected terms are indexed over concrete inputs,
+and from which we are able to extract proofs directly.  Unlike Coq's and Agda's
+ring provers, we do not have the duplication that arise when separating the
+computational content from the proof of correctness. Instead, construction of
+the proof is done step by step, following the construction of the normalised
+terms.  In addition to avoiding redundancy, this simplifies the proof
+generation considerably. This development shows that if dependent types
 effectively bring some new problems, they are also very expressive tools for
@@
 structures and for more specific structures. For example, regular expressions
-form a ``pre semi-ring" with some extra axioms. We will refactor the semi-ring
+form a ``pre semi-ring'' with some extra axioms. We will refactor the semi-ring
 level with the creation of the intermediate structure of pre semi-ring, that

### 2017-05-20 paper-CICM-17/content/3-hierarchyOfTactics.tex
Automatic reflection : the extension if now on the right to avoid the change of the index of the variables that were already there
CICM17 : Final version for Franck

@@
 \Gamma\ c1)$ is transformed into the corresponding constant
-$(\code{ConstMo}\ p\ \Gamma\ c1)$, a \code{PlusG} into the corresponding
+$(\code{ConstMo}\ \_\ \Gamma\ c1)$, a \code{PlusG} into the corresponding
 \code{PlusMo}, a variable into the same variable, the negation of a
 constant into the resulting constant, and finally the negation of a variable
-$i$ into a $(\code{VarMo}\ p\ (\code{EncodingNegOfVar}\ \Gamma\ i))$.
+$i$ into a $(\code{VarMo}\ \_\ (\code{EncodingNeg}\ \Gamma\ i))$.
 
@@
 %reflection reflectGroup : (p : Group c) $\rightarrow$ ($\Gamma$ : Vect n c) $\rightarrow$ 
-                           (x : c) $\rightarrow$ ($\Gamma'$ ** ExprG p ($\Gamma'$ ++ $\Gamma$) x)
+                           (x : c) $\rightarrow$ ($\Gamma'$ ** ExprG p ($\Gamma$ ++ $\Gamma'$) x)
 \end{lstlisting}
@@
 \texttt{c}, in some context $\Gamma$, it constructs a reflected expression 
-in an extended context \texttt{$\Gamma'$ ++ $\Gamma$}. The context
+in an extended context \texttt{$\Gamma$ ++ $\Gamma'$}. The context
 contains references to subexpressions 

### 2018-05-03 Old/OldPapers/paper-JFP-DTP/content/3-generalProblem.tex
moving an old folder

--- 2016-01-19 paper-JFP-DTP/content/3-generalProblem.tex
@@
 
-Because there is no additional operations in a SemiGroup or in a Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
+Because there is no additional operations in SemiGroup and Monoid, the datatypes for reflected terms in these two structures will have exactly the same shape as the one for Magma that we've given above.
 However, the one for $Group$ will introduce two new constructors for the $Neg$ and $Minus$ operations :
@@
 Without this property, it would become a lot more complicated to decide the equality without brute-forcing a serie of rewriting, that might even not terminate.
-These normalisation functions will compute this canonical represent, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the real value indexing the original term and the real value indexing the produced term.
+These normalisation functions will compute the canonical representation of the reflected input term, by rewriting the given term multiple times, and at the same time, it will produce the proof of equality between the original real value (indexing the input term) and the produced real value (indexing the produced term).
+Thus, the type of the reduction function for group is :
 
-\begin{figure}[H]
-\figrule
-\begin{center}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{Type of the reduction function for terms reflecting elements in a Group}
-\label{groupReduce}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Type of the function for deciding equality between elements in a Group}
-\label{groupDecideEqType}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Decides if two elements of a group are equal}
-\label{groupDecideEq}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Composes the two proofs if the normal forms are the same}
-\label{buildProofGroup}
-\figrule
-\end{figure}
+
 We've used here the function $exprG\_eq$ to decide the syntactical equality between the normalised reflected terms $ex'$ and $ey'$.
@@
 \end{lstlisting}
-\end{center}
-\caption{Type of the function doing the reflection for constants}
-\label{ConstantsRingReflector}
-\figrule
-\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Automatic reflection to ring}
-\label{reflectRingTermType}
-\figrule
-\end{figure}
 
 This function will work very similarly to the reflection function for $Nat$ that we've writen in section~\ref{sect:ReflectNat}.
-Here, there are just more constructions to deal with, because $c$ is equipped with negations, subtractions and multiplications.
-We just show the pattern for the multiplication.
+Here, there are simply more constructions to deal with, because $c$ is also equipped with negations, subtractions and multiplications, and not only with additions.
+We show one of these new pattern on figure~{\ref{reflectRingTerm_pattern1}.
 

### 2018-05-03 Old/OldPapers/paper-JFP-DTP/content/4-detailsNormalisation.tex
moving an old folder

--- 2016-01-19 paper-JFP-DTP/content/4-detailsNormalisation.tex
@@
 \end{lstlisting}
+\end{center}
+\caption{Computing with associativity in a Semi-Group, first pattern}
+\label{assoc}
+\figrule
+\end{figure}
 
@@
 \end{lstlisting}
-\end{center}
-\caption{Real variables and encodings}
-\label{Variable}
-\figrule
-\end{figure}
 
@@
 The constructor for variable now takes a Variable, instead of taking directly a $(Fin\ n)$.
-\begin{figure}[H]
-\figrule
-\begin{center}
+
 \begin{lstlisting}
@@
 \end{lstlisting}
-\end{center}
-\caption{new Var constructor}
-\label{VarG}
-\figrule
-\end{figure}
 
@@
 Once this encoding is computed, the antepenultimate step is to call the monoid prover and to let it finish the normalisation.
-Finally, we need to convert back the result into an $ExprG$ by interpreting the $EncodingNegOfVar$ into a $Neg$ of the corresponding variable.
+Finally, we need to convert back the result into an $ExprG$ by interpreting the potential $EncodingNegOfVar$ into $Neg$ of the corresponding variable.
 
